### 2019-04-05 paper/spinifex-paper.tex
init commit. figure and caption references wrong.

% !TeX root = RJwrapper.tex
\title{spinifex; manual control of dynamic linear projections of high-dimensional data}
\author{by Nicholas Spyrison, Dianne Cook}

\maketitle

\abstract{%
The class of dynamic linear projections that are collectively known as `tours' provide a unique dynamic visualization of numeric multivariate data. Tours are particularly useful for understanding the structure held within multivariate data, and in association with techniques for dimension reduction, supervised, and unsupervised classification. The \emph{R} package \emph{tourr} offers a variety of path generators and geometric displays for conducting tours. This paper discusses an extension package, \emph{spinifex}, that adds support for the path generation of manual tours and extends the display of tours to use with the contemporary animation packages, \emph{plotly} and \emph{gganimate}. Manual tours are used to explore the sensitivity of structure as the contributions of a manipulation variable are changed. This particularly useful after identifying a feature of interest.
A recent paper \{Wang et al. (2018)\} visualizes the sensitivey of the hadronic experiments to nucleon structure. Sensitivity was characterized in non-linear 3D embeddings of the first 10 principal components. This research applies manual tours to this data showing that manual tours resolves more structrual information that is orthogonal to the original viewing plane.
}

% Any extra LaTeX you need in the preamble

\bibliography{RJrefs}

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

A tour is a multivariate data analysis technique in which a sequence of linear (orthogonal) projections are viewed as an animation while the orientation of the projection basis is rotated across time. Each frame of the sequence corresponds to a small change in the projection for a smooth transition that perseveres continuity.

While there are numerous methods that generate tour paths, this research focuses on the manual tour. The manual tour was described in Cook and Buja (1997) and allows a user to control the projection coefficients of a select variable has in a 2D projection. The manipulation of these coefficients allows the analyst to explore how sensitive the projections structure is to these changes. This makes manual tours particularly useful once a feature of interest has been identified, for example, with the use of a guided tour (Cook et al. 1995). The path of a guided tour is selected via projection pursuit, the optimization of an index function on the projection via a hill climbing algorithm. This allows guided tours to identify interesting projection features rapidly given the relatively large parameter-space. Once the given projection has been provided, it is time to define the path of rotation.

Ideally, the path would be intuitively user-generated from physical movement, be it through mouse or motion capture. Unfortunately, this type of dynamic control has proven difficult to capture for in R. Because of this, manual tours were not implemented in \emph{tourr}. This research allows for the consumption, but not the generation, of such dynamic input. After the capture of an oblique user motion, the rotation needs to be applied to step 3 (rotation sequence) of the algorithm discussed below. In the section below we stick with a radial rotation where, \(\theta\), the angle of in-projection-plane rotation is held constant.

Spinifex utilizes two new animation packages, \emph{plotly} (Sievert 2018) and \emph{gganimate} (Pedersen and Robinson 2019), to display tours, manual or other saved tours. From a given projection, the user can choose which variable to control, and the animation sequence is generated to remove the variable from the projection, and then extend its contribution to be the sole variable in one direction. This allows the viewer to assess the change in structure induced in the projection by the variable's contribution.

The paper is organized as follows. Section \ref{sec:algorithm} explains the algorithm using a toy dataset. Section \ref{sec:display} discussed the display of the animation after the path has been generated. Section \ref{sec:application} illustrates how this can be used for sensitivity analysis applied to contemporary high energy physics. The last section, \ref{sec:discussion} summarizes the work and discusses future research.

\hypertarget{sec:algorithm}{%
\subsection{Algorithm}\label{sec:algorithm}}

The section below describes the algorithm for performing a 2D radial manual tour:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Provided with a 2D projection, choose a variable to explore. This is called the ``manip'' variable.
\item
  Create a 3D manipulation space, where the manip variable has the full contribution.
\item
  Generate a rotation sequence which increases the norm of the coefficient to 1 and zeros it.
\end{enumerate}

The steps are described in more detail below. The R functions used below mentioned briefly, but more complete code example can be found in section \ref{sec:usage}

\hypertarget{notation}{%
\subsubsection{Notation}\label{notation}}

This section describes the notation used in the algorithm for a 2D radial manual tour.

\begin{itemize}
  \item $\textbf{X}$, the data, an $n \times p$ numeric matrix to be embedded in two dimensions.
  \item $\textbf{B} = (B_1,~ B_2)$, any orthonormal projection basis set, $p \times 2$ matrix, describing the projection from $p$ to two dimensions.
  \item $\textbf{e}$, a zero column vector of length $p$ with the $k-$th element set to one, where $k$ is the number of the variable to manipulate.
  \item $\theta$, the angle of in-projection-plane rotation, for example, on the reference axes. 
  \item $\phi$, the angle of out-of-projection-plane rotation, coming into the manipulation space.
\end{itemize}

The algorithm primarily operates on the projection basis and utilizes the data only when making a display. The projection space can be viewed at any point in the process by pre-multiplying the data and plotting the first 2 variables.

\hypertarget{toy-data-set}{%
\subsubsection{Toy data set}\label{toy-data-set}}

The flea data, originally from Lubischew (1962), available in the R package \emph{tourr} (Wickham et al. 2011) is used to illustrate the algorithm. The data contains 74 observations across 6 variables, physical measurements of the flea beetles. Each observation belonging to one of three species.

The data is defined. A basis set (ideally that views an interesting feature) should be provided to explore the sensitivity of the variables to the structure. To identify a projection containing an interesting feature, apply
a guided tour(Cook, Swayne, and Buja 2007) on the flea data. In a guided tour the projection sequence is selected by optimizing an index via hill-climbing. In this case, the holes index is selected. The holes index is maximized by when the projected observations are furthest from the center. Figure @ref(fig:step0) shows a locally optimized projection for this data. The left plot displays the reference axes of the projection basis, a visual indication of the magnitude and direction each variable contributed to the projections. The right plot shows the projection of the data through the basis set described by the reference axes (left). Data points are colored and given point characters according to the species of the flea (the guided tour was unsupervised with this information).

\begin{Schunk}
\begin{figure}

{\centering \includegraphics[width=0.95\linewidth]{spinifex_paper_files/figure-latex/step0-1}

}

\caption[(ref:step0-cap)]{(ref:step0-cap)}(\#fig:step0)
\end{figure}
\end{Schunk}

(ref:step0-cap) Basis reference axes (left) and projected data (right) of standardized flea data. Data points color and shape are mapped to beetle species. Basis identified by a holes-index guided tour. The variables \texttt{aede2} and \texttt{tars1} contribute mostly orthogonal to the other variables. We'll select \texttt{aede2} as our manipulation variable to see how the structure of the projection changes as we rotate \texttt{aede2} into and out of the projection.

Call \texttt{view\_basis()} on a basis to produce a \emph{ggplot2} graphic similar to @ref(fig:step0). Projection space is always available for display via the matrix multiplication \(\textbf{X}_{[n,~p]} ~*~ \textbf{B}_{[p,~d]} ~=~ \textbf{P}_{[n,~d]}\).

\hypertarget{step-1-choose-variable-of-interest}{%
\subsubsection{Step 1) Choose variable of interest}\label{step-1-choose-variable-of-interest}}

In figure @ref(fig:step0), above, the contributions of the variables \texttt{tars1} and \texttt{aede2} are mostly orthogonal to the contributions of the other four variables. These two variables explain the variation of the data between the purple and green species. We select \texttt{aede2} as the manip var, the variable to be manipulated as it typically has a larger contribution after the optimizing the holes index. The question that will be explored in the explanation of the algorithm is how important the variable \texttt{aede2} is to the separation of the clusters.

\hypertarget{step-2-create-the-manip-space}{%
\subsubsection{Step 2) Create the manip space}\label{step-2-create-the-manip-space}}

Initialize a zero vector \(e\) of \(p\) elements. Because \texttt{aede2} is the fifth variable in the data, set the \(k=5\)-th element to one giving the manip var a full contribution in this dimension. Use the Gram-Schmidt process to orthonormalize the zero vector onto the basis yielding the 3D manipulation space, \textbf{M}.

\begin{align*}
  \textbf{e} &\leftarrow Orthonormalize_{GS}(\textbf{e}) w.r.t. Basis \\
  &= \textbf{e} - \langle \textbf{e},\textbf{B}_1 \rangle \textbf{B}_1 - \langle \textbf{e}, \textbf{B}_2 \rangle \textbf{B}_2 \\
  \\
  \textbf{M}_{[p,~3]} &= (\textbf{B}_1,\textbf{B}_2,\textbf{e})
\end{align*}

Adding this extra dimension to our basis plane allows for the coefficients of the specified variable to be changed. For example, the ability to lift a piece of paper, rather than being constrained to the motion on a table top. Orthonormalizing rescales the new depth vector while the projection down to 2D is the original basis, that is the first \(d\) vectors remain constant. Imagine the reference axes (and projection plane) laying flat on a table, while a new dimension exists with axes projecting back onto the reference axes. An illustration of such can be seen below in figure @ref(fig:step2). The manip var is highlighted, while the depths of the other variables are not depicted.

(ref:step2-cap) Manipulation space for controlling the contribution of \texttt{aede2} of standardized flea data. Basis selected by a holes-index guided tour. The Projection plane is shown in blue. The manipulation axis, in red, allows the coefficients of the manip var to be changed.

\begin{Schunk}
\begin{figure}

{\centering \includegraphics[width=1\linewidth]{spinifex_paper_files/figure-latex/step2-1}

}

\caption[(ref:step2-cap)]{(ref:step2-cap)}(\#fig:step2)
\end{figure}
\end{Schunk}

The representation in @ref(fig:step2) can be duplicated by calling the function \texttt{view\_manip\_space()}.

\hypertarget{step-3-generate-rotation}{%
\subsubsection{Step 3) Generate rotation}\label{step-3-generate-rotation}}

Imagine holding the red axis it is fixed to the origin. As it is manipulated the projection back onto the projection plane correspondingly moves. This is what happens in a manual tour. For a radial tour, fix \(\theta\), the angle within the blue plane, and vary the sequence of \(\phi\), the angle coming out of the projection plane. Conceptually, live manipulation on a 2D plane allows the user to dynamically control these angles, effectively changing the coefficients of the manip var, which then performs a constrained rotation on the remaining variables.

For the demonstration of the radial tour, we define a sequence for \(\phi\) that brings the initial contribution of the manip var to be maximized and then zeroed before returning to the initial position.

\textbf{For } \(i\) \textbf{in 1 to n\_slides:}

Post-multiply the manipulation space by the pre-defined rotation matrix producing \textbf{RM}, the rotated manip space.

Let:

\begin{description}
  \item[$c_\theta$] be the cosine of $\theta$
  \item[$c_\phi$]   be the cosine of $\phi$
  \item[$s_\theta$] be the sine of   $\theta$
  \item[$s_\phi$]   be the sine of   $\phi$
\end{description}

then
\begin{align*}
  \textbf{RM}_{[p,~3,~i]}
  &= \textbf{M}_{[p,~3]} ~*~ \textbf{R}_{[3,~3]} \\
  &= \begin{bmatrix}
    M_{1,~1} & M_{1,~2} & M_{1,~3} \\
    M_{2,~1} & M_{2,~2} & M_{2,~3} \\
    \vdots   & \vdots   \\
    M_{p,~1} & M_{p,~2} & M_{p,~3}
  \end{bmatrix}_{[p,~3]}
    ~*~
  \begin{bmatrix}
    c_\theta^2 c_\phi s_\theta^2 &
    -c_\theta s_\theta (1 - c_\phi) &
    -c_\theta s_\phi \\
    -c_\theta s_\theta (1 - c_\phi) &
    s_\theta^2 c_\phi + c_\theta^2 &
    -s_\theta s_\phi \\
    c_\theta s_\phi &
    s_\theta s_\phi &
    c_\phi
  \end{bmatrix}_{[3,~3]}
\end{align*}

A note on application: compile the sequence of \(\phi_i\) and create an array/long table for each rotated manipulation space. \(\phi\) is the angle relative to the initial value of \(\phi\), we find the transformation \(\phi_i\) - \(\phi_1\) useful to think about \(\phi\) relative to the basis plane. Additionally, the value of \(\phi\) may be offset by a factor of pi. If the manip variable doesn't move as expected these are the first places to check.

\begin{Schunk}
\begin{Sinput}
for (phi in seq(seq_start, seq_end, phi_inc_sign)) {
  slide <- slide + 1
  tour[,, slide] <- rotate_manip_space(manip_space, theta, phi)[, 1:2]
}
\end{Sinput}
\end{Schunk}

Figure @ref(fig:step3) illustrates a sequence with 15 projected bases and highlight the manip variable on top while showing the corresponding projected data points on the bottom. Take note of how the changes in the manip var change the distance between the purple and green cluster of points, \texttt{aede2} is crucial in distinguishing between these groups. Tours are typically viewed as an animation such a dynamic version of this tour can be viewed online at \url{https://nspyrison.netlify.com/thesis/flea_manualtour_mvar5/}. The page may take a moment to load. The format of this figure and linking to an HTML animation will be used again in the Application, section \ref{sec:application}.

(ref:step3-cap) Radial manual tour changing the contributions from \texttt{aede2} of standardized flea data. The contributions increase from its initial contribution to a full contribution to the projection before decreasing to zero and then returning to its initial value. The change in the projected data shows that \texttt{aede2} is important for distinguishing between the purple and green clusters. An animated version can be viewed at \url{https://nspyrison.netlify.com/thesis/flea_manualtour_mvar5/}.

\begin{Schunk}
\begin{figure}

{\centering \includegraphics[width=6in,height=7.2in]{spinifex_paper_files/figure-latex/step3-1}

}

\caption[(ref:step3-cap)]{(ref:step3-cap)}(\#fig:step3)
\end{figure}
\end{Schunk}

Animations can be produced using the function \texttt{play\_manual\_tour()}. This function defaults to an HTML5 widget produced from \emph{plotly}.

\hypertarget{sec:display}{%
\subsection{Data in projection-space}\label{sec:display}}

In light of performance, the above operations are performed on the bases without the use of the larger datasets. After the bases are brought into the projection-space, however, it is helpful to observe them with data in the same space. Pre-multiply the data by basis frame bringing the data into the projection space.

\begin{align}
  \textbf{P}_{[n,~3]}
    &= \textbf{X}_{[n,~p]} ~*~ \textbf{RM}_{[p,~3]} \\
    &=
      \begin{bmatrix}
          X_{1,~1} & \dots & X_{1,~p} \\
          X_{2,~1} & \dots & X_{2,~p} \\
          \vdots   & \vdots & \vdots  \\
          X_{n,~1} & \dots & X_{n,~p}
      \end{bmatrix}_{[n,~p]}
      ~*~
      \begin{bmatrix}
        RM_{1,~1} & RM_{1,~2} & RM_{1,~3} \\
        RM_{2,~1} & RM_{2,~2} & RM_{2,~3} \\
        \vdots     & \vdots     & \vdots  \\
        RM_{p,~1} & RM_{p,~2} & RM_{p,~3}
      \end{bmatrix}_{[p,~3]}
\end{align}

For a 2D scatterplot, plot the first two variables from each frame statically as in the previous figure, or in sequence, producing an animated scatterplot. The remaining variable is sometimes linked to a data point aesthetic (such as size or color) to produce depth cues used in conjunction with the \(XY\) scatterplot.

\hypertarget{rendering-and-sharing}{%
\subsubsection{Rendering and sharing}\label{rendering-and-sharing}}

The \emph{tourr} package utilizes R's base graphics for the display of tours. \emph{spinifex} allows tours to be used in rendered in \emph{plotly} Sievert (2018) as an HTML5 object or \emph{gganimate} Pedersen and Robinson (2019) as .gif or .mp4 objects. Both of which build off \emph{ggplot2} objects in internal functions. Sharing of animations is not trivial especially in print and static formats such as .pdf. Even with the use of computers and dynamic file formats capturing the correct resolution, aspect, and display is challenging and many formats quickly bloat file sizes. Keep in mind hosting options and exporting functions from \emph{plotly}, \emph{gganimate} and \emph{tourr}.

\hypertarget{storage}{%
\subsubsection{Storage}\label{storage}}

Storing each data point for every frame of the animation is very inefficient. Just as operations are performed on the bases, so too should tour paths be stored as bases. Consider a radial manual tour, we can store the salient features in 3 bases, where \(\phi\) is at its starting, minimum, and maximum values. The frames in between can be interpolated by supplying angular speed. With the use of the \texttt{tourr::save\_history()} function, the target bases can be saved. From there geodesic interpolation can be used to populate the intermittent frames. This type of interpolation should not be used on manual tours, which have already been initialized into a 3D manipulation space where direct linear interpolation is appropriate.

\hypertarget{sec:application}{%
\subsection{Application}\label{sec:application}}

In a recent paper, Wang et al. (2018), the authors aggregate and visualize the sensitivity of hadronic experiments to ncleon structure. The authors introduce a new tool, PDFSense, to aid in the visualization of Parton distribution functions (PDF). The parameter-space of these experiments lies in 56 dimensions, \(\delta \in \mathbb{R}^{56}\), and are visualized as 3D subspaces of the 10 first principal components in linear (PCA) and non-linear (t-SNE) embeddings.

Using the same data, another study, Cook, Laa, and Valencia (2018), applies grand tours to the same subspaces. Grand tours are able to better resolve the distribution shape of clusters, intra-cluter detail, better outlier detection, and exonerate a claim persened from TFEP (TensorFlow embedded projections). Table 1 of Cook et al.~summarizes the key findings of observations made with PDFSense \& TFEP and those from grand tours.

Without getting too domain-specific the data has three primary groupings; DIS, VBP, and jet Each group is a particular class of experiments and each with many experimental datasets which inturn have many observation. Inconsideration of data density and business of the data We conduct manual tours on a subsets of the DIS and jet clusters. This explores the sensitivity of the structue to each of the variables in turn, and we present the subjectively best and worst manaul tour identifying structure in the respective data sets.

\hypertarget{jet-cluster}{%
\subsubsection{Jet cluster}\label{jet-cluster}}

The jet cluster is of interest as it contains the largest data sets and is found to be important in Wang et al. (2018) The jet cluster resides in a smaller dimensionality than the full set of experiments with 4 principal components explaining 95\% of the variation in the jet cluster (Cook, Laa, and Valencia 2018). The data is subset down to ATLAS7old and ATLAS7new to focus in on two groups with a reasonable number of observations that occupy different parts of the subspace. Below, we perform radial manual tours all four principal components within this scope. Visualizing PC3 and PC4 in figure @ref(fig:JetClusterGood) (more sturcturally insightful) and figure @ref(fig:JetClusterBad) (less sturcturally insightful) respectively, and list links to dynamic animation of all variables.

(ref:JetClusterGood-cap) Jet cluster, a radial manual tour of PC3. Colored by experiment type: `ATLAS7new' in green and `ATLAS7old' in orange. When PC3 fully contributes to the projection ATLAS7new (green) occupies unique space and several outliers are identifiable. Zeroing the contribution from PC3 to the projection hides the outliers and indeed all observations with ATLAS7new are contained within ATLAS7old (orange). A dynamic version can be viewed at \url{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/}.

\begin{Schunk}
\begin{figure}

{\centering \includegraphics[width=6in,height=7.2in]{spinifex_paper_files/figure-latex/JetClusterGood-1}

}

\caption[(ref:JetClusterGood-cap)]{(ref:JetClusterGood-cap)}(\#fig:JetClusterGood)
\end{figure}
\end{Schunk}

(ref:JetClusterBad-cap) Jet cluster, a radial manual tour of PC4. Colored by experiment type: ATLAS7new in green and ATLAS7old in orange. This manual tour contains less interesting information ATLAS7new (green) has points that are right and left of ATLAS7old, while most points occupy the same projection space, regardless of the contribution of PC4. A dynamic version can be viewed at \url{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/}.

\begin{Schunk}
\begin{figure}

{\centering \includegraphics[width=6in,height=7.2in]{spinifex_paper_files/figure-latex/JetClusterBad-1}

}

\caption[(ref:JetClusterBad-cap)]{(ref:JetClusterBad-cap)}(\#fig:JetClusterBad)
\end{figure}
\end{Schunk}

\hypertarget{todo-edit-application-grammarly-and-word-checks.}{%
\subsection{TODO: edit application, Grammarly and word checks.}\label{todo-edit-application-grammarly-and-word-checks.}}

Manipulating PC3, where varying the angle of rotation brings interesting features into and out of the center mass of the data, is more interesting than the manipulation of PC4, where the features are mostly independent of the contribution of PC4.

Jet cluster manual tours manipulating each of the principal components can be viewed from the links: \href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc1/}{PC1}, \href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc2/}{PC2}, \href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/}{PC3}, and \href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc4/}{PC4}.

\hypertarget{dis-cluster}{%
\subsubsection{DIS cluster}\label{dis-cluster}}

We perform a manual tour on this data, manipulating PC6 as depicted in figure @ref(fig:DISclusterGood). Looking at several frames we see that DIS HERA data lies mostly on a plane. When PC6 has full contributions, we see the dimuon SIDIS in purple is almost orthogonal to the DIS HERA (green). Yet the contribution of PC6 has zeroed the dimuon SIDIS data occupy the same space as the DIS HERA data. A dynamic version of this manual tour can be found at:
\url{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/}.
The page may some time to load, as the animation is several megabytes.

(ref:DISclusterGood-cap) DIS cluster, a radial manual tour of PC6. colored by experiment type: `DIS HERA1+2' in green, `dimuon SIDIS' in purple, and `charm SIDIS' in orange. When the contribution PC 6 is large we see that dimuon SIDIS (purple) data are nearly orthogonal to DIS HERA (green) data. As the projection is rotated, we can also see that DIS HERA (green) practically lies on a plane in this 6D subspace. When the contribution of PC6 is near zero, dimonSIDIS (purple) occupies the same space as the DIS HERA data. A dynamic version can be viewed at \url{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/}.

\begin{Schunk}
\begin{figure}

{\centering \includegraphics[width=6in,height=7.2in]{spinifex_paper_files/figure-latex/DISclusterGood-1}

}

\caption[(ref:DISclusterGood-cap)]{(ref:DISclusterGood-cap)}(\#fig:DISclusterGood)
\end{figure}
\end{Schunk}

The selection of the correct manip variable is important as the manipulation spaces convey different information. For example, in figure @ref(fig:DISclusterBad) we select PC2 as the manip variable finding it to be less insightful than PC6.

(ref:DISclusterBad-cap) DIS cluster, a radial manual tour of PC2. Colored by experiment type: `DIS HERA1+2' in green, `dimuon SIDIS' in purple, and `charm SIDIS' in orange. The structure of previously described plane of DIS HERA (green) and nearly orthogonal dimuon SIDIS (purple) is present, however, the manipulating PC2 does not give a head-on view of either, a less useful manual tour than that of PC6. A dynamic version can be viewed at \url{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc2/}.

\begin{Schunk}
\begin{figure}

{\centering \includegraphics[width=6in,height=7.2in]{spinifex_paper_files/figure-latex/DISclusterBad-1}

}

\caption[(ref:DISclusterBad-cap)]{(ref:DISclusterBad-cap)}(\#fig:DISclusterBad)
\end{figure}
\end{Schunk}

DIS cluster manual tours manipulating each of the principal components can be viewed from the links: \href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc1/}{PC1}, \href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc2/}{PC2}, \href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc3/}{PC3}, \href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc4/}{PC4}, \href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc5/}{PC5}, and \href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/}{PC6}.

\hypertarget{sec:usage}{%
\subsection{Source code and usage}\label{sec:usage}}

Use the below code as a guide for installation and finding the vignette. The vignette offers a less technical discussion opting to focus on code usage and goes through a couple more use cases. If you prefer to follow along with the example in the algorithm then simplified code is also listed below.

\begin{Schunk}
\begin{Sinput}
# devtools::install_github("nspyrison/spinifex") # Development version
install.package("spinifex")

# Also see vignette:
vignette("spinifex") # vignette ‘spinifex’ not found

## manual tour of std flea from holes-index:
library(spinifex)
f_dat  <- tourr::rescale(flea[,1:6])
f_cat  <- factor(flea$species)
f_path <- save_history(f_dat, guided_tour(holes()))
f_bas  <- matrix(f_path[,, max(dim(f_path)[3])], ncol=2)
f_mvar <- 5
f_lab  <- colnames(f_dat)

# View the basis
view_basis(f_bas, data = f_dat, labels = f_lab)
# View the manip space
view_manip_space(basis = f_bas, manip_var = f_mvar, labels = f_lab)
# Play animation as HTML5 from plotly
play_manual_tour(data = f_dat, basis = f_bas, manip_var = f_mvar, 
                 col = f_cat, angle = f_angle)
\end{Sinput}
\end{Schunk}

\hypertarget{acknowledgments}{%
\subsubsection{Acknowledgments}\label{acknowledgments}}

This article was created in \emph{R} (R Core Team 2018), using \emph{bookdown} (Xie 2016) and \emph{rmarkdown} (Xie, Allaire, and Grolemund 2018), with code generating the examples inline. The source files for this article be found at \href{https://github.com/nspyrison/confirmation/}{github.com/nspyrison/confirmation/}.
The source code for the \emph{spinifex} package can be found at \href{https://github.com/nspyrison/spinifex/}{github.com/nspyrison/spinifex/}.

\hypertarget{sec:discussion}{%
\subsection{Discussion}\label{sec:discussion}}

Tours, the dynamic linear projection of multivariate data, is an important aspect of data visualization extending the display of data-space as data dimensionality increases. This research has modified the algorithm producing manual tours, applied this functionality in \emph{R} and offers extends the graphics offerings that can be used to display tours. The paragraphs below explore how this work might be extended.

Future research on the algorithm would include extending it for use in 3D projections. The addition of another dimension theoretically allows for improved perception. This could explore interactions in immersive virtual reality or mixed reality, which may further allow for a better perception of structure and aid in higher-dimensional function visualization. Functions with many parameters suffer from the same dimensionality problem as data while their possible values lie on a plane of values rather than discrete points. Occulation, or the closer surface blocking further surfaces, will likely be an issue that may be alleviated by the use of wire mesh, changing opacity, or looking at sections of the projections (Furnas and Buja 1994).

The \emph{tourr} package provides many other geometric displays with the \texttt{tourr::display\_*()} family. These geometric options could be integrated into the \emph{ggplot2} framework for display on \emph{plotly} and \emph{gganimate}. Additionally, the \emph{animation} package Xie et al. (2018) could be implemented for another graphics framework. However, \emph{animation} builds from base graphs while \emph{spinifex} utilizes \emph{ggplot2} graphics.

The Givens rotations and Householder reflections as outlined in Buja et al. (2005) could also be added. Currently, Gram-Schmidt is the only form of frame interpolation used (not used in manual tours). In a Givens rotation, the \(x\) and \(y\) components (for example \(\theta~= 0,~pi/2\)) of the in-plane rotation are calculated separately and would be applied sequentially to produce the radial rotation. Householder reflections define reflection axes to project points on to the axes and generate rotations.

Having a script only interaction with tours causes a significant barrier to entry. To a lesser extent, \emph{plotly} offers some static interactions with the contained object, such as tooltips, brushing, and linking without communicating back to the R console. The development of a dynamic graphical user interface, perhaps with the use of a \emph{shiny} (Chang et al. 2018) application, would mitigate the barrier to entry, allow for more rapid analysis, and offer an approachable demo tool. The user could easily switch between variables to control, adjust interpolation step angle, or flag/save specific frame basis sets.

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-buja_computational_2005}{}%
Buja, Andreas, Dianne Cook, Daniel Asimov, and Catherine Hurley. 2005. ``Computational Methods for High-Dimensional Rotations in Data Visualization.'' In \emph{Handbook of Statistics}, 24:391--413. Elsevier. \url{https://doi.org/10.1016/S0169-7161(04)24014-7}.

\leavevmode\hypertarget{ref-chang_shiny:_2018}{}%
Chang, Winston, Joe Cheng, J. J. Allaire, Yihui Xie, and Jonathan McPherson. 2018. \emph{Shiny: Web Application Framework for R}. \url{https://CRAN.R-project.org/package=shiny}.

\leavevmode\hypertarget{ref-cook_manual_1997}{}%
Cook, Dianne, and Andreas Buja. 1997. ``Manual Controls for High-Dimensional Data Projections.'' \emph{Journal of Computational and Graphical Statistics} 6 (4): 464--80. \url{https://doi.org/10.2307/1390747}.

\leavevmode\hypertarget{ref-cook_grand_1995}{}%
Cook, Dianne, Andreas Buja, Javier Cabrera, and Catherine Hurley. 1995. ``Grand Tour and Projection Pursuit.'' \emph{Journal of Computational and Graphical Statistics} 4 (3): 155. \url{https://doi.org/10.2307/1390844}.

\leavevmode\hypertarget{ref-cook_dynamical_2018}{}%
Cook, Dianne, Ursula Laa, and German Valencia. 2018. ``Dynamical Projections for the Visualization of PDFSense Data.'' \emph{Eur. Phys. J. C} 78 (9): 742.

\leavevmode\hypertarget{ref-cook_interactive_2007}{}%
Cook, Dianne, Deborah F. Swayne, and A. Buja. 2007. \emph{Interactive and Dynamic Graphics for Data Analysis: With R and GGobi}. Springer Science \& Business Media.

\leavevmode\hypertarget{ref-furnas_prosection_1994}{}%
Furnas, George W., and Andreas Buja. 1994. ``Prosection Views: Dimensional Inference Through Sections and Projections.'' \emph{Journal of Computational and Graphical Statistics} 3 (4): 323--53. \url{https://doi.org/10.2307/1390897}.

\leavevmode\hypertarget{ref-lubischew_use_1962}{}%
Lubischew, Alexander A. 1962. ``On the Use of Discriminant Functions in Taxonomy.'' \emph{Biometrics}, 455--77.

\leavevmode\hypertarget{ref-pedersen_gganimate:_2019}{}%
Pedersen, Thomas Lin, and David Robinson. 2019. \emph{Gganimate: A Grammar of Animated Graphics}. \url{http://github.com/thomasp85/gganimate}.

\leavevmode\hypertarget{ref-r_core_team_r:_2018}{}%
R Core Team. 2018. \emph{R: A Language and Environment for Statistical Computing}. Vienna, Austria: R Foundation for Statistical Computing. \url{https://www.R-project.org/}.

\leavevmode\hypertarget{ref-sievert_plotly_2018}{}%
Sievert, Carson. 2018. \emph{Plotly for R}. \url{https://plotly-book.cpsievert.me}.

\leavevmode\hypertarget{ref-wang_mapping_2018}{}%
Wang, Bo-Ting, T. J. Hobbs, Sean Doyle, Jun Gao, Tie-Jiun Hou, Pavel M. Nadolsky, and Fredrick I. Olness. 2018. ``Mapping the Sensitivity of Hadronic Experiments to Nucleon Structure.'' \emph{Physical Review D} 98 (9): 094030.

\leavevmode\hypertarget{ref-wickham_tourr_2011}{}%
Wickham, Hadley, Dianne Cook, Heike Hofmann, and Andreas Buja. 2011. ``\textbf{Tourr} : An \emph{R} Package for Exploring Multivariate Data with Projections.'' \emph{Journal of Statistical Software} 40 (2). \url{https://doi.org/10.18637/jss.v040.i02}.

\leavevmode\hypertarget{ref-xie_bookdown:_2016}{}%
Xie, Yihui. 2016. \emph{Bookdown: Authoring Books and Technical Documents with R Markdown}. Boca Raton, Florida: Chapman; Hall/CRC. \url{https://github.com/rstudio/bookdown}.

\leavevmode\hypertarget{ref-xie_r_2018}{}%
Xie, Yihui, J. J. Allaire, and Garrett Grolemund. 2018. \emph{R Markdown: The Definitive Guide}. Boca Raton, Florida: Chapman; Hall/CRC. \url{https://bookdown.org/yihui/rmarkdown}.

\leavevmode\hypertarget{ref-xie_animation:_2018}{}%
Xie, Yihui, Christian Mueller, Lijia Yu, and Weicheng Zhu. 2018. \emph{Animation: A Gallery of Animations in Statistics and Utilities to Create Animations}. \url{https://yihui.name/animation}.

\bibliography{RJrefs.bib}

\address{%
Nicholas Spyrison\\
Monash University\\
line 1\\ line 2\\
}
\href{mailto:nicholas.spyrison@monash.edu}{\nolinkurl{nicholas.spyrison@monash.edu}}

\address{%
Dianne Cook\\
Monash University\\
line 1\\ line 2\\
}
\href{mailto:author2@work}{\nolinkurl{author2@work}}

### 2019-04-05 paper/spinifex_paper.Rmd
init commit. figure and caption references wrong.

---
title: spinifex; manual control of dynamic linear projections of high-dimensional data
author:
  - name: Nicholas Spyrison
    affiliation: Monash University 
    address:
    - line 1
    - line 2
    email:  nicholas.spyrison@monash.edu
  - name: Dianne Cook
    affiliation: Monash University 
    address:
    - line 1
    - line 2
    email:  author2@work
abstract: >
  The class of dynamic linear projections that are collectively known as 'tours' provide a unique dynamic visualization of numeric multivariate data. Tours are particularly useful for understanding the structure held within multivariate data, and in association with techniques for dimension reduction, supervised, and unsupervised classification. The *R* package *tourr* offers a variety of path generators and geometric displays for conducting tours. This paper discusses an extension package, *spinifex*, that adds support for the path generation of manual tours and extends the display of tours to use with the contemporary animation packages, *plotly* and *gganimate*. Manual tours are used to explore the sensitivity of structure as the contributions of a manipulation variable are changed. This particularly useful after identifying a feature of interest.

  A recent paper {@wang_mapping_2018} visualizes the sensitivey of the hadronic experiments to nucleon structure. Sensitivity was characterized in non-linear 3D embeddings of the first 10 principal components. This research applies manual tours to this data showing that manual tours resolves more structrual information that is orthogonal to the original viewing plane.
output:
  bookdown::pdf_book:
    base_format: rticles::rjournal_article
    includes:
      in_header: preamble.tex
bibliography: RJrefs.bib
---
\bibliography{RJrefs}
```{r setup_03, include=FALSE}
knitr::opts_chunk$set(
  fig.align = "center",
  echo = TRUE, 
  message = FALSE,
  warning = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "",
  cache = F, 
  cache.lazy = FALSE
)

library(spinifex)
library(ggplot2)

## FUNCTION FOR STATIC OUTPPUT, consider going to scalable facets.
array2static <- function(.m_tour, .data, .m_var, .cat, .mag =2.2)
{
  slides       <- array2df(array = .m_tour, data = .data)
  basis_slides <- slides$basis_slides
  data_slides  <- slides$data_slides
  n_slides     <- max(basis_slides$slide)
  if (n_slides != 15) 
    stop(paste0("stop: n_slides !=15 !!!!! n_slides = ", n_slides))

  # Initialize
  ## manip var asethetics
  p             <- nrow(basis_slides) / n_slides
  col_v         <- rep("grey80", p)
  col_v[.m_var] <- "blue"
  col_v         <- rep(col_v, n_slides)
  siz_v         <- rep(0.3, p)
  siz_v[.m_var] <- 1
  siz_v         <- rep(siz_v, n_slides)
  cat           <- rep(as.factor(.cat), n_slides)

  ## circle
  angle <- seq(0, 2 * pi, length = 180)
  circ  <- data.frame(c_x = cos(angle), c_y = sin(angle))
  circ[nrow(circ)+1, ] <- NA
  ## data asethetics
  data_slides <- data.frame(data_slides, cat = rep(.cat, n_slides))

  grid_b <- grid_t <-
    data.frame(slide = 1:n_slides, 
               x = .mag*rep(1:5, 3), y = .mag*rep(3:1, each = 5))
  grid_t$y <- grid_t$y + max(grid_t$y)
  # OUTER JOIN
  basis_grid <- merge(x = basis_slides, y = grid_t, by = "slide", all = TRUE)
  # CROSS JOIN
  circ_grid  <- merge(x = circ, y = grid_t, by = NULL)
  # OUTER JOIN
  data_grid  <- merge(x = data_slides, y = grid_b, by = "slide", all = TRUE)

  # BASIS
  gg1 <-
    ggplot(data = basis_grid) +
    # AXES LINE SEGMETNS
    geom_segment(aes(x = V1 + x, y = V2 + y, xend = x, yend = y),
                 color = col_v, size = siz_v) +
    # AXES TEXT LABELS
    geom_text(aes(x = V1 + x, y = V2 + y, label = lab_abbr),
              color = col_v, vjust = "outward", hjust = "outward") +
    # AXES FRAME NUM
    geom_text(aes(x = x - .7, y = y + 1.1, 
                  label = paste0("frame: ",slide)), color = "grey50") +
    # AXES CIRCLE PATH
    suppressWarnings( # Suppress for "Removed 1 rows containing missing values."
      geom_path(data = circ_grid, color = "grey80",
                mapping = aes(x = x+c_x, y = y+c_y))
    )

  # PROJECTION
  gg2 <- gg1 +
    # PROJ DATA POINTS
    geom_point(data = data_grid, size = .7,
               mapping = aes(x = V1 + x, y = V2 + y, color = cat),
               shape = as.integer(cat) + 15) +
    # PROJ DATA FRAME NUM
    geom_text(data = data_grid, color = "grey50",
              mapping = aes(x = x - .7, y = y + 1.1, 
                            label = paste0("frame: ",slide))) +
    theme_void() +
    scale_color_brewer(palette = "Dark2") + 
    # coord_fixed() +
    theme(legend.position="none",
          panel.border = element_rect(colour = "black", fill = NA))

  gg2
}
```

## Introduction

<!-- Tours -->
A tour is a multivariate data analysis technique in which a sequence of linear (orthogonal) projections are viewed as an animation while the orientation of the projection basis is rotated across time. Each frame of the sequence corresponds to a small change in the projection for a smooth transition that perseveres continuity.

<!-- Manual tour -->
While there are numerous methods that generate tour paths, this research focuses on the manual tour. The manual tour was described in @cook_manual_1997 and allows a user to control the projection coefficients of a select variable has in a 2D projection. The manipulation of these coefficients allows the analyst to explore how sensitive the projections structure is to these changes. This makes manual tours particularly useful once a feature of interest has been identified, for example, with the use of a guided tour [@cook_grand_1995]. The path of a guided tour is selected via projection pursuit, the optimization of an index function on the projection via a hill climbing algorithm. This allows guided tours to identify interesting projection features rapidly given the relatively large parameter-space. Once the given projection has been provided, it is time to define the path of rotation.

<!-- Manual tour paths -->
Ideally, the path would be intuitively user-generated from physical movement, be it through mouse or motion capture. Unfortunately, this type of dynamic control has proven difficult to capture for in R. Because of this, manual tours were not implemented in *tourr*. This research allows for the consumption, but not the generation, of such dynamic input. After the capture of an oblique user motion, the rotation needs to be applied to step 3 (rotation sequence) of the algorithm discussed below. In the section below we stick with a radial rotation where, $\theta$, the angle of in-projection-plane rotation is held constant.

<!-- Display -->
Spinifex utilizes two new animation packages, *plotly* [@sievert_plotly_2018] and *gganimate* [@pedersen_gganimate:_2019], to display tours, manual or other saved tours. From a given projection, the user can choose which variable to control, and the animation sequence is generated to remove the variable from the projection, and then extend its contribution to be the sole variable in one direction. This allows the viewer to assess the change in structure induced in the projection by the variable's contribution.

<!-- Paper outline -->
The paper is organized as follows. Section \ref{sec:algorithm} explains the algorithm using a toy dataset. Section \ref{sec:display} discussed the display of the animation after the path has been generated. Section \ref{sec:application} illustrates how this can be used for sensitivity analysis applied to contemporary high energy physics. The last section, \ref{sec:discussion} summarizes the work and discusses future research.

<!-- Algorithm outline -->
## Algorithm {#sec:algorithm}

The section below describes the algorithm for performing a 2D radial manual tour:

1. Provided with a 2D projection, choose a variable to explore. This is called the "manip" variable.
2. Create a 3D manipulation space, where the manip variable has the full contribution.
3. Generate a rotation sequence which increases the norm of the coefficient to 1 and zeros it.

The steps are described in more detail below. The R functions used below mentioned briefly, but more complete code example can be found in section \ref{sec:usage}

### Notation

This section describes the notation used in the algorithm for a 2D radial manual tour.

<!-- remember mixing latex and rmarkdown bullets isn't trivial. -->
\begin{itemize}
  \item $\textbf{X}$, the data, an $n \times p$ numeric matrix to be embedded in two dimensions.
  \item $\textbf{B} = (B_1,~ B_2)$, any orthonormal projection basis set, $p \times 2$ matrix, describing the projection from $p$ to two dimensions.
  \item $\textbf{e}$, a zero column vector of length $p$ with the $k-$th element set to one, where $k$ is the number of the variable to manipulate.
  \item $\theta$, the angle of in-projection-plane rotation, for example, on the reference axes. 
  \item $\phi$, the angle of out-of-projection-plane rotation, coming into the manipulation space.
\end{itemize}

The algorithm primarily operates on the projection basis and utilizes the data only when making a display. The projection space can be viewed at any point in the process by pre-multiplying the data and plotting the first 2 variables.

### Toy data set

<!-- describe data. -->
The flea data, originally from @lubischew_use_1962, available in the R package *tourr* [@wickham_tourr_2011] is used to illustrate the algorithm. The data contains 74 observations across 6 variables, physical measurements of the flea beetles. Each observation belonging to one of three species.

<!-- Projection basis -->
The data is defined. A basis set (ideally that views an interesting feature) should be provided to explore the sensitivity of the variables to the structure. To identify a projection containing an interesting feature, apply
a guided tour[@cook_interactive_2007] on the flea data. In a guided tour the projection sequence is selected by optimizing an index via hill-climbing. In this case, the holes index is selected. The holes index is maximized by when the projected observations are furthest from the center. Figure \@ref(fig:step0) shows a locally optimized projection for this data. The left plot displays the reference axes of the projection basis, a visual indication of the magnitude and direction each variable contributed to the projections. The right plot shows the projection of the data through the basis set described by the reference axes (left). Data points are colored and given point characters according to the species of the flea (the guided tour was unsupervised with this information).

```{r step0, echo=F, results='hide', out.width='95%', fig.cap = "(ref:step0-cap)"}
## Flea holes tour
f_dat  <- tourr::rescale(flea[,1:6])
f_cat  <- factor(flea$species)
f_path <- save_history(f_dat, guided_tour(holes()))
f_bas  <- matrix(f_path[,, max(dim(f_path)[3])], ncol=2)
f_mvar <- 5
f_msp  <- create_manip_space(basis = f_bas, manip_var = f_mvar)
f_proj <- data.frame(tourr::rescale(f_dat %*% f_msp[, 1:2]))

# step0, output
# Adjust centering to make axes and data side by side
view_basis(f_bas, labels = colnames(f_dat)) +
  geom_point(data = f_proj,
             mapping = aes(x = X1 + .75, y = X2 - .5, color = f_cat),
             pch = as.integer(f_cat) + 15) +
  theme(panel.border = element_rect(colour = "black", fill = NA))
```

(ref:step0-cap) Basis reference axes (left) and projected data (right) of standardized flea data. Data points color and shape are mapped to beetle species. Basis identified by a holes-index guided tour. The variables `aede2` and `tars1` contribute mostly orthogonal to the other variables. We'll select `aede2` as our manipulation variable to see how the structure of the projection changes as we rotate `aede2` into and out of the projection.

Call `view_basis()` on a basis to produce a *ggplot2* graphic similar to \@ref(fig:step0). Projection space is always available for display via the matrix multiplication $\textbf{X}_{[n,~p]} ~*~ \textbf{B}_{[p,~d]} ~=~ \textbf{P}_{[n,~d]}$.

### Step 1) Choose variable of interest

<!-- select a manip var-->
In figure \@ref(fig:step0), above, the contributions of the variables `tars1` and `aede2` are mostly orthogonal to the contributions of the other four variables. These two variables explain the variation of the data between the purple and green species. We select `aede2` as the manip var, the variable to be manipulated as it typically has a larger contribution after the optimizing the holes index. The question that will be explored in the explanation of the algorithm is how important the variable `aede2` is to the separation of the clusters.

### Step 2) Create the manip space

<!-- Zero Vect, manip sp -->
Initialize a zero vector $e$ of $p$ elements. Because `aede2` is the fifth variable in the data, set the $k=5$-th element to one giving the manip var a full contribution in this dimension. Use the Gram-Schmidt process to orthonormalize the zero vector onto the basis yielding the 3D manipulation space, **M**.

\begin{align*}
  \textbf{e} &\leftarrow Orthonormalize_{GS}(\textbf{e}) w.r.t. Basis \\
  &= \textbf{e} - \langle \textbf{e},\textbf{B}_1 \rangle \textbf{B}_1 - \langle \textbf{e}, \textbf{B}_2 \rangle \textbf{B}_2 \\
  \\
  \textbf{M}_{[p,~3]} &= (\textbf{B}_1,\textbf{B}_2,\textbf{e})
\end{align*}

<!-- What the manip space facilitates -->
Adding this extra dimension to our basis plane allows for the coefficients of the specified variable to be changed. For example, the ability to lift a piece of paper, rather than being constrained to the motion on a table top. Orthonormalizing rescales the new depth vector while the projection down to 2D is the original basis, that is the first $d$ vectors remain constant. Imagine the reference axes (and projection plane) laying flat on a table, while a new dimension exists with axes projecting back onto the reference axes. An illustration of such can be seen below in figure \@ref(fig:step2). The manip var is highlighted, while the depths of the other variables are not depicted.

(ref:step2-cap) Manipulation space for controlling the contribution of `aede2` of standardized flea data. Basis selected by a holes-index guided tour. The Projection plane is shown in blue. The manipulation axis, in red, allows the coefficients of the manip var to be changed.

```{r step2, echo=F, out.width='100%', fig.cap = "(ref:step2-cap)"}
view_manip_space(basis = f_bas, manip_var = f_mvar, labels = colnames(f_dat)) +
    theme(panel.border = element_rect(colour = "black", fill = NA))
```

The representation in \@ref(fig:step2) can be duplicated by calling the function `view_manip_space()`.

### Step 3) Generate rotation

<!-- illustration of axis manip -->
Imagine holding the red axis it is fixed to the origin. As it is manipulated the projection back onto the projection plane correspondingly moves. This is what happens in a manual tour. For a radial tour, fix $\theta$, the angle within the blue plane, and vary the sequence of $\phi$, the angle coming out of the projection plane. Conceptually, live manipulation on a 2D plane allows the user to dynamically control these angles, effectively changing the coefficients of the manip var, which then performs a constrained rotation on the remaining variables.

<!-- describe manip var path -->
For the demonstration of the radial tour, we define a sequence for $\phi$ that brings the initial contribution of the manip var to be maximized and then zeroed before returning to the initial position.

**For ** $i$ **in 1 to n_slides:**
<!-- don't mix rmd and latex... -->

Post-multiply the manipulation space by the pre-defined rotation matrix producing **RM**, the rotated manip space.

Let:
\begin{description}
  \item[$c_\theta$] be the cosine of $\theta$
  \item[$c_\phi$]   be the cosine of $\phi$
  \item[$s_\theta$] be the sine of   $\theta$
  \item[$s_\phi$]   be the sine of   $\phi$
\end{description}

then
\begin{align*}
  \textbf{RM}_{[p,~3,~i]}
  &= \textbf{M}_{[p,~3]} ~*~ \textbf{R}_{[3,~3]} \\
  &= \begin{bmatrix}
    M_{1,~1} & M_{1,~2} & M_{1,~3} \\
    M_{2,~1} & M_{2,~2} & M_{2,~3} \\
    \vdots   & \vdots   \\
    M_{p,~1} & M_{p,~2} & M_{p,~3}
  \end{bmatrix}_{[p,~3]}
    ~*~
  \begin{bmatrix}
    c_\theta^2 c_\phi s_\theta^2 &
    -c_\theta s_\theta (1 - c_\phi) &
    -c_\theta s_\phi \\
    -c_\theta s_\theta (1 - c_\phi) &
    s_\theta^2 c_\phi + c_\theta^2 &
    -s_\theta s_\phi \\
    c_\theta s_\phi &
    s_\theta s_\phi &
    c_\phi
  \end{bmatrix}_{[3,~3]}
\end{align*}

<!--  phi transform and application note -->
A note on application: compile the sequence of $\phi_i$ and create an array/long table for each rotated manipulation space. $\phi$ is the angle relative to the initial value of $\phi$, we find the transformation $\phi_i$ - $\phi_1$ useful to think about $\phi$ relative to the basis plane. Additionally, the value of $\phi$ may be offset by a factor of pi. If the manip variable doesn't move as expected these are the first places to check.

```{r STEP3CODEEXAMPLE, eval=F, echo=T}
for (phi in seq(seq_start, seq_end, phi_inc_sign)) {
  slide <- slide + 1
  tour[,, slide] <- rotate_manip_space(manip_space, theta, phi)[, 1:2]
}
```

<!-- description of slides, on the technical side we haven't gone back to data space yet. -->
Figure \@ref(fig:step3) illustrates a sequence with 15 projected bases and highlight the manip variable on top while showing the corresponding projected data points on the bottom. Take note of how the changes in the manip var change the distance between the purple and green cluster of points, `aede2` is crucial in distinguishing between these groups. Tours are typically viewed as an animation such a dynamic version of this tour can be viewed online at https://nspyrison.netlify.com/thesis/flea_manualtour_mvar5/. The page may take a moment to load. The format of this figure and linking to an HTML animation will be used again in the Application, section \ref{sec:application}.

<!-- step3 caption -->
(ref:step3-cap) Radial manual tour changing the contributions from `aede2` of standardized flea data. The contributions increase from its initial contribution to a full contribution to the projection before decreasing to zero and then returning to its initial value. The change in the projected data shows that `aede2` is important for distinguishing between the purple and green clusters. An animated version can be viewed at https://nspyrison.netlify.com/thesis/flea_manualtour_mvar5/.

```{r step3, echo=F, warning=F, fig.height=7.2, fig.width=6, out.height='7.2in', out.width='6in', fig.cap = "(ref:step3-cap)"}
f_angle <- .26
f_mtour <- manual_tour(f_bas, manip_var = f_mvar, angle = f_angle)
if(dim(f_mtour)[3] != 15) message(
  paste0("!!!!! step3 n_slides != 15 !!!!! n_slides = ", dim(f_mtour)[3]))

#play_manual_tour(f_dat, f_bas, f_mvar, col = f_cat, angle = f_angle)
array2static(.m_tour = f_mtour, .data = f_dat, .m_var = f_mvar, .cat = f_cat)
```

Animations can be produced using the function `play_manual_tour()`. This function defaults to an HTML5 widget produced from *plotly*.

## Data in projection-space {#sec:display}

<!-- reminder of basis operation and now apply data -->
In light of performance, the above operations are performed on the bases without the use of the larger datasets. After the bases are brought into the projection-space, however, it is helpful to observe them with data in the same space. Pre-multiply the data by basis frame bringing the data into the projection space.

\begin{align}
  \textbf{P}_{[n,~3]}
    &= \textbf{X}_{[n,~p]} ~*~ \textbf{RM}_{[p,~3]} \\
    &=
      \begin{bmatrix}
          X_{1,~1} & \dots & X_{1,~p} \\
          X_{2,~1} & \dots & X_{2,~p} \\
          \vdots   & \vdots & \vdots  \\
          X_{n,~1} & \dots & X_{n,~p}
      \end{bmatrix}_{[n,~p]}
      ~*~
      \begin{bmatrix}
        RM_{1,~1} & RM_{1,~2} & RM_{1,~3} \\
        RM_{2,~1} & RM_{2,~2} & RM_{2,~3} \\
        \vdots     & \vdots     & \vdots  \\
        RM_{p,~1} & RM_{p,~2} & RM_{p,~3}
      \end{bmatrix}_{[p,~3]}
\end{align}

<!-- plot 2 var in seq for animation -->
For a 2D scatterplot, plot the first two variables from each frame statically as in the previous figure, or in sequence, producing an animated scatterplot. The remaining variable is sometimes linked to a data point aesthetic (such as size or color) to produce depth cues used in conjunction with the $XY$ scatterplot.

### Rendering and sharing

The *tourr* package utilizes R's base graphics for the display of tours. *spinifex* allows tours to be used in rendered in *plotly* @sievert_plotly_2018 as an HTML5 object or *gganimate* @pedersen_gganimate:_2019 as .gif or .mp4 objects. Both of which build off *ggplot2* objects in internal functions. Sharing of animations is not trivial especially in print and static formats such as .pdf. Even with the use of computers and dynamic file formats capturing the correct resolution, aspect, and display is challenging and many formats quickly bloat file sizes. Keep in mind hosting options and exporting functions from *plotly*, *gganimate* and *tourr*.

### Storage

Storing each data point for every frame of the animation is very inefficient. Just as operations are performed on the bases, so too should tour paths be stored as bases. Consider a radial manual tour, we can store the salient features in 3 bases, where $\phi$ is at its starting, minimum, and maximum values. The frames in between can be interpolated by supplying angular speed. With the use of the `tourr::save_history()` function, the target bases can be saved. From there geodesic interpolation can be used to populate the intermittent frames. This type of interpolation should not be used on manual tours, which have already been initialized into a 3D manipulation space where direct linear interpolation is appropriate.

## Application {#sec:application}

<!-- Introduction of data and original paper  -->
In a recent paper, @wang_mapping_2018, the authors aggregate and visualize the sensitivity of hadronic experiments to ncleon structure. The authors introduce a new tool, PDFSense, to aid in the visualization of Parton distribution functions (PDF). The parameter-space of these experiments lies in 56 dimensions, $\delta \in \mathbb{R}^{56}$, and are visualized as 3D subspaces of the 10 first principal components in linear (PCA) and non-linear (t-SNE) embeddings.

<!-- grand tours on the same data, summarize findings better -->
Using the same data, another study, @cook_dynamical_2018, applies grand tours to the same subspaces. Grand tours are able to better resolve the distribution shape of clusters, intra-cluter detail, better outlier detection, and exonerate a claim persened from TFEP (TensorFlow embedded projections). Table 1 of Cook et al. summarizes the key findings of observations made with PDFSense & TFEP and those from grand tours.

<!-- is this the best place for this level of detail or is it suited better later on? -->
Without getting too domain-specific the data has three primary groupings; DIS, VBP, and jet Each group is a particular class of experiments and each with many experimental datasets which inturn have many observation. Inconsideration of data density and business of the data We conduct manual tours on a subsets of the DIS and jet clusters. This explores the sensitivity of the structue to each of the variables in turn, and we present the subjectively best and worst manaul tour identifying structure in the respective data sets.

### Jet cluster

<!-- motivate jet cluster, explain dimensionality  -->
The jet cluster is of interest as it contains the largest data sets and is found to be important in @wang_mapping_2018 The jet cluster resides in a smaller dimensionality than the full set of experiments with 4 principal components explaining 95% of the variation in the jet cluster [@cook_dynamical_2018]. The data is subset down to ATLAS7old and ATLAS7new to focus in on two groups with a reasonable number of observations that occupy different parts of the subspace. Below, we perform radial manual tours all four principal components within this scope. Visualizing PC3 and PC4 in figure \@ref(fig:JetClusterGood) (more sturcturally insightful) and figure \@ref(fig:JetClusterBad) (less sturcturally insightful) respectively, and list links to dynamic animation of all variables.

(ref:JetClusterGood-cap) Jet cluster, a radial manual tour of PC3. Colored by experiment type: 'ATLAS7new' in green and 'ATLAS7old' in orange. When PC3 fully contributes to the projection ATLAS7new (green) occupies unique space and several outliers are identifiable. Zeroing the contribution from PC3 to the projection hides the outliers and indeed all observations with ATLAS7new are contained within ATLAS7old (orange). A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/.

```{r JetClusterGood, echo=F, warning=F, fig.height=7.2, fig.width=6, out.height='7.2in', out.width='6in', fig.cap="(ref:JetClusterGood-cap)"}
# Jet cluster, Fig 7 of cook_dynamical_2018, subset to ATLAS7old and ATLAS7new
load("./data/JetCluster_sub.rda")
load("./data/JetCluster_basis.rda")

jet_bas   <- JetCluster_basis
jet_dat   <- tourr::rescale(JetCluster_sub[, 1:4])
jet_cat   <- factor(JetCluster_sub$exp)
jet_mvar  <- 4
jet_ang   <- .26
jet_mtour <- manual_tour(basis = jet_bas, manip_var = jet_mvar, angle = jet_ang)
#dim(jet_mtour)

array2static(.m_tour = jet_mtour, .data = jet_dat,
             .m_var = jet_mvar, .cat = jet_cat)
# play_manual_tour(data = jet_dat, basis = jet_bas, manip_var = jet_mvar,
#                  angle = jet_ang, axes = "bottomleft", col = jet_cat)

##M_VAR, RESULT
# 1, good
# 2, poor
# 3, *worst
# 4, *best
```

(ref:JetClusterBad-cap) Jet cluster, a radial manual tour of PC4. Colored by experiment type: ATLAS7new in green and ATLAS7old in orange. This manual tour contains less interesting information ATLAS7new (green) has points that are right and left of ATLAS7old, while most points occupy the same projection space, regardless of the contribution of PC4. A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/.

```{r JetClusterBad, echo=F, warning=F, fig.height=7.2, fig.width=6, out.height='7.2in', out.width='6in', fig.cap = "(ref:JetClusterBad-cap)"}
# Jet cluster, Fig 7 of cook_dynalical_2018, subset to ATLAS7old and ATLAS7new
jet_mvar  <- 3
jet_mtour <- manual_tour(basis = jet_bas, manip_var = jet_mvar, angle = jet_ang)

array2static(.m_tour = jet_mtour, .data = jet_dat,
             .m_var = jet_mvar, .cat = jet_cat)
```

## TODO: edit application, Grammarly and word checks.
<!-- disscusion of findings and which is better -->
Manipulating PC3, where varying the angle of rotation brings interesting features into and out of the center mass of the data, is more interesting than the manipulation of PC4, where the features are mostly independent of the contribution of PC4.

Jet cluster manual tours manipulating each of the principal components can be viewed from the links: [PC1](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc1/), [PC2](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc2/), [PC3](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/), and [PC4](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc4/).

### DIS cluster

We perform a manual tour on this data, manipulating PC6 as depicted in figure \@ref(fig:DISclusterGood). Looking at several frames we see that DIS HERA data lies mostly on a plane. When PC6 has full contributions, we see the dimuon SIDIS in purple is almost orthogonal to the DIS HERA (green). Yet the contribution of PC6 has zeroed the dimuon SIDIS data occupy the same space as the DIS HERA data. A dynamic version of this manual tour can be found at:
https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/.
The page may some time to load, as the animation is several megabytes.

(ref:DISclusterGood-cap) DIS cluster, a radial manual tour of PC6. colored by experiment type: 'DIS HERA1+2' in green, 'dimuon SIDIS' in purple, and 'charm SIDIS' in orange. When the contribution PC 6 is large we see that dimuon SIDIS (purple) data are nearly orthogonal to DIS HERA (green) data. As the projection is rotated, we can also see that DIS HERA (green) practically lies on a plane in this 6D subspace. When the contribution of PC6 is near zero, dimonSIDIS (purple) occupies the same space as the DIS HERA data. A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/.

```{r DISclusterGood, echo=F, warning=F, fig.height=7.2, fig.width=6, out.height='7.2in', out.width='6in', fig.cap = "(ref:DISclusterGood-cap)"}
##DIS cluster, fig 8 of cook_dynamical_2018
load("./data/grDIScenter.rda")
load("./data/DIScluster_centered_basis.rda")
DIS_bas   <- DIScluster_centered_basis
DIS_dat   <- tourr::rescale(grDIScenter[, 1:6])
DIS_cat   <- factor(grDIScenter$disID)
DIS_mvar  <- 6
DIS_ang   <- .27
DIS_mtour <- manual_tour(basis = DIS_bas, manip_var = DIS_mvar, angle = DIS_ang)

array2static(.m_tour = DIS_mtour, .data = DIS_dat,
             .m_var = DIS_mvar, .cat = DIS_cat)
# play_manual_tour(data = DIS_dat, basis = DIS_bas, manip_var = DIS_mvar,
#                  angle = DIS_ang, axes="bottomleft", col = DIS_cat)

##M_VAR, RESULT
# 1, purple jet
# 2, *poor
# 3, black
# 4, plane
# 5, black and plane
# 6, best*purple and plane
```

The selection of the correct manip variable is important as the manipulation spaces convey different information. For example, in figure \@ref(fig:DISclusterBad) we select PC2 as the manip variable finding it to be less insightful than PC6.

(ref:DISclusterBad-cap) DIS cluster, a radial manual tour of PC2. Colored by experiment type: 'DIS HERA1+2' in green, 'dimuon SIDIS' in purple, and 'charm SIDIS' in orange. The structure of previously described plane of DIS HERA (green) and nearly orthogonal dimuon SIDIS (purple) is present, however, the manipulating PC2 does not give a head-on view of either, a less useful manual tour than that of PC6. A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/discluster_manualtour_pc2/.

```{r DISclusterBad, echo=F, warning=F, fig.height=7.2, fig.width=6, out.height='7.2in', out.width='6in', fig.cap = "(ref:DISclusterBad-cap)"}
##DIS cluster, fig 8 of cook_dynamical_2018
DIS_mvar  <- 2
DIS_mtour <- manual_tour(basis = DIS_bas, manip_var = DIS_mvar, angle = DIS_ang)

array2static(.m_tour = DIS_mtour, .data = DIS_dat,
             .m_var = DIS_mvar, .cat = DIS_cat)
```

DIS cluster manual tours manipulating each of the principal components can be viewed from the links: [PC1](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc1/), [PC2](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc2/), [PC3](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc3/), [PC4](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc4/), [PC5](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc5/), and [PC6](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/).

## Source code and usage {#sec:usage}

Use the below code as a guide for installation and finding the vignette. The vignette offers a less technical discussion opting to focus on code usage and goes through a couple more use cases. If you prefer to follow along with the example in the algorithm then simplified code is also listed below.

```{r INSTALLCODEEXAMPLE, eval=F, echo=T}
# devtools::install_github("nspyrison/spinifex") # Development version
install.package("spinifex")

# Also see vignette:
vignette("spinifex") # vignette ‘spinifex’ not found

## manual tour of std flea from holes-index:
library(spinifex)
f_dat  <- tourr::rescale(flea[,1:6])
f_cat  <- factor(flea$species)
f_path <- save_history(f_dat, guided_tour(holes()))
f_bas  <- matrix(f_path[,, max(dim(f_path)[3])], ncol=2)
f_mvar <- 5
f_lab  <- colnames(f_dat)

# View the basis
view_basis(f_bas, data = f_dat, labels = f_lab)
# View the manip space
view_manip_space(basis = f_bas, manip_var = f_mvar, labels = f_lab)
# Play animation as HTML5 from plotly
play_manual_tour(data = f_dat, basis = f_bas, manip_var = f_mvar, 
                 col = f_cat, angle = f_angle)
```

### Acknowledgments

This article was created in *R* [@r_core_team_r:_2018], using *bookdown* [@xie_bookdown:_2016] and *rmarkdown* [@xie_r_2018], with code generating the examples inline. The source files for this article be found at [github.com/nspyrison/confirmation/](https://github.com/nspyrison/confirmation/).
The source code for the *spinifex* package can be found at [github.com/nspyrison/spinifex/](https://github.com/nspyrison/spinifex/).

## Discussion {#sec:discussion}

<!-- summary and seague  -->
Tours, the dynamic linear projection of multivariate data, is an important aspect of data visualization extending the display of data-space as data dimensionality increases. This research has modified the algorithm producing manual tours, applied this functionality in *R* and offers extends the graphics offerings that can be used to display tours. The paragraphs below explore how this work might be extended.

<!-- 3D and function vis -->
Future research on the algorithm would include extending it for use in 3D projections. The addition of another dimension theoretically allows for improved perception. This could explore interactions in immersive virtual reality or mixed reality, which may further allow for a better perception of structure and aid in higher-dimensional function visualization. Functions with many parameters suffer from the same dimensionality problem as data while their possible values lie on a plane of values rather than discrete points. Occulation, or the closer surface blocking further surfaces, will likely be an issue that may be alleviated by the use of wire mesh, changing opacity, or looking at sections of the projections [@furnas_prosection_1994].

<!-- geometric and graphics display -->
The *tourr* package provides many other geometric displays with the `tourr::display_*()` family. These geometric options could be integrated into the *ggplot2* framework for display on *plotly* and *gganimate*. Additionally, the *animation* package @xie_animation:_2018 could be implemented for another graphics framework. However, *animation* builds from base graphs while *spinifex* utilizes *ggplot2* graphics.

<!-- interpolation -->
The Givens rotations and Householder reflections as outlined in @buja_computational_2005 could also be added. Currently, Gram-Schmidt is the only form of frame interpolation used (not used in manual tours). In a Givens rotation, the $x$ and $y$ components (for example $\theta~= 0,~pi/2$) of the in-plane rotation are calculated separately and would be applied sequentially to produce the radial rotation. Householder reflections define reflection axes to project points on to the axes and generate rotations.

<!-- dynamic interaction/gui -->
Having a script only interaction with tours causes a significant barrier to entry. To a lesser extent, *plotly* offers some static interactions with the contained object, such as tooltips, brushing, and linking without communicating back to the R console. The development of a dynamic graphical user interface, perhaps with the use of a *shiny* [@chang_shiny:_2018] application, would mitigate the barrier to entry, allow for more rapid analysis, and offer an approachable demo tool. The user could easily switch between variables to control, adjust interpolation step angle, or flag/save specific frame basis sets.

### 2019-04-05 paper/spinifex-paper.tex
wrestling references

@@
 % !TeX root = RJwrapper.tex
-\title{spinifex; manual control of dynamic linear projections of high-dimensional data}
+\title{spinifex: manual control of dynamic linear projections of high-dimensional data}
 \author{by Nicholas Spyrison, Dianne Cook}
@@
 
-A tour is a multivariate data analysis technique in which a sequence of linear (orthogonal) projections are viewed as an animation while the orientation of the projection basis is rotated across time.
-Each frame of the sequence corresponds to a small change in the projection for a smooth transition that perseveres continuity.
+--\textgreater{}
 
-While there are numerous methods that generate tour paths, this research focuses on the manual tour.
-The manual tour was described in Cook and Buja (1997) and allows a user to control the projection coefficients of a select variable has in a 2D projection.
-The manipulation of these coefficients allows the analyst to explore how sensitive the projections structure is to these changes.
-This makes manual tours particularly useful once a feature of interest has been identified, for example, with the use of a guided tour (Cook et al.
-1995).
-The path of a guided tour is selected via projection pursuit, the optimization of an index function on the projection via a hill climbing algorithm.
-This allows guided tours to identify interesting projection features rapidly given the relatively large parameter-space.
-Once the given projection has been provided, it is time to define the path of rotation.
+--\textgreater{}
 
-Ideally, the path would be intuitively user-generated from physical movement, be it through mouse or motion capture.
-Unfortunately, this type of dynamic control has proven difficult to capture for in R.
-Because of this, manual tours were not implemented in \emph{tourr}.
-This research allows for the consumption, but not the generation, of such dynamic input.
-After the capture of an oblique user motion, the rotation needs to be applied to step 3 (rotation sequence) of the algorithm discussed below.
-In the section below we stick with a radial rotation where, \(\theta\), the angle of in-projection-plane rotation is held constant.
+--\textgreater{}
 
-Spinifex utilizes two new animation packages, \emph{plotly} (Sievert 2018) and \emph{gganimate} (Pedersen and Robinson 2019), to display tours, manual or other saved tours.
-From a given projection, the user can choose which variable to control, and the animation sequence is generated to remove the variable from the projection, and then extend its contribution to be the sole variable in one direction.
-This allows the viewer to assess the change in structure induced in the projection by the variable's contribution.
+--\textgreater{}
 
-The paper is organized as follows.
-Section \ref{sec:algorithm} explains the algorithm using a toy dataset.
-Section \ref{sec:display} discussed the display of the animation after the path has been generated.
-Section \ref{sec:application} illustrates how this can be used for sensitivity analysis applied to contemporary high energy physics.
-The last section, \ref{sec:discussion} summarizes the work and discusses future research.
+--\textgreater{}
 
-\hypertarget{sec:algorithm}{%
-\subsection{Algorithm}\label{sec:algorithm}}
+--\textgreater{}
 
-The section below describes the algorithm for performing a 2D radial manual tour:
+--\textgreater{}
 
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Provided with a 2D projection, choose a variable to explore. This is called the ``manip'' variable.
-\item
-  Create a 3D manipulation space, where the manip variable has the full contribution.
-\item
-  Generate a rotation sequence which increases the norm of the coefficient to 1 and zeros it.
-\end{enumerate}
-
-The steps are described in more detail below. The R functions used below mentioned briefly, but more complete code example can be found in section \ref{sec:usage}
-
-\hypertarget{notation}{%
-\subsubsection{Notation}\label{notation}}
-
-This section describes the notation used in the algorithm for a 2D radial manual tour.
-
-\begin{itemize}
-  \item $\textbf{X}$, the data, an $n \times p$ numeric matrix to be embedded in two dimensions.
-  \item $\textbf{B} = (B_1,~ B_2)$, any orthonormal projection basis set, $p \times 2$ matrix, describing the projection from $p$ to two dimensions.
-  \item $\textbf{e}$, a zero column vector of length $p$ with the $k-$th element set to one, where $k$ is the number of the variable to manipulate.
-  \item $\theta$, the angle of in-projection-plane rotation, for example, on the reference axes. 
-  \item $\phi$, the angle of out-of-projection-plane rotation, coming into the manipulation space.
-\end{itemize}
-
-The algorithm primarily operates on the projection basis and utilizes the data only when making a display.
-The projection space can be viewed at any point in the process by pre-multiplying the data and plotting the first 2 variables.
-
-\hypertarget{toy-data-set}{%
-\subsubsection{Toy data set}\label{toy-data-set}}
-
-The flea data, originally from Lubischew (1962), available in the R package \emph{tourr} (Wickham et al.
-2011) is used to illustrate the algorithm.
-The data contains 74 observations across 6 variables, physical measurements of the flea beetles.
-Each observation belonging to one of three species.
+--\textgreater{}
 
@@
 The holes index is maximized by when the projected observations are furthest from the center.
-Figure @ref(fig:step0) shows a locally optimized projection for this data.
+Figure @ref\{fig:step0\} shows a locally optimized projection for this data.
 The left plot displays the reference axes of the projection basis, a visual indication of the magnitude and direction each variable contributed to the projections.
@@
 Data points are colored and given point characters according to the species of the flea (the guided tour was unsupervised with this information).
+
+(ref:step0-cap) Basis reference axes (left) and projected data (right) of standardized flea data.
+Data points color and shape are mapped to beetle species.
+Basis identified by a holes-index guided tour.
+The variables \texttt{aede2} and \texttt{tars1} contribute mostly orthogonal to the other variables.
+We'll select \texttt{aede2} as our manipulation variable to see how the structure of the projection changes as we rotate \texttt{aede2} into and out of the projection.
 
@@
 
-\caption[(ref:step0-cap)]{(ref:step0-cap)}(\#fig:step0)
+\caption{\label{fig:step0} and some caption}(\#fig:step0)
 \end{figure}
 \end{Schunk}
-
-(ref:step0-cap) Basis reference axes (left) and projected data (right) of standardized flea data.
-Data points color and shape are mapped to beetle species.
-Basis identified by a holes-index guided tour.
-The variables \texttt{aede2} and \texttt{tars1} contribute mostly orthogonal to the other variables.
-We'll select \texttt{aede2} as our manipulation variable to see how the structure of the projection changes as we rotate \texttt{aede2} into and out of the projection.
 
@@
 Using the same data, another study, Cook, Laa, and Valencia (2018), applies grand tours to the same subspaces.
-Grand tours are able to better resolve the distribution shape of clusters, intra-cluter detail, better outlier detection, and exonerate a claim persened from TFEP (TensorFlow embedded projections).
-Table 1 of Cook et al.~summarizes the key findings of observations made with PDFSense \& TFEP and those from grand tours.
+Grand tours are able to better resolve the distribution shape of clusters, intra-cluter detail, better outlier detection, and challenges a claim persented from TFEP (TensorFlow embedded projections).
+Table 1 of Cook \emph{et al.} summarizes the key findings of observations made with PDFSense \& TFEP and those from grand tours.
 
-Without getting too domain-specific the data has three primary groupings; DIS, VBP, and jet Each group is a particular class of experiments and each with many experimental datasets which inturn have many observation.
-Inconsideration of data density and business of the data We conduct manual tours on a subsets of the DIS and jet clusters.
-This explores the sensitivity of the structue to each of the variables in turn, and we present the subjectively best and worst manaul tour identifying structure in the respective data sets.
+Without getting too domain-specific the data has three primary groupings; \texttt{DIS}, \texttt{VBP}, and \texttt{jet}.
+Each group is a particular class of experiments and each with many experimental datasets which, inturn, have many observation.
+In the consideration of data density of the data we conduct manual tours on a subsets of the \texttt{DIS} and \texttt{jet} clusters.
+This explores the sensitivity of the structue to each of the variables in turn and we present the subjectively best and worst variable to manipulate for identifying structure and outliers.
 
@@
 
-The jet cluster is of interest as it contains the largest data sets and is found to be important in Wang et al.
-(2018) The jet cluster resides in a smaller dimensionality than the full set of experiments with 4 principal components explaining 95\% of the variation in the jet cluster (Cook, Laa, and Valencia 2018).
-The data is subset down to ATLAS7old and ATLAS7new to focus in on two groups with a reasonable number of observations that occupy different parts of the subspace.
+The \texttt{jet} cluster is of interest as it contains the largest data sets and is found to be important in Wang et al.
+(2018).
+The jet cluster resides in a smaller dimensionality than the full set of experiments with 4 principal components explaining 95\% of the variation in the jet cluster (Cook, Laa, and Valencia 2018).
+The data within this 4D embedding is subset down to \texttt{ATLAS7old} and \texttt{ATLAS7new} to focus in on two groups with a reasonable number of observations that occupy different parts of the subspace.
 Below, we perform radial manual tours all four principal components within this scope.
@@
 
-(ref:JetClusterGood-cap) Jet cluster, a radial manual tour of PC3.
-Colored by experiment type: `ATLAS7new' in green and `ATLAS7old' in orange.
-When PC3 fully contributes to the projection ATLAS7new (green) occupies unique space and several outliers are identifiable.
-Zeroing the contribution from PC3 to the projection hides the outliers and indeed all observations with ATLAS7new are contained within ATLAS7old (orange).
+(ref:JetClusterGood-cap) \texttt{Jet} cluster, a radial manual tour of PC3.
+Colored by experiment type: \texttt{ATLAS7new} in green and \texttt{ATLAS7old} in orange.
+When PC3 fully contributes to the projection \texttt{ATLAS7new} (green) occupies unique space and several outliers are identifiable.
+Zeroing the contribution from PC3 to the projection hides the outliers and all observations with \texttt{ATLAS7new} are contained within the area spanned by \texttt{ATLAS7old} (orange).
 A dynamic version can be viewed at \url{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/}.
@@
 (ref:DISclusterGood-cap) DIS cluster, a radial manual tour of PC6.
-colored by experiment type: `DIS HERA1+2' in green, `dimuon SIDIS' in purple, and `charm SIDIS' in orange.
+colored by experiment type: DIS HERA1+2 in green, dimuon SIDIS in purple, and charm SIDIS in orange.
 When the contribution PC 6 is large we see that dimuon SIDIS (purple) data are nearly orthogonal to DIS HERA (green) data.
@@
 (ref:DISclusterBad-cap) DIS cluster, a radial manual tour of PC2.
-Colored by experiment type: `DIS HERA1+2' in green, `dimuon SIDIS' in purple, and `charm SIDIS' in orange.
+Colored by experiment type: DIS HERA1+2 in green, dimuon SIDIS in purple, and charm SIDIS in orange.
 The structure of previously described plane of DIS HERA (green) and nearly orthogonal dimuon SIDIS (purple) is present, however, the manipulating PC2 does not give a head-on view of either, a less useful manual tour than that of PC6.
@@
 
-\leavevmode\hypertarget{ref-cook_manual_1997}{}%
-Cook, Dianne, and Andreas Buja.
-1997.
-``Manual Controls for High-Dimensional Data Projections.'' \emph{Journal of Computational and Graphical Statistics} 6 (4): 464--80.
-\url{https://doi.org/10.2307/1390747}.
-
-\leavevmode\hypertarget{ref-cook_grand_1995}{}%
-Cook, Dianne, Andreas Buja, Javier Cabrera, and Catherine Hurley.
-1995.
-``Grand Tour and Projection Pursuit.'' \emph{Journal of Computational and Graphical Statistics} 4 (3): 155.
-\url{https://doi.org/10.2307/1390844}.
-
 \leavevmode\hypertarget{ref-cook_dynamical_2018}{}%
@@
 \url{https://doi.org/10.2307/1390897}.
-
-\leavevmode\hypertarget{ref-lubischew_use_1962}{}%
-Lubischew, Alexander A. 1962. ``On the Use of Discriminant Functions in Taxonomy.'' \emph{Biometrics}, 455--77.
 
@@
 
-\leavevmode\hypertarget{ref-wickham_tourr_2011}{}%
-Wickham, Hadley, Dianne Cook, Heike Hofmann, and Andreas Buja.
-2011.
-``\textbf{Tourr} : An \emph{R} Package for Exploring Multivariate Data with Projections.'' \emph{Journal of Statistical Software} 40 (2).
-\url{https://doi.org/10.18637/jss.v040.i02}.
-
 \leavevmode\hypertarget{ref-xie_bookdown:_2016}{}%
@@
 Monash University\\
-line 1\\ line 2\\
+School of Information Technology\\ Monash University, Melbourne, VIC 3800, Australia\\
 }
-\href{mailto:nicholas.spyrison@monash.edu}{\nolinkurl{nicholas.spyrison@monash.edu}}
+\href{mailto:Nicholas.Spyrison@monash.edu}{\nolinkurl{Nicholas.Spyrison@monash.edu}}
 
@@
 Monash University\\
-line 1\\ line 2\\
+School of Econometrics and Business Statistics\\ Monash University, Melbourne, VIC 3800, Australia\\
 }
-\href{mailto:author2@work}{\nolinkurl{author2@work}}
+\href{mailto:DiCook@monash.edu}{\nolinkurl{DiCook@monash.edu}}

### 2019-04-05 paper/spinifex_paper.Rmd
wrestling references

@@
 ---
-title: spinifex; manual control of dynamic linear projections of high-dimensional data
+title: > 
+  spinifex: manual control of dynamic linear projections of high-dimensional data
 author:
@@
     address:
-    - line 1
-    - line 2
-    email:  nicholas.spyrison@monash.edu
+    - School of Information Technology
+    - Monash University, Melbourne, VIC 3800, Australia
+    email:  Nicholas.Spyrison@monash.edu
   - name: Dianne Cook
@@
     address:
-    - line 1
-    - line 2
-    email:  author2@work
+    - School of Econometrics and Business Statistics
+    - Monash University, Melbourne, VIC 3800, Australia
+    email:  DiCook@monash.edu
 abstract: >
@@
 This research applies manual tours to this data showing that manual tours resolves more structrual information that is orthogonal to the original viewing plane.
+keywords: manual tour, guided tour, grand tour, projection pursuit, high dimensional data, multivariate data, data visualization, statistical graphics, data science
 output:
   bookdown::pdf_book:
-    base_format: rticles::rjournal_article
+    #base_format: rticles::rjournal_article #see sand box for hard code solution.
+    fig_caption: yes
     includes:
@@
 ---
+
 \bibliography{RJrefs}
-```{r setup_03, include=FALSE}
+```{r setup_paper, include=FALSE}
 knitr::opts_chunk$set(
-  fig.align = "center",
-  echo = TRUE, 
-  message = FALSE,
-  warning = FALSE,
-  error = FALSE,
-  collapse = TRUE,
-  comment = "",
-  cache = F, 
-  cache.lazy = FALSE
+  fig.align = "center", echo = TRUE, collapse = TRUE,
+  message = FALSE, warning = FALSE, error = FALSE,
+  cache = F, cache.lazy = FALSE
 )
-
 library(spinifex)
 library(ggplot2)
-
+```
+```{r localFunction, include=FALSE}
 ## FUNCTION FOR STATIC OUTPPUT, consider going to scalable facets.
@@
 
-<!-- Tours -->
-A tour is a multivariate data analysis technique in which a sequence of linear (orthogonal) projections are viewed as an animation while the orientation of the projection basis is rotated across time.
+<!-- <!-- Tours --> -->
+<!-- A tour is a multivariate data analysis technique in which a sequence of linear (orthogonal) projections are viewed as an animation while the orientation of the projection basis is rotated across time.
 Each frame of the sequence corresponds to a small change in the projection for a smooth transition that perseveres continuity.
+-->
 
-<!-- Manual tour -->
-While there are numerous methods that generate tour paths, this research focuses on the manual tour.
+<!-- <!-- Manual tour --> -->
+<!-- While there are numerous methods that generate tour paths, this research focuses on the manual tour.
 The manual tour was described in @cook_manual_1997 and allows a user to control the projection coefficients of a select variable has in a 2D projection.
@@
 Once the given projection has been provided, it is time to define the path of rotation.
+-->
 
-<!-- Manual tour paths -->
-Ideally, the path would be intuitively user-generated from physical movement, be it through mouse or motion capture.
+<!-- <!-- Manual tour paths --> -->
+<!-- Ideally, the path would be intuitively user-generated from physical movement, be it through mouse or motion capture.
 Unfortunately, this type of dynamic control has proven difficult to capture for in R.
@@
 In the section below we stick with a radial rotation where, $\theta$, the angle of in-projection-plane rotation is held constant.
+-->
 
-<!-- Display -->
-Spinifex utilizes two new animation packages, *plotly* [@sievert_plotly_2018] and *gganimate* [@pedersen_gganimate:_2019], to display tours, manual or other saved tours.
+<!-- <!-- Display --> -->
+<!-- Spinifex utilizes two new animation packages, *plotly* [@sievert_plotly_2018] and *gganimate* [@pedersen_gganimate:_2019], to display tours, manual or other saved tours.
 From a given projection, the user can choose which variable to control, and the animation sequence is generated to remove the variable from the projection, and then extend its contribution to be the sole variable in one direction.
 This allows the viewer to assess the change in structure induced in the projection by the variable's contribution.
+-->
 
-<!-- Paper outline -->
-The paper is organized as follows.
+<!-- <!-- Paper outline --> -->
+<!-- The paper is organized as follows.
 Section \ref{sec:algorithm} explains the algorithm using a toy dataset.
@@
 The last section, \ref{sec:discussion} summarizes the work and discusses future research.
+-->
 
-<!-- Algorithm outline -->
-## Algorithm {#sec:algorithm}
+<!-- <!-- Algorithm outline --> -->
+<!-- ## Algorithm {#sec:algorithm} -->
 
-The section below describes the algorithm for performing a 2D radial manual tour:
+<!-- The section below describes the algorithm for performing a 2D radial manual tour: -->
 
-1. Provided with a 2D projection, choose a variable to explore. This is called the "manip" variable.
-2. Create a 3D manipulation space, where the manip variable has the full contribution.
-3. Generate a rotation sequence which increases the norm of the coefficient to 1 and zeros it.
+<!-- 1. Provided with a 2D projection, choose a variable to explore. This is called the "manip" variable. -->
+<!-- 2. Create a 3D manipulation space, where the manip variable has the full contribution. -->
+<!-- 3. Generate a rotation sequence which increases the norm of the coefficient to 1 and zeros it. -->
 
-The steps are described in more detail below. The R functions used below mentioned briefly, but more complete code example can be found in section \ref{sec:usage}
+<!-- The steps are described in more detail below. The R functions used below mentioned briefly, but more complete code example can be found in section \ref{sec:usage} -->
 
-### Notation
+<!-- ### Notation -->
 
-This section describes the notation used in the algorithm for a 2D radial manual tour.
+<!-- This section describes the notation used in the algorithm for a 2D radial manual tour. -->
 
-<!-- remember mixing latex and rmarkdown bullets isn't trivial. -->
-\begin{itemize}
-  \item $\textbf{X}$, the data, an $n \times p$ numeric matrix to be embedded in two dimensions.
-  \item $\textbf{B} = (B_1,~ B_2)$, any orthonormal projection basis set, $p \times 2$ matrix, describing the projection from $p$ to two dimensions.
-  \item $\textbf{e}$, a zero column vector of length $p$ with the $k-$th element set to one, where $k$ is the number of the variable to manipulate.
-  \item $\theta$, the angle of in-projection-plane rotation, for example, on the reference axes. 
-  \item $\phi$, the angle of out-of-projection-plane rotation, coming into the manipulation space.
-\end{itemize}
+<!-- <!-- remember mixing latex and rmarkdown bullets isn't trivial. --> -->
+<!-- \begin{itemize} -->
+<!--   \item $\textbf{X}$, the data, an $n \times p$ numeric matrix to be embedded in two dimensions. -->
+<!--   \item $\textbf{B} = (B_1,~ B_2)$, any of orthonormal projection basis set, $p \times 2$ matrix, describing the projection from $p$ to two dimensions -->
+<!--   \item $\textbf{e}$, a zero column vector of length $p$ with the $k-$th element set to one, where $k$ is the number of the variable to manipulate. -->
+<!--   \item $\theta$, the angle of in-projection-plane rotation, for example, on the reference axes.  -->
+<!--   \item $\phi$, the angle of out-of-projection-plane rotation, coming into the manipulation space. -->
+<!-- \end{itemize} -->
 
-The algorithm primarily operates on the projection basis and utilizes the data only when making a display.
+<!-- The algorithm primarily operates on the projection basis and utilizes the data only when making a display.
 The projection space can be viewed at any point in the process by pre-multiplying the data and plotting the first 2 variables.
+-->
 
-### Toy data set
+<!-- ### Toy data set -->
 
-<!-- describe data. -->
-The flea data, originally from @lubischew_use_1962, available in the R package *tourr* [@wickham_tourr_2011] is used to illustrate the algorithm.
+<!-- <!-- describe data. --> -->
+<!-- The flea data, originally from @lubischew_use_1962, available in the R package *tourr* [@wickham_tourr_2011] is used to illustrate the algorithm.
 The data contains 74 observations across 6 variables, physical measurements of the flea beetles.
 Each observation belonging to one of three species.
+-->
 
@@
 The holes index is maximized by when the projected observations are furthest from the center.
-Figure \@ref(fig:step0) shows a locally optimized projection for this data.
+Figure \@ref{fig:step0} shows a locally optimized projection for this data.
 The left plot displays the reference axes of the projection basis, a visual indication of the magnitude and direction each variable contributed to the projections.
@@
 
-```{r step0, echo=F, results='hide', out.width='95%', fig.cap = "(ref:step0-cap)"}
+(ref:step0-cap) Basis reference axes (left) and projected data (right) of standardized flea data.
+Data points color and shape are mapped to beetle species.
+Basis identified by a holes-index guided tour.
+The variables `aede2` and `tars1` contribute mostly orthogonal to the other variables.
+We'll select `aede2` as our manipulation variable to see how the structure of the projection changes as we rotate `aede2` into and out of the projection.
+
+```{r step0, echo=F, results='hide', out.width='95%', fig.cap = "\\label{fig:step0} and some caption"}
+
 ## Flea holes tour
@@
 ```
-
-(ref:step0-cap) Basis reference axes (left) and projected data (right) of standardized flea data.
-Data points color and shape are mapped to beetle species.
-Basis identified by a holes-index guided tour.
-The variables `aede2` and `tars1` contribute mostly orthogonal to the other variables.
-We'll select `aede2` as our manipulation variable to see how the structure of the projection changes as we rotate `aede2` into and out of the projection.
 
@@
 Using the same data, another study, @cook_dynamical_2018, applies grand tours to the same subspaces.
-Grand tours are able to better resolve the distribution shape of clusters, intra-cluter detail, better outlier detection, and exonerate a claim persened from TFEP (TensorFlow embedded projections).
-Table 1 of Cook et al.
-summarizes the key findings of observations made with PDFSense & TFEP and those from grand tours.
+Grand tours are able to better resolve the distribution shape of clusters, intra-cluter detail, better outlier detection, and challenges a claim persented from TFEP (TensorFlow embedded projections).
+Table 1 of Cook *et al.* summarizes the key findings of observations made with PDFSense & TFEP and those from grand tours.
 
 <!-- is this the best place for this level of detail or is it suited better later on? -->
-Without getting too domain-specific the data has three primary groupings; DIS, VBP, and jet Each group is a particular class of experiments and each with many experimental datasets which inturn have many observation.
-Inconsideration of data density and business of the data We conduct manual tours on a subsets of the DIS and jet clusters.
-This explores the sensitivity of the structue to each of the variables in turn, and we present the subjectively best and worst manaul tour identifying structure in the respective data sets.
+Without getting too domain-specific the data has three primary groupings; `DIS`, `VBP`, and `jet`.
+Each group is a particular class of experiments and each with many experimental datasets which, inturn, have many observation.
+In the consideration of data density of the data we conduct manual tours on a subsets of the `DIS` and `jet` clusters.
+This explores the sensitivity of the structue to each of the variables in turn and we present the subjectively best and worst variable to manipulate for identifying structure and outliers.
 
@@
 <!-- motivate jet cluster, explain dimensionality  -->
-The jet cluster is of interest as it contains the largest data sets and is found to be important in @wang_mapping_2018 The jet cluster resides in a smaller dimensionality than the full set of experiments with 4 principal components explaining 95% of the variation in the jet cluster [@cook_dynamical_2018].
-The data is subset down to ATLAS7old and ATLAS7new to focus in on two groups with a reasonable number of observations that occupy different parts of the subspace.
+The `jet` cluster is of interest as it contains the largest data sets and is found to be important in @wang_mapping_2018.
+The jet cluster resides in a smaller dimensionality than the full set of experiments with 4 principal components explaining 95% of the variation in the jet cluster [@cook_dynamical_2018].
+The data within this 4D embedding is subset down to `ATLAS7old` and `ATLAS7new` to focus in on two groups with a reasonable number of observations that occupy different parts of the subspace.
 Below, we perform radial manual tours all four principal components within this scope.
@@
 
-(ref:JetClusterGood-cap) Jet cluster, a radial manual tour of PC3.
-Colored by experiment type: 'ATLAS7new' in green and 'ATLAS7old' in orange.
-When PC3 fully contributes to the projection ATLAS7new (green) occupies unique space and several outliers are identifiable.
-Zeroing the contribution from PC3 to the projection hides the outliers and indeed all observations with ATLAS7new are contained within ATLAS7old (orange).
+(ref:JetClusterGood-cap) `Jet` cluster, a radial manual tour of PC3.
+Colored by experiment type: `ATLAS7new` in green and `ATLAS7old` in orange.
+When PC3 fully contributes to the projection `ATLAS7new` (green) occupies unique space and several outliers are identifiable.
+Zeroing the contribution from PC3 to the projection hides the outliers and all observations with `ATLAS7new` are contained within the area spanned by `ATLAS7old` (orange).
 A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/.
@@
 (ref:DISclusterGood-cap) DIS cluster, a radial manual tour of PC6.
-colored by experiment type: 'DIS HERA1+2' in green, 'dimuon SIDIS' in purple, and 'charm SIDIS' in orange.
+colored by experiment type: DIS HERA1+2 in green, dimuon SIDIS in purple, and charm SIDIS in orange.
 When the contribution PC 6 is large we see that dimuon SIDIS (purple) data are nearly orthogonal to DIS HERA (green) data.
@@
 (ref:DISclusterBad-cap) DIS cluster, a radial manual tour of PC2.
-Colored by experiment type: 'DIS HERA1+2' in green, 'dimuon SIDIS' in purple, and 'charm SIDIS' in orange.
+Colored by experiment type: DIS HERA1+2 in green, dimuon SIDIS in purple, and charm SIDIS in orange.
 The structure of previously described plane of DIS HERA (green) and nearly orthogonal dimuon SIDIS (purple) is present, however, the manipulating PC2 does not give a head-on view of either, a less useful manual tour than that of PC6.

### 2019-04-05 paper/spinifex_paper.Rmd
changes before envigilating exam.

@@
 
-<!-- <!-- Tours --> -->
-<!-- A tour is a multivariate data analysis technique in which a sequence of linear (orthogonal) projections are viewed as an animation while the orientation of the projection basis is rotated across time.
+<!-- Tours -->
+A tour is a multivariate data analysis technique in which a sequence of linear (orthogonal) projections are viewed as an animation while the orientation of the projection basis is rotated across time.
 Each frame of the sequence corresponds to a small change in the projection for a smooth transition that perseveres continuity.
--->
 
-<!-- <!-- Manual tour --> -->
-<!-- While there are numerous methods that generate tour paths, this research focuses on the manual tour.
+<!-- Manual tour -->
+While there are numerous methods that generate tour paths, this research focuses on the manual tour.
 The manual tour was described in @cook_manual_1997 and allows a user to control the projection coefficients of a select variable has in a 2D projection.
@@
 Once the given projection has been provided, it is time to define the path of rotation.
--->
 
-<!-- <!-- Manual tour paths --> -->
-<!-- Ideally, the path would be intuitively user-generated from physical movement, be it through mouse or motion capture.
+<!-- Manual tour paths -->
+Ideally, the path would be intuitively user-generated from physical movement, be it through mouse or motion capture.
 Unfortunately, this type of dynamic control has proven difficult to capture for in R.
@@
 In the section below we stick with a radial rotation where, $\theta$, the angle of in-projection-plane rotation is held constant.
--->
 
-<!-- <!-- Display --> -->
-<!-- Spinifex utilizes two new animation packages, *plotly* [@sievert_plotly_2018] and *gganimate* [@pedersen_gganimate:_2019], to display tours, manual or other saved tours.
+<!-- Display -->
+Spinifex utilizes two new animation packages, *plotly* [@sievert_plotly_2018] and *gganimate* [@pedersen_gganimate:_2019], to display tours, manual or other saved tours.
 From a given projection, the user can choose which variable to control, and the animation sequence is generated to remove the variable from the projection, and then extend its contribution to be the sole variable in one direction.
 This allows the viewer to assess the change in structure induced in the projection by the variable's contribution.
--->
 
-<!-- <!-- Paper outline --> -->
-<!-- The paper is organized as follows.
+<!-- Paper outline -->
+The paper is organized as follows.
 Section \ref{sec:algorithm} explains the algorithm using a toy dataset.
@@
 The last section, \ref{sec:discussion} summarizes the work and discusses future research.
--->
 
-<!-- <!-- Algorithm outline --> -->
-<!-- ## Algorithm {#sec:algorithm} -->
+<!-- Algorithm outline -->
+## Algorithm {#sec:algorithm}
 
-<!-- The section below describes the algorithm for performing a 2D radial manual tour: -->
+The section below describes the algorithm for performing a 2D radial manual tour:
 
-<!-- 1. Provided with a 2D projection, choose a variable to explore. This is called the "manip" variable. -->
-<!-- 2. Create a 3D manipulation space, where the manip variable has the full contribution. -->
-<!-- 3. Generate a rotation sequence which increases the norm of the coefficient to 1 and zeros it. -->
+1. Provided with a 2D projection, choose a variable to explore. This is called the "manip" variable.
+2. Create a 3D manipulation space, where the manip variable has the full contribution.
+3. Generate a rotation sequence which increases the norm of the coefficient to 1 and zeros it.
 
-<!-- The steps are described in more detail below. The R functions used below mentioned briefly, but more complete code example can be found in section \ref{sec:usage} -->
+The steps are described in more detail below. The R functions used below mentioned briefly, but more complete code example can be found in section \ref{sec:usage}
 
-<!-- ### Notation -->
+### Notation
 
-<!-- This section describes the notation used in the algorithm for a 2D radial manual tour. -->
+This section describes the notation used in the algorithm for a 2D radial manual tour.
 
-<!-- <!-- remember mixing latex and rmarkdown bullets isn't trivial. --> -->
-<!-- \begin{itemize} -->
-<!--   \item $\textbf{X}$, the data, an $n \times p$ numeric matrix to be embedded in two dimensions. -->
-<!--   \item $\textbf{B} = (B_1,~ B_2)$, any of orthonormal projection basis set, $p \times 2$ matrix, describing the projection from $p$ to two dimensions -->
-<!--   \item $\textbf{e}$, a zero column vector of length $p$ with the $k-$th element set to one, where $k$ is the number of the variable to manipulate. -->
-<!--   \item $\theta$, the angle of in-projection-plane rotation, for example, on the reference axes.  -->
-<!--   \item $\phi$, the angle of out-of-projection-plane rotation, coming into the manipulation space. -->
-<!-- \end{itemize} -->
+<!-- remember mixing latex and rmarkdown bullets isn't trivial. -->
+\begin{itemize}
+  \item $\textbf{X}$, the data, an $n \times p$ numeric matrix to be embedded in two dimensions.
+  \item $\textbf{B} = (B_1,~ B_2)$, any of orthonormal projection basis set, $p \times 2$ matrix, describing the projection from $p$ to two dimensions
+  \item $\textbf{e}$, a zero column vector of length $p$ with the $k-$th element set to one, where $k$ is the number of the variable to manipulate.
+  \item $\theta$, the angle of in-projection-plane rotation, for example, on the reference axes.
+  \item $\phi$, the angle of out-of-projection-plane rotation, coming into the manipulation space.
+\end{itemize}
 
-<!-- The algorithm primarily operates on the projection basis and utilizes the data only when making a display.
+The algorithm primarily operates on the projection basis and utilizes the data only when making a display.
 The projection space can be viewed at any point in the process by pre-multiplying the data and plotting the first 2 variables.
--->
 
-<!-- ### Toy data set -->
+### Toy data set
 
-<!-- <!-- describe data. --> -->
-<!-- The flea data, originally from @lubischew_use_1962, available in the R package *tourr* [@wickham_tourr_2011] is used to illustrate the algorithm.
+<!-- describe data. -->
+The flea data, originally from @lubischew_use_1962, available in the R package *tourr* [@wickham_tourr_2011] is used to illustrate the algorithm.
 The data contains 74 observations across 6 variables, physical measurements of the flea beetles.
 Each observation belonging to one of three species.
--->
 
@@
 The variables `aede2` and `tars1` contribute mostly orthogonal to the other variables.
-We'll select `aede2` as our manipulation variable to see how the structure of the projection changes as we rotate `aede2` into and out of the projection.
+Select `aede2` as the manipulation variable to see how the structure of the projection changes as `aede2` is  rotated into and out of the projection.
 
-```{r step0, echo=F, results='hide', out.width='95%', fig.cap = "\\label{fig:step0} and some caption"}
+```{r step0, echo=F, results='hide', out.width='95%', fig.cap = "(ref:step0-cap)"} 
+#"\\label{fig:step0} and some caption"}
 
@@
 \begin{align*}
-  \textbf{e} &\leftarrow Orthonormalize_{GS}(\textbf{e}) w.r.t. Basis \\
+  \textbf{e} &\leftarrow Orthonormalize_{GS}(\textbf{e}) w.r.t. the basis \\
   &= \textbf{e} - \langle \textbf{e},\textbf{B}_1 \rangle \textbf{B}_1 - \langle \textbf{e}, \textbf{B}_2 \rangle \textbf{B}_2 \\

### 2019-04-05 doi_helper.r

@@ -1,0 +1,137 @@
+# The R Journal, 2017 Roger Bivand
+# file share/doi_helpers.R
+#
+library(bibtex)
+library(stringr)
+library(rcrossref)
+
+# Helper function to check whether a declared doi exists and points
+# to the correct object; the existing doi format must follow:
+# https://www.crossref.org/display-guidelines/
+# for example: url = {https://doi.org/10.1080/10618600.1996.10474713}
+#
+# input *.bib file sent by copy-editor
+# output a nested list of length = # articles with doi as urls, and nested
+# lists of length 3 giving the article bibtex key, a simple text rendering of
+# the original article metadata from the *.bib file, and the simple text
+# rendering of retrieved article metadata corresponding to the doi value.
+#
+# Visual inspection of the output $orig and $found values is required!!
+#
+# If the $found metadata indicate that the doi points to a different article
+# or if the $found component is the empty string, the reference at the bibtex
+# key needs urgent attention
+
+check_doi <- function(bibfile) {
+  biblist <- bibtex::read.bib(bibfile)
+  stopifnot(inherits(biblist, "bibentry"))
+  stopifnot(length(biblist) >= 1L)
+  res <- list()
+  for(i in 1:length(biblist)) {
+    if(tolower(biblist[i]$bibtype) == "article" && !is.null(biblist[i]$url)) {
+      url <- biblist[i]$url
+      str_search_doi <- "((https://doi.org/)([-a-zA-Z0-9/_\\+\\.]*))"
+      doi_str <- c(na.omit(str_trim(str_extract(url, str_search_doi))))
+      doi <- str_sub(doi_str, 17L, str_length(doi_str))
+      cat(biblist[i]$key, doi, "\n")
+      j <- 0
+      ok <- FALSE
+      while(j < 5 & !ok) {
+        j <- j + 1
+        ret <- try(rcrossref::cr_cn(dois = doi, format = "text",
+                                    style = "apa"), silent = TRUE)
+        if(inherits(ret, "try-error")) {
+          Sys.sleep(5)
+        } else {
+          ok <- TRUE
+        }
+      }
+      if(!ok) ret <- ""
+      resi <- list(list(key=biblist[i]$key, orig=format(biblist[i]),
+                        found=ret))
+      names(resi) <- biblist[i]$key
+      res <- c(res, resi)
+    }
+  }
+  res
+}
+
+# Helper function to try to guess article doi strings from article metadata,
+# based on jss package code on R-Forge:
[…]

### 2019-04-05 paper/spinifex_paper.Rmd
paper edits through step 3

@@
 abstract: >
-  The class of dynamic linear projections that are collectively known as 'tours' provide a unique dynamic visualization of numeric multivariate data.
+  The set of dynamic linear projections of multivariate data collectively known as **tours** provide a unique dynamic visualization as the dimensionality of the data increase.
 Tours are particularly useful for understanding the structure held within multivariate data, and in association with techniques for dimension reduction, supervised, and unsupervised classification.
-The *R* package *tourr* offers a variety of path generators and geometric displays for conducting tours.
-This paper discusses an extension package, *spinifex*, that adds support for the path generation of manual tours and extends the display of tours to use with the contemporary animation packages, *plotly* and *gganimate*.
+The R package **tourr** offers a variety of path generators and geometric displays for conducting tours.
+This paper discusses an extension package, **spinifex**, that adds support for the path generation of manual tours and extends the display of tours to use with the contemporary animation packages, **plotly** and **gganimate**.
 Manual tours are used to explore the sensitivity of structure as the contributions of a manipulation variable are changed.
@@
     fig_caption: yes
+    toc: false
     includes:
@@
 Each frame of the sequence corresponds to a small change in the projection for a smooth transition that perseveres continuity.
+This is a crutial tool extending the dimensionality of vizualization which is large boon as datasets and models exist in increasing cardinality.
 
@@
 Unfortunately, this type of dynamic control has proven difficult to capture for in R.
-Because of this, manual tours were not implemented in *tourr*.
+Because of this, manual tours were not implemented in **tourr** [@wickham_tourr_2011].
 This research allows for the consumption, but not the generation, of such dynamic input.
@@
 <!-- Display -->
-Spinifex utilizes two new animation packages, *plotly* [@sievert_plotly_2018] and *gganimate* [@pedersen_gganimate:_2019], to display tours, manual or other saved tours.
+Spinifex utilizes two new animation packages, **plotly** [@sievert_plotly_2018] and **gganimate** [@pedersen_gganimate:_2019], to display tours, manual or other saved tours.
 From a given projection, the user can choose which variable to control, and the animation sequence is generated to remove the variable from the projection, and then extend its contribution to be the sole variable in one direction.
@@
   \item $\textbf{X}$, the data, an $n \times p$ numeric matrix to be embedded in two dimensions.
-  \item $\textbf{B} = (B_1,~ B_2)$, any of orthonormal projection basis set, $p \times 2$ matrix, describing the projection from $p$ to two dimensions
+  \item $\textbf{B} = (B_1,~ B_2)$, any orthonormal projection basis set, $p \times 2$ matrix, describing the projection from $p$ to two dimensions
   \item $\textbf{e}$, a zero column vector of length $p$ with the $k-$th element set to one, where $k$ is the number of the variable to manipulate.
@@
 
-The algorithm primarily operates on the projection basis and utilizes the data only when making a display.
-The projection space can be viewed at any point in the process by pre-multiplying the data and plotting the first 2 variables.
+The algorithm primarily operates on projection bases and utilizes the data only when making a display.
+The projection space can be viewed at any point in the process by pre-multiplying the data and plotting the first two variables.
 
@@
 <!-- describe data. -->
-The flea data, originally from @lubischew_use_1962, available in the R package *tourr* [@wickham_tourr_2011] is used to illustrate the algorithm.
+The flea data, originally from @lubischew_use_1962, made available in **tourr**  is used to illustrate the algorithm.
 The data contains 74 observations across 6 variables, physical measurements of the flea beetles.
@@
 The data is defined.
-A basis set (ideally that views an interesting feature) should be provided to explore the sensitivity of the variables to the structure.
-To identify a projection containing an interesting feature, apply
-a guided tour[@cook_interactive_2007] on the flea data.
-In a guided tour the projection sequence is selected by optimizing an index via hill-climbing.
-In this case, the holes index is selected.
-The holes index is maximized by when the projected observations are furthest from the center.
-Figure \@ref{fig:step0} shows a locally optimized projection for this data.
-The left plot displays the reference axes of the projection basis, a visual indication of the magnitude and direction each variable contributed to the projections.
-The right plot shows the projection of the data through the basis set described by the reference axes (left).
-Data points are colored and given point characters according to the species of the flea (the guided tour was unsupervised with this information).
+A basis set (ideally orienting to an interesting feature) must be provided as an initial orientation.
+One way to identify a projection containing interesting features is to apply a guided tour [@cook_interactive_2007].
+In a guided tour the projection sequence is selected by optimizing an index fucntion via hill-climbing on the projection space.
+In this case, the holes index is selected and applied to standardized flea data.
+The holes index is maximized when the projected observations are furthest from the center.
+Figure \@ref{fig:step0} shows a locally optimized projection of the data.
+The left side displays the reference axes of the projection basis, a visual indication of the magnitude and direction each variable contributed to the projection.
+The right side shows the data as projectioned through the basis set described by the reference axes (left).
+Data points are colored and given shape according to the species (while the guided tour was unsupervised with this information).
 
@@
 The variables `aede2` and `tars1` contribute mostly orthogonal to the other variables.
-Select `aede2` as the manipulation variable to see how the structure of the projection changes as `aede2` is  rotated into and out of the projection.
 
@@
 
-Call `view_basis()` on a basis to produce a *ggplot2* graphic similar to \@ref(fig:step0).
+Call `view_basis()` on a basis to produce a **ggplot2** graphic similar to \@ref(fig:step0).
 Projection space is always available for display via the matrix multiplication $\textbf{X}_{[n,~p]} ~*~ \textbf{B}_{[p,~d]} ~=~ \textbf{P}_{[n,~d]}$.
@@
 <!-- select a manip var-->
-In figure \@ref(fig:step0), above, the contributions of the variables `tars1` and `aede2` are mostly orthogonal to the contributions of the other four variables.
-These two variables explain the variation of the data between the purple and green species.
-We select `aede2` as the manip var, the variable to be manipulated as it typically has a larger contribution after the optimizing the holes index.
-The question that will be explored in the explanation of the algorithm is how important the variable `aede2` is to the separation of the clusters.
+In figure \@ref(fig:step0) the contributions of the variables `tars1` and `aede2` are mostly orthogonal to the contributions of the other four variables.
+These two variables explain the variation of the data distinguishing the purple group from the rest of the sample.
+Select `aede2` as the manip var, the variable to be manipulated, as it has a larger contribution to the projection.
+The question that will be explored is how important the variable `aede2` is to the separation of the clusters.
 
@@
 <!-- Zero Vect, manip sp -->
-Initialize a zero vector $e$ of $p$ elements.
-Because `aede2` is the fifth variable in the data, set the $k=5$-th element to one giving the manip var a full contribution in this dimension.
+Initialize a zero vector **e** of length $p$.
+Set the fifth element to on, as  `aede2` is the fifth variable in the data, giving the manip var a full contribution in this dimension.
 Use the Gram-Schmidt process to orthonormalize the zero vector onto the basis yielding the 3D manipulation space, **M**.
@@
 \begin{align*}
-  \textbf{e} &\leftarrow Orthonormalize_{GS}(\textbf{e}) w.r.t. the basis \\
+  \textbf{e} &\leftarrow Orthonormalize_{GS}(\textbf{e}) ~~~~~ w.r.t. the basis \\
   &= \textbf{e} - \langle \textbf{e},\textbf{B}_1 \rangle \textbf{B}_1 - \langle \textbf{e}, \textbf{B}_2 \rangle \textbf{B}_2 \\
@@
 <!-- What the manip space facilitates -->
-Adding this extra dimension to our basis plane allows for the coefficients of the specified variable to be changed.
-For example, the ability to lift a piece of paper, rather than being constrained to the motion on a table top.
-Orthonormalizing rescales the new depth vector while the projection down to 2D is the original basis, that is the first $d$ vectors remain constant.
-Imagine the reference axes (and projection plane) laying flat on a table, while a new dimension exists with axes projecting back onto the reference axes.
-An illustration of such can be seen below in figure \@ref(fig:step2).
+Adding this new dimension to our projection space allows for the coefficients of the manip var to be changed via rotation about the origin.
+For example, the ability to lift a piece of paper, rather than being constrained to a 2D table top.
+Orthonormalizing rescales the new vector while the projection down to 2D remains the original basis.
+Imagine the reference axes laying flat on a table, the a new dimension extends above the table, with axes projecting back onto the reference axes.
+An illustration of this is depicted in figure \@ref(fig:step2).
 The manip var is highlighted, while the depths of the other variables are not depicted.
@@
 Basis selected by a holes-index guided tour.
-The Projection plane is shown in blue.
-The manipulation axis, in red, allows the coefficients of the manip var to be changed.
+The projection plane is shown in blue.
+The manipulation axis, in red, allows the coefficients of the manip var to be changed by rotation around the origin.
 
@@
 
-The representation in \@ref(fig:step2) can be duplicated by calling the function `view_manip_space()`.
+The representation in figure \@ref(fig:step2) can be produced by calling the function `view_manip_space()`.
 
@@
 <!-- illustration of axis manip -->
-Imagine holding the red axis it is fixed to the origin.
+Imagine holding the red axis, one end fixed to the origin.
 As it is manipulated the projection back onto the projection plane correspondingly moves.
@@
 <!-- describe manip var path -->
-For the demonstration of the radial tour, we define a sequence for $\phi$ that brings the initial contribution of the manip var to be maximized and then zeroed before returning to the initial position.
+For the demonstration of the radial tour, a sequence is define for $\phi$ that brings the `XY` contributions of the manip var to a maximum and to zero before returning to the initial position.
 
@@
 
-Animations can be produced using the function `play_manual_tour()`. This function defaults to an HTML5 widget produced from *plotly*.
+Animations can be produced using the function `play_manual_tour()`. This function defaults to an HTML5 widget produced from **plotly**.
 
@@
 
-The *tourr* package utilizes R's base graphics for the display of tours.
-*spinifex* allows tours to be used in rendered in *plotly* @sievert_plotly_2018 as an HTML5 object or *gganimate* @pedersen_gganimate:_2019 as .gif or .mp4 objects.
-Both of which build off *ggplot2* objects in internal functions.
+The **tourr** package utilizes R's base graphics for the display of tours.
+**spinifex** allows tours to be used in rendered in **plotly** @sievert_plotly_2018 as an HTML5 object or **gganimate** @pedersen_gganimate:_2019 as .gif or .mp4 objects.
+Both of which build off **ggplot2** objects in internal functions.
 Sharing of animations is not trivial especially in print and static formats such as .pdf.
 Even with the use of computers and dynamic file formats capturing the correct resolution, aspect, and display is challenging and many formats quickly bloat file sizes.
-Keep in mind hosting options and exporting functions from *plotly*, *gganimate* and *tourr*.
+Keep in mind hosting options and exporting functions from **plotly**, **gganimate** and **tourr**.
 
@@
 
-This article was created in *R* [@r_core_team_r:_2018], using *bookdown* [@xie_bookdown:_2016] and *rmarkdown* [@xie_r_2018], with code generating the examples inline.
+This article was created in R [@r_core_team_r:_2018], using **bookdown** [@xie_bookdown:_2016] and **rmarkdown** [@xie_r_2018], with code generating the examples inline.
 The source files for this article be found at [github.com/nspyrison/confirmation/](https://github.com/nspyrison/confirmation/).
-The source code for the *spinifex* package can be found at [github.com/nspyrison/spinifex/](https://github.com/nspyrison/spinifex/).
+The source code for the **spinifex** package can be found at [github.com/nspyrison/spinifex/](https://github.com/nspyrison/spinifex/).
 
@@
 Tours, the dynamic linear projection of multivariate data, is an important aspect of data visualization extending the display of data-space as data dimensionality increases.
-This research has modified the algorithm producing manual tours, applied this functionality in *R* and offers extends the graphics offerings that can be used to display tours.
+This research has modified the algorithm producing manual tours, applied this functionality in R and offers extends the graphics offerings that can be used to display tours.
 The paragraphs below explore how this work might be extended.
@@
 <!-- geometric and graphics display -->
-The *tourr* package provides many other geometric displays with the `tourr::display_*()` family.
-These geometric options could be integrated into the *ggplot2* framework for display on *plotly* and *gganimate*.
-Additionally, the *animation* package @xie_animation:_2018 could be implemented for another graphics framework.
-However, *animation* builds from base graphs while *spinifex* utilizes *ggplot2* graphics.
+The **tourr** package provides many other geometric displays with the `tourr::display_*()` family.
+These geometric options could be integrated into the **ggplot2** framework for display on **plotly** and **gganimate**.
+Additionally, the **animation** package @xie_animation:_2018 could be implemented for another graphics framework.
+However, **animation** builds from base graphs while **spinifex** utilizes **ggplot2** graphics.
 
@@
 Having a script only interaction with tours causes a significant barrier to entry.
-To a lesser extent, *plotly* offers some static interactions with the contained object, such as tooltips, brushing, and linking without communicating back to the R console.
-The development of a dynamic graphical user interface, perhaps with the use of a *shiny* [@chang_shiny:_2018] application, would mitigate the barrier to entry, allow for more rapid analysis, and offer an approachable demo tool.
+To a lesser extent, **plotly** offers some static interactions with the contained object, such as tooltips, brushing, and linking without communicating back to the R console.
+The development of a dynamic graphical user interface, perhaps with the use of a **shiny** [@chang_shiny:_2018] application, would mitigate the barrier to entry, allow for more rapid analysis, and offer an approachable demo tool.
 The user could easily switch between variables to control, adjust interpolation step angle, or flag/save specific frame basis sets.

### 2019-04-07 paper/spinifex_paper.Rmd
paper edits through data in projection space. grammerly, knit.

@@
 
-  A recent paper {@wang_mapping_2018} visualizes the sensitivey of the hadronic experiments to nucleon structure.
+  A recent paper {@wang_mapping_2018} visualizes the sensitivity of the hadronic experiments to nucleon structure.
 Sensitivity was characterized in non-linear 3D embeddings of the first 10 principal components.
-This research applies manual tours to this data showing that manual tours resolves more structrual information that is orthogonal to the original viewing plane.
+This research applies manual tours to this data showing that manual tours resolve more structural information that is orthogonal to the original viewing plane.
 keywords: manual tour, guided tour, grand tour, projection pursuit, high dimensional data, multivariate data, data visualization, statistical graphics, data science
@@
   bookdown::pdf_book:
-    #base_format: rticles::rjournal_article #see sand box for hard code solution.
+    #base_format: rticles::rjournal_article #see sandbox for hard code solution.
     fig_caption: yes
@@
 Each frame of the sequence corresponds to a small change in the projection for a smooth transition that perseveres continuity.
-This is a crutial tool extending the dimensionality of vizualization which is large boon as datasets and models exist in increasing cardinality.
+This is a crucial tool extending the dimensionality of visualization which is a large boon as datasets and models exist in increasing cardinality.
 
@@
 One way to identify a projection containing interesting features is to apply a guided tour [@cook_interactive_2007].
-In a guided tour the projection sequence is selected by optimizing an index fucntion via hill-climbing on the projection space.
+In a guided tour the projection sequence is selected by optimizing an index function via hill-climbing on the projection space.
 In this case, the holes index is selected and applied to standardized flea data.
@@
 The left side displays the reference axes of the projection basis, a visual indication of the magnitude and direction each variable contributed to the projection.
-The right side shows the data as projectioned through the basis set described by the reference axes (left).
+The right side shows the data as projected through the basis set described by the reference axes (left).
 Data points are colored and given shape according to the species (while the guided tour was unsupervised with this information).
@@
 <!-- Zero Vect, manip sp -->
-Initialize a zero vector **e** of length $p$.
+Initialize a zero vector **e** of length *p*.
 Set the fifth element to on, as  `aede2` is the fifth variable in the data, giving the manip var a full contribution in this dimension.
@@
 Adding this new dimension to our projection space allows for the coefficients of the manip var to be changed via rotation about the origin.
-For example, the ability to lift a piece of paper, rather than being constrained to a 2D table top.
+For example, the ability to lift a piece of paper, rather than being constrained to a 2D plane.
 Orthonormalizing rescales the new vector while the projection down to 2D remains the original basis.
-Imagine the reference axes laying flat on a table, the a new dimension extends above the table, with axes projecting back onto the reference axes.
-An illustration of this is depicted in figure \@ref(fig:step2).
-The manip var is highlighted, while the depths of the other variables are not depicted.
+Place the plane horizontally, with the new dimension extending vertically, with axes projecting back onto the reference axes.
+Figure \@ref(fig:step2) illustrates this 3D manipulation space with the manip var highlighted (height of the other variables are not depicted.)
 
 (ref:step2-cap) Manipulation space for controlling the contribution of `aede2` of standardized flea data.
-Basis selected by a holes-index guided tour.
-The projection plane is shown in blue.
-The manipulation axis, in red, allows the coefficients of the manip var to be changed by rotation around the origin.
+Basis selected by a holes-index optimized guided tour lies on the horizontal projection plane, shown in blue.
+The manip var axis, in red, extends into the vertical manipulation space, allowing the coefficients of the manip var to be changed by rotation around the origin.
 
@@
 <!-- illustration of axis manip -->
-Imagine holding the red axis, one end fixed to the origin.
-As it is manipulated the projection back onto the projection plane correspondingly moves.
+Imagine holding the manip var, the red axis, one end fixed to the origin.
+As it is controlled the manipulation space rotates about the origin, the projection onto the horizontal projection plane correspondingly moves.
 This is what happens in a manual tour.
-For a radial tour, fix $\theta$, the angle within the blue plane, and vary the sequence of $\phi$, the angle coming out of the projection plane.
-Conceptually, live manipulation on a 2D plane allows the user to dynamically control these angles, effectively changing the coefficients of the manip var, which then performs a constrained rotation on the remaining variables.
+Generating a sequence of values for the horizontal and vertical, angles produces a path for the rotation of the manipulation space.
+This defines the (orthonormally-constrained) rotation on the coefficients of the variables.
 
 <!-- describe manip var path -->
-For the demonstration of the radial tour, a sequence is define for $\phi$ that brings the `XY` contributions of the manip var to a maximum and to zero before returning to the initial position.
+For a radial tour fix the (horizontal) angle within the projection plane, $\theta$, and define a sequence for the (vertical) angle coming out of the projection plane, $\phi$, bringing the initial `XY` contributions of the manip var to a maximum and then to zero before returning to the initial position.
+Dynamic capture of user manipulation is typically performed directly on the projection plane (without depiction of the manipulation space.)
 
@@
   \textbf{RM}_{[p,~3,~i]}
-  &= \textbf{M}_{[p,~3]} ~*~ \textbf{R}_{[3,~3]} \\
+  &= \textbf{M}_{[p,~3]} ~*~ \textbf{R}_{[3,~3,~i]} \\
   &= \begin{bmatrix}
@@
     c_\phi
-  \end{bmatrix}_{[3,~3]}
+  \end{bmatrix}_{[3,~3,~i]}
 \end{align*}
@@
 <!--  phi transform and application note -->
-A note on application: compile the sequence of $\phi_i$ and create an array/long table for each rotated manipulation space.
-$\phi$ is the angle relative to the initial value of $\phi$, we find the transformation $\phi_i$ - $\phi_1$ useful to think about $\phi$ relative to the basis plane.
-Additionally, the value of $\phi$ may be offset by a factor of pi.
+A note on application: $\phi$ is the angle relative to the initial value of $\phi$, we find the transformation $\phi_i$ - $\phi_1$ useful to think about $\phi$ relative to the basis plane.
+Additionally, the value of $\phi$ may be out of phase by a factor of pi.
 If the manip variable doesn't move as expected these are the first places to check.
 
-```{r STEP3CODEEXAMPLE, eval=F, echo=T}
-for (phi in seq(seq_start, seq_end, phi_inc_sign)) {
-  slide <- slide + 1
-  tour[,, slide] <- rotate_manip_space(manip_space, theta, phi)[, 1:2]
-}
-```
+<!-- ```{r STEP3CODEEXAMPLE, eval=F, echo=T} -->
+<!-- for (phi in seq(seq_start, seq_end, phi_inc_sign)) { -->
+<!--   slide <- slide + 1 -->
+<!--   tour[,, slide] <- rotate_manip_space(manip_space, theta, phi)[, 1:2] -->
+<!-- } -->
+<!-- ``` -->
 
 <!-- description of slides, on the technical side we haven't gone back to data space yet. -->
-Figure \@ref(fig:step3) illustrates a sequence with 15 projected bases and highlight the manip variable on top while showing the corresponding projected data points on the bottom.
-Take note of how the changes in the manip var change the distance between the purple and green cluster of points, `aede2` is crucial in distinguishing between these groups.
+Figure \@ref(fig:step3) illustrates a sequence with 15 projected bases, showing the reference axes on top the corresponding projected data points on the below.
+As a result, we can see that changes in the manip var controlled the distance between the purple cluster and the remaining sample, `aede2` is crucial in distinguishing this species.
 Tours are typically viewed as an animation such a dynamic version of this tour can be viewed online at https://nspyrison.netlify.com/thesis/flea_manualtour_mvar5/.
 The page may take a moment to load.
-The format of this figure and linking to an HTML animation will be used again in the Application, section \ref{sec:application}.
 
 <!-- step3 caption -->
-(ref:step3-cap) Radial manual tour changing the contributions from `aede2` of standardized flea data.
+(ref:step3-cap) Radial manual tour manipulating `aede2` of standardized flea data.
 The contributions increase from its initial contribution to a full contribution to the projection before decreasing to zero and then returning to its initial value.
-The change in the projected data shows that `aede2` is important for distinguishing between the purple and green clusters.
+The change in the projected data shows that `aede2` is important for distinguishing between the purple species.
 An animated version can be viewed at https://nspyrison.netlify.com/thesis/flea_manualtour_mvar5/.
@@
 
-Animations can be produced using the function `play_manual_tour()`. This function defaults to an HTML5 widget produced from **plotly**.
+Animations can be produced using the function `play_manual_tour()`.
+This function defaults to an HTML5 widget produced from **plotly**.
+The `render_type` argument can be changed to `render_gganimate` for exporting to .gif or .mp4 files.
 
@@
 
-<!-- reminder of basis operation and now apply data -->
+<!-- the reminder of basis operation and now apply data -->
 In light of performance, the above operations are performed on the bases without the use of the larger datasets.
@@
 <!-- Introduction of data and original paper  -->
-In a recent paper, @wang_mapping_2018, the authors aggregate and visualize the sensitivity of hadronic experiments to ncleon structure.
+In a recent paper, @wang_mapping_2018, the authors aggregate and visualize the sensitivity of hadronic experiments to nucleon structure.
 The authors introduce a new tool, PDFSense, to aid in the visualization of Parton distribution functions (PDF).
@@
 Using the same data, another study, @cook_dynamical_2018, applies grand tours to the same subspaces.
-Grand tours are able to better resolve the distribution shape of clusters, intra-cluter detail, better outlier detection, and challenges a claim persented from TFEP (TensorFlow embedded projections).
+Grand tours are able to better resolve the distribution shape of clusters, intra-cluster detail, better outlier detection, and challenges a claim presented from TFEP (TensorFlow embedded projections).
 Table 1 of Cook *et al.* summarizes the key findings of observations made with PDFSense & TFEP and those from grand tours.
@@
 Without getting too domain-specific the data has three primary groupings; `DIS`, `VBP`, and `jet`.
-Each group is a particular class of experiments and each with many experimental datasets which, inturn, have many observation.
-In the consideration of data density of the data we conduct manual tours on a subsets of the `DIS` and `jet` clusters.
-This explores the sensitivity of the structue to each of the variables in turn and we present the subjectively best and worst variable to manipulate for identifying structure and outliers.
+Each group is a particular class of experiments and each with many experimental datasets which, in turn, have many observations.
+In the consideration of data density, we conduct manual tours on subsets of the `DIS` and `jet` clusters.
+This explores the sensitivity of the structure to each of the variables in turn and we present the subjectively best and worst variable to manipulate for identifying structure and outliers.
 
@@
 Below, we perform radial manual tours all four principal components within this scope.
-Visualizing PC3 and PC4 in figure \@ref(fig:JetClusterGood) (more sturcturally insightful) and figure \@ref(fig:JetClusterBad) (less sturcturally insightful) respectively, and list links to dynamic animation of all variables.
+Visualizing PC3 and PC4 in figure \@ref(fig:JetClusterGood) (more structurally insightful) and figure \@ref(fig:JetClusterBad) (less structurally insightful) respectively, and list links to dynamic animation of all variables.
 
@@
 ## TODO: edit application, Grammarly and word checks.
-<!-- disscusion of findings and which is better -->
+<!-- discussion of findings and which is better -->
 Manipulating PC3, where varying the angle of rotation brings interesting features into and out of the center mass of the data, is more interesting than the manipulation of PC4, where the features are mostly independent of the contribution of PC4.
@@
 
-<!-- summary and seague  -->
+<!-- summary and segue  -->
 Tours, the dynamic linear projection of multivariate data, is an important aspect of data visualization extending the display of data-space as data dimensionality increases.

### 2019-04-08 paper/spinifex_paper.Rmd
Paper/digital edits though jet cluster. Knit (no grammerly)

@@
 <!-- the reminder of basis operation and now apply data -->
-In light of performance, the above operations are performed on the bases without the use of the larger datasets.
+In an appeal to performance, the above operations are performed on the bases without the use of the larger datasets.
 After the bases are brought into the projection-space, however, it is helpful to observe them with data in the same space.
@@
 
-The **tourr** package utilizes R's base graphics for the display of tours.
-**spinifex** allows tours to be used in rendered in **plotly** @sievert_plotly_2018 as an HTML5 object or **gganimate** @pedersen_gganimate:_2019 as .gif or .mp4 objects.
+The **tourr** package utilizes **base** graphics for the display of tours.
+**spinifex** allows tours to be used in rendered in **plotly** as an HTML5 object or **gganimate** as .gif or .mp4 objects.
 Both of which build off **ggplot2** objects in internal functions.
-Sharing of animations is not trivial especially in print and static formats such as .pdf.
+Sharing of animations is not trivial especially in print and static file formats such as .pdf.
 Even with the use of computers and dynamic file formats capturing the correct resolution, aspect, and display is challenging and many formats quickly bloat file sizes.
@@
 
-Storing each data point for every frame of the animation is very inefficient.
-Just as operations are performed on the bases, so too should tour paths be stored as bases.
+Storing each data point for every frame of the animation is redundant.
+Just as operations are performed on the bases, so too should tour paths be stored as bases and a single instance of the data.
 Consider a radial manual tour, we can store the salient features in 3 bases, where $\phi$ is at its starting, minimum, and maximum values.
@@
 
-<!-- grand tours on the same data, summarize findings better -->
+<!-- grand tours on the same data -->
 Using the same data, another study, @cook_dynamical_2018, applies grand tours to the same subspaces.
-Grand tours are able to better resolve the distribution shape of clusters, intra-cluster detail, better outlier detection, and challenges a claim presented from TFEP (TensorFlow embedded projections).
-Table 1 of Cook *et al.* summarizes the key findings of observations made with PDFSense & TFEP and those from grand tours.
+Grand tours are dynamic subspace projections of high dimensional where frames and selected at random and linked with geodesically interpolation of the intermidiate frames.
+Grand tours are able to better resolve the distribution shape of clusters, intra-cluster detail, and better outlier detection than use of PDFSense & TFEP (TensorFlow embedded projections).
+Before applying manual tours let's discuss the structure of the data.
 
-<!-- is this the best place for this level of detail or is it suited better later on? -->
-Without getting too domain-specific the data has three primary groupings; `DIS`, `VBP`, and `jet`.
-Each group is a particular class of experiments and each with many experimental datasets which, in turn, have many observations.
+<!-- Data structure -->
+The data has a hierachrical structure with top-level groupings; `DIS`, `VBP`, and `jet`.
+Each group is a particular class of experiments, each with many experimental datasets which, in turn, have many observations.
 In the consideration of data density, we conduct manual tours on subsets of the `DIS` and `jet` clusters.
-This explores the sensitivity of the structure to each of the variables in turn and we present the subjectively best and worst variable to manipulate for identifying structure and outliers.
+This explores the sensitivity of the structure to each of the variables in turn and we present the subjectively best and worst variable to manipulate for identifying dimensionality of the clusters and describing the span of the clusters.
 
@@
 The data within this 4D embedding is subset down to `ATLAS7old` and `ATLAS7new` to focus in on two groups with a reasonable number of observations that occupy different parts of the subspace.
-Below, we perform radial manual tours all four principal components within this scope.
-Visualizing PC3 and PC4 in figure \@ref(fig:JetClusterGood) (more structurally insightful) and figure \@ref(fig:JetClusterBad) (less structurally insightful) respectively, and list links to dynamic animation of all variables.
+Radial manual tours controlling contributions from PC4 and PC3 are shown in figure \@ref(fig:JetClusterGood) and figure \@ref(fig:JetClusterBad) respectively.
+These components are selected to contrast the difference of information conveyed by touring different variables.
+Links to dynamic HTML5 animations controlling each of the four variables are also provided.
 

### 2019-04-08 paper/spinifex_paper.Rmd
Changed abstract, moved code betweeen steps. consistency. Gammerly and knit.

@@
 abstract: >
-  The set of dynamic linear projections of multivariate data collectively known as **tours** provide a unique dynamic visualization as the dimensionality of the data increase.
-Tours are particularly useful for understanding the structure held within multivariate data, and in association with techniques for dimension reduction, supervised, and unsupervised classification.
+  The set of dynamic linear projections of multivariate data collectively known as **tours** provides an important tool for extending the dimensionality of visuals.
 The R package **tourr** offers a variety of path generators and geometric displays for conducting tours.
-This paper discusses an extension package, **spinifex**, that adds support for the path generation of manual tours and extends the display of tours to use with the contemporary animation packages, **plotly** and **gganimate**.
+This paper discusses an extension package, **spinifex**, that adds support for the path generation of manual tours and extends the display of tours to use with the animation packages, **plotly** and **gganimate**.
 Manual tours are used to explore the sensitivity of structure as the contributions of a manipulation variable are changed.
-This particularly useful after identifying a feature of interest.
 
@@
 Sensitivity was characterized in non-linear 3D embeddings of the first 10 principal components.
-This research applies manual tours to this data showing that manual tours resolve more structural information that is orthogonal to the original viewing plane.
+This research applies manual tours to this data showing that manual tours resolve information about cluster structure and identify outliers, information that is orthogonal to the original static viewing plane.
 keywords: manual tour, guided tour, grand tour, projection pursuit, high dimensional data, multivariate data, data visualization, statistical graphics, data science
@@
 This is a crucial tool extending the dimensionality of visualization which is a large boon as datasets and models exist in increasing cardinality.
+The package **tourr** [@wickham_tourr_2011] builds a platform for generating tour paths and applying to a wide number of display geometrics in **base** graphics.
 
@@
 While there are numerous methods that generate tour paths, this research focuses on the manual tour.
-The manual tour was described in @cook_manual_1997 and allows a user to control the projection coefficients of a select variable has in a 2D projection.
+The manual tour was described in @cook_manual_1997 and allows a user to control the projection coefficients of a selected variable in a 2D projection.
 The manipulation of these coefficients allows the analyst to explore how sensitive the projections structure is to these changes.
-This makes manual tours particularly useful once a feature of interest has been identified, for example, with the use of a guided tour [@cook_grand_1995].
-The path of a guided tour is selected via projection pursuit, the optimization of an index function on the projection via a hill climbing algorithm.
-This allows guided tours to identify interesting projection features rapidly given the relatively large parameter-space.
-Once the given projection has been provided, it is time to define the path of rotation.
-
-<!-- Manual tour paths -->
-Ideally, the path would be intuitively user-generated from physical movement, be it through mouse or motion capture.
-Unfortunately, this type of dynamic control has proven difficult to capture for in R.
-Because of this, manual tours were not implemented in **tourr** [@wickham_tourr_2011].
-This research allows for the consumption, but not the generation, of such dynamic input.
-After the capture of an oblique user motion, the rotation needs to be applied to step 3 (rotation sequence) of the algorithm discussed below.
-In the section below we stick with a radial rotation where, $\theta$, the angle of in-projection-plane rotation is held constant.
+This makes manual tours particularly useful once a feature of interest has been identified, for example, with the use of a guided tour [@hurley_analyzing_1990].
+Rather than exploring the large phase space of the data, a guided tour selects a very specific path via projection pursuit[@diaconis_asymptotics_1984], the optimization of an index function on the projection via a hill climbing algorithm.
+This allows guided tours to rapidly identify interesting projection features given the relatively large parameter-space.
+After a projection of interest is identified an analyst can then use the "finer brush" of the manual tour, by controlling the contributions of individual variables.
 
@@
 
+<!-- operate on bases -->
 The algorithm primarily operates on projection bases and utilizes the data only when making a display.
@@
 Figure \@ref{fig:step0} shows a locally optimized projection of the data.
-The left side displays the reference axes of the projection basis, a visual indication of the magnitude and direction each variable contributed to the projection.
-The right side shows the data as projected through the basis set described by the reference axes (left).
+The left panel displays the reference axes of the projection basis, a visual indication of the magnitude and direction each variable contributed to the projection.
+The right panel shows the data as projected through the basis set described by the reference axes (left).
 Data points are colored and given shape according to the species (while the guided tour was unsupervised with this information).
@@
 
-```{r step0, echo=F, results='hide', out.width='95%', fig.cap = "(ref:step0-cap)"}
+```{r step0, echo=F, results='hide', out.width='60%', fig.cap = "(ref:step0-cap)"}
 #"\\label{fig:step0} and some caption"}
@@
 
+<!-- step 0 disp code. -->
+```{r step0CODEEXAMPLE, eval=F, echo=T}
+# devtools::install_github("nspyrison/spinifex") # Development version
+install.package("spinifex")
+library("spinifex")
+
+# Also see vignette: 
+vignette("spinifex") # vignette ‘spinifex’ not found
+
+f_data  <- tourr::rescale(flea[,1:6])
+f_path  <- save_history(f_data, guided_tour(holes()))
+f_basis <- matrix(f_path[,, max(dim(f_path)[3])], ncol=2)
+f_cat   <- factor(flea$species)
+
+view_basis(basis = f_basis, 
+           data = f_data,
+           labels = colnames(f_data))
+```
+
 ### Step 1) Choose variable of interest
@@
 Initialize a zero vector **e** of length *p*.
-Set the fifth element to on, as  `aede2` is the fifth variable in the data, giving the manip var a full contribution in this dimension.
+Set the fifth element to one, as  `aede2` is the fifth variable in the data, giving the manip var a full contribution in this dimension.
 Use the Gram-Schmidt process to orthonormalize the zero vector onto the basis yielding the 3D manipulation space, **M**.
@@
 \begin{align*}
-  \textbf{e} &\leftarrow Orthonormalize_{GS}(\textbf{e}) ~~~~~ w.r.t. the basis \\
-  &= \textbf{e} - \langle \textbf{e},\textbf{B}_1 \rangle \textbf{B}_1 - \langle \textbf{e}, \textbf{B}_2 \rangle \textbf{B}_2 \\
+  \textbf{e} &\leftarrow Orthonormalize_{GS}(\textbf{e}) ~~~ w.r.t.~the~basis \\
+  &\leftarrow \textbf{e} - \langle \textbf{e},\textbf{B}_1 \rangle \textbf{B}_1 - \langle \textbf{e}, \textbf{B}_2 \rangle \textbf{B}_2 \\
   \\
@@
 
-```{r step2, echo=F, out.width='100%', fig.cap = "(ref:step2-cap)"}
+```{r step2, echo=F, out.width='40%', fig.cap = "(ref:step2-cap)"}
 view_manip_space(basis = f_bas, manip_var = f_mvar, labels = colnames(f_dat)) +
@@
 The representation in figure \@ref(fig:step2) can be produced by calling the function `view_manip_space()`.
+
+<!-- step 2 disp code. -->
+```{r step2CODEEXAMPLE, eval=F, echo=T}
+view_manip_space(basis = f_basis, 
+                 manip_var = 5, 
+                 labels = colnames(f_data))
+```
 
@@
 Let:
-\begin{description}
+\begin{itemize}
   \item[$c_\theta$] be the cosine of $\theta$
@@
   \item[$s_\phi$]   be the sine of   $\phi$
-\end{description}
+\end{itemize}
 
@@
 
-<!-- ```{r STEP3CODEEXAMPLE, eval=F, echo=T} -->
-<!-- for (phi in seq(seq_start, seq_end, phi_inc_sign)) { -->
-<!--   slide <- slide + 1 -->
-<!--   tour[,, slide] <- rotate_manip_space(manip_space, theta, phi)[, 1:2] -->
-<!-- } -->
-<!-- ``` -->
-
 <!-- description of slides, on the technical side we haven't gone back to data space yet. -->
@@
 
-## Data in projection-space {#sec:display}
+<!-- step 3 disp code. -->
+```{r step3CODEEXAMPLE, eval=F, echo=T}
+angle_speed <- .26
+
+play_manual_tour(data = f_data,
+                 basis = f_basis, 
+                 manip_var = 5, 
+                 angle = angle_speed,
+                 col = f_cat,
+                 pch = f_cat)
+```
+
+### Data in projection-space {#sec:display}
 
@@
 <!-- grand tours on the same data -->
-Using the same data, another study, @cook_dynamical_2018, applies grand tours to the same subspaces.
-Grand tours are dynamic subspace projections of high dimensional where frames and selected at random and linked with geodesically interpolation of the intermidiate frames.
-Grand tours are able to better resolve the distribution shape of clusters, intra-cluster detail, and better outlier detection than use of PDFSense & TFEP (TensorFlow embedded projections).
+Using the same data, another study, @cook_dynamical_2018, applies grand tours [@asimov_grand_1985] to the same subspaces.
+Grand tours are dynamic subspace projections of high dimensional where frames and selected at random and linked with geodesically interpolation of the intermediate frames.
+Grand tours are able to better resolve the distribution shape of clusters, intra-cluster detail, and better outlier detection than the use of PDFSense & TFEP (TensorFlow embedded projections).
 Before applying manual tours let's discuss the structure of the data.
@@
 <!-- Data structure -->
-The data has a hierachrical structure with top-level groupings; `DIS`, `VBP`, and `jet`.
-Each group is a particular class of experiments, each with many experimental datasets which, in turn, have many observations.
+The data has a hierarchical structure with top-level clusters; `DIS`, `VBP`, and `jet`.
+Each cluster is a particular class of experiments, each with many experimental datasets which, in turn, have many observations.
 In the consideration of data density, we conduct manual tours on subsets of the `DIS` and `jet` clusters.
@@
 
-<!-- motivate jet cluster, explain dimensionality  -->
-The `jet` cluster is of interest as it contains the largest data sets and is found to be important in @wang_mapping_2018.
-The jet cluster resides in a smaller dimensionality than the full set of experiments with 4 principal components explaining 95% of the variation in the jet cluster [@cook_dynamical_2018].
+<!-- jet cluster, explain dimensionality  -->
+The `jet` cluster resides in a smaller dimensionality than the full set of experiments with four principal components explaining 95% of the variation in the cluster [@cook_dynamical_2018].
 The data within this 4D embedding is subset down to `ATLAS7old` and `ATLAS7new` to focus in on two groups with a reasonable number of observations that occupy different parts of the subspace.
@@
 
-(ref:JetClusterBad-cap) Jet cluster, a radial manual tour of PC4.
+(ref:JetClusterBad-cap) `Jet` cluster, a radial manual tour of PC4.
 Colored by experiment type: ATLAS7new in green and ATLAS7old in orange.
-This manual tour contains less interesting information ATLAS7new (green) has points that are right and left of ATLAS7old, while most points occupy the same projection space, regardless of the contribution of PC4.
+This manual tour contains less interesting information ATLAS7new (green) has points that are right and left of ATLAS7old, while most points occupy the same projection space, independent of the contribution of PC4.
 A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/.
@@
 
-## TODO: edit application, Grammarly and word checks.
-<!-- discussion of findings and which is better -->
+## TODO: edit DIS cluster, Grammarly and word checks.
+<!-- discussion of findings and which is more insightful -->
 Manipulating PC3, where varying the angle of rotation brings interesting features into and out of the center mass of the data, is more interesting than the manipulation of PC4, where the features are mostly independent of the contribution of PC4.
 
-Jet cluster manual tours manipulating each of the principal components can be viewed from the links: [PC1](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc1/), [PC2](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc2/), [PC3](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/), and [PC4](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc4/).
+Radial manual tours manipulating each of the principal components in the `jet` cluster can be viewed by following the links: [PC1](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc1/), [PC2](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc2/), [PC3](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/), and [PC4](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc4/).
 
@@
 We perform a manual tour on this data, manipulating PC6 as depicted in figure \@ref(fig:DISclusterGood).
-Looking at several frames we see that DIS HERA data lies mostly on a plane.
-When PC6 has full contributions, we see the dimuon SIDIS in purple is almost orthogonal to the DIS HERA (green).
-Yet the contribution of PC6 has zeroed the dimuon SIDIS data occupy the same space as the DIS HERA data.
+Looking at several frames we see that `DIS HERA` (green) data lie mostly on a plane.
+When PC6 has full contributions, `dimuon SIDIS` (purple) is almost orthogonal to the `DIS HERA` (green).
+Yet the contribution of PC6 has zeroed the dimuon SIDIS data occupy the same space as the `DIS HERA` data.
 A dynamic version of this manual tour can be found at:
@@
 
-(ref:DISclusterGood-cap) DIS cluster, a radial manual tour of PC6.
-colored by experiment type: DIS HERA1+2 in green, dimuon SIDIS in purple, and charm SIDIS in orange.
-When the contribution PC 6 is large we see that dimuon SIDIS (purple) data are nearly orthogonal to DIS HERA (green) data.
-As the projection is rotated, we can also see that DIS HERA (green) practically lies on a plane in this 6D subspace.
-When the contribution of PC6 is near zero, dimonSIDIS (purple) occupies the same space as the DIS HERA data.
+(ref:DISclusterGood-cap) `DIS` cluster, a radial manual tour of PC6.
+colored by experiment type: `DIS HERA1+2` in green, `dimuon SIDIS` in purple, and `charm SIDIS` in orange.
+When the contribution PC 6 is large we see that dimuon SIDIS (purple) data are nearly orthogonal to `DIS HERA` (green) data.
+As the projection is rotated, we can also see that `DIS HERA` (green) practically lies on a plane in this 6D subspace.
+When the contribution of PC6 is near zero, `dimonSIDIS` (purple) occupies the same space as the `DIS HERA` data.
 A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/.
@@
 
-(ref:DISclusterBad-cap) DIS cluster, a radial manual tour of PC2.
-Colored by experiment type: DIS HERA1+2 in green, dimuon SIDIS in purple, and charm SIDIS in orange.
-The structure of previously described plane of DIS HERA (green) and nearly orthogonal dimuon SIDIS (purple) is present, however, the manipulating PC2 does not give a head-on view of either, a less useful manual tour than that of PC6.
+(ref:DISclusterBad-cap) `DIS` cluster, a radial manual tour of PC2.
+Colored by experiment type: `DIS HERA1+2` in green, `dimuon SIDIS` in purple, and `charm SIDIS` in orange.
+The plane of `DIS HERA` (green) and nearly orthogonal `dimuon SIDIS` (purple) is present, however, manipulating PC2 does not give a head-on view of either, offers less information about these features than manipulating PC6.
 A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/discluster_manualtour_pc2/.
@@
 
-## Source code and usage {#sec:usage}
+<!-- ## Source code and usage {#sec:usage} -->
 
-Use the below code as a guide for installation and finding the vignette.
+<!-- Use the below code as a guide for installation and finding the vignette.
 The vignette offers a less technical discussion opting to focus on code usage and goes through a couple more use cases.
 If you prefer to follow along with the example in the algorithm then simplified code is also listed below.
+-->
 
-```{r INSTALLCODEEXAMPLE, eval=F, echo=T}
-# devtools::install_github("nspyrison/spinifex") # Development version
-install.package("spinifex")
+<!-- ```{r INSTALLCODEEXAMPLE, eval=F, echo=T} -->
+<!-- # devtools::install_github("nspyrison/spinifex") # Development version -->
+<!-- install.package("spinifex") -->
+<!-- library("spinifex") -->
 
-# Also see vignette:
-vignette("spinifex") # vignette ‘spinifex’ not found
+<!-- # Also see vignette: -->
+<!-- vignette("spinifex") # vignette ‘spinifex’ not found -->
 
-## manual tour of std flea from holes-index:
-library(spinifex)
-f_dat  <- tourr::rescale(flea[,1:6])
-f_cat  <- factor(flea$species)
-f_path <- save_history(f_dat, guided_tour(holes()))
-f_bas  <- matrix(f_path[,, max(dim(f_path)[3])], ncol=2)
-f_mvar <- 5
-f_lab  <- colnames(f_dat)
+<!-- ## manual tour of std flea from holes-index: -->
+<!-- f_dat  <- tourr::rescale(flea[,1:6]) -->
+<!-- f_cat  <- factor(flea$species) -->
+<!-- f_path <- save_history(f_dat, guided_tour(holes())) -->
+<!-- f_bas  <- matrix(f_path[,, max(dim(f_path)[3])], ncol=2) -->
+<!-- f_mvar <- 5 -->
+<!-- f_lab  <- colnames(f_dat) -->
 
-# View the basis
-view_basis(f_bas, data = f_dat, labels = f_lab)
-# View the manip space
-view_manip_space(basis = f_bas, manip_var = f_mvar, labels = f_lab)
-# Play animation as HTML5 from plotly
-play_manual_tour(data = f_dat, basis = f_bas, manip_var = f_mvar, 
-                 col = f_cat, angle = f_angle)
-```
-
-### Acknowledgments
-
-This article was created in R [@r_core_team_r:_2018], using **bookdown** [@xie_bookdown:_2016] and **rmarkdown** [@xie_r_2018], with code generating the examples inline.
-The source files for this article be found at [github.com/nspyrison/confirmation/](https://github.com/nspyrison/confirmation/).
-The source code for the **spinifex** package can be found at [github.com/nspyrison/spinifex/](https://github.com/nspyrison/spinifex/).
+<!-- # View the basis -->
+<!-- view_basis(f_bas, data = f_dat, labels = f_lab) -->
+<!-- # View the manip space -->
+<!-- view_manip_space(basis = f_bas, manip_var = f_mvar, labels = f_lab) -->
+<!-- # Play animation as HTML5 from plotly -->
+<!-- play_manual_tour(data = f_dat, basis = f_bas, manip_var = f_mvar,  -->
+<!--                  col = f_cat, angle = f_angle) -->
+<!-- ``` -->
 
@@
 Additionally, the **animation** package @xie_animation:_2018 could be implemented for another graphics framework.
-However, **animation** builds from base graphs while **spinifex** utilizes **ggplot2** graphics.
+However, **animation** builds from **base** graphs while **spinifex** utilizes **ggplot2** graphics.
 
@@
 The user could easily switch between variables to control, adjust interpolation step angle, or flag/save specific frame basis sets.
+
+## Acknowledgments
+
+This article was created in R [@r_core_team_r:_2018], using **bookdown** [@xie_bookdown:_2016] and **rmarkdown** [@xie_r_2018], with code generating the examples inline.
+The source files for this article be found at [github.com/nspyrison/confirmation/](https://github.com/nspyrison/confirmation/).
+The source code for the **spinifex** package can be found at [github.com/nspyrison/spinifex/](https://github.com/nspyrison/spinifex/).

### 2019-04-09 paper/spinifex-paper.tex
remainder of changes to DIS. Grammerly and knit. add a wrap up summary.

--- 2019-04-05 paper/spinifex-paper.tex
@@
 \abstract{%
-The class of dynamic linear projections that are collectively known as `tours' provide a unique dynamic visualization of numeric multivariate data.
-Tours are particularly useful for understanding the structure held within multivariate data, and in association with techniques for dimension reduction, supervised, and unsupervised classification.
-The \emph{R} package \emph{tourr} offers a variety of path generators and geometric displays for conducting tours.
-This paper discusses an extension package, \emph{spinifex}, that adds support for the path generation of manual tours and extends the display of tours to use with the contemporary animation packages, \emph{plotly} and \emph{gganimate}.
+The set of dynamic linear projections of multivariate data collectively known as \textbf{tours} provides an important tool for extending the dimensionality of visuals.
+The R package \textbf{tourr} offers a variety of path generators and geometric displays for conducting tours.
+This paper discusses an extension package, \textbf{spinifex}, that adds support for the path generation of manual tours and extends the display of tours to use with the animation packages, \textbf{plotly} and \textbf{gganimate}.
 Manual tours are used to explore the sensitivity of structure as the contributions of a manipulation variable are changed.
-This particularly useful after identifying a feature of interest.
 A recent paper \{Wang et al.
-(2018)\} visualizes the sensitivey of the hadronic experiments to nucleon structure.
+(2018)\} visualizes the sensitivity of the hadronic experiments to nucleon structure.
 Sensitivity was characterized in non-linear 3D embeddings of the first 10 principal components.
-This research applies manual tours to this data showing that manual tours resolves more structrual information that is orthogonal to the original viewing plane.
+This research applies manual tours to this data showing that manual tours resolve information about cluster structure and identify outliers, information that is orthogonal to the original static viewing plane.
 }
@@
 Each frame of the sequence corresponds to a small change in the projection for a smooth transition that perseveres continuity.
+This is a crucial tool extending the dimensionality of visualization which is a large boon as datasets and models exist in increasing cardinality.
+The package \textbf{tourr} (Wickham et al.
+2011) builds a platform for generating tour paths and applying to a wide number of display geometrics in \textbf{base} graphics.
 
 While there are numerous methods that generate tour paths, this research focuses on the manual tour.
-The manual tour was described in Cook and Buja (1997) and allows a user to control the projection coefficients of a select variable has in a 2D projection.
+The manual tour was described in Cook and Buja (1997) and allows a user to control the projection coefficients of a selected variable in a 2D projection.
 The manipulation of these coefficients allows the analyst to explore how sensitive the projections structure is to these changes.
-This makes manual tours particularly useful once a feature of interest has been identified, for example, with the use of a guided tour (Cook et al.
-1995).
-The path of a guided tour is selected via projection pursuit, the optimization of an index function on the projection via a hill climbing algorithm.
-This allows guided tours to identify interesting projection features rapidly given the relatively large parameter-space.
-Once the given projection has been provided, it is time to define the path of rotation.
+This makes manual tours particularly useful once a feature of interest has been identified, for example, with the use of a guided tour (Hurley and Buja 1990).
+Rather than exploring the large phase space of the data, a guided tour selects a very specific path via projection pursuit(Diaconis and Freedman 1984), the optimization of an index function on the projection via a hill climbing algorithm.
+This allows guided tours to rapidly identify interesting projection features given the relatively large parameter-space.
+After a projection of interest is identified an analyst can then use the ``finer brush'' of the manual tour, by controlling the contributions of individual variables.
 
-Ideally, the path would be intuitively user-generated from physical movement, be it through mouse or motion capture.
-Unfortunately, this type of dynamic control has proven difficult to capture for in R.
-Because of this, manual tours were not implemented in \emph{tourr}.
-This research allows for the consumption, but not the generation, of such dynamic input.
-After the capture of an oblique user motion, the rotation needs to be applied to step 3 (rotation sequence) of the algorithm discussed below.
-In the section below we stick with a radial rotation where, \(\theta\), the angle of in-projection-plane rotation is held constant.
-
-Spinifex utilizes two new animation packages, \emph{plotly} (Sievert 2018) and \emph{gganimate} (Pedersen and Robinson 2019), to display tours, manual or other saved tours.
+Spinifex utilizes two new animation packages, \textbf{plotly} (Sievert 2018) and \textbf{gganimate} (Pedersen and Robinson 2019), to display tours, manual or other saved tours.
 From a given projection, the user can choose which variable to control, and the animation sequence is generated to remove the variable from the projection, and then extend its contribution to be the sole variable in one direction.
@@
   \item $\textbf{X}$, the data, an $n \times p$ numeric matrix to be embedded in two dimensions.
-  \item $\textbf{B} = (B_1,~ B_2)$, any orthonormal projection basis set, $p \times 2$ matrix, describing the projection from $p$ to two dimensions.
+  \item $\textbf{B} = (B_1,~ B_2)$, any orthonormal projection basis set, $p \times 2$ matrix, describing the projection from $p$ to two dimensions
   \item $\textbf{e}$, a zero column vector of length $p$ with the $k-$th element set to one, where $k$ is the number of the variable to manipulate.
@@
 
-The flea data, originally from Lubischew (1962), available in the R package \emph{tourr} (Wickham et al.
-2011) is used to illustrate the algorithm.
+The flea data, originally from Lubischew (1962), made available in \textbf{tourr} is used to illustrate the algorithm.
 The data contains 74 observations across 6 variables, physical measurements of the flea beetles.
@@
 The data is defined.
-A basis set (ideally that views an interesting feature) should be provided to explore the sensitivity of the variables to the structure.
-To identify a projection containing an interesting feature, apply
-a guided tour(Cook, Swayne, and Buja 2007) on the flea data.
-In a guided tour the projection sequence is selected by optimizing an index via hill-climbing.
-In this case, the holes index is selected.
-The holes index is maximized by when the projected observations are furthest from the center.
-Figure @ref(fig:step0) shows a locally optimized projection for this data.
-The left plot displays the reference axes of the projection basis, a visual indication of the magnitude and direction each variable contributed to the projections.
-The right plot shows the projection of the data through the basis set described by the reference axes (left).
-Data points are colored and given point characters according to the species of the flea (the guided tour was unsupervised with this information).
+A basis set (ideally orienting to an interesting feature) must be provided as an initial orientation.
+One way to identify a projection containing interesting features is to apply a guided tour (Cook, Swayne, and Buja 2007).
+In a guided tour the projection sequence is selected by optimizing an index function via hill-climbing on the projection space.
+In this case, the holes index is selected and applied to standardized flea data.
+The holes index is maximized when the projected observations are furthest from the center.
+Figure @ref\{fig:step0\} shows a locally optimized projection of the data.
+The left panel displays the reference axes of the projection basis, a visual indication of the magnitude and direction each variable contributed to the projection.
+The right panel shows the data as projected through the basis set described by the reference axes (left).
+Data points are colored and given shape according to the species (while the guided tour was unsupervised with this information).
+
+(ref:step0-cap) Basis reference axes (left) and projected data (right) of standardized flea data.
+Data points color and shape are mapped to beetle species.
+Basis identified by a holes-index guided tour.
+The variables \texttt{aede2} and \texttt{tars1} contribute mostly orthogonal to the other variables.
 
@@
 
-{\centering \includegraphics[width=0.95\linewidth]{spinifex_paper_files/figure-latex/step0-1}
+{\centering \includegraphics[width=0.6\linewidth]{spinifex_paper_files/figure-latex/step0-1}
 
@@
 
-(ref:step0-cap) Basis reference axes (left) and projected data (right) of standardized flea data.
-Data points color and shape are mapped to beetle species.
-Basis identified by a holes-index guided tour.
-The variables \texttt{aede2} and \texttt{tars1} contribute mostly orthogonal to the other variables.
-We'll select \texttt{aede2} as our manipulation variable to see how the structure of the projection changes as we rotate \texttt{aede2} into and out of the projection.
+Call \texttt{view\_basis()} on a basis to produce a \textbf{ggplot2} graphic similar to @ref(fig:step0).
+Projection space is always available for display via the matrix multiplication \(\textbf{X}_{[n,~p]} ~*~ \textbf{B}_{[p,~d]} ~=~ \textbf{P}_{[n,~d]}\).
 
-Call \texttt{view\_basis()} on a basis to produce a \emph{ggplot2} graphic similar to @ref(fig:step0).
-Projection space is always available for display via the matrix multiplication \(\textbf{X}_{[n,~p]} ~*~ \textbf{B}_{[p,~d]} ~=~ \textbf{P}_{[n,~d]}\).
+\begin{Schunk}
+\begin{Sinput}
+# devtools::install_github("nspyrison/spinifex") # Development version
+install.package("spinifex")
+library("spinifex")
+
+# Also see vignette: 
+vignette("spinifex") # vignette ‘spinifex’ not found
+
+f_data  <- tourr::rescale(flea[,2:6])
+f_path  <- save_history(f_data, guided_tour(holes()))
+f_basis <- matrix(f_path[,, max(dim(f_path)[3])], ncol=2)
+f_cat   <- factor(flea$species)
+
+view_basis(basis = f_basis, 
+           data = f_data,
+           labels = colnames(f_data))
+\end{Sinput}
+\end{Schunk}
 
@@
 
-In figure @ref(fig:step0), above, the contributions of the variables \texttt{tars1} and \texttt{aede2} are mostly orthogonal to the contributions of the other four variables.
-These two variables explain the variation of the data between the purple and green species.
-We select \texttt{aede2} as the manip var, the variable to be manipulated as it typically has a larger contribution after the optimizing the holes index.
-The question that will be explored in the explanation of the algorithm is how important the variable \texttt{aede2} is to the separation of the clusters.
+In figure @ref(fig:step0) the contributions of the variables \texttt{tars1} and \texttt{aede2} are mostly orthogonal to the contributions of the other four variables.
+These two variables explain the variation of the data distinguishing the purple group from the rest of the sample.
+Select \texttt{aede2} as the manip var, the variable to be manipulated, as it has a larger contribution to the projection.
+The question that will be explored is how important the variable \texttt{aede2} is to the separation of the clusters.
 
@@
 
-Initialize a zero vector \(e\) of \(p\) elements.
-Because \texttt{aede2} is the fifth variable in the data, set the \(k=5\)-th element to one giving the manip var a full contribution in this dimension.
+Initialize a zero vector \textbf{e} of length \emph{p}.
+Set the fifth element to one, as \texttt{aede2} is the fifth variable in the data, giving the manip var a full contribution in this dimension.
 Use the Gram-Schmidt process to orthonormalize the zero vector onto the basis yielding the 3D manipulation space, \textbf{M}.
@@
 \begin{align*}
-  \textbf{e} &\leftarrow Orthonormalize_{GS}(\textbf{e}) w.r.t. Basis \\
-  &= \textbf{e} - \langle \textbf{e},\textbf{B}_1 \rangle \textbf{B}_1 - \langle \textbf{e}, \textbf{B}_2 \rangle \textbf{B}_2 \\
+  \textbf{e} &\leftarrow Orthonormalize_{GS}(\textbf{e}) ~~~ w.r.t.~the~basis \\
+  &\leftarrow \textbf{e} - \langle \textbf{e},\textbf{B}_1 \rangle \textbf{B}_1 - \langle \textbf{e}, \textbf{B}_2 \rangle \textbf{B}_2 \\
   \\
@@
 
-Adding this extra dimension to our basis plane allows for the coefficients of the specified variable to be changed.
-For example, the ability to lift a piece of paper, rather than being constrained to the motion on a table top.
-Orthonormalizing rescales the new depth vector while the projection down to 2D is the original basis, that is the first \(d\) vectors remain constant.
-Imagine the reference axes (and projection plane) laying flat on a table, while a new dimension exists with axes projecting back onto the reference axes.
-An illustration of such can be seen below in figure @ref(fig:step2).
-The manip var is highlighted, while the depths of the other variables are not depicted.
+Adding this new dimension to our projection space allows for the coefficients of the manip var to be changed via rotation about the origin.
+For example, the ability to lift a piece of paper, rather than being constrained to a 2D plane.
+Orthonormalizing rescales the new vector while the projection down to 2D remains the original basis.
+Place the plane horizontally, with the new dimension extending vertically, with axes projecting back onto the reference axes.
+Figure @ref(fig:step2) illustrates this 3D manipulation space with the manip var highlighted (height of the other variables are not depicted.)
 
 (ref:step2-cap) Manipulation space for controlling the contribution of \texttt{aede2} of standardized flea data.
-Basis selected by a holes-index guided tour.
-The Projection plane is shown in blue.
-The manipulation axis, in red, allows the coefficients of the manip var to be changed.
+Basis selected by a holes-index optimized guided tour lies on the horizontal projection plane, shown in blue.
+The manip var axis, in red, extends into the vertical manipulation space, allowing the coefficients of the manip var to be changed by rotation around the origin.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{spinifex_paper_files/figure-latex/step2-1}
+{\centering \includegraphics[width=0.6\linewidth]{spinifex_paper_files/figure-latex/step2-1}
 
@@
 
-The representation in @ref(fig:step2) can be duplicated by calling the function \texttt{view\_manip\_space()}.
+The representation in figure @ref(fig:step2) can be produced by calling the function \texttt{view\_manip\_space()}.
+
+\begin{Schunk}
+\begin{Sinput}
+view_manip_space(basis = f_basis, 
+                 manip_var = 4, 
+                 labels = colnames(f_data))
+\end{Sinput}
+\end{Schunk}
 
@@
 
-Imagine holding the red axis it is fixed to the origin.
-As it is manipulated the projection back onto the projection plane correspondingly moves.
+Imagine holding the manip var, the red axis, one end fixed to the origin.
+As it is controlled the manipulation space rotates about the origin, the projection onto the horizontal projection plane correspondingly moves.
 This is what happens in a manual tour.
-For a radial tour, fix \(\theta\), the angle within the blue plane, and vary the sequence of \(\phi\), the angle coming out of the projection plane.
-Conceptually, live manipulation on a 2D plane allows the user to dynamically control these angles, effectively changing the coefficients of the manip var, which then performs a constrained rotation on the remaining variables.
+Generating a sequence of values for the horizontal and vertical, angles produces a path for the rotation of the manipulation space.
+This defines the (orthonormally-constrained) rotation on the coefficients of the variables.
 
-For the demonstration of the radial tour, we define a sequence for \(\phi\) that brings the initial contribution of the manip var to be maximized and then zeroed before returning to the initial position.
+For a radial tour fix the (horizontal) angle within the projection plane, \(\theta\), and define a sequence for the (vertical) angle coming out of the projection plane, \(\phi\), bringing the initial \texttt{XY} contributions of the manip var to a maximum and then to zero before returning to the initial position.
+Dynamic capture of user manipulation is typically performed directly on the projection plane (without depiction of the manipulation space.)
 
@@
 
-\begin{description}
+\begin{itemize}
   \item[$c_\theta$] be the cosine of $\theta$
@@
     c_\phi
-  \end{bmatrix}_{[3,~3]}
+  \end{bmatrix}_{[3,~3,~i]}
 \end{align*}
 
-A note on application: compile the sequence of \(\phi_i\) and create an array/long table for each rotated manipulation space.
-\(\phi\) is the angle relative to the initial value of \(\phi\), we find the transformation \(\phi_i\) - \(\phi_1\) useful to think about \(\phi\) relative to the basis plane.
-Additionally, the value of \(\phi\) may be offset by a factor of pi.
+A note on application: \(\phi\) is the angle relative to the initial value of \(\phi\), we find the transformation \(\phi_i\) - \(\phi_1\) useful to think about \(\phi\) relative to the basis plane.
+Additionally, the value of \(\phi\) may be out of phase by a factor of pi.
 If the manip variable doesn't move as expected these are the first places to check.
 
-\begin{Schunk}
-\begin{Sinput}
-for (phi in seq(seq_start, seq_end, phi_inc_sign)) {
-  slide <- slide + 1
-  tour[,, slide] <- rotate_manip_space(manip_space, theta, phi)[, 1:2]
-}
-\end{Sinput}
-\end{Schunk}
-
-Figure @ref(fig:step3) illustrates a sequence with 15 projected bases and highlight the manip variable on top while showing the corresponding projected data points on the bottom.
-Take note of how the changes in the manip var change the distance between the purple and green cluster of points, \texttt{aede2} is crucial in distinguishing between these groups.
+Figure @ref(fig:step3) illustrates a sequence with 15 projected bases, showing the reference axes on top the corresponding projected data points on the below.
+As a result, we can see that changes in the manip var controlled the distance between the purple cluster and the remaining sample, \texttt{aede2} is crucial in distinguishing this species.
 Tours are typically viewed as an animation such a dynamic version of this tour can be viewed online at \url{https://nspyrison.netlify.com/thesis/flea_manualtour_mvar5/}.
 The page may take a moment to load.
-The format of this figure and linking to an HTML animation will be used again in the Application, section \ref{sec:application}.
 
-(ref:step3-cap) Radial manual tour changing the contributions from \texttt{aede2} of standardized flea data.
+(ref:step3-cap) Radial manual tour manipulating \texttt{aede2} of standardized flea data.
 The contributions increase from its initial contribution to a full contribution to the projection before decreasing to zero and then returning to its initial value.
-The change in the projected data shows that \texttt{aede2} is important for distinguishing between the purple and green clusters.
+The change in the projected data shows that \texttt{aede2} is important for distinguishing the purple species.
 An animated version can be viewed at \url{https://nspyrison.netlify.com/thesis/flea_manualtour_mvar5/}.
@@
 
-Animations can be produced using the function \texttt{play\_manual\_tour()}. This function defaults to an HTML5 widget produced from \emph{plotly}.
+Animations can be produced using the function \texttt{play\_manual\_tour()}.
+This function defaults to an HTML5 widget produced from \textbf{plotly}.
+The \texttt{render\_type} argument can be changed to \texttt{render\_gganimate} for exporting to .gif or .mp4 files.
+
+\begin{Schunk}
+\begin{Sinput}
+angle_speed <- .26
+
+play_manual_tour(data = f_data,
+                 basis = f_basis, 
+                 manip_var = 5, 
+                 angle = angle_speed,
+                 col = f_cat,
+                 pch = f_cat)
+\end{Sinput}
+\end{Schunk}
 
@@
 
-In light of performance, the above operations are performed on the bases without the use of the larger datasets.
+In an appeal to performance, the above operations are performed on the bases without the use of the larger datasets.
 After the bases are brought into the projection-space, however, it is helpful to observe them with data in the same space.
@@
 
-The \emph{tourr} package utilizes R's base graphics for the display of tours.
-\emph{spinifex} allows tours to be used in rendered in \emph{plotly} Sievert (2018) as an HTML5 object or \emph{gganimate} Pedersen and Robinson (2019) as .gif or .mp4 objects.
-Both of which build off \emph{ggplot2} objects in internal functions.
-Sharing of animations is not trivial especially in print and static formats such as .pdf.
+The \textbf{tourr} package utilizes \textbf{base} graphics for the display of tours.
+\textbf{spinifex} allows tours to be used in rendered in \textbf{plotly} as an HTML5 object or \textbf{gganimate} as .gif or .mp4 objects.
+Both of which build off \textbf{ggplot2} objects in internal functions.
+Sharing of animations is not trivial especially in print and static file formats such as .pdf.
 Even with the use of computers and dynamic file formats capturing the correct resolution, aspect, and display is challenging and many formats quickly bloat file sizes.
-Keep in mind hosting options and exporting functions from \emph{plotly}, \emph{gganimate} and \emph{tourr}.
+Keep in mind hosting options and exporting functions from \textbf{plotly}, \textbf{gganimate} and \textbf{tourr}.
 
@@
 
-Storing each data point for every frame of the animation is very inefficient.
-Just as operations are performed on the bases, so too should tour paths be stored as bases.
+Storing each data point for every frame of the animation is redundant.
+Just as operations are performed on the bases, so too should tour paths be stored as bases and a single instance of the data.
 Consider a radial manual tour, we can store the salient features in 3 bases, where \(\phi\) is at its starting, minimum, and maximum values.
@@
 In a recent paper, Wang et al.
-(2018), the authors aggregate and visualize the sensitivity of hadronic experiments to ncleon structure.
+(2018), the authors aggregate and visualize the sensitivity of hadronic experiments to nucleon structure.
 The authors introduce a new tool, PDFSense, to aid in the visualization of Parton distribution functions (PDF).
@@
 
-Using the same data, another study, Cook, Laa, and Valencia (2018), applies grand tours to the same subspaces.
-Grand tours are able to better resolve the distribution shape of clusters, intra-cluter detail, better outlier detection, and exonerate a claim persened from TFEP (TensorFlow embedded projections).
-Table 1 of Cook et al.~summarizes the key findings of observations made with PDFSense \& TFEP and those from grand tours.
+Using the same data, another study, Cook, Laa, and Valencia (2018), applies grand tours (Asimov 1985) to the same subspaces.
+Grand tours are dynamic subspace projections of high dimensional where frames and selected at random and linked with geodesically interpolation of the intermediate frames.
+Grand tours are able to better resolve the distribution shape of clusters, intra-cluster detail, and better outlier detection than the use of PDFSense \& TFEP (TensorFlow embedded projections).
+Before applying manual tours let's discuss the structure of the data.
 
-Without getting too domain-specific the data has three primary groupings; DIS, VBP, and jet Each group is a particular class of experiments and each with many experimental datasets which inturn have many observation.
-Inconsideration of data density and business of the data We conduct manual tours on a subsets of the DIS and jet clusters.
-This explores the sensitivity of the structue to each of the variables in turn, and we present the subjectively best and worst manaul tour identifying structure in the respective data sets.
+The data has a hierarchical structure with top-level clusters; \texttt{DIS}, \texttt{VBP}, and \texttt{jet}.
+Each cluster is a particular class of experiments, each with many experimental datasets which, in turn, have many observations.
+In the consideration of data density, we conduct manual tours on subsets of the \texttt{DIS} and \texttt{jet} clusters.
+This explores the sensitivity of the structure to each of the variables in turn and we present the subjectively best and worst variable to manipulate for identifying dimensionality of the clusters and describing the span of the clusters.
 
@@
 
-The jet cluster is of interest as it contains the largest data sets and is found to be important in Wang et al.
-(2018) The jet cluster resides in a smaller dimensionality than the full set of experiments with 4 principal components explaining 95\% of the variation in the jet cluster (Cook, Laa, and Valencia 2018).
-The data is subset down to ATLAS7old and ATLAS7new to focus in on two groups with a reasonable number of observations that occupy different parts of the subspace.
-Below, we perform radial manual tours all four principal components within this scope.
-Visualizing PC3 and PC4 in figure @ref(fig:JetClusterGood) (more sturcturally insightful) and figure @ref(fig:JetClusterBad) (less sturcturally insightful) respectively, and list links to dynamic animation of all variables.
+The \texttt{jet} cluster resides in a smaller dimensionality than the full set of experiments with four principal components explaining 95\% of the variation in the cluster (Cook, Laa, and Valencia 2018).
+The data within this 4D embedding is subset down to \texttt{ATLAS7old} and \texttt{ATLAS7new} to focus in on two groups with a reasonable number of observations that occupy different parts of the subspace.
+Radial manual tours controlling contributions from PC4 and PC3 are shown in figure @ref(fig:JetClusterGood) and figure @ref(fig:JetClusterBad) respectively.
+These components are selected to contrast the difference of information conveyed by touring different variables.
+Links to dynamic HTML5 animations controlling each of the four variables are also provided.
 
-(ref:JetClusterGood-cap) Jet cluster, a radial manual tour of PC3.
-Colored by experiment type: `ATLAS7new' in green and `ATLAS7old' in orange.
-When PC3 fully contributes to the projection ATLAS7new (green) occupies unique space and several outliers are identifiable.
-Zeroing the contribution from PC3 to the projection hides the outliers and indeed all observations with ATLAS7new are contained within ATLAS7old (orange).
-A dynamic version can be viewed at \url{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/}.
+When manipulating PC4, there is clear difference in the parameter space spanned by the experiment types \texttt{ATLAS7new} and \texttt{ATLAS7old}.
+Yet, when PC3 is manipulated there is no indication that the diffent expiments probe different parameter space.
+Performing a radial manual tour on PC4 is more insightful than for PC3.
+
+Radial manual tours manipulating each of the principal components in the \texttt{jet} cluster can be viewed by following the links: \href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc1/}{PC1}, \href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc2/}{PC2}, \href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/}{PC3}, and \href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc4/}{PC4}.
+
+(ref:JetClusterGood-cap) A radial manual tour of PC4 within the \texttt{jet} cluster.
+Colored by experiment type: \texttt{ATLAS7new} in green and \texttt{ATLAS7old} in orange.
+When PC4 fully/negligably contributes to the projection \texttt{ATLAS7new} (green) spans the same space as the orange points.
+During the intermediate frames the \texttt{ATLAS7new} is compressed in the direction radial to PC4.
+A dynamic version can be viewed at \url{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc4/}.
 
@@
 
-(ref:JetClusterBad-cap) Jet cluster, a radial manual tour of PC4.
-Colored by experiment type: ATLAS7new in green and ATLAS7old in orange.
-This manual tour contains less interesting information ATLAS7new (green) has points that are right and left of ATLAS7old, while most points occupy the same projection space, regardless of the contribution of PC4.
+(ref:JetClusterBad-cap) A radial manual tour of PC4 within the \texttt{jet} cluster.
+Colored by experiment type: \texttt{ATLAS7new} in green and \texttt{ATLAS7old} in orange.
+Data from \texttt{ATLAS7new} (green) spans the mostly the same space as \texttt{ALTLAS7old} (orange) with no evident differnce in cluster structure across varying contributions of PC3.
 A dynamic version can be viewed at \url{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/}.
@@
 
-\hypertarget{todo-edit-application-grammarly-and-word-checks.}{%
-\subsection{TODO: edit application, Grammarly and word checks.}\label{todo-edit-application-grammarly-and-word-checks.}}
-
-Manipulating PC3, where varying the angle of rotation brings interesting features into and out of the center mass of the data, is more interesting than the manipulation of PC4, where the features are mostly independent of the contribution of PC4.
-
-Jet cluster manual tours manipulating each of the principal components can be viewed from the links: \href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc1/}{PC1}, \href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc2/}{PC2}, \href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/}{PC3}, and \href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc4/}{PC4}.
-
 \hypertarget{dis-cluster}{%
@@
 
-We perform a manual tour on this data, manipulating PC6 as depicted in figure @ref(fig:DISclusterGood).
-Looking at several frames we see that DIS HERA data lies mostly on a plane.
-When PC6 has full contributions, we see the dimuon SIDIS in purple is almost orthogonal to the DIS HERA (green).
-Yet the contribution of PC6 has zeroed the dimuon SIDIS data occupy the same space as the DIS HERA data.
-A dynamic version of this manual tour can be found at:
-\url{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/}.
-The page may some time to load, as the animation is several megabytes.
+A different space is used to explore the \texttt{DIS} cluster; specfically the first six pricipal components, which explain 48\% of the variation contained within the aggregated data(Cook, Laa, and Valencia 2018).
+Radial manual tours are performed on PC6 and PC2 in figure @ref(fig:DISclusterGood) and figure @ref(fig:DISclusterBad) respectively.
 
-(ref:DISclusterGood-cap) DIS cluster, a radial manual tour of PC6.
-colored by experiment type: `DIS HERA1+2' in green, `dimuon SIDIS' in purple, and `charm SIDIS' in orange.
-When the contribution PC 6 is large we see that dimuon SIDIS (purple) data are nearly orthogonal to DIS HERA (green) data.
-As the projection is rotated, we can also see that DIS HERA (green) practically lies on a plane in this 6D subspace.
-When the contribution of PC6 is near zero, dimonSIDIS (purple) occupies the same space as the DIS HERA data.
+The selection of the manip variable is important as the manipulation spaces convey substantially different information.
+The manual tour of PC6 offers information about the dimensionality, shape, and orientations of the differnet experiment classes.
+whereas manipulating the contributions of PC2 only shows a subset of the dimensionality and shape information.
+Manipulating the contributions of PC6 turned out to be much more insightful than that of PC2.
+This result might seem counter-intuitive at first as PC2 should explain much more of the variation in the data.
+However, features and structure in the data regularly reside in finer detail that can be lost when looking only at static projections.
+
+\texttt{DIS} cluster manual tours manipulating each of the principal components can be viewed from the links: \href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc1/}{PC1}, \href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc2/}{PC2}, \href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc3/}{PC3}, \href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc4/}{PC4}, \href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc5/}{PC5}, and \href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/}{PC6}.
+
+(ref:DISclusterGood-cap) A radial manual tour manipulating the contribution of PC6 within the \texttt{DIS} cluster.
+Points are colored by experiment type: \texttt{DIS\ HERA1+2} in green, \texttt{dimuon\ SIDIS} in purple, and \texttt{charm\ SIDIS} in orange.
+The cluster \texttt{DIS\ HERA1+2} (green) is distributed in a cross shaped plane, \texttt{charm\ SIDIS} (orange) occupies the the center space of this cross, with the plane projecting into the field of view when the contribution of PC6 is max.
+Less evident is the linear \texttt{dimuon\ SIDIS} (purple) observations approaching the line of view for intermediate values of PC6.
 A dynamic version can be viewed at \url{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/}.
@@
 
-The selection of the correct manip variable is important as the manipulation spaces convey different information.
-For example, in figure @ref(fig:DISclusterBad) we select PC2 as the manip variable finding it to be less insightful than PC6.
-
-(ref:DISclusterBad-cap) DIS cluster, a radial manual tour of PC2.
-Colored by experiment type: `DIS HERA1+2' in green, `dimuon SIDIS' in purple, and `charm SIDIS' in orange.
-The structure of previously described plane of DIS HERA (green) and nearly orthogonal dimuon SIDIS (purple) is present, however, the manipulating PC2 does not give a head-on view of either, a less useful manual tour than that of PC6.
+(ref:DISclusterBad-cap) A radial manual tour manipulating the contribution of PC2 within the \texttt{DIS} cluster.
+Points are colored by experiment type: \texttt{DIS\ HERA1+2} in green, \texttt{dimuon\ SIDIS} in purple, and \texttt{charm\ SIDIS} in orange.
+The plane of cross distributed \texttt{DIS\ HERA} data (green) and nearly orthogonal jet of \texttt{dimuon\ SIDIS} (purple) is present.
+This jet does entend more in the plane of view when the contribution of PC2 is full.
+However, less information about the shape of \texttt{DIS\ HERA} (green) and \texttt{charm\ SIDIS} (orange) is available.
 A dynamic version can be viewed at \url{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc2/}.
@@
 
-DIS cluster manual tours manipulating each of the principal components can be viewed from the links: \href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc1/}{PC1}, \href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc2/}{PC2}, \href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc3/}{PC3}, \href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc4/}{PC4}, \href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc5/}{PC5}, and \href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/}{PC6}.
+\hypertarget{todo}{%
+\subsection{TODO:}\label{todo}}
 
-\hypertarget{sec:usage}{%
-\subsection{Source code and usage}\label{sec:usage}}
-
-Use the below code as a guide for installation and finding the vignette.
-The vignette offers a less technical discussion opting to focus on code usage and goes through a couple more use cases.
-If you prefer to follow along with the example in the algorithm then simplified code is also listed below.
-
-\begin{Schunk}
-\begin{Sinput}
-# devtools::install_github("nspyrison/spinifex") # Development version
-install.package("spinifex")
-
-# Also see vignette:
-vignette("spinifex") # vignette ‘spinifex’ not found
-
-## manual tour of std flea from holes-index:
-library(spinifex)
-f_dat  <- tourr::rescale(flea[,1:6])
-f_cat  <- factor(flea$species)
-f_path <- save_history(f_dat, guided_tour(holes()))
-f_bas  <- matrix(f_path[,, max(dim(f_path)[3])], ncol=2)
-f_mvar <- 5
-f_lab  <- colnames(f_dat)
-
-# View the basis
-view_basis(f_bas, data = f_dat, labels = f_lab)
-# View the manip space
-view_manip_space(basis = f_bas, manip_var = f_mvar, labels = f_lab)
-# Play animation as HTML5 from plotly
-play_manual_tour(data = f_dat, basis = f_bas, manip_var = f_mvar, 
-                 col = f_cat, angle = f_angle)
-\end{Sinput}
-\end{Schunk}
-
-\hypertarget{acknowledgments}{%
-\subsubsection{Acknowledgments}\label{acknowledgments}}
-
-This article was created in \emph{R} (R Core Team 2018), using \emph{bookdown} (Xie 2016) and \emph{rmarkdown} (Xie, Allaire, and Grolemund 2018), with code generating the examples inline.
-The source files for this article be found at \href{https://github.com/nspyrison/confirmation/}{github.com/nspyrison/confirmation/}.
-The source code for the \emph{spinifex} package can be found at \href{https://github.com/nspyrison/spinifex/}{github.com/nspyrison/spinifex/}.
+\hypertarget{summary}{%
+\subsubsection{Summary}\label{summary}}
 
@@
 Tours, the dynamic linear projection of multivariate data, is an important aspect of data visualization extending the display of data-space as data dimensionality increases.
-This research has modified the algorithm producing manual tours, applied this functionality in \emph{R} and offers extends the graphics offerings that can be used to display tours.
+This research has modified the algorithm producing manual tours, applied this functionality in R and offers extends the graphics offerings that can be used to display tours.
 The paragraphs below explore how this work might be extended.
@@
 
-The \emph{tourr} package provides many other geometric displays with the \texttt{tourr::display\_*()} family.
-These geometric options could be integrated into the \emph{ggplot2} framework for display on \emph{plotly} and \emph{gganimate}.
-Additionally, the \emph{animation} package Xie et al.
+The \textbf{tourr} package provides many other geometric displays with the \texttt{tourr::display\_*()} family.
+These geometric options could be integrated into the \textbf{ggplot2} framework for display on \textbf{plotly} and \textbf{gganimate}.
+Additionally, the \textbf{animation} package Xie et al.
 (2018) could be implemented for another graphics framework.
-However, \emph{animation} builds from base graphs while \emph{spinifex} utilizes \emph{ggplot2} graphics.
+However, \textbf{animation} builds from \textbf{base} graphs while \textbf{spinifex} utilizes \textbf{ggplot2} graphics.
 
@@
 Having a script only interaction with tours causes a significant barrier to entry.
-To a lesser extent, \emph{plotly} offers some static interactions with the contained object, such as tooltips, brushing, and linking without communicating back to the R console.
-The development of a dynamic graphical user interface, perhaps with the use of a \emph{shiny} (Chang et al.
+To a lesser extent, \textbf{plotly} offers some static interactions with the contained object, such as tooltips, brushing, and linking without communicating back to the R console.
+The development of a dynamic graphical user interface, perhaps with the use of a \textbf{shiny} (Chang et al.
 2018) application, would mitigate the barrier to entry, allow for more rapid analysis, and offer an approachable demo tool.
@@
 
+\hypertarget{acknowledgments}{%
+\subsection{Acknowledgments}\label{acknowledgments}}
+
+This article was created in R (R Core Team 2018), using \textbf{bookdown} (Xie 2016) and \textbf{rmarkdown} (Xie, Allaire, and Grolemund 2018), with code generating the examples inline.
+The source files for this article be found at \href{https://github.com/nspyrison/confirmation/}{github.com/nspyrison/confirmation/}.
+The source code for the \textbf{spinifex} package can be found at \href{https://github.com/nspyrison/spinifex/}{github.com/nspyrison/spinifex/}.
+
 \hypertarget{refs}{}
+\leavevmode\hypertarget{ref-asimov_grand_1985}{}%
+Asimov, Daniel. 1985. ``The Grand Tour: A Tool for Viewing Multidimensional Data.'' \emph{SIAM Journal on Scientific and Statistical Computing} 6 (1): 128--43.
+
 \leavevmode\hypertarget{ref-buja_computational_2005}{}%
@@
 
-\leavevmode\hypertarget{ref-cook_grand_1995}{}%
-Cook, Dianne, Andreas Buja, Javier Cabrera, and Catherine Hurley.
-1995.
-``Grand Tour and Projection Pursuit.'' \emph{Journal of Computational and Graphical Statistics} 4 (3): 155.
-\url{https://doi.org/10.2307/1390844}.
-
 \leavevmode\hypertarget{ref-cook_dynamical_2018}{}%
@@
 
+\leavevmode\hypertarget{ref-diaconis_asymptotics_1984}{}%
+Diaconis, Persi, and David Freedman. 1984. ``Asymptotics of Graphical Projection Pursuit.'' \emph{The Annals of Statistics} 12 (3): 793--815.
+
 \leavevmode\hypertarget{ref-furnas_prosection_1994}{}%
@@
 \url{https://doi.org/10.2307/1390897}.
+
+\leavevmode\hypertarget{ref-hurley_analyzing_1990}{}%
+Hurley, C., and A.
+Buja.
+1990.
+``Analyzing High-Dimensional Data with Motion Graphics.'' \emph{SIAM Journal on Scientific and Statistical Computing} 11 (6): 1193--1211.
+\url{https://doi.org/10.1137/0911068}.
 

### 2019-04-09 paper/spinifex_paper.Rmd
remainder of changes to DIS. Grammerly and knit. add a wrap up summary.

@@
 ---
+\bibliography{RJrefs}
 
-\bibliography{RJrefs}
 ```{r setup_paper, include=FALSE}
@@
 
-f_data  <- tourr::rescale(flea[,1:6])
+f_data  <- tourr::rescale(flea[,2:6])
 f_path  <- save_history(f_data, guided_tour(holes()))
@@
 
-```{r step2, echo=F, out.width='40%', fig.cap = "(ref:step2-cap)"}
+```{r step2, echo=F, out.width='60%', fig.cap = "(ref:step2-cap)"}
 view_manip_space(basis = f_bas, manip_var = f_mvar, labels = colnames(f_dat)) +
@@
 view_manip_space(basis = f_basis, 
-                 manip_var = 5, 
+                 manip_var = 4, 
                  labels = colnames(f_data))
@@
 The contributions increase from its initial contribution to a full contribution to the projection before decreasing to zero and then returning to its initial value.
-The change in the projected data shows that `aede2` is important for distinguishing between the purple species.
+The change in the projected data shows that `aede2` is important for distinguishing the purple species.
 An animated version can be viewed at https://nspyrison.netlify.com/thesis/flea_manualtour_mvar5/.
@@
 
-### Data in projection-space {#sec:display}
+## Data in projection-space {#sec:display}
 
@@
 
-(ref:JetClusterGood-cap) `Jet` cluster, a radial manual tour of PC3.
+<!-- discussion of findings and which is more insightful -->
+When manipulating PC4, there is a clear difference in the parameter space spanned by the experiment types `ATLAS7new` and `ATLAS7old`.
+Yet, when PC3 is manipulated there is no indication that the different experiments probe different parameter space.
+Performing a radial manual tour on PC4 is more insightful than for PC3.
+
+Radial manual tours manipulating each of the principal components in the `jet` cluster can be viewed by following the links: [PC1](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc1/), [PC2](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc2/), [PC3](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/), and [PC4](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc4/).
+
+(ref:JetClusterGood-cap) A radial manual tour of PC4 within the `jet` cluster.
 Colored by experiment type: `ATLAS7new` in green and `ATLAS7old` in orange.
-When PC3 fully contributes to the projection `ATLAS7new` (green) occupies unique space and several outliers are identifiable.
-Zeroing the contribution from PC3 to the projection hides the outliers and all observations with `ATLAS7new` are contained within the area spanned by `ATLAS7old` (orange).
-A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/.
+When PC4 fully/negligibly contributes to the projection `ATLAS7new` (green) spans the same space as the orange points.
+During the intermediate frames, the `ATLAS7new` is compressed in the direction radial to PC4.
+A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc4/.
 
@@
 
-(ref:JetClusterBad-cap) `Jet` cluster, a radial manual tour of PC4.
-Colored by experiment type: ATLAS7new in green and ATLAS7old in orange.
-This manual tour contains less interesting information ATLAS7new (green) has points that are right and left of ATLAS7old, while most points occupy the same projection space, independent of the contribution of PC4.
+(ref:JetClusterBad-cap) A radial manual tour of PC4 within the `jet` cluster.
+Colored by experiment type: `ATLAS7new` in green and `ATLAS7old` in orange.
+Data from `ATLAS7new` (green) spans mostly the same space as `ALTLAS7old` (orange) with no evident difference in cluster structure across varying contributions of PC3.
 A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/.
@@
 
-## TODO: edit DIS cluster, Grammarly and word checks.
-<!-- discussion of findings and which is more insightful -->
-Manipulating PC3, where varying the angle of rotation brings interesting features into and out of the center mass of the data, is more interesting than the manipulation of PC4, where the features are mostly independent of the contribution of PC4.
-
-Radial manual tours manipulating each of the principal components in the `jet` cluster can be viewed by following the links: [PC1](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc1/), [PC2](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc2/), [PC3](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/), and [PC4](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc4/).
-
 ### DIS cluster
 
-We perform a manual tour on this data, manipulating PC6 as depicted in figure \@ref(fig:DISclusterGood).
-Looking at several frames we see that `DIS HERA` (green) data lie mostly on a plane.
-When PC6 has full contributions, `dimuon SIDIS` (purple) is almost orthogonal to the `DIS HERA` (green).
-Yet the contribution of PC6 has zeroed the dimuon SIDIS data occupy the same space as the `DIS HERA` data.
-A dynamic version of this manual tour can be found at:
-https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/.
-The page may some time to load, as the animation is several megabytes.
+<!-- introduce DIS cluster -->
+A different space is used to explore the `DIS` cluster; specifically the first six principal components, which explain 48% of the variation contained within the aggregated data[@cook_dynamical_2018].
+Radial manual tours are performed on PC6 and PC2 in figure \@ref(fig:DISclusterGood) and figure \@ref(fig:DISclusterBad) respectively.
 
-(ref:DISclusterGood-cap) `DIS` cluster, a radial manual tour of PC6.
-colored by experiment type: `DIS HERA1+2` in green, `dimuon SIDIS` in purple, and `charm SIDIS` in orange.
-When the contribution PC 6 is large we see that dimuon SIDIS (purple) data are nearly orthogonal to `DIS HERA` (green) data.
-As the projection is rotated, we can also see that `DIS HERA` (green) practically lies on a plane in this 6D subspace.
-When the contribution of PC6 is near zero, `dimonSIDIS` (purple) occupies the same space as the `DIS HERA` data.
+<!-- comparison of the DIS cluster -->
+The selection of the manip variable is important as the manipulation spaces convey substantially different information.
+The manual tour of PC6 offers information about the dimensionality, shape, and orientations of the different experiment classes.
+whereas manipulating the contributions of PC2 only shows a subset of the dimensionality and shape information.
+Manipulating the contributions of PC6 turned out to be much more insightful than that of PC2.
+This result might seem counter-intuitive at first as PC2 should explain much more of the variation in the data.
+However, features and structure in the data regularly reside in finer detail that can be lost when looking only at static projections.
+
+`DIS` cluster manual tours manipulating each of the principal components can be viewed from the links: [PC1](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc1/), [PC2](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc2/), [PC3](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc3/), [PC4](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc4/), [PC5](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc5/), and [PC6](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/).
+
+(ref:DISclusterGood-cap) A radial manual tour manipulating the contribution of  PC6 within the `DIS` cluster.
+Points are colored by experiment type: `DIS HERA1+2` in green, `dimuon SIDIS` in purple, and `charm SIDIS` in orange.
+The cluster `DIS HERA1+2` (green) is distributed in a cross-shaped plane, `charm SIDIS` (orange) occupies the center space of this cross, with the plane projecting into the field of view when the contribution of PC6 is max.
+Less evident is the linear `dimuon SIDIS` (purple) observations approaching the line of view for intermediate values of PC6.
 A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/.
@@
 
-The selection of the correct manip variable is important as the manipulation spaces convey different information.
-For example, in figure \@ref(fig:DISclusterBad) we select PC2 as the manip variable finding it to be less insightful than PC6.
-
-(ref:DISclusterBad-cap) `DIS` cluster, a radial manual tour of PC2.
-Colored by experiment type: `DIS HERA1+2` in green, `dimuon SIDIS` in purple, and `charm SIDIS` in orange.
-The plane of `DIS HERA` (green) and nearly orthogonal `dimuon SIDIS` (purple) is present, however, manipulating PC2 does not give a head-on view of either, offers less information about these features than manipulating PC6.
+(ref:DISclusterBad-cap) A radial manual tour manipulating the contribution of PC2 within the `DIS` cluster.
+Points are colored by experiment type: `DIS HERA1+2` in green, `dimuon SIDIS` in purple, and `charm SIDIS` in orange.
+The plane of cross distributed `DIS HERA` data (green) and a nearly orthogonal jet of `dimuon SIDIS` (purple) is present.
+This jet does extend more in the plane of view when the contribution of PC2 is full, giving insight to its orientation.
+However, less information about the shape of `DIS HERA` (green) and `charm SIDIS` (orange) is available.
 A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/discluster_manualtour_pc2/.
@@
 
-DIS cluster manual tours manipulating each of the principal components can be viewed from the links: [PC1](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc1/), [PC2](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc2/), [PC3](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc3/), [PC4](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc4/), [PC5](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc5/), and [PC6](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/).
-
-<!-- ## Source code and usage {#sec:usage} -->
-
-<!-- Use the below code as a guide for installation and finding the vignette.
-The vignette offers a less technical discussion opting to focus on code usage and goes through a couple more use cases.
-If you prefer to follow along with the example in the algorithm then simplified code is also listed below.
--->
-
-<!-- ```{r INSTALLCODEEXAMPLE, eval=F, echo=T} -->
-<!-- # devtools::install_github("nspyrison/spinifex") # Development version -->
-<!-- install.package("spinifex") -->
-<!-- library("spinifex") -->
-
-<!-- # Also see vignette: -->
-<!-- vignette("spinifex") # vignette ‘spinifex’ not found -->
-
-<!-- ## manual tour of std flea from holes-index: -->
-<!-- f_dat  <- tourr::rescale(flea[,1:6]) -->
-<!-- f_cat  <- factor(flea$species) -->
-<!-- f_path <- save_history(f_dat, guided_tour(holes())) -->
-<!-- f_bas  <- matrix(f_path[,, max(dim(f_path)[3])], ncol=2) -->
-<!-- f_mvar <- 5 -->
-<!-- f_lab  <- colnames(f_dat) -->
-
-<!-- # View the basis -->
-<!-- view_basis(f_bas, data = f_dat, labels = f_lab) -->
-<!-- # View the manip space -->
-<!-- view_manip_space(basis = f_bas, manip_var = f_mvar, labels = f_lab) -->
-<!-- # Play animation as HTML5 from plotly -->
-<!-- play_manual_tour(data = f_dat, basis = f_bas, manip_var = f_mvar,  -->
-<!--                  col = f_cat, angle = f_angle) -->
-<!-- ``` -->
+## TODO:
+### Summary
 
@@
 The source code for the **spinifex** package can be found at [github.com/nspyrison/spinifex/](https://github.com/nspyrison/spinifex/).
+
+## Bibliography

### 2019-04-09 paper/spinifex-paper.tex
References to RJ format, touched up equations.

@@
 % !TeX root = RJwrapper.tex
-\title{spinifex: manual control of dynamic linear projections of high-dimensional data}
+\title{spinifex: manual control of dynamic linear projections of
+high-dimensional data}
 \author{by Nicholas Spyrison, Dianne Cook}
@@
 \abstract{%
-The set of dynamic linear projections of multivariate data collectively known as \textbf{tours} provides an important tool for extending the dimensionality of visuals.
-The R package \textbf{tourr} offers a variety of path generators and geometric displays for conducting tours.
-This paper discusses an extension package, \textbf{spinifex}, that adds support for the path generation of manual tours and extends the display of tours to use with the animation packages, \textbf{plotly} and \textbf{gganimate}.
-Manual tours are used to explore the sensitivity of structure as the contributions of a manipulation variable are changed.
-A recent paper \{Wang et al.
-(2018)\} visualizes the sensitivity of the hadronic experiments to nucleon structure.
-Sensitivity was characterized in non-linear 3D embeddings of the first 10 principal components.
-This research applies manual tours to this data showing that manual tours resolve information about cluster structure and identify outliers, information that is orthogonal to the original static viewing plane.
+The set of dynamic linear projections of multivariate data collectively
+known as \textbf{tours} provides an important tool for extending the
+dimensionality of visuals. The R package \textbf{tourr} offers a variety
+of path generators and geometric displays for conducting tours. This
+paper discusses an extension package, \textbf{spinifex}, that adds
+support for the path generation of manual tours and extends the display
+of tours to use with the animation packages, \textbf{plotly} and
+\textbf{gganimate}. Manual tours are used to explore the sensitivity of
+structure as the contributions of a manipulation variable are changed. A
+recent paper (Wang et al.
+\protect\hyperlink{ref-wang_mapping_2018}{2018}) visualizes the
+sensitivity of the hadronic experiments to nucleon structure.
+Sensitivity was characterized in non-linear 3D embeddings of the first
+10 principal components. This research applies manual tours to this data
+showing that manual tours resolve information about cluster structure
+and identify outliers, information that is orthogonal to the original
+static viewing plane.
 }
-
-% Any extra LaTeX you need in the preamble
 
@@
 
-A tour is a multivariate data analysis technique in which a sequence of linear (orthogonal) projections are viewed as an animation while the orientation of the projection basis is rotated across time.
-Each frame of the sequence corresponds to a small change in the projection for a smooth transition that perseveres continuity.
-This is a crucial tool extending the dimensionality of visualization which is a large boon as datasets and models exist in increasing cardinality.
-The package \textbf{tourr} (Wickham et al.
-2011) builds a platform for generating tour paths and applying to a wide number of display geometrics in \textbf{base} graphics.
+A tour is a multivariate data analysis technique in which a sequence of
+linear (orthogonal) projections are viewed as an animation while the
+orientation of the projection basis is rotated across time. Each frame
+of the sequence corresponds to a small change in the projection for a
+smooth transition that perseveres continuity. This is a crucial tool
+extending the dimensionality of visualization which is a large boon as
+datasets and models exist in increasing cardinality. The package
+\textbf{tourr} (Wickham et al.
+\protect\hyperlink{ref-wickham_tourr_2011}{2011}) builds a platform for
+generating tour paths and applying to a wide number of display
+geometrics in \textbf{base} graphics.
 
-While there are numerous methods that generate tour paths, this research focuses on the manual tour.
-The manual tour was described in Cook and Buja (1997) and allows a user to control the projection coefficients of a selected variable in a 2D projection.
-The manipulation of these coefficients allows the analyst to explore how sensitive the projections structure is to these changes.
-This makes manual tours particularly useful once a feature of interest has been identified, for example, with the use of a guided tour (Hurley and Buja 1990).
-Rather than exploring the large phase space of the data, a guided tour selects a very specific path via projection pursuit(Diaconis and Freedman 1984), the optimization of an index function on the projection via a hill climbing algorithm.
-This allows guided tours to rapidly identify interesting projection features given the relatively large parameter-space.
-After a projection of interest is identified an analyst can then use the ``finer brush'' of the manual tour, by controlling the contributions of individual variables.
+While there are numerous methods that generate tour paths, this research
+focuses on the manual tour. The manual tour was described in Cook and
+Buja (\protect\hyperlink{ref-cook_manual_1997}{1997}) and allows a user
+to control the projection coefficients of a selected variable in a 2D
+projection. The manipulation of these coefficients allows the analyst to
+explore how sensitive the projections structure is to these changes.
+This makes manual tours particularly useful once a feature of interest
+has been identified, for example, with the use of a guided tour (Hurley
+and Buja \protect\hyperlink{ref-hurley_analyzing_1990}{1990}). Rather
+than exploring the large phase space of the data, a guided tour selects
+a very specific path via projection pursuit(Diaconis and Freedman
+\protect\hyperlink{ref-diaconis_asymptotics_1984}{1984}), the
+optimization of an index function on the projection via a hill climbing
+algorithm. This allows guided tours to rapidly identify interesting
+projection features given the relatively large parameter-space. After a
+projection of interest is identified an analyst can then use the ``finer
+brush'' of the manual tour, by controlling the contributions of
+individual variables.
 
-Spinifex utilizes two new animation packages, \textbf{plotly} (Sievert 2018) and \textbf{gganimate} (Pedersen and Robinson 2019), to display tours, manual or other saved tours.
-From a given projection, the user can choose which variable to control, and the animation sequence is generated to remove the variable from the projection, and then extend its contribution to be the sole variable in one direction.
-This allows the viewer to assess the change in structure induced in the projection by the variable's contribution.
+Spinifex utilizes two new animation packages, \textbf{plotly} (Sievert
+\protect\hyperlink{ref-sievert_plotly_2018}{2018}) and
+\textbf{gganimate} (Pedersen and Robinson
+\protect\hyperlink{ref-pedersen_gganimate:_2019}{2019}), to display
+tours, manual or other saved tours. From a given projection, the user
+can choose which variable to control, and the animation sequence is
+generated to remove the variable from the projection, and then extend
+its contribution to be the sole variable in one direction. This allows
+the viewer to assess the change in structure induced in the projection
+by the variable's contribution.
 
-The paper is organized as follows.
-Section \ref{sec:algorithm} explains the algorithm using a toy dataset.
-Section \ref{sec:display} discussed the display of the animation after the path has been generated.
-Section \ref{sec:application} illustrates how this can be used for sensitivity analysis applied to contemporary high energy physics.
-The last section, \ref{sec:discussion} summarizes the work and discusses future research.
+The paper is organized as follows. Section \ref{sec:algorithm} explains
+the algorithm using a toy dataset. Section \ref{sec:display} discussed
+the display of the animation after the path has been generated. Section
+\ref{sec:application} illustrates how this can be used for sensitivity
+analysis applied to contemporary high energy physics. The last section,
+\ref{sec:discussion} summarizes the work and discusses future research.
 
@@
 
-The section below describes the algorithm for performing a 2D radial manual tour:
+The section below describes the algorithm for performing a 2D radial
+manual tour:
 
@@
 \item
-  Provided with a 2D projection, choose a variable to explore. This is called the ``manip'' variable.
+  Provided with a 2D projection, choose a variable to explore. This is
+  called the ``manip'' variable.
 \item
-  Create a 3D manipulation space, where the manip variable has the full contribution.
+  Create a 3D manipulation space, where the manip variable has the full
+  contribution.
 \item
-  Generate a rotation sequence which increases the norm of the coefficient to 1 and zeros it.
+  Generate a rotation sequence which increases the norm of the
+  coefficient to 1 and zeros it.
 \end{enumerate}
 
-The steps are described in more detail below. The R functions used below mentioned briefly, but more complete code example can be found in section \ref{sec:usage}
+\hypertarget{fix-here}{%
+\subsection{FIX here}\label{fix-here}}
+
+The steps are described in more detail below. The R functions used below
+mentioned briefly, but more complete code examples are given with
 
@@
 
-This section describes the notation used in the algorithm for a 2D radial manual tour.
+This section describes the notation used in the algorithm for a 2D
+radial manual tour.
 
@@
 
-The algorithm primarily operates on projection bases and utilizes the data only when making a display.
-The projection space can be viewed at any point in the process by pre-multiplying the data and plotting the first two variables.
+The algorithm primarily operates on projection bases and utilizes the
+data only when making a display. The projection space can be viewed at
+any point in the process by pre-multiplying the data and plotting the
+first two variables.
 
@@
 
-The flea data, originally from Lubischew (1962), made available in \textbf{tourr} is used to illustrate the algorithm.
-The data contains 74 observations across 6 variables, physical measurements of the flea beetles.
-Each observation belonging to one of three species.
+The flea data, originally from Lubischew
+(\protect\hyperlink{ref-lubischew_use_1962}{1962}), made available in
+\textbf{tourr} is used to illustrate the algorithm. The data contains 74
+observations across 6 variables, physical measurements of the flea
+beetles. Each observation belonging to one of three species.
 
-The data is defined.
-A basis set (ideally orienting to an interesting feature) must be provided as an initial orientation.
-One way to identify a projection containing interesting features is to apply a guided tour (Cook, Swayne, and Buja 2007).
-In a guided tour the projection sequence is selected by optimizing an index function via hill-climbing on the projection space.
-In this case, the holes index is selected and applied to standardized flea data.
-The holes index is maximized when the projected observations are furthest from the center.
-Figure @ref\{fig:step0\} shows a locally optimized projection of the data.
-The left panel displays the reference axes of the projection basis, a visual indication of the magnitude and direction each variable contributed to the projection.
-The right panel shows the data as projected through the basis set described by the reference axes (left).
-Data points are colored and given shape according to the species (while the guided tour was unsupervised with this information).
-
-(ref:step0-cap) Basis reference axes (left) and projected data (right) of standardized flea data.
-Data points color and shape are mapped to beetle species.
-Basis identified by a holes-index guided tour.
-The variables \texttt{aede2} and \texttt{tars1} contribute mostly orthogonal to the other variables.
+The data is defined. A basis set (ideally orienting to an interesting
+feature) must be provided as an initial orientation. One way to identify
+a projection containing interesting features is to apply a guided tour
+(Cook, Swayne, and Buja
+\protect\hyperlink{ref-cook_interactive_2007}{2007}). In a guided tour
+the projection sequence is selected by optimizing an index function via
+hill-climbing on the projection space. In this case, the holes index is
+selected and applied to standardized flea data. The holes index is
+maximized when the projected observations are furthest from the center.
+Figure \ref{fig:step0} shows a locally optimized projection of the data.
+The left panel displays the reference axes of the projection basis, a
+visual indication of the magnitude and direction each variable
+contributed to the projection. The right panel shows the data as
+projected through the basis set described by the reference axes (left).
+Data points are colored and given shape according to the species (while
+the guided tour was unsupervised with this information).
 
@@
 
-\caption[(ref:step0-cap)]{(ref:step0-cap)}(\#fig:step0)
+\caption[Basis reference axes (left) and projected data (right) of standardized flea data]{Basis reference axes (left) and projected data (right) of standardized flea data.
+Data points color and shape are mapped to beetle species.
+Basis identified by a holes-index guided tour.
+The variables aede2 and tars1 contribute mostly orthogonal to the other variables.}\label{fig:step0}
 \end{figure}
@@
 
-Call \texttt{view\_basis()} on a basis to produce a \textbf{ggplot2} graphic similar to @ref(fig:step0).
-Projection space is always available for display via the matrix multiplication \(\textbf{X}_{[n,~p]} ~*~ \textbf{B}_{[p,~d]} ~=~ \textbf{P}_{[n,~d]}\).
+Call \texttt{view\_basis()} on a basis to produce a \textbf{ggplot2}
+graphic similar to \textbackslash ref\{fig:step0). Projection space is
+always available for display via the matrix multiplication
+\(\textbf{X}_{[n,~p]} ~*~ \textbf{B}_{[p,~d]} ~=~ \textbf{P}_{[n,~d]}\).
 
@@
 
-f_data  <- tourr::rescale(flea[,2:6])
+f_data  <- tourr::rescale(flea[, 1:6])
 f_path  <- save_history(f_data, guided_tour(holes()))
@@
 \hypertarget{step-1-choose-variable-of-interest}{%
-\subsubsection{Step 1) Choose variable of interest}\label{step-1-choose-variable-of-interest}}
+\subsubsection{Step 1) Choose variable of
+interest}\label{step-1-choose-variable-of-interest}}
 
-In figure @ref(fig:step0) the contributions of the variables \texttt{tars1} and \texttt{aede2} are mostly orthogonal to the contributions of the other four variables.
-These two variables explain the variation of the data distinguishing the purple group from the rest of the sample.
-Select \texttt{aede2} as the manip var, the variable to be manipulated, as it has a larger contribution to the projection.
-The question that will be explored is how important the variable \texttt{aede2} is to the separation of the clusters.
+In figure \ref{fig:step0} the contributions of the variables
+\texttt{tars1} and \texttt{aede2} are mostly orthogonal to the
+contributions of the other four variables. These two variables explain
+the variation of the data distinguishing the purple group from the rest
+of the sample. Select \texttt{aede2} as the manip var, the variable to
+be manipulated, as it has a larger contribution to the projection. The
+question that will be explored is how important the variable
+\texttt{aede2} is to the separation of the clusters.
 
 \hypertarget{step-2-create-the-manip-space}{%
-\subsubsection{Step 2) Create the manip space}\label{step-2-create-the-manip-space}}
+\subsubsection{Step 2) Create the manip
+space}\label{step-2-create-the-manip-space}}
 
-Initialize a zero vector \textbf{e} of length \emph{p}.
-Set the fifth element to one, as \texttt{aede2} is the fifth variable in the data, giving the manip var a full contribution in this dimension.
-Use the Gram-Schmidt process to orthonormalize the zero vector onto the basis yielding the 3D manipulation space, \textbf{M}.
+Initialize a zero vector \textbf{e} of length \emph{p}. Set the fifth
+element to one, as \texttt{aede2} is the fifth variable in the data,
+giving the manip var a full contribution in this dimension. Use the
+Gram-Schmidt process to orthonormalize the zero vector onto the basis
+yielding the 3D manipulation space, \textbf{M}.
 
@@
 
-Adding this new dimension to our projection space allows for the coefficients of the manip var to be changed via rotation about the origin.
-For example, the ability to lift a piece of paper, rather than being constrained to a 2D plane.
-Orthonormalizing rescales the new vector while the projection down to 2D remains the original basis.
-Place the plane horizontally, with the new dimension extending vertically, with axes projecting back onto the reference axes.
-Figure @ref(fig:step2) illustrates this 3D manipulation space with the manip var highlighted (height of the other variables are not depicted.)
-
-(ref:step2-cap) Manipulation space for controlling the contribution of \texttt{aede2} of standardized flea data.
-Basis selected by a holes-index optimized guided tour lies on the horizontal projection plane, shown in blue.
-The manip var axis, in red, extends into the vertical manipulation space, allowing the coefficients of the manip var to be changed by rotation around the origin.
+Adding this new dimension to our projection space allows for the
+coefficients of the manip var to be changed via rotation about the
+origin. For example, the ability to lift a piece of paper, rather than
+being constrained to a 2D plane. Orthonormalizing rescales the new
+vector while the projection down to 2D remains the original basis. Place
+the plane horizontally, with the new dimension extending vertically,
+with axes projecting back onto the reference axes. Figure
+\ref{fig:step2} illustrates this 3D manipulation space with the manip
+var highlighted (height of the other variables are not depicted.)
 
@@
 
-\caption[(ref:step2-cap)]{(ref:step2-cap)}(\#fig:step2)
+\caption[Manipulation space for controlling the contribution of aede2 of standardized flea data]{Manipulation space for controlling the contribution of aede2 of standardized flea data.
+Basis selected by a holes-index optimized guided tour lies on the horizontal projection plane, shown in blue.
+The manip var axis, in red, extends into the vertical manipulation space, allowing the coefficients of the manip var to be changed by rotation around the origin.}\label{fig:step2}
 \end{figure}
@@
 
-The representation in figure @ref(fig:step2) can be produced by calling the function \texttt{view\_manip\_space()}.
+The representation in figure \ref{fig:step2} can be produced by calling
+the function \texttt{view\_manip\_space()}.
 
@@
 view_manip_space(basis = f_basis, 
-                 manip_var = 4, 
+                 manip_var = 5, 
                  labels = colnames(f_data))
@@
 \hypertarget{step-3-generate-rotation}{%
-\subsubsection{Step 3) Generate rotation}\label{step-3-generate-rotation}}
+\subsubsection{Step 3) Generate
+rotation}\label{step-3-generate-rotation}}
 
-Imagine holding the manip var, the red axis, one end fixed to the origin.
-As it is controlled the manipulation space rotates about the origin, the projection onto the horizontal projection plane correspondingly moves.
-This is what happens in a manual tour.
-Generating a sequence of values for the horizontal and vertical, angles produces a path for the rotation of the manipulation space.
-This defines the (orthonormally-constrained) rotation on the coefficients of the variables.
+Imagine holding the manip var, the red axis, one end fixed to the
+origin. As it is controlled the manipulation space rotates about the
+origin, the projection onto the horizontal projection plane
+correspondingly moves. This is what happens in a manual tour. Generating
+a sequence of values for the horizontal and vertical, angles produces a
+path for the rotation of the manipulation space. This defines the
+(orthonormally-constrained) rotation on the coefficients of the
+variables.
 
-For a radial tour fix the (horizontal) angle within the projection plane, \(\theta\), and define a sequence for the (vertical) angle coming out of the projection plane, \(\phi\), bringing the initial \texttt{XY} contributions of the manip var to a maximum and then to zero before returning to the initial position.
-Dynamic capture of user manipulation is typically performed directly on the projection plane (without depiction of the manipulation space.)
+For a radial tour fix the (horizontal) angle within the projection
+plane, \(\theta\), and define a sequence for the (vertical) angle coming
+out of the projection plane, \(\phi\), bringing the initial \texttt{XY}
+contributions of the manip var to a maximum and then to zero before
+returning to the initial position. Dynamic capture of user manipulation
+is typically performed directly on the projection plane (without
+depiction of the manipulation space.)
 
@@
 
-Post-multiply the manipulation space by the pre-defined rotation matrix producing \textbf{RM}, the rotated manip space.
+Post-multiply the manipulation space by the pre-defined rotation matrix
+producing \textbf{RM}, the rotated manip space.
 
@@
 
-then
-\begin{align*}
+then \begin{align*}
   \textbf{RM}_{[p,~3,~i]}
   &= \textbf{M}_{[p,~3]} ~*~ \textbf{R}_{[3,~3,~i]} \\
-  &= \begin{bmatrix}
-    M_{1,~1} & M_{1,~2} & M_{1,~3} \\
-    M_{2,~1} & M_{2,~2} & M_{2,~3} \\
-    \vdots   & \vdots   \\
-    M_{p,~1} & M_{p,~2} & M_{p,~3}
-  \end{bmatrix}_{[p,~3]}
+  &= \textbf{M}_{[p,~3]}
     ~*~
@@
 
-A note on application: \(\phi\) is the angle relative to the initial value of \(\phi\), we find the transformation \(\phi_i\) - \(\phi_1\) useful to think about \(\phi\) relative to the basis plane.
-Additionally, the value of \(\phi\) may be out of phase by a factor of pi.
-If the manip variable doesn't move as expected these are the first places to check.
+A note on application: \(\phi\) is the angle relative to the initial
+value of \(\phi\), we find the transformation \(\phi_i\) - \(\phi_1\)
+useful to think about \(\phi\) relative to the basis plane.
+Additionally, the value of \(\phi\) may be out of phase by a factor of
+pi. If the manip variable doesn't move as expected these are the first
+places to check.
 
-Figure @ref(fig:step3) illustrates a sequence with 15 projected bases, showing the reference axes on top the corresponding projected data points on the below.
-As a result, we can see that changes in the manip var controlled the distance between the purple cluster and the remaining sample, \texttt{aede2} is crucial in distinguishing this species.
-Tours are typically viewed as an animation such a dynamic version of this tour can be viewed online at \url{https://nspyrison.netlify.com/thesis/flea_manualtour_mvar5/}.
-The page may take a moment to load.
-
-(ref:step3-cap) Radial manual tour manipulating \texttt{aede2} of standardized flea data.
-The contributions increase from its initial contribution to a full contribution to the projection before decreasing to zero and then returning to its initial value.
-The change in the projected data shows that \texttt{aede2} is important for distinguishing the purple species.
-An animated version can be viewed at \url{https://nspyrison.netlify.com/thesis/flea_manualtour_mvar5/}.
+Figure \ref{fig:step3} illustrates a sequence with 15 projected bases,
+showing the reference axes on top the corresponding projected data
+points on the below. As a result, we can see that changes in the manip
+var controlled the distance between the purple cluster and the remaining
+sample, \texttt{aede2} is crucial in distinguishing this species. Tours
+are typically viewed as an animation such a dynamic version of this tour
+can be viewed online at
+\url{https://nspyrison.netlify.com/thesis/flea_manualtour_mvar5/}. The
+page may take a moment to load.
 
@@
 
-\caption[(ref:step3-cap)]{(ref:step3-cap)}(\#fig:step3)
+\caption[Radial manual tour manipulating aede2 of standardized flea data]{Radial manual tour manipulating aede2 of standardized flea data.
+The contributions increase from its initial contribution to a full contribution to the projection before decreasing to zero and then returning to its initial value.
+The change in the projected data shows that aede2 is important for distinguishing the purple species.
+An animated version can be viewed at  https://nspyrison.netlify.com/thesis/flea\_manualtour\_mvar5/.}\label{fig:step3}
 \end{figure}
@@
 
-Animations can be produced using the function \texttt{play\_manual\_tour()}.
-This function defaults to an HTML5 widget produced from \textbf{plotly}.
-The \texttt{render\_type} argument can be changed to \texttt{render\_gganimate} for exporting to .gif or .mp4 files.
+Animations can be produced using the function
+\texttt{play\_manual\_tour()}. This function defaults to an HTML5 widget
+produced from \textbf{plotly}. The \texttt{render\_type} argument can be
+changed to \texttt{render\_gganimate} for exporting to .gif or .mp4
+files.
 
@@
 
-In an appeal to performance, the above operations are performed on the bases without the use of the larger datasets.
-After the bases are brought into the projection-space, however, it is helpful to observe them with data in the same space.
-Pre-multiply the data by basis frame bringing the data into the projection space.
+In an appeal to performance, the above operations are performed on the
+bases without the use of the larger datasets. After the bases are
+brought into the projection-space, however, it is helpful to observe
+them with data in the same space. Pre-multiply the data by basis frame
+bringing the data into the projection space.
 
@@
 
-For a 2D scatterplot, plot the first two variables from each frame statically as in the previous figure, or in sequence, producing an animated scatterplot.
-The remaining variable is sometimes linked to a data point aesthetic (such as size or color) to produce depth cues used in conjunction with the \(XY\) scatterplot.
+For a 2D scatterplot, plot the first two variables from each frame
+statically as in the previous figure, or in sequence, producing an
+animated scatterplot. The remaining variable is sometimes linked to a
+data point aesthetic (such as size or color) to produce depth cues used
+in conjunction with the \(XY\) scatterplot.
 
@@
 
-The \textbf{tourr} package utilizes \textbf{base} graphics for the display of tours.
-\textbf{spinifex} allows tours to be used in rendered in \textbf{plotly} as an HTML5 object or \textbf{gganimate} as .gif or .mp4 objects.
-Both of which build off \textbf{ggplot2} objects in internal functions.
-Sharing of animations is not trivial especially in print and static file formats such as .pdf.
-Even with the use of computers and dynamic file formats capturing the correct resolution, aspect, and display is challenging and many formats quickly bloat file sizes.
-Keep in mind hosting options and exporting functions from \textbf{plotly}, \textbf{gganimate} and \textbf{tourr}.
+The \textbf{tourr} package utilizes \textbf{base} graphics for the
+display of tours. \textbf{spinifex} allows tours to be used in rendered
+in \textbf{plotly} as an HTML5 object or \textbf{gganimate} as .gif or
+.mp4 objects. Both of which build off \textbf{ggplot2} objects in
+internal functions. Sharing of animations is not trivial especially in
+print and static file formats such as .pdf. Even with the use of
+computers and dynamic file formats capturing the correct resolution,
+aspect, and display is challenging and many formats quickly bloat file
+sizes. Keep in mind hosting options and exporting functions from
+\textbf{plotly}, \textbf{gganimate} and \textbf{tourr}.
 
@@
 Storing each data point for every frame of the animation is redundant.
-Just as operations are performed on the bases, so too should tour paths be stored as bases and a single instance of the data.
-Consider a radial manual tour, we can store the salient features in 3 bases, where \(\phi\) is at its starting, minimum, and maximum values.
-The frames in between can be interpolated by supplying angular speed.
-With the use of the \texttt{tourr::save\_history()} function, the target bases can be saved.
-From there geodesic interpolation can be used to populate the intermittent frames.
-This type of interpolation should not be used on manual tours, which have already been initialized into a 3D manipulation space where direct linear interpolation is appropriate.
+Just as operations are performed on the bases, so too should tour paths
+be stored as bases and a single instance of the data. Consider a radial
+manual tour, we can store the salient features in 3 bases, where
+\(\phi\) is at its starting, minimum, and maximum values. The frames in
+between can be interpolated by supplying angular speed. With the use of
+the \texttt{tourr::save\_history()} function, the target bases can be
+saved. From there geodesic interpolation can be used to populate the
+intermittent frames. This type of interpolation should not be used on
+manual tours, which have already been initialized into a 3D manipulation
+space where direct linear interpolation is appropriate.
 
@@
 In a recent paper, Wang et al.
-(2018), the authors aggregate and visualize the sensitivity of hadronic experiments to nucleon structure.
-The authors introduce a new tool, PDFSense, to aid in the visualization of Parton distribution functions (PDF).
-The parameter-space of these experiments lies in 56 dimensions, \(\delta \in \mathbb{R}^{56}\), and are visualized as 3D subspaces of the 10 first principal components in linear (PCA) and non-linear (t-SNE) embeddings.
+(\protect\hyperlink{ref-wang_mapping_2018}{2018}), the authors aggregate
+and visualize the sensitivity of hadronic experiments to nucleon
+structure. The authors introduce a new tool, PDFSense, to aid in the
+visualization of Parton distribution functions (PDF). The
+parameter-space of these experiments lies in 56 dimensions,
+\(\delta \in \mathbb{R}^{56}\), and are visualized as 3D subspaces of
+the 10 first principal components in linear (PCA) and non-linear (t-SNE)
+embeddings.
 
-Using the same data, another study, Cook, Laa, and Valencia (2018), applies grand tours (Asimov 1985) to the same subspaces.
-Grand tours are dynamic subspace projections of high dimensional where frames and selected at random and linked with geodesically interpolation of the intermediate frames.
-Grand tours are able to better resolve the distribution shape of clusters, intra-cluster detail, and better outlier detection than the use of PDFSense \& TFEP (TensorFlow embedded projections).
-Before applying manual tours let's discuss the structure of the data.
+Using the same data, another study, Cook, Laa, and Valencia
+(\protect\hyperlink{ref-cook_dynamical_2018}{2018}), applies grand tours
+(Asimov \protect\hyperlink{ref-asimov_grand_1985}{1985}) to the same
+subspaces. Grand tours are dynamic subspace projections of high
+dimensional where frames and selected at random and linked with
+geodesically interpolation of the intermediate frames. Grand tours are
+able to better resolve the distribution shape of clusters, intra-cluster
+detail, and better outlier detection than the use of PDFSense \& TFEP
+(TensorFlow embedded projections). Before applying manual tours let's
+discuss the structure of the data.
 
-The data has a hierarchical structure with top-level clusters; \texttt{DIS}, \texttt{VBP}, and \texttt{jet}.
-Each cluster is a particular class of experiments, each with many experimental datasets which, in turn, have many observations.
-In the consideration of data density, we conduct manual tours on subsets of the \texttt{DIS} and \texttt{jet} clusters.
-This explores the sensitivity of the structure to each of the variables in turn and we present the subjectively best and worst variable to manipulate for identifying dimensionality of the clusters and describing the span of the clusters.
+The data has a hierarchical structure with top-level clusters;
+\texttt{DIS}, \texttt{VBP}, and \texttt{jet}. Each cluster is a
+particular class of experiments, each with many experimental datasets
+which, in turn, have many observations. In the consideration of data
+density, we conduct manual tours on subsets of the \texttt{DIS} and
+\texttt{jet} clusters. This explores the sensitivity of the structure to
+each of the variables in turn and we present the subjectively best and
+worst variable to manipulate for identifying dimensionality of the
+clusters and describing the span of the clusters.
 
@@
 
-The \texttt{jet} cluster resides in a smaller dimensionality than the full set of experiments with four principal components explaining 95\% of the variation in the cluster (Cook, Laa, and Valencia 2018).
-The data within this 4D embedding is subset down to \texttt{ATLAS7old} and \texttt{ATLAS7new} to focus in on two groups with a reasonable number of observations that occupy different parts of the subspace.
-Radial manual tours controlling contributions from PC4 and PC3 are shown in figure @ref(fig:JetClusterGood) and figure @ref(fig:JetClusterBad) respectively.
-These components are selected to contrast the difference of information conveyed by touring different variables.
-Links to dynamic HTML5 animations controlling each of the four variables are also provided.
+The \texttt{jet} cluster resides in a smaller dimensionality than the
+full set of experiments with four principal components explaining 95\%
+of the variation in the cluster (Cook, Laa, and Valencia
+\protect\hyperlink{ref-cook_dynamical_2018}{2018}). The data within this
+4D embedding is subset down to \texttt{ATLAS7old} and \texttt{ATLAS7new}
+to focus in on two groups with a reasonable number of observations that
+occupy different parts of the subspace. Radial manual tours controlling
+contributions from PC4 and PC3 are shown in figure
+\ref{fig:JetClusterGood} and figure \ref{fig:JetClusterBad}
+respectively. These components are selected to contrast the difference
+of information conveyed by touring different variables. Links to dynamic
+HTML5 animations controlling each of the four variables are also
+provided.
 
-When manipulating PC4, there is clear difference in the parameter space spanned by the experiment types \texttt{ATLAS7new} and \texttt{ATLAS7old}.
-Yet, when PC3 is manipulated there is no indication that the diffent expiments probe different parameter space.
+When manipulating PC4, there is a clear difference in the parameter
+space spanned by the experiment types \texttt{ATLAS7new} and
+\texttt{ATLAS7old}. Yet, when PC3 is manipulated there is no indication
+that the different experiments probe different parameter space.
 Performing a radial manual tour on PC4 is more insightful than for PC3.
 
-Radial manual tours manipulating each of the principal components in the \texttt{jet} cluster can be viewed by following the links: \href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc1/}{PC1}, \href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc2/}{PC2}, \href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/}{PC3}, and \href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc4/}{PC4}.
-
-(ref:JetClusterGood-cap) A radial manual tour of PC4 within the \texttt{jet} cluster.
-Colored by experiment type: \texttt{ATLAS7new} in green and \texttt{ATLAS7old} in orange.
-When PC4 fully/negligably contributes to the projection \texttt{ATLAS7new} (green) spans the same space as the orange points.
-During the intermediate frames the \texttt{ATLAS7new} is compressed in the direction radial to PC4.
-A dynamic version can be viewed at \url{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc4/}.
+Radial manual tours manipulating each of the principal components in the
+\texttt{jet} cluster can be viewed by following the links:
+\href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc1/}{PC1},
+\href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc2/}{PC2},
+\href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/}{PC3},
+and
+\href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc4/}{PC4}.
 
@@
 
-\caption[(ref:JetClusterGood-cap)]{(ref:JetClusterGood-cap)}(\#fig:JetClusterGood)
+\caption[A radial manual tour of PC4 within the jet cluster]{A radial manual tour of PC4 within the jet cluster.
+Colored by experiment type: ATLAS7new in green and ATLAS7old in orange.
+When PC4 fully/negligibly contributes to the projection ATLAS7new (green) spans the same space as the orange points.
+During the intermediate frames, the ATLAS7new is compressed in the direction radial to PC4.
+A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/jetcluster\_manualtour\_pc4/.}\label{fig:JetClusterGood}
 \end{figure}
@@
 
-(ref:JetClusterBad-cap) A radial manual tour of PC4 within the \texttt{jet} cluster.
-Colored by experiment type: \texttt{ATLAS7new} in green and \texttt{ATLAS7old} in orange.
-Data from \texttt{ATLAS7new} (green) spans the mostly the same space as \texttt{ALTLAS7old} (orange) with no evident differnce in cluster structure across varying contributions of PC3.
-A dynamic version can be viewed at \url{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/}.
+(ref:JetClusterBad-cap) A radial manual tour of PC4 within the
+\texttt{jet} cluster. Colored by experiment type: \texttt{ATLAS7new} in
+green and \texttt{ATLAS7old} in orange. Data from \texttt{ATLAS7new}
+(green) spans mostly the same space as \texttt{ALTLAS7old} (orange) with
+no evident difference in cluster structure across varying contributions
+of PC3. A dynamic version can be viewed at
+\url{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/}.
 
@@
 
-\caption[(ref:JetClusterBad-cap)]{(ref:JetClusterBad-cap)}(\#fig:JetClusterBad)
+\caption[(ref:JetClusterBad-cap)]{(ref:JetClusterBad-cap)}\label{fig:JetClusterBad}
 \end{figure}
@@
 
-A different space is used to explore the \texttt{DIS} cluster; specfically the first six pricipal components, which explain 48\% of the variation contained within the aggregated data(Cook, Laa, and Valencia 2018).
-Radial manual tours are performed on PC6 and PC2 in figure @ref(fig:DISclusterGood) and figure @ref(fig:DISclusterBad) respectively.
+A different space is used to explore the \texttt{DIS} cluster;
+specifically the first six principal components, which explain 48\% of
+the variation contained within the aggregated data(Cook, Laa, and
+Valencia \protect\hyperlink{ref-cook_dynamical_2018}{2018}). Radial
+manual tours are performed on PC6 and PC2 in figure
+\ref{fig:DISclusterGood} and figure \ref{fig:DISclusterBad}
+respectively.
 
-The selection of the manip variable is important as the manipulation spaces convey substantially different information.
-The manual tour of PC6 offers information about the dimensionality, shape, and orientations of the differnet experiment classes.
-whereas manipulating the contributions of PC2 only shows a subset of the dimensionality and shape information.
-Manipulating the contributions of PC6 turned out to be much more insightful than that of PC2.
-This result might seem counter-intuitive at first as PC2 should explain much more of the variation in the data.
-However, features and structure in the data regularly reside in finer detail that can be lost when looking only at static projections.
+The selection of the manip variable is important as the manipulation
+spaces convey substantially different information. The manual tour of
+PC6 offers information about the dimensionality, shape, and orientations
+of the different experiment classes. whereas manipulating the
+contributions of PC2 only shows a subset of the dimensionality and shape
+information. Manipulating the contributions of PC6 turned out to be much
+more insightful than that of PC2. This result might seem
+counter-intuitive at first as PC2 should explain much more of the
+variation in the data. However, features and structure in the data
+regularly reside in finer detail that can be lost when looking only at
+static projections.
 
-\texttt{DIS} cluster manual tours manipulating each of the principal components can be viewed from the links: \href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc1/}{PC1}, \href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc2/}{PC2}, \href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc3/}{PC3}, \href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc4/}{PC4}, \href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc5/}{PC5}, and \href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/}{PC6}.
-
-(ref:DISclusterGood-cap) A radial manual tour manipulating the contribution of PC6 within the \texttt{DIS} cluster.
-Points are colored by experiment type: \texttt{DIS\ HERA1+2} in green, \texttt{dimuon\ SIDIS} in purple, and \texttt{charm\ SIDIS} in orange.
-The cluster \texttt{DIS\ HERA1+2} (green) is distributed in a cross shaped plane, \texttt{charm\ SIDIS} (orange) occupies the the center space of this cross, with the plane projecting into the field of view when the contribution of PC6 is max.
-Less evident is the linear \texttt{dimuon\ SIDIS} (purple) observations approaching the line of view for intermediate values of PC6.
-A dynamic version can be viewed at \url{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/}.
+\texttt{DIS} cluster manual tours manipulating each of the principal
+components can be viewed from the links:
+\href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc1/}{PC1},
+\href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc2/}{PC2},
+\href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc3/}{PC3},
+\href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc4/}{PC4},
+\href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc5/}{PC5},
+and
+\href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/}{PC6}.
 
@@
 
-\caption[(ref:DISclusterGood-cap)]{(ref:DISclusterGood-cap)}(\#fig:DISclusterGood)
+\caption[A radial manual tour manipulating the contribution of  PC6 within the DIS cluster]{A radial manual tour manipulating the contribution of  PC6 within the DIS cluster.
+Points are colored by experiment type: DIS HERA1+2 in green, dimuon SIDIS in purple, and charm SIDIS in orange.
+The cluster DIS HERA1+2 (green) is distributed in a cross-shaped plane, charm SIDIS (orange) occupies the center space of this cross, with the plane projecting into the field of view when the contribution of PC6 is max.
+Less evident is the linear dimuon SIDIS (purple) observations approaching the line of view for intermediate values of PC6.
+A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/discluster\_manualtour\_pc6/.}\label{fig:DISclusterGood}
 \end{figure}
 \end{Schunk}
-
-(ref:DISclusterBad-cap) A radial manual tour manipulating the contribution of PC2 within the \texttt{DIS} cluster.
-Points are colored by experiment type: \texttt{DIS\ HERA1+2} in green, \texttt{dimuon\ SIDIS} in purple, and \texttt{charm\ SIDIS} in orange.
-The plane of cross distributed \texttt{DIS\ HERA} data (green) and nearly orthogonal jet of \texttt{dimuon\ SIDIS} (purple) is present.
-This jet does entend more in the plane of view when the contribution of PC2 is full.
-However, less information about the shape of \texttt{DIS\ HERA} (green) and \texttt{charm\ SIDIS} (orange) is available.
-A dynamic version can be viewed at \url{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc2/}.
 
@@
 
-\caption[(ref:DISclusterBad-cap)]{(ref:DISclusterBad-cap)}(\#fig:DISclusterBad)
+\caption[A radial manual tour manipulating the contribution of PC2 within the DIS cluster]{A radial manual tour manipulating the contribution of PC2 within the DIS cluster.
+Points are colored by experiment type: DIS HERA1+2 in green, dimuon SIDIS in purple, and charm SIDIS in orange.
+The plane of cross distributed DIS HERA data (green) and a nearly orthogonal jet of dimuon SIDIS (purple) is present.
+This jet does extend more in the plane of view when the contribution of PC2 is full, giving insight to its orientation.
+However, less information about the shape of DIS HERA (green) and charm SIDIS (orange) is available.
+A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/discluster\_manualtour\_pc2/.}\label{fig:DISclusterBad}
 \end{figure}
@@
 
-Tours, the dynamic linear projection of multivariate data, is an important aspect of data visualization extending the display of data-space as data dimensionality increases.
-This research has modified the algorithm producing manual tours, applied this functionality in R and offers extends the graphics offerings that can be used to display tours.
-The paragraphs below explore how this work might be extended.
+Tours, the dynamic linear projection of multivariate data, is an
+important aspect of data visualization extending the display of
+data-space as data dimensionality increases. This research has modified
+the algorithm producing manual tours, applied this functionality in R
+and offers extends the graphics offerings that can be used to display
+tours. The paragraphs below explore how this work might be extended.
 
-Future research on the algorithm would include extending it for use in 3D projections.
-The addition of another dimension theoretically allows for improved perception.
-This could explore interactions in immersive virtual reality or mixed reality, which may further allow for a better perception of structure and aid in higher-dimensional function visualization.
-Functions with many parameters suffer from the same dimensionality problem as data while their possible values lie on a plane of values rather than discrete points.
-Occulation, or the closer surface blocking further surfaces, will likely be an issue that may be alleviated by the use of wire mesh, changing opacity, or looking at sections of the projections (Furnas and Buja 1994).
+Future research on the algorithm would include extending it for use in
+3D projections. The addition of another dimension theoretically allows
+for improved perception. This could explore interactions in immersive
+virtual reality or mixed reality, which may further allow for a better
+perception of structure and aid in higher-dimensional function
+visualization. Functions with many parameters suffer from the same
+dimensionality problem as data while their possible values lie on a
+plane of values rather than discrete points. Occulation, or the closer
+surface blocking further surfaces, will likely be an issue that may be
+alleviated by the use of wire mesh, changing opacity, or looking at
+sections of the projections (Furnas and Buja
+\protect\hyperlink{ref-furnas_prosection_1994}{1994}).
 
-The \textbf{tourr} package provides many other geometric displays with the \texttt{tourr::display\_*()} family.
-These geometric options could be integrated into the \textbf{ggplot2} framework for display on \textbf{plotly} and \textbf{gganimate}.
-Additionally, the \textbf{animation} package Xie et al.
-(2018) could be implemented for another graphics framework.
-However, \textbf{animation} builds from \textbf{base} graphs while \textbf{spinifex} utilizes \textbf{ggplot2} graphics.
+The \textbf{tourr} package provides many other geometric displays with
+the \texttt{tourr::display\_*()} family. These geometric options could
+be integrated into the \textbf{ggplot2} framework for display on
+\textbf{plotly} and \textbf{gganimate}. Additionally, the
+\textbf{animation} package Xie et al.
+(\protect\hyperlink{ref-xie_animation:_2018}{2018}) could be implemented
+for another graphics framework. However, \textbf{animation} builds from
+\textbf{base} graphs while \textbf{spinifex} utilizes \textbf{ggplot2}
+graphics.
 
-The Givens rotations and Householder reflections as outlined in Buja et al.
-(2005) could also be added.
-Currently, Gram-Schmidt is the only form of frame interpolation used (not used in manual tours).
-In a Givens rotation, the \(x\) and \(y\) components (for example \(\theta~= 0,~pi/2\)) of the in-plane rotation are calculated separately and would be applied sequentially to produce the radial rotation.
-Householder reflections define reflection axes to project points on to the axes and generate rotations.
+The Givens rotations and Householder reflections as outlined in Buja et
+al. (\protect\hyperlink{ref-buja_computational_2005}{2005}) could also
+be added. Currently, Gram-Schmidt is the only form of frame
+interpolation used (not used in manual tours). In a Givens rotation, the
+\(x\) and \(y\) components (for example \(\theta~= 0,~pi/2\)) of the
+in-plane rotation are calculated separately and would be applied
+sequentially to produce the radial rotation. Householder reflections
+define reflection axes to project points on to the axes and generate
+rotations.
 
-Having a script only interaction with tours causes a significant barrier to entry.
-To a lesser extent, \textbf{plotly} offers some static interactions with the contained object, such as tooltips, brushing, and linking without communicating back to the R console.
-The development of a dynamic graphical user interface, perhaps with the use of a \textbf{shiny} (Chang et al.
-2018) application, would mitigate the barrier to entry, allow for more rapid analysis, and offer an approachable demo tool.
-The user could easily switch between variables to control, adjust interpolation step angle, or flag/save specific frame basis sets.
+Having a script only interaction with tours causes a significant barrier
+to entry. To a lesser extent, \textbf{plotly} offers some static
+interactions with the contained object, such as tooltips, brushing, and
+linking without communicating back to the R console. The development of
+a dynamic graphical user interface, perhaps with the use of a
+\textbf{shiny} (Chang et al.
+\protect\hyperlink{ref-chang_shiny:_2018}{2018}) application, would
+mitigate the barrier to entry, allow for more rapid analysis, and offer
+an approachable demo tool. The user could easily switch between
+variables to control, adjust interpolation step angle, or flag/save
+specific frame basis sets.
 
@@
 
-This article was created in R (R Core Team 2018), using \textbf{bookdown} (Xie 2016) and \textbf{rmarkdown} (Xie, Allaire, and Grolemund 2018), with code generating the examples inline.
-The source files for this article be found at \href{https://github.com/nspyrison/confirmation/}{github.com/nspyrison/confirmation/}.
-The source code for the \textbf{spinifex} package can be found at \href{https://github.com/nspyrison/spinifex/}{github.com/nspyrison/spinifex/}.
+This article was created in R (R Core Team
+\protect\hyperlink{ref-r_core_team_r:_2018}{2018}), using
+\textbf{bookdown} (Xie \protect\hyperlink{ref-xie_bookdown:_2016}{2016})
+and \textbf{rmarkdown} (Xie, Allaire, and Grolemund
+\protect\hyperlink{ref-xie_r_2018}{2018}), with code generating the
+examples inline. The source files for this article be found at
+\href{https://github.com/nspyrison/confirmation/}{github.com/nspyrison/confirmation/}.
+The source code for the \textbf{spinifex} package can be found at
+\href{https://github.com/nspyrison/spinifex/}{github.com/nspyrison/spinifex/}.
+
+\hypertarget{bibliography}{%
+\subsection*{Bibliography}\label{bibliography}}
+\addcontentsline{toc}{subsection}{Bibliography}
 
@@
 \leavevmode\hypertarget{ref-asimov_grand_1985}{}%
-Asimov, Daniel. 1985. ``The Grand Tour: A Tool for Viewing Multidimensional Data.'' \emph{SIAM Journal on Scientific and Statistical Computing} 6 (1): 128--43.
+Asimov, Daniel. 1985. ``The Grand Tour: A Tool for Viewing
+Multidimensional Data.'' \emph{SIAM Journal on Scientific and
+Statistical Computing} 6 (1): 128--43.
 
 \leavevmode\hypertarget{ref-buja_computational_2005}{}%
-Buja, Andreas, Dianne Cook, Daniel Asimov, and Catherine Hurley.
-2005.
-``Computational Methods for High-Dimensional Rotations in Data Visualization.'' In \emph{Handbook of Statistics}, 24:391--413.
-Elsevier.
-\url{https://doi.org/10.1016/S0169-7161(04)24014-7}.
+Buja, Andreas, Dianne Cook, Daniel Asimov, and Catherine Hurley. 2005.
+``Computational Methods for High-Dimensional Rotations in Data
+Visualization.'' In \emph{Handbook of Statistics}, 24:391--413.
+Elsevier. \url{https://doi.org/10.1016/S0169-7161(04)24014-7}.
 
 \leavevmode\hypertarget{ref-chang_shiny:_2018}{}%
-Chang, Winston, Joe Cheng, J. J. Allaire, Yihui Xie, and Jonathan McPherson. 2018. \emph{Shiny: Web Application Framework for R}. \url{https://CRAN.R-project.org/package=shiny}.
+Chang, Winston, Joe Cheng, J. J. Allaire, Yihui Xie, and Jonathan
+McPherson. 2018. \emph{Shiny: Web Application Framework for R}.
+\url{https://CRAN.R-project.org/package=shiny}.
 
 \leavevmode\hypertarget{ref-cook_manual_1997}{}%
-Cook, Dianne, and Andreas Buja.
-1997.
-``Manual Controls for High-Dimensional Data Projections.'' \emph{Journal of Computational and Graphical Statistics} 6 (4): 464--80.
+Cook, Dianne, and Andreas Buja. 1997. ``Manual Controls for
+High-Dimensional Data Projections.'' \emph{Journal of Computational and
+Graphical Statistics} 6 (4): 464--80.
 \url{https://doi.org/10.2307/1390747}.
@@
 \leavevmode\hypertarget{ref-cook_dynamical_2018}{}%
-Cook, Dianne, Ursula Laa, and German Valencia. 2018. ``Dynamical Projections for the Visualization of PDFSense Data.'' \emph{Eur. Phys. J. C} 78 (9): 742.
+Cook, Dianne, Ursula Laa, and German Valencia. 2018. ``Dynamical
+Projections for the Visualization of PDFSense Data.'' \emph{Eur. Phys.
+J. C} 78 (9): 742.
 
 \leavevmode\hypertarget{ref-cook_interactive_2007}{}%
-Cook, Dianne, Deborah F. Swayne, and A. Buja. 2007. \emph{Interactive and Dynamic Graphics for Data Analysis: With R and GGobi}. Springer Science \& Business Media.
+Cook, Dianne, Deborah F. Swayne, and A. Buja. 2007. \emph{Interactive
+and Dynamic Graphics for Data Analysis: With R and GGobi}. Springer
+Science \& Business Media.
 
 \leavevmode\hypertarget{ref-diaconis_asymptotics_1984}{}%
-Diaconis, Persi, and David Freedman. 1984. ``Asymptotics of Graphical Projection Pursuit.'' \emph{The Annals of Statistics} 12 (3): 793--815.
+Diaconis, Persi, and David Freedman. 1984. ``Asymptotics of Graphical
+Projection Pursuit.'' \emph{The Annals of Statistics} 12 (3): 793--815.
 
 \leavevmode\hypertarget{ref-furnas_prosection_1994}{}%
-Furnas, George W., and Andreas Buja.
-1994.
-``Prosection Views: Dimensional Inference Through Sections and Projections.'' \emph{Journal of Computational and Graphical Statistics} 3 (4): 323--53.
+Furnas, George W., and Andreas Buja. 1994. ``Prosection Views:
+Dimensional Inference Through Sections and Projections.'' \emph{Journal
+of Computational and Graphical Statistics} 3 (4): 323--53.
 \url{https://doi.org/10.2307/1390897}.
@@
 \leavevmode\hypertarget{ref-hurley_analyzing_1990}{}%
-Hurley, C., and A.
-Buja.
-1990.
-``Analyzing High-Dimensional Data with Motion Graphics.'' \emph{SIAM Journal on Scientific and Statistical Computing} 11 (6): 1193--1211.
-\url{https://doi.org/10.1137/0911068}.
+Hurley, C., and A. Buja. 1990. ``Analyzing High-Dimensional Data with
+Motion Graphics.'' \emph{SIAM Journal on Scientific and Statistical
+Computing} 11 (6): 1193--1211. \url{https://doi.org/10.1137/0911068}.
 
 \leavevmode\hypertarget{ref-lubischew_use_1962}{}%
-Lubischew, Alexander A. 1962. ``On the Use of Discriminant Functions in Taxonomy.'' \emph{Biometrics}, 455--77.
+Lubischew, Alexander A. 1962. ``On the Use of Discriminant Functions in
+Taxonomy.'' \emph{Biometrics}, 455--77.
 
 \leavevmode\hypertarget{ref-pedersen_gganimate:_2019}{}%
-Pedersen, Thomas Lin, and David Robinson. 2019. \emph{Gganimate: A Grammar of Animated Graphics}. \url{http://github.com/thomasp85/gganimate}.
+Pedersen, Thomas Lin, and David Robinson. 2019. \emph{Gganimate: A
+Grammar of Animated Graphics}.
+\url{http://github.com/thomasp85/gganimate}.
 
 \leavevmode\hypertarget{ref-r_core_team_r:_2018}{}%
-R Core Team. 2018. \emph{R: A Language and Environment for Statistical Computing}. Vienna, Austria: R Foundation for Statistical Computing. \url{https://www.R-project.org/}.
+R Core Team. 2018. \emph{R: A Language and Environment for Statistical
+Computing}. Vienna, Austria: R Foundation for Statistical Computing.
+\url{https://www.R-project.org/}.
 
 \leavevmode\hypertarget{ref-sievert_plotly_2018}{}%
-Sievert, Carson. 2018. \emph{Plotly for R}. \url{https://plotly-book.cpsievert.me}.
+Sievert, Carson. 2018. \emph{Plotly for R}.
+\url{https://plotly-book.cpsievert.me}.
 
 \leavevmode\hypertarget{ref-wang_mapping_2018}{}%
-Wang, Bo-Ting, T.
-J.
-Hobbs, Sean Doyle, Jun Gao, Tie-Jiun Hou, Pavel M.
-Nadolsky, and Fredrick I.
-Olness.
-2018.
-``Mapping the Sensitivity of Hadronic Experiments to Nucleon Structure.'' \emph{Physical Review D} 98 (9): 094030.
+Wang, Bo-Ting, T. J. Hobbs, Sean Doyle, Jun Gao, Tie-Jiun Hou, Pavel M.
+Nadolsky, and Fredrick I. Olness. 2018. ``Mapping the Sensitivity of
+Hadronic Experiments to Nucleon Structure.'' \emph{Physical Review D} 98
+(9): 094030.
 
 \leavevmode\hypertarget{ref-wickham_tourr_2011}{}%
-Wickham, Hadley, Dianne Cook, Heike Hofmann, and Andreas Buja.
-2011.
-``\textbf{Tourr} : An \emph{R} Package for Exploring Multivariate Data with Projections.'' \emph{Journal of Statistical Software} 40 (2).
+Wickham, Hadley, Dianne Cook, Heike Hofmann, and Andreas Buja. 2011.
+``\textbf{Tourr} : An \emph{R} Package for Exploring Multivariate Data
+with Projections.'' \emph{Journal of Statistical Software} 40 (2).
 \url{https://doi.org/10.18637/jss.v040.i02}.
@@
 \leavevmode\hypertarget{ref-xie_bookdown:_2016}{}%
-Xie, Yihui. 2016. \emph{Bookdown: Authoring Books and Technical Documents with R Markdown}. Boca Raton, Florida: Chapman; Hall/CRC. \url{https://github.com/rstudio/bookdown}.
+Xie, Yihui. 2016. \emph{Bookdown: Authoring Books and Technical
+Documents with R Markdown}. Boca Raton, Florida: Chapman; Hall/CRC.
+\url{https://github.com/rstudio/bookdown}.
 
 \leavevmode\hypertarget{ref-xie_r_2018}{}%
-Xie, Yihui, J. J. Allaire, and Garrett Grolemund. 2018. \emph{R Markdown: The Definitive Guide}. Boca Raton, Florida: Chapman; Hall/CRC. \url{https://bookdown.org/yihui/rmarkdown}.
+Xie, Yihui, J. J. Allaire, and Garrett Grolemund. 2018. \emph{R
+Markdown: The Definitive Guide}. Boca Raton, Florida: Chapman; Hall/CRC.
+\url{https://bookdown.org/yihui/rmarkdown}.
 
 \leavevmode\hypertarget{ref-xie_animation:_2018}{}%
-Xie, Yihui, Christian Mueller, Lijia Yu, and Weicheng Zhu. 2018. \emph{Animation: A Gallery of Animations in Statistics and Utilities to Create Animations}. \url{https://yihui.name/animation}.
+Xie, Yihui, Christian Mueller, Lijia Yu, and Weicheng Zhu. 2018.
+\emph{Animation: A Gallery of Animations in Statistics and Utilities to
+Create Animations}. \url{https://yihui.name/animation}.
 

### 2019-04-09 paper/spinifex_paper.Rmd
References to RJ format, touched up equations.

@@
 
-  A recent paper {@wang_mapping_2018} visualizes the sensitivity of the hadronic experiments to nucleon structure.
+  A recent paper [@wang_mapping_2018] visualizes the sensitivity of the hadronic experiments to nucleon structure.
 Sensitivity was characterized in non-linear 3D embeddings of the first 10 principal components.
@@
 keywords: manual tour, guided tour, grand tour, projection pursuit, high dimensional data, multivariate data, data visualization, statistical graphics, data science
-output:
-  bookdown::pdf_book:
-    #base_format: rticles::rjournal_article #see sandbox for hard code solution.
+preamble: >
+  % preamble packages
+link-citations: true
 fig_caption: yes
 toc: false
-    includes:
-      in_header: preamble.tex
 bibliography: RJrefs.bib
+output: rticles::rjournal_article
+  #bookdown::pdf_book:
+   # base_format: rticles::rjournal_article #see sandbox for hard code solution.
 ---
 \bibliography{RJrefs}
-
+<!-- note that Di/Ursula have different setup and options. use static images.  -->
 ```{r setup_paper, include=FALSE}
@@
 ```
+
 ```{r localFunction, include=FALSE}
@@
 
-The steps are described in more detail below. The R functions used below mentioned briefly, but more complete code example can be found in section \ref{sec:usage}
+## FIX here
+The steps are described in more detail below. The R functions used below mentioned briefly, but more complete code examples are given with
 
@@
 The holes index is maximized when the projected observations are furthest from the center.
-Figure \@ref{fig:step0} shows a locally optimized projection of the data.
+Figure \ref{fig:step0} shows a locally optimized projection of the data.
 The left panel displays the reference axes of the projection basis, a visual indication of the magnitude and direction each variable contributed to the projection.
@@
 
-(ref:step0-cap) Basis reference axes (left) and projected data (right) of standardized flea data.
+```{r step0, echo=F, results='hide', out.width='60%',                             fig.cap = "Basis reference axes (left) and projected data (right) of standardized flea data.
 Data points color and shape are mapped to beetle species.
 Basis identified by a holes-index guided tour.
-The variables `aede2` and `tars1` contribute mostly orthogonal to the other variables.
-
-```{r step0, echo=F, results='hide', out.width='60%', fig.cap = "(ref:step0-cap)"}
-#"\\label{fig:step0} and some caption"}
-
+The variables aede2 and tars1 contribute mostly orthogonal to the other variables."}
 ## Flea holes tour
@@
 
-Call `view_basis()` on a basis to produce a **ggplot2** graphic similar to \@ref(fig:step0).
+Call `view_basis()` on a basis to produce a **ggplot2** graphic similar to \ref{fig:step0).
 Projection space is always available for display via the matrix multiplication $\textbf{X}_{[n,~p]} ~*~ \textbf{B}_{[p,~d]} ~=~ \textbf{P}_{[n,~d]}$.
@@
 <!-- select a manip var-->
-In figure \@ref(fig:step0) the contributions of the variables `tars1` and `aede2` are mostly orthogonal to the contributions of the other four variables.
+In figure \ref{fig:step0} the contributions of the variables `tars1` and `aede2` are mostly orthogonal to the contributions of the other four variables.
 These two variables explain the variation of the data distinguishing the purple group from the rest of the sample.
@@
 \begin{align*}
-  \textbf{e} &\leftarrow Orthonormalize_{GS}(\textbf{e}) ~~~ w.r.t.~the~basis \\
-  &\leftarrow \textbf{e} - \langle \textbf{e},\textbf{B}_1 \rangle \textbf{B}_1 - \langle \textbf{e}, \textbf{B}_2 \rangle \textbf{B}_2 \\
+  \textbf{e_k} &\leftarrow 1 \\
+  \textbf{e}   &\leftarrow \textbf{e} - \langle \textbf{e},\textbf{B}_1 \rangle \textbf{B}_1 - \langle \textbf{e}, \textbf{B}_2 \rangle \textbf{B}_2 \\
   \\
@@
 Place the plane horizontally, with the new dimension extending vertically, with axes projecting back onto the reference axes.
-Figure \@ref(fig:step2) illustrates this 3D manipulation space with the manip var highlighted (height of the other variables are not depicted.)
+Figure \ref{fig:step2} illustrates this 3D manipulation space with the manip var highlighted (height of the other variables are not depicted.)
 
-(ref:step2-cap) Manipulation space for controlling the contribution of `aede2` of standardized flea data.
+```{r step2, echo=F, out.width='60%',                                           fig.cap = "Manipulation space for controlling the contribution of aede2 of standardized flea data.
 Basis selected by a holes-index optimized guided tour lies on the horizontal projection plane, shown in blue.
-The manip var axis, in red, extends into the vertical manipulation space, allowing the coefficients of the manip var to be changed by rotation around the origin.
+The manip var axis, in red, extends into the vertical manipulation space, allowing the coefficients of the manip var to be changed by rotation around the origin."}
 
-```{r step2, echo=F, out.width='60%', fig.cap = "(ref:step2-cap)"}
 view_manip_space(basis = f_bas, manip_var = f_mvar, labels = colnames(f_dat)) +
@@
 
-The representation in figure \@ref(fig:step2) can be produced by calling the function `view_manip_space()`.
+The representation in figure \ref{fig:step2} can be produced by calling the function `view_manip_space()`.
 
@@
   &= \textbf{M}_{[p,~3]} ~*~ \textbf{R}_{[3,~3,~i]} \\
-  &= \begin{bmatrix}
-    M_{1,~1} & M_{1,~2} & M_{1,~3} \\
-    M_{2,~1} & M_{2,~2} & M_{2,~3} \\
-    \vdots   & \vdots   \\
-    M_{p,~1} & M_{p,~2} & M_{p,~3}
-  \end{bmatrix}_{[p,~3]}
+  &= \textbf{M}_{[p,~3]}
     ~*~
@@
 <!-- description of slides, on the technical side we haven't gone back to data space yet. -->
-Figure \@ref(fig:step3) illustrates a sequence with 15 projected bases, showing the reference axes on top the corresponding projected data points on the below.
+Figure \ref{fig:step3} illustrates a sequence with 15 projected bases, showing the reference axes on top the corresponding projected data points on the below.
 As a result, we can see that changes in the manip var controlled the distance between the purple cluster and the remaining sample, `aede2` is crucial in distinguishing this species.
@@
 <!-- step3 caption -->
-(ref:step3-cap) Radial manual tour manipulating `aede2` of standardized flea data.
+
+```{r step3, echo=F, warning=F, fig.height=7.2, fig.width=6, out.height='7.2in', out.width='6in',                                                                fig.cap = "Radial manual tour manipulating aede2 of standardized flea data.
 The contributions increase from its initial contribution to a full contribution to the projection before decreasing to zero and then returning to its initial value.
-The change in the projected data shows that `aede2` is important for distinguishing the purple species.
-An animated version can be viewed at https://nspyrison.netlify.com/thesis/flea_manualtour_mvar5/.
-
-```{r step3, echo=F, warning=F, fig.height=7.2, fig.width=6, out.height='7.2in', out.width='6in', fig.cap = "(ref:step3-cap)"}
+The change in the projected data shows that aede2 is important for distinguishing the purple species.
+An animated version can be viewed at  https://nspyrison.netlify.com/thesis/flea\\_manualtour\\_mvar5/."}
 f_angle <- .26
@@
 
-\begin{align}
-  \textbf{P}_{[n,~3]}
-    &= \textbf{X}_{[n,~p]} ~*~ \textbf{RM}_{[p,~3]} \\
-    &=
-      \begin{bmatrix}
-          X_{1,~1} & \dots & X_{1,~p} \\
-          X_{2,~1} & \dots & X_{2,~p} \\
-          \vdots   & \vdots & \vdots  \\
-          X_{n,~1} & \dots & X_{n,~p}
-      \end{bmatrix}_{[n,~p]}
-      ~*~
-      \begin{bmatrix}
-        RM_{1,~1} & RM_{1,~2} & RM_{1,~3} \\
-        RM_{2,~1} & RM_{2,~2} & RM_{2,~3} \\
-        \vdots     & \vdots     & \vdots  \\
-        RM_{p,~1} & RM_{p,~2} & RM_{p,~3}
-      \end{bmatrix}_{[p,~3]}
-\end{align}
+\begin{align*}
+  \textbf{P}_{[n,~3]} &= \textbf{X}_{[n,~p]} ~*~ \textbf{RM}_{[p,~3]}
+\end{align*}
 
@@
 The data within this 4D embedding is subset down to `ATLAS7old` and `ATLAS7new` to focus in on two groups with a reasonable number of observations that occupy different parts of the subspace.
-Radial manual tours controlling contributions from PC4 and PC3 are shown in figure \@ref(fig:JetClusterGood) and figure \@ref(fig:JetClusterBad) respectively.
+Radial manual tours controlling contributions from PC4 and PC3 are shown in figure \ref{fig:JetClusterGood} and figure \ref{fig:JetClusterBad} respectively.
 These components are selected to contrast the difference of information conveyed by touring different variables.
@@
 
-(ref:JetClusterGood-cap) A radial manual tour of PC4 within the `jet` cluster.
-Colored by experiment type: `ATLAS7new` in green and `ATLAS7old` in orange.
-When PC4 fully/negligibly contributes to the projection `ATLAS7new` (green) spans the same space as the orange points.
-During the intermediate frames, the `ATLAS7new` is compressed in the direction radial to PC4.
-A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc4/.
-
-```{r JetClusterGood, echo=F, warning=F, fig.height=7.2, fig.width=6, out.height='7.2in', out.width='6in', fig.cap="(ref:JetClusterGood-cap)"}
+```{r JetClusterGood, echo=F, warning=F, fig.height=7.2, fig.width=6, out.height='7.2in', out.width='6in',                                            fig.cap="A radial manual tour of PC4 within the jet cluster.
+Colored by experiment type: ATLAS7new in green and ATLAS7old in orange.
+When PC4 fully/negligibly contributes to the projection ATLAS7new (green) spans the same space as the orange points.
+During the intermediate frames, the ATLAS7new is compressed in the direction radial to PC4.
+A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/jetcluster\\_manualtour\\_pc4/."}
 # Jet cluster, Fig 7 of cook_dynamical_2018, subset to ATLAS7old and ATLAS7new
@@
 A different space is used to explore the `DIS` cluster; specifically the first six principal components, which explain 48% of the variation contained within the aggregated data[@cook_dynamical_2018].
-Radial manual tours are performed on PC6 and PC2 in figure \@ref(fig:DISclusterGood) and figure \@ref(fig:DISclusterBad) respectively.
+Radial manual tours are performed on PC6 and PC2 in figure \ref{fig:DISclusterGood} and figure \ref{fig:DISclusterBad} respectively.
 
@@
 
-(ref:DISclusterGood-cap) A radial manual tour manipulating the contribution of  PC6 within the `DIS` cluster.
-Points are colored by experiment type: `DIS HERA1+2` in green, `dimuon SIDIS` in purple, and `charm SIDIS` in orange.
-The cluster `DIS HERA1+2` (green) is distributed in a cross-shaped plane, `charm SIDIS` (orange) occupies the center space of this cross, with the plane projecting into the field of view when the contribution of PC6 is max.
-Less evident is the linear `dimuon SIDIS` (purple) observations approaching the line of view for intermediate values of PC6.
-A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/.
+```{r DISclusterGood, echo=F, warning=F, fig.height=7.2, fig.width=6, out.height='7.2in', out.width='6in',                                            fig.cap = "A radial manual tour manipulating the contribution of  PC6 within the DIS cluster.
+Points are colored by experiment type: DIS HERA1+2 in green, dimuon SIDIS in purple, and charm SIDIS in orange.
+The cluster DIS HERA1+2 (green) is distributed in a cross-shaped plane, charm SIDIS (orange) occupies the center space of this cross, with the plane projecting into the field of view when the contribution of PC6 is max.
+Less evident is the linear dimuon SIDIS (purple) observations approaching the line of view for intermediate values of PC6.
+A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/discluster\\_manualtour\\_pc6/."}
 
-```{r DISclusterGood, echo=F, warning=F, fig.height=7.2, fig.width=6, out.height='7.2in', out.width='6in', fig.cap = "(ref:DISclusterGood-cap)"}
 ##DIS cluster, fig 8 of cook_dynamical_2018
@@
 
-(ref:DISclusterBad-cap) A radial manual tour manipulating the contribution of PC2 within the `DIS` cluster.
-Points are colored by experiment type: `DIS HERA1+2` in green, `dimuon SIDIS` in purple, and `charm SIDIS` in orange.
-The plane of cross distributed `DIS HERA` data (green) and a nearly orthogonal jet of `dimuon SIDIS` (purple) is present.
+```{r DISclusterBad, echo=F, warning=F, fig.height=7.2, fig.width=6, out.height='7.2in', out.width='6in',                                            fig.cap = "A radial manual tour manipulating the contribution of PC2 within the DIS cluster.
+Points are colored by experiment type: DIS HERA1+2 in green, dimuon SIDIS in purple, and charm SIDIS in orange.
+The plane of cross distributed DIS HERA data (green) and a nearly orthogonal jet of dimuon SIDIS (purple) is present.
 This jet does extend more in the plane of view when the contribution of PC2 is full, giving insight to its orientation.
-However, less information about the shape of `DIS HERA` (green) and `charm SIDIS` (orange) is available.
-A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/discluster_manualtour_pc2/.
-
-```{r DISclusterBad, echo=F, warning=F, fig.height=7.2, fig.width=6, out.height='7.2in', out.width='6in', fig.cap = "(ref:DISclusterBad-cap)"}
+However, less information about the shape of DIS HERA (green) and charm SIDIS (orange) is available.
+A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/discluster\\_manualtour\\_pc2/."}
 ##DIS cluster, fig 8 of cook_dynamical_2018

### 2019-04-09 paper/spinifex-paper.tex
remove variable backticks. fixed latex, summary section. grammerly,  knit.

@@
 
-\hypertarget{fix-here}{%
-\subsection{FIX here}\label{fix-here}}
-
-The steps are described in more detail below. The R functions used below
-mentioned briefly, but more complete code examples are given with
+The steps are described in more detail below with description,
+equations, figures, and their code usage.
 
@@
   \item $\textbf{B} = (B_1,~ B_2)$, any orthonormal projection basis set, $p \times 2$ matrix, describing the projection from $p$ to two dimensions
-  \item $\textbf{e}$, a zero column vector of length $p$ with the $k-$th element set to one, where $k$ is the number of the variable to manipulate.
+  \item $\textbf{e}$, a zero column vector of length $p$ with the $k-$th element set to one, where $k$ is the number of the variable to manipulate, the manip var.
   \item $\theta$, the angle of in-projection-plane rotation, for example, on the reference axes.
@@
 
-In figure \ref{fig:step0} the contributions of the variables
-\texttt{tars1} and \texttt{aede2} are mostly orthogonal to the
-contributions of the other four variables. These two variables explain
-the variation of the data distinguishing the purple group from the rest
-of the sample. Select \texttt{aede2} as the manip var, the variable to
-be manipulated, as it has a larger contribution to the projection. The
-question that will be explored is how important the variable
-\texttt{aede2} is to the separation of the clusters.
+In figure \ref{fig:step0} the contributions of the variables tars1 and
+aede2 are mostly orthogonal to the contributions of the other four
+variables. These two variables explain the variation of the data
+distinguishing the purple group from the rest of the sample. Select
+aede2 as the manip var, the variable to be manipulated, as it has a
+larger contribution to the projection. The question that will be
+explored is how important the variable aede2 is to the separation of the
+clusters.
 
@@
 Initialize a zero vector \textbf{e} of length \emph{p}. Set the fifth
-element to one, as \texttt{aede2} is the fifth variable in the data,
-giving the manip var a full contribution in this dimension. Use the
-Gram-Schmidt process to orthonormalize the zero vector onto the basis
-yielding the 3D manipulation space, \textbf{M}.
+element to one, as aede2 is the fifth variable in the data, giving the
+manip var a full contribution in this dimension. Use the Gram-Schmidt
+process to orthonormalize the zero vector onto the basis yielding the 3D
+manipulation space, \textbf{M}.
 
 \begin{align*}
-  \textbf{e} &\leftarrow Orthonormalize_{GS}(\textbf{e}) ~~~ w.r.t.~the~basis \\
-  &\leftarrow \textbf{e} - \langle \textbf{e},\textbf{B}_1 \rangle \textbf{B}_1 - \langle \textbf{e}, \textbf{B}_2 \rangle \textbf{B}_2 \\
-  \\
+  \textbf{e}_k &\leftarrow 1 \\ 
+  \textbf{e}   &\leftarrow \textbf{e} - \langle \textbf{e}, \textbf{B}_1 \rangle \textbf{B}_1 - \langle \textbf{e}, \textbf{B}_2 \rangle \textbf{B}_2 \\ 
+  &\\
   \textbf{M}_{[p,~3]} &= (\textbf{B}_1,\textbf{B}_2,\textbf{e})
@@
 plane, \(\theta\), and define a sequence for the (vertical) angle coming
-out of the projection plane, \(\phi\), bringing the initial \texttt{XY}
+out of the projection plane, \(\phi\), bringing the initial \(XY\)
 contributions of the manip var to a maximum and then to zero before
@@
 var controlled the distance between the purple cluster and the remaining
-sample, \texttt{aede2} is crucial in distinguishing this species. Tours
-are typically viewed as an animation such a dynamic version of this tour
-can be viewed online at
+sample, aede2 is crucial in distinguishing this species. Tours are
+typically viewed as an animation such a dynamic version of this tour can
+be viewed online at
 \url{https://nspyrison.netlify.com/thesis/flea_manualtour_mvar5/}. The
@@
 
-The data has a hierarchical structure with top-level clusters;
-\texttt{DIS}, \texttt{VBP}, and \texttt{jet}. Each cluster is a
-particular class of experiments, each with many experimental datasets
-which, in turn, have many observations. In the consideration of data
-density, we conduct manual tours on subsets of the \texttt{DIS} and
-\texttt{jet} clusters. This explores the sensitivity of the structure to
-each of the variables in turn and we present the subjectively best and
-worst variable to manipulate for identifying dimensionality of the
+The data has a hierarchical structure with top-level clusters; DIS, VBP,
+and jet. Each cluster is a particular class of experiments, each with
+many experimental datasets which, in turn, have many observations. In
+the consideration of data density, we conduct manual tours on subsets of
+the DIS and jet clusters. This explores the sensitivity of the structure
+to each of the variables in turn and we present the subjectively best
+and worst variable to manipulate for identifying dimensionality of the
 clusters and describing the span of the clusters.
@@
 
-The \texttt{jet} cluster resides in a smaller dimensionality than the
-full set of experiments with four principal components explaining 95\%
-of the variation in the cluster (Cook, Laa, and Valencia
+The jet cluster resides in a smaller dimensionality than the full set of
+experiments with four principal components explaining 95\% of the
+variation in the cluster (Cook, Laa, and Valencia
 \protect\hyperlink{ref-cook_dynamical_2018}{2018}). The data within this
-4D embedding is subset down to \texttt{ATLAS7old} and \texttt{ATLAS7new}
-to focus in on two groups with a reasonable number of observations that
-occupy different parts of the subspace. Radial manual tours controlling
+4D embedding is subset down to ATLAS7old and ATLAS7new to focus in on
+two groups with a reasonable number of observations that occupy
+different parts of the subspace. Radial manual tours controlling
 contributions from PC4 and PC3 are shown in figure
@@
 When manipulating PC4, there is a clear difference in the parameter
-space spanned by the experiment types \texttt{ATLAS7new} and
-\texttt{ATLAS7old}. Yet, when PC3 is manipulated there is no indication
-that the different experiments probe different parameter space.
-Performing a radial manual tour on PC4 is more insightful than for PC3.
+space spanned by the experiment types ATLAS7new and ATLAS7old. Yet, when
+PC3 is manipulated there is no indication that the different experiments
+probe different parameter space. Performing a radial manual tour on PC4
+is more insightful than for PC3.
 
 Radial manual tours manipulating each of the principal components in the
-\texttt{jet} cluster can be viewed by following the links:
+jet cluster can be viewed by following the links:
 \href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc1/}{PC1},
@@
 
-(ref:JetClusterBad-cap) A radial manual tour of PC4 within the
-\texttt{jet} cluster. Colored by experiment type: \texttt{ATLAS7new} in
-green and \texttt{ATLAS7old} in orange. Data from \texttt{ATLAS7new}
-(green) spans mostly the same space as \texttt{ALTLAS7old} (orange) with
-no evident difference in cluster structure across varying contributions
-of PC3. A dynamic version can be viewed at
+(ref:JetClusterBad-cap) A radial manual tour of PC4 within the jet
+cluster. Colored by experiment type: ATLAS7new in green and ATLAS7old in
+orange. Data from ATLAS7new (green) spans mostly the same space as
+ALTLAS7old (orange) with no evident difference in cluster structure
+across varying contributions of PC3. A dynamic version can be viewed at
 \url{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/}.
@@
 
-A different space is used to explore the \texttt{DIS} cluster;
-specifically the first six principal components, which explain 48\% of
-the variation contained within the aggregated data(Cook, Laa, and
-Valencia \protect\hyperlink{ref-cook_dynamical_2018}{2018}). Radial
-manual tours are performed on PC6 and PC2 in figure
-\ref{fig:DISclusterGood} and figure \ref{fig:DISclusterBad}
-respectively.
+A different space is used to explore the DIS cluster; specifically the
+first six principal components, which explain 48\% of the variation
+contained within the aggregated data(Cook, Laa, and Valencia
+\protect\hyperlink{ref-cook_dynamical_2018}{2018}). Radial manual tours
+are performed on PC6 and PC2 in figure \ref{fig:DISclusterGood} and
+figure \ref{fig:DISclusterBad} respectively.
 
@@
 
-\texttt{DIS} cluster manual tours manipulating each of the principal
-components can be viewed from the links:
+DIS cluster manual tours manipulating each of the principal components
+can be viewed from the links:
 \href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc1/}{PC1},
@@
 
-\hypertarget{todo}{%
-\subsection{TODO:}\label{todo}}
-
 \hypertarget{summary}{%
@@
 
+Tours, the dynamic linear projection of multivariate data, play an
+important role in data visualization; they extending the dimensionality
+of visuals while data- and parameter-spaces become ever larger. This
+research has modified the algorithm producing manual tours and has made
+this functionality available in package \textbf{spinifex}. The package
+adds to \textbf{tourr}, extending the graphics offerings that can be
+used to display tours.
+
+Radial manual tours were applied to a dataset across different
+experiments of hadronic collisions. The importance of selecting the
+correct variable to manipulate is demonstrated by comparing tours of
+varying quality. The manual tours convey a better picture of the
+structure of the clustering than static linear or non-linear projections
+are able to. Giving the full contribution of the manipulation space to
+the manip var enables analysts to explore the sensitivity of the
+structure with the selected variable. This information can be used by
+physicists to which experiments are probing which parameter spaces,
+which can be indispensable for smarter planning of effort and funding.
+
 \hypertarget{sec:discussion}{%
 \subsection{Discussion}\label{sec:discussion}}
-
-Tours, the dynamic linear projection of multivariate data, is an
-important aspect of data visualization extending the display of
-data-space as data dimensionality increases. This research has modified
-the algorithm producing manual tours, applied this functionality in R
-and offers extends the graphics offerings that can be used to display
-tours. The paragraphs below explore how this work might be extended.
 

### 2019-04-09 paper/spinifex_paper.Rmd
remove variable backticks. fixed latex, summary section. grammerly,  knit.

@@
 
-## FIX here
-The steps are described in more detail below. The R functions used below mentioned briefly, but more complete code examples are given with
+The steps are described in more detail below with description, equations, figures, and their code usage.
 
@@
 <!-- select a manip var-->
-In figure \ref{fig:step0} the contributions of the variables `tars1` and `aede2` are mostly orthogonal to the contributions of the other four variables.
+In figure \ref{fig:step0} the contributions of the variables tars1 and aede2 are mostly orthogonal to the contributions of the other four variables.
 These two variables explain the variation of the data distinguishing the purple group from the rest of the sample.
-Select `aede2` as the manip var, the variable to be manipulated, as it has a larger contribution to the projection.
-The question that will be explored is how important the variable `aede2` is to the separation of the clusters.
+Select aede2 as the manip var, the variable to be manipulated, as it has a larger contribution to the projection.
+The question that will be explored is how important the variable aede2 is to the separation of the clusters.
 
@@
 Initialize a zero vector **e** of length *p*.
-Set the fifth element to one, as  `aede2` is the fifth variable in the data, giving the manip var a full contribution in this dimension.
+Set the fifth element to one, as  aede2 is the fifth variable in the data, giving the manip var a full contribution in this dimension.
 Use the Gram-Schmidt process to orthonormalize the zero vector onto the basis yielding the 3D manipulation space, **M**.
@@
 \begin{align*}
-  \textbf{e_k} &\leftarrow 1 \\
-  \textbf{e}   &\leftarrow \textbf{e} - \langle \textbf{e},\textbf{B}_1 \rangle \textbf{B}_1 - \langle \textbf{e}, \textbf{B}_2 \rangle \textbf{B}_2 \\
-  \\
+  \textbf{e}_k &\leftarrow 1 \\ 
+  \textbf{e}   &\leftarrow \textbf{e} - \langle \textbf{e}, \textbf{B}_1 \rangle \textbf{B}_1 - \langle \textbf{e}, \textbf{B}_2 \rangle \textbf{B}_2 \\ 
+  &\\
   \textbf{M}_{[p,~3]} &= (\textbf{B}_1,\textbf{B}_2,\textbf{e})
@@
 <!-- describe manip var path -->
-For a radial tour fix the (horizontal) angle within the projection plane, $\theta$, and define a sequence for the (vertical) angle coming out of the projection plane, $\phi$, bringing the initial `XY` contributions of the manip var to a maximum and then to zero before returning to the initial position.
+For a radial tour fix the (horizontal) angle within the projection plane, $\theta$, and define a sequence for the (vertical) angle coming out of the projection plane, $\phi$, bringing the initial $XY$ contributions of the manip var to a maximum and then to zero before returning to the initial position.
 Dynamic capture of user manipulation is typically performed directly on the projection plane (without depiction of the manipulation space.)
@@
 Figure \ref{fig:step3} illustrates a sequence with 15 projected bases, showing the reference axes on top the corresponding projected data points on the below.
-As a result, we can see that changes in the manip var controlled the distance between the purple cluster and the remaining sample, `aede2` is crucial in distinguishing this species.
+As a result, we can see that changes in the manip var controlled the distance between the purple cluster and the remaining sample, aede2 is crucial in distinguishing this species.
 Tours are typically viewed as an animation such a dynamic version of this tour can be viewed online at https://nspyrison.netlify.com/thesis/flea_manualtour_mvar5/.
@@
 <!-- Data structure -->
-The data has a hierarchical structure with top-level clusters; `DIS`, `VBP`, and `jet`.
+The data has a hierarchical structure with top-level clusters; DIS, VBP, and jet.
 Each cluster is a particular class of experiments, each with many experimental datasets which, in turn, have many observations.
-In the consideration of data density, we conduct manual tours on subsets of the `DIS` and `jet` clusters.
+In the consideration of data density, we conduct manual tours on subsets of the DIS and jet clusters.
 This explores the sensitivity of the structure to each of the variables in turn and we present the subjectively best and worst variable to manipulate for identifying dimensionality of the clusters and describing the span of the clusters.
@@
 <!-- jet cluster, explain dimensionality  -->
-The `jet` cluster resides in a smaller dimensionality than the full set of experiments with four principal components explaining 95% of the variation in the cluster [@cook_dynamical_2018].
-The data within this 4D embedding is subset down to `ATLAS7old` and `ATLAS7new` to focus in on two groups with a reasonable number of observations that occupy different parts of the subspace.
+The jet cluster resides in a smaller dimensionality than the full set of experiments with four principal components explaining 95% of the variation in the cluster [@cook_dynamical_2018].
+The data within this 4D embedding is subset down to ATLAS7old and ATLAS7new to focus in on two groups with a reasonable number of observations that occupy different parts of the subspace.
 Radial manual tours controlling contributions from PC4 and PC3 are shown in figure \ref{fig:JetClusterGood} and figure \ref{fig:JetClusterBad} respectively.
@@
 <!-- discussion of findings and which is more insightful -->
-When manipulating PC4, there is a clear difference in the parameter space spanned by the experiment types `ATLAS7new` and `ATLAS7old`.
+When manipulating PC4, there is a clear difference in the parameter space spanned by the experiment types ATLAS7new and ATLAS7old.
 Yet, when PC3 is manipulated there is no indication that the different experiments probe different parameter space.
@@
 
-Radial manual tours manipulating each of the principal components in the `jet` cluster can be viewed by following the links: [PC1](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc1/), [PC2](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc2/), [PC3](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/), and [PC4](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc4/).
+Radial manual tours manipulating each of the principal components in the jet cluster can be viewed by following the links: [PC1](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc1/), [PC2](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc2/), [PC3](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/), and [PC4](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc4/).
 
@@
 
-(ref:JetClusterBad-cap) A radial manual tour of PC4 within the `jet` cluster.
-Colored by experiment type: `ATLAS7new` in green and `ATLAS7old` in orange.
-Data from `ATLAS7new` (green) spans mostly the same space as `ALTLAS7old` (orange) with no evident difference in cluster structure across varying contributions of PC3.
+(ref:JetClusterBad-cap) A radial manual tour of PC4 within the jet cluster.
+Colored by experiment type: ATLAS7new in green and ATLAS7old in orange.
+Data from ATLAS7new (green) spans mostly the same space as ALTLAS7old (orange) with no evident difference in cluster structure across varying contributions of PC3.
 A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/.
@@
 <!-- introduce DIS cluster -->
-A different space is used to explore the `DIS` cluster; specifically the first six principal components, which explain 48% of the variation contained within the aggregated data[@cook_dynamical_2018].
+A different space is used to explore the DIS cluster; specifically the first six principal components, which explain 48% of the variation contained within the aggregated data[@cook_dynamical_2018].
 Radial manual tours are performed on PC6 and PC2 in figure \ref{fig:DISclusterGood} and figure \ref{fig:DISclusterBad} respectively.
@@
 
-`DIS` cluster manual tours manipulating each of the principal components can be viewed from the links: [PC1](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc1/), [PC2](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc2/), [PC3](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc3/), [PC4](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc4/), [PC5](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc5/), and [PC6](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/).
+DIS cluster manual tours manipulating each of the principal components can be viewed from the links: [PC1](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc1/), [PC2](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc2/), [PC3](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc3/), [PC4](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc4/), [PC5](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc5/), and [PC6](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/).
 
@@
 
-## TODO:
 ### Summary
 
+<!-- summary and segue  -->
+Tours, the dynamic linear projection of multivariate data, play an important role in data visualization; they extending the dimensionality of visuals while data- and parameter-spaces become ever larger.
+This research has modified the algorithm producing manual tours and has made this functionality available in package **spinifex**.
+The package adds to **tourr**, extending the graphics offerings that can be used to display tours.
+
+Radial manual tours were applied to a dataset across different experiments of hadronic collisions.
+The importance of selecting the correct variable to manipulate is demonstrated by comparing tours of varying quality.
+The manual tours convey a better picture of the structure of the clustering than static linear or non-linear projections are able to.
+Giving the full contribution of the manipulation space to the manip var enables analysts to explore the sensitivity of the structure with the selected variable.
+This information can be used by physicists to which experiments are probing which parameter spaces, which can be indispensable for smarter planning of effort and funding.
+
 ## Discussion {#sec:discussion}
-
-<!-- summary and segue  -->
-Tours, the dynamic linear projection of multivariate data, is an important aspect of data visualization extending the display of data-space as data dimensionality increases.
-This research has modified the algorithm producing manual tours, applied this functionality in R and offers extends the graphics offerings that can be used to display tours.
-The paragraphs below explore how this work might be extended.
 

### 2019-04-09 paper/spinifex-paper.tex
polish and proofreading checks.

@@
 datasets and models exist in increasing cardinality. The package
-\textbf{tourr} (Wickham et al.
+\CRANpkg{tourr} (Wickham et al.
 \protect\hyperlink{ref-wickham_tourr_2011}{2011}) builds a platform for
@@
 than exploring the large phase space of the data, a guided tour selects
-a very specific path via projection pursuit(Diaconis and Freedman
+a very specific path via projection pursuit (Diaconis and Freedman
 \protect\hyperlink{ref-diaconis_asymptotics_1984}{1984}), the
@@
 
-Spinifex utilizes two new animation packages, \textbf{plotly} (Sievert
-\protect\hyperlink{ref-sievert_plotly_2018}{2018}) and
-\textbf{gganimate} (Pedersen and Robinson
+\CRANpkg{spinifex} utilizes two new animation packages, \CRANpkg{plotly}
+(Sievert \protect\hyperlink{ref-sievert_plotly_2018}{2018}) and
+\CRANpkg{gganimate} (Pedersen and Robinson
 \protect\hyperlink{ref-pedersen_gganimate:_2019}{2019}), to display
@@
 
-{\centering \includegraphics[width=0.6\linewidth]{spinifex_paper_files/figure-latex/step0-1}
+{\centering \includegraphics[width=0.7\linewidth]{spinifex_paper_files/figure-latex/step0-1}
 
@@
 
-Call \texttt{view\_basis()} on a basis to produce a \textbf{ggplot2}
-graphic similar to \textbackslash ref\{fig:step0). Projection space is
-always available for display via the matrix multiplication
+Call \texttt{view\_basis()} on a basis to produce a \CRANpkg{ggplot2}
+(Wickham \protect\hyperlink{ref-wickham_ggplot2:_2016}{2016}) graphic
+similar to \textbackslash ref\{fig:step0). Projection space is always
+available for display via the matrix multiplication
 \(\textbf{X}_{[n,~p]} ~*~ \textbf{B}_{[p,~d]} ~=~ \textbf{P}_{[n,~d]}\).
@@
 
-# Also see vignette: 
-vignette("spinifex") # vignette ‘spinifex’ not found
+# Also see the vignette: 
+vignette("spinifex")
 
@@
 
-{\centering \includegraphics[width=0.6\linewidth]{spinifex_paper_files/figure-latex/step2-1}
+{\centering \includegraphics[width=0.7\linewidth]{spinifex_paper_files/figure-latex/step2-1}
 
@@
 produced from \textbf{plotly}. The \texttt{render\_type} argument can be
-changed to \texttt{render\_gganimate} for exporting to .gif or .mp4
-files.
+changed to \texttt{render\_gganimate} for exporting to GIF or MP4 files.
 
@@
 display of tours. \textbf{spinifex} allows tours to be used in rendered
-in \textbf{plotly} as an HTML5 object or \textbf{gganimate} as .gif or
-.mp4 objects. Both of which build off \textbf{ggplot2} objects in
-internal functions. Sharing of animations is not trivial especially in
-print and static file formats such as .pdf. Even with the use of
-computers and dynamic file formats capturing the correct resolution,
-aspect, and display is challenging and many formats quickly bloat file
-sizes. Keep in mind hosting options and exporting functions from
-\textbf{plotly}, \textbf{gganimate} and \textbf{tourr}.
+in \textbf{plotly} as an HTML5 object or \textbf{gganimate} as GIF or
+MP4 files. Both of which build off \textbf{ggplot2} objects in internal
+functions. Sharing of animations is not trivial especially in print and
+static file formats such as PDF. Even with the use of computers and
+dynamic file formats capturing the correct resolution, aspect, and
+display is challenging and many formats quickly bloat file sizes. Keep
+in mind hosting options and exporting functions from \textbf{plotly},
+\textbf{gganimate} and \textbf{tourr}.
 
@@
 \textbf{plotly} and \textbf{gganimate}. Additionally, the
-\textbf{animation} package Xie et al.
+\CRANpkg{animation} package Xie et al.
 (\protect\hyperlink{ref-xie_animation:_2018}{2018}) could be implemented
@@
 a dynamic graphical user interface, perhaps with the use of a
-\textbf{shiny} (Chang et al.
+\CRANpkg{shiny} (Chang et al.
 \protect\hyperlink{ref-chang_shiny:_2018}{2018}) application, would
@@
 \protect\hyperlink{ref-r_core_team_r:_2018}{2018}), using
-\textbf{bookdown} (Xie \protect\hyperlink{ref-xie_bookdown:_2016}{2016})
-and \textbf{rmarkdown} (Xie, Allaire, and Grolemund
+\CRANpkg{knitr} (Xie \protect\hyperlink{ref-stodden_knitr:_2014}{2014})
+and \CRANpkg{rmarkdown} (Xie, Allaire, and Grolemund
 \protect\hyperlink{ref-xie_r_2018}{2018}), with code generating the
 examples inline. The source files for this article be found at
-\href{https://github.com/nspyrison/confirmation/}{github.com/nspyrison/confirmation/}.
+\href{https://github.com/nspyrison/spinifex_paper/}{github.com/nspyrison/spinifex\_paper/}.
 The source code for the \textbf{spinifex} package can be found at
@@
 
+\leavevmode\hypertarget{ref-wickham_ggplot2:_2016}{}%
+Wickham, Hadley. 2016. \emph{Ggplot2: Elegant Graphics for Data
+Analysis}. Springer-Verlag New York. \url{http://ggplot2.org}.
+
 \leavevmode\hypertarget{ref-wickham_tourr_2011}{}%
@@
 
-\leavevmode\hypertarget{ref-xie_bookdown:_2016}{}%
-Xie, Yihui. 2016. \emph{Bookdown: Authoring Books and Technical
-Documents with R Markdown}. Boca Raton, Florida: Chapman; Hall/CRC.
-\url{https://github.com/rstudio/bookdown}.
+\leavevmode\hypertarget{ref-stodden_knitr:_2014}{}%
+Xie, Yihui. 2014. ``Knitr: A Comprehensive Tool for Reproducible
+Research in R.'' In \emph{Implementing Reproducible Computational
+Research}, edited by Victoria Stodden, Friedrich Leisch, and Roger D.
+Peng. Chapman; Hall/CRC.
+\url{http://www.crcpress.com/product/isbn/9781466561595}.
 
@@
 Monash University\\
-School of Information Technology\\ Monash University, Melbourne, VIC 3800, Australia\\
+Faculty of Information Technology\\
 }
@@
 Monash University\\
-School of Econometrics and Business Statistics\\ Monash University, Melbourne, VIC 3800, Australia\\
+Department of Econometrics and Business Statistics\\
 }

### 2019-04-09 paper/spinifex_paper.Rmd
polish and proofreading checks.

@@
     address:
-    - School of Information Technology
-    - Monash University, Melbourne, VIC 3800, Australia
+    - Faculty of Information Technology
     email:  Nicholas.Spyrison@monash.edu
@@
     address:
-    - School of Econometrics and Business Statistics
-    - Monash University, Melbourne, VIC 3800, Australia
+    - Department of Econometrics and Business Statistics
     email:  DiCook@monash.edu
@@
 This is a crucial tool extending the dimensionality of visualization which is a large boon as datasets and models exist in increasing cardinality.
-The package **tourr** [@wickham_tourr_2011] builds a platform for generating tour paths and applying to a wide number of display geometrics in **base** graphics.
+The package \CRANpkg{tourr} [@wickham_tourr_2011] builds a platform for generating tour paths and applying to a wide number of display geometrics in **base** graphics.
 
@@
 This makes manual tours particularly useful once a feature of interest has been identified, for example, with the use of a guided tour [@hurley_analyzing_1990].
-Rather than exploring the large phase space of the data, a guided tour selects a very specific path via projection pursuit[@diaconis_asymptotics_1984], the optimization of an index function on the projection via a hill climbing algorithm.
+Rather than exploring the large phase space of the data, a guided tour selects a very specific path via projection pursuit [@diaconis_asymptotics_1984], the optimization of an index function on the projection via a hill climbing algorithm.
 This allows guided tours to rapidly identify interesting projection features given the relatively large parameter-space.
@@
 <!-- Display -->
-Spinifex utilizes two new animation packages, **plotly** [@sievert_plotly_2018] and **gganimate** [@pedersen_gganimate:_2019], to display tours, manual or other saved tours.
+\CRANpkg{spinifex} utilizes two new animation packages, \CRANpkg{plotly} [@sievert_plotly_2018] and \CRANpkg{gganimate} [@pedersen_gganimate:_2019], to display tours, manual or other saved tours.
 From a given projection, the user can choose which variable to control, and the animation sequence is generated to remove the variable from the projection, and then extend its contribution to be the sole variable in one direction.
@@
 
-```{r step0, echo=F, results='hide', out.width='60%',                             fig.cap = "Basis reference axes (left) and projected data (right) of standardized flea data.
+```{r step0, echo=F, results='hide', out.width='70%',                             fig.cap = "Basis reference axes (left) and projected data (right) of standardized flea data.
 Data points color and shape are mapped to beetle species.
@@
 
-Call `view_basis()` on a basis to produce a **ggplot2** graphic similar to \ref{fig:step0).
+Call `view_basis()` on a basis to produce a \CRANpkg{ggplot2} [@wickham_ggplot2:_2016] graphic similar to \ref{fig:step0).
 Projection space is always available for display via the matrix multiplication $\textbf{X}_{[n,~p]} ~*~ \textbf{B}_{[p,~d]} ~=~ \textbf{P}_{[n,~d]}$.
@@
 
-```{r step2, echo=F, out.width='60%',                                           fig.cap = "Manipulation space for controlling the contribution of aede2 of standardized flea data.
+```{r step2, echo=F, out.width='70%',                                           fig.cap = "Manipulation space for controlling the contribution of aede2 of standardized flea data.
 Basis selected by a holes-index optimized guided tour lies on the horizontal projection plane, shown in blue.
@@
 if(dim(f_mtour)[3] != 15) message(
-  paste0("!!!!! step3 n_slides != 15 !!!!! n_slides = ", dim(f_mtour)[3]))
+  paste0("!!!!! step3 n_slides != 15 !!!!! Please knit again. n_slides = ", dim(f_mtour)[3]))
 
@@
 This function defaults to an HTML5 widget produced from **plotly**.
-The `render_type` argument can be changed to `render_gganimate` for exporting to .gif or .mp4 files.
+The `render_type` argument can be changed to `render_gganimate` for exporting to GIF or MP4 files.
 
@@
 The **tourr** package utilizes **base** graphics for the display of tours.
-**spinifex** allows tours to be used in rendered in **plotly** as an HTML5 object or **gganimate** as .gif or .mp4 objects.
+**spinifex** allows tours to be used in rendered in **plotly** as an HTML5 object or **gganimate** as GIF or MP4 files.
 Both of which build off **ggplot2** objects in internal functions.
-Sharing of animations is not trivial especially in print and static file formats such as .pdf.
+Sharing of animations is not trivial especially in print and static file formats such as PDF.
 Even with the use of computers and dynamic file formats capturing the correct resolution, aspect, and display is challenging and many formats quickly bloat file sizes.
@@
 These geometric options could be integrated into the **ggplot2** framework for display on **plotly** and **gganimate**.
-Additionally, the **animation** package @xie_animation:_2018 could be implemented for another graphics framework.
+Additionally, the \CRANpkg{animation} package @xie_animation:_2018 could be implemented for another graphics framework.
 However, **animation** builds from **base** graphs while **spinifex** utilizes **ggplot2** graphics.
@@
 To a lesser extent, **plotly** offers some static interactions with the contained object, such as tooltips, brushing, and linking without communicating back to the R console.
-The development of a dynamic graphical user interface, perhaps with the use of a **shiny** [@chang_shiny:_2018] application, would mitigate the barrier to entry, allow for more rapid analysis, and offer an approachable demo tool.
+The development of a dynamic graphical user interface, perhaps with the use of a \CRANpkg{shiny} [@chang_shiny:_2018] application, would mitigate the barrier to entry, allow for more rapid analysis, and offer an approachable demo tool.
 The user could easily switch between variables to control, adjust interpolation step angle, or flag/save specific frame basis sets.
@@
 
-This article was created in R [@r_core_team_r:_2018], using **bookdown** [@xie_bookdown:_2016] and **rmarkdown** [@xie_r_2018], with code generating the examples inline.
-The source files for this article be found at [github.com/nspyrison/confirmation/](https://github.com/nspyrison/confirmation/).
+This article was created in R [@r_core_team_r:_2018], using \CRANpkg{knitr} [@stodden_knitr:_2014] and \CRANpkg{rmarkdown} [@xie_r_2018], with code generating the examples inline.
+The source files for this article be found at [github.com/nspyrison/spinifex_paper/](https://github.com/nspyrison/spinifex_paper/).
 The source code for the **spinifex** package can be found at [github.com/nspyrison/spinifex/](https://github.com/nspyrison/spinifex/).

### 2019-04-11 paper/spinifex_paper.Rmd
Writing changes after meeting with Julie.

@@
 
-  A recent paper [@wang_mapping_2018] visualizes the sensitivity of the hadronic experiments to nucleon structure.
+  A recent paper [@wang_mapping_2018] visualized the sensitivity of the hadronic experiments to nucleon structure.
 Sensitivity was characterized in non-linear 3D embeddings of the first 10 principal components.
-This research applies manual tours to this data showing that manual tours resolve information about cluster structure and identify outliers, information that is orthogonal to the original static viewing plane.
+This research applies manual tours to this data showing that manual tours resolve information about the structure of cluster distribution by bringing to light information that is orthogonal to static viewing planes.
 keywords: manual tour, guided tour, grand tour, projection pursuit, high dimensional data, multivariate data, data visualization, statistical graphics, data science
@@
 \bibliography{RJrefs}
-<!-- note that Di/Ursula have different setup and options. use static images.  -->
+<!-- note that Di/Ursula have different setup and options and use static images.  -->
 ```{r setup_paper, include=FALSE}
@@
 A tour is a multivariate data analysis technique in which a sequence of linear (orthogonal) projections are viewed as an animation while the orientation of the projection basis is rotated across time.
-Each frame of the sequence corresponds to a small change in the projection for a smooth transition that perseveres continuity.
-This is a crucial tool extending the dimensionality of visualization which is a large boon as datasets and models exist in increasing cardinality.
+Each frame of the sequence corresponds to a small change in the projection for a smooth transition that preserves continuity.
+Tours are a crucial tool extending the dimensionality of visualization which is increasingly important as datasets and models exist in increasing cardinality.
 The package \CRANpkg{tourr} [@wickham_tourr_2011] builds a platform for generating tour paths and applying to a wide number of display geometrics in **base** graphics.
@@
 The manual tour was described in @cook_manual_1997 and allows a user to control the projection coefficients of a selected variable in a 2D projection.
-The manipulation of these coefficients allows the analyst to explore how sensitive the projections structure is to these changes.
-This makes manual tours particularly useful once a feature of interest has been identified, for example, with the use of a guided tour [@hurley_analyzing_1990].
-Rather than exploring the large phase space of the data, a guided tour selects a very specific path via projection pursuit [@diaconis_asymptotics_1984], the optimization of an index function on the projection via a hill climbing algorithm.
-This allows guided tours to rapidly identify interesting projection features given the relatively large parameter-space.
-After a projection of interest is identified an analyst can then use the "finer brush" of the manual tour, by controlling the contributions of individual variables.
+The manipulation of these coefficients allows the analyst to explore how sensitive the projection structure is to these manipulations.
+As manual tours operate on only one variable at a time, they particularly useful once a feature of interest has been identified.
+One way to identify "interesting" features is with the use of a guided tour [@hurley_analyzing_1990].
+Guided tours select a very specific path, that which approachs a projection that optimizes an objective-function, similar to simulated annealing [@kirkpatrick_optimization_1983].
+The direct optimization of a function allows guided tours to rapidly identify interesting projection features given the relatively large parameter-space.
+After a projection of interest is identified an analyst can then use the "finer brush" of the manual tour, by controlling the contributions of individual variables to explore the sensitivity they have to the projection structure.
 
-<!-- Display -->
-\CRANpkg{spinifex} utilizes two new animation packages, \CRANpkg{plotly} [@sievert_plotly_2018] and \CRANpkg{gganimate} [@pedersen_gganimate:_2019], to display tours, manual or other saved tours.
-From a given projection, the user can choose which variable to control, and the animation sequence is generated to remove the variable from the projection, and then extend its contribution to be the sole variable in one direction.
-This allows the viewer to assess the change in structure induced in the projection by the variable's contribution.
+<!-- Contributions -->
+This paper offers an implementation of manual tours with the package
+\CRANpkg{spinifex}.
+This package also exends the graphics display of **tourr** to the animation frameworks offered by \CRANpkg{plotly} [@sievert_plotly_2018] and \CRANpkg{gganimate} [@pedersen_gganimate:_2019].
+The algorithm used to perform a radial manual tour is demonstrated on a toy data set.
+Followed by an application and disscussion of the radial manual tour on an aggregation of high energy physics experiments in a large multivariate space.
 
@@
 Section \ref{sec:algorithm} explains the algorithm using a toy dataset.
-Section \ref{sec:display} discussed the display of the animation after the path has been generated.
+Section \ref{sec:display} discusses the display of the animation after the path has been generated.
 Section \ref{sec:application} illustrates how this can be used for sensitivity analysis applied to contemporary high energy physics.
@@
 The flea data, originally from @lubischew_use_1962, made available in **tourr**  is used to illustrate the algorithm.
-The data contains 74 observations across 6 variables, physical measurements of the flea beetles.
-Each observation belonging to one of three species.
+The data contains 74 observations across 6 variables, which are physical measurements of the flea beetles.
+Each observation belong to one of three species.
 
@@
 A basis set (ideally orienting to an interesting feature) must be provided as an initial orientation.
-One way to identify a projection containing interesting features is to apply a guided tour [@cook_interactive_2007].
-In a guided tour the projection sequence is selected by optimizing an index function via hill-climbing on the projection space.
+One way to identify a projection containing interesting features is to apply a guided tour.
+In a guided tour, the projection sequence is selected by optimizing an index function via hill-climbing on the projection space.
 In this case, the holes index is selected and applied to standardized flea data.
@@
 Data points color and shape are mapped to beetle species.
-Basis identified by a holes-index guided tour.
+The basis was identified by a holes-index guided tour.
 The variables aede2 and tars1 contribute mostly orthogonal to the other variables."}
@@
 
-(ref:JetClusterBad-cap) A radial manual tour of PC4 within the jet cluster.
+```{r JetClusterBad, echo=F, warning=F, fig.height=7.2, fig.width=6, out.height='7.2in', out.width='6in',                                       fig.cap = "A radial manual tour of PC4 within the jet cluster.
 Colored by experiment type: ATLAS7new in green and ATLAS7old in orange.
 Data from ATLAS7new (green) spans mostly the same space as ALTLAS7old (orange) with no evident difference in cluster structure across varying contributions of PC3.
-A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/.
-
-```{r JetClusterBad, echo=F, warning=F, fig.height=7.2, fig.width=6, out.height='7.2in', out.width='6in', fig.cap = "(ref:JetClusterBad-cap)"}
+A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/jetcluster\\_manualtour\\_pc3/."}
 # Jet cluster, Fig 7 of cook_dynalical_2018, subset to ATLAS7old and ATLAS7new
@@
 <!-- introduce DIS cluster -->
-A different space is used to explore the DIS cluster; specifically the first six principal components, which explain 48% of the variation contained within the aggregated data[@cook_dynamical_2018].
+A different space is used to explore the DIS cluster; specifically the first six principal components, which explains 48% of the variation contained within the aggregated data [@cook_dynamical_2018].
 Radial manual tours are performed on PC6 and PC2 in figure \ref{fig:DISclusterGood} and figure \ref{fig:DISclusterBad} respectively.
@@
 The manual tour of PC6 offers information about the dimensionality, shape, and orientations of the different experiment classes.
-whereas manipulating the contributions of PC2 only shows a subset of the dimensionality and shape information.
+Whereas manipulating the contributions of PC2 only shows a subset of the dimensionality and shape information.
 Manipulating the contributions of PC6 turned out to be much more insightful than that of PC2.
@@
 <!-- summary and segue  -->
-Tours, the dynamic linear projection of multivariate data, play an important role in data visualization; they extending the dimensionality of visuals while data- and parameter-spaces become ever larger.
-This research has modified the algorithm producing manual tours and has made this functionality available in package **spinifex**.
+Tours, which are dynamic linear projection of multivariate data, play an important role in data visualization; they extend the dimensionality of visuals while data- and parameter-spaces become ever larger.
+This research has modified the algorithm producing manual tours which and has made this functionality available in package **spinifex**.
 The package adds to **tourr**, extending the graphics offerings that can be used to display tours.
@@
 The importance of selecting the correct variable to manipulate is demonstrated by comparing tours of varying quality.
-The manual tours convey a better picture of the structure of the clustering than static linear or non-linear projections are able to.
+The manual tours convey a better picture of the structure of the clustering than static linear or non-linear projections.
 Giving the full contribution of the manipulation space to the manip var enables analysts to explore the sensitivity of the structure with the selected variable.
-This information can be used by physicists to which experiments are probing which parameter spaces, which can be indispensable for smarter planning of effort and funding.
+This information can be used by physicists to identify which experiments are probing which parameter spaces, which can be indispensable for smarter planning of effort and funding.
 
@@
 Additionally, the \CRANpkg{animation} package @xie_animation:_2018 could be implemented for another graphics framework.
-However, **animation** builds from **base** graphs while **spinifex** utilizes **ggplot2** graphics.
+However, **animation** builds from **base** graphics while **spinifex** utilizes **ggplot2** graphics, a significant paradigm shift.
 

### 2019-04-11 paper/spinifex-paper.tex
Knit of last changes

@@
 recent paper (Wang et al.
-\protect\hyperlink{ref-wang_mapping_2018}{2018}) visualizes the
+\protect\hyperlink{ref-wang_mapping_2018}{2018}) visualized the
 sensitivity of the hadronic experiments to nucleon structure.
@@
 10 principal components. This research applies manual tours to this data
-showing that manual tours resolve information about cluster structure
-and identify outliers, information that is orthogonal to the original
-static viewing plane.
+showing that manual tours resolve information about the structure of
+cluster distribution by bringing to light information that is orthogonal
+to static viewing planes.
 }
@@
 of the sequence corresponds to a small change in the projection for a
-smooth transition that perseveres continuity. This is a crucial tool
-extending the dimensionality of visualization which is a large boon as
-datasets and models exist in increasing cardinality. The package
-\CRANpkg{tourr} (Wickham et al.
+smooth transition that preserves continuity. Tours are a crucial tool
+extending the dimensionality of visualization which is increasingly
+important as datasets and models exist in increasing cardinality. The
+package \CRANpkg{tourr} (Wickham et al.
 \protect\hyperlink{ref-wickham_tourr_2011}{2011}) builds a platform for
@@
 projection. The manipulation of these coefficients allows the analyst to
-explore how sensitive the projections structure is to these changes.
-This makes manual tours particularly useful once a feature of interest
-has been identified, for example, with the use of a guided tour (Hurley
-and Buja \protect\hyperlink{ref-hurley_analyzing_1990}{1990}). Rather
-than exploring the large phase space of the data, a guided tour selects
-a very specific path via projection pursuit (Diaconis and Freedman
-\protect\hyperlink{ref-diaconis_asymptotics_1984}{1984}), the
-optimization of an index function on the projection via a hill climbing
-algorithm. This allows guided tours to rapidly identify interesting
-projection features given the relatively large parameter-space. After a
-projection of interest is identified an analyst can then use the ``finer
-brush'' of the manual tour, by controlling the contributions of
-individual variables.
+explore how sensitive the projection structure is to these
+manipulations. As manual tours operate on only one variable at a time,
+they particularly useful once a feature of interest has been identified.
+One way to identify ``interesting'' features is with the use of a guided
+tour (Hurley and Buja
+\protect\hyperlink{ref-hurley_analyzing_1990}{1990}). Guided tours
+select a very specific path, that which approachs a projection that
+optimizes an objective-function, similar to simulated annealing
+(Kirkpatrick, Gelatt, and Vecchi
+\protect\hyperlink{ref-kirkpatrick_optimization_1983}{1983}). The direct
+optimization of a function allows guided tours to rapidly identify
+interesting projection features given the relatively large
+parameter-space. After a projection of interest is identified an analyst
+can then use the ``finer brush'' of the manual tour, by controlling the
+contributions of individual variables to explore the sensitivity they
+have to the projection structure.
 
-\CRANpkg{spinifex} utilizes two new animation packages, \CRANpkg{plotly}
+This paper offers an implementation of manual tours with the package
+\CRANpkg{spinifex}. This package also exends the graphics display of
+\textbf{tourr} to the animation frameworks offered by \CRANpkg{plotly}
 (Sievert \protect\hyperlink{ref-sievert_plotly_2018}{2018}) and
 \CRANpkg{gganimate} (Pedersen and Robinson
-\protect\hyperlink{ref-pedersen_gganimate:_2019}{2019}), to display
-tours, manual or other saved tours. From a given projection, the user
-can choose which variable to control, and the animation sequence is
-generated to remove the variable from the projection, and then extend
-its contribution to be the sole variable in one direction. This allows
-the viewer to assess the change in structure induced in the projection
-by the variable's contribution.
+\protect\hyperlink{ref-pedersen_gganimate:_2019}{2019}). The algorithm
+used to perform a radial manual tour is demonstrated on a toy data set.
+Followed by an application and disscussion of the radial manual tour on
+an aggregation of high energy physics experiments in a large
+multivariate space.
 
 The paper is organized as follows. Section \ref{sec:algorithm} explains
-the algorithm using a toy dataset. Section \ref{sec:display} discussed
+the algorithm using a toy dataset. Section \ref{sec:display} discusses
 the display of the animation after the path has been generated. Section
@@
 \textbf{tourr} is used to illustrate the algorithm. The data contains 74
-observations across 6 variables, physical measurements of the flea
-beetles. Each observation belonging to one of three species.
+observations across 6 variables, which are physical measurements of the
+flea beetles. Each observation belong to one of three species.
 
@@
 feature) must be provided as an initial orientation. One way to identify
-a projection containing interesting features is to apply a guided tour
-(Cook, Swayne, and Buja
-\protect\hyperlink{ref-cook_interactive_2007}{2007}). In a guided tour
-the projection sequence is selected by optimizing an index function via
-hill-climbing on the projection space. In this case, the holes index is
-selected and applied to standardized flea data. The holes index is
-maximized when the projected observations are furthest from the center.
-Figure \ref{fig:step0} shows a locally optimized projection of the data.
-The left panel displays the reference axes of the projection basis, a
-visual indication of the magnitude and direction each variable
-contributed to the projection. The right panel shows the data as
-projected through the basis set described by the reference axes (left).
-Data points are colored and given shape according to the species (while
-the guided tour was unsupervised with this information).
+a projection containing interesting features is to apply a guided tour.
+In a guided tour, the projection sequence is selected by optimizing an
+index function via hill-climbing on the projection space. In this case,
+the holes index is selected and applied to standardized flea data. The
+holes index is maximized when the projected observations are furthest
+from the center. Figure \ref{fig:step0} shows a locally optimized
+projection of the data. The left panel displays the reference axes of
+the projection basis, a visual indication of the magnitude and direction
+each variable contributed to the projection. The right panel shows the
+data as projected through the basis set described by the reference axes
+(left). Data points are colored and given shape according to the species
+(while the guided tour was unsupervised with this information).
 
@@
 Data points color and shape are mapped to beetle species.
-Basis identified by a holes-index guided tour.
+The basis was identified by a holes-index guided tour.
 The variables aede2 and tars1 contribute mostly orthogonal to the other variables.}\label{fig:step0}
@@
 
-(ref:JetClusterBad-cap) A radial manual tour of PC4 within the jet
-cluster. Colored by experiment type: ATLAS7new in green and ATLAS7old in
-orange. Data from ATLAS7new (green) spans mostly the same space as
-ALTLAS7old (orange) with no evident difference in cluster structure
-across varying contributions of PC3. A dynamic version can be viewed at
-\url{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/}.
-
 \begin{Schunk}
@@
 
-\caption[(ref:JetClusterBad-cap)]{(ref:JetClusterBad-cap)}\label{fig:JetClusterBad}
+\caption[A radial manual tour of PC4 within the jet cluster]{A radial manual tour of PC4 within the jet cluster.
+Colored by experiment type: ATLAS7new in green and ATLAS7old in orange.
+Data from ATLAS7new (green) spans mostly the same space as ALTLAS7old (orange) with no evident difference in cluster structure across varying contributions of PC3.
+A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/jetcluster\_manualtour\_pc3/.}\label{fig:JetClusterBad}
 \end{figure}
@@
 A different space is used to explore the DIS cluster; specifically the
-first six principal components, which explain 48\% of the variation
-contained within the aggregated data(Cook, Laa, and Valencia
+first six principal components, which explains 48\% of the variation
+contained within the aggregated data (Cook, Laa, and Valencia
 \protect\hyperlink{ref-cook_dynamical_2018}{2018}). Radial manual tours
@@
 PC6 offers information about the dimensionality, shape, and orientations
-of the different experiment classes. whereas manipulating the
+of the different experiment classes. Whereas manipulating the
 contributions of PC2 only shows a subset of the dimensionality and shape
@@
 
-Tours, the dynamic linear projection of multivariate data, play an
-important role in data visualization; they extending the dimensionality
-of visuals while data- and parameter-spaces become ever larger. This
-research has modified the algorithm producing manual tours and has made
-this functionality available in package \textbf{spinifex}. The package
-adds to \textbf{tourr}, extending the graphics offerings that can be
-used to display tours.
+Tours, which are dynamic linear projection of multivariate data, play an
+important role in data visualization; they extend the dimensionality of
+visuals while data- and parameter-spaces become ever larger. This
+research has modified the algorithm producing manual tours which and has
+made this functionality available in package \textbf{spinifex}. The
+package adds to \textbf{tourr}, extending the graphics offerings that
+can be used to display tours.
 
@@
 varying quality. The manual tours convey a better picture of the
-structure of the clustering than static linear or non-linear projections
-are able to. Giving the full contribution of the manipulation space to
+structure of the clustering than static linear or non-linear
+projections. Giving the full contribution of the manipulation space to
 the manip var enables analysts to explore the sensitivity of the
 structure with the selected variable. This information can be used by
-physicists to which experiments are probing which parameter spaces,
-which can be indispensable for smarter planning of effort and funding.
+physicists to identify which experiments are probing which parameter
+spaces, which can be indispensable for smarter planning of effort and
+funding.
 
@@
 for another graphics framework. However, \textbf{animation} builds from
-\textbf{base} graphs while \textbf{spinifex} utilizes \textbf{ggplot2}
-graphics.
+\textbf{base} graphics while \textbf{spinifex} utilizes \textbf{ggplot2}
+graphics, a significant paradigm shift.
 
@@
 
-\leavevmode\hypertarget{ref-cook_interactive_2007}{}%
-Cook, Dianne, Deborah F. Swayne, and A. Buja. 2007. \emph{Interactive
-and Dynamic Graphics for Data Analysis: With R and GGobi}. Springer
-Science \& Business Media.
-
-\leavevmode\hypertarget{ref-diaconis_asymptotics_1984}{}%
-Diaconis, Persi, and David Freedman. 1984. ``Asymptotics of Graphical
-Projection Pursuit.'' \emph{The Annals of Statistics} 12 (3): 793--815.
-
 \leavevmode\hypertarget{ref-furnas_prosection_1994}{}%
@@
 Computing} 11 (6): 1193--1211. \url{https://doi.org/10.1137/0911068}.
+
+\leavevmode\hypertarget{ref-kirkpatrick_optimization_1983}{}%
+Kirkpatrick, Scott, C. Daniel Gelatt, and Mario P. Vecchi. 1983.
+``Optimization by Simulated Annealing.'' \emph{Science} 220 (4598):
+671--80.
 

### 2019-04-12 paper/spinifex-paper.tex
Section level changes. DOI's to ref. building out code/package section.

@@
 % !TeX root = RJwrapper.tex
-\title{spinifex: manual control of dynamic linear projections of
-high-dimensional data}
+\title{spinifex: A Package for Manual Control of Dynamic Linear Projections of
+Multivariate Data}
 \author{by Nicholas Spyrison, Dianne Cook}
@@
 
-\bibliography{RJrefs}
+\bibliography{spinifex_paper}
 
 \hypertarget{introduction}{%
-\subsection{Introduction}\label{introduction}}
+\section{Introduction}\label{introduction}}
 
@@
 \hypertarget{sec:algorithm}{%
-\subsection{Algorithm}\label{sec:algorithm}}
+\section{Algorithm}\label{sec:algorithm}}
 
@@
 \hypertarget{notation}{%
-\subsubsection{Notation}\label{notation}}
+\subsection{Notation}\label{notation}}
 
@@
 \hypertarget{toy-data-set}{%
-\subsubsection{Toy data set}\label{toy-data-set}}
+\subsection{Toy data set}\label{toy-data-set}}
 
@@
 \begin{Sinput}
-# devtools::install_github("nspyrison/spinifex") # Development version
+# remotes::install_github("nspyrison/spinifex") # Development version
 install.package("spinifex")
@@
 # Also see the vignette: 
-vignette("spinifex")
+vignette("spinifex_vignette")
 
@@
 \hypertarget{step-1-choose-variable-of-interest}{%
-\subsubsection{Step 1) Choose variable of
+\subsection{Step 1) Choose variable of
 interest}\label{step-1-choose-variable-of-interest}}
@@
 \hypertarget{step-2-create-the-manip-space}{%
-\subsubsection{Step 2) Create the manip
+\subsection{Step 2) Create the manip
 space}\label{step-2-create-the-manip-space}}
@@
   \textbf{e}   &\leftarrow \textbf{e} - \langle \textbf{e}, \textbf{B}_1 \rangle \textbf{B}_1 - \langle \textbf{e}, \textbf{B}_2 \rangle \textbf{B}_2 \\ 
-  &\\
   \textbf{M}_{[p,~3]} &= (\textbf{B}_1,\textbf{B}_2,\textbf{e})
@@
 \hypertarget{step-3-generate-rotation}{%
-\subsubsection{Step 3) Generate
-rotation}\label{step-3-generate-rotation}}
+\subsection{Step 3) Generate rotation}\label{step-3-generate-rotation}}
 
@@
 
-\textbf{For } \(i\) \textbf{in 1 to n\_slides:}
-
 Post-multiply the manipulation space by the pre-defined rotation matrix
-producing \textbf{RM}, the rotated manip space.
+producing \textbf{RM}, the rotated manip space. This is one frame of
+basis values, repeat this procress for each value in the sequences of
+\(\theta\) and \(\phi\) for the complete animation.
 
@@
 
-then \begin{align*}
+\textbf{For each value of } \(i\) \textbf{:}
+
+\begin{align*}
   \textbf{RM}_{[p,~3,~i]}
@@
 \hypertarget{sec:display}{%
-\subsection{Data in projection-space}\label{sec:display}}
+\section{Data in projection-space}\label{sec:display}}
 
@@
 
+\hypertarget{package-structure-and-functionality}{%
+\section{Package structure and
+functionality}\label{package-structure-and-functionality}}
+
+\hypertarget{installation}{%
+\subsection{Installation}\label{installation}}
+
+\hypertarget{walk-though-package-functions-in-terms-of-work-flow.}{%
+\subsection{Walk though package functions in terms of work
+flow.}\label{walk-though-package-functions-in-terms-of-work-flow.}}
+
 \hypertarget{rendering-and-sharing}{%
-\subsubsection{Rendering and sharing}\label{rendering-and-sharing}}
+\subsection{Rendering and sharing}\label{rendering-and-sharing}}
 
@@
 \hypertarget{storage}{%
-\subsubsection{Storage}\label{storage}}
+\subsection{Storage}\label{storage}}
 
@@
 \hypertarget{sec:application}{%
-\subsection{Application}\label{sec:application}}
+\section{Application}\label{sec:application}}
 
@@
 \hypertarget{jet-cluster}{%
-\subsubsection{Jet cluster}\label{jet-cluster}}
+\subsection{Jet cluster}\label{jet-cluster}}
 
@@
 \hypertarget{dis-cluster}{%
-\subsubsection{DIS cluster}\label{dis-cluster}}
+\subsection{DIS cluster}\label{dis-cluster}}
 
@@
 \hypertarget{summary}{%
-\subsubsection{Summary}\label{summary}}
+\subsection{Summary}\label{summary}}
 
@@
 \hypertarget{sec:discussion}{%
-\subsection{Discussion}\label{sec:discussion}}
+\section{Discussion}\label{sec:discussion}}
 
@@
 \hypertarget{acknowledgments}{%
-\subsection{Acknowledgments}\label{acknowledgments}}
+\section{Acknowledgments}\label{acknowledgments}}
 
@@
 \hypertarget{bibliography}{%
-\subsection*{Bibliography}\label{bibliography}}
-\addcontentsline{toc}{subsection}{Bibliography}
+\section*{Bibliography}\label{bibliography}}
+\addcontentsline{toc}{section}{Bibliography}
 
@@
 Statistical Computing} 6 (1): 128--43.
+\url{https://doi.org/https://doi.org/10.1137/0906011}.
 
@@
 Projections for the Visualization of PDFSense Data.'' \emph{Eur. Phys.
-J. C} 78 (9): 742.
+J. C} 78 (9): 742. \url{https://doi.org/10.1140/epjc/s10052-018-6205-2}.
 
@@
 ``Optimization by Simulated Annealing.'' \emph{Science} 220 (4598):
-671--80.
+671--80. \url{https://doi.org/10.1126/science.220.4598.671}.
 
@@
 Taxonomy.'' \emph{Biometrics}, 455--77.
+\url{https://doi.org/10.2307/2527894}.
 
@@
 Hadronic Experiments to Nucleon Structure.'' \emph{Physical Review D} 98
-(9): 094030.
+(9): 094030. \url{https://doi.org/10.1103/PhysRevD.98.094030}.
 
@@
 
-\bibliography{RJrefs.bib}
+\bibliography{spinifex\_paper.bib}
 

### 2019-04-12 paper/spinifex_paper.Rmd
Section level changes. DOI's to ref. building out code/package section.

@@
 title: > 
-  spinifex: manual control of dynamic linear projections of high-dimensional data
+  spinifex: A Package for Manual Control of Dynamic Linear Projections of Multivariate Data
 author:
@@
 toc: false
-bibliography: RJrefs.bib
+bibliography: spinifex_paper.bib
 output: rticles::rjournal_article
@@
 ---
-\bibliography{RJrefs}
+\bibliography{spinifex_paper}
 <!-- note that Di/Ursula have different setup and options and use static images.  -->
@@
 
-## Introduction
+# Introduction
 
@@
 Section \ref{sec:display} discusses the display of the animation after the path has been generated.
+Section \ref{ref:usage} explains the though the code used in the algorithm section and discusses function structure for the rest of the package.
 Section \ref{sec:application} illustrates how this can be used for sensitivity analysis applied to contemporary high energy physics.
@@
 <!-- Algorithm outline -->
-## Algorithm {#sec:algorithm}
+# Algorithm {#sec:algorithm}
 
@@
 
-The steps are described in more detail below with description, equations, figures, and their code usage.
+The steps are described in more detail below with description, equations, and figures. Function names are mentioned inline, but the code is aggregated in
 
-### Notation
+## Notation
 
@@
 
-### Toy data set
+## Toy data set
 
@@
 
-<!-- step 0 disp code. -->
-```{r step0CODEEXAMPLE, eval=F, echo=T}
-# devtools::install_github("nspyrison/spinifex") # Development version
-install.package("spinifex")
-library("spinifex")
-
-# Also see the vignette: 
-vignette("spinifex")
-
-f_data  <- tourr::rescale(flea[, 1:6])
-f_path  <- save_history(f_data, guided_tour(holes()))
-f_basis <- matrix(f_path[,, max(dim(f_path)[3])], ncol=2)
-f_cat   <- factor(flea$species)
-
-view_basis(basis = f_basis, 
-           data = f_data,
-           labels = colnames(f_data))
-```
-
-### Step 1) Choose variable of interest
+## Step 1) Choose variable of interest
 
@@
 
-### Step 2) Create the manip space
+## Step 2) Create the manip space
 
@@
 
-<!-- step 2 disp code. -->
-```{r step2CODEEXAMPLE, eval=F, echo=T}
-view_manip_space(basis = f_basis, 
-                 manip_var = 5, 
-                 labels = colnames(f_data))
-```
-
-### Step 3) Generate rotation
+## Step 3) Generate rotation
 
@@
 
-**For ** $i$ **in 1 to n_slides:**
-<!-- don't mix rmd and latex... -->
-
 Post-multiply the manipulation space by the pre-defined rotation matrix producing **RM**, the rotated manip space.
+This is one frame of basis values, repeat this procress for each value in the sequences of $\theta$ and $\phi$ for the complete animation.
 
@@
 
-then
+**For each value of ** $i$ **:**
+<!-- don't mix rmd and latex... -->
+
 \begin{align*}
@@
 
-<!-- step 3 disp code. -->
-```{r step3CODEEXAMPLE, eval=F, echo=T}
-angle_speed <- .26
-
-play_manual_tour(data = f_data,
-                 basis = f_basis, 
-                 manip_var = 5, 
-                 angle = angle_speed,
-                 col = f_cat,
-                 pch = f_cat)
-```
-
-## Data in projection-space {#sec:display}
+# Data in projection-space {#sec:display}
 
@@
 
-### Rendering and sharing
+# Package structure and functionality {#sec:usage}
+
+## Installation
+
+<!-- step 0 disp code. -->
+```{r installCODEEXAMPLE, eval=F, echo=T}
+# remotes::install_github("nspyrison/spinifex") # Development version
+install.package("spinifex")
+library("spinifex")
+
+# Also see the vignette: 
+vignette("spinifex_vignette")
+```
+
+## Algorithm
+
+We'll start by initializing values including: standardizing our data set (numeric coulmns only), provide a basis, a categorical variable for point asthetics, and specifiy the manip var.
+Then we the references axes of the basis with `view_basis()`.
+
+<!-- setup -->
+```{r step0CODEEXAMPLE, eval=F, echo=T}
+f_data  <- tourr::rescale(flea[, 1:6])                    ## standardize data
+f_path  <- save_history(f_data, guided_tour(holes()))     ## produce guided tour
+f_basis <- matrix(f_path[,, max(dim(f_path)[3])], ncol=2) ## end of guided tour
+f_cat   <- factor(flea$species)                           ## categorical var
+
+view_basis(basis = f_basis, 
+           data = f_data,
+           labels = colnames(f_data))
+```
+
+After becoming familar with the space, select a
+
+<!-- step 2 disp code. -->
+```{r step2CODEEXAMPLE, eval=F, echo=T}
+f_mvar  <- 5  ## manip var numebr
+
+view_manip_space(basis = f_basis, 
+                 manip_var = f_mvar, 
+                 labels = colnames(f_data))
+```
+
+<!-- step 3 disp code. -->
+```{r step3CODEEXAMPLE, eval=F, echo=T}
+angle_speed <- .26
+
+play_manual_tour(data = f_data,
+                 basis = f_basis, 
+                 manip_var = f_mvar, 
+                 angle = angle_speed,
+                 col = f_cat,
+                 pch = f_cat)
+```
+
+## Walk though package functions in terms of work flow.
+
+## Rendering and sharing
 
@@
 
-### Storage
+## Storage
 
@@
 
-## Application {#sec:application}
+# Application {#sec:application}
 
@@
 
-### Jet cluster
+## Jet cluster
 
@@
 
-### DIS cluster
+## DIS cluster
 
@@
 
-### Summary
+## Summary
 
@@
 
-## Discussion {#sec:discussion}
+# Discussion {#sec:discussion}
 
@@
 
-## Acknowledgments
+# Acknowledgments
 
@@
 
-## Bibliography
+# Bibliography

### 2019-04-15 paper/spinifex-paper.tex
pull the code usage into a new section, references pointing to the useage. grammerly check and knit.

@@
 \protect\hyperlink{ref-hurley_analyzing_1990}{1990}). Guided tours
-select a very specific path, that which approachs a projection that
+select a very specific path, that which approaches a projection that
 optimizes an objective-function, similar to simulated annealing
@@
 This paper offers an implementation of manual tours with the package
-\CRANpkg{spinifex}. This package also exends the graphics display of
+\CRANpkg{spinifex}. This package also extends the graphics display of
 \textbf{tourr} to the animation frameworks offered by \CRANpkg{plotly}
@@
 used to perform a radial manual tour is demonstrated on a toy data set.
-Followed by an application and disscussion of the radial manual tour on
+Followed by an application and discussion of the radial manual tour on
 an aggregation of high energy physics experiments in a large
@@
 the display of the animation after the path has been generated. Section
+\ref{sec:usage} explains the package functionality and code usage
+following the order applied in the algorithm. Section
 \ref{sec:application} illustrates how this can be used for sensitivity
@@
 \item
-  Provided with a 2D projection, choose a variable to explore. This is
-  called the ``manip'' variable.
+  Provided with a 2D projection, choose a variable to manipulate. This
+  is called the ``manip var''.
 \item
@@
 
-The steps are described in more detail below with description,
-equations, figures, and their code usage.
+The steps are described in more detail below with descriptions,
+equations, and figures are given inline. Function names are mentioned,
+but code usage is discussed in more detail in section \ref{sec:usage}.
 
@@
 observations across 6 variables, which are physical measurements of the
-flea beetles. Each observation belong to one of three species.
+flea beetles. Each observation belongs to one of three species.
 
@@
 \(\textbf{X}_{[n,~p]} ~*~ \textbf{B}_{[p,~d]} ~=~ \textbf{P}_{[n,~d]}\).
-
-\begin{Schunk}
-\begin{Sinput}
-# remotes::install_github("nspyrison/spinifex") # Development version
-install.package("spinifex")
-library("spinifex")
-
-# Also see the vignette: 
-vignette("spinifex_vignette")
-
-f_data  <- tourr::rescale(flea[, 1:6])
-f_path  <- save_history(f_data, guided_tour(holes()))
-f_basis <- matrix(f_path[,, max(dim(f_path)[3])], ncol=2)
-f_cat   <- factor(flea$species)
-
-view_basis(basis = f_basis, 
-           data = f_data,
-           labels = colnames(f_data))
-\end{Sinput}
-\end{Schunk}
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-view_manip_space(basis = f_basis, 
-                 manip_var = 5, 
-                 labels = colnames(f_data))
-\end{Sinput}
-\end{Schunk}
-
 \hypertarget{step-3-generate-rotation}{%
@@
 producing \textbf{RM}, the rotated manip space. This is one frame of
-basis values, repeat this procress for each value in the sequences of
+basis values, repeat this process for each value in the sequences of
 \(\theta\) and \(\phi\) for the complete animation.
@@
 
-\begin{Schunk}
-\begin{Sinput}
-angle_speed <- .26
-
-play_manual_tour(data = f_data,
-                 basis = f_basis, 
-                 manip_var = 5, 
-                 angle = angle_speed,
-                 col = f_cat,
-                 pch = f_cat)
-\end{Sinput}
-\end{Schunk}
-
 \hypertarget{sec:display}{%
@@
 
-\hypertarget{package-structure-and-functionality}{%
-\section{Package structure and
-functionality}\label{package-structure-and-functionality}}
+\hypertarget{sec:usage}{%
+\section{Package structure and functionality}\label{sec:usage}}
+
+This section discusses the functions and code usage of the package
+before delving into a domain-specific application of the manual tour.
 
@@
 
-\hypertarget{walk-though-package-functions-in-terms-of-work-flow.}{%
-\subsection{Walk though package functions in terms of work
-flow.}\label{walk-though-package-functions-in-terms-of-work-flow.}}
+\begin{Schunk}
+\begin{Sinput}
+# remotes::install_github("nspyrison/spinifex") # Development version
+install.package("spinifex")
+library("spinifex")
+
+# Also see the vignette: 
+vignette("spinifex_vignette")
+\end{Sinput}
+\end{Schunk}
+
+\hypertarget{primary-package-functions}{%
+\subsection{Primary package functions}\label{primary-package-functions}}
+
+The primary functions that will aid users in performing their own manual
+tours are broken into the following groupings: tour preparation, tour
+animation, and utility.
+
+\begin{Schunk}
+\begin{table}[t]
+
+\caption{\label{tab:functionsTable}Primary functions}
+\centering
+\begin{tabular}{lll}
+\toprule
+Class & Function & Description\\
+\midrule
+tour preparation & view\_basis & displays the reference frame of the basis\\
+tour preparation & view\_manip\_space & displays the manipulation space\\
+tour animation & play\_tour\_path & animates given tour path\\
+tour animation & play\_manual\_tour & animates the manual tour algorithm\\
+utility & render\_gganimate & uses gganimate to render the animation\\
+\addlinespace
+utility & render\_plotly & uses plotly to render the animation\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\end{Schunk}
+
+In the following section describes and shows usage of the tour related
+functions in parallel with the algorithm. The utility functions are
+called by the tour animation functions, passed to the arguments
+\texttt{render\_type}. Usage is demonstrated below and in the example
+documentation.
+
+\hypertarget{algorithm-code}{%
+\subsection{Algorithm code}\label{algorithm-code}}
+
+We'll start by initializing values including a standardized data set
+(numeric columns only), a starting basis, a categorical variable for
+point aesthetics, and a manip var. To get bearings on the projection,
+start by observing the reference axes of the basis with
+\texttt{view\_basis()}.
+
+\begin{Schunk}
+\begin{Sinput}
+f_data  <- tourr::rescale(flea[, 1:6])                    ## standardize data
+f_path  <- save_history(f_data, guided_tour(holes()))     ## produce guided tour
+f_basis <- matrix(f_path[,, max(dim(f_path)[3])], ncol=2) ## end of guided tour
+f_cat   <- factor(flea$species)                           ## categorical var
+
+view_basis(basis = f_basis, 
+           data = f_data,
+           labels = colnames(f_data))
+\end{Sinput}
+\end{Schunk}
+
+After becoming familiar with this space, select a manip var, the
+variable to change the contributions of. Use
+\texttt{view\_manip\_space()} View the new space with a dimension
+orthogonal to the projection plane where the manip var has a full
+contribution. This illustrates how the manip var is manipulated in a
+high space not depicted in the manual tour.
+
+\begin{Schunk}
+\begin{Sinput}
+f_mvar  <- 5  ## manip var numebr
+
+view_manip_space(basis = f_basis, 
+                 manip_var = f_mvar, 
+                 labels = colnames(f_data))
+\end{Sinput}
+\end{Schunk}
+
+Now we are ready to perform a manual tour on the selected variable. Use
+\texttt{play\_manual\_tour()} to perform the algorithm as discussed
+above, in section \ref{sec:algorithm}. The argument
+\texttt{render\_type} can be changed to \texttt{render\_gganimate} for a
+respective animation.
+
+\begin{Schunk}
+\begin{Sinput}
+angle_speed <- .26
+
+play_manual_tour(data = f_data,
+                 basis = f_basis, 
+                 manip_var = f_mvar, 
+                 angle = angle_speed,
+                 col = f_cat,
+                 pch = f_cat)
+\end{Sinput}
+\end{Schunk}
+
+Aside from the code used in the algorithm section a previously generated
+tour path can also be animated via \texttt{play\_tour\_path()}. Utility
+functions can also be passed into arguments: for example animation
+format can be changed by setting
+\texttt{render\_type\ =\ render\_gganimate}. View the guided tour path
+as a gganimate function:
+
+\begin{Schunk}
+\begin{Sinput}
+play_tour_path(tour_path = f_path,
+               data = f_data,
+               render_type = render_gganimate, 
+               angle = angle_speed
+)
+\end{Sinput}
+\end{Schunk}
 
@@
 
-Tours, which are dynamic linear projection of multivariate data, play an
-important role in data visualization; they extend the dimensionality of
-visuals while data- and parameter-spaces become ever larger. This
+Tours, which are a dynamic linear projection of multivariate data, play
+an important role in data visualization; they extend the dimensionality
+of visuals while data- and parameter-spaces become ever larger. This
 research has modified the algorithm producing manual tours which and has

### 2019-04-15 paper/spinifex_paper.Rmd
pull the code usage into a new section, references pointing to the useage. grammerly check and knit.

@@
 One way to identify "interesting" features is with the use of a guided tour [@hurley_analyzing_1990].
-Guided tours select a very specific path, that which approachs a projection that optimizes an objective-function, similar to simulated annealing [@kirkpatrick_optimization_1983].
+Guided tours select a very specific path, that which approaches a projection that optimizes an objective-function, similar to simulated annealing [@kirkpatrick_optimization_1983].
 The direct optimization of a function allows guided tours to rapidly identify interesting projection features given the relatively large parameter-space.
@@
 \CRANpkg{spinifex}.
-This package also exends the graphics display of **tourr** to the animation frameworks offered by \CRANpkg{plotly} [@sievert_plotly_2018] and \CRANpkg{gganimate} [@pedersen_gganimate:_2019].
+This package also extends the graphics display of **tourr** to the animation frameworks offered by \CRANpkg{plotly} [@sievert_plotly_2018] and \CRANpkg{gganimate} [@pedersen_gganimate:_2019].
 The algorithm used to perform a radial manual tour is demonstrated on a toy data set.
-Followed by an application and disscussion of the radial manual tour on an aggregation of high energy physics experiments in a large multivariate space.
+Followed by an application and discussion of the radial manual tour on an aggregation of high energy physics experiments in a large multivariate space.
 
@@
 Section \ref{sec:display} discusses the display of the animation after the path has been generated.
-Section \ref{ref:usage} explains the though the code used in the algorithm section and discusses function structure for the rest of the package.
+Section \ref{sec:usage} explains the package functionality and code usage following the order applied in the algorithm.
 Section \ref{sec:application} illustrates how this can be used for sensitivity analysis applied to contemporary high energy physics.
@@
 
-1. Provided with a 2D projection, choose a variable to explore. This is called the "manip" variable.
+1. Provided with a 2D projection, choose a variable to manipulate. This is called the "manip var".
 2. Create a 3D manipulation space, where the manip variable has the full contribution.
@@
 
-The steps are described in more detail below with description, equations, and figures. Function names are mentioned inline, but the code is aggregated in
+The steps are described in more detail below with descriptions, equations, and figures are given inline.
+Function names are mentioned, but code usage is discussed in more detail in section \ref{sec:usage}.
 
@@
 The data contains 74 observations across 6 variables, which are physical measurements of the flea beetles.
-Each observation belong to one of three species.
+Each observation belongs to one of three species.
 
@@
 Post-multiply the manipulation space by the pre-defined rotation matrix producing **RM**, the rotated manip space.
-This is one frame of basis values, repeat this procress for each value in the sequences of $\theta$ and $\phi$ for the complete animation.
+This is one frame of basis values, repeat this process for each value in the sequences of $\theta$ and $\phi$ for the complete animation.
 
@@
 
+This section discusses the functions and code usage of the package before delving into a domain-specific application of the manual tour.
+
 ## Installation
@@
 
-## Algorithm
+## Primary package functions 
+<!-- in terms of work flow. -->
 
-We'll start by initializing values including: standardizing our data set (numeric coulmns only), provide a basis, a categorical variable for point asthetics, and specifiy the manip var.
-Then we the references axes of the basis with `view_basis()`.
+The primary functions that will aid users in performing their own manual tours are broken into the following groupings: tour preparation, tour animation, and utility.
+
+```{r functionsTable, echo=F}
+library(kableExtra)
+
+funcs <- rbind(
+  c("tour preparation", "view_basis", "displays the reference frame of the basis"),
+  c("tour preparation", "view_manip_space", "displays the manipulation space"),
+  c("tour animation", "play_tour_path", "animates given tour path"),
+  c("tour animation", "play_manual_tour", "animates the manual tour algorithm"),
+  # c("utility", "col_of", "maps a variable to the point color"),
+  # c("utility", "pch_of", "maps a variable to the point character (shape)"),
+  c("utility", "render_gganimate", "uses gganimate to render the animation"),
+  c("utility", "render_plotly", "uses plotly to render the animation")
+)
+colnames(funcs) <- c("Class", "Function", "Description")
+
+kable(funcs, "latex", caption = "Primary functions", booktabs = T)
+```
+
+In the following section describes and shows usage of the tour related functions in parallel with the algorithm.
+The utility functions are called by the tour animation functions, passed to the arguments `render_type`.
+Usage is demonstrated below and in the example documentation.
+
+## Algorithm code
+
+We'll start by initializing values including a standardized data set (numeric columns only), a starting basis, a categorical variable for point aesthetics, and a manip var.
+To get bearings on the projection, start by observing the reference axes of the basis with `view_basis()`.
 
@@
 
-After becoming familar with the space, select a
+After becoming familiar with this space, select a manip var, the variable to change the contributions of.
+Use `view_manip_space()` View the new space with a dimension orthogonal to the projection plane where the manip var has a full contribution.
+This illustrates how the manip var is manipulated in a high space not depicted in the manual tour.
 
@@
 ```
+
+Now we are ready to perform a manual tour on the selected variable.
+Use `play_manual_tour()` to perform the algorithm as discussed above, in section  \ref{sec:algorithm}.
+The argument `render_type` can be changed to `render_gganimate` for a respective animation.
 
@@
 
-## Walk though package functions in terms of work flow.
+Aside from the code used in the algorithm section a previously generated tour path can also be animated via `play_tour_path()`.
+Utility functions can also be passed into arguments: for example animation format can be changed by setting `render_type = render_gganimate`.
+View the guided tour path as a gganimate function:
+
+```{r guidedTourCODEEXAMPLE, eval=F, echo=T}
+play_tour_path(tour_path = f_path,
+               data = f_data,
+               render_type = render_gganimate, 
+               angle = angle_speed
+)
+```
 
@@
 <!-- summary and segue  -->
-Tours, which are dynamic linear projection of multivariate data, play an important role in data visualization; they extend the dimensionality of visuals while data- and parameter-spaces become ever larger.
+Tours, which are a dynamic linear projection of multivariate data, play an important role in data visualization; they extend the dimensionality of visuals while data- and parameter-spaces become ever larger.
 This research has modified the algorithm producing manual tours which and has made this functionality available in package **spinifex**.

### 2019-04-17 paper/spinifex-paper.tex
application wording, figure dim. Grammerly, Knit, print for edit.

@@
 contributions of the manip var to a maximum and then to zero before
-returning to the initial position. Dynamic capture of user manipulation
-is typically performed directly on the projection plane (without
-depiction of the manipulation space.)
+returning to the initial position. For an oblique manual tour, capture
+the user manipulation directly on the projection plane (not the
+manipulation space.)
 
@@
 
-{\centering \includegraphics[width=6in,height=7.2in]{spinifex_paper_files/figure-latex/step3-1}
+{\centering \includegraphics[width=5.83in,height=7in]{spinifex_paper_files/figure-latex/step3-1}
 
@@
 \begin{Sinput}
-f_mvar  <- 5  ## manip var numebr
+f_mvar  <- 5  ## manip var number
 
@@
 The \textbf{tourr} package utilizes \textbf{base} graphics for the
-display of tours. \textbf{spinifex} allows tours to be used in rendered
-in \textbf{plotly} as an HTML5 object or \textbf{gganimate} as GIF or
-MP4 files. Both of which build off \textbf{ggplot2} objects in internal
+display of tours. \textbf{spinifex} allows tours to be rendered in
+\textbf{plotly} as an HTML5 object or \textbf{gganimate} as GIF or MP4
+files. Both of which build off \textbf{ggplot2} objects in internal
 functions. Sharing of animations is not trivial especially in print and
@@
 structure. The authors introduce a new tool, PDFSense, to aid in the
-visualization of Parton distribution functions (PDF). The
+visualization of parton distribution functions (PDF). The
 parameter-space of these experiments lies in 56 dimensions,
@@
 Using the same data, another study, Cook, Laa, and Valencia
-(\protect\hyperlink{ref-cook_dynamical_2018}{2018}), applies grand tours
+(\protect\hyperlink{ref-cook_dynamical_2018}{2018}), applied grand tours
 (Asimov \protect\hyperlink{ref-asimov_grand_1985}{1985}) to the same
-subspaces. Grand tours are dynamic subspace projections of high
-dimensional where frames and selected at random and linked with
-geodesically interpolation of the intermediate frames. Grand tours are
-able to better resolve the distribution shape of clusters, intra-cluster
-detail, and better outlier detection than the use of PDFSense \& TFEP
-(TensorFlow embedded projections). Before applying manual tours let's
-discuss the structure of the data.
+subspaces. Grand tours are dynamic linear projections of high
+dimensional spaces where basis sets are selected at random and animated
+with geodesic interpolation of the intermediate frames. Because of the
+change in the basis, or orientation to the subspace, tours are able to
+better resolve the distribution shape of clusters, intra-cluster detail,
+and better outlier detection than the use of PDFSense \& TFEP
+(TensorFlow embedded projections) or traditional static embeddings.
+Before applying manual tours let's discuss the structure of the data.
 
@@
 many experimental datasets which, in turn, have many observations. In
-the consideration of data density, we conduct manual tours on subsets of
-the DIS and jet clusters. This explores the sensitivity of the structure
-to each of the variables in turn and we present the subjectively best
-and worst variable to manipulate for identifying dimensionality of the
+consideration of data density, we conduct manual tours on subsets of the
+DIS and jet clusters. This explores the sensitivity of the structure to
+each of the variables in turn and we present the subjectively best and
+worst variable to manipulate for identifying dimensionality of the
 clusters and describing the span of the clusters.
@@
 \ref{fig:JetClusterGood} and figure \ref{fig:JetClusterBad}
-respectively. These components are selected to contrast the difference
-of information conveyed by touring different variables. Links to dynamic
+respectively. These variables are selected to contrast the difference of
+information conveyed by touring different variables. Links to dynamic
 HTML5 animations controlling each of the four variables are also
@@
 When manipulating PC4, there is a clear difference in the parameter
-space spanned by the experiment types ATLAS7new and ATLAS7old. Yet, when
-PC3 is manipulated there is no indication that the different experiments
-probe different parameter space. Performing a radial manual tour on PC4
-is more insightful than for PC3.
-
-Radial manual tours manipulating each of the principal components in the
-jet cluster can be viewed by following the links:
+space spanned by the experiment types ATLAS7new and ATLAS7old.
+Specifically, the variation of ATLAS7new becomes more singular. The
+experiments are probing different parameter space and PC4 is important
+to demonstrate this. Yet, when PC3 is manipulated there is no clear
+indication that the different experiments probe different parameter
+space. Performing a radial manual tour on PC4 is more insightful than
+for PC3. Radial manual tours manipulating each of the principal
+components in the jet cluster can be viewed by following the links:
 \href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc1/}{PC1},
@@
 
-{\centering \includegraphics[width=6in,height=7.2in]{spinifex_paper_files/figure-latex/JetClusterGood-1}
+{\centering \includegraphics[width=5.83in,height=7in]{spinifex_paper_files/figure-latex/JetClusterGood-1}
 
@@
 During the intermediate frames, the ATLAS7new is compressed in the direction radial to PC4.
-A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/jetcluster\_manualtour\_pc4/.}\label{fig:JetClusterGood}
+The differece in distribution shape demonstrates the experiments probe different phase-space, which has a linear mapping back to the original dimensions.
+An HTML5 version can be viewed at https://nspyrison.netlify.com/thesis/jetcluster\_manualtour\_pc4/.}\label{fig:JetClusterGood}
 \end{figure}
@@
 
-{\centering \includegraphics[width=6in,height=7.2in]{spinifex_paper_files/figure-latex/JetClusterBad-1}
+{\centering \includegraphics[width=5.83in,height=7in]{spinifex_paper_files/figure-latex/JetClusterBad-1}
 
@@
 Data from ATLAS7new (green) spans mostly the same space as ALTLAS7old (orange) with no evident difference in cluster structure across varying contributions of PC3.
-A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/jetcluster\_manualtour\_pc3/.}\label{fig:JetClusterBad}
+An HTML5 version can be viewed at https://nspyrison.netlify.com/thesis/jetcluster\_manualtour\_pc3/.}\label{fig:JetClusterBad}
 \end{figure}
@@
 
-The selection of the manip variable is important as the manipulation
+The selection of the manip variable is important, as the manipulation
 spaces convey substantially different information. The manual tour of
 PC6 offers information about the dimensionality, shape, and orientations
-of the different experiment classes. Whereas manipulating the
-contributions of PC2 only shows a subset of the dimensionality and shape
-information. Manipulating the contributions of PC6 turned out to be much
-more insightful than that of PC2. This result might seem
-counter-intuitive at first as PC2 should explain much more of the
-variation in the data. However, features and structure in the data
-regularly reside in finer detail that can be lost when looking only at
-static projections.
-
+of the different experiment classes. PC6 is particularly important to
+describe the variations of DIS HERA1+2 and charm SIDIS observations,
+Whereas manipulating the contributions of PC2 only shows a subset of the
+dimensionality and shape information. Manipulating the contributions of
+PC6 turned out to be much more insightful than that of PC2. This result
+might seem counter-intuitive at first as PC2 should explain much more of
+the variation in the data. However, features and structures in the data
+regularly reside in smaller dimensionality. The details finer resolution
+of these features can be lost when looking only at static projections.
 DIS cluster manual tours manipulating each of the principal components
@@
 
-{\centering \includegraphics[width=6in,height=7.2in]{spinifex_paper_files/figure-latex/DISclusterGood-1}
+{\centering \includegraphics[width=5.83in,height=7in]{spinifex_paper_files/figure-latex/DISclusterGood-1}
 
@@
 Points are colored by experiment type: DIS HERA1+2 in green, dimuon SIDIS in purple, and charm SIDIS in orange.
-The cluster DIS HERA1+2 (green) is distributed in a cross-shaped plane, charm SIDIS (orange) occupies the center space of this cross, with the plane projecting into the field of view when the contribution of PC6 is max.
+The cluster DIS HERA1+2 (green) is distributed in a cross-shaped plane, charm SIDIS (orange) occupies the center space of this cross.
+As the contribution of PC6 becomes whole the distributions of DIS HERA1+2 (green) and charm SIDIS (orange) become singular but offset by a small angle.
 Less evident is the linear dimuon SIDIS (purple) observations approaching the line of view for intermediate values of PC6.
-A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/discluster\_manualtour\_pc6/.}\label{fig:DISclusterGood}
+An HTML5 version can be viewed at https://nspyrison.netlify.com/thesis/discluster\_manualtour\_pc6/.}\label{fig:DISclusterGood}
 \end{figure}
@@
 
-{\centering \includegraphics[width=6in,height=7.2in]{spinifex_paper_files/figure-latex/DISclusterBad-1}
+{\centering \includegraphics[width=5.83in,height=7in]{spinifex_paper_files/figure-latex/DISclusterBad-1}
 
@@
 However, less information about the shape of DIS HERA (green) and charm SIDIS (orange) is available.
-A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/discluster\_manualtour\_pc2/.}\label{fig:DISclusterBad}
+An HTML5 version can be viewed at https://nspyrison.netlify.com/thesis/discluster\_manualtour\_pc2/.}\label{fig:DISclusterBad}
 \end{figure}

### 2019-04-17 paper/spinifex_paper.Rmd
application wording, figure dim. Grammerly, Knit, print for edit.

@@
 For a radial tour fix the (horizontal) angle within the projection plane, $\theta$, and define a sequence for the (vertical) angle coming out of the projection plane, $\phi$, bringing the initial $XY$ contributions of the manip var to a maximum and then to zero before returning to the initial position.
-Dynamic capture of user manipulation is typically performed directly on the projection plane (without depiction of the manipulation space.)
+For an oblique manual tour, capture the user manipulation directly on the projection plane (not the manipulation space.)
 
@@
 
-```{r step3, echo=F, warning=F, fig.height=7.2, fig.width=6, out.height='7.2in', out.width='6in',                                                                fig.cap = "Radial manual tour manipulating aede2 of standardized flea data.
+```{r step3, echo=F, warning=F, fig.height=7.0, fig.width=5.83, out.height='7in', out.width='5.83in',                                                                fig.cap = "Radial manual tour manipulating aede2 of standardized flea data.
 The contributions increase from its initial contribution to a full contribution to the projection before decreasing to zero and then returning to its initial value.
@@
 ```{r step2CODEEXAMPLE, eval=F, echo=T}
-f_mvar  <- 5  ## manip var numebr
+f_mvar  <- 5  ## manip var number
 
@@
 The **tourr** package utilizes **base** graphics for the display of tours.
-**spinifex** allows tours to be used in rendered in **plotly** as an HTML5 object or **gganimate** as GIF or MP4 files.
+**spinifex** allows tours to be rendered in **plotly** as an HTML5 object or **gganimate** as GIF or MP4 files.
 Both of which build off **ggplot2** objects in internal functions.
@@
 In a recent paper, @wang_mapping_2018, the authors aggregate and visualize the sensitivity of hadronic experiments to nucleon structure.
-The authors introduce a new tool, PDFSense, to aid in the visualization of Parton distribution functions (PDF).
+The authors introduce a new tool, PDFSense, to aid in the visualization of parton distribution functions (PDF).
 The parameter-space of these experiments lies in 56 dimensions, $\delta \in \mathbb{R}^{56}$, and are visualized as 3D subspaces of the 10 first principal components in linear (PCA) and non-linear (t-SNE) embeddings.
@@
 <!-- grand tours on the same data -->
-Using the same data, another study, @cook_dynamical_2018, applies grand tours [@asimov_grand_1985] to the same subspaces.
-Grand tours are dynamic subspace projections of high dimensional where frames and selected at random and linked with geodesically interpolation of the intermediate frames.
-Grand tours are able to better resolve the distribution shape of clusters, intra-cluster detail, and better outlier detection than the use of PDFSense & TFEP (TensorFlow embedded projections).
+Using the same data, another study, @cook_dynamical_2018, applied grand tours [@asimov_grand_1985] to the same subspaces.
+Grand tours are dynamic linear projections of high dimensional spaces where basis sets are selected at random and animated with geodesic interpolation of the intermediate frames.
+Because of the change in the basis, or orientation to the subspace, tours are able to better resolve the distribution shape of clusters, intra-cluster detail, and better outlier detection than the use of PDFSense & TFEP (TensorFlow embedded projections) or traditional static embeddings.
 Before applying manual tours let's discuss the structure of the data.
@@
 Each cluster is a particular class of experiments, each with many experimental datasets which, in turn, have many observations.
-In the consideration of data density, we conduct manual tours on subsets of the DIS and jet clusters.
+In consideration of data density, we conduct manual tours on subsets of the DIS and jet clusters.
 This explores the sensitivity of the structure to each of the variables in turn and we present the subjectively best and worst variable to manipulate for identifying dimensionality of the clusters and describing the span of the clusters.
@@
 Radial manual tours controlling contributions from PC4 and PC3 are shown in figure \ref{fig:JetClusterGood} and figure \ref{fig:JetClusterBad} respectively.
-These components are selected to contrast the difference of information conveyed by touring different variables.
+These variables are selected to contrast the difference of information conveyed by touring different variables.
 Links to dynamic HTML5 animations controlling each of the four variables are also provided.
@@
 When manipulating PC4, there is a clear difference in the parameter space spanned by the experiment types ATLAS7new and ATLAS7old.
-Yet, when PC3 is manipulated there is no indication that the different experiments probe different parameter space.
+Specifically, the variation of ATLAS7new becomes more singular.
+The experiments are probing different parameter space and PC4 is important to demonstrate this.
+Yet, when PC3 is manipulated there is no clear indication that the different experiments probe different parameter space.
 Performing a radial manual tour on PC4 is more insightful than for PC3.
-
 Radial manual tours manipulating each of the principal components in the jet cluster can be viewed by following the links: [PC1](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc1/), [PC2](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc2/), [PC3](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/), and [PC4](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc4/).
 
-```{r JetClusterGood, echo=F, warning=F, fig.height=7.2, fig.width=6, out.height='7.2in', out.width='6in',                                     fig.cap="A radial manual tour of PC4 within the jet cluster.
+```{r JetClusterGood, echo=F, warning=F, fig.height=7.0, fig.width=5.83, out.height='7in', out.width='5.83in',                                     fig.cap="A radial manual tour of PC4 within the jet cluster.
 Colored by experiment type: ATLAS7new in green and ATLAS7old in orange.
@@
 During the intermediate frames, the ATLAS7new is compressed in the direction radial to PC4.
-A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/jetcluster\\_manualtour\\_pc4/."}
+The differece in distribution shape demonstrates the experiments probe different phase-space, which has a linear mapping back to the original dimensions.
+An HTML5 version can be viewed at https://nspyrison.netlify.com/thesis/jetcluster\\_manualtour\\_pc4/."}
 # Jet cluster, Fig 7 of cook_dynamical_2018, subset to ATLAS7old and ATLAS7new
@@
 
-```{r JetClusterBad, echo=F, warning=F, fig.height=7.2, fig.width=6, out.height='7.2in', out.width='6in',                                       fig.cap = "A radial manual tour of PC4 within the jet cluster.
+```{r JetClusterBad, echo=F, warning=F, fig.height=7.0, fig.width=5.83, out.height='7in', out.width='5.83in',                                       fig.cap = "A radial manual tour of PC4 within the jet cluster.
 Colored by experiment type: ATLAS7new in green and ATLAS7old in orange.
 Data from ATLAS7new (green) spans mostly the same space as ALTLAS7old (orange) with no evident difference in cluster structure across varying contributions of PC3.
-A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/jetcluster\\_manualtour\\_pc3/."}
+An HTML5 version can be viewed at https://nspyrison.netlify.com/thesis/jetcluster\\_manualtour\\_pc3/."}
 # Jet cluster, Fig 7 of cook_dynalical_2018, subset to ATLAS7old and ATLAS7new
@@
 <!-- comparison of the DIS cluster -->
-The selection of the manip variable is important as the manipulation spaces convey substantially different information.
+The selection of the manip variable is important, as the manipulation spaces convey substantially different information.
 The manual tour of PC6 offers information about the dimensionality, shape, and orientations of the different experiment classes.
-Whereas manipulating the contributions of PC2 only shows a subset of the dimensionality and shape information.
+PC6 is particularly important to describe the variations of DIS HERA1+2 and charm SIDIS observations,  Whereas manipulating the contributions of PC2 only shows a subset of the dimensionality and shape information.
 Manipulating the contributions of PC6 turned out to be much more insightful than that of PC2.
 This result might seem counter-intuitive at first as PC2 should explain much more of the variation in the data.
-However, features and structure in the data regularly reside in finer detail that can be lost when looking only at static projections.
-
+However, features and structures in the data regularly reside in smaller dimensionality.
+The details finer resolution of these features can be lost when looking only at static projections.
 DIS cluster manual tours manipulating each of the principal components can be viewed from the links: [PC1](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc1/), [PC2](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc2/), [PC3](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc3/), [PC4](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc4/), [PC5](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc5/), and [PC6](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/).
 
-```{r DISclusterGood, echo=F, warning=F, fig.height=7.2, fig.width=6, out.height='7.2in', out.width='6in',                                            fig.cap = "A radial manual tour manipulating the contribution of  PC6 within the DIS cluster.
+```{r DISclusterGood, echo=F, warning=F, fig.height=7.0, fig.width=5.83, out.height='7in', out.width='5.83in',                                            fig.cap = "A radial manual tour manipulating the contribution of PC6 within the DIS cluster.
 Points are colored by experiment type: DIS HERA1+2 in green, dimuon SIDIS in purple, and charm SIDIS in orange.
-The cluster DIS HERA1+2 (green) is distributed in a cross-shaped plane, charm SIDIS (orange) occupies the center space of this cross, with the plane projecting into the field of view when the contribution of PC6 is max.
+The cluster DIS HERA1+2 (green) is distributed in a cross-shaped plane, charm SIDIS (orange) occupies the center space of this cross.
+As the contribution of PC6 becomes whole the distributions of DIS HERA1+2 (green) and charm SIDIS (orange) become singular but offset by a small angle.
 Less evident is the linear dimuon SIDIS (purple) observations approaching the line of view for intermediate values of PC6.
-A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/discluster\\_manualtour\\_pc6/."}
+An HTML5 version can be viewed at https://nspyrison.netlify.com/thesis/discluster\\_manualtour\\_pc6/."}
 
@@
 
-```{r DISclusterBad, echo=F, warning=F, fig.height=7.2, fig.width=6, out.height='7.2in', out.width='6in',                                            fig.cap = "A radial manual tour manipulating the contribution of PC2 within the DIS cluster.
+```{r DISclusterBad, echo=F, warning=F, fig.height=7.0, fig.width=5.83, out.height='7in', out.width='5.83in',                                            fig.cap = "A radial manual tour manipulating the contribution of PC2 within the DIS cluster.
 Points are colored by experiment type: DIS HERA1+2 in green, dimuon SIDIS in purple, and charm SIDIS in orange.
@@
 However, less information about the shape of DIS HERA (green) and charm SIDIS (orange) is available.
-A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/discluster\\_manualtour\\_pc2/."}
+An HTML5 version can be viewed at https://nspyrison.netlify.com/thesis/discluster\\_manualtour\\_pc2/."}
 ##DIS cluster, fig 8 of cook_dynamical_2018

### 2019-04-18 paper/spinifex-paper.tex
pg1-6 paper edits. grammerly, knit.

@@
 showing that manual tours resolve information about the structure of
-cluster distribution by bringing to light information that is orthogonal
-to static viewing planes.
+cluster distribution by bringing revealing information that is
+orthogonal to static viewing planes.
 }
@@
 
-A tour is a multivariate data analysis technique in which a sequence of
-linear (orthogonal) projections are viewed as an animation while the
-orientation of the projection basis is rotated across time. Each frame
-of the sequence corresponds to a small change in the projection for a
-smooth transition that preserves continuity. Tours are a crucial tool
-extending the dimensionality of visualization which is increasingly
-important as datasets and models exist in increasing cardinality. The
-package \CRANpkg{tourr} (Wickham et al.
+Exploring multivariate spaces is a challenging task and increasingly so
+as dimensionality increases. Traditionally, static projections display
+an embedding of multivariate data in two dimensions. Such static views
+cannot provide depth cues or convey the information of other
+orientations. Alternatively, tours are a class of dynamic linear
+projections viewed as an animation while the orientation of the
+projection is rotated. Tours are a crucial tool extending the
+dimensionality of visualization which is increasingly important as
+datasets and models exist in increasing cardinality. The package
+\CRANpkg{tourr} (Wickham et al.
 \protect\hyperlink{ref-wickham_tourr_2011}{2011}) builds a platform for
@@
 projection. The manipulation of these coefficients allows the analyst to
-explore how sensitive the projection structure is to these
-manipulations. As manual tours operate on only one variable at a time,
-they particularly useful once a feature of interest has been identified.
-One way to identify ``interesting'' features is with the use of a guided
-tour (Hurley and Buja
-\protect\hyperlink{ref-hurley_analyzing_1990}{1990}). Guided tours
-select a very specific path, that which approaches a projection that
-optimizes an objective-function, similar to simulated annealing
-(Kirkpatrick, Gelatt, and Vecchi
+explore their sensitivity to the structure within the projection. As
+manual tours operate on only one variable at a time, they particularly
+useful once a feature of interest has been identified. One way to
+identify ``interesting'' features is with the use of a guided tour
+(Hurley and Buja \protect\hyperlink{ref-hurley_analyzing_1990}{1990}).
+Guided tours select a very specific path, that which approaches a
+projection that optimizes an objective-function, in a manner similar to
+simulated annealing (Kirkpatrick, Gelatt, and Vecchi
 \protect\hyperlink{ref-kirkpatrick_optimization_1983}{1983}). The direct
@@
 contributions of individual variables to explore the sensitivity they
-have to the projection structure.
+have on the structure within the projection.
 
@@
 the algorithm using a toy dataset. Section \ref{sec:display} discusses
-the display of the animation after the path has been generated. Section
-\ref{sec:usage} explains the package functionality and code usage
-following the order applied in the algorithm. Section
+the display of the animation after the path has been generated. Package
+functionality and code usage following the order applied in the
+algorithm follows in section \ref{sec:usage}. Section
 \ref{sec:application} illustrates how this can be used for sensitivity
-analysis applied to contemporary high energy physics. The last section,
-\ref{sec:discussion} summarizes the work and discusses future research.
+analysis applied to contemporary high energy physics. Then section
+\ref{sec:summary} summarizes this paper. Lastly, section
+\ref{sec:discussion} discusses possible future research.
 
@@
 
-The steps are described in more detail below with descriptions,
-equations, and figures are given inline. Function names are mentioned,
-but code usage is discussed in more detail in section \ref{sec:usage}.
+These algorithm steps are described in more detail below with
+descriptions, equations, and figures are given inline. Function names
+are mentioned, but code usage is discussed in more detail in section
+\ref{sec:usage}.
 
@@
 \begin{itemize}
-  \item $\textbf{X}$, the data, an $n \times p$ numeric matrix to be embedded in two dimensions.
+  \item $\textbf{X}$, the data, an $n \times p$ numeric matrix to be embedded in two dimensions
   \item $\textbf{B} = (B_1,~ B_2)$, any orthonormal projection basis set, $p \times 2$ matrix, describing the projection from $p$ to two dimensions
-  \item $\textbf{e}$, a zero column vector of length $p$ with the $k-$th element set to one, where $k$ is the number of the variable to manipulate, the manip var.
-  \item $\theta$, the angle of in-projection-plane rotation, for example, on the reference axes.
-  \item $\phi$, the angle of out-of-projection-plane rotation, coming into the manipulation space.
+  \item $\textbf{e}$, a zero column vector of length $p$ with the $k-$th element set to one, where $k$ is the number of the variable to manipulate, the manip var
+  \item $\theta_i$, the sequence of angles of in-projection-plane rotation, for example, on the reference axes
+  \item $\phi_i$, the sequence of angles of out-of-projection-plane rotation, coming into the manipulation space
 \end{itemize}
@@
 
-The data is defined. A basis set (ideally orienting to an interesting
-feature) must be provided as an initial orientation. One way to identify
-a projection containing interesting features is to apply a guided tour.
-In a guided tour, the projection sequence is selected by optimizing an
-index function via hill-climbing on the projection space. In this case,
-the holes index is selected and applied to standardized flea data. The
-holes index is maximized when the projected observations are furthest
-from the center. Figure \ref{fig:step0} shows a locally optimized
-projection of the data. The left panel displays the reference axes of
-the projection basis, a visual indication of the magnitude and direction
-each variable contributed to the projection. The right panel shows the
-data as projected through the basis set described by the reference axes
-(left). Data points are colored and given shape according to the species
-(while the guided tour was unsupervised with this information).
+The data is defined. A basis set, the orientation of the projection to
+explore, must be provided. One way to identify a projection containing
+interesting features is to apply a guided tour. In a guided tour, the
+projection sequence is selected by optimizing an index function via
+hill-climbing on the projection space. In this case, the holes index is
+selected and applied to standardized flea data. The holes index is
+maximized when the projected observations are furthest from the center.
+Figure \ref{fig:step0} shows a locally optimized projection of the data.
+The left panel displays the reference axes of the projection basis, a
+visual indication of the magnitude and direction each variable
+contributes to the projection. The right panel shows the data as
+projected through the basis set described by the reference axes (left).
+Data points are colored and given shape according to the species (while
+the guided tour was unsupervised with this information).
 
@@
 (Wickham \protect\hyperlink{ref-wickham_ggplot2:_2016}{2016}) graphic
-similar to \textbackslash ref\{fig:step0). Projection space is always
-available for display via the matrix multiplication
+similar to figure \ref{fig:step0}. Projection space is always available
+for display via the matrix multiplication
 \(\textbf{X}_{[n,~p]} ~*~ \textbf{B}_{[p,~d]} ~=~ \textbf{P}_{[n,~d]}\).
@@
 variables. These two variables explain the variation of the data
-distinguishing the purple group from the rest of the sample. Select
+distinguishing the purple cluster from the rest of the sample. Select
 aede2 as the manip var, the variable to be manipulated, as it has a
@@
 being constrained to a 2D plane. Orthonormalizing rescales the new
-vector while the projection down to 2D remains the original basis. Place
-the plane horizontally, with the new dimension extending vertically,
-with axes projecting back onto the reference axes. Figure
-\ref{fig:step2} illustrates this 3D manipulation space with the manip
-var highlighted (height of the other variables are not depicted.)
+vector, while the projection down to 2D remains as the original basis.
+Place the projection plane horizontally, with the new dimension
+extending vertically, with axes projecting back onto the reference axes.
+Figure \ref{fig:step2} illustrates this 3D manipulation space with the
+manip var highlighted (height of the other variables are not depicted).
+This representation is produced by calling the
+\texttt{view\_manip\_space()} function. Functionally, the manipulation
+space is more of a mathematical construct that needn't be visualized but
+does help to digest what is happening mathematically.
 
@@
 \caption[Manipulation space for controlling the contribution of aede2 of standardized flea data]{Manipulation space for controlling the contribution of aede2 of standardized flea data.
-Basis selected by a holes-index optimized guided tour lies on the horizontal projection plane, shown in blue.
+The basis was selected by a holes-index optimized guided tour lies on the horizontal projection plane, shown in blue.
 The manip var axis, in red, extends into the vertical manipulation space, allowing the coefficients of the manip var to be changed by rotation around the origin.}\label{fig:step2}
@@
 \end{Schunk}
-
-The representation in figure \ref{fig:step2} can be produced by calling
-the function \texttt{view\_manip\_space()}.
 
@@
 returning to the initial position. For an oblique manual tour, capture
-the user manipulation directly on the projection plane (not the
-manipulation space.)
+the user manipulation directly on the \(XY\) of the projection plane.
 
@@
 \begin{itemize}
-  \item[$c_\theta$] be the cosine of $\theta$
-  \item[$c_\phi$]   be the cosine of $\phi$
-  \item[$s_\theta$] be the sine of   $\theta$
-  \item[$s_\phi$]   be the sine of   $\phi$
+  \item[$c_\theta$] be the cosine of $\theta_i$
+  \item[$c_\phi$]   be the cosine of $\phi_i$
+  \item[$s_\theta$] be the sine of   $\theta_i$
+  \item[$s_\phi$]   be the sine of   $\phi_i$
 \end{itemize}
@@
 Figure \ref{fig:step3} illustrates a sequence with 15 projected bases,
-showing the reference axes on top the corresponding projected data
-points on the below. As a result, we can see that changes in the manip
-var controlled the distance between the purple cluster and the remaining
-sample, aede2 is crucial in distinguishing this species. Tours are
-typically viewed as an animation such a dynamic version of this tour can
-be viewed online at
+showing the reference axes on the top half with the corresponding
+projected data points below. Changes in the basis coefficients of the
+manip var correspond to the distance separating the purple cluster and
+the remaining sample, aede2 is crucial in distinguishing the variation
+of species. Tours are typically viewed as an animation. The animation of
+this tour can be viewed online at
 \url{https://nspyrison.netlify.com/thesis/flea_manualtour_mvar5/}. The
-page may take a moment to load.
+page may take a moment to load. Animations can be produced using the
+function \texttt{play\_manual\_tour()}.
 
@@
 The contributions increase from its initial contribution to a full contribution to the projection before decreasing to zero and then returning to its initial value.
-The change in the projected data shows that aede2 is important for distinguishing the purple species.
+Manipulating the coefficeients of aede2 changes the distance separating the purple cluster.
+This shows that aede2 is important for distinguishing this species from the ramaining data.
 An animated version can be viewed at  https://nspyrison.netlify.com/thesis/flea\_manualtour\_mvar5/.}\label{fig:step3}
@@
 
-Animations can be produced using the function
-\texttt{play\_manual\_tour()}. This function defaults to an HTML5 widget
-produced from \textbf{plotly}. The \texttt{render\_type} argument can be
-changed to \texttt{render\_gganimate} for exporting to GIF or MP4 files.
-
 \hypertarget{sec:display}{%
@@
 
-In an appeal to performance, the above operations are performed on the
-bases without the use of the larger datasets. After the bases are
-brought into the projection-space, however, it is helpful to observe
-them with data in the same space. Pre-multiply the data by basis frame
-bringing the data into the projection space.
+In an appeal to performance, the operations of a tour are performed on
+the bases while the larger datasets are not operated on until the
+display of data points are needed. After the bases are brought into the
+projection-space, however, it is helpful to observe them with projected
+data in the same space. Pre-multiply the data by basis frame bringing
+the data into the projection space.
 
@@
 
-For a 2D scatterplot, plot the first two variables from each frame
-statically as in the previous figure, or in sequence, producing an
-animated scatterplot. The remaining variable is sometimes linked to a
-data point aesthetic (such as size or color) to produce depth cues used
-in conjunction with the \(XY\) scatterplot.
+For a 2D scatterplot, plot the first two variables within each frame.
+Show each frame in sequence to form an animation. The remaining
+manipulation space variable can be linked to point aesthetic (such as
+size or color) to produce depth cues used in conjunction with the \(XY\)
+scatterplot.
 
@@
 
-In the following section describes and shows usage of the tour related
-functions in parallel with the algorithm. The utility functions are
-called by the tour animation functions, passed to the arguments
+In the following section describes and demonstrates usage of the above
+functions following the order in the algorithm. The utility functions
+are called by the tour animation functions, passed to the arguments
 \texttt{render\_type}. Usage is demonstrated below and in the example
@@
 (numeric columns only), a starting basis, a categorical variable for
-point aesthetics, and a manip var. To get bearings on the projection,
-start by observing the reference axes of the basis with
-\texttt{view\_basis()}.
+point aesthetics (optional), and a manip var. To get bearings on the
+projection, start by observing the reference axes of the basis with
+\texttt{view\_basis()} producing figure \ref{fig:step0}.
 
@@
 variable to change the contributions of. Use
-\texttt{view\_manip\_space()} View the new space with a dimension
+\texttt{view\_manip\_space()} to view the new space with a dimension
 orthogonal to the projection plane where the manip var has a full
-contribution. This illustrates how the manip var is manipulated in a
-high space not depicted in the manual tour.
+contribution. This illustrates how the manip var is manipulated with the
+addition of the manipulation space as shown in figure \ref{fig:step2}.
 
@@
 \texttt{play\_manual\_tour()} to perform the algorithm as discussed
-above, in section \ref{sec:algorithm}. The argument
-\texttt{render\_type} can be changed to \texttt{render\_gganimate} for a
-respective animation.
+above, in section \ref{sec:algorithm}. This is the animated equivalent
+of figure \ref{fig:step3}.
 
@@
 
-Aside from the code used in the algorithm section a previously generated
-tour path can also be animated via \texttt{play\_tour\_path()}. Utility
-functions can also be passed into arguments: for example animation
-format can be changed by setting
-\texttt{render\_type\ =\ render\_gganimate}. View the guided tour path
-as a gganimate function:
+This concludes the content of the algorithm section, however, lets cover
+animating paths generated in \textbf{tourr}. Animate the previously
+generated guided tour path via \texttt{play\_tour\_path()}. Utility
+functions can also be passed as arguments into either of the tour
+animation function to change the resulting graphics object, set
+\texttt{render\_type\ =\ render\_gganimate} to view the animation as a
+GIF.
 
@@
 
-\hypertarget{summary}{%
-\subsection{Summary}\label{summary}}
+\hypertarget{sec:summary}{%
+\subsection{Summary}\label{sec:summary}}
 

### 2019-04-18 paper/spinifex_paper.Rmd
pg1-6 paper edits. grammerly, knit.

@@
 Sensitivity was characterized in non-linear 3D embeddings of the first 10 principal components.
-This research applies manual tours to this data showing that manual tours resolve information about the structure of cluster distribution by bringing to light information that is orthogonal to static viewing planes.
+This research applies manual tours to this data showing that manual tours resolve information about the structure of cluster distribution by bringing revealing information that is orthogonal to static viewing planes.
 keywords: manual tour, guided tour, grand tour, projection pursuit, high dimensional data, multivariate data, data visualization, statistical graphics, data science
@@
 <!-- Tours -->
-A tour is a multivariate data analysis technique in which a sequence of linear (orthogonal) projections are viewed as an animation while the orientation of the projection basis is rotated across time.
-Each frame of the sequence corresponds to a small change in the projection for a smooth transition that preserves continuity.
+Exploring multivariate spaces is a challenging task and increasingly so as dimensionality increases.
+Traditionally, static projections display an embedding of multivariate data in two dimensions.
+Such static views cannot provide depth cues or convey the information of other orientations.
+Alternatively, tours are a class of dynamic linear projections viewed as an animation while the orientation of the projection is rotated.
 Tours are a crucial tool extending the dimensionality of visualization which is increasingly important as datasets and models exist in increasing cardinality.
@@
 The manual tour was described in @cook_manual_1997 and allows a user to control the projection coefficients of a selected variable in a 2D projection.
-The manipulation of these coefficients allows the analyst to explore how sensitive the projection structure is to these manipulations.
+The manipulation of these coefficients allows the analyst to explore their sensitivity to the structure within the projection.
 As manual tours operate on only one variable at a time, they particularly useful once a feature of interest has been identified.
 One way to identify "interesting" features is with the use of a guided tour [@hurley_analyzing_1990].
-Guided tours select a very specific path, that which approaches a projection that optimizes an objective-function, similar to simulated annealing [@kirkpatrick_optimization_1983].
+Guided tours select a very specific path, that which approaches a projection that optimizes an objective-function, in a manner similar to simulated annealing [@kirkpatrick_optimization_1983].
 The direct optimization of a function allows guided tours to rapidly identify interesting projection features given the relatively large parameter-space.
-After a projection of interest is identified an analyst can then use the "finer brush" of the manual tour, by controlling the contributions of individual variables to explore the sensitivity they have to the projection structure.
+After a projection of interest is identified an analyst can then use the "finer brush" of the manual tour, by controlling the contributions of individual variables to explore the sensitivity they have on the structure within the projection.
 
@@
 Section \ref{sec:display} discusses the display of the animation after the path has been generated.
-Section \ref{sec:usage} explains the package functionality and code usage following the order applied in the algorithm.
+Package functionality and code usage following the order applied in the algorithm follows in section \ref{sec:usage}.
 Section \ref{sec:application} illustrates how this can be used for sensitivity analysis applied to contemporary high energy physics.
-The last section, \ref{sec:discussion} summarizes the work and discusses future research.
+Then section \ref{sec:summary} summarizes this paper.
+Lastly, section \ref{sec:discussion} discusses possible future research.
 
@@
 
-The steps are described in more detail below with descriptions, equations, and figures are given inline.
+These algorithm steps are described in more detail below with descriptions, equations, and figures are given inline.
 Function names are mentioned, but code usage is discussed in more detail in section \ref{sec:usage}.
@@
 The data is defined.
-A basis set (ideally orienting to an interesting feature) must be provided as an initial orientation.
+A basis set, the orientation of the projection to explore, must be provided.
 One way to identify a projection containing interesting features is to apply a guided tour.
@@
 Figure \ref{fig:step0} shows a locally optimized projection of the data.
-The left panel displays the reference axes of the projection basis, a visual indication of the magnitude and direction each variable contributed to the projection.
+The left panel displays the reference axes of the projection basis, a visual indication of the magnitude and direction each variable contributes to the projection.
 The right panel shows the data as projected through the basis set described by the reference axes (left).
@@
 
-Call `view_basis()` on a basis to produce a \CRANpkg{ggplot2} [@wickham_ggplot2:_2016] graphic similar to \ref{fig:step0).
+Call `view_basis()` on a basis to produce a \CRANpkg{ggplot2} [@wickham_ggplot2:_2016] graphic similar to figure \ref{fig:step0}.
 Projection space is always available for display via the matrix multiplication $\textbf{X}_{[n,~p]} ~*~ \textbf{B}_{[p,~d]} ~=~ \textbf{P}_{[n,~d]}$.
@@
 In figure \ref{fig:step0} the contributions of the variables tars1 and aede2 are mostly orthogonal to the contributions of the other four variables.
-These two variables explain the variation of the data distinguishing the purple group from the rest of the sample.
+These two variables explain the variation of the data distinguishing the purple cluster from the rest of the sample.
 Select aede2 as the manip var, the variable to be manipulated, as it has a larger contribution to the projection.
@@
 For example, the ability to lift a piece of paper, rather than being constrained to a 2D plane.
-Orthonormalizing rescales the new vector while the projection down to 2D remains the original basis.
-Place the plane horizontally, with the new dimension extending vertically, with axes projecting back onto the reference axes.
-Figure \ref{fig:step2} illustrates this 3D manipulation space with the manip var highlighted (height of the other variables are not depicted.)
+Orthonormalizing rescales the new vector, while the projection down to 2D remains as the original basis.
+Place the projection plane horizontally, with the new dimension extending vertically, with axes projecting back onto the reference axes.
+Figure \ref{fig:step2} illustrates this 3D manipulation space with the manip var highlighted (height of the other variables are not depicted).
+This representation is produced by calling the `view_manip_space()` function.
+Functionally, the manipulation space is more of a mathematical construct that needn't be visualized but does help to digest what is happening mathematically.
 
 ```{r step2, echo=F, out.width='70%',                                           fig.cap = "Manipulation space for controlling the contribution of aede2 of standardized flea data.
-Basis selected by a holes-index optimized guided tour lies on the horizontal projection plane, shown in blue.
+The basis was selected by a holes-index optimized guided tour lies on the horizontal projection plane, shown in blue.
 The manip var axis, in red, extends into the vertical manipulation space, allowing the coefficients of the manip var to be changed by rotation around the origin."}
@@
 ```
-
-The representation in figure \ref{fig:step2} can be produced by calling the function `view_manip_space()`.
 
@@
 For a radial tour fix the (horizontal) angle within the projection plane, $\theta$, and define a sequence for the (vertical) angle coming out of the projection plane, $\phi$, bringing the initial $XY$ contributions of the manip var to a maximum and then to zero before returning to the initial position.
-For an oblique manual tour, capture the user manipulation directly on the projection plane (not the manipulation space.)
+For an oblique manual tour, capture the user manipulation directly on the $XY$ of the projection plane.
 
@@
 <!-- description of slides, on the technical side we haven't gone back to data space yet. -->
-Figure \ref{fig:step3} illustrates a sequence with 15 projected bases, showing the reference axes on top the corresponding projected data points on the below.
-As a result, we can see that changes in the manip var controlled the distance between the purple cluster and the remaining sample, aede2 is crucial in distinguishing this species.
-Tours are typically viewed as an animation such a dynamic version of this tour can be viewed online at https://nspyrison.netlify.com/thesis/flea_manualtour_mvar5/.
+Figure \ref{fig:step3} illustrates a sequence with 15 projected bases, showing the reference axes on the top half with the corresponding projected data points below.
+Changes in the basis coefficients of the manip var correspond to the distance separating the purple cluster and the remaining sample, aede2 is crucial in distinguishing the variation of species.
+Tours are typically viewed as an animation.
+The animation of this tour can be viewed online at https://nspyrison.netlify.com/thesis/flea_manualtour_mvar5/.
 The page may take a moment to load.
+Animations can be produced using the function `play_manual_tour()`.
 
@@
 The contributions increase from its initial contribution to a full contribution to the projection before decreasing to zero and then returning to its initial value.
-The change in the projected data shows that aede2 is important for distinguishing the purple species.
+Manipulating the coefficeients of aede2 changes the distance separating the purple cluster.
+This shows that aede2 is important for distinguishing this species from the ramaining data.
 An animated version can be viewed at  https://nspyrison.netlify.com/thesis/flea\\_manualtour\\_mvar5/."}
@@
 
-Animations can be produced using the function `play_manual_tour()`.
-This function defaults to an HTML5 widget produced from **plotly**.
-The `render_type` argument can be changed to `render_gganimate` for exporting to GIF or MP4 files.
-
 # Data in projection-space {#sec:display}
@@
 <!-- the reminder of basis operation and now apply data -->
-In an appeal to performance, the above operations are performed on the bases without the use of the larger datasets.
-After the bases are brought into the projection-space, however, it is helpful to observe them with data in the same space.
+In an appeal to performance, the operations of a tour are performed on the bases while the larger datasets are not operated on until the display of data points are needed.
+After the bases are brought into the projection-space, however, it is helpful to observe them with projected data in the same space.
 Pre-multiply the data by basis frame bringing the data into the projection space.
@@
 <!-- plot 2 var in seq for animation -->
-For a 2D scatterplot, plot the first two variables from each frame statically as in the previous figure, or in sequence, producing an animated scatterplot.
-The remaining variable is sometimes linked to a data point aesthetic (such as size or color) to produce depth cues used in conjunction with the $XY$ scatterplot.
+For a 2D scatterplot, plot the first two variables within each frame.
+Show each frame in sequence to form an animation.
+The remaining manipulation space variable can be linked to point aesthetic (such as size or color) to produce depth cues used in conjunction with the $XY$ scatterplot.
 
@@
 
-In the following section describes and shows usage of the tour related functions in parallel with the algorithm.
+In the following section describes and demonstrates usage of the above functions following the order in the algorithm.
 The utility functions are called by the tour animation functions, passed to the arguments `render_type`.
@@
 
-We'll start by initializing values including a standardized data set (numeric columns only), a starting basis, a categorical variable for point aesthetics, and a manip var.
-To get bearings on the projection, start by observing the reference axes of the basis with `view_basis()`.
+We'll start by initializing values including a standardized data set (numeric columns only), a starting basis, a categorical variable for point aesthetics (optional), and a manip var.
+To get bearings on the projection, start by observing the reference axes of the basis with `view_basis()` producing figure \ref{fig:step0}.
 
@@
 After becoming familiar with this space, select a manip var, the variable to change the contributions of.
-Use `view_manip_space()` View the new space with a dimension orthogonal to the projection plane where the manip var has a full contribution.
-This illustrates how the manip var is manipulated in a high space not depicted in the manual tour.
+Use `view_manip_space()` to view the new space with a dimension orthogonal to the projection plane where the manip var has a full contribution.
+This illustrates how the manip var is manipulated with the addition of the manipulation space as shown in figure \ref{fig:step2}.
 
@@
 Use `play_manual_tour()` to perform the algorithm as discussed above, in section  \ref{sec:algorithm}.
-The argument `render_type` can be changed to `render_gganimate` for a respective animation.
+This is the animated equivalent of figure \ref{fig:step3}.
 
@@
 
-Aside from the code used in the algorithm section a previously generated tour path can also be animated via `play_tour_path()`.
-Utility functions can also be passed into arguments: for example animation format can be changed by setting `render_type = render_gganimate`.
-View the guided tour path as a gganimate function:
+This concludes the content of the algorithm section, however, lets cover animating paths generated in **tourr**.
+Animate the previously generated guided tour path via `play_tour_path()`.
+Utility functions can also be passed as arguments into either of the tour animation function to change the resulting graphics object, set `render_type = render_gganimate` to view the animation as a GIF.
 
@@
 
-## Summary
+## Summary {#sec:summary}
 

### 2019-04-22 paper/spinifex-paper.tex
revised title and abstrsact

@@
 % !TeX root = RJwrapper.tex
-\title{spinifex: A Package for Manual Control of Dynamic Linear Projections of
-Multivariate Data}
+\title{spinifex: A Package for Creating an Animation of a Manually Controlled
+Tour of Linear Projections of Multivariate Data, to Explore the
+Sensitivity of Structure to Specific Variables}
 \author{by Nicholas Spyrison, Dianne Cook}
@@
 \abstract{%
-The set of dynamic linear projections of multivariate data collectively
-known as \textbf{tours} provides an important tool for extending the
-dimensionality of visuals. The R package \textbf{tourr} offers a variety
-of path generators and geometric displays for conducting tours. This
-paper discusses an extension package, \textbf{spinifex}, that adds
-support for the path generation of manual tours and extends the display
-of tours to use with the animation packages, \textbf{plotly} and
-\textbf{gganimate}. Manual tours are used to explore the sensitivity of
-structure as the contributions of a manipulation variable are changed. A
-recent paper (Wang et al.
-\protect\hyperlink{ref-wang_mapping_2018}{2018}) visualized the
-sensitivity of the hadronic experiments to nucleon structure.
-Sensitivity was characterized in non-linear 3D embeddings of the first
-10 principal components. This research applies manual tours to this data
-showing that manual tours resolve information about the structure of
-cluster distribution by bringing revealing information that is
-orthogonal to static viewing planes.
+Dynamic linear projections of multivariate data collectively known as
+\textbf{tours} provide an important tool for exploring mulitvariate data
+and models. The R package \textbf{tourr} provides functions for several
+types of tours: grand, guided, little, local and frozen. Each of these
+can be viewed dynamically, or saved into a data object for animation.
+This paper describes a new package, \textbf{spinifex}, to provide a
+manual tour, that allows the coefficient for a single variable to be
+varied, with the result displayed using animation, with the packages,
+\textbf{plotly} and \textbf{gganimate}. By varying the coefficient of a
+single variable, it is possible to explore the sensitivity of structure
+in the projection to that variable. It is particularly useful when used
+with a projection pursuit guided tour to simplify and understand the
+solution. The manual tour is applied to a problem from particle physics.
 }
@@
 }
-\href{mailto:DiCook@monash.edu}{\nolinkurl{DiCook@monash.edu}}
+\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}

### 2019-04-22 paper/spinifex_paper.Rmd
revised title and abstrsact

@@
 title: > 
-  spinifex: A Package for Manual Control of Dynamic Linear Projections of Multivariate Data
+  spinifex: A Package for Creating an Animation of a Manually Controlled Tour of Linear Projections of Multivariate Data, to Explore the Sensitivity of Structure to Specific Variables
 author:
@@
     - Department of Econometrics and Business Statistics
-    email:  DiCook@monash.edu
+    email:  dicook@monash.edu
 abstract: >
-  The set of dynamic linear projections of multivariate data collectively known as **tours** provides an important tool for extending the dimensionality of visuals.
-The R package **tourr** offers a variety of path generators and geometric displays for conducting tours.
-This paper discusses an extension package, **spinifex**, that adds support for the path generation of manual tours and extends the display of tours to use with the animation packages, **plotly** and **gganimate**.
-Manual tours are used to explore the sensitivity of structure as the contributions of a manipulation variable are changed.
+  Dynamic linear projections of multivariate data collectively known as **tours** provide an important tool for exploring mulitvariate data and models.
+The R package **tourr** provides functions for several types of tours: grand, guided, little, local and frozen.
+Each of these can be viewed dynamically, or saved into a data object for animation.
+This paper describes a new package, **spinifex**, to provide a manual tour, that allows the coefficient for a single variable to be varied, with the result displayed using animation, with the packages, **plotly** and **gganimate**.
+By varying the coefficient of a single variable, it is possible to explore the sensitivity of structure in the projection to that variable.
+It is particularly useful when used with a projection pursuit guided tour to simplify and understand the solution.
+The manual tour is applied to a problem from particle physics.
 
-  A recent paper [@wang_mapping_2018] visualized the sensitivity of the hadronic experiments to nucleon structure.
-Sensitivity was characterized in non-linear 3D embeddings of the first 10 principal components.
-This research applies manual tours to this data showing that manual tours resolve information about the structure of cluster distribution by bringing revealing information that is orthogonal to static viewing planes.
-keywords: manual tour, guided tour, grand tour, projection pursuit, high dimensional data, multivariate data, data visualization, statistical graphics, data science
+keywords: exploratory data anlaysis, projection pursuit, high dimensional data, data visualization, statistical graphics, data science
 preamble: >

### 2019-04-22 paper/spinifex-paper.tex
introduction revised

@@
 % !TeX root = RJwrapper.tex
-\title{spinifex: A Package for Creating an Animation of a Manually Controlled
-Tour of Linear Projections of Multivariate Data, to Explore the
-Sensitivity of Structure to Specific Variables}
+\title{spinifex: An R Package for Creating a Manual Tour of Low-dimensional
+Projections of Multivariate Data}
 \author{by Nicholas Spyrison, Dianne Cook}
@@
 \abstract{%
-Dynamic linear projections of multivariate data collectively known as
-\textbf{tours} provide an important tool for exploring mulitvariate data
-and models. The R package \textbf{tourr} provides functions for several
-types of tours: grand, guided, little, local and frozen. Each of these
-can be viewed dynamically, or saved into a data object for animation.
-This paper describes a new package, \textbf{spinifex}, to provide a
-manual tour, that allows the coefficient for a single variable to be
-varied, with the result displayed using animation, with the packages,
-\textbf{plotly} and \textbf{gganimate}. By varying the coefficient of a
-single variable, it is possible to explore the sensitivity of structure
-in the projection to that variable. It is particularly useful when used
-with a projection pursuit guided tour to simplify and understand the
-solution. The manual tour is applied to a problem from particle physics.
+Dynamic low-dimensional linear projections of multivariate data
+collectively known as \textbf{tours} provide an important tool for
+exploring mulitvariate data and models. The R package \textbf{tourr}
+provides functions for several types of tours: grand, guided, little,
+local and frozen. Each of these can be viewed dynamically, or saved into
+a data object for animation. This paper describes a new package,
+\textbf{spinifex}, to provide a manual tour, that allows the coefficient
+for a single variable to be controlled. The variable is rotated fully
+into the projection with a coefficient of 1 or -1, or completely out of
+the projection with a coefficient of 0. The resulting sequence of
+projections can be displayed using animation, with functions from either
+the \textbf{plotly} and \textbf{gganimate} packages. By varying the
+coefficient of a single variable, it is possible to explore the
+sensitivity of structure in the projection to that variable. It is
+particularly useful when used with a projection pursuit guided tour to
+simplify and understand the solution. The use of the manual tour is
+illustrated using a problem from particle physics.
 }
@@
 
-Exploring multivariate spaces is a challenging task and increasingly so
-as dimensionality increases. Traditionally, static projections display
-an embedding of multivariate data in two dimensions. Such static views
-cannot provide depth cues or convey the information of other
-orientations. Alternatively, tours are a class of dynamic linear
-projections viewed as an animation while the orientation of the
-projection is rotated. Tours are a crucial tool extending the
-dimensionality of visualization which is increasingly important as
-datasets and models exist in increasing cardinality. The package
-\CRANpkg{tourr} (Wickham et al.
-\protect\hyperlink{ref-wickham_tourr_2011}{2011}) builds a platform for
-generating tour paths and applying to a wide number of display
-geometrics in \textbf{base} graphics.
+Exploring multivariate spaces is a challenging task, increasingly so as
+dimensionality increases. Traditionally, static low-dimensional
+projections are used to display multivariate data in two dimensions,
+such as principal component analysis, linear discriminant spaces or
+projection pursuit. These are useful for finding relationships between
+multiple variables, but they are limited because they show only a
+glimpse of the high-dimensional space. An alternative approach, is to
+use a tour of dynamic linear projections, to look at many different
+low-dimensional proections. Tours can be considered to extend the
+dimensionality of visualization, which is important as data and models
+exist in more than 3D. The package \CRANpkg{tourr} (Wickham et al.
+\protect\hyperlink{ref-wickham_tourr_2011}{2011}) provides a platform
+for generating tours. It has many types of tours available, and many
+types of displays possible. A user can make a grand, guided, little,
+local or frozen tour, and display the resulting projected data as a
+scatterplot, density plot, histogram, or even as Chernoff faces if the
+projection dimension is more than 3.
 
-While there are numerous methods that generate tour paths, this research
-focuses on the manual tour. The manual tour was described in Cook and
-Buja (\protect\hyperlink{ref-cook_manual_1997}{1997}) and allows a user
-to control the projection coefficients of a selected variable in a 2D
+This work adds a manual tour to the collection. The manual tour was
+described in Cook and Buja
+(\protect\hyperlink{ref-cook_manual_1997}{1997}) and allows a user to
+control the projection coefficients of a selected variable in a 2D
 projection. The manipulation of these coefficients allows the analyst to
 explore their sensitivity to the structure within the projection. As
-manual tours operate on only one variable at a time, they particularly
-useful once a feature of interest has been identified. One way to
-identify ``interesting'' features is with the use of a guided tour
-(Hurley and Buja \protect\hyperlink{ref-hurley_analyzing_1990}{1990}).
-Guided tours select a very specific path, that which approaches a
-projection that optimizes an objective-function, in a manner similar to
-simulated annealing (Kirkpatrick, Gelatt, and Vecchi
+manual tours operate on only one variable at a time, they are
+particularly useful once a feature of interest has been identified.
+
+One way to identify ``interesting'' features is with the use of a guided
+tour (Cook et al. \protect\hyperlink{ref-cook_grand_1995}{1995}). Guided
+tours select a very specific path, that which approaches a projection
+that optimizes an objective-function. The optimization is conducted in a
+manner similar to simulated annealing (Kirkpatrick, Gelatt, and Vecchi
 \protect\hyperlink{ref-kirkpatrick_optimization_1983}{1983}). The direct
@@
 contributions of individual variables to explore the sensitivity they
-have on the structure within the projection.
+have on the structure visible in the projection.
 
-This paper offers an implementation of manual tours with the package
-\CRANpkg{spinifex}. This package also extends the graphics display of
-\textbf{tourr} to the animation frameworks offered by \CRANpkg{plotly}
-(Sievert \protect\hyperlink{ref-sievert_plotly_2018}{2018}) and
+The paper is organized as follows. Section \ref{sec:algorithm} describes
+the algorithm used to perform a radial manual tour as implemented in the
+package \CRANpkg{spinifex}. Section \ref{sec:display} explains how to
+generate an animation of the manual tour using the animation frameworks
+offered by \CRANpkg{plotly} (Sievert
+\protect\hyperlink{ref-sievert_plotly_2018}{2018}) and
 \CRANpkg{gganimate} (Pedersen and Robinson
-\protect\hyperlink{ref-pedersen_gganimate:_2019}{2019}). The algorithm
-used to perform a radial manual tour is demonstrated on a toy data set.
-Followed by an application and discussion of the radial manual tour on
-an aggregation of high energy physics experiments in a large
-multivariate space.
-
-The paper is organized as follows. Section \ref{sec:algorithm} explains
-the algorithm using a toy dataset. Section \ref{sec:display} discusses
-the display of the animation after the path has been generated. Package
+\protect\hyperlink{ref-pedersen_gganimate:_2019}{2019}). Package
 functionality and code usage following the order applied in the
@@
 \ref{sec:application} illustrates how this can be used for sensitivity
-analysis applied to contemporary high energy physics. Then section
-\ref{sec:summary} summarizes this paper. Lastly, section
-\ref{sec:discussion} discusses possible future research.
+analysis applied to multivariate data collected on high energy physics
+experiments (Wang et al.
+\protect\hyperlink{ref-wang_mapping_2018}{2018}). Section
+\ref{sec:discussion} summarizes this paper and discusses potential
+future directions.
 
@@
 
-\hypertarget{sec:summary}{%
-\subsection{Summary}\label{sec:summary}}
+\hypertarget{sec:discussion}{%
+\subsection{Discussion}\label{sec:discussion}}
 
@@
 funding.
-
-\hypertarget{sec:discussion}{%
-\section{Discussion}\label{sec:discussion}}
 
@@
 
+\leavevmode\hypertarget{ref-cook_grand_1995}{}%
+Cook, Dianne, Andreas Buja, Javier Cabrera, and Catherine Hurley. 1995.
+``Grand Tour and Projection Pursuit.'' \emph{Journal of Computational
+and Graphical Statistics} 4 (3): 155.
+\url{https://doi.org/10.2307/1390844}.
+
 \leavevmode\hypertarget{ref-cook_dynamical_2018}{}%
@@
 \url{https://doi.org/10.2307/1390897}.
-
-\leavevmode\hypertarget{ref-hurley_analyzing_1990}{}%
-Hurley, C., and A. Buja. 1990. ``Analyzing High-Dimensional Data with
-Motion Graphics.'' \emph{SIAM Journal on Scientific and Statistical
-Computing} 11 (6): 1193--1211. \url{https://doi.org/10.1137/0911068}.
 

### 2019-04-22 paper/spinifex_paper.Rmd
introduction revised

@@
 title: > 
-  spinifex: A Package for Creating an Animation of a Manually Controlled Tour of Linear Projections of Multivariate Data, to Explore the Sensitivity of Structure to Specific Variables
+  spinifex: An R Package for Creating a Manual Tour of Low-dimensional  Projections of Multivariate Data
 author:
@@
 abstract: >
-  Dynamic linear projections of multivariate data collectively known as **tours** provide an important tool for exploring mulitvariate data and models.
+  Dynamic low-dimensional linear projections of multivariate data collectively known as **tours** provide an important tool for exploring mulitvariate data and models.
 The R package **tourr** provides functions for several types of tours: grand, guided, little, local and frozen.
 Each of these can be viewed dynamically, or saved into a data object for animation.
-This paper describes a new package, **spinifex**, to provide a manual tour, that allows the coefficient for a single variable to be varied, with the result displayed using animation, with the packages, **plotly** and **gganimate**.
+This paper describes a new package, **spinifex**, to provide a manual tour, that allows the coefficient for a single variable to be controlled.
+The variable is rotated fully into the projection with a coefficient of 1 or -1, or completely out of the projection with a coefficient of 0.
+The resulting sequence of projections can be displayed using animation, with functions from either the **plotly** and **gganimate** packages.
 By varying the coefficient of a single variable, it is possible to explore the sensitivity of structure in the projection to that variable.
 It is particularly useful when used with a projection pursuit guided tour to simplify and understand the solution.
-The manual tour is applied to a problem from particle physics.
+The use of the manual tour is illustrated using a problem from particle physics.
 
@@
 <!-- Tours -->
-Exploring multivariate spaces is a challenging task and increasingly so as dimensionality increases.
-Traditionally, static projections display an embedding of multivariate data in two dimensions.
-Such static views cannot provide depth cues or convey the information of other orientations.
-Alternatively, tours are a class of dynamic linear projections viewed as an animation while the orientation of the projection is rotated.
-Tours are a crucial tool extending the dimensionality of visualization which is increasingly important as datasets and models exist in increasing cardinality.
-The package \CRANpkg{tourr} [@wickham_tourr_2011] builds a platform for generating tour paths and applying to a wide number of display geometrics in **base** graphics.
+Exploring multivariate spaces is a challenging task, increasingly so as dimensionality increases.
+Traditionally, static low-dimensional projections are used to display multivariate data in two dimensions, such as principal component analysis, linear discriminant spaces or projection pursuit.
+These are useful for finding relationships between multiple variables, but they are limited because they show only a glimpse of the high-dimensional space.
+An alternative approach, is to use a tour of dynamic linear projections, to look at many different low-dimensional proections.
+Tours can be considered to extend the dimensionality of visualization, which is important as data and models exist in more than 3D.
+The package \CRANpkg{tourr} [@wickham_tourr_2011] provides a platform for generating tours.
+It has many types of tours available, and many types of displays possible.
+A user can make a grand, guided, little, local or frozen tour, and display the resulting projected data as a scatterplot, density plot, histogram, or even as Chernoff faces if the projection dimension is more than 3.
 
 <!-- Manual tour -->
-While there are numerous methods that generate tour paths, this research focuses on the manual tour.
+This work adds a manual tour to the collection.
 The manual tour was described in @cook_manual_1997 and allows a user to control the projection coefficients of a selected variable in a 2D projection.
 The manipulation of these coefficients allows the analyst to explore their sensitivity to the structure within the projection.
-As manual tours operate on only one variable at a time, they particularly useful once a feature of interest has been identified.
-One way to identify "interesting" features is with the use of a guided tour [@hurley_analyzing_1990].
-Guided tours select a very specific path, that which approaches a projection that optimizes an objective-function, in a manner similar to simulated annealing [@kirkpatrick_optimization_1983].
+As manual tours operate on only one variable at a time, they are particularly useful once a feature of interest has been identified.
+
+One way to identify "interesting" features is with the use of a guided tour [@cook_grand_1995].
+Guided tours select a very specific path, that which approaches a projection that optimizes an objective-function.
+The optimization is conducted in a manner similar to simulated annealing [@kirkpatrick_optimization_1983].
 The direct optimization of a function allows guided tours to rapidly identify interesting projection features given the relatively large parameter-space.
-After a projection of interest is identified an analyst can then use the "finer brush" of the manual tour, by controlling the contributions of individual variables to explore the sensitivity they have on the structure within the projection.
+After a projection of interest is identified an analyst can then use the "finer brush" of the manual tour, by controlling the contributions of individual variables to explore the sensitivity they have on the structure visible in the projection.
 
-<!-- Contributions -->
-This paper offers an implementation of manual tours with the package
-\CRANpkg{spinifex}.
-This package also extends the graphics display of **tourr** to the animation frameworks offered by \CRANpkg{plotly} [@sievert_plotly_2018] and \CRANpkg{gganimate} [@pedersen_gganimate:_2019].
-The algorithm used to perform a radial manual tour is demonstrated on a toy data set.
-Followed by an application and discussion of the radial manual tour on an aggregation of high energy physics experiments in a large multivariate space.
+<!-- Contributions 
+This paper describes the algorithm used to perform a radial manual tour as implemented in the package \CRANpkg{spinifex}.
+It also describes how to generate an animation of the manual tour using the animation frameworks offered by \CRANpkg{plotly} [@sievert_plotly_2018] and \CRANpkg{gganimate} [@pedersen_gganimate:_2019].
+The method is applied to multivariate data collected on high energy physics experiments (REF).
+-->
 
@@
 The paper is organized as follows.
-Section \ref{sec:algorithm} explains the algorithm using a toy dataset.
-Section \ref{sec:display} discusses the display of the animation after the path has been generated.
+Section \ref{sec:algorithm} describes the algorithm used to perform a radial manual tour as implemented in the package \CRANpkg{spinifex}.
+Section \ref{sec:display} explains how to generate an animation of the manual tour using the animation frameworks offered by \CRANpkg{plotly} [@sievert_plotly_2018] and \CRANpkg{gganimate} [@pedersen_gganimate:_2019].
 Package functionality and code usage following the order applied in the algorithm follows in section \ref{sec:usage}.
-Section \ref{sec:application} illustrates how this can be used for sensitivity analysis applied to contemporary high energy physics.
-Then section \ref{sec:summary} summarizes this paper.
-Lastly, section \ref{sec:discussion} discusses possible future research.
+Section \ref{sec:application} illustrates how this can be used for sensitivity analysis applied to multivariate data collected on high energy physics experiments [@wang_mapping_2018].
+Section \ref{sec:discussion} summarizes this paper and discusses potential future directions.
 
@@
 
-## Summary {#sec:summary}
+## Discussion {#sec:discussion}
 
@@
 This information can be used by physicists to identify which experiments are probing which parameter spaces, which can be indispensable for smarter planning of effort and funding.
-
-# Discussion {#sec:discussion}
 

### 2019-04-22 paper/spinifex-paper.tex
first two sections of algorithm section

@@
 
-The section below describes the algorithm for performing a 2D radial
-manual tour:
+This section describes the algorithm for performing a 2D radial manual
+tour, containing these steps:
 
@@
 
-These algorithm steps are described in more detail below with
-descriptions, equations, and figures are given inline. Function names
-are mentioned, but code usage is discussed in more detail in section
+More detail is provided in the following sections. Function names are
+provided, but code structure and usage is discussed in section
 \ref{sec:usage}.
@@
 \begin{itemize}
-  \item $\textbf{X}$, the data, an $n \times p$ numeric matrix to be embedded in two dimensions
-  \item $\textbf{B} = (B_1,~ B_2)$, any orthonormal projection basis set, $p \times 2$ matrix, describing the projection from $p$ to two dimensions
-  \item $\textbf{e}$, a zero column vector of length $p$ with the $k-$th element set to one, where $k$ is the number of the variable to manipulate, the manip var
-  \item $\theta_i$, the sequence of angles of in-projection-plane rotation, for example, on the reference axes
-  \item $\phi_i$, the sequence of angles of out-of-projection-plane rotation, coming into the manipulation space
+\tightlist
+\item
+  \(\textbf{X}\), the data, an \(n \times p\) numeric matrix to be
+  projected to 2D.
+\item
+  \(\textbf{B} = (B_1,~ B_2)\), any 2D orthonormal projection basis,
+  \(p \times 2\) matrix, describing the projection from \(p\) to two
+  dimensions
+\item
+  \(\textbf{e}\), a 1D coordinate basis vector of length \(p\) with the
+  \(k-\)th element set to one, where \(k\) is the index of the variable
+  to manipulate, called the ``manip var''.
+\item
+  \(\theta_i\), the sequence of angles of in-projection-plane rotation,
+  for example, on the reference axes.
+\item
+  \(\phi_i\), the sequence of angles of out-of-projection-plane
+  rotation, coming into the manipulation space.
 \end{itemize}
@@
 
-\hypertarget{toy-data-set}{%
-\subsection{Toy data set}\label{toy-data-set}}
+\hypertarget{set-up}{%
+\subsection{Set up}\label{set-up}}
 
@@
 
-The data is defined. A basis set, the orientation of the projection to
-explore, must be provided. One way to identify a projection containing
-interesting features is to apply a guided tour. In a guided tour, the
-projection sequence is selected by optimizing an index function via
-hill-climbing on the projection space. In this case, the holes index is
-selected and applied to standardized flea data. The holes index is
-maximized when the projected observations are furthest from the center.
-Figure \ref{fig:step0} shows a locally optimized projection of the data.
-The left panel displays the reference axes of the projection basis, a
-visual indication of the magnitude and direction each variable
-contributes to the projection. The right panel shows the data as
-projected through the basis set described by the reference axes (left).
-Data points are colored and given shape according to the species (while
-the guided tour was unsupervised with this information).
+An initial 2D projection basis set must be provided. One way to identify
+a projection containing interesting features is to apply a guided tour.
+In a guided tour, the projection sequence is selected by optimizing an
+index function via hill-climbing on the projection space. In this case,
+the holes index is selected and applied to standardized flea data. The
+holes index is maximized when the projected observations are furthest
+from the center. Figure \ref{fig:step0} shows a locally optimized
+projection of the data. The left panel displays the reference axes of
+the projection basis, a visual indication of the magnitude and direction
+each variable contributes to the projection. The right panel shows the
+data as projected through the basis set described by the reference axes
+(left). Data points are colored and given shape according to the
+species.
+
+Call \texttt{view\_basis()} on a basis to produce a \CRANpkg{ggplot2}
+(Wickham \protect\hyperlink{ref-wickham_ggplot2:_2016}{2016}) graphic
+similar to Figure \ref{fig:step0}. The projected data is always
+available for display via matrix multiplication
+\(\textbf{X}_{[n,~p]} ~*~ \textbf{B}_{[p,~d]} ~=~ \textbf{P}_{[n,~d]}\).
 
@@
 
-\caption[Basis reference axes (left) and projected data (right) of standardized flea data]{Basis reference axes (left) and projected data (right) of standardized flea data.
-Data points color and shape are mapped to beetle species.
-The basis was identified by a holes-index guided tour.
-The variables aede2 and tars1 contribute mostly orthogonal to the other variables.}\label{fig:step0}
+\caption[Initial 2D projection]{Initial 2D projection: representation of the basis  (left) and resulting data projection (right) of standardized flea data.
+The color and shape of data points are mapped to beetle species.
+The basis was identified using a projection pursuit guided tour, with the holes index.
+The contribution of the variables aede2 and tars1 approximately contrasts the other variables.
+The visible structure in the projection are the three clusters corresponding to the three species.}\label{fig:step0}
 \end{figure}
 \end{Schunk}
-
-Call \texttt{view\_basis()} on a basis to produce a \CRANpkg{ggplot2}
-(Wickham \protect\hyperlink{ref-wickham_ggplot2:_2016}{2016}) graphic
-similar to figure \ref{fig:step0}. Projection space is always available
-for display via the matrix multiplication
-\(\textbf{X}_{[n,~p]} ~*~ \textbf{B}_{[p,~d]} ~=~ \textbf{P}_{[n,~d]}\).
 

### 2019-04-22 paper/spinifex_paper.Rmd
first two sections of algorithm section

@@
 
-The section below describes the algorithm for performing a 2D radial manual tour:
+This section describes the algorithm for performing a 2D radial manual tour, containing these steps:
 
@@
 
-These algorithm steps are described in more detail below with descriptions, equations, and figures are given inline.
-Function names are mentioned, but code usage is discussed in more detail in section \ref{sec:usage}.
+More detail is provided in the following sections. Function names are provided, but code structure and usage is discussed in section \ref{sec:usage}.
 
@@
 <!-- remember mixing latex and rmarkdown bullets isn't trivial. -->
-\begin{itemize}
-  \item $\textbf{X}$, the data, an $n \times p$ numeric matrix to be embedded in two dimensions
-  \item $\textbf{B} = (B_1,~ B_2)$, any orthonormal projection basis set, $p \times 2$ matrix, describing the projection from $p$ to two dimensions
-  \item $\textbf{e}$, a zero column vector of length $p$ with the $k-$th element set to one, where $k$ is the number of the variable to manipulate, the manip var
-  \item $\theta_i$, the sequence of angles of in-projection-plane rotation, for example, on the reference axes
-  \item $\phi_i$, the sequence of angles of out-of-projection-plane rotation, coming into the manipulation space
-\end{itemize}
+- $\textbf{X}$, the data, an $n \times p$ numeric matrix to be projected to 2D.
+- $\textbf{B} = (B_1,~ B_2)$, any 2D orthonormal projection basis, $p \times 2$ matrix, describing the projection from $p$ to two dimensions
+- $\textbf{e}$, a 1D coordinate basis vector of length $p$ with the $k-$th element set to one, where $k$ is the index of the variable to manipulate, called the "manip var".
+- $\theta_i$, the sequence of angles of in-projection-plane rotation, for example, on the reference axes.
+- $\phi_i$, the sequence of angles of out-of-projection-plane rotation, coming into the manipulation space.
 
@@
 
-## Toy data set
+## Set up
 
@@
 <!-- Projection basis -->
-The data is defined.
-A basis set, the orientation of the projection to explore, must be provided.
+An initial 2D projection basis set must be provided.
 One way to identify a projection containing interesting features is to apply a guided tour.
@@
 The right panel shows the data as projected through the basis set described by the reference axes (left).
-Data points are colored and given shape according to the species (while the guided tour was unsupervised with this information).
+Data points are colored and given shape according to the species.
 
-```{r step0, echo=F, results='hide', out.width='70%',                             fig.cap = "Basis reference axes (left) and projected data (right) of standardized flea data.
-Data points color and shape are mapped to beetle species.
-The basis was identified by a holes-index guided tour.
-The variables aede2 and tars1 contribute mostly orthogonal to the other variables."}
+Call `view_basis()` on a basis to produce a \CRANpkg{ggplot2} [@wickham_ggplot2:_2016] graphic similar to Figure \ref{fig:step0}.
+The projected data is always available for display via matrix multiplication $\textbf{X}_{[n,~p]} ~*~ \textbf{B}_{[p,~d]} ~=~ \textbf{P}_{[n,~d]}$.
+
+```{r step0, echo=F, results='hide', out.width='70%',                             fig.cap = "Initial 2D projection: representation of the basis  (left) and resulting data projection (right) of standardized flea data.
+The color and shape of data points are mapped to beetle species.
+The basis was identified using a projection pursuit guided tour, with the holes index.
+The contribution of the variables aede2 and tars1 approximately contrasts the other variables.
+The visible structure in the projection are the three clusters corresponding to the three species."}
 ## Flea holes tour
@@
 ```
-
-Call `view_basis()` on a basis to produce a \CRANpkg{ggplot2} [@wickham_ggplot2:_2016] graphic similar to figure \ref{fig:step0}.
-Projection space is always available for display via the matrix multiplication $\textbf{X}_{[n,~p]} ~*~ \textbf{B}_{[p,~d]} ~=~ \textbf{P}_{[n,~d]}$.
 

### 2019-04-22 paper/spinifex-paper.tex
capitalised figure

@@
 
-In figure \ref{fig:step0} the contributions of the variables tars1 and
+In Figure \ref{fig:step0} the contributions of the variables tars1 and
 aede2 are mostly orthogonal to the contributions of the other four
@@
 projection, start by observing the reference axes of the basis with
-\texttt{view\_basis()} producing figure \ref{fig:step0}.
+\texttt{view\_basis()} producing Figure \ref{fig:step0}.
 
@@
 contribution. This illustrates how the manip var is manipulated with the
-addition of the manipulation space as shown in figure \ref{fig:step2}.
+addition of the manipulation space as shown in Figure \ref{fig:step2}.
 
@@
 above, in section \ref{sec:algorithm}. This is the animated equivalent
-of figure \ref{fig:step3}.
+of Figure \ref{fig:step3}.
 
@@
 different parts of the subspace. Radial manual tours controlling
-contributions from PC4 and PC3 are shown in figure
-\ref{fig:JetClusterGood} and figure \ref{fig:JetClusterBad}
+contributions from PC4 and PC3 are shown in Figure
+\ref{fig:JetClusterGood} and Figure \ref{fig:JetClusterBad}
 respectively. These variables are selected to contrast the difference of
@@
 \protect\hyperlink{ref-cook_dynamical_2018}{2018}). Radial manual tours
-are performed on PC6 and PC2 in figure \ref{fig:DISclusterGood} and
-figure \ref{fig:DISclusterBad} respectively.
+are performed on PC6 and PC2 in Figure \ref{fig:DISclusterGood} and
+Figure \ref{fig:DISclusterBad} respectively.
 

### 2019-04-22 paper/spinifex_paper.Rmd
capitalised figure

@@
 <!-- select a manip var-->
-In figure \ref{fig:step0} the contributions of the variables tars1 and aede2 are mostly orthogonal to the contributions of the other four variables.
+In Figure \ref{fig:step0} the contributions of the variables tars1 and aede2 are mostly orthogonal to the contributions of the other four variables.
 These two variables explain the variation of the data distinguishing the purple cluster from the rest of the sample.
@@
 We'll start by initializing values including a standardized data set (numeric columns only), a starting basis, a categorical variable for point aesthetics (optional), and a manip var.
-To get bearings on the projection, start by observing the reference axes of the basis with `view_basis()` producing figure \ref{fig:step0}.
+To get bearings on the projection, start by observing the reference axes of the basis with `view_basis()` producing Figure \ref{fig:step0}.
 
@@
 Use `view_manip_space()` to view the new space with a dimension orthogonal to the projection plane where the manip var has a full contribution.
-This illustrates how the manip var is manipulated with the addition of the manipulation space as shown in figure \ref{fig:step2}.
+This illustrates how the manip var is manipulated with the addition of the manipulation space as shown in Figure \ref{fig:step2}.
 
@@
 Use `play_manual_tour()` to perform the algorithm as discussed above, in section  \ref{sec:algorithm}.
-This is the animated equivalent of figure \ref{fig:step3}.
+This is the animated equivalent of Figure \ref{fig:step3}.
 
@@
 The data within this 4D embedding is subset down to ATLAS7old and ATLAS7new to focus in on two groups with a reasonable number of observations that occupy different parts of the subspace.
-Radial manual tours controlling contributions from PC4 and PC3 are shown in figure \ref{fig:JetClusterGood} and figure \ref{fig:JetClusterBad} respectively.
+Radial manual tours controlling contributions from PC4 and PC3 are shown in Figure \ref{fig:JetClusterGood} and Figure \ref{fig:JetClusterBad} respectively.
 These variables are selected to contrast the difference of information conveyed by touring different variables.
@@
 A different space is used to explore the DIS cluster; specifically the first six principal components, which explains 48% of the variation contained within the aggregated data [@cook_dynamical_2018].
-Radial manual tours are performed on PC6 and PC2 in figure \ref{fig:DISclusterGood} and figure \ref{fig:DISclusterBad} respectively.
+Radial manual tours are performed on PC6 and PC2 in Figure \ref{fig:DISclusterGood} and Figure \ref{fig:DISclusterBad} respectively.
 

### 2019-04-22 paper/spinifex-paper.tex
step 1

@@
 
-\hypertarget{set-up}{%
-\subsection{Set up}\label{set-up}}
+\hypertarget{step-0-set-up}{%
+\subsection{Step 0) Set up}\label{step-0-set-up}}
 
@@
 
-In Figure \ref{fig:step0} the contributions of the variables tars1 and
-aede2 are mostly orthogonal to the contributions of the other four
-variables. These two variables explain the variation of the data
-distinguishing the purple cluster from the rest of the sample. Select
-aede2 as the manip var, the variable to be manipulated, as it has a
-larger contribution to the projection. The question that will be
-explored is how important the variable aede2 is to the separation of the
-clusters.
+In Figure \ref{fig:step0} the contribution of the variables
+\texttt{tars1} and \texttt{aede2} mostly contrast the contribution of
+the other four variables. These two variables combined contribute in the
+direction of the projection where the purple cluster is separated from
+the other two clusters. The variable \texttt{aede2} is selected as the
+manip var, the variable to be controlled in the tour. The question that
+will be explored is how important this variable is to the separation of
+the clusters.
 

### 2019-04-22 paper/spinifex_paper.Rmd
step 1

@@
 
-## Set up
+## Step 0) Set up
 
@@
 <!-- select a manip var-->
-In Figure \ref{fig:step0} the contributions of the variables tars1 and aede2 are mostly orthogonal to the contributions of the other four variables.
-These two variables explain the variation of the data distinguishing the purple cluster from the rest of the sample.
-Select aede2 as the manip var, the variable to be manipulated, as it has a larger contribution to the projection.
-The question that will be explored is how important the variable aede2 is to the separation of the clusters.
+In Figure \ref{fig:step0} the contribution of the variables `tars1` and `aede2` mostly contrast the contribution of the other four variables.
+These two variables combined contribute in the direction of the projection where the purple cluster is separated from the other two clusters.
+The variable `aede2` is selected as the manip var, the variable to be controlled in the tour.
+The question that will be explored is how important this variable is to the separation of the clusters.
 

### 2019-04-23 paper/spinifex-paper.tex
working on algorithm

@@
 
-This section describes the algorithm for performing a 2D radial manual
-tour, containing these steps:
+The algorithm to conduct a manual tour interactively, by recording
+mouse/cursor motion, is described in detail in Cook and Buja
+(\protect\hyperlink{ref-cook_manual_1997}{1997}). Movement can be in any
+direction and magnitude, but it can also be constrained in several ways:
+
+\begin{itemize}
+\tightlist
+\item
+  \emph{radial}: fix the direction of contribution, and allow the
+  magnitude to change.
+\item
+  \emph{angular}: fix the magnitude, and allow the angle or direction of
+  the contribution to vary.
+\item
+  \emph{horizontal}, \emph{vertical}: allow rotation only around the
+  horizontal or vertical axis of the current 2D projection.
+\end{itemize}
+
+The algorithm described here produces a \textbf{radial} tour as an
+\emph{animation sequence}. It takes the current contribution of the
+chosen variable, and using rotation brings this variable fully into the
+projection, completely removes it, and then returns to the original
+position. This is an overview of the steps:
 
@@
 
-This section describes the notation used in the algorithm for a 2D
-radial manual tour.
+The notation used to describe the algorithm for a 2D radial manual tour
+is as follows:
 
@@
   \(p \times 2\) matrix, describing the projection from \(p\) to two
-  dimensions
+  dimensions. We will call this the ``original projection'' because it
+  is the starting point for the manual tour.
 \item
@@
 
-The algorithm primarily operates on projection bases and utilizes the
-data only when making a display. The projection space can be viewed at
-any point in the process by pre-multiplying the data and plotting the
-first two variables.
+The algorithm operates entirely on projection bases, and utilizes the
+data only later when making a display.
 
@@
 index function via hill-climbing on the projection space. In this case,
-the holes index is selected and applied to standardized flea data. The
-holes index is maximized when the projected observations are furthest
-from the center. Figure \ref{fig:step0} shows a locally optimized
-projection of the data. The left panel displays the reference axes of
-the projection basis, a visual indication of the magnitude and direction
-each variable contributes to the projection. The right panel shows the
-data as projected through the basis set described by the reference axes
-(left). Data points are colored and given shape according to the
-species.
+the holes index is selected and applied to the standardized flea data.
+The holes index is maximized when the projected observations are
+furthest from the center. Figure \ref{fig:step0} shows a locally
+optimized projection of the data. The left panel displays the reference
+axes of the projection basis, a visual indication of the magnitude and
+direction each variable contributes to the projection. The right panel
+shows the data as projected through the basis set described by the
+reference axes (left). Data points are colored and given shape according
+to the species.
 
@@
 (Wickham \protect\hyperlink{ref-wickham_ggplot2:_2016}{2016}) graphic
-similar to Figure \ref{fig:step0}. The projected data is always
-available for display via matrix multiplication
+similar to Figure \ref{fig:step0}. The projected data is processed for
+display using matrix multiplication
 \(\textbf{X}_{[n,~p]} ~*~ \textbf{B}_{[p,~d]} ~=~ \textbf{P}_{[n,~d]}\).
@@
 
-\hypertarget{step-1-choose-variable-of-interest}{%
-\subsection{Step 1) Choose variable of
-interest}\label{step-1-choose-variable-of-interest}}
+\hypertarget{step-1-choose-manip-variable}{%
+\subsection{Step 1) Choose manip
+variable}\label{step-1-choose-manip-variable}}
 
@@
 
-\hypertarget{step-2-create-the-manip-space}{%
-\subsection{Step 2) Create the manip
-space}\label{step-2-create-the-manip-space}}
+\hypertarget{step-2-create-the-3d-manip-space}{%
+\subsection{Step 2) Create the 3D manip
+space}\label{step-2-create-the-3d-manip-space}}
 
-Initialize a zero vector \textbf{e} of length \emph{p}. Set the fifth
-element to one, as aede2 is the fifth variable in the data, giving the
-manip var a full contribution in this dimension. Use the Gram-Schmidt
-process to orthonormalize the zero vector onto the basis yielding the 3D
-manipulation space, \textbf{M}.
+Initialize the coordinate basis vector as a zero vector \textbf{e} of
+length \emph{p}, and set the \(k^{th}\) element to 1. In the example
+data, \texttt{aede2} is the fifth variable in the data, so \(k=5\). Use
+a Gram-Schmidt process to orthonormalize the coordinate basis vector on
+the original 2D projection to describe a 3D manipulation space,
+\textbf{M}.
 
@@
   \textbf{e}_k &\leftarrow 1 \\ 
-  \textbf{e}   &\leftarrow \textbf{e} - \langle \textbf{e}, \textbf{B}_1 \rangle \textbf{B}_1 - \langle \textbf{e}, \textbf{B}_2 \rangle \textbf{B}_2 \\ 
-  \textbf{M}_{[p,~3]} &= (\textbf{B}_1,\textbf{B}_2,\textbf{e})
+  \textbf{e}^*   &\leftarrow \textbf{e} - \langle \textbf{e}, \textbf{B}_1 \rangle \textbf{B}_1 - \langle \textbf{e}, \textbf{B}_2 \rangle \textbf{B}_2 \\ 
+  \textbf{M}_{[p,~3]} &= (\textbf{B}_1,\textbf{B}_2,\textbf{e}^*)
 \end{align*}
 
-Adding this new dimension to our projection space allows for the
-coefficients of the manip var to be changed via rotation about the
-origin. For example, the ability to lift a piece of paper, rather than
-being constrained to a 2D plane. Orthonormalizing rescales the new
-vector, while the projection down to 2D remains as the original basis.
-Place the projection plane horizontally, with the new dimension
-extending vertically, with axes projecting back onto the reference axes.
-Figure \ref{fig:step2} illustrates this 3D manipulation space with the
-manip var highlighted (height of the other variables are not depicted).
-This representation is produced by calling the
+The manip space provides a 3D projection from \(p\)-dimensional space,
+where the coefficient of the manip var can range completely between
+{[}-1, 1{]}. 3D rotation can be used to rotate the mani variable
+completely into or completely out of a 2D projection. Figure
+\ref{fig:step2} illustrates this 3D manipulation space with the manip
+var highlighted. This representation is produced by calling the
 \texttt{view\_manip\_space()} function. Functionally, the manipulation
 space is more of a mathematical construct that needn't be visualized but
-does help to digest what is happening mathematically.
+this diagram provides intuition for the technique.
 
@@
 
-\caption[Manipulation space for controlling the contribution of aede2 of standardized flea data]{Manipulation space for controlling the contribution of aede2 of standardized flea data.
-The basis was selected by a holes-index optimized guided tour lies on the horizontal projection plane, shown in blue.
-The manip var axis, in red, extends into the vertical manipulation space, allowing the coefficients of the manip var to be changed by rotation around the origin.}\label{fig:step2}
+\caption[Illustration of a 3D manipulation space, constructed to change the contribution of the variable aede2 in the example data]{Illustration of a 3D manipulation space, constructed to change the contribution of the variable aede2 in the example data.
+The red circle indicates a unit sphere.
+The 2D projection is represented by the blue circle, with the projection coefficients represented by grey lines and text.
+The manip var axis, in red, has length 1 touching the sphere, extends the projection to a third dimension.
+The shadow of this axis (blue) is its contribution in the original 2D projection.}\label{fig:step2}
+\end{figure}
+\end{Schunk}
+
+\begin{Schunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{spinifex_paper_files/figure-latex/step3-1}
+
+}
+
+\caption[Snapshots of a radial manual tour manipulating aede2]{Snapshots of a radial manual tour manipulating aede2: (1) original projection, (2) full contribution, (3) zero contribution, (4) back to original.
+}\label{fig:step3}
 \end{figure}
@@
 
-Imagine holding the manip var, the red axis, one end fixed to the
-origin. As it is controlled the manipulation space rotates about the
-origin, the projection onto the horizontal projection plane
-correspondingly moves. This is what happens in a manual tour. Generating
-a sequence of values for the horizontal and vertical, angles produces a
-path for the rotation of the manipulation space. This defines the
-(orthonormally-constrained) rotation on the coefficients of the
-variables.
+Imagine holding the manip var, the red axis in Figure \ref{fig:step2},
+at one end fixed to the origin and moving it like a lever. This is
+effectively the process of the radial manual to rotate the manip
+variable into and out of the 2D projection (Figure \ref{fig:step3}).
+
+As it is controlled the manipulation space rotates about the origin, the
+projection onto the horizontal projection plane correspondingly moves.
+This is what happens in a manual tour. Generating a sequence of values
+for the horizontal and vertical, angles produces a path for the rotation
+of the manipulation space. This defines the (orthonormally-constrained)
+rotation on the coefficients of the variables.
 
@@
 
-Figure \ref{fig:step3} illustrates a sequence with 15 projected bases,
+Figure \ref{fig:step4} illustrates a sequence with 15 projected bases,
 showing the reference axes on the top half with the corresponding
@@
 
-{\centering \includegraphics[width=5.83in,height=7in]{spinifex_paper_files/figure-latex/step3-1}
+{\centering \includegraphics[width=5.83in,height=7in]{spinifex_paper_files/figure-latex/step4-1}
 
@@
 This shows that aede2 is important for distinguishing this species from the ramaining data.
-An animated version can be viewed at  https://nspyrison.netlify.com/thesis/flea\_manualtour\_mvar5/.}\label{fig:step3}
+An animated version can be viewed at  https://nspyrison.netlify.com/thesis/flea\_manualtour\_mvar5/.}\label{fig:step4}
 \end{figure}

### 2019-04-23 paper/spinifex_paper.Rmd
working on algorithm

@@
 library(ggplot2)
+library(gridExtra)
 ```
@@
 
-This section describes the algorithm for performing a 2D radial manual tour, containing these steps:
+The algorithm to conduct a manual tour interactively, by recording mouse/cursor motion, is described in detail in @cook_manual_1997.
+Movement can be in any direction and magnitude, but it can also be constrained in several ways:
+
+- *radial*: fix the direction of contribution, and allow the magnitude to change.
+- *angular*: fix the magnitude, and allow the angle or direction of the contribution to vary.
+- *horizontal*, *vertical*: allow rotation only around the horizontal or vertical axis of the current 2D projection.
+
+The algorithm described here produces a **radial** tour as an *animation sequence*.
+It takes the current contribution of the chosen variable, and using rotation brings this variable fully into the projection, completely removes it, and then returns to the original position.
+This is an overview of the steps:
 
@@
 
-This section describes the notation used in the algorithm for a 2D radial manual tour.
+The notation used to describe the algorithm for a 2D radial manual tour is as follows:
 
@@
 - $\textbf{X}$, the data, an $n \times p$ numeric matrix to be projected to 2D.
-- $\textbf{B} = (B_1,~ B_2)$, any 2D orthonormal projection basis, $p \times 2$ matrix, describing the projection from $p$ to two dimensions
+- $\textbf{B} = (B_1,~ B_2)$, any 2D orthonormal projection basis, $p \times 2$ matrix, describing the projection from $p$ to two dimensions.
+We will call this the "original projection" because it is the starting point for the manual tour.
 - $\textbf{e}$, a 1D coordinate basis vector of length $p$ with the $k-$th element set to one, where $k$ is the index of the variable to manipulate, called the "manip var".
@@
 <!-- operate on bases -->
-The algorithm primarily operates on projection bases and utilizes the data only when making a display.
-The projection space can be viewed at any point in the process by pre-multiplying the data and plotting the first two variables.
+The algorithm operates entirely on projection bases, and utilizes the data only later when making a display.
+<!--The projection space can be viewed at any point in the process by pre-multiplying the data and plotting the first two variables.-->
 
@@
 In a guided tour, the projection sequence is selected by optimizing an index function via hill-climbing on the projection space.
-In this case, the holes index is selected and applied to standardized flea data.
+In this case, the holes index is selected and applied to the standardized flea data.
 The holes index is maximized when the projected observations are furthest from the center.
@@
 Call `view_basis()` on a basis to produce a \CRANpkg{ggplot2} [@wickham_ggplot2:_2016] graphic similar to Figure \ref{fig:step0}.
-The projected data is always available for display via matrix multiplication $\textbf{X}_{[n,~p]} ~*~ \textbf{B}_{[p,~d]} ~=~ \textbf{P}_{[n,~d]}$.
+The projected data is processed for display using matrix multiplication $\textbf{X}_{[n,~p]} ~*~ \textbf{B}_{[p,~d]} ~=~ \textbf{P}_{[n,~d]}$.
 
@@
 ## Flea holes tour
+set.seed(20190425)
 f_dat  <- tourr::rescale(flea[,1:6])
@@
 
-## Step 1) Choose variable of interest
+## Step 1) Choose manip variable
 
@@
 
-## Step 2) Create the manip space
+## Step 2) Create the 3D manip space
 
 <!-- Zero Vect, manip sp -->
-Initialize a zero vector **e** of length *p*.
-Set the fifth element to one, as  aede2 is the fifth variable in the data, giving the manip var a full contribution in this dimension.
-Use the Gram-Schmidt process to orthonormalize the zero vector onto the basis yielding the 3D manipulation space, **M**.
+Initialize the coordinate basis vector as a zero vector **e** of length *p*, and set the $k^{th}$ element to 1.
+In the example data, `aede2`  is the fifth variable in the data, so $k=5$.
+Use a Gram-Schmidt process to orthonormalize the coordinate basis vector on the original 2D projection to describe a 3D manipulation space, **M**.
 
@@
   \textbf{e}_k &\leftarrow 1 \\ 
-  \textbf{e}   &\leftarrow \textbf{e} - \langle \textbf{e}, \textbf{B}_1 \rangle \textbf{B}_1 - \langle \textbf{e}, \textbf{B}_2 \rangle \textbf{B}_2 \\ 
-  \textbf{M}_{[p,~3]} &= (\textbf{B}_1,\textbf{B}_2,\textbf{e})
+  \textbf{e}^*   &\leftarrow \textbf{e} - \langle \textbf{e}, \textbf{B}_1 \rangle \textbf{B}_1 - \langle \textbf{e}, \textbf{B}_2 \rangle \textbf{B}_2 \\ 
+  \textbf{M}_{[p,~3]} &= (\textbf{B}_1,\textbf{B}_2,\textbf{e}^*)
 \end{align*}
@@
 <!-- What the manip space facilitates -->
-Adding this new dimension to our projection space allows for the coefficients of the manip var to be changed via rotation about the origin.
-For example, the ability to lift a piece of paper, rather than being constrained to a 2D plane.
-Orthonormalizing rescales the new vector, while the projection down to 2D remains as the original basis.
-Place the projection plane horizontally, with the new dimension extending vertically, with axes projecting back onto the reference axes.
-Figure \ref{fig:step2} illustrates this 3D manipulation space with the manip var highlighted (height of the other variables are not depicted).
+The manip space provides a 3D projection from $p$-dimensional space, where the coefficient of the manip var can range completely between [-1, 1].
+3D rotation can be used to rotate the mani variable completely into or completely out of a 2D projection.
+Figure \ref{fig:step2} illustrates this 3D manipulation space with the manip var highlighted.
 This representation is produced by calling the `view_manip_space()` function.
-Functionally, the manipulation space is more of a mathematical construct that needn't be visualized but does help to digest what is happening mathematically.
+Functionally, the manipulation space is more of a mathematical construct that needn't be visualized but this diagram provides intuition for the technique.
 
-```{r step2, echo=F, out.width='70%',                                           fig.cap = "Manipulation space for controlling the contribution of aede2 of standardized flea data.
-The basis was selected by a holes-index optimized guided tour lies on the horizontal projection plane, shown in blue.
-The manip var axis, in red, extends into the vertical manipulation space, allowing the coefficients of the manip var to be changed by rotation around the origin."}
+```{r step2, echo=F, out.width='70%',                                           fig.cap = "Illustration of a 3D manipulation space, constructed to change the contribution of the variable aede2 in the example data.
+The red circle indicates a unit sphere.
+The 2D projection is represented by the blue circle, with the projection coefficients represented by grey lines and text.
+The manip var axis, in red, has length 1 touching the sphere, extends the projection to a third dimension.
+The shadow of this axis (blue) is its contribution in the original 2D projection."}
 
@@
 
+```{r step3, echo=F, warning=F, fig.height=2, fig.width=8, out.width='100%',  fig.cap = "Snapshots of a radial manual tour manipulating aede2: (1) original projection, (2) full contribution, (3) zero contribution, (4) back to original.
+"}
+f_angle <- .26
+f_mtour <- manual_tour(f_bas, manip_var = f_mvar, angle = f_angle)
+if(dim(f_mtour)[3] != 15) message(
+  paste0("!!!!! step3 n_slides != 15 !!!!! Please knit again. n_slides = ", dim(f_mtour)[3]))
+
+p1 <- view_basis(f_mtour[,,1], labels=colnames(f_dat)) + ggtitle("1")
+p2 <- view_basis(f_mtour[,,6], labels=colnames(f_dat)) + ggtitle("2")
+p3 <- view_basis(f_mtour[,,12], labels=colnames(f_dat)) + ggtitle("3")
+p4 <- view_basis(f_mtour[,,15], labels=colnames(f_dat)) + ggtitle("4")
+grid.arrange(p1, p2, p3, p4, ncol=4)
+```
+
 ## Step 3) Generate rotation
@@
 <!-- illustration of axis manip -->
-Imagine holding the manip var, the red axis, one end fixed to the origin.
+Imagine holding the manip var, the red axis in Figure \ref{fig:step2}, at one end fixed to the origin and moving it like a lever.
+This is effectively the process of the radial manual to rotate the manip variable into and out of the 2D projection (Figure \ref{fig:step3}).
+
 As it is controlled the manipulation space rotates about the origin, the projection onto the horizontal projection plane correspondingly moves.
@@
 <!-- description of slides, on the technical side we haven't gone back to data space yet. -->
-Figure \ref{fig:step3} illustrates a sequence with 15 projected bases, showing the reference axes on the top half with the corresponding projected data points below.
+Figure \ref{fig:step4} illustrates a sequence with 15 projected bases, showing the reference axes on the top half with the corresponding projected data points below.
 Changes in the basis coefficients of the manip var correspond to the distance separating the purple cluster and the remaining sample, aede2 is crucial in distinguishing the variation of species.
@@
 
-```{r step3, echo=F, warning=F, fig.height=7.0, fig.width=5.83, out.height='7in', out.width='5.83in',                                                                fig.cap = "Radial manual tour manipulating aede2 of standardized flea data.
+```{r step4, echo=F, warning=F, fig.height=7.0, fig.width=5.83, out.height='7in', out.width='5.83in',                                                                fig.cap = "Radial manual tour manipulating aede2 of standardized flea data.
 The contributions increase from its initial contribution to a full contribution to the projection before decreasing to zero and then returning to its initial value.
@@
 An animated version can be viewed at  https://nspyrison.netlify.com/thesis/flea\\_manualtour\\_mvar5/."}
-f_angle <- .26
-f_mtour <- manual_tour(f_bas, manip_var = f_mvar, angle = f_angle)
-if(dim(f_mtour)[3] != 15) message(
-  paste0("!!!!! step3 n_slides != 15 !!!!! Please knit again. n_slides = ", dim(f_mtour)[3]))
-
 #play_manual_tour(f_dat, f_bas, f_mvar, col = f_cat, angle = f_angle)

### 2019-04-24 zRotationMatrix.r
exploring rotation matrix, and 3rd euler angle.

@@ -1,0 +1,84 @@
+## scratchpad for changign rotation matrix
+# https://en.wikipedia.org/wiki/3D_rotation_group
+# https://en.wikipedia.org/wiki/Rotation_matrix#Basic_rotations
+
+library(spinifex)
+
+
+if (F) #from rotate_manip_space
+{
+  rotate_manip_space
+
+  theta = 1
+  phi   = 2
+
+  s_theta <- sin(theta)
+  c_theta <- cos(theta)
+  s_phi <- sin(phi)
+  c_phi <- cos(phi)
+  R <- matrix(c(c_theta^2 * c_phi + s_theta^2,
+                -c_theta * s_theta * (1 - c_phi),
+                -c_theta * s_phi,  # 3
+                -c_theta * s_theta * (1 - c_phi),
+                s_theta^2 * c_phi + c_theta^2,
+                -s_theta * s_phi, # 6
+                c_theta * s_phi,
+                s_theta * s_phi, c_phi), # 9
+              nrow = 3, ncol = 3,
+              byrow = TRUE)
+}
+
+
+# https://en.wikipedia.org/wiki/Rotation_matrix#Basic_rotations
+
+p=3
+#m <- diag(p)
+a <- 1/1*pi
+b <- 1/2*pi
+c <- 0 #1/3*pi # 0 is the null hypothesis for current state.
+cos_a <- cos(a)
+sin_a <- sin(a)
+cos_b <- cos(b)
+sin_b <- sin(b)
+cos_c <- cos(c)
+sin_c <- sin(c)
+
+# cos -sin
+# sin  cos
+r_a <- matrix(c(1,
+                0,
+                0,  # 3
+                0,
+                cos_a,
+                -sin_a, # 6
+                0,
+                sin_a,
+                cos_a), # 9
+              nrow = 3, ncol = 3,
+              byrow = TRUE)
+
+r_b <- matrix(c(cos_b,
+                0,
+                sin_b, # 3
+                0,
+                1,
+                0, # 6
+                -sin_a,
+                0,
+                cos_a), # 9
+              nrow = 3, ncol = 3,
+              byrow = TRUE)
+
+r_c <- matrix(c(cos_c,
+                -sin_c,
+                0, # 3
+                sin_c,
+                cos_c,
+                0, # 6
+                0,
+                0,
+                1), # 9
+              nrow = 3, ncol = 3,
+              byrow = TRUE)
+
+R <- r_a * r_b * r_c

### 2019-04-25 zRotationMatrix.r
using 3 euler angles doesn't behave at all like the previous method.

@@ -4,81 +4,171 @@
 
 library(spinifex)
 
+set.seed(20190425)
+f_dat  <- tourr::rescale(flea[,1:6])
+f_cat  <- factor(flea$species)
+f_bas  <- basis_random(6,2)
+f_mvar <- 4
 
-if (F) #from rotate_manip_space
+view_manip_space(basis = f_bas,manip_var = f_mvar)
+#play_manual_tour(data = f_dat, basis = f_bas, manip_var = f_mvar,
+#                 col = col_of(flea$species), axes = "bottomleft")
+
+load_funcs <- T
+if (load_funcs) # overwrites the spinifex functions: rotate_manip_space and manual_tour
 {
-  rotate_manip_space
+  #rotate_manip_space
+  rotate_manip_space <- function(manip_space,
+                                    theta=NULL,
+                                    phi=NULL,
+                                    omega=NULL # if null, use Cook's 2 angle rotation matrix.
+  )
+  {
+    # if(is.null(theta)) theta = 1
+    # if(is.null(phi))   phi   = 2
+    s_theta <- sin(theta)
+    c_theta <- cos(theta)
+    s_phi <- sin(phi)
+    c_phi <- cos(phi)
+    if(!is.null(omega))
+      {s_omega <- sin(omega)
+      c_omega <- cos(omega)}
 
-  theta = 1
-  phi   = 2
+    if(is.null(omega)){ # if omega not passed, use Cook defined (2 angle) rotation.
+      R <- matrix(c(c_theta^2 * c_phi + s_theta^2,
+                    -c_theta * s_theta * (1 - c_phi),
+                    -c_theta * s_phi,  # 3
+                    -c_theta * s_theta * (1 - c_phi),
+                    s_theta^2 * c_phi + c_theta^2,
+                    -s_theta * s_phi, # 6
+                    c_theta * s_phi,
+                    s_theta * s_phi, c_phi), # 9
+                  nrow = 3, ncol = 3,
+                  byrow = TRUE)
+    } else { #if omega is passed, use 3 euler angle rotation
+      r_theta <- matrix(c(1,
+                      0,
+                      0,  # 3
+                      0,
+                      c_theta,
+                      -s_theta, # 6
+                      0,
+                      s_theta,
+                      c_theta), # 9
+                    nrow = 3, ncol = 3,
[…]

### 2019-04-25 zRotationMatrix.r
fixed matrix multiplication. the 3 angle rotation still not making sense in the same way that Gram schmidt is.

@@ -82,7 +82,8 @@ if (load_funcs) # overwrites the spinifex functions: rotate_manip_space and manu
                     nrow = 3, ncol = 3,
                     byrow = TRUE)
 
-      R <- r_theta * r_phi * r_omega
+      R <- r_theta %*% r_omega%*% r_phi
+      # Order matters, 3rd angle is projection plane rotation.
     }
 
 
@@ -92,83 +93,62 @@ if (load_funcs) # overwrites the spinifex functions: rotate_manip_space and manu
     rotation_space
   }
 
-  # # paste over spinifex function
-  # manual_tour <- function (basis = NULL, manip_var, theta = NULL, phi_min = 0,
-  #                          phi_max = 0.5 * pi, omega=NULL, angle = 0.05)
-  # {
-  #   basis <- as.matrix(basis)
-  #   p <- nrow(basis)
-  #   d <- ncol(basis)
-  #   manip_space <- create_manip_space(basis = basis, manip_var = manip_var)
-  #   if (is.null(theta))
-  #     theta <- atan(basis[manip_var, 2]/basis[manip_var, 1])
-  #   phi_start <- acos(sqrt(basis[manip_var, 1]^2 + basis[manip_var, 2]^2))
-  #   stopifnot(phi_min <= phi_start & phi_max >= phi_start)
-  #   find_path <- function(start, end) {
-  #     mvar_xsign <- -sign(basis[manip_var, 1])
-  #     start <- mvar_xsign * (start - phi_start)
-  #     end <- mvar_xsign * (end - phi_start)
-  #     dist <- abs(end - start)
-  #     steps <- round(dist/angle)
-  #     angle <- dist/steps
-  #     sign <- ifelse(end > start, 1, -1)
-  #     seq(from = start, to = end, by = sign * angle)
-  #   }
-  #   phi_path <- c(find_path(start = phi_start, end = phi_min),
-  #                 find_path(start = phi_min, end = phi_max), find_path(start = phi_max,
-  #                                                                      end = phi_start))
-  #   n_frames <- length(phi_path)
-  #   basis_set <- array(NA, dim = c(p, d, n_frames))
-  #   for (i in 1:n_frames) {
[…]

### 2019-04-25 zRotationMatrix.r
changing fixed values.

@@ -98,22 +98,24 @@ if (load_funcs) # overwrites the spinifex functions: rotate_manip_space and manu
 #runif(1, max = 2 * pi)
 #my_t <- 1
 rad_theta <- atan(f_bas[f_mvar, 2]/f_bas[f_mvar, 1])
-my_p <- 2
 f_msp <- create_manip_space(basis = f_bas, manip_var = f_mvar)
 b <- rotate_manip_space(f_msp, theta = my_t, phi = my_p) #omega is Null)
-b0 <- rotate_manip_space(f_msp, theta = my_p, phi = my_p, omega = 0)
+b_a  <- rotate_manip_space(f_msp, theta = rad_theta, phi = rad_theta, omega = 0)
 # comparing 2 Euler angle rotaion with 3 angle.
 view_basis(b)
-view_basis(o0) # far cry from equal, values too small, orthonormality broken?
+view_basis(b_a) # far cry from equal, values too small, orthonormality broken?
 is_orthonormal(b) # neither are orthonormal
-is_orthonormal(o0)
+is_orthonormal(b_a)
 # changing omega; fixed theta, phi. manip_var = 4. rotating whole plane.
-o05 <- rotate_manip_space(f_msp, theta = rad_theta, phi = my_p, omega = .5)
-o1  <- rotate_manip_space(f_msp, theta = rad_theta, phi = my_p, omega = 1)
-o15 <- rotate_manip_space(f_msp, theta = rad_theta, phi = my_p, omega = 1.5)
-o2  <- rotate_manip_space(f_msp, theta = rad_theta, phi = my_p, omega = 2)
-o25 <- rotate_manip_space(f_msp, theta = rad_theta, phi = my_p, omega = 2.5)
-o3  <- rotate_manip_space(f_msp, theta = rad_theta, phi = my_p, omega = 3)
+my_t <- rad_theta
+my_p <- rad_theta
+o0  <- rotate_manip_space(f_msp, theta = my_t, phi = my_p, omega = 0)
+o05 <- rotate_manip_space(f_msp, theta = my_t, phi = my_p, omega = .5)
+o1  <- rotate_manip_space(f_msp, theta = my_t, phi = my_p, omega = 1)
+o15 <- rotate_manip_space(f_msp, theta = my_t, phi = my_p, omega = 1.5)
[…]

### 2019-04-26 paper/spinifex-paper.tex
Rotation matrix explained. grammerly, knit.

@@
 collectively known as \textbf{tours} provide an important tool for
-exploring mulitvariate data and models. The R package \textbf{tourr}
+exploring multivariate data and models. The R package \textbf{tourr}
 provides functions for several types of tours: grand, guided, little,
@@
 multiple variables, but they are limited because they show only a
-glimpse of the high-dimensional space. An alternative approach, is to
-use a tour of dynamic linear projections, to look at many different
-low-dimensional proections. Tours can be considered to extend the
+glimpse of the high-dimensional space. An alternative approach is to use
+a tour of dynamic linear projections, to look at many different
+low-dimensional projections. Tours can be considered to extend the
 dimensionality of visualization, which is important as data and models
@@
 tour (Cook et al. \protect\hyperlink{ref-cook_grand_1995}{1995}). Guided
-tours select a very specific path, that which approaches a projection
-that optimizes an objective-function. The optimization is conducted in a
+tours select a very specific path, which approaches a projection that
+optimizes an objective-function. The optimization is conducted in a
 manner similar to simulated annealing (Kirkpatrick, Gelatt, and Vecchi
@@
 
-The algorithm operates entirely on projection bases, and utilizes the
+The algorithm operates entirely on projection bases and utilizes the
 data only later when making a display.
@@
 Imagine holding the manip var, the red axis in Figure \ref{fig:step2},
-at one end fixed to the origin and moving it like a lever. This is
+one end fixed to the origin and moving it like a lever. This is
 effectively the process of the radial manual to rotate the manip
-variable into and out of the 2D projection (Figure \ref{fig:step3}).
-
-As it is controlled the manipulation space rotates about the origin, the
-projection onto the horizontal projection plane correspondingly moves.
-This is what happens in a manual tour. Generating a sequence of values
-for the horizontal and vertical, angles produces a path for the rotation
-of the manipulation space. This defines the (orthonormally-constrained)
-rotation on the coefficients of the variables.
+variable into and out of the 2D projection (Figure \ref{fig:step3}). As
+the variable is controlled, the manipulation space rotates about the
+origin, the projection onto the horizontal projection plane
+correspondingly moves. This is what happens in a manual tour. Generating
+a sequence of values for the horizontal and vertical, angles produces a
+path for the rotation of the manipulation space. This defines the
+(orthonormally-constrained) rotation on the basis coefficients.
 
@@
 contributions of the manip var to a maximum and then to zero before
-returning to the initial position. For an oblique manual tour, capture
-the user manipulation directly on the \(XY\) of the projection plane.
+returning to the initial position. Alternatively, for an oblique manual
+tour, capture the user manipulation directly on the \(XY\) of the
+projection plane.
 
-Post-multiply the manipulation space by the pre-defined rotation matrix
-producing \textbf{RM}, the rotated manip space. This is one frame of
-basis values, repeat this process for each value in the sequences of
-\(\theta\) and \(\phi\) for the complete animation.
+Rotations are defined by the axes they pivot on, \(\theta\) then is the
+rotation around the \(Z\) axis. While \(\phi\) will be the rotation
+around, \(U\), a new axis on the \(XY\) plane, orthogonal to \(\phi\).
+Given these axes, the rotation matrix can be derived by the following.
 
@@
 \begin{itemize}
-  \item[$c_\theta$] be the cosine of $\theta_i$
+  \item[$c_\theta$] be the cosine of $\theta$
   \item[$c_\phi$]   be the cosine of $\phi_i$
-  \item[$s_\theta$] be the sine of   $\theta_i$
+  \item[$s_\theta$] be the sine of   $\theta$
   \item[$s_\phi$]   be the sine of   $\phi_i$
@@
 
-\textbf{For each value of } \(i\) \textbf{:}
-
 \begin{align*}
-  \textbf{RM}_{[p,~3,~i]}
-  &= \textbf{M}_{[p,~3]} ~*~ \textbf{R}_{[3,~3,~i]} \\
-  &= \textbf{M}_{[p,~3]}
-    ~*~
+  \textbf{B}_k &= (c_\theta, s_\theta, 0) \\
+  \vec{u} &= (s_\theta, -c_\theta, 0) \\ &=
   \begin{bmatrix}
-    c_\theta^2 c_\phi s_\theta^2 &
+     0 & 0 & -c_\theta \\
+     0 & 0 & -s_\theta \\
+     c_\theta & s_\theta & 0 \\
+   \end{bmatrix} \\
+  \text{Apply Rodrigues' rotation formula:} \\
+    \textbf{R}_{[3,~3]} 
+    &= \textbf{I}_3 + s_\phi\*\vec{u} + (1-c_\phi)\*\vec{u}^2 \\
+    &= 
+    \begin{bmatrix}
+      c_\theta^2 c_\phi + s_\theta^2 &
       -c_\theta s_\theta (1 - c_\phi) &
@@
       c_\phi
-  \end{bmatrix}_{[3,~3,~i]}
+    \end{bmatrix}
+\end{align*}
+
+Post-multiply the manipulation space by the pre-defined rotation matrix
+producing \textbf{RM}, the rotated manip space. This is one frame of
+basis values, repeat this process for each value in the sequence
+\(\phi_i\) for the complete animation.
+
+\textbf{For each value of } \(\phi_i\) \textbf{:}
+
+\begin{align*}
+  \textbf{RM}_{[p,~3,~i]} &= \textbf{M}_{[p,~3]} ~*~ \textbf{R}_{[3,~3,~i]} 
 \end{align*}
@@
 
-In the following section describes and demonstrates usage of the above
-functions following the order in the algorithm. The utility functions
-are called by the tour animation functions, passed to the arguments
-\texttt{render\_type}. Usage is demonstrated below and in the example
-documentation.
+In the following section describes and demonstrates the usage of the
+above functions following the order in the algorithm. The utility
+functions are called by the tour animation functions, passed to the
+arguments \texttt{render\_type}. Usage is demonstrated below and in the
+example documentation.
 
@@
 an approachable demo tool. The user could easily switch between
-variables to control, adjust interpolation step angle, or flag/save
+variables to control, adjust the interpolation step angle, or flag/save
 specific frame basis sets.

### 2019-04-26 paper/spinifex_paper.Rmd
note about angles
Rotation matrix explained. grammerly, knit.

@@
 abstract: >
-  Dynamic low-dimensional linear projections of multivariate data collectively known as **tours** provide an important tool for exploring mulitvariate data and models.
+  Dynamic low-dimensional linear projections of multivariate data collectively known as **tours** provide an important tool for exploring multivariate data and models.
 The R package **tourr** provides functions for several types of tours: grand, guided, little, local and frozen.
@@
 
-keywords: exploratory data anlaysis, projection pursuit, high dimensional data, data visualization, statistical graphics, data science
+keywords: exploratory data analysis, projection pursuit, high dimensional data, data visualization, statistical graphics, data science
 preamble: >
@@
 These are useful for finding relationships between multiple variables, but they are limited because they show only a glimpse of the high-dimensional space.
-An alternative approach, is to use a tour of dynamic linear projections, to look at many different low-dimensional proections.
+An alternative approach is to use a tour of dynamic linear projections, to look at many different low-dimensional projections.
 Tours can be considered to extend the dimensionality of visualization, which is important as data and models exist in more than 3D.
@@
 One way to identify "interesting" features is with the use of a guided tour [@cook_grand_1995].
-Guided tours select a very specific path, that which approaches a projection that optimizes an objective-function.
+Guided tours select a very specific path, which approaches a projection that optimizes an objective-function.
 The optimization is conducted in a manner similar to simulated annealing [@kirkpatrick_optimization_1983].
@@
 
-<!-- remember mixing latex and rmarkdown bullets isn't trivial. -->
 - $\textbf{X}$, the data, an $n \times p$ numeric matrix to be projected to 2D.
@@
 <!-- operate on bases -->
-The algorithm operates entirely on projection bases, and utilizes the data only later when making a display.
+The algorithm operates entirely on projection bases and utilizes the data only later when making a display.
 <!--The projection space can be viewed at any point in the process by pre-multiplying the data and plotting the first two variables.-->
@@
 <!-- illustration of axis manip -->
-Imagine holding the manip var, the red axis in Figure \ref{fig:step2}, at one end fixed to the origin and moving it like a lever.
+Imagine holding the manip var, the red axis in Figure \ref{fig:step2}, one end fixed to the origin and moving it like a lever.
 This is effectively the process of the radial manual to rotate the manip variable into and out of the 2D projection (Figure \ref{fig:step3}).
-
-As it is controlled the manipulation space rotates about the origin, the projection onto the horizontal projection plane correspondingly moves.
+As the variable is controlled, the manipulation space rotates about the origin, the projection onto the horizontal projection plane correspondingly moves.
 This is what happens in a manual tour.
 Generating a sequence of values for the horizontal and vertical, angles produces a path for the rotation of the manipulation space.
-This defines the (orthonormally-constrained) rotation on the coefficients of the variables.
+This defines the (orthonormally-constrained) rotation on the basis coefficients.
 
@@
 For a radial tour fix the (horizontal) angle within the projection plane, $\theta$, and define a sequence for the (vertical) angle coming out of the projection plane, $\phi$, bringing the initial $XY$ contributions of the manip var to a maximum and then to zero before returning to the initial position.
-For an oblique manual tour, capture the user manipulation directly on the $XY$ of the projection plane.
+Alternatively, for an oblique manual tour, capture the user manipulation directly on the $XY$ of the projection plane.
 
-Post-multiply the manipulation space by the pre-defined rotation matrix producing **RM**, the rotated manip space.
-This is one frame of basis values, repeat this process for each value in the sequences of $\theta$ and $\phi$ for the complete animation.
+<!-- define the rotation matrix -->
+Rotations are defined by the axes they pivot on, $\theta$ then is the rotation around the $Z$ axis.
+While $\phi$ will be the rotation around, $U$, a new axis on the $XY$ plane, orthogonal to $\phi$.
+Given these axes, the rotation matrix can be derived by the following.
 
@@
 
-**For each value of ** $i$ **:**
-<!-- don't mix rmd and latex... -->
-
 \begin{align*}
-  \textbf{RM}_{[p,~3,~i]}
-  &= \textbf{M}_{[p,~3]} ~*~ \textbf{R}_{[3,~3,~i]} \\
-  &= \textbf{M}_{[p,~3]}
-    ~*~
+  \textbf{B}_k &= (c_\theta, s_\theta, 0) \\
+  \vec{u} &= (s_\theta, -c_\theta, 0) \\ &=
   \begin{bmatrix}
-    c_\theta^2 c_\phi s_\theta^2 &
+     0 & 0 & -c_\theta \\
+     0 & 0 & -s_\theta \\
+     c_\theta & s_\theta & 0 \\
+   \end{bmatrix} \\
+  \text{Apply Rodrigues' rotation formula:} \\
+    \textbf{R}_{[3,~3]} 
+    &= \textbf{I}_3 + s_\phi\*\vec{u} + (1-c_\phi)\*\vec{u}^2 \\
+    &= 
+    \begin{bmatrix}
+      c_\theta^2 c_\phi + s_\theta^2 &
       -c_\theta s_\theta (1 - c_\phi) &
@@
       c_\phi
-  \end{bmatrix}_{[3,~3,~i]}
+    \end{bmatrix}
+\end{align*}
+
+<!-- rotate manip space -->
+Post-multiply the manipulation space by the pre-defined rotation matrix producing **RM**, the rotated manip space.
+This is one frame of basis values, repeat this process for each value in the sequence $\phi_i$ for the complete animation.
+
+**For each value of ** $\phi_i$ **:**
+<!-- don't mix rmd and latex... -->
+
+\begin{align*}
+  \textbf{RM}_{[p,~3,~i]} &= \textbf{M}_{[p,~3]} ~*~ \textbf{R}_{[3,~3,~i]} 
 \end{align*}
@@
 <!-- step3 caption -->
-
 ```{r step4, echo=F, warning=F, fig.height=7.0, fig.width=5.83, out.height='7in', out.width='5.83in',                                                                fig.cap = "Radial manual tour manipulating aede2 of standardized flea data.
@@
 
-In the following section describes and demonstrates usage of the above functions following the order in the algorithm.
+In the following section describes and demonstrates the usage of the above functions following the order in the algorithm.
 The utility functions are called by the tour animation functions, passed to the arguments `render_type`.
@@
 The development of a dynamic graphical user interface, perhaps with the use of a \CRANpkg{shiny} [@chang_shiny:_2018] application, would mitigate the barrier to entry, allow for more rapid analysis, and offer an approachable demo tool.
-The user could easily switch between variables to control, adjust interpolation step angle, or flag/save specific frame basis sets.
+The user could easily switch between variables to control, adjust the interpolation step angle, or flag/save specific frame basis sets.
 

### 2019-04-26 paper/spinifex-paper.tex
remainder of hard copy edits. grammerly, knit.

@@
 \textbf{plotly} as an HTML5 object or \textbf{gganimate} as GIF or MP4
-files. Both of which build off \textbf{ggplot2} objects in internal
-functions. Sharing of animations is not trivial especially in print and
-static file formats such as PDF. Even with the use of computers and
-dynamic file formats capturing the correct resolution, aspect, and
-display is challenging and many formats quickly bloat file sizes. Keep
-in mind hosting options and exporting functions from \textbf{plotly},
+files. Sharing of animations is not trivial, especially static formats
+like print and PDF. Even with dynamic display capturing the correct
+resolution and aspect ratio can be challenging, while many formats
+quickly bloat file sizes limiting sharing options. Keep in mind hosting
+options and exporting functions offered in \textbf{plotly},
 \textbf{gganimate} and \textbf{tourr}.
@@
 (TensorFlow embedded projections) or traditional static embeddings.
-Before applying manual tours let's discuss the structure of the data.
+Before applying manual tours the structure of the data is discussed.
 
@@
 \ref{fig:JetClusterGood} and Figure \ref{fig:JetClusterBad}
-respectively. These variables are selected to contrast the difference of
-information conveyed by touring different variables. Links to dynamic
-HTML5 animations controlling each of the four variables are also
-provided.
+respectively. These variables are selected to contrast the information
+conveyed by different manip variables. Links to dynamic HTML5 animations
+controlling each of the four variables are also provided.
 
@@
 space spanned by the experiment types ATLAS7new and ATLAS7old.
-Specifically, the variation of ATLAS7new becomes more singular. The
-experiments are probing different parameter space and PC4 is important
-to demonstrate this. Yet, when PC3 is manipulated there is no clear
-indication that the different experiments probe different parameter
-space. Performing a radial manual tour on PC4 is more insightful than
-for PC3. Radial manual tours manipulating each of the principal
-components in the jet cluster can be viewed by following the links:
+Specifically, the variation of ATLAS7new (green) becomes more singular.
+The experiments are probing different parameter space and PC4 is
+important to demonstrate this. Yet, when PC3 is manipulated there is no
+clear indication that the different experiments probe different
+parameter space. Performing a radial manual tour on PC4 is more
+insightful than for PC3. Radial manual tours manipulating each of the
+principal components in the jet cluster can be viewed by following the
+links:
 \href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc1/}{PC1},
@@
 During the intermediate frames, the ATLAS7new is compressed in the direction radial to PC4.
-The differece in distribution shape demonstrates the experiments probe different phase-space, which has a linear mapping back to the original dimensions.
+The differece in distribution shape demonstrates the experiments probe different phase-space, which has a linear mapping back to the original variables for interpretation and further exploration.
 An HTML5 version can be viewed at https://nspyrison.netlify.com/thesis/jetcluster\_manualtour\_pc4/.}\label{fig:JetClusterGood}
@@
 dimensionality and shape information. Manipulating the contributions of
-PC6 turned out to be much more insightful than that of PC2. This result
-might seem counter-intuitive at first as PC2 should explain much more of
-the variation in the data. However, features and structures in the data
-regularly reside in smaller dimensionality. The details finer resolution
-of these features can be lost when looking only at static projections.
-DIS cluster manual tours manipulating each of the principal components
-can be viewed from the links:
+PC6 turned out to be much more insightful than PC2. This result might
+seem counter-intuitive at first as PC2 should explain much more of the
+variation in the data. However, features and structures in the data
+regularly reside in smaller dimensionality which can be overlooked when
+optimizing on full sample statistics. DIS cluster manual tours
+manipulating each of the principal components can be viewed from the
+links:
 \href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc1/}{PC1},
@@
 This jet does extend more in the plane of view when the contribution of PC2 is full, giving insight to its orientation.
-However, less information about the shape of DIS HERA (green) and charm SIDIS (orange) is available.
+However, less information about the shape of DIS HERA (green) and charm SIDIS (orange) is available compared to PC6 as the manip var.
 An HTML5 version can be viewed at https://nspyrison.netlify.com/thesis/discluster\_manualtour\_pc2/.}\label{fig:DISclusterBad}
@@
 
-Tours, which are a dynamic linear projection of multivariate data, play
-an important role in data visualization; they extend the dimensionality
-of visuals while data- and parameter-spaces become ever larger. This
-research has modified the algorithm producing manual tours which and has
-made this functionality available in package \textbf{spinifex}. The
-package adds to \textbf{tourr}, extending the graphics offerings that
-can be used to display tours.
+Tours, which are a dynamic linear projection of numeric multivariate
+data, play an important role in data visualization; they extend the
+dimensionality of visuals while data- and parameter-spaces become ever
+larger. This research has modified the algorithm producing manual tours
+which and has made this functionality available in package
+\textbf{spinifex}. The package adds to \textbf{tourr}, extending the
+graphics offerings that can be used to display tours.
 
@@
 experiments of hadronic collisions. The importance of selecting the
-correct variable to manipulate is demonstrated by comparing tours of
-varying quality. The manual tours convey a better picture of the
-structure of the clustering than static linear or non-linear
-projections. Giving the full contribution of the manipulation space to
-the manip var enables analysts to explore the sensitivity of the
-structure with the selected variable. This information can be used by
-physicists to identify which experiments are probing which parameter
-spaces, which can be indispensable for smarter planning of effort and
-funding.
+correct manip var, as demonstrated by comparing tours of varying amounts
+of structural information. Manual tours, by giving the full contribution
+of the manipulation space to the manip var, enable analysts to explore
+the sensitivity of the structure with the selected variable. This
+information can be used by domain experts to identify which variables
+are probing which parameter spaces and how sensitive those variables are
+to structural features. This insight can be indispensable for variable
+inclusion/exclusions in a model, or in high-level decisions, such as
+budgeting time and funding.
 
@@
 3D projections. The addition of another dimension theoretically allows
-for improved perception. This could explore interactions in immersive
+for improved perception. In some frameworks, such as the game engine
+\textbf{unity}, this would allow for the exploration in immersive
 virtual reality or mixed reality, which may further allow for a better
@@
 dimensionality problem as data while their possible values lie on a
-plane of values rather than discrete points. Occulation, or the closer
+surface of values rather than discrete points. Occulation, or the closer
 surface blocking further surfaces, will likely be an issue that may be
 alleviated by the use of wire mesh, changing opacity, or looking at
-sections of the projections (Furnas and Buja
+sections of the projections known as prosections (Furnas and Buja
 \protect\hyperlink{ref-furnas_prosection_1994}{1994}).
@@
 \textbf{base} graphics while \textbf{spinifex} utilizes \textbf{ggplot2}
-graphics, a significant paradigm shift.
+graphics, a significant paradigm shift, which may have a low return on
+investment.
 
@@
 be added. Currently, Gram-Schmidt is the only form of frame
-interpolation used (not used in manual tours). In a Givens rotation, the
-\(x\) and \(y\) components (for example \(\theta~= 0,~pi/2\)) of the
-in-plane rotation are calculated separately and would be applied
-sequentially to produce the radial rotation. Householder reflections
-define reflection axes to project points on to the axes and generate
-rotations.
+interpolation used. In a Givens rotation, the \(x\) and \(y\) components
+(for example \(\theta~= 0,~pi/2\)) of the in-plane rotation are
+calculated separately and would be applied sequentially to produce the
+radial rotation. Householder reflections define reflection axes to
+project points on to the axes to generate rotations.
 
-Having a script only interaction with tours causes a significant barrier
+Having script-only interaction with tours causes a significant barrier
 to entry. To a lesser extent, \textbf{plotly} offers some static

### 2019-04-26 paper/spinifex_paper.Rmd
remainder of hard copy edits. grammerly, knit.

@@
 
+<!-- interesting features and guided tours -->
 One way to identify "interesting" features is with the use of a guided tour [@cook_grand_1995].
@@
 After a projection of interest is identified an analyst can then use the "finer brush" of the manual tour, by controlling the contributions of individual variables to explore the sensitivity they have on the structure visible in the projection.
-
-<!-- Contributions 
-This paper describes the algorithm used to perform a radial manual tour as implemented in the package \CRANpkg{spinifex}.
-It also describes how to generate an animation of the manual tour using the animation frameworks offered by \CRANpkg{plotly} [@sievert_plotly_2018] and \CRANpkg{gganimate} [@pedersen_gganimate:_2019].
-The method is applied to multivariate data collected on high energy physics experiments (REF).
--->
 
@@
 **spinifex** allows tours to be rendered in **plotly** as an HTML5 object or **gganimate** as GIF or MP4 files.
-Both of which build off **ggplot2** objects in internal functions.
-Sharing of animations is not trivial especially in print and static file formats such as PDF.
-Even with the use of computers and dynamic file formats capturing the correct resolution, aspect, and display is challenging and many formats quickly bloat file sizes.
-Keep in mind hosting options and exporting functions from **plotly**, **gganimate** and **tourr**.
+Sharing of animations is not trivial, especially static formats like print and PDF.
+Even with dynamic display capturing the correct resolution and aspect ratio can be challenging, while many formats quickly bloat file sizes limiting sharing options.
+Keep in mind hosting options and exporting functions offered in **plotly**, **gganimate** and **tourr**.
 
@@
 Because of the change in the basis, or orientation to the subspace, tours are able to better resolve the distribution shape of clusters, intra-cluster detail, and better outlier detection than the use of PDFSense & TFEP (TensorFlow embedded projections) or traditional static embeddings.
-Before applying manual tours let's discuss the structure of the data.
+Before applying manual tours the structure of the data is discussed.
 
@@
 Radial manual tours controlling contributions from PC4 and PC3 are shown in Figure \ref{fig:JetClusterGood} and Figure \ref{fig:JetClusterBad} respectively.
-These variables are selected to contrast the difference of information conveyed by touring different variables.
+These variables are selected to contrast the information conveyed by different manip variables.
 Links to dynamic HTML5 animations controlling each of the four variables are also provided.
@@
 When manipulating PC4, there is a clear difference in the parameter space spanned by the experiment types ATLAS7new and ATLAS7old.
-Specifically, the variation of ATLAS7new becomes more singular.
+Specifically, the variation of ATLAS7new (green) becomes more singular.
 The experiments are probing different parameter space and PC4 is important to demonstrate this.
@@
 During the intermediate frames, the ATLAS7new is compressed in the direction radial to PC4.
-The differece in distribution shape demonstrates the experiments probe different phase-space, which has a linear mapping back to the original dimensions.
+The differece in distribution shape demonstrates the experiments probe different phase-space, which has a linear mapping back to the original variables for interpretation and further exploration.
 An HTML5 version can be viewed at https://nspyrison.netlify.com/thesis/jetcluster\\_manualtour\\_pc4/."}
@@
 PC6 is particularly important to describe the variations of DIS HERA1+2 and charm SIDIS observations,  Whereas manipulating the contributions of PC2 only shows a subset of the dimensionality and shape information.
-Manipulating the contributions of PC6 turned out to be much more insightful than that of PC2.
+Manipulating the contributions of PC6 turned out to be much more insightful than PC2.
 This result might seem counter-intuitive at first as PC2 should explain much more of the variation in the data.
-However, features and structures in the data regularly reside in smaller dimensionality.
-The details finer resolution of these features can be lost when looking only at static projections.
+However, features and structures in the data regularly reside in smaller dimensionality which can be overlooked when optimizing on full sample statistics.
 DIS cluster manual tours manipulating each of the principal components can be viewed from the links: [PC1](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc1/), [PC2](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc2/), [PC3](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc3/), [PC4](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc4/), [PC5](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc5/), and [PC6](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/).
@@
 This jet does extend more in the plane of view when the contribution of PC2 is full, giving insight to its orientation.
-However, less information about the shape of DIS HERA (green) and charm SIDIS (orange) is available.
+However, less information about the shape of DIS HERA (green) and charm SIDIS (orange) is available compared to PC6 as the manip var.
 An HTML5 version can be viewed at https://nspyrison.netlify.com/thesis/discluster\\_manualtour\\_pc2/."}
@@
 
-<!-- summary and segue  -->
-Tours, which are a dynamic linear projection of multivariate data, play an important role in data visualization; they extend the dimensionality of visuals while data- and parameter-spaces become ever larger.
+<!-- summary of spinifex  -->
+Tours, which are a dynamic linear projection of numeric multivariate data, play an important role in data visualization; they extend the dimensionality of visuals while data- and parameter-spaces become ever larger.
 This research has modified the algorithm producing manual tours which and has made this functionality available in package **spinifex**.
@@
 
+<!-- summary of application  -->
 Radial manual tours were applied to a dataset across different experiments of hadronic collisions.
-The importance of selecting the correct variable to manipulate is demonstrated by comparing tours of varying quality.
-The manual tours convey a better picture of the structure of the clustering than static linear or non-linear projections.
-Giving the full contribution of the manipulation space to the manip var enables analysts to explore the sensitivity of the structure with the selected variable.
-This information can be used by physicists to identify which experiments are probing which parameter spaces, which can be indispensable for smarter planning of effort and funding.
+The importance of selecting the correct manip var, as demonstrated by comparing tours of varying amounts of structural information.
+Manual tours, by giving the full contribution of the manipulation space to the manip var, enable analysts to explore the sensitivity of the structure with the selected variable.
+This information can be used by domain experts to identify which variables are probing which parameter spaces and how sensitive those variables are to structural features.
+This insight can be indispensable for variable inclusion/exclusions in a model, or in high-level decisions, such as budgeting time and funding.
 
@@
 The addition of another dimension theoretically allows for improved perception.
-This could explore interactions in immersive virtual reality or mixed reality, which may further allow for a better perception of structure and aid in higher-dimensional function visualization.
-Functions with many parameters suffer from the same dimensionality problem as data while their possible values lie on a plane of values rather than discrete points.
-Occulation, or the closer surface blocking further surfaces, will likely be an issue that may be alleviated by the use of wire mesh, changing opacity, or looking at sections of the projections [@furnas_prosection_1994].
+In some frameworks, such as the game engine **unity**, this would allow for the exploration in immersive virtual reality or mixed reality, which may further allow for a better perception of structure and aid in higher-dimensional function visualization.
+Functions with many parameters suffer from the same dimensionality problem as data while their possible values lie on a surface of values rather than discrete points.
+Occulation, or the closer surface blocking further surfaces, will likely be an issue that may be alleviated by the use of wire mesh, changing opacity, or looking at sections of the projections known as prosections [@furnas_prosection_1994].
 
@@
 Additionally, the \CRANpkg{animation} package @xie_animation:_2018 could be implemented for another graphics framework.
-However, **animation** builds from **base** graphics while **spinifex** utilizes **ggplot2** graphics, a significant paradigm shift.
+However, **animation** builds from **base** graphics while **spinifex** utilizes **ggplot2** graphics, a significant paradigm shift, which may have  a low return on investment.
 
@@
 The Givens rotations and Householder reflections as outlined in @buja_computational_2005 could also be added.
-Currently, Gram-Schmidt is the only form of frame interpolation used (not used in manual tours).
+Currently, Gram-Schmidt is the only form of frame interpolation used.
 In a Givens rotation, the $x$ and $y$ components (for example $\theta~= 0,~pi/2$) of the in-plane rotation are calculated separately and would be applied sequentially to produce the radial rotation.
-Householder reflections define reflection axes to project points on to the axes and generate rotations.
+Householder reflections define reflection axes to project points on to the axes to generate rotations.
 
 <!-- dynamic interaction/gui -->
-Having a script only interaction with tours causes a significant barrier to entry.
+Having script-only interaction with tours causes a significant barrier to entry.
 To a lesser extent, **plotly** offers some static interactions with the contained object, such as tooltips, brushing, and linking without communicating back to the R console.

### 2019-05-03 paper/spinifex-paper.tex
expanded rotation matrix and it's explaination.

@@
 rotation around the \(Z\) axis. While \(\phi\) will be the rotation
-around, \(U\), a new axis on the \(XY\) plane, orthogonal to \(\phi\).
-Given these axes, the rotation matrix can be derived by the following.
+around, \(\textbf{U}\), a new axis on the \(XY\) plane, orthogonal to
+\(\phi\). Given these axes, the SO(3) rotation matrix can be derived by
+the following.
 
@@
   \item[$s_\phi$]   be the sine of   $\phi_i$
+  \item[$\textbf{U}$] be the axis of rotation for $\phi$.
 \end{itemize}
@@
 \begin{align*}
-  \textbf{B}_k &= (c_\theta, s_\theta, 0) \\
-  \vec{u} &= (s_\theta, -c_\theta, 0) \\ &=
+  \textbf{U} &= (u_x, u_y, u_z) =
+  (s_\theta, -c_\theta, 0) \\ 
+  &=
+  \begin{bmatrix}
+  0 & -u_z & u_y  \\
+  u_z & 0 & -u_x \\
+  -u_y & u_x & 0 \\
+  \end{bmatrix} =
   \begin{bmatrix}
@@
   \end{bmatrix} \\
-  \text{Apply Rodrigues' rotation formula:} \\
+  \end{align*} Apply Rodrigues' rotation formula: \begin{align*}
     \textbf{R}_{[3,~3]} 
-    &= \textbf{I}_3 + s_\phi\*\vec{u} + (1-c_\phi)\*\vec{u}^2 \\
+    &= \textbf{I}_3 + s_\phi\*\textbf{U} + (1-c_\phi)\*\textbf{U}^2 \\
+        &=
+    \begin{bmatrix}
+      1&0&0 \\ 0&1&0 \\ 0&0&1 \\
+    \end{bmatrix} +
+    \begin{bmatrix}
+      0 & 0 & c_\theta s_\phi \\
+      0 & 0 & s_\theta s_\phi \\
+      -c_\theta s_\phi & -s_\theta s_\phi & 0 \\
+    \end{bmatrix} +
+    \begin{bmatrix}
+      -c_\theta (1-c_\phi) & s^2_\theta (1-c_\phi) & 0 \\
+      -c_\theta s_\theta (1-c_\phi) & -s^2_\theta (1-c_\phi) & 0 \\
+      0 & 0 & c_\phi-1 \\
+    \end{bmatrix} \\
     &= 
@@
       c_\phi
-    \end{bmatrix}
+    \end{bmatrix} \\
 \end{align*}
 
-Post-multiply the manipulation space by the pre-defined rotation matrix
-producing \textbf{RM}, the rotated manip space. This is one frame of
-basis values, repeat this process for each value in the sequence
-\(\phi_i\) for the complete animation.
+Defining the axis of rotation for \(\phi\) and the application of
+Rogrigues' rotation formula results in the rotation matrix, the rotation
+group SO(3). Continue with the rotation of the manipulation space,
+post-multiply the space by the above rotation matrix producing
+\textbf{RM}, the rotated manip space. This is basis values, the linear
+combination of variables, for one frame of the animation, repeat this
+process for each value in the sequence \(\phi_i\) for the complete
+animation.
 

### 2019-05-03 paper/spinifex_paper.Rmd
expanded rotation matrix and it's explaination.

@@
 Rotations are defined by the axes they pivot on, $\theta$ then is the rotation around the $Z$ axis.
-While $\phi$ will be the rotation around, $U$, a new axis on the $XY$ plane, orthogonal to $\phi$.
-Given these axes, the rotation matrix can be derived by the following.
+While $\phi$ will be the rotation around, $\textbf{U}$, a new axis on the $XY$ plane, orthogonal to $\phi$.
+Given these axes, the SO(3) rotation matrix can be derived by the following.
 
@@
   \end{bmatrix} \\
-  \text{Apply Rodrigues' rotation formula:} \\
+  \end{align*}
+Apply Rodrigues' rotation formula:
+  \begin{align*}
     \textbf{R}_{[3,~3]} 
-    &= \textbf{I}_3 + s_\phi\*\vec{u} + (1-c_\phi)\*\vec{u}^2 \\
+    &= \textbf{I}_3 + s_\phi\*\textbf{U} + (1-c_\phi)\*\textbf{U}^2 \\
+        &=
+    \begin{bmatrix}
+      1&0&0 \\ 0&1&0 \\ 0&0&1 \\
+    \end{bmatrix} +
+    \begin{bmatrix}
+      0 & 0 & c_\theta s_\phi \\
+      0 & 0 & s_\theta s_\phi \\
+      -c_\theta s_\phi & -s_\theta s_\phi & 0 \\
+    \end{bmatrix} +
+    \begin{bmatrix}
+      -c_\theta (1-c_\phi) & s^2_\theta (1-c_\phi) & 0 \\
+      -c_\theta s_\theta (1-c_\phi) & -s^2_\theta (1-c_\phi) & 0 \\
+      0 & 0 & c_\phi-1 \\
+    \end{bmatrix} \\
     &= 
@@
       c_\phi
-    \end{bmatrix}
+    \end{bmatrix} \\
 \end{align*}
+<!-- The term $(1-c_\phi)$ is used by convention, but $2sin^2(\phi/2)$ is more computationally robust. -->
 
 <!-- rotate manip space -->
-Post-multiply the manipulation space by the pre-defined rotation matrix producing **RM**, the rotated manip space.
-This is one frame of basis values, repeat this process for each value in the sequence $\phi_i$ for the complete animation.
+Defining the axis of rotation for $\phi$ and the application of Rogrigues' rotation formula results in the rotation matrix, the rotation group SO(3).
+Continue with the rotation of the manipulation space, post-multiply the space by the above rotation matrix producing **RM**, the rotated manip space.
+This is basis values, the linear combination of variables, for one frame of the animation, repeat this process for each value in the sequence $\phi_i$ for the complete animation.
 

### 2019-05-03 paper/spinifex_paper.Rmd
fix text pos and color of radial manual tour

@@
 
-```{r localFunction, include=FALSE}
-## FUNCTION FOR STATIC OUTPPUT, consider going to scalable facets.
+```{r localFunctions, include=FALSE}
+## FUNCTION FOR STATIC OUTPPUT, 
+# going to facets loses control of output size, simpler to just live with this.
 array2static <- function(.m_tour, .data, .m_var, .cat, .mag =2.2)
@@
   gg2
+}
+
+view_basis2 <- function(basis, labels = paste0("V", 1:nrow(basis)),
+                        manip_var, data = NULL, axes = "center"){
+  basis <- as.data.frame(basis)
+  angle <- seq(0, 2 * pi, length = 360)
+  circ <- data.frame(x = cos(angle), y = sin(angle))
+  zero <- set_axes_position(0, axes)
+  basis <- set_axes_position(basis, axes)
+  circ <- set_axes_position(circ, axes)
+  col <- rep("black", nrow(basis))
+  col[manip_var] <- "blue"
+
+  gg <- ggplot2::ggplot() + ggplot2::scale_color_brewer(palette = "Dark2") + 
+    ggplot2::theme_void() + ggplot2::theme(legend.position = "none") + 
+    ggplot2::coord_fixed()
+  if (axes != "off") {
+    gg <- gg + ggplot2::geom_path(
+      data = circ, mapping = ggplot2::aes(x = x, y = y), 
+      color = "grey80", size = 0.3, inherit.aes = F) + 
+      ggplot2::geom_segment(
+        data = basis, mapping = ggplot2::aes(x = V1, y = V2, xend = zero, yend = zero), color = col) + 
+      ggplot2::geom_text(
+        data = basis, mapping = ggplot2::aes(x = V1, y = V2, label = labels), 
+        size = 4, color = col, hjust = 1.1, vjust=.4)
+  }
+  if (!is.null(data)) {
+        proj <- as.data.frame(tourr::rescale(as.matrix(data) %*% 
+            as.matrix(basis)) - 0.5)
+        gg <- gg + ggplot2::geom_point(data = proj, mapping = ggplot2::aes(x = V1, 
+            y = V2))
+    }
+    gg
 }
@@
   paste0("!!!!! step3 n_slides != 15 !!!!! Please knit again. n_slides = ", dim(f_mtour)[3]))
-
-p1 <- view_basis(f_mtour[,,1], labels=colnames(f_dat)) + ggtitle("1")
-p2 <- view_basis(f_mtour[,,6], labels=colnames(f_dat)) + ggtitle("2")
-p3 <- view_basis(f_mtour[,,12], labels=colnames(f_dat)) + ggtitle("3")
-p4 <- view_basis(f_mtour[,,15], labels=colnames(f_dat)) + ggtitle("4")
+p1 <- view_basis2(f_mtour[,,1],  colnames(f_dat), manip_var = f_mvar) + ggtitle("1")
+p2 <- view_basis2(f_mtour[,,6],  colnames(f_dat), manip_var = f_mvar) + ggtitle("2")
+p3 <- view_basis2(f_mtour[,,12], colnames(f_dat), manip_var = f_mvar) + ggtitle("3")
+p4 <- view_basis2(f_mtour[,,15], colnames(f_dat), manip_var = f_mvar) + ggtitle("4")
 grid.arrange(p1, p2, p3, p4, ncol=4)

### 2019-05-03 paper/spinifex_paper.Rmd
format, cleanup, working on view_manip_space2() locally

@@
 library(gridExtra)
+library(magrittr)
 ```
 
-```{r localFunctions, include=FALSE}
+```{r paperFunctions, include=FALSE}
 ## FUNCTION FOR STATIC OUTPPUT, 
@@
 The shadow of this axis (blue) is its contribution in the original 2D projection."}
+mvar <- as.data.frame(f_msp[manip_var, ])
+mvar <- .6*data.frame(mvar, norm2 = sqrt(mvar$V1^2 + mvar$V2^2))
 
@@
     \begin{bmatrix}
-      1&0&0 \\ 0&1&0 \\ 0&0&1 \\
+      1&0&0 \\ 
+      0&1&0 \\ 
+      0&0&1 \\
     \end{bmatrix} +

### 2019-05-06 paper/spinifex_paper.Rmd
Add angle notation to view_manip_space2().

@@
 library(gridExtra)
-library(magrittr)
+library(dplyr)
 ```
@@
     }
+    gg
+}
+
+view_manip_space2 <- function (basis,
+                               manip_var,
+                               manip_col = "blue",
+                               tilt = pi *
+                                 5 / 12,
+                               z_col = "red",
+                               labels = paste0("V", 1:nrow(basis)))
+{
+    find_angle <-
+      function(a, b)
+        acos(sum(a * b) / (sqrt(sum(a * a)) * sqrt(sum(b * b))) )
+    make_curve <-
+      function(ang_st = 0,
+               ang_stop = 2 * pi) {
+        # angle is atan(y/x)
+        angle <- seq(ang_st, ang_stop,
+                     length = round(360 / (2 * pi) * abs(ang_st - ang_stop)) )
+        data.frame(x = cos(angle), y = sin(angle), z = sin(angle))
+      }
+    xyz <- function(x) { colnames(x)<-c("x","y","z");x }
+
+    p <- nrow(basis)
+    col_v <- rep("grey80", p)
+    col_v[manip_var] <- manip_col
+    siz_v <- rep(0.3, p)
+    siz_v[manip_var] <- 1
+
+    circ_r <- make_curve() %>%
+      mutate( y=y*cos(tilt), z=z*sin(tilt))
+    m_sp   <- xyz(as.data.frame(create_manip_space(basis, manip_var)))
+    m_sp_r <- m_sp %>% mutate( y=y*cos(tilt), z=z*sin(tilt))
+    mvar   <- m_sp[manip_var, ]
+    mvar_r <- m_sp_r[manip_var, ]
+
+    phi_ang     <- find_angle(c(m_sp_r$x, m_sp_r$y), c(m_sp_r$x, m_sp_r$z))
+    phi_offset  <- find_angle(c(mvar_r$x, mvar_r$y), c(1,0))
+    phi_curve_r <- .4*make_curve(ang_st = phi_offset, ang_stop = phi_ang) %>%
+      mutate( y=y*cos(tilt), z=z*sin(tilt))
+
+    thata_ang     <- find_angle(c(mvar$x, mvar$y),c(1, 0))
+    theta_curve_r <- .5*make_curve(ang_st = 0, ang_stop = thata_ang) %>%
+      mutate( y=y*cos(tilt), z=z*sin(tilt))
+
+    gg <-
+      ggplot2::ggplot() + ggplot2::scale_color_brewer(palette = "Dark2") +
+      ggplot2::theme_void() + ggplot2::theme(legend.position = "none") +
+      ggplot2::coord_fixed() +
+        ggplot2::geom_path(data = circ_r,
+                           mapping = ggplot2::aes(x = x, y = y),
+                           color = manip_col,
+                           size = 0.3,
+                           inherit.aes = F
+        ) +
+        #projection plane:
+        ggplot2::geom_segment(data = m_sp_r,
+                              mapping = ggplot2::aes(
+                                x = x,
+                                y = y,
+                                xend = 0,
+                                yend = 0
+                              ),
+                              size = siz_v,
+                              colour = col_v
+        ) +
+        ggplot2::geom_text(data = m_sp_r,
+                           mapping = ggplot2::aes(x = x, y = y, label = labels),
+                           size = 4,
+                           colour = col_v,
+                           vjust = "outward",
+                           hjust = "outward"
+        ) +
+        #Z direction
+        ggplot2::geom_path(data = circ_r,
+                           mapping = ggplot2::aes(x = x, y = z),
+                           color = z_col,
+                           size = 0.3,
+                           inherit.aes = F
+        ) +
+        ggplot2::geom_segment(data = mvar_r,
+                              mapping = ggplot2::aes(x = x,y = z,
+                                                     xend = 0,yend = 0),
+                              size = 1,
+                              colour = z_col
+        ) +
+        ggplot2::geom_segment(data = mvar_r,
+                              mapping = ggplot2::aes(x = x, y = z,
+                                                     xend = x,yend = y),
+                              size = 0.3,
+                              colour = "grey80",
+                              linetype = 2
+        ) +
+        ggplot2::geom_text(data = mvar_r,
+                           mapping = ggplot2::aes(x = x, y = z, label = labels[manip_var]),
+                           size = 4,
+                           colour = z_col,
+                           vjust = "outward",
+                           hjust = "outward"
+        ) +
+        # label angles
+        geom_path(data = phi_curve_r,
+                  mapping = aes(x=x, y=z),
+                  color = z_col,
+                  size = 0.2
+        ) +
+        geom_text(data = 1.2 * phi_curve_r[ceiling(nrow(phi_curve_r)/2), ],
+                  mapping = ggplot2::aes(x=x, y=z, label = "phi"),
+                  color = z_col,
+                  parse = T,
+                  size = 4
+        ) +
+        geom_path(data = theta_curve_r,
+                  mapping = aes(x, y),
+                  color = manip_col,
+                  size = 0.2
+        ) +
+        geom_text(data = 1.2 * theta_curve_r[ceiling(nrow(theta_curve_r)/2), ],
+                  mapping = ggplot2::aes(x=x, y=y-.02, label = "theta"),
+                  color = manip_col,
+                  parse = T,
+                  size = 4
+        )
     gg
@@
 The shadow of this axis (blue) is its contribution in the original 2D projection."}
-mvar <- as.data.frame(f_msp[manip_var, ])
-mvar <- .6*data.frame(mvar, norm2 = sqrt(mvar$V1^2 + mvar$V2^2))
-
-view_manip_space(basis = f_bas, manip_var = f_mvar, labels = colnames(f_dat)) +
+view_manip_space2(basis = f_bas, manip_var = f_mvar, labels = colnames(f_dat)) +
     theme(panel.border = element_rect(colour = "black", fill = NA))

### 2019-05-10 paper/spinifex-paper.tex
Add angle notation to view_manip_space2().
algorithm details

@@
 \item
-  \(\theta_i\), the sequence of angles of in-projection-plane rotation,
-  for example, on the reference axes.
+  \(\theta\), the angle of in-projection-plane rotation, for example, on
+  the reference axes; \(c_\theta, s_\theta\) are the cosine and sine.
 \item
-  \(\phi_i\), the sequence of angles of out-of-projection-plane
-  rotation, coming into the manipulation space.
+  \(\phi\), the angle of out-of-projection-plane rotation, into the
+  manipulation space; \(c_\phi, s_\phi\) are the cosine and sine.
+\item
+  \(\textbf{U}\) is the axis of rotation for \(\phi\).
 \end{itemize}
@@
 
+\hypertarget{steps}{%
+\subsection{Steps}\label{steps}}
+
 \hypertarget{step-0-set-up}{%
-\subsection{Step 0) Set up}\label{step-0-set-up}}
+\subsubsection{Step 0) Set up}\label{step-0-set-up}}
 
@@
 \hypertarget{step-1-choose-manip-variable}{%
-\subsection{Step 1) Choose manip
+\subsubsection{Step 1) Choose manip
 variable}\label{step-1-choose-manip-variable}}
@@
 \hypertarget{step-2-create-the-3d-manip-space}{%
-\subsection{Step 2) Create the 3D manip
+\subsubsection{Step 2) Create the 3D manip
 space}\label{step-2-create-the-3d-manip-space}}
@@
 \hypertarget{step-3-generate-rotation}{%
-\subsection{Step 3) Generate rotation}\label{step-3-generate-rotation}}
+\subsubsection{Step 3) Generate
+rotation}\label{step-3-generate-rotation}}
 
@@
 
-Let:
-
-\begin{itemize}
-  \item[$c_\theta$] be the cosine of $\theta$
-  \item[$c_\phi$]   be the cosine of $\phi_i$
-  \item[$s_\theta$] be the sine of   $\theta$
-  \item[$s_\phi$]   be the sine of   $\phi_i$
-  \item[$\textbf{U}$] be the axis of rotation for $\phi$.
-\end{itemize}
-
 \begin{align*}
@@
 
-A note on application: \(\phi\) is the angle relative to the initial
-value of \(\phi\), we find the transformation \(\phi_i\) - \(\phi_1\)
-useful to think about \(\phi\) relative to the basis plane.
-Additionally, the value of \(\phi\) may be out of phase by a factor of
-pi. If the manip variable doesn't move as expected these are the first
-places to check.
+\hypertarget{creating-the-animation}{%
+\subsection{Creating the animation}\label{creating-the-animation}}
+
+The steps outlined above can be used to create any arbitrary rotation in
+the manip space. To use these for sensitivity analysis, the radial
+rotation is built into an animation where the manip variable is rotated
+fully into the projection, completely out, and then back to the start.
+This involves changing \(\phi\), the angle relative to its initial
+value.
+
+When \(\phi\) is XXX the manip var is completely in the projection, and
+when \(\phi\) is not contributing at all.
+
+\hypertarget{sequence-of-phi_i}{%
+\subsubsection{\texorpdfstring{Sequence of
+\(\phi_i\)}{Sequence of \textbackslash phi\_i}}\label{sequence-of-phi_i}}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Set initial value to be \(\phi_1\).
+\item
+  Set an angle value (\(\Delta_\phi\)) that sets the step size to make
+  in rotating the manip var into and out of the projection. (Note: Using
+  angle value to control this is consistent with the tour algorithm, as
+  implemented in the \CRANpkg{tourr}. )
+\item
+  Step towards \(0\), where the manip var has a completely in the
+  projection.
+\item
+  Step towards XXX where the manip variable has no contribution.
+\item
+  Step back to \(\phi_1\).
+\end{enumerate}
+
+In each of steps 3-5, a small step may be made to bring it exactly to
+the end position.
+
+\hypertarget{checks-that-need-to-be-made}{%
+\subsubsection{Checks that need to be
+made}\label{checks-that-need-to-be-made}}
+
+find the transformation \(\phi_i\) - \(\phi_1\) useful to think about
+\(\phi\) relative to the basis plane. Additionally, the value of
+\(\phi\) may be out of phase by a factor of pi. If the manip variable
+doesn't move as expected these are the first places to check.
 
@@
 \hypertarget{sec:display}{%
-\section{Data in projection-space}\label{sec:display}}
+\subsection{Data in projection-space}\label{sec:display}}
 

### 2019-05-10 paper/spinifex_paper.Rmd
algorithm details

@@
 - $\textbf{e}$, a 1D coordinate basis vector of length $p$ with the $k-$th element set to one, where $k$ is the index of the variable to manipulate, called the "manip var".
-- $\theta_i$, the sequence of angles of in-projection-plane rotation, for example, on the reference axes.
-- $\phi_i$, the sequence of angles of out-of-projection-plane rotation, coming into the manipulation space.
+- $\theta$, the angle of in-projection-plane rotation, for example, on the reference axes; $c_\theta, s_\theta$ are the cosine and sine.
+- $\phi$, the angle of out-of-projection-plane rotation, into the manipulation space; $c_\phi, s_\phi$ are the cosine and sine.
+- $\textbf{U}$ is the axis of rotation for $\phi$.
 
@@
 
-## Step 0) Set up
+## Steps
+### Step 0) Set up
 
@@
 
-## Step 1) Choose manip variable
+### Step 1) Choose manip variable
 
@@
 
-## Step 2) Create the 3D manip space
+### Step 2) Create the 3D manip space
 
@@
 
-## Step 3) Generate rotation
+### Step 3) Generate rotation
 
@@
 Given these axes, the SO(3) rotation matrix can be derived by the following.
-
-Let:
-\begin{itemize}
-  \item[$c_\theta$] be the cosine of $\theta$
-  \item[$c_\phi$]   be the cosine of $\phi_i$
-  \item[$s_\theta$] be the sine of   $\theta$
-  \item[$s_\phi$]   be the sine of   $\phi_i$
-  \item[$\textbf{U}$] be the axis of rotation for $\phi$.
-\end{itemize}
 
@@
 
-<!--  phi transform and application note -->
-A note on application: $\phi$ is the angle relative to the initial value of $\phi$, we find the transformation $\phi_i$ - $\phi_1$ useful to think about $\phi$ relative to the basis plane.
+## Creating the animation
+
+<!--  phi transform and animation -->
+The steps outlined above can be used to create any arbitrary rotation in the manip space.
+To use these for sensitivity analysis, the radial rotation is built into an animation where the manip variable is rotated fully into the projection, completely out, and then back to the start.
+This involves changing $\phi$, the angle relative to its initial value.
+
+When $\phi$ is XXX the manip var is completely in the projection, and when $\phi$ is not contributing at all.
+
+### Sequence of $\phi_i$
+
+1. Set initial value to be $\phi_1$.
+2.
+Set an angle value ($\Delta_\phi$) that sets the step size to make in rotating the manip var into and out of the projection.
+(Note: Using angle value to control this is consistent with the tour algorithm, as implemented in the \CRANpkg{tourr}.
+)
+3. Step towards $0$, where the manip var has a completely in the projection.
+4. Step towards XXX where the manip variable has no contribution.
+5. Step back to $\phi_1$.
+
+In each of steps 3-5, a small step may be made to bring it exactly to the end position.
+
+### Checks that need to be made
+
+find the transformation $\phi_i$ - $\phi_1$ useful to think about $\phi$ relative to the basis plane.
 Additionally, the value of $\phi$ may be out of phase by a factor of pi.
@@
 
-# Data in projection-space {#sec:display}
+## Data in projection-space {#sec:display}
 

### 2019-05-10 paper/spinifex-paper.tex
\pkg and \code, Rodrigues rotation citation, consistency

@@
 collectively known as \textbf{tours} provide an important tool for
-exploring multivariate data and models. The R package \textbf{tourr}
+exploring multivariate data and models. The R package \pkg{tourr}
 provides functions for several types of tours: grand, guided, little,
@@
 a data object for animation. This paper describes a new package,
-\textbf{spinifex}, to provide a manual tour, that allows the coefficient
+\pkg{spinifex}, to provide a manual tour, that allows the coefficient
 for a single variable to be controlled. The variable is rotated fully
@@
 projections can be displayed using animation, with functions from either
-the \textbf{plotly} and \textbf{gganimate} packages. By varying the
+the \pkg{plotly} and \pkg{gganimate} packages. By varying the
 coefficient of a single variable, it is possible to explore the
@@
 \item
+  \(k\), is the index of the variable to manipulate, called the ``manip
+  var''.
+\item
+  \(|\textbf{B}_k|\), the norm of the manip var, \(k\) 
+\item
   \(\textbf{e}\), a 1D coordinate basis vector of length \(p\) with the
-  \(k-\)th element set to one, where \(k\) is the index of the variable
-  to manipulate, called the ``manip var''.
+  \(k\)-th element set to one.
 \item
@@
 (\protect\hyperlink{ref-lubischew_use_1962}{1962}), made available in
-\textbf{tourr} is used to illustrate the algorithm. The data contains 74
+\pkg{tourr} is used to illustrate the algorithm. The data contains 74
 observations across 6 variables, which are physical measurements of the
@@
 
-Call \texttt{view\_basis()} on a basis to produce a \CRANpkg{ggplot2}
+Call \code{view\_basis()} on a basis to produce a \CRANpkg{ggplot2}
 (Wickham \protect\hyperlink{ref-wickham_ggplot2:_2016}{2016}) graphic
@@
 
-In Figure \ref{fig:step0} the contribution of the variables
-\texttt{tars1} and \texttt{aede2} mostly contrast the contribution of
-the other four variables. These two variables combined contribute in the
-direction of the projection where the purple cluster is separated from
-the other two clusters. The variable \texttt{aede2} is selected as the
-manip var, the variable to be controlled in the tour. The question that
-will be explored is how important this variable is to the separation of
-the clusters.
+In Figure \ref{fig:step0} the contribution of the variables tars1 and
+aede2 mostly contrast the contribution of the other four variables.
+These two variables combined contribute in the direction of the
+projection where the purple cluster is separated from the other two
+clusters. The variable aede2 is selected as the manip var, the variable
+to be controlled in the tour. The question that will be explored is how
+important this variable is to the separation of the clusters.
 
@@
 
-Initialize the coordinate basis vector as a zero vector \textbf{e} of
-length \emph{p}, and set the \(k^{th}\) element to 1. In the example
-data, \texttt{aede2} is the fifth variable in the data, so \(k=5\). Use
+Initialize the coordinate basis vector as a zero vector \(\textbf{e}\)
+of length \(p\), and set the \(k\)-th element to 1. In the example data,
+aede2 is the fifth variable in the data, so \(k=5\), set \(e_5=1\) . Use
 a Gram-Schmidt process to orthonormalize the coordinate basis vector on
 the original 2D projection to describe a 3D manipulation space,
-\textbf{M}.
+\(\textbf{M}\).
 
 \begin{align*}
-  \textbf{e}_k &\leftarrow 1 \\ 
+  e_k &\leftarrow 1 \\ 
   \textbf{e}^*   &\leftarrow \textbf{e} - \langle \textbf{e}, \textbf{B}_1 \rangle \textbf{B}_1 - \langle \textbf{e}, \textbf{B}_2 \rangle \textbf{B}_2 \\ 
@@
 var highlighted. This representation is produced by calling the
-\texttt{view\_manip\_space()} function. Functionally, the manipulation
+\code{view\_manip\_space()} function. Functionally, the manipulation
 space is more of a mathematical construct that needn't be visualized but
@@
   \end{bmatrix} \\
-  \end{align*} Apply Rodrigues' rotation formula: \begin{align*}
+  \end{align*} Apply Rodrigues' rotation
+\href{mailto:formula\%7B@rodrigues_lois_1840}{\nolinkurl{formula\{@rodrigues\_lois\_1840}}\}:
+\begin{align*}
     \textbf{R}_{[3,~3]} 
@@
 combination of variables, for one frame of the animation, repeat this
-process for each value in the sequence \(\phi_i\) for the complete
-animation.
+process over a number of values of \(\phi\) to produce an animation.
 
@@
 
-When \(\phi\) is XXX the manip var is completely in the projection, and
-when \(\phi\) is not contributing at all.
+When \(\phi\) is 0, the manip var is completely in the projection, and
+when \(\phi\) is \(\pi\), \(\textbf{X}_k\) is not contributing at all.
 
@@
   angle value to control this is consistent with the tour algorithm, as
-  implemented in the \CRANpkg{tourr}. )
+  implemented in the \CRANpkg{tourr}).
 \item
@@
 \item
-  Step towards XXX where the manip variable has no contribution.
+  Step towards \(\pi/2\) where the manip variable has no contribution.
 \item
@@
 page may take a moment to load. Animations can be produced using the
-function \texttt{play\_manual\_tour()}.
+function \code{play\_manual\_tour()}.
 
@@
 functions are called by the tour animation functions, passed to the
-arguments \texttt{render\_type}. Usage is demonstrated below and in the
+argument \code{render\_type}. Usage is demonstrated below and in the
 example documentation.
@@
 projection, start by observing the reference axes of the basis with
-\texttt{view\_basis()} producing Figure \ref{fig:step0}.
+\code{view\_basis()} producing Figure \ref{fig:step0}.
 
@@
 After becoming familiar with this space, select a manip var, the
-variable to change the contributions of. Use
-\texttt{view\_manip\_space()} to view the new space with a dimension
-orthogonal to the projection plane where the manip var has a full
-contribution. This illustrates how the manip var is manipulated with the
-addition of the manipulation space as shown in Figure \ref{fig:step2}.
+variable to change the contributions of. Use \code{view\_manip\_space()}
+to view the new space with a dimension orthogonal to the projection
+plane where the manip var has a full contribution. This illustrates how
+the manip var is manipulated with the addition of the manipulation space
+as shown in Figure \ref{fig:step2}.
 
@@
 Now we are ready to perform a manual tour on the selected variable. Use
-\texttt{play\_manual\_tour()} to perform the algorithm as discussed
-above, in section \ref{sec:algorithm}. This is the animated equivalent
-of Figure \ref{fig:step3}.
+\code{play\_manual\_tour()} to perform the algorithm as discussed above,
+in section \ref{sec:algorithm}. This is the animated equivalent of
+Figure \ref{fig:step3}.
 
@@
 This concludes the content of the algorithm section, however, lets cover
-animating paths generated in \textbf{tourr}. Animate the previously
-generated guided tour path via \texttt{play\_tour\_path()}. Utility
+animating paths generated in \pkg{tourr}. Animate the previously
+generated guided tour path via \code{play\_tour\_path()}. Utility
 functions can also be passed as arguments into either of the tour
 animation function to change the resulting graphics object, set
-\texttt{render\_type\ =\ render\_gganimate} to view the animation as a
-GIF.
+\code{render\_type = render\_gganimate} to view the animation as a GIF.
 
@@
 
-The \textbf{tourr} package utilizes \textbf{base} graphics for the
-display of tours. \textbf{spinifex} allows tours to be rendered in
-\textbf{plotly} as an HTML5 object or \textbf{gganimate} as GIF or MP4
-files. Sharing of animations is not trivial, especially static formats
-like print and PDF. Even with dynamic display capturing the correct
-resolution and aspect ratio can be challenging, while many formats
-quickly bloat file sizes limiting sharing options. Keep in mind hosting
-options and exporting functions offered in \textbf{plotly},
-\textbf{gganimate} and \textbf{tourr}.
+The \pkg{tourr} package utilizes \pkg{base} graphics for the display of
+tours. \pkg{spinifex} allows tours to be rendered in \pkg{plotly} as an
+HTML5 object or \pkg{gganimate} as GIF or MP4 files. Sharing of
+animations is not trivial, especially static formats like print and PDF.
+Even with dynamic display capturing the correct resolution and aspect
+ratio can be challenging, while many formats quickly bloat file sizes
+limiting sharing options. Keep in mind hosting options and exporting
+functions offered in \pkg{plotly}, \pkg{gganimate} and \pkg{tourr}.
 
@@
 between can be interpolated by supplying angular speed. With the use of
-the \texttt{tourr::save\_history()} function, the target bases can be
+the \code{tourr::save\_history()} function, the target bases can be
 saved. From there geodesic interpolation can be used to populate the
@@
 which and has made this functionality available in package
-\textbf{spinifex}. The package adds to \textbf{tourr}, extending the
-graphics offerings that can be used to display tours.
+\pkg{spinifex}. The package adds to \pkg{tourr}, extending the graphics
+offerings that can be used to display tours.
 
@@
 
-The \textbf{tourr} package provides many other geometric displays with
-the \texttt{tourr::display\_*()} family. These geometric options could
-be integrated into the \textbf{ggplot2} framework for display on
-\textbf{plotly} and \textbf{gganimate}. Additionally, the
-\CRANpkg{animation} package Xie et al.
-(\protect\hyperlink{ref-xie_animation:_2018}{2018}) could be implemented
-for another graphics framework. However, \textbf{animation} builds from
-\textbf{base} graphics while \textbf{spinifex} utilizes \textbf{ggplot2}
-graphics, a significant paradigm shift, which may have a low return on
-investment.
+The \pkg{tourr} package provides many other geometric displays with the
+\code{tourr::display\_*()} family. These geometric options could be
+integrated into the \pkg{ggplot2} framework for display on \pkg{plotly}
+and \pkg{gganimate}. Additionally, the \CRANpkg{animation} package Xie
+et al. (\protect\hyperlink{ref-xie_animation:_2018}{2018}) could be
+implemented for another graphics framework. However, \pkg{animation}
+builds from \pkg{base} graphics while \pkg{spinifex} utilizes
+\pkg{ggplot2} graphics, a significant paradigm shift, which may have a
+low return on investment.
 
@@
 Having script-only interaction with tours causes a significant barrier
-to entry. To a lesser extent, \textbf{plotly} offers some static
+to entry. To a lesser extent, \pkg{plotly} offers some static
 interactions with the contained object, such as tooltips, brushing, and
@@
 \href{https://github.com/nspyrison/spinifex_paper/}{github.com/nspyrison/spinifex\_paper/}.
-The source code for the \textbf{spinifex} package can be found at
+The source code for the \pkg{spinifex} package can be found at
 \href{https://github.com/nspyrison/spinifex/}{github.com/nspyrison/spinifex/}.

### 2019-05-10 paper/spinifex_paper.Rmd
Auto stash before revert of "algorithm details"
\pkg and \code, Rodrigues rotation citation, consistency

@@
   Dynamic low-dimensional linear projections of multivariate data collectively known as **tours** provide an important tool for exploring multivariate data and models.
-The R package **tourr** provides functions for several types of tours: grand, guided, little, local and frozen.
+The R package \pkg{tourr} provides functions for several types of tours: grand, guided, little, local and frozen.
 Each of these can be viewed dynamically, or saved into a data object for animation.
-This paper describes a new package, **spinifex**, to provide a manual tour, that allows the coefficient for a single variable to be controlled.
+This paper describes a new package, \pkg{spinifex}, to provide a manual tour, that allows the coefficient for a single variable to be controlled.
 The variable is rotated fully into the projection with a coefficient of 1 or -1, or completely out of the projection with a coefficient of 0.
-The resulting sequence of projections can be displayed using animation, with functions from either the **plotly** and **gganimate** packages.
+The resulting sequence of projections can be displayed using animation, with functions from either the \pkg{plotly} and \pkg{gganimate} packages.
 By varying the coefficient of a single variable, it is possible to explore the sensitivity of structure in the projection to that variable.
@@
 We will call this the "original projection" because it is the starting point for the manual tour.
-- $\textbf{e}$, a 1D coordinate basis vector of length $p$ with the $k-$th element set to one, where $k$ is the index of the variable to manipulate, called the "manip var".
+- $k$, is the index of the variable to manipulate, called the "manip var".
+- $|\textbf{B}_k|$, the norm of the manip var, $k$
+<!-- - $\textbf{B}_k$, the manipulation variable (manip var), the variable whoes contribution will be controled. -->
+- $\textbf{e}$, a 1D coordinate basis vector of length $p$ with the $k$-th element set to one. 
 - $\theta$, the angle of in-projection-plane rotation, for example, on the reference axes; $c_\theta, s_\theta$ are the cosine and sine.
@@
 <!-- describe data. -->
-The flea data, originally from @lubischew_use_1962, made available in **tourr**  is used to illustrate the algorithm.
+The flea data, originally from @lubischew_use_1962, made available in \pkg{tourr} is used to illustrate the algorithm.
 The data contains 74 observations across 6 variables, which are physical measurements of the flea beetles.
@@
 
-Call `view_basis()` on a basis to produce a \CRANpkg{ggplot2} [@wickham_ggplot2:_2016] graphic similar to Figure \ref{fig:step0}.
+Call \code{view\_basis()} on a basis to produce a \CRANpkg{ggplot2} [@wickham_ggplot2:_2016] graphic similar to Figure \ref{fig:step0}.
 The projected data is processed for display using matrix multiplication $\textbf{X}_{[n,~p]} ~*~ \textbf{B}_{[p,~d]} ~=~ \textbf{P}_{[n,~d]}$.
@@
 <!-- select a manip var-->
-In Figure \ref{fig:step0} the contribution of the variables `tars1` and `aede2` mostly contrast the contribution of the other four variables.
+In Figure \ref{fig:step0} the contribution of the variables tars1 and aede2 mostly contrast the contribution of the other four variables.
 These two variables combined contribute in the direction of the projection where the purple cluster is separated from the other two clusters.
-The variable `aede2` is selected as the manip var, the variable to be controlled in the tour.
+The variable aede2 is selected as the manip var, the variable to be controlled in the tour.
 The question that will be explored is how important this variable is to the separation of the clusters.
@@
 <!-- Zero Vect, manip sp -->
-Initialize the coordinate basis vector as a zero vector **e** of length *p*, and set the $k^{th}$ element to 1.
-In the example data, `aede2`  is the fifth variable in the data, so $k=5$.
-Use a Gram-Schmidt process to orthonormalize the coordinate basis vector on the original 2D projection to describe a 3D manipulation space, **M**.
+Initialize the coordinate basis vector as a zero vector $\textbf{e}$ of length $p$, and set the $k$-th element to 1.
+In the example data, aede2 is the fifth variable in the data, so $k=5$, set $e_5=1$ .
+Use a Gram-Schmidt process to orthonormalize the coordinate basis vector on the original 2D projection to describe a 3D manipulation space, $\textbf{M}$.
 
 \begin{align*}
-  \textbf{e}_k &\leftarrow 1 \\ 
+  e_k &\leftarrow 1 \\ 
   \textbf{e}^*   &\leftarrow \textbf{e} - \langle \textbf{e}, \textbf{B}_1 \rangle \textbf{B}_1 - \langle \textbf{e}, \textbf{B}_2 \rangle \textbf{B}_2 \\ 
@@
 Figure \ref{fig:step2} illustrates this 3D manipulation space with the manip var highlighted.
-This representation is produced by calling the `view_manip_space()` function.
+This representation is produced by calling the \code{view\_manip\_space()} function.
 Functionally, the manipulation space is more of a mathematical construct that needn't be visualized but this diagram provides intuition for the technique.
@@
   \end{align*}
-Apply Rodrigues' rotation formula:
+Apply Rodrigues' rotation formula{@rodrigues_lois_1840}:
   \begin{align*}
@@
 Continue with the rotation of the manipulation space, post-multiply the space by the above rotation matrix producing **RM**, the rotated manip space.
-This is basis values, the linear combination of variables, for one frame of the animation, repeat this process for each value in the sequence $\phi_i$ for the complete animation.
+This is basis values, the linear combination of variables, for one frame of the animation, repeat this process over a number of values of $\phi$ to produce an animation.
 
@@
 
-When $\phi$ is XXX the manip var is completely in the projection, and when $\phi$ is not contributing at all.
+When $\phi$ is 0, the manip var is completely in the projection, and when $\phi$ is $\pi$, $\textbf{X}_k$ is not contributing at all.
 
@@
 Set an angle value ($\Delta_\phi$) that sets the step size to make in rotating the manip var into and out of the projection.
-(Note: Using angle value to control this is consistent with the tour algorithm, as implemented in the \CRANpkg{tourr}.
-)
+(Note: Using angle value to control this is consistent with the tour algorithm, as implemented in the \CRANpkg{tourr}).
 3. Step towards $0$, where the manip var has a completely in the projection.
-4. Step towards XXX where the manip variable has no contribution.
+4. Step towards $\pi/2$ where the manip variable has no contribution.
 5. Step back to $\phi_1$.
@@
 The page may take a moment to load.
-Animations can be produced using the function `play_manual_tour()`.
+Animations can be produced using the function \code{play\_manual\_tour()}.
 
@@
 In the following section describes and demonstrates the usage of the above functions following the order in the algorithm.
-The utility functions are called by the tour animation functions, passed to the arguments `render_type`.
+The utility functions are called by the tour animation functions, passed to the argument \code{render\_type}.
 Usage is demonstrated below and in the example documentation.
@@
 We'll start by initializing values including a standardized data set (numeric columns only), a starting basis, a categorical variable for point aesthetics (optional), and a manip var.
-To get bearings on the projection, start by observing the reference axes of the basis with `view_basis()` producing Figure \ref{fig:step0}.
+To get bearings on the projection, start by observing the reference axes of the basis with \code{view\_basis()} producing Figure \ref{fig:step0}.
 
@@
 After becoming familiar with this space, select a manip var, the variable to change the contributions of.
-Use `view_manip_space()` to view the new space with a dimension orthogonal to the projection plane where the manip var has a full contribution.
+Use \code{view\_manip\_space()} to view the new space with a dimension orthogonal to the projection plane where the manip var has a full contribution.
 This illustrates how the manip var is manipulated with the addition of the manipulation space as shown in Figure \ref{fig:step2}.
@@
 Now we are ready to perform a manual tour on the selected variable.
-Use `play_manual_tour()` to perform the algorithm as discussed above, in section  \ref{sec:algorithm}.
+Use \code{play\_manual\_tour()} to perform the algorithm as discussed above, in section \ref{sec:algorithm}.
 This is the animated equivalent of Figure \ref{fig:step3}.
@@
 
-This concludes the content of the algorithm section, however, lets cover animating paths generated in **tourr**.
-Animate the previously generated guided tour path via `play_tour_path()`.
-Utility functions can also be passed as arguments into either of the tour animation function to change the resulting graphics object, set `render_type = render_gganimate` to view the animation as a GIF.
+This concludes the content of the algorithm section, however, lets cover animating paths generated in \pkg{tourr}.
+Animate the previously generated guided tour path via \code{play\_tour\_path()}.
+Utility functions can also be passed as arguments into either of the tour animation function to change the resulting graphics object, set \code{render\_type = render\_gganimate} to view the animation as a GIF.
 
@@
 
-The **tourr** package utilizes **base** graphics for the display of tours.
-**spinifex** allows tours to be rendered in **plotly** as an HTML5 object or **gganimate** as GIF or MP4 files.
+The \pkg{tourr} package utilizes \pkg{base} graphics for the display of tours.
+\pkg{spinifex} allows tours to be rendered in \pkg{plotly} as an HTML5 object or \pkg{gganimate} as GIF or MP4 files.
 Sharing of animations is not trivial, especially static formats like print and PDF.
 Even with dynamic display capturing the correct resolution and aspect ratio can be challenging, while many formats quickly bloat file sizes limiting sharing options.
-Keep in mind hosting options and exporting functions offered in **plotly**, **gganimate** and **tourr**.
+Keep in mind hosting options and exporting functions offered in \pkg{plotly}, \pkg{gganimate} and \pkg{tourr}.
 
@@
 The frames in between can be interpolated by supplying angular speed.
-With the use of the `tourr::save_history()` function, the target bases can be saved.
+With the use of the \code{tourr::save\_history()} function, the target bases can be saved.
 From there geodesic interpolation can be used to populate the intermittent frames.
@@
 Tours, which are a dynamic linear projection of numeric multivariate data, play an important role in data visualization; they extend the dimensionality of visuals while data- and parameter-spaces become ever larger.
-This research has modified the algorithm producing manual tours which and has made this functionality available in package **spinifex**.
-The package adds to **tourr**, extending the graphics offerings that can be used to display tours.
+This research has modified the algorithm producing manual tours which and has made this functionality available in package \pkg{spinifex}.
+The package adds to \pkg{tourr}, extending the graphics offerings that can be used to display tours.
 
@@
 <!-- geometric and graphics display -->
-The **tourr** package provides many other geometric displays with the `tourr::display_*()` family.
-These geometric options could be integrated into the **ggplot2** framework for display on **plotly** and **gganimate**.
+The \pkg{tourr} package provides many other geometric displays with the \code{tourr::display\_*()} family.
+These geometric options could be integrated into the \pkg{ggplot2} framework for display on \pkg{plotly} and \pkg{gganimate}.
 Additionally, the \CRANpkg{animation} package @xie_animation:_2018 could be implemented for another graphics framework.
-However, **animation** builds from **base** graphics while **spinifex** utilizes **ggplot2** graphics, a significant paradigm shift, which may have  a low return on investment.
+However, \pkg{animation} builds from \pkg{base} graphics while \pkg{spinifex} utilizes \pkg{ggplot2} graphics, a significant paradigm shift, which may have  a low return on investment.
 
@@
 Having script-only interaction with tours causes a significant barrier to entry.
-To a lesser extent, **plotly** offers some static interactions with the contained object, such as tooltips, brushing, and linking without communicating back to the R console.
+To a lesser extent, \pkg{plotly} offers some static interactions with the contained object, such as tooltips, brushing, and linking without communicating back to the R console.
 The development of a dynamic graphical user interface, perhaps with the use of a \CRANpkg{shiny} [@chang_shiny:_2018] application, would mitigate the barrier to entry, allow for more rapid analysis, and offer an approachable demo tool.
@@
 The source files for this article be found at [github.com/nspyrison/spinifex_paper/](https://github.com/nspyrison/spinifex_paper/).
-The source code for the **spinifex** package can be found at [github.com/nspyrison/spinifex/](https://github.com/nspyrison/spinifex/).
+The source code for the \pkg{spinifex} package can be found at [github.com/nspyrison/spinifex/](https://github.com/nspyrison/spinifex/).
 

### 2019-05-13 paper/spinifex-paper.tex
functions in figure captions,

@@
 The contribution of the variables aede2 and tars1 approximately contrasts the other variables.
-The visible structure in the projection are the three clusters corresponding to the three species.}\label{fig:step0}
+The visible structure in the projection are the three clusters corresponding to the three species.
+Produced with the function \code{view\_basis()}.}\label{fig:step0}
 \end{figure}
@@
 The manip var axis, in red, has length 1 touching the sphere, extends the projection to a third dimension.
-The shadow of this axis (blue) is its contribution in the original 2D projection.}\label{fig:step2}
+The shadow of this axis (blue) is its contribution in the original 2D projection.
+Illustrated with the \code{view\_manip\_space()} function.}\label{fig:step2}
 \end{figure}
@@
 combination of variables, for one frame of the animation, repeat this
-process over a number of values of \(\phi\) to produce an animation.
-
-\textbf{For each value of } \(\phi_i\) \textbf{:}
-
-\begin{align*}
-  \textbf{RM}_{[p,~3,~i]} &= \textbf{M}_{[p,~3]} ~*~ \textbf{R}_{[3,~3,~i]} 
-\end{align*}
+process over a sequence of values for \(\phi\) to produce an animation.
 
@@
 the manip space. To use these for sensitivity analysis, the radial
-rotation is built into an animation where the manip variable is rotated
-fully into the projection, completely out, and then back to the start.
-This involves changing \(\phi\), the angle relative to its initial
-value.
+rotation is built into an animation where the manip var is rotated fully
+into the projection, completely out, and then back to the start. This
+involves changing \(\phi\), the angle relative to its initial value.
 
 When \(\phi\) is 0, the manip var is completely in the projection, and
-when \(\phi\) is \(\pi\), \(\textbf{X}_k\) is not contributing at all.
+when \(\phi\) is \(\pi/2\), \(\textbf{X}_k\) is not contributing at all.
 
@@
   angle value to control this is consistent with the tour algorithm, as
-  implemented in the \CRANpkg{tourr}).
+  implemented in the \pkg{tourr}).
 \item
@@
 
-find the transformation \(\phi_i\) - \(\phi_1\) useful to think about
-\(\phi\) relative to the basis plane. Additionally, the value of
-\(\phi\) may be out of phase by a factor of pi. If the manip variable
-doesn't move as expected these are the first places to check.
+The value of \(\phi\) is relative to its initial value. The
+transformation \(\phi_i\) - \(\phi_1\) is useful to think about \(\phi\)
+relative to the basis plane. Additionally, the value of \(\phi\) may be
+out of phase by a factor of pi. If the manip variable doesn't move as
+expected these are the first places to check.
 
@@
 This shows that aede2 is important for distinguishing this species from the ramaining data.
-An animated version can be viewed at  https://nspyrison.netlify.com/thesis/flea\_manualtour\_mvar5/.}\label{fig:step4}
+Animated versions made with the \code{play\_manual\_tour()} function and can be viewed at https://nspyrison.netlify.com/thesis/flea\_manualtour\_mvar5/.}\label{fig:step4}
 \end{figure}

### 2019-05-13 paper/spinifex_paper.Rmd
functions in figure captions,

@@
 The contribution of the variables aede2 and tars1 approximately contrasts the other variables.
-The visible structure in the projection are the three clusters corresponding to the three species."}
+The visible structure in the projection are the three clusters corresponding to the three species.
+Produced with the function \\code{view\\_basis()}."}
 ## Flea holes tour
@@
 The manip var axis, in red, has length 1 touching the sphere, extends the projection to a third dimension.
-The shadow of this axis (blue) is its contribution in the original 2D projection."}
+The shadow of this axis (blue) is its contribution in the original 2D projection.
+Illustrated with the \\code{view\\_manip\\_space()} function."}
 view_manip_space2(basis = f_bas, manip_var = f_mvar, labels = colnames(f_dat)) +
@@
 Continue with the rotation of the manipulation space, post-multiply the space by the above rotation matrix producing **RM**, the rotated manip space.
-This is basis values, the linear combination of variables, for one frame of the animation, repeat this process over a number of values of $\phi$ to produce an animation.
+This is basis values, the linear combination of variables, for one frame of the animation, repeat this process over a sequence of values for $\phi$ to produce an animation.
 
-**For each value of ** $\phi_i$ **:**
-<!-- don't mix rmd and latex... -->
+<!-- **For each value of ** $\phi_i$ **:** -->
 
-\begin{align*}
-  \textbf{RM}_{[p,~3,~i]} &= \textbf{M}_{[p,~3]} ~*~ \textbf{R}_{[3,~3,~i]} 
-\end{align*}
+<!-- \begin{align*} -->
+<!--   \textbf{RM}_{[p,~3,~i]} &= \textbf{M}_{[p,~3]} ~*~ \textbf{R}_{[3,~3,~i]}  -->
+<!-- \end{align*} -->
 
@@
 The steps outlined above can be used to create any arbitrary rotation in the manip space.
-To use these for sensitivity analysis, the radial rotation is built into an animation where the manip variable is rotated fully into the projection, completely out, and then back to the start.
+To use these for sensitivity analysis, the radial rotation is built into an animation where the manip var is rotated fully into the projection, completely out, and then back to the start.
 This involves changing $\phi$, the angle relative to its initial value.
 
-When $\phi$ is 0, the manip var is completely in the projection, and when $\phi$ is $\pi$, $\textbf{X}_k$ is not contributing at all.
+When $\phi$ is 0, the manip var is completely in the projection, and when $\phi$ is $\pi/2$, $\textbf{X}_k$ is not contributing at all.
 
@@
 Set an angle value ($\Delta_\phi$) that sets the step size to make in rotating the manip var into and out of the projection.
-(Note: Using angle value to control this is consistent with the tour algorithm, as implemented in the \CRANpkg{tourr}).
+(Note: Using angle value to control this is consistent with the tour algorithm, as implemented in the \pkg{tourr}).
 3. Step towards $0$, where the manip var has a completely in the projection.
@@
 
-find the transformation $\phi_i$ - $\phi_1$ useful to think about $\phi$ relative to the basis plane.
+The value of $\phi$ is relative to its initial value.
+The transformation $\phi_i$ - $\phi_1$ is useful to think about $\phi$ relative to the basis plane.
 Additionally, the value of $\phi$ may be out of phase by a factor of pi.
@@
 This shows that aede2 is important for distinguishing this species from the ramaining data.
-An animated version can be viewed at  https://nspyrison.netlify.com/thesis/flea\\_manualtour\\_mvar5/."}
+Animated versions made with the \\code{play\\_manual\\_tour()} function and can be viewed at https://nspyrison.netlify.com/thesis/flea\\_manualtour\\_mvar5/."}
 #play_manual_tour(f_dat, f_bas, f_mvar, col = f_cat, angle = f_angle)

### 2019-05-13 paper/spinifex-paper.tex

@@
 \item
-  \(|\textbf{B}_k|\), the norm of the manip var, \(k\) 
+  \(|\textbf{B}_k|\), the norm of the manip var, \(k\), the length of
+  this basis vector between 0 and 1 inclusive.
 \item
   \(\textbf{e}\), a 1D coordinate basis vector of length \(p\) with the
-  \(k\)-th element set to one.
+  \(k\)-th element set to one and orthonormalized on the end of
+  \(\textbf{B}\).
 \item
@@
   \end{bmatrix} \\
-  \end{align*} Apply Rodrigues' rotation
-\href{mailto:formula\%7B@rodrigues_lois_1840}{\nolinkurl{formula\{@rodrigues\_lois\_1840}}\}:
-\begin{align*}
+  \end{align*} Apply Rodrigues' rotation formula (Rodrigues
+\protect\hyperlink{ref-rodrigues_lois_1840}{1840}): \begin{align*}
     \textbf{R}_{[3,~3]} 
@@
 Defining the axis of rotation for \(\phi\) and the application of
-Rogrigues' rotation formula results in the rotation matrix, the rotation
+Rodrigues' rotation formula results in the rotation matrix, the rotation
 group SO(3). Continue with the rotation of the manipulation space,
@@
 involves changing \(\phi\), the angle relative to its initial value.
-
-When \(\phi\) is 0, the manip var is completely in the projection, and
-when \(\phi\) is \(\pi/2\), \(\textbf{X}_k\) is not contributing at all.
+Apply the transformation \(\phi_i \leftarrow \phi_i-\phi_1\). When
+\(\phi\) is 0, the manip var is completely in the projection, and when
+\(\phi\) is \(\pi/2\), \(\textbf{B}_k\) is not contributing at all.
 
@@
 
-In each of steps 3-5, a small step may be made to bring it exactly to
-the end position.
-
-\hypertarget{checks-that-need-to-be-made}{%
-\subsubsection{Checks that need to be
-made}\label{checks-that-need-to-be-made}}
-
-The value of \(\phi\) is relative to its initial value. The
-transformation \(\phi_i\) - \(\phi_1\) is useful to think about \(\phi\)
-relative to the basis plane. Additionally, the value of \(\phi\) may be
-out of phase by a factor of pi. If the manip variable doesn't move as
-expected these are the first places to check.
+In each of the steps 3-5, a small step may be made to bring it exactly
+to the end position.
 
@@
 showing the reference axes on the top half with the corresponding
-projected data points below. Changes in the basis coefficients of the
-manip var correspond to the distance separating the purple cluster and
-the remaining sample, aede2 is crucial in distinguishing the variation
-of species. Tours are typically viewed as an animation. The animation of
-this tour can be viewed online at
+projected data points below. Changes in \(\textbf{B}_k\) correspond to
+the distance separating the purple cluster and the remaining sample.
+This shows aede2 is crucial in distinguishing the variation of species.
+Tours are typically viewed as an animation. The animation of this tour
+can be viewed online at
 \url{https://nspyrison.netlify.com/thesis/flea_manualtour_mvar5/}. The
@@
 \end{Schunk}
+
+\hypertarget{checks-that-need-to-be-made}{%
+\subsubsection{Checks that need to be
+made}\label{checks-that-need-to-be-made}}
+
+The value of \(\phi\) is relative to its initial value. The
+transformation \(\phi_i\) - \(\phi_1\) is useful to think about \(\phi\)
+relative to the basis plane. Additionally, the value of \(\phi\) may be
+out of phase by a factor of \(\pi\). If the manip variable doesn't move
+as expected these are the first places to check.
 
@@
 
-# Also see the vignette: 
+# Also see the vignette for further code usage: 
 vignette("spinifex_vignette")
@@
 
+\leavevmode\hypertarget{ref-rodrigues_lois_1840}{}%
+Rodrigues, Olinde. 1840. \emph{Des Lois Géométriques Qui Régissent Les
+Déplacements d'un Système Solide Dans L'espace: Et de La Variation Des
+Cordonnées Provenant de Ces Déplacements Considérés Indépendamment Des
+Causes Qui Peuvent Les Produire}.
+
 \leavevmode\hypertarget{ref-sievert_plotly_2018}{}%

### 2019-05-13 paper/spinifex_paper.Rmd

@@
 - $k$, is the index of the variable to manipulate, called the "manip var".
-- $|\textbf{B}_k|$, the norm of the manip var, $k$
-<!-- - $\textbf{B}_k$, the manipulation variable (manip var), the variable whoes contribution will be controled. -->
-- $\textbf{e}$, a 1D coordinate basis vector of length $p$ with the $k$-th element set to one. 
+- $|\textbf{B}_k|$, the norm of the manip var, $k$, the length of this basis vector between 0 and 1 inclusive.
+- $\textbf{e}$, a 1D coordinate basis vector of length $p$ with the $k$-th element set to one and orthonormalized on the end of $\textbf{B}$. 
 - $\theta$, the angle of in-projection-plane rotation, for example, on the reference axes; $c_\theta, s_\theta$ are the cosine and sine.
@@
   \end{align*}
-Apply Rodrigues' rotation formula{@rodrigues_lois_1840}:
+Apply Rodrigues' rotation formula [@rodrigues_lois_1840]:
   \begin{align*}
@@
 <!-- rotate manip space -->
-Defining the axis of rotation for $\phi$ and the application of Rogrigues' rotation formula results in the rotation matrix, the rotation group SO(3).
+Defining the axis of rotation for $\phi$ and the application of Rodrigues' rotation formula results in the rotation matrix, the rotation group SO(3).
 Continue with the rotation of the manipulation space, post-multiply the space by the above rotation matrix producing **RM**, the rotated manip space.
@@
 This involves changing $\phi$, the angle relative to its initial value.
-
-When $\phi$ is 0, the manip var is completely in the projection, and when $\phi$ is $\pi/2$, $\textbf{X}_k$ is not contributing at all.
+Apply the transformation $\phi_i \leftarrow \phi_i-\phi_1$.
+When $\phi$ is 0, the manip var is completely in the projection, and when $\phi$ is $\pi/2$, $\textbf{B}_k$ is not contributing at all.
 
@@
 
-In each of steps 3-5, a small step may be made to bring it exactly to the end position.
+In each of the steps 3-5, a small step may be made to bring it exactly to the end position.
 
-### Checks that need to be made
-
-The value of $\phi$ is relative to its initial value.
-The transformation $\phi_i$ - $\phi_1$ is useful to think about $\phi$ relative to the basis plane.
-Additionally, the value of $\phi$ may be out of phase by a factor of pi.
-If the manip variable doesn't move as expected these are the first places to check.
-
-<!-- description of slides, on the technical side we haven't gone back to data space yet. -->
+<!-- Animation on flea data -->
 Figure \ref{fig:step4} illustrates a sequence with 15 projected bases, showing the reference axes on the top half with the corresponding projected data points below.
-Changes in the basis coefficients of the manip var correspond to the distance separating the purple cluster and the remaining sample, aede2 is crucial in distinguishing the variation of species.
+Changes in $\textbf{B}_k$ correspond to the distance separating the purple cluster and the remaining sample.
+This shows aede2 is crucial in distinguishing the variation of species.
 Tours are typically viewed as an animation.
@@
 
-<!-- step3 caption -->
 ```{r step4, echo=F, warning=F, fig.height=7.0, fig.width=5.83, out.height='7in', out.width='5.83in',                                                                fig.cap = "Radial manual tour manipulating aede2 of standardized flea data.
@@
 ```
+
+### Checks that need to be made
+
+The value of $\phi$ is relative to its initial value.
+The transformation $\phi_i$ - $\phi_1$ is useful to think about $\phi$ relative to the basis plane.
+Additionally, the value of $\phi$ may be out of phase by a factor of $\pi$.
+If the manip variable doesn't move as expected these are the first places to check.
 

### 2019-05-15 paper/spinifex-paper.tex
Structure, wording, and notaion

@@
 \item
-  \(\textbf{e}\), a 1D coordinate basis vector of length \(p\) with the
-  \(k\)-th element set to one and orthonormalized on the end of
-  \(\textbf{B}\).
+  \(\textbf{e}\), a 1D coordinate zero basis vector of length \(p\) with
+  the \(k\)-th element set to one.
+\item
+  \(\textbf{M} = orthonormalize(\textbf{B}|\textbf{e})\), the
+  manipulation space, a \(p \times 3\) matrix where out-of-plane
+  rotation takes place.
 \item
@@
 \item
-  \(\textbf{U}\) is the axis of rotation for \(\phi\).
+  \(\textbf{U}\), the axis of rotation for \(\phi\).
+\item
+  \(\textbf{R}\), the rotation matrix, containing the rotaions of
+  \(\theta\) and \(\phi\) within the manipulation space, \(\textbf{M}\).
+\item
+  \(\textbf{RM} = \textbf{M} * \textbf{R}\), the result of the
+  manipulation space after the application of the rotation matrix.
+\item
+  \(\textbf{P} = \textbf{X} * \textbf{RM}\), the data projection, the
+  data as projected through the rotated manipulation space.
 \end{itemize}
@@
 The holes index is maximized when the projected observations are
-furthest from the center. Figure \ref{fig:step0} shows a locally
-optimized projection of the data. The left panel displays the reference
-axes of the projection basis, a visual indication of the magnitude and
-direction each variable contributes to the projection. The right panel
-shows the data as projected through the basis set described by the
-reference axes (left). Data points are colored and given shape according
-to the species.
+furthest from the center.
 
-Call \code{view\_basis()} on a basis to produce a \CRANpkg{ggplot2}
-(Wickham \protect\hyperlink{ref-wickham_ggplot2:_2016}{2016}) graphic
-similar to Figure \ref{fig:step0}. The projected data is processed for
-display using matrix multiplication
-\(\textbf{X}_{[n,~p]} ~*~ \textbf{B}_{[p,~d]} ~=~ \textbf{P}_{[n,~d]}\).
+Figure \ref{fig:step0} shows a locally optimized projection of the data.
+The left panel displays the reference axes of the projection basis, a
+visual indication of the magnitude and direction each variable
+contributes to the projection. The right panel shows the data as
+projected through the basis set described by the reference axes (left).
+Data points are colored and given shape according to the species. Call
+\code{view\_basis()} on a basis to produce a \CRANpkg{ggplot2} (Wickham
+\protect\hyperlink{ref-wickham_ggplot2:_2016}{2016}) object similar to
+Figure \ref{fig:step0}. The projected data is processed for display
+using matrix multiplication,
+\(\textbf{P}_{[n,~d]} ~=~ \textbf{X}_{[n,~p]} ~*~ \textbf{B}_{[p,~d]}\).
 
@@
 of length \(p\), and set the \(k\)-th element to 1. In the example data,
-aede2 is the fifth variable in the data, so \(k=5\), set \(e_5=1\) . Use
+aede2 is the fifth variable in the data, so \(k=5\), set \(e_5=1\). Use
 a Gram-Schmidt process to orthonormalize the coordinate basis vector on
@@
 where the coefficient of the manip var can range completely between
-{[}-1, 1{]}. 3D rotation can be used to rotate the mani variable
+{[}-1, 1{]}. 3D rotation can be used to rotate the manip variable
 completely into or completely out of a 2D projection. Figure
@@
 \code{view\_manip\_space()} function. Functionally, the manipulation
-space is more of a mathematical construct that needn't be visualized but
-this diagram provides intuition for the technique.
+space adds another layer of complexity that needn't be displayed in the
+final animation, but this diagram aids in comprehension of the
+technique.
 
@@
 plane, \(\theta\), and define a sequence for the (vertical) angle coming
-out of the projection plane, \(\phi\), bringing the initial \(XY\)
-contributions of the manip var to a maximum and then to zero before
-returning to the initial position. Alternatively, for an oblique manual
-tour, capture the user manipulation directly on the \(XY\) of the
-projection plane.
+out of the projection plane, \(\phi\), bringing the initial value of
+\(|\textbf{B}_k|\) to a maximum and then to zero before returning to the
+initial position. Alternatively, for an \emph{oblique} manual tour,
+capture the user manipulation directly on the \(XY\) of the projection
+plane.
 
@@
 rotation is built into an animation where the manip var is rotated fully
-into the projection, completely out, and then back to the start. This
-involves changing \(\phi\), the angle relative to its initial value.
-Apply the transformation \(\phi_i \leftarrow \phi_i-\phi_1\). When
-\(\phi\) is 0, the manip var is completely in the projection, and when
-\(\phi\) is \(\pi/2\), \(\textbf{B}_k\) is not contributing at all.
+into the projection, completely out, and then back to the initial value.
+This involves changing \(\phi\), the angle relative to its initial
+value. Apply the transformation \(\phi_i \leftarrow \phi_i-\phi_1\) such
+that \(\phi\) is relative to the projection plane. When \(\phi\) is 0,
+\(|\textbf{B}_k| = 1\), and when \(\phi\) is \(\pi/2\), the manip var
+has no contribution to the projection.
 
@@
 
-In each of the steps 3-5, a small step may be made to bring it exactly
-to the end position.
+In each of the steps 3-5, a small step may be appended to bring \(\phi\)
+exactly to the desired value.
 
@@
 \caption[Radial manual tour manipulating aede2 of standardized flea data]{Radial manual tour manipulating aede2 of standardized flea data.
-The contributions increase from its initial contribution to a full contribution to the projection before decreasing to zero and then returning to its initial value.
+The contributions increase from its initial vlaue to a full contribution to the projection before decreasing to zero and then returning to the initial value.
 Manipulating the coefficeients of aede2 changes the distance separating the purple cluster.
@@
 relative to the basis plane. Additionally, the value of \(\phi\) may be
-out of phase by a factor of \(\pi\). If the manip variable doesn't move
-as expected these are the first places to check.
+out of phase by a factor of \(\pi\). If the rotation does not behave as
+expected these are the first places to check.
 
 \hypertarget{sec:display}{%
-\subsection{Data in projection-space}\label{sec:display}}
+\subsection{Projecting the data}\label{sec:display}}
 
@@
 tours are broken into the following groupings: tour preparation, tour
-animation, and utility.
+animation, and utility. In the following section describes and
+demonstrates the usage of the above functions following the order in the
+algorithm. The utility functions are called by the tour animation
+functions, passed to the argument \code{render\_type}. Usage is
+demonstrated below and in the example documentation.
 
@@
 \end{Schunk}
-
-In the following section describes and demonstrates the usage of the
-above functions following the order in the algorithm. The utility
-functions are called by the tour animation functions, passed to the
-argument \code{render\_type}. Usage is demonstrated below and in the
-example documentation.
 
@@
 correct manip var, as demonstrated by comparing tours of varying amounts
-of structural information. Manual tours, by giving the full contribution
-of the manipulation space to the manip var, enable analysts to explore
-the sensitivity of the structure with the selected variable. This
+of structural information. Manual tours, by controlling the contribution
+of the manip var on the projection, enable analysts to explore the
+sensitivity that variable has on the structure of the projection. This
 information can be used by domain experts to identify which variables
-are probing which parameter spaces and how sensitive those variables are
-to structural features. This insight can be indispensable for variable
-inclusion/exclusions in a model, or in high-level decisions, such as
-budgeting time and funding.
+are probing which parameter spaces and how sensitive structural features
+are to different variables. This insight can be indispensable for
+variable inclusion/exclusion, and in higher-level decisions, such as
+meta-analysis, to suggest directions of future research.
 

### 2019-05-15 paper/spinifex_paper.Rmd
Structure, wording, and notaion

@@
 - $|\textbf{B}_k|$, the norm of the manip var, $k$, the length of this basis vector between 0 and 1 inclusive.
-- $\textbf{e}$, a 1D coordinate basis vector of length $p$ with the $k$-th element set to one and orthonormalized on the end of $\textbf{B}$. 
+- $\textbf{e}$, a 1D coordinate zero basis vector of length $p$ with the $k$-th element set to one.
+- $\textbf{M} = orthonormalize(\textbf{B}|\textbf{e})$, the manipulation space, a $p \times 3$ matrix where out-of-plane rotation takes place. 
 - $\theta$, the angle of in-projection-plane rotation, for example, on the reference axes; $c_\theta, s_\theta$ are the cosine and sine.
 - $\phi$, the angle of out-of-projection-plane rotation, into the manipulation space; $c_\phi, s_\phi$ are the cosine and sine.
-- $\textbf{U}$ is the axis of rotation for $\phi$.
+- $\textbf{U}$, the axis of rotation for $\phi$.
+- $\textbf{R}$, the rotation matrix, containing the rotaions of $\theta$ and $\phi$ within the manipulation space, $\textbf{M}$.
+- $\textbf{RM} = \textbf{M} * \textbf{R}$, the result of the manipulation space after the application of the rotation matrix.
+- $\textbf{P} = \textbf{X} * \textbf{RM}$, the data projection, the data as projected through the rotated manipulation space.
 
@@
 The algorithm operates entirely on projection bases and utilizes the data only later when making a display.
-<!--The projection space can be viewed at any point in the process by pre-multiplying the data and plotting the first two variables.-->
 
@@
 The holes index is maximized when the projected observations are furthest from the center.
+
 Figure \ref{fig:step0} shows a locally optimized projection of the data.
@@
 Data points are colored and given shape according to the species.
-
-Call \code{view\_basis()} on a basis to produce a \CRANpkg{ggplot2} [@wickham_ggplot2:_2016] graphic similar to Figure \ref{fig:step0}.
-The projected data is processed for display using matrix multiplication $\textbf{X}_{[n,~p]} ~*~ \textbf{B}_{[p,~d]} ~=~ \textbf{P}_{[n,~d]}$.
+Call \code{view\_basis()} on a basis to produce a \CRANpkg{ggplot2} [@wickham_ggplot2:_2016] object similar to Figure \ref{fig:step0}.
+The projected data is processed for display using matrix multiplication, $\textbf{P}_{[n,~d]} ~=~ \textbf{X}_{[n,~p]} ~*~ \textbf{B}_{[p,~d]}$.
 
@@
 Initialize the coordinate basis vector as a zero vector $\textbf{e}$ of length $p$, and set the $k$-th element to 1.
-In the example data, aede2 is the fifth variable in the data, so $k=5$, set $e_5=1$ .
+In the example data, aede2 is the fifth variable in the data, so $k=5$, set $e_5=1$.
 Use a Gram-Schmidt process to orthonormalize the coordinate basis vector on the original 2D projection to describe a 3D manipulation space, $\textbf{M}$.
@@
 
-<!-- What the manip space facilitates -->
+<!-- What the manip space provides -->
 The manip space provides a 3D projection from $p$-dimensional space, where the coefficient of the manip var can range completely between [-1, 1].
-3D rotation can be used to rotate the mani variable completely into or completely out of a 2D projection.
+3D rotation can be used to rotate the manip variable completely into or completely out of a 2D projection.
 Figure \ref{fig:step2} illustrates this 3D manipulation space with the manip var highlighted.
 This representation is produced by calling the \code{view\_manip\_space()} function.
-Functionally, the manipulation space is more of a mathematical construct that needn't be visualized but this diagram provides intuition for the technique.
+Functionally, the manipulation space adds another layer of complexity that needn't be displayed in the final animation, but this diagram aids in comprehension of the technique.
 
@@
 <!-- describe manip var path -->
-For a radial tour fix the (horizontal) angle within the projection plane, $\theta$, and define a sequence for the (vertical) angle coming out of the projection plane, $\phi$, bringing the initial $XY$ contributions of the manip var to a maximum and then to zero before returning to the initial position.
-Alternatively, for an oblique manual tour, capture the user manipulation directly on the $XY$ of the projection plane.
+For a radial tour fix the (horizontal) angle within the projection plane, $\theta$, and define a sequence for the (vertical) angle coming out of the projection plane, $\phi$, bringing the initial value of $|\textbf{B}_k|$ to a maximum and then to zero before returning to the initial position.
+Alternatively, for an *oblique* manual tour, capture the user manipulation directly on the $XY$ of the projection plane.
 
@@
 The steps outlined above can be used to create any arbitrary rotation in the manip space.
-To use these for sensitivity analysis, the radial rotation is built into an animation where the manip var is rotated fully into the projection, completely out, and then back to the start.
+To use these for sensitivity analysis, the radial rotation is built into an animation where the manip var is rotated fully into the projection, completely out, and then back to the initial value.
 This involves changing $\phi$, the angle relative to its initial value.
-Apply the transformation $\phi_i \leftarrow \phi_i-\phi_1$.
-When $\phi$ is 0, the manip var is completely in the projection, and when $\phi$ is $\pi/2$, $\textbf{B}_k$ is not contributing at all.
+Apply the transformation $\phi_i \leftarrow \phi_i-\phi_1$ such that $\phi$ is relative to the projection plane.
+When $\phi$ is 0, $|\textbf{B}_k| = 1$, and when $\phi$ is $\pi/2$, the manip var has no contribution to the projection.
 
@@
 
-In each of the steps 3-5, a small step may be made to bring it exactly to the end position.
+In each of the steps 3-5, a small step may be appended to bring $\phi$ exactly to the desired value.
 
@@
 ```{r step4, echo=F, warning=F, fig.height=7.0, fig.width=5.83, out.height='7in', out.width='5.83in',                                                                fig.cap = "Radial manual tour manipulating aede2 of standardized flea data.
-The contributions increase from its initial contribution to a full contribution to the projection before decreasing to zero and then returning to its initial value.
+The contributions increase from its initial vlaue to a full contribution to the projection before decreasing to zero and then returning to the initial value.
 Manipulating the coefficeients of aede2 changes the distance separating the purple cluster.
@@
 Additionally, the value of $\phi$ may be out of phase by a factor of $\pi$.
-If the manip variable doesn't move as expected these are the first places to check.
+If the rotation does not behave as expected these are the first places to check.
 
-## Data in projection-space {#sec:display}
+## Projecting the data {#sec:display}
 
@@
 
-<!-- step 0 disp code. -->
 ```{r installCODEEXAMPLE, eval=F, echo=T}
@@
 The primary functions that will aid users in performing their own manual tours are broken into the following groupings: tour preparation, tour animation, and utility.
+In the following section describes and demonstrates the usage of the above functions following the order in the algorithm.
+The utility functions are called by the tour animation functions, passed to the argument \code{render\_type}.
+Usage is demonstrated below and in the example documentation.
 
@@
 ```
-
-In the following section describes and demonstrates the usage of the above functions following the order in the algorithm.
-The utility functions are called by the tour animation functions, passed to the argument \code{render\_type}.
-Usage is demonstrated below and in the example documentation.
 
@@
 The importance of selecting the correct manip var, as demonstrated by comparing tours of varying amounts of structural information.
-Manual tours, by giving the full contribution of the manipulation space to the manip var, enable analysts to explore the sensitivity of the structure with the selected variable.
-This information can be used by domain experts to identify which variables are probing which parameter spaces and how sensitive those variables are to structural features.
-This insight can be indispensable for variable inclusion/exclusions in a model, or in high-level decisions, such as budgeting time and funding.
+Manual tours, by controlling the contribution of the manip var on the projection, enable analysts to explore the sensitivity that variable has on the structure of the projection.
+This information can be used by domain experts to identify which variables are probing which parameter spaces and how sensitive structural features are to different variables.
+This insight can be indispensable for variable inclusion/exclusion, and in higher-level decisions, such as meta-analysis, to suggest directions of future research.
 

### 2019-05-22 paper/spinifex-paper.tex
checking that the few changes made are uploaded

@@
 
-Imagine holding the manip var, the red axis in Figure \ref{fig:step2},
-one end fixed to the origin and moving it like a lever. This is
-effectively the process of the radial manual to rotate the manip
-variable into and out of the 2D projection (Figure \ref{fig:step3}). As
-the variable is controlled, the manipulation space rotates about the
-origin, the projection onto the horizontal projection plane
-correspondingly moves. This is what happens in a manual tour. Generating
-a sequence of values for the horizontal and vertical, angles produces a
-path for the rotation of the manipulation space. This defines the
-(orthonormally-constrained) rotation on the basis coefficients.
+Imagine moving the axis of manip var, the red line in Figure
+\ref{fig:step2}, like a lever. This is effectively the process of the
+radial manual control, that rotates the manip variable into and out of
+the 2D projection (Figure \ref{fig:step3}). As the variable is
+controlled, the manipulation space rotates, and the projection onto the
+horizontal projection plane correspondingly changes. This is a manual
+tour. Generating a sequence of values for the horizontal and vertical
+angles produces a path for the rotation of the manipulation space. This
+defines the (orthonormally-constrained) rotation on the basis
+coefficients.
 
-For a radial tour fix the (horizontal) angle within the projection
-plane, \(\theta\), and define a sequence for the (vertical) angle coming
-out of the projection plane, \(\phi\), bringing the initial value of
-\(|\textbf{B}_k|\) to a maximum and then to zero before returning to the
-initial position. Alternatively, for an \emph{oblique} manual tour,
-capture the user manipulation directly on the \(XY\) of the projection
-plane.
+For a radial tour, fix the (horizontal) angle within the projection
+plane, \(\theta\), and define a sequence for the (vertical) angle,
+\(\phi\), defining movement out of the plane, that changes the initial
+value of \(|\textbf{B}_k|\) to a maximum of 1, and then to 0, before
+returning to the initial value.
 

### 2019-05-22 paper/spinifex_paper.Rmd
checking that the few changes made are uploaded

@@
 <!-- illustration of axis manip -->
-Imagine holding the manip var, the red axis in Figure \ref{fig:step2}, one end fixed to the origin and moving it like a lever.
-This is effectively the process of the radial manual to rotate the manip variable into and out of the 2D projection (Figure \ref{fig:step3}).
-As the variable is controlled, the manipulation space rotates about the origin, the projection onto the horizontal projection plane correspondingly moves.
-This is what happens in a manual tour.
-Generating a sequence of values for the horizontal and vertical, angles produces a path for the rotation of the manipulation space.
+Imagine moving the axis of manip var, the red line in Figure \ref{fig:step2}, like a lever.
+This is effectively the process of the radial manual control, that rotates the manip variable into and out of the 2D projection (Figure \ref{fig:step3}).
+As the variable is controlled, the manipulation space rotates, and the projection onto the horizontal projection plane correspondingly changes.
+This is a manual tour.
+Generating a sequence of values for the horizontal and vertical angles produces a path for the rotation of the manipulation space.
 This defines the (orthonormally-constrained) rotation on the basis coefficients.
@@
 <!-- describe manip var path -->
-For a radial tour fix the (horizontal) angle within the projection plane, $\theta$, and define a sequence for the (vertical) angle coming out of the projection plane, $\phi$, bringing the initial value of $|\textbf{B}_k|$ to a maximum and then to zero before returning to the initial position.
-Alternatively, for an *oblique* manual tour, capture the user manipulation directly on the $XY$ of the projection plane.
+For a radial tour, fix the (horizontal) angle within the projection plane, $\theta$, and define a sequence for the (vertical) angle, $\phi$, defining movement out of the plane, that changes the initial value of $|\textbf{B}_k|$ to a maximum of 1, and then to 0, before returning to the initial value.
 

### 2019-05-29 paper/spinifex-paper.tex
di tweaks on plane

@@
   \(\phi\), the angle of out-of-projection-plane rotation, into the
-  manipulation space; \(c_\phi, s_\phi\) are the cosine and sine.
+  manipulation space; \(c_\phi, s_\phi\) are the cosine and sine. The
+  initial value for animation purposes is \(\phi_1\).
 \item
@@
 \end{itemize}
+
+XXX Clean up the notation, not all of this is needed or useful. The
+shadow of \(e\) in the projection may need some notation. The projected
+data should be \(Y\) not \(P\).
 
@@
 
-\hypertarget{step-3-generate-rotation}{%
-\subsubsection{Step 3) Generate
-rotation}\label{step-3-generate-rotation}}
+\hypertarget{step-3-defining-a-3d-rotation}{%
+\subsubsection{Step 3) Defining a 3D
+rotation}\label{step-3-defining-a-3d-rotation}}
 
-Imagine moving the axis of manip var, the red line in Figure
-\ref{fig:step2}, like a lever. This is effectively the process of the
-radial manual control, that rotates the manip variable into and out of
-the 2D projection (Figure \ref{fig:step3}). As the variable is
-controlled, the manipulation space rotates, and the projection onto the
-horizontal projection plane correspondingly changes. This is a manual
-tour. Generating a sequence of values for the horizontal and vertical
-angles produces a path for the rotation of the manipulation space. This
-defines the (orthonormally-constrained) rotation on the basis
-coefficients.
+The basis vector corresponding to the manip var (red line in Figure
+\ref{fig:step2}), can be operated like a lever. This is the process of
+the manual control, that rotates the manip variable into and out of the
+2D projection (Figure \ref{fig:step3}). As the variable is controlled,
+the manipulation space rotates, and the projection onto the horizontal
+projection plane correspondingly changes. This is a manual tour.
+Generating a sequence of values for the horizontal and vertical angles
+produces a path for the rotation of the manipulation space. This defines
+the (orthonormally-constrained) rotation on the basis coefficients.
 
@@
 \(\phi\), defining movement out of the plane, that changes the initial
-value of \(|\textbf{B}_k|\) to a maximum of 1, and then to 0, before
-returning to the initial value.
+value of \(\phi_1\), the angle between \(\textbf{e}\) and its shadow in
+\(\textbf{B}\), to a maximum of 1, and then to 0, before returning to
+\(\phi_1\).
 
-Rotations are defined by the axes they pivot on, \(\theta\) then is the
-rotation around the \(Z\) axis. While \(\phi\) will be the rotation
-around, \(\textbf{U}\), a new axis on the \(XY\) plane, orthogonal to
-\(\phi\). Given these axes, the SO(3) rotation matrix can be derived by
-the following.
+Rotations in 3D can be defined by the axes they pivot on. \(\theta\)
+describes the rotation around the \(Z\) axis. While \(\phi\) is the
+rotation around \(\textbf{U}\), an axis on the \(XY\) plane, orthogonal
+to \(\textbf{e}\). Given these axes, the SO(3) rotation matrix can be
+derived as follows:
 
@@
 
-\hypertarget{creating-the-animation}{%
-\subsection{Creating the animation}\label{creating-the-animation}}
+\hypertarget{creating-the-animation-of-the-radial-rotation}{%
+\subsection{Creating the animation of the radial
+rotation}\label{creating-the-animation-of-the-radial-rotation}}
 
@@
 into the projection, completely out, and then back to the initial value.
-This involves changing \(\phi\), the angle relative to its initial
-value. Apply the transformation \(\phi_i \leftarrow \phi_i-\phi_1\) such
-that \(\phi\) is relative to the projection plane. When \(\phi\) is 0,
-\(|\textbf{B}_k| = 1\), and when \(\phi\) is \(\pi/2\), the manip var
-has no contribution to the projection.
+This involves allowing \(\phi\), to vary between 0 and 1.
 
@@
 \item
-  Set initial value to be \(\phi_1\).
+  Set initial value of \(\phi_1\) and \(\theta\):
+  \(\phi_1 = \cos^{-1}{\sqrt{B_{k1}^2+B_{k2}^2}}\),
+  \(\theta=\tan^{-1}\frac{B_{k2}}{B_{k1}}\). \(\phi_1\) is the angle
+  between \(\textbf{e}\) and its shadow in \(\textbf{B}\). Check that
+  \(\phi_1\) is between 0 and \(\pi/2\), and if not, transform it into
+  this range, by subtracting or adding \(\pi\) XXX.
 \item
-  Set an angle value (\(\Delta_\phi\)) that sets the step size to make
-  in rotating the manip var into and out of the projection. (Note: Using
-  angle value to control this is consistent with the tour algorithm, as
-  implemented in the \pkg{tourr}).
+  Set an angle increment (\(\Delta_\phi\)) that sets the step size to
+  make, to rotate the manip var into and out of the projection. (Note:
+  Using angle increment, rather than a number of steps, to control the
+  movement, is consistent with the tour algorithm as implemented in the
+  \pkg{tourr}).
 \item
@@
 \end{Schunk}
-
-\hypertarget{checks-that-need-to-be-made}{%
-\subsubsection{Checks that need to be
-made}\label{checks-that-need-to-be-made}}
-
-The value of \(\phi\) is relative to its initial value. The
-transformation \(\phi_i\) - \(\phi_1\) is useful to think about \(\phi\)
-relative to the basis plane. Additionally, the value of \(\phi\) may be
-out of phase by a factor of \(\pi\). If the rotation does not behave as
-expected these are the first places to check.
 

### 2019-05-29 paper/spinifex_paper.Rmd
di tweaks on plane

@@
 - $\theta$, the angle of in-projection-plane rotation, for example, on the reference axes; $c_\theta, s_\theta$ are the cosine and sine.
-- $\phi$, the angle of out-of-projection-plane rotation, into the manipulation space; $c_\phi, s_\phi$ are the cosine and sine.
+- $\phi$, the angle of out-of-projection-plane rotation, into the manipulation space; $c_\phi, s_\phi$ are the cosine and sine. The initial value for animation purposes is $\phi_1$.
 - $\textbf{U}$, the axis of rotation for $\phi$.
@@
 - $\textbf{P} = \textbf{X} * \textbf{RM}$, the data projection, the data as projected through the rotated manipulation space.
+
+XXX Clean up the notation, not all of this is needed or useful. The shadow of $e$ in the projection may need some notation. The projected data should be $Y$ not $P$.
 
@@
 
-### Step 3) Generate rotation
+### Step 3) Defining a 3D rotation
 
 <!-- illustration of axis manip -->
-Imagine moving the axis of manip var, the red line in Figure \ref{fig:step2}, like a lever.
-This is effectively the process of the radial manual control, that rotates the manip variable into and out of the 2D projection (Figure \ref{fig:step3}).
+The basis vector corresponding to the manip var (red line in Figure \ref{fig:step2}), can be operated like a lever.
+This is the process of the manual control, that rotates the manip variable into and out of the 2D projection (Figure \ref{fig:step3}).
 As the variable is controlled, the manipulation space rotates, and the projection onto the horizontal projection plane correspondingly changes.
@@
 <!-- describe manip var path -->
-For a radial tour, fix the (horizontal) angle within the projection plane, $\theta$, and define a sequence for the (vertical) angle, $\phi$, defining movement out of the plane, that changes the initial value of $|\textbf{B}_k|$ to a maximum of 1, and then to 0, before returning to the initial value.
+For a radial tour, fix the (horizontal) angle within the projection plane, $\theta$, and define a sequence for the (vertical) angle, $\phi$, defining movement out of the plane, that changes the initial value of $\phi_1$, the angle between $\textbf{e}$ and its shadow in $\textbf{B}$, to a maximum of 1, and then to 0, before returning to $\phi_1$.
 
 <!-- define the rotation matrix -->
-Rotations are defined by the axes they pivot on, $\theta$ then is the rotation around the $Z$ axis.
-While $\phi$ will be the rotation around, $\textbf{U}$, a new axis on the $XY$ plane, orthogonal to $\phi$.
-Given these axes, the SO(3) rotation matrix can be derived by the following.
+Rotations in 3D can be defined by the axes they pivot on.
+$\theta$ describes the rotation around the $Z$ axis.
+While $\phi$ is the rotation around $\textbf{U}$, an axis on the $XY$ plane, orthogonal to $\textbf{e}$.
+Given these axes, the SO(3) rotation matrix can be derived as follows:
 
@@
 
-## Creating the animation
+## Creating the animation of the radial rotation
 
@@
 To use these for sensitivity analysis, the radial rotation is built into an animation where the manip var is rotated fully into the projection, completely out, and then back to the initial value.
-This involves changing $\phi$, the angle relative to its initial value.
-Apply the transformation $\phi_i \leftarrow \phi_i-\phi_1$ such that $\phi$ is relative to the projection plane.
-When $\phi$ is 0, $|\textbf{B}_k| = 1$, and when $\phi$ is $\pi/2$, the manip var has no contribution to the projection.
+This involves allowing $\phi$, to vary between 0 and 1.
+
+<!--Apply the transformation $\phi_i \leftarrow \phi_i-\phi_1$ such that $\phi$ is relative to the projection plane.
+When $\phi$ is 0, $|\textbf{B}_k| = 1$, and when $\phi$ is $\pi/2$, the manip var has no contribution to the projection.-->
 
@@
 
-1. Set initial value to be $\phi_1$.
+1.
+Set initial value of $\phi_1$ and $\theta$: $\phi_1 = \cos^{-1}{\sqrt{B_{k1}^2+B_{k2}^2}}$, $\theta=\tan^{-1}\frac{B_{k2}}{B_{k1}}$.
+$\phi_1$ is the angle between $\textbf{e}$ and its shadow in $\textbf{B}$.
+Check that  $\phi_1$ is between 0 and $\pi/2$, and if not,  transform it into this range, by subtracting or adding $\pi$ XXX.
 2.
-Set an angle value ($\Delta_\phi$) that sets the step size to make in rotating the manip var into and out of the projection.
-(Note: Using angle value to control this is consistent with the tour algorithm, as implemented in the \pkg{tourr}).
+Set an angle increment ($\Delta_\phi$) that sets the step size to make, to rotate the manip var into and out of the projection.
+(Note: Using angle increment, rather than a number of steps, to control the movement, is consistent with the tour algorithm as implemented in the \pkg{tourr}).
 3. Step towards $0$, where the manip var has a completely in the projection.
@@
 
+<!--
 ### Checks that need to be made
 
-The value of $\phi$ is relative to its initial value.
+The initial value of $\phi$ is determined by the angle between $\textbf{e}$ and its shadow in $\textbf{B}$.
+
+relative to its initial value.
 The transformation $\phi_i$ - $\phi_1$ is useful to think about $\phi$ relative to the basis plane.
@@
 If the rotation does not behave as expected these are the first places to check.
+-->
 

### 2019-05-29 paper/spinifex-paper.tex
step 0 and notation

@@
 projection, completely removes it, and then returns to the original
-position. This is an overview of the steps:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Provided with a 2D projection, choose a variable to manipulate. This
-  is called the ``manip var''.
-\item
-  Create a 3D manipulation space, where the manip variable has the full
-  contribution.
-\item
-  Generate a rotation sequence which increases the norm of the
-  coefficient to 1 and zeros it.
-\end{enumerate}
-
-More detail is provided in the following sections. Function names are
-provided, but code structure and usage is discussed in section
-\ref{sec:usage}.
+position.
 
@@
   \(\textbf{B} = (B_1,~ B_2)\), any 2D orthonormal projection basis,
-  \(p \times 2\) matrix, describing the projection from \(p\) to two
-  dimensions. We will call this the ``original projection'' because it
-  is the starting point for the manual tour.
+  \(p \times 2\) matrix, describing the projection from
+  \(\mathbb{R}^p \Rightarrow \mathbb{R}^2\). This is called this the
+  ``original projection'' because it is the starting point for the
+  manual tour.
 \item
@@
 \item
-  \(|\textbf{B}_k|\), the norm of the manip var, \(k\), the length of
-  this basis vector between 0 and 1 inclusive.
+  \(\textbf{e}\), a 1D basis vector of length \(p\), with 1 in the
+  \(k\)-th position and 0 elsewhere.
 \item
-  \(\textbf{e}\), a 1D coordinate zero basis vector of length \(p\) with
-  the \(k\)-th element set to one.
-\item
-  \(\textbf{M} = orthonormalize(\textbf{B}|\textbf{e})\), the
-  manipulation space, a \(p \times 3\) matrix where out-of-plane
-  rotation takes place.
+  \(\textbf{M}\) is a \(p \times 3\) matrix, defining the 3D subspace
+  where data rotation occurs, and is called the manip(ulation) space.
 \item
@@
   \(\phi\), the angle of out-of-projection-plane rotation, into the
-  manipulation space; \(c_\phi, s_\phi\) are the cosine and sine. The
-  initial value for animation purposes is \(\phi_1\).
+  manip space; \(c_\phi, s_\phi\) are the cosine and sine. The initial
+  value for animation purposes is \(\phi_1\).
 \item
-  \(\textbf{U}\), the axis of rotation for \(\phi\).
+  \(\textbf{U}\), the axis of rotation for out-of-projection rotation.
 \item
-  \(\textbf{R}\), the rotation matrix, containing the rotaions of
-  \(\theta\) and \(\phi\) within the manipulation space, \(\textbf{M}\).
-\item
-  \(\textbf{RM} = \textbf{M} * \textbf{R}\), the result of the
-  manipulation space after the application of the rotation matrix.
-\item
-  \(\textbf{P} = \textbf{X} * \textbf{RM}\), the data projection, the
-  data as projected through the rotated manipulation space.
+  \(\textbf{R}\), the 3D rotation matrix, for performing unconstrained
+  3D rotations within the manip space, \(\textbf{M}\). 
 \end{itemize}
 
-XXX Clean up the notation, not all of this is needed or useful. The
-shadow of \(e\) in the projection may need some notation. The projected
-data should be \(Y\) not \(P\).
-
-The algorithm operates entirely on projection bases and utilizes the
-data only later when making a display.
+The algorithm operates entirely on projection bases, and incorporates
+the data only when making the projected data plots.
 
@@
 
-The flea data, originally from Lubischew
-(\protect\hyperlink{ref-lubischew_use_1962}{1962}), made available in
-\pkg{tourr} is used to illustrate the algorithm. The data contains 74
-observations across 6 variables, which are physical measurements of the
-flea beetles. Each observation belongs to one of three species.
+The flea data (Lubischew
+(\protect\hyperlink{ref-lubischew_use_1962}{1962})), available in the
+\pkg{tourr} package, is used to illustrate the algorithm. The data
+contains 74 observations on 6 variables, which are physical measurements
+made on beetles. Each observation belongs to one of three species.
 
-An initial 2D projection basis set must be provided. One way to identify
-a projection containing interesting features is to apply a guided tour.
-In a guided tour, the projection sequence is selected by optimizing an
-index function via hill-climbing on the projection space. In this case,
-the holes index is selected and applied to the standardized flea data.
-The holes index is maximized when the projected observations are
-furthest from the center.
-
-Figure \ref{fig:step0} shows a locally optimized projection of the data.
-The left panel displays the reference axes of the projection basis, a
-visual indication of the magnitude and direction each variable
+An initial 2D projection basis must be provided. A suggested way to
+start is to identify an interesting projection using a projection
+pursuit guided tour. Here the holes index is used to find a 2D
+projection of the flea data, which shows three separated groups. Figure
+\ref{fig:step0} shows the initial projection of the data. The left panel
+displays the projection basis (\(\textbf{B}\)), and can be used as a
+visual guide of the magnitude and direction that each variable
 contributes to the projection. The right panel shows the data as
-projected through the basis set described by the reference axes (left).
-Data points are colored and given shape according to the species. Call
-\code{view\_basis()} on a basis to produce a \CRANpkg{ggplot2} (Wickham
-\protect\hyperlink{ref-wickham_ggplot2:_2016}{2016}) object similar to
-Figure \ref{fig:step0}. The projected data is processed for display
-using matrix multiplication,
-\(\textbf{P}_{[n,~d]} ~=~ \textbf{X}_{[n,~p]} ~*~ \textbf{B}_{[p,~d]}\).
+projected in \(\textbf{B}\). The colour and shape of points is mapped to
+the flea species. This plot is made using the \code{view\_basis()}
+function in \CRANpkg{spinifex}, which generated a \CRANpkg{ggplot2}
+(Wickham \protect\hyperlink{ref-wickham_ggplot2:_2016}{2016}) object.
+The projected data is processed for display using matrix multiplication,
+\(\textbf{Y}_{[n,~2]} ~=~ \textbf{X}_{[n,~p]} \textbf{B}_{[p,~2]}\).
 

### 2019-05-29 paper/spinifex_paper.Rmd
step 0 and notation

@@
 It takes the current contribution of the chosen variable, and using rotation brings this variable fully into the projection, completely removes it, and then returns to the original position.
-This is an overview of the steps:
+<!--This is an overview of the steps:
 
@@
 More detail is provided in the following sections. Function names are provided, but code structure and usage is discussed in section \ref{sec:usage}.
+-->
 
@@
 - $\textbf{X}$, the data, an $n \times p$ numeric matrix to be projected to 2D.
-- $\textbf{B} = (B_1,~ B_2)$, any 2D orthonormal projection basis, $p \times 2$ matrix, describing the projection from $p$ to two dimensions.
-We will call this the "original projection" because it is the starting point for the manual tour.
+- $\textbf{B} = (B_1,~ B_2)$, any 2D orthonormal projection basis, $p \times 2$ matrix, describing the projection from $\mathbb{R}^p \Rightarrow \mathbb{R}^2$.
+This is called this the "original projection" because it is the starting point for the manual tour.
 - $k$, is the index of the variable to manipulate, called the "manip var".
-- $|\textbf{B}_k|$, the norm of the manip var, $k$, the length of this basis vector between 0 and 1 inclusive.
-- $\textbf{e}$, a 1D coordinate zero basis vector of length $p$ with the $k$-th element set to one.
-- $\textbf{M} = orthonormalize(\textbf{B}|\textbf{e})$, the manipulation space, a $p \times 3$ matrix where out-of-plane rotation takes place. 
+<!--- $|\textbf{B}_k|$, the norm of the manip var, $k$, the length of this basis vector between 0 and 1 inclusive.
+-->
+- $\textbf{e}$, a 1D basis vector of length $p$, with 1 in the $k$-th position and 0 elsewhere.
+- $\textbf{M}$ is a $p \times 3$ matrix, defining the 3D subspace where data rotation occurs, and is called the manip(ulation) space. 
 - $\theta$, the angle of in-projection-plane rotation, for example, on the reference axes; $c_\theta, s_\theta$ are the cosine and sine.
-- $\phi$, the angle of out-of-projection-plane rotation, into the manipulation space; $c_\phi, s_\phi$ are the cosine and sine. The initial value for animation purposes is $\phi_1$.
-- $\textbf{U}$, the axis of rotation for $\phi$.
-- $\textbf{R}$, the rotation matrix, containing the rotaions of $\theta$ and $\phi$ within the manipulation space, $\textbf{M}$.
+- $\phi$, the angle of out-of-projection-plane rotation, into the manip space; $c_\phi, s_\phi$ are the cosine and sine. The initial value for animation purposes is $\phi_1$.
+- $\textbf{U}$, the axis of rotation for out-of-projection rotation.
+- $\textbf{R}$, the 3D rotation matrix, for performing unconstrained 3D rotations within the manip space, $\textbf{M}$.
+<!--
 - $\textbf{RM} = \textbf{M} * \textbf{R}$, the result of the manipulation space after the application of the rotation matrix.
@@
 XXX Clean up the notation, not all of this is needed or useful. The shadow of $e$ in the projection may need some notation. The projected data should be $Y$ not $P$. 
+-->
 
 <!-- operate on bases -->
-The algorithm operates entirely on projection bases and utilizes the data only later when making a display.
+The algorithm operates entirely on projection bases, and incorporates the data only when making the projected data plots.
 
@@
 <!-- describe data. -->
-The flea data, originally from @lubischew_use_1962, made available in \pkg{tourr} is used to illustrate the algorithm.
-The data contains 74 observations across 6 variables, which are physical measurements of the flea beetles.
+The flea data (@lubischew_use_1962), available in the \pkg{tourr} package, is used to illustrate the algorithm.
+The data contains 74 observations on 6 variables, which are physical measurements made on beetles.
 Each observation belongs to one of three species.
@@
 <!-- Projection basis -->
-An initial 2D projection basis set must be provided.
-One way to identify a projection containing interesting features is to apply a guided tour.
-In a guided tour, the projection sequence is selected by optimizing an index function via hill-climbing on the projection space.
-In this case, the holes index is selected and applied to the standardized flea data.
-The holes index is maximized when the projected observations are furthest from the center.
-
-Figure \ref{fig:step0} shows a locally optimized projection of the data.
-The left panel displays the reference axes of the projection basis, a visual indication of the magnitude and direction each variable contributes to the projection.
-The right panel shows the data as projected through the basis set described by the reference axes (left).
-Data points are colored and given shape according to the species.
-Call \code{view\_basis()} on a basis to produce a \CRANpkg{ggplot2} [@wickham_ggplot2:_2016] object similar to Figure \ref{fig:step0}.
-The projected data is processed for display using matrix multiplication, $\textbf{P}_{[n,~d]} ~=~ \textbf{X}_{[n,~p]} ~*~ \textbf{B}_{[p,~d]}$.
+An initial 2D projection basis must be provided.
+A suggested way to start is to identify an interesting projection using a projection pursuit guided tour.
+<!--In a guided tour, the projection sequence is selected by optimizing an index function via hill-climbing on the projection space.--> Here the holes index is used to find a 2D projection of the flea data, which shows three separated groups.
+Figure \ref{fig:step0} shows the initial projection of the data.
+The left panel displays the projection basis ($\textbf{B}$), and can be used as a visual guide of the magnitude and direction that each variable contributes to the projection.
+The right panel shows the data as projected in $\textbf{B}$.
+The colour and shape of points is mapped to the flea species.
+This plot is made using the  \code{view\_basis()} function in \CRANpkg{spinifex}, which generated a \CRANpkg{ggplot2} [@wickham_ggplot2:_2016] object.
+The projected data is processed for display using matrix multiplication, $\textbf{Y}_{[n,~2]} ~=~ \textbf{X}_{[n,~p]} \textbf{B}_{[p,~2]}$.
 

### 2019-05-29 paper/spinifex-paper.tex
step 1 and 2

@@
 visual guide of the magnitude and direction that each variable
-contributes to the projection. The right panel shows the data as
-projected in \(\textbf{B}\). The colour and shape of points is mapped to
-the flea species. This plot is made using the \code{view\_basis()}
-function in \CRANpkg{spinifex}, which generated a \CRANpkg{ggplot2}
-(Wickham \protect\hyperlink{ref-wickham_ggplot2:_2016}{2016}) object.
-The projected data is processed for display using matrix multiplication,
-\(\textbf{Y}_{[n,~2]} ~=~ \textbf{X}_{[n,~p]} \textbf{B}_{[p,~2]}\).
+contributes to the projection. The right panel shows the projected data,
+\(\textbf{Y}_{[n,~2]} ~=~ \textbf{X}_{[n,~p]} \textbf{B}_{[p,~2]}\). The
+colour and shape of points is mapped to the flea species. This plot is
+made using the \code{view\_basis()} function in \CRANpkg{spinifex},
+which generates a \CRANpkg{ggplot2} (Wickham
+\protect\hyperlink{ref-wickham_ggplot2:_2016}{2016}) object.
 
@@
 clusters. The variable aede2 is selected as the manip var, the variable
-to be controlled in the tour. The question that will be explored is how
-important this variable is to the separation of the clusters.
+to be controlled in the tour. The question that will be explored, is,
+how important this variable is to the separation of the clusters.
 
@@
 a Gram-Schmidt process to orthonormalize the coordinate basis vector on
-the original 2D projection to describe a 3D manipulation space,
-\(\textbf{M}\).
+the original 2D projection to describe a 3D manip space, \(\textbf{M}\).
 
@@
 where the coefficient of the manip var can range completely between
-{[}-1, 1{]}. 3D rotation can be used to rotate the manip variable
+{[}0, 1{]}. 3D rotation can be used to rotate the manip variable
 completely into or completely out of a 2D projection. Figure
-\ref{fig:step2} illustrates this 3D manipulation space with the manip
-var highlighted. This representation is produced by calling the
-\code{view\_manip\_space()} function. Functionally, the manipulation
-space adds another layer of complexity that needn't be displayed in the
-final animation, but this diagram aids in comprehension of the
-technique.
+\ref{fig:step2} illustrates this 3D manip space with the manip var
+highlighted. This representation is produced by calling the
+\code{view\_manip\_space()} function. Functionally, the manip space adds
+another layer of complexity that needn't be displayed in the final
+animation, but this diagram aids in comprehension of the technique.
 
@@
 
-\caption[Illustration of a 3D manipulation space, constructed to change the contribution of the variable aede2 in the example data]{Illustration of a 3D manipulation space, constructed to change the contribution of the variable aede2 in the example data.
+\caption[Illustration of a 3D manip space, constructed to change the contribution of the variable aede2 in the example data]{Illustration of a 3D manip space, constructed to change the contribution of the variable aede2 in the example data.
 The red circle indicates a unit sphere.
@@
 2D projection (Figure \ref{fig:step3}). As the variable is controlled,
-the manipulation space rotates, and the projection onto the horizontal
+the manip space rotates, and the projection onto the horizontal
 projection plane correspondingly changes. This is a manual tour.
 Generating a sequence of values for the horizontal and vertical angles
-produces a path for the rotation of the manipulation space. This defines
-the (orthonormally-constrained) rotation on the basis coefficients.
+produces a path for the rotation of the manip space. This defines the
+(orthonormally-constrained) rotation on the basis coefficients.
 
@@
 Rodrigues' rotation formula results in the rotation matrix, the rotation
-group SO(3). Continue with the rotation of the manipulation space,
+group SO(3). Continue with the rotation of the manip space,
 post-multiply the space by the above rotation matrix producing
@@
 For a 2D scatterplot, plot the first two variables within each frame.
-Show each frame in sequence to form an animation. The remaining
-manipulation space variable can be linked to point aesthetic (such as
-size or color) to produce depth cues used in conjunction with the \(XY\)
-scatterplot.
+Show each frame in sequence to form an animation. The remaining manip
+space variable can be linked to point aesthetic (such as size or color)
+to produce depth cues used in conjunction with the \(XY\) scatterplot.
 
@@
 tour preparation & view\_basis & displays the reference frame of the basis\\
-tour preparation & view\_manip\_space & displays the manipulation space\\
+tour preparation & view\_manip\_space & displays the manip space\\
 tour animation & play\_tour\_path & animates given tour path\\
@@
 plane where the manip var has a full contribution. This illustrates how
-the manip var is manipulated with the addition of the manipulation space
-as shown in Figure \ref{fig:step2}.
+the manip var is manipulated with the addition of the manip space as
+shown in Figure \ref{fig:step2}.
 
@@
 intermittent frames. This type of interpolation should not be used on
-manual tours, which have already been initialized into a 3D manipulation
-space where direct linear interpolation is appropriate.
+manual tours, which have already been initialized into a 3D manip space
+where direct linear interpolation is appropriate.
 
@@
 
-The selection of the manip variable is important, as the manipulation
-spaces convey substantially different information. The manual tour of
-PC6 offers information about the dimensionality, shape, and orientations
-of the different experiment classes. PC6 is particularly important to
+The selection of the manip variable is important, as the manip spaces
+convey substantially different information. The manual tour of PC6
+offers information about the dimensionality, shape, and orientations of
+the different experiment classes. PC6 is particularly important to
 describe the variations of DIS HERA1+2 and charm SIDIS observations,

### 2019-05-29 paper/spinifex_paper.Rmd
step 1 and 2

@@
 1. Provided with a 2D projection, choose a variable to manipulate. This is called the "manip var".
-2. Create a 3D manipulation space, where the manip variable has the full contribution.
+2. Create a 3D manip space, where the manip variable has the full contribution.
 3. Generate a rotation sequence which increases the norm of the coefficient to 1 and zeros it.
@@
 <!--
-- $\textbf{RM} = \textbf{M} * \textbf{R}$, the result of the manipulation space after the application of the rotation matrix.
-- $\textbf{P} = \textbf{X} * \textbf{RM}$, the data projection, the data as projected through the rotated manipulation space.
+- $\textbf{RM} = \textbf{M} * \textbf{R}$, the result of the manip space after the application of the rotation matrix.
+- $\textbf{P} = \textbf{X} * \textbf{RM}$, the data projection, the data as projected through the rotated manip space.
 
@@
 The left panel displays the projection basis ($\textbf{B}$), and can be used as a visual guide of the magnitude and direction that each variable contributes to the projection.
-The right panel shows the data as projected in $\textbf{B}$.
+The right panel shows the projected data, $\textbf{Y}_{[n,~2]} ~=~ \textbf{X}_{[n,~p]} \textbf{B}_{[p,~2]}$.
 The colour and shape of points is mapped to the flea species.
-This plot is made using the  \code{view\_basis()} function in \CRANpkg{spinifex}, which generated a \CRANpkg{ggplot2} [@wickham_ggplot2:_2016] object.
-The projected data is processed for display using matrix multiplication, $\textbf{Y}_{[n,~2]} ~=~ \textbf{X}_{[n,~p]} \textbf{B}_{[p,~2]}$.
+This plot is made using the  \code{view\_basis()} function in \CRANpkg{spinifex}, which generates a \CRANpkg{ggplot2} [@wickham_ggplot2:_2016] object.
 
@@
 The variable aede2 is selected as the manip var, the variable to be controlled in the tour.
-The question that will be explored is how important this variable is to the separation of the clusters.
+The question that will be explored, is, how important this variable is to the separation of the clusters.
 
@@
 In the example data, aede2 is the fifth variable in the data, so $k=5$, set $e_5=1$.
-Use a Gram-Schmidt process to orthonormalize the coordinate basis vector on the original 2D projection to describe a 3D manipulation space, $\textbf{M}$.
+Use a Gram-Schmidt process to orthonormalize the coordinate basis vector on the original 2D projection to describe a 3D manip space, $\textbf{M}$.
 
@@
 <!-- What the manip space provides -->
-The manip space provides a 3D projection from $p$-dimensional space, where the coefficient of the manip var can range completely between [-1, 1].
+The manip space provides a 3D projection from $p$-dimensional space, where the coefficient of the manip var can range completely between [0, 1].
 3D rotation can be used to rotate the manip variable completely into or completely out of a 2D projection.
-Figure \ref{fig:step2} illustrates this 3D manipulation space with the manip var highlighted.
+Figure \ref{fig:step2} illustrates this 3D manip space with the manip var highlighted.
 This representation is produced by calling the \code{view\_manip\_space()} function.
-Functionally, the manipulation space adds another layer of complexity that needn't be displayed in the final animation, but this diagram aids in comprehension of the technique.
+Functionally, the manip space adds another layer of complexity that needn't be displayed in the final animation, but this diagram aids in comprehension of the technique.
 
-```{r step2, echo=F, out.width='70%',                                           fig.cap = "Illustration of a 3D manipulation space, constructed to change the contribution of the variable aede2 in the example data.
+```{r step2, echo=F, out.width='70%',                                           fig.cap = "Illustration of a 3D manip space, constructed to change the contribution of the variable aede2 in the example data.
 The red circle indicates a unit sphere.
@@
 This is the process of the manual control, that rotates the manip variable into and out of the 2D projection (Figure \ref{fig:step3}).
-As the variable is controlled, the manipulation space rotates, and the projection onto the horizontal projection plane correspondingly changes.
+As the variable is controlled, the manip space rotates, and the projection onto the horizontal projection plane correspondingly changes.
 This is a manual tour.
-Generating a sequence of values for the horizontal and vertical angles produces a path for the rotation of the manipulation space.
+Generating a sequence of values for the horizontal and vertical angles produces a path for the rotation of the manip space.
 This defines the (orthonormally-constrained) rotation on the basis coefficients.
@@
 Defining the axis of rotation for $\phi$ and the application of Rodrigues' rotation formula results in the rotation matrix, the rotation group SO(3).
-Continue with the rotation of the manipulation space, post-multiply the space by the above rotation matrix producing **RM**, the rotated manip space.
+Continue with the rotation of the manip space, post-multiply the space by the above rotation matrix producing **RM**, the rotated manip space.
 This is basis values, the linear combination of variables, for one frame of the animation, repeat this process over a sequence of values for $\phi$ to produce an animation.
@@
 Show each frame in sequence to form an animation.
-The remaining manipulation space variable can be linked to point aesthetic (such as size or color) to produce depth cues used in conjunction with the $XY$ scatterplot.
+The remaining manip space variable can be linked to point aesthetic (such as size or color) to produce depth cues used in conjunction with the $XY$ scatterplot.
 
@@
   c("tour preparation", "view_basis", "displays the reference frame of the basis"),
-  c("tour preparation", "view_manip_space", "displays the manipulation space"),
+  c("tour preparation", "view_manip_space", "displays the manip space"),
   c("tour animation", "play_tour_path", "animates given tour path"),
@@
 Use \code{view\_manip\_space()} to view the new space with a dimension orthogonal to the projection plane where the manip var has a full contribution.
-This illustrates how the manip var is manipulated with the addition of the manipulation space as shown in Figure \ref{fig:step2}.
+This illustrates how the manip var is manipulated with the addition of the manip space as shown in Figure \ref{fig:step2}.
 
@@
 From there geodesic interpolation can be used to populate the intermittent frames.
-This type of interpolation should not be used on manual tours, which have already been initialized into a 3D manipulation space where direct linear interpolation is appropriate.
+This type of interpolation should not be used on manual tours, which have already been initialized into a 3D manip space where direct linear interpolation is appropriate.
 
@@
 <!-- comparison of the DIS cluster -->
-The selection of the manip variable is important, as the manipulation spaces convey substantially different information.
+The selection of the manip variable is important, as the manip spaces convey substantially different information.
 The manual tour of PC6 offers information about the dimensionality, shape, and orientations of the different experiment classes.

### 2019-05-29 paper/spinifex-paper.tex
step 3, 4

@@
 highlighted. This representation is produced by calling the
-\code{view\_manip\_space()} function. Functionally, the manip space adds
-another layer of complexity that needn't be displayed in the final
-animation, but this diagram aids in comprehension of the technique.
+\code{view\_manip\_space()} function. This diagram is purely used to
+help explain the algorithm.
 
@@
 
-\begin{Schunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{spinifex_paper_files/figure-latex/step3-1}
-
-}
-
-\caption[Snapshots of a radial manual tour manipulating aede2]{Snapshots of a radial manual tour manipulating aede2: (1) original projection, (2) full contribution, (3) zero contribution, (4) back to original.
-}\label{fig:step3}
-\end{figure}
-\end{Schunk}
-
 \hypertarget{step-3-defining-a-3d-rotation}{%
@@
 the manual control, that rotates the manip variable into and out of the
-2D projection (Figure \ref{fig:step3}). As the variable is controlled,
-the manip space rotates, and the projection onto the horizontal
-projection plane correspondingly changes. This is a manual tour.
-Generating a sequence of values for the horizontal and vertical angles
-produces a path for the rotation of the manip space. This defines the
-(orthonormally-constrained) rotation on the basis coefficients.
+2D projection (Figure \ref{fig:step3}). As the variable contribution is
+controlled, the manip space rotates, and the projection onto the
+horizontal projection plane correspondingly changes. This is a manual
+tour. Generating a sequence of values for the rotation angles produces a
+path for the rotation of the manip space.
 
-For a radial tour, fix the (horizontal) angle within the projection
-plane, \(\theta\), and define a sequence for the (vertical) angle,
-\(\phi\), defining movement out of the plane, that changes the initial
-value of \(\phi_1\), the angle between \(\textbf{e}\) and its shadow in
-\(\textbf{B}\), to a maximum of 1, and then to 0, before returning to
-\(\phi_1\).
+For a radial tour, fix \(\theta\), the angle describing rotation within
+the projection plane, and compute a sequence for \(\phi\), defining
+movement out of the plane. This will change the initial value,
+\(\phi_1\), the angle between \(\textbf{e}\) and its shadow in
+\(\textbf{B}\), to a maximum of \(0\) (manip var fully in projection),
+then to a minimum of \(\pi/2\) (manip var out of projection), before
+returning to \(\phi_1\).
 
-Rotations in 3D can be defined by the axes they pivot on. \(\theta\)
-describes the rotation around the \(Z\) axis. While \(\phi\) is the
-rotation around \(\textbf{U}\), an axis on the \(XY\) plane, orthogonal
-to \(\textbf{e}\). Given these axes, the SO(3) rotation matrix can be
-derived as follows:
+Rotations in 3D can be defined by the axes they pivot on. Rotation
+within the projection, \(\theta\), is rotation around the \(Z\) axis.
+Out-of-projection rotation, \(\phi\), is the rotation around an axis on
+the \(XY\) plane, \(\textbf{U}\), orthogonal to \(\textbf{e}\). Given
+these axes, the rotation matrix, \(\textbf{R}\) can be written as
+follows, using Rodrigues' rotation formula (originally pubished in
+Rodrigues (\protect\hyperlink{ref-rodrigues_lois_1840}{1840})):
 
 \begin{align*}
-  \textbf{U} &= (u_x, u_y, u_z) =
-  (s_\theta, -c_\theta, 0) \\ 
-  &=
-  \begin{bmatrix}
-  0 & -u_z & u_y  \\
-  u_z & 0 & -u_x \\
-  -u_y & u_x & 0 \\
-  \end{bmatrix} =
-  \begin{bmatrix}
-    0 & 0 & -c_\theta \\
-    0 & 0 & -s_\theta \\
-    c_\theta & s_\theta & 0 \\
-  \end{bmatrix} \\
-  \end{align*} Apply Rodrigues' rotation formula (Rodrigues
-\protect\hyperlink{ref-rodrigues_lois_1840}{1840}): \begin{align*}
     \textbf{R}_{[3,~3]} 
@@
 
-Defining the axis of rotation for \(\phi\) and the application of
-Rodrigues' rotation formula results in the rotation matrix, the rotation
-group SO(3). Continue with the rotation of the manip space,
-post-multiply the space by the above rotation matrix producing
-\textbf{RM}, the rotated manip space. This is basis values, the linear
-combination of variables, for one frame of the animation, repeat this
-process over a sequence of values for \(\phi\) to produce an animation.
+\noindent where
 
-\hypertarget{creating-the-animation-of-the-radial-rotation}{%
-\subsection{Creating the animation of the radial
-rotation}\label{creating-the-animation-of-the-radial-rotation}}
+\begin{align*}
+  \textbf{U} &= (u_x, u_y, u_z) =
+  (s_\theta, -c_\theta, 0) \\ 
+  &=
+  \begin{bmatrix}
+  0 & -u_z & u_y  \\
+  u_z & 0 & -u_x \\
+  -u_y & u_x & 0 \\
+  \end{bmatrix} =
+  \begin{bmatrix}
+    0 & 0 & -c_\theta \\
+    0 & 0 & -s_\theta \\
+    c_\theta & s_\theta & 0 \\
+  \end{bmatrix} \\
+  \end{align*}
+
+\hypertarget{step-4-creating-an-animation-of-the-radial-rotation}{%
+\subsubsection{Step 4) Creating an animation of the radial
+rotation}\label{step-4-creating-an-animation-of-the-radial-rotation}}
 
@@
 into the projection, completely out, and then back to the initial value.
-This involves allowing \(\phi\), to vary between 0 and 1.
+This involves allowing \(\phi\), to vary between 0 and \(\pi/2\), call
+the steps \(\phi_i\).
 
-\hypertarget{sequence-of-phi_i}{%
-\subsubsection{\texorpdfstring{Sequence of
-\(\phi_i\)}{Sequence of \textbackslash phi\_i}}\label{sequence-of-phi_i}}
+\begin{Schunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{spinifex_paper_files/figure-latex/step3-1}
+
+}
+
+\caption[Snapshots of a radial manual tour manipulating aede2]{Snapshots of a radial manual tour manipulating aede2: (1) original projection, (2) full contribution, (3) zero contribution, (4) back to original.
+}\label{fig:step3}
+\end{figure}
+\end{Schunk}
 
@@
   \(\phi_1\) is between 0 and \(\pi/2\), and if not, transform it into
-  this range, by subtracting or adding \(\pi\) XXX.
+  this range, by subtracting or adding \(\pi\). (XXX Nick needs to check
+  this math, and the ways to check and fix if angle produced by the
+  calculation given fall in the right range.)
 \item
-  Set an angle increment (\(\Delta_\phi\)) that sets the step size to
-  make, to rotate the manip var into and out of the projection. (Note:
-  Using angle increment, rather than a number of steps, to control the
-  movement, is consistent with the tour algorithm as implemented in the
-  \pkg{tourr}).
+  Set an angle increment (\(\Delta_\phi\)) that sets the step size for
+  the animation, to rotate the manip var into and out of the projection.
+  (Note: Using angle increment, rather than a number of steps, to
+  control the movement, is consistent with the tour algorithm as
+  implemented in the \pkg{tourr}).
 \item
-  Step towards \(0\), where the manip var has a completely in the
+  Step towards \(0\), where the manip var is a completely in the
   projection.
 \item
-  Step towards \(\pi/2\) where the manip variable has no contribution.
+  Step towards \(\pi/2\), where the manip variable has no contribution.
 \item
@@
 
-In each of the steps 3-5, a small step may be appended to bring \(\phi\)
-exactly to the desired value.
+In each of the steps 3-5, a small step may be added to ensure that the
+end points of \(\phi\) (0, \(\pi/2\)) are reached.
 
-Figure \ref{fig:step4} illustrates a sequence with 15 projected bases,
-showing the reference axes on the top half with the corresponding
-projected data points below. Changes in \(\textbf{B}_k\) correspond to
-the distance separating the purple cluster and the remaining sample.
-This shows aede2 is crucial in distinguishing the variation of species.
+\hypertarget{sec:display}{%
+\subsubsection{Step 4) Projecting the data}\label{sec:display}}
+
+The operation of a manual tour is defined on the projection bases. Only
+when the data plot needs to be made is the data projected into the
+relevant basis.
+
+\begin{align*}
+  \textbf{Y}_{[n,~3]} &= \textbf{X}_{[n,~p]} \textbf{M}_{[p,~3]} \textbf{R}^{(i)}_{[3,3]}
+\end{align*}
+
+Plot the first two columns of \textbf{Y} to make the data plot. Show the
+projected data for each frame in sequence to form an animation.
+
+Figure \ref{fig:step4} illustrates a manual tour sequence having 15
+steps. The projection axes are displayed on the top half, which
+correspond to the projected data in the bottom half. When aede2 is
+removed from the projection, the purple cluster overlaps with the green
+cluster. This suggests that aede2 is important for distinguishing this
+species.
+
 Tours are typically viewed as an animation. The animation of this tour
@@
 \caption[Radial manual tour manipulating aede2 of standardized flea data]{Radial manual tour manipulating aede2 of standardized flea data.
-The contributions increase from its initial vlaue to a full contribution to the projection before decreasing to zero and then returning to the initial value.
-Manipulating the coefficeients of aede2 changes the distance separating the purple cluster.
-This shows that aede2 is important for distinguishing this species from the ramaining data.
-Animated versions made with the \code{play\_manual\_tour()} function and can be viewed at https://nspyrison.netlify.com/thesis/flea\_manualtour\_mvar5/.}\label{fig:step4}
+The axis for aede2 increases in contribution to the projection, from its initial value to 1, decreasing to 0 and then returning to the initial value.
+This affects the separatiion between the purple and green clusters.
+This shows that aede2 is important for distinguishing the purple species, because the separation disappears when aede2 is not contributing to the projection.
+An animation can be viewed at https://nspyrison.netlify.com/thesis/flea\_manualtour\_mvar5/.}\label{fig:step4}
 \end{figure}
 \end{Schunk}
-
-\hypertarget{sec:display}{%
-\subsection{Projecting the data}\label{sec:display}}
-
-In an appeal to performance, the operations of a tour are performed on
-the bases while the larger datasets are not operated on until the
-display of data points are needed. After the bases are brought into the
-projection-space, however, it is helpful to observe them with projected
-data in the same space. Pre-multiply the data by basis frame bringing
-the data into the projection space.
-
-\begin{align*}
-  \textbf{P}_{[n,~3]} &= \textbf{X}_{[n,~p]} ~*~ \textbf{RM}_{[p,~3]}
-\end{align*}
-
-For a 2D scatterplot, plot the first two variables within each frame.
-Show each frame in sequence to form an animation. The remaining manip
-space variable can be linked to point aesthetic (such as size or color)
-to produce depth cues used in conjunction with the \(XY\) scatterplot.
 

### 2019-05-29 paper/spinifex_paper.Rmd
step 3, 4

@@
 This representation is produced by calling the \code{view\_manip\_space()} function.
-Functionally, the manip space adds another layer of complexity that needn't be displayed in the final animation, but this diagram aids in comprehension of the technique.
+This diagram is purely used to help explain the algorithm.
 
@@
 
-```{r step3, echo=F, warning=F, fig.height=2, fig.width=8, out.width='100%',  fig.cap = "Snapshots of a radial manual tour manipulating aede2: (1) original projection, (2) full contribution, (3) zero contribution, (4) back to original.
-"}
-f_angle <- .26
-f_mtour <- manual_tour(f_bas, manip_var = f_mvar, angle = f_angle)
-if(dim(f_mtour)[3] != 15) message(
-  paste0("!!!!! step3 n_slides != 15 !!!!! Please knit again. n_slides = ", dim(f_mtour)[3]))
-p1 <- view_basis2(f_mtour[,,1],  colnames(f_dat), manip_var = f_mvar) + ggtitle("1")
-p2 <- view_basis2(f_mtour[,,6],  colnames(f_dat), manip_var = f_mvar) + ggtitle("2")
-p3 <- view_basis2(f_mtour[,,12], colnames(f_dat), manip_var = f_mvar) + ggtitle("3")
-p4 <- view_basis2(f_mtour[,,15], colnames(f_dat), manip_var = f_mvar) + ggtitle("4")
-grid.arrange(p1, p2, p3, p4, ncol=4)
-```
-
 ### Step 3) Defining a 3D rotation
@@
 This is the process of the manual control, that rotates the manip variable into and out of the 2D projection (Figure \ref{fig:step3}).
-As the variable is controlled, the manip space rotates, and the projection onto the horizontal projection plane correspondingly changes.
+As the variable contribution is controlled, the manip space rotates, and the projection onto the horizontal projection plane correspondingly changes.
 This is a manual tour.
-Generating a sequence of values for the horizontal and vertical angles produces a path for the rotation of the manip space.
-This defines the (orthonormally-constrained) rotation on the basis coefficients.
+Generating a sequence of values for the rotation angles produces a path for the rotation of the manip space.
+<!--This defines the (orthonormally-constrained) rotation on the basis coefficients.-->
 
 <!-- describe manip var path -->
-For a radial tour, fix the (horizontal) angle within the projection plane, $\theta$, and define a sequence for the (vertical) angle, $\phi$, defining movement out of the plane, that changes the initial value of $\phi_1$, the angle between $\textbf{e}$ and its shadow in $\textbf{B}$, to a maximum of 1, and then to 0, before returning to $\phi_1$.
+For a radial tour, fix $\theta$, the angle describing rotation within the projection plane, and compute a sequence for $\phi$, defining movement out of the plane.
+This will change the initial value, $\phi_1$, the angle between $\textbf{e}$ and its shadow in $\textbf{B}$, to a maximum of $0$ (manip var fully in projection), then to a minimum of $\pi/2$ (manip var out of projection), before returning to $\phi_1$.
 
@@
 Rotations in 3D can be defined by the axes they pivot on.
-$\theta$ describes the rotation around the $Z$ axis.
-While $\phi$ is the rotation around $\textbf{U}$, an axis on the $XY$ plane, orthogonal to $\textbf{e}$.
-Given these axes, the SO(3) rotation matrix can be derived as follows:
+Rotation within the projection, $\theta$, is rotation around the $Z$ axis.
+Out-of-projection rotation, $\phi$, is the rotation around an axis on the $XY$ plane, $\textbf{U}$, orthogonal to $\textbf{e}$.
+Given these axes, the rotation matrix, $\textbf{R}$ can be written as follows, using Rodrigues' rotation formula (originally pubished in @rodrigues_lois_1840):
 
-\begin{align*}
-  \textbf{U} &= (u_x, u_y, u_z) =
-  (s_\theta, -c_\theta, 0) \\ 
-  &=
-  \begin{bmatrix}
-  0 & -u_z & u_y  \\
-  u_z & 0 & -u_x \\
-  -u_y & u_x & 0 \\
-  \end{bmatrix} =
-  \begin{bmatrix}
-    0 & 0 & -c_\theta \\
-    0 & 0 & -s_\theta \\
-    c_\theta & s_\theta & 0 \\
-  \end{bmatrix} \\
-  \end{align*}
-Apply Rodrigues' rotation formula [@rodrigues_lois_1840]:
   \begin{align*}
@@
 \end{align*}
+
+\noindent where
+
+\begin{align*}
+  \textbf{U} &= (u_x, u_y, u_z) =
+  (s_\theta, -c_\theta, 0) \\ 
+  &=
+  \begin{bmatrix}
+  0 & -u_z & u_y  \\
+  u_z & 0 & -u_x \\
+  -u_y & u_x & 0 \\
+  \end{bmatrix} =
+  \begin{bmatrix}
+    0 & 0 & -c_\theta \\
+    0 & 0 & -s_\theta \\
+    c_\theta & s_\theta & 0 \\
+  \end{bmatrix} \\
+  \end{align*}
+
 <!-- The term $(1-c_\phi)$ is used by convention, but $2sin^2(\phi/2)$ is more computationally robust. -->
@@
 <!-- rotate manip space -->
-Defining the axis of rotation for $\phi$ and the application of Rodrigues' rotation formula results in the rotation matrix, the rotation group SO(3).
+<!--Defining the axis of rotation for $\phi$ and the application of Rodrigues' rotation formula results in the rotation matrix, the rotation group SO(3).
 Continue with the rotation of the manip space, post-multiply the space by the above rotation matrix producing **RM**, the rotated manip space.
 This is basis values, the linear combination of variables, for one frame of the animation, repeat this process over a sequence of values for $\phi$ to produce an animation.
+-->
 
@@
 
-## Creating the animation of the radial rotation
+### Step 4) Creating an animation of the radial rotation
 
@@
 To use these for sensitivity analysis, the radial rotation is built into an animation where the manip var is rotated fully into the projection, completely out, and then back to the initial value.
-This involves allowing $\phi$, to vary between 0 and 1.
+This involves allowing $\phi$, to vary between 0 and $\pi/2$, call the steps $\phi_i$.
 
@@
 
-### Sequence of $\phi_i$
+```{r step3, echo=F, warning=F, fig.height=2, fig.width=8, out.width='100%',  fig.cap = "Snapshots of a radial manual tour manipulating aede2: (1) original projection, (2) full contribution, (3) zero contribution, (4) back to original.
+"}
+f_angle <- .26
+f_mtour <- manual_tour(f_bas, manip_var = f_mvar, angle = f_angle)
+if(dim(f_mtour)[3] != 15) message(
+  paste0("!!!!! step3 n_slides != 15 !!!!! Please knit again. n_slides = ", dim(f_mtour)[3]))
+p1 <- view_basis2(f_mtour[,,1],  colnames(f_dat), manip_var = f_mvar) + ggtitle("1 phi=0.78")
+p2 <- view_basis2(f_mtour[,,6],  colnames(f_dat), manip_var = f_mvar) + ggtitle("2  phi=1.57")
+p3 <- view_basis2(f_mtour[,,12], colnames(f_dat), manip_var = f_mvar) + ggtitle("3 phi=0")
+p4 <- view_basis2(f_mtour[,,15], colnames(f_dat), manip_var = f_mvar) + ggtitle("4 phi=0.78")
+grid.arrange(p1, p2, p3, p4, ncol=4)
+```
+
+<!--### Sequence of $\phi_i$-->
 
@@
 $\phi_1$ is the angle between $\textbf{e}$ and its shadow in $\textbf{B}$.
-Check that  $\phi_1$ is between 0 and $\pi/2$, and if not,  transform it into this range, by subtracting or adding $\pi$ XXX.
+Check that  $\phi_1$ is between 0 and $\pi/2$, and if not, transform it into this range, by subtracting or adding $\pi$.
+(XXX Nick needs to check this math, and the ways to check and fix if angle produced by the calculation given fall in the right range.)
 2.
-Set an angle increment ($\Delta_\phi$) that sets the step size to make, to rotate the manip var into and out of the projection.
+Set an angle increment ($\Delta_\phi$) that sets the step size for the animation, to rotate the manip var into and out of the projection.
 (Note: Using angle increment, rather than a number of steps, to control the movement, is consistent with the tour algorithm as implemented in the \pkg{tourr}).
-3. Step towards $0$, where the manip var has a completely in the projection.
-4. Step towards $\pi/2$ where the manip variable has no contribution.
+3. Step towards $0$, where the manip var is a completely in the projection.
+4. Step towards $\pi/2$, where the manip variable has no contribution.
 5. Step back to $\phi_1$.
 
-In each of the steps 3-5, a small step may be appended to bring $\phi$ exactly to the desired value.
+In each of the steps 3-5, a small step may be added to ensure that the end points of $\phi$ (0, $\pi/2$) are reached.
+
+### Step 4) Projecting the data {#sec:display}
+
+<!-- the reminder of basis operation and now apply data -->
+The operation of a manual tour is defined on the projection bases. Only when the data plot needs to be made is the data projected into the relevant basis.
+
+\begin{align*}
+  \textbf{Y}_{[n,~3]} &= \textbf{X}_{[n,~p]} \textbf{M}_{[p,~3]} \textbf{R}^{(i)}_{[3,3]}
+\end{align*}
+
+<!-- plot 2 var in seq for animation -->
+Plot the first two columns of \textbf{Y} to make the data plot.
+Show the projected data for each frame in sequence to form an animation.
+<!--The remaining manip space variable can be linked to point aesthetic (such as size or color) to produce depth cues used in conjunction with the $XY$ scatterplot.-->
 
 <!-- Animation on flea data -->
-Figure \ref{fig:step4} illustrates a sequence with 15 projected bases, showing the reference axes on the top half with the corresponding projected data points below.
-Changes in $\textbf{B}_k$ correspond to the distance separating the purple cluster and the remaining sample.
-This shows aede2 is crucial in distinguishing the variation of species.
+Figure \ref{fig:step4} illustrates a manual tour sequence having 15 steps.
+The projection axes are displayed on the top half, which correspond to the projected data in the bottom half.
+When aede2 is removed from the projection, the purple cluster overlaps with the green cluster.
+This suggests that aede2 is important for distinguishing this species.
+
 Tours are typically viewed as an animation.
@@
 ```{r step4, echo=F, warning=F, fig.height=7.0, fig.width=5.83, out.height='7in', out.width='5.83in',                                                                fig.cap = "Radial manual tour manipulating aede2 of standardized flea data.
-The contributions increase from its initial vlaue to a full contribution to the projection before decreasing to zero and then returning to the initial value.
-Manipulating the coefficeients of aede2 changes the distance separating the purple cluster.
-This shows that aede2 is important for distinguishing this species from the ramaining data.
-Animated versions made with the \\code{play\\_manual\\_tour()} function and can be viewed at https://nspyrison.netlify.com/thesis/flea\\_manualtour\\_mvar5/."}
+The axis for aede2 increases in contribution to the projection, from its initial value to 1, decreasing to 0 and then returning to the initial value.
+This affects the separatiion between the purple and green clusters.
+This shows that aede2 is important for distinguishing the purple species, because the separation disappears when aede2 is not contributing to the projection.
+An animation can be viewed at https://nspyrison.netlify.com/thesis/flea\\_manualtour\\_mvar5/."}
 #play_manual_tour(f_dat, f_bas, f_mvar, col = f_cat, angle = f_angle)
@@
 -->
-
-## Projecting the data {#sec:display}
-
-<!-- the reminder of basis operation and now apply data -->
-In an appeal to performance, the operations of a tour are performed on the bases while the larger datasets are not operated on until the display of data points are needed.
-After the bases are brought into the projection-space, however, it is helpful to observe them with projected data in the same space.
-Pre-multiply the data by basis frame bringing the data into the projection space.
-
-\begin{align*}
-  \textbf{P}_{[n,~3]} &= \textbf{X}_{[n,~p]} ~*~ \textbf{RM}_{[p,~3]}
-\end{align*}
-
-<!-- plot 2 var in seq for animation -->
-For a 2D scatterplot, plot the first two variables within each frame.
-Show each frame in sequence to form an animation.
-The remaining manip space variable can be linked to point aesthetic (such as size or color) to produce depth cues used in conjunction with the $XY$ scatterplot.
 

### 2019-05-30 paper/spinifex_paper.Rmd
step 3, 4
start of usage section

@@
 p1 <- view_basis2(f_mtour[,,1],  colnames(f_dat), manip_var = f_mvar) + ggtitle("1 phi=0.78")
-p2 <- view_basis2(f_mtour[,,6],  colnames(f_dat), manip_var = f_mvar) + ggtitle("2  phi=1.57")
-p3 <- view_basis2(f_mtour[,,12], colnames(f_dat), manip_var = f_mvar) + ggtitle("3 phi=0")
+p2 <- view_basis2(f_mtour[,,6],  colnames(f_dat), manip_var = f_mvar) + ggtitle("2  phi=0")
+p3 <- view_basis2(f_mtour[,,12], colnames(f_dat), manip_var = f_mvar) + ggtitle("3 phi=1.57")
 p4 <- view_basis2(f_mtour[,,15], colnames(f_dat), manip_var = f_mvar) + ggtitle("4 phi=0.78")
@@
 <!-- plot 2 var in seq for animation -->
-Plot the first two columns of \textbf{Y} to make the data plot.
+\noindent where $\textbf{R}^{(i)}_{[3,3]}$ is the incremental rotation matrix, using $\phi_i$.
+To make the data plot, use the first two columns of \textbf{Y}.
 Show the projected data for each frame in sequence to form an animation.
@@
 
-This section discusses the functions and code usage of the package before delving into a domain-specific application of the manual tour.
+This section describes the functions available in the package, and how to use them.
 
@@
 
+The \CRANpkg{spinifex} is available from CRAN, and can be installed by: 
 ```{r installCODEEXAMPLE, eval=F, echo=T}
-# remotes::install_github("nspyrison/spinifex") # Development version
 install.package("spinifex")
@@
 
-# Also see the vignette for further code usage: 
+# Also see the vignette for details on using the code: 
 vignette("spinifex_vignette")
+```
+
+The development version can be installed from github:
+
+```
+remotes::install_github("nspyrison/spinifex") 
 ```

### 2019-05-30 paper/spinifex-paper.tex
step 3, 4
usages section

@@
 
-Plot the first two columns of \textbf{Y} to make the data plot. Show the
-projected data for each frame in sequence to form an animation.
+\noindent where \(\textbf{R}^{(i)}_{[3,3]}\) is the incremental rotation
+matrix, using \(\phi_i\). To make the data plot, use the first two
+columns of \textbf{Y}. Show the projected data for each frame in
+sequence to form an animation.
 
@@
 
-This section discusses the functions and code usage of the package
-before delving into a domain-specific application of the manual tour.
+This section describes the functions available in the package, and how
+to use them.
 
@@
 
+The \CRANpkg{spinifex} is available from CRAN, and can be installed by:
+
 \begin{Schunk}
 \begin{Sinput}
-# remotes::install_github("nspyrison/spinifex") # Development version
 install.package("spinifex")
 library("spinifex")
-
-# Also see the vignette for further code usage: 
-vignette("spinifex_vignette")
 \end{Sinput}
@@
 
-\hypertarget{primary-package-functions}{%
-\subsection{Primary package functions}\label{primary-package-functions}}
+\noindent Also see the vignette for examples of usage:
 
-The primary functions that will aid users in performing their own manual
-tours are broken into the following groupings: tour preparation, tour
-animation, and utility. In the following section describes and
-demonstrates the usage of the above functions following the order in the
-algorithm. The utility functions are called by the tour animation
-functions, passed to the argument \code{render\_type}. Usage is
-demonstrated below and in the example documentation.
+\begin{verbatim}
+vignette("spinifex_vignette")
+\end{verbatim}
+
+\noindent The development version can be installed from github:
+
+\begin{verbatim}
+remotes::install_github("nspyrison/spinifex") 
+\end{verbatim}
+
+\hypertarget{functions}{%
+\subsection{Functions}\label{functions}}
+
+Table \ref{tab:functionsTable} lists the primary functions and their
+purpose. These are grouped into four types: construction for building a
+tour path, render to make the plot objects, animation for running the
+animation, and specialty for providing illustrations used in the
+algorithm description.
 
@@
 
-\caption{\label{tab:functionsTable}Primary functions}
+\caption{\label{tab:functionsTable}Summary of available functions.}
 \centering
@@
 \midrule
-tour preparation & view\_basis & displays the reference frame of the basis\\
-tour preparation & view\_manip\_space & displays the manip space\\
-tour animation & play\_tour\_path & animates given tour path\\
-tour animation & play\_manual\_tour & animates the manual tour algorithm\\
-utility & render\_gganimate & uses gganimate to render the animation\\
-\addlinespace
-utility & render\_plotly & uses plotly to render the animation\\
+construction & create\_manip\_space & forms the 3D space of rotation\\
+construction & rotate\_manip\_space & performs 3D rotation\\
+construction & manual\_tour & generates sequence of 2D frames\\
+ &  & \\
+render & render\_ & constructs the ggplot object to feed to animation\\
+render & render\_plotly & converts the ggplot object to a plotly animation\\
+render & render\_gganimate & converts the ggplot object to a gganimate  animation\\
+render & array2df & turns a tour path array into long form, for plotting\\
+ &  & \\
+animation & play\_tour\_path & animates given tour path\\
+animation & play\_manual\_tour & animates the manual tour algorithm\\
+ &  & \\
+specialty & view\_basis & displays the reference frame of a given basis\\
+specialty & view\_manip\_space & illustrative display of any manip space\\
 \bottomrule

### 2019-05-30 paper/spinifex_paper.Rmd
usages section

@@
 library("spinifex")
+```
 
-# Also see the vignette for details on using the code: 
+\noindent Also see the vignette for examples of usage:
+
+```
 vignette("spinifex_vignette")
@@
 
-The development version can be installed from github:
+\noindent The development version can be installed from github:
 
@@
 
-## Primary package functions 
+## Functions 
 <!-- in terms of work flow. -->
 
+Table \ref{tab:functionsTable} lists the primary functions and their purpose.
+These are grouped into four types: construction for building a tour path, render to make the plot objects, animation for running the animation, and specialty for providing illustrations used in the algorithm description.
+
+<!--
 The primary functions that will aid users in performing their own manual tours are broken into the following groupings: tour preparation, tour animation, and utility.
@@
 The utility functions are called by the tour animation functions, passed to the argument \code{render\_type}.
-Usage is demonstrated below and in the example documentation.
+Usage is demonstrated below and in the example documentation.-->
 
@@
 funcs <- rbind(
-  c("tour preparation", "view_basis", "displays the reference frame of the basis"),
-  c("tour preparation", "view_manip_space", "displays the manip space"),
-  c("tour animation", "play_tour_path", "animates given tour path"),
-  c("tour animation", "play_manual_tour", "animates the manual tour algorithm"),
+  c("construction", "create_manip_space", "forms the 3D space of rotation"),
+  c("construction", "rotate_manip_space", "performs 3D rotation"),
+  c("construction", "manual_tour", "generates sequence of 2D frames"),
+  c("", "", ""),
+  c("render", "render_", "constructs the ggplot object to feed to animation"),
+  c("render", "render_plotly", "converts the ggplot object to a plotly animation"),
+  c("render", "render_gganimate", "converts the ggplot object to a gganimate  animation"),
+  c("render", "array2df", "turns a tour path array into long form, for plotting"),
+  c("", "", ""),
+  c("animation", "play_tour_path", "animates given tour path"),
+  c("animation", "play_manual_tour", "animates the manual tour algorithm"),
+  c("", "", ""),
   # c("utility", "col_of", "maps a variable to the point color"),
   # c("utility", "pch_of", "maps a variable to the point character (shape)"),
-  c("utility", "render_gganimate", "uses gganimate to render the animation"),
-  c("utility", "render_plotly", "uses plotly to render the animation")
+  c("specialty", "view_basis", "displays the reference frame of a given basis"),
+  c("specialty", "view_manip_space", "illustrative display of any manip space")
 )
-colnames(funcs) <- c("Class", "Function", "Description")
+colnames(funcs) <- c("Type", "Function", "Description")
 
-kable(funcs, "latex", caption = "Primary functions", booktabs = T)
+kable(funcs, "latex", caption = "Summary of available functions.", booktabs = T, linesep="") 
 ```
 
-## Algorithm code
+## Usage
 

### 2019-05-30 paper/spinifex-paper.tex
tidying usage

@@
 \toprule
-Class & Function & Description\\
+Type & Function & Description\\
 \midrule
@@
 
-\hypertarget{algorithm-code}{%
-\subsection{Algorithm code}\label{algorithm-code}}
+\hypertarget{usage}{%
+\subsection{Usage}\label{usage}}
 
-We'll start by initializing values including a standardized data set
-(numeric columns only), a starting basis, a categorical variable for
-point aesthetics (optional), and a manip var. To get bearings on the
-projection, start by observing the reference axes of the basis with
-\code{view\_basis()} producing Figure \ref{fig:step0}.
+Using the \texttt{flea} data from the \CRANpkg{tourr} package, we will
+illustrate generating a manual tour to explore the sensitivity of the
+cluster structure is to the variable \texttt{aede2}.
 
@@
 \begin{Sinput}
+library(spinifex)
 f_data  <- tourr::rescale(flea[, 1:6])                    ## standardize data
-f_path  <- save_history(f_data, guided_tour(holes()))     ## produce guided tour
-f_basis <- matrix(f_path[,, max(dim(f_path)[3])], ncol=2) ## end of guided tour
+f_path  <- save_history(f_data, guided_tour(holes()))     ## guided tour using holes index
+f_basis <- matrix(f_path[,, max(dim(f_path)[3])], ncol=2) ## optimal projection
 f_cat   <- factor(flea$species)                           ## categorical var
-
-view_basis(basis = f_basis, 
-           data = f_data,
-           labels = colnames(f_data))
-\end{Sinput}
-\end{Schunk}
-
-After becoming familiar with this space, select a manip var, the
-variable to change the contributions of. Use \code{view\_manip\_space()}
-to view the new space with a dimension orthogonal to the projection
-plane where the manip var has a full contribution. This illustrates how
-the manip var is manipulated with the addition of the manip space as
-shown in Figure \ref{fig:step2}.
-
-\begin{Schunk}
-\begin{Sinput}
-f_mvar  <- 5  ## manip var number
-
-view_manip_space(basis = f_basis, 
-                 manip_var = f_mvar, 
-                 labels = colnames(f_data))
-\end{Sinput}
-\end{Schunk}
-
-Now we are ready to perform a manual tour on the selected variable. Use
-\code{play\_manual\_tour()} to perform the algorithm as discussed above,
-in section \ref{sec:algorithm}. This is the animated equivalent of
-Figure \ref{fig:step3}.
-
-\begin{Schunk}
-\begin{Sinput}
+f_mvar  <- 5                                              ## set manip var
 angle_speed <- .26
-
-play_manual_tour(data = f_data,
-                 basis = f_basis, 
+play_manual_tour(data = f_data,                           ## Generates animation
+                 basis = f_basis,                         ## using plotly
                  manip_var = f_mvar, 
@@
 
-This concludes the content of the algorithm section, however, lets cover
-animating paths generated in \pkg{tourr}. Animate the previously
-generated guided tour path via \code{play\_tour\_path()}. Utility
-functions can also be passed as arguments into either of the tour
-animation function to change the resulting graphics object, set
-\code{render\_type = render\_gganimate} to view the animation as a GIF.
+The \texttt{play\_manual\_tour} function is a composite function for
+user convenience that does these operations:
 
@@
 \begin{Sinput}
-play_tour_path(tour_path = f_path,
+manual_tour()
+array2df()
+render_plotly()
+\end{Sinput}
+\end{Schunk}
+
+\hypertarget{extras}{%
+\subsubsection{Extras}\label{extras}}
+
+\begin{Schunk}
+\begin{Sinput}
+view_basis(basis = f_basis, 
            data = f_data,
-               render_type = render_gganimate, 
-               angle = angle_speed
-)
+           labels = colnames(f_data))
+\end{Sinput}
+\end{Schunk}
+
+\begin{Schunk}
+\begin{Sinput}
+
+view_manip_space(basis = f_basis, 
+                 manip_var = f_mvar, 
+                 labels = colnames(f_data))
 \end{Sinput}

### 2019-05-30 paper/spinifex_paper.Rmd
tidying usage

@@
 
+Using the `flea` data from the \CRANpkg{tourr} package, we will illustrate generating a manual tour to explore the sensitivity of the cluster structure is to the variable `aede2`.
+
+<!--
 We'll start by initializing values including a standardized data set (numeric columns only), a starting basis, a categorical variable for point aesthetics (optional), and a manip var.
-To get bearings on the projection, start by observing the reference axes of the basis with \code{view\_basis()} producing Figure \ref{fig:step0}.
+To get bearings on the projection, start by observing the reference axes of the basis with \code{view\_basis()} producing Figure \ref{fig:step0}.-->
 
@@
 ```{r step0CODEEXAMPLE, eval=F, echo=T}
+library(spinifex)
 f_data  <- tourr::rescale(flea[, 1:6])                    ## standardize data
-f_path  <- save_history(f_data, guided_tour(holes()))     ## produce guided tour
-f_basis <- matrix(f_path[,, max(dim(f_path)[3])], ncol=2) ## end of guided tour
+f_path  <- save_history(f_data, guided_tour(holes()))     ## guided tour using holes index
+f_basis <- matrix(f_path[,, max(dim(f_path)[3])], ncol=2) ## optimal projection
 f_cat   <- factor(flea$species)                           ## categorical var
-
-view_basis(basis = f_basis, 
-           data = f_data,
-           labels = colnames(f_data))
-```
-
-After becoming familiar with this space, select a manip var, the variable to change the contributions of.
-Use \code{view\_manip\_space()} to view the new space with a dimension orthogonal to the projection plane where the manip var has a full contribution.
-This illustrates how the manip var is manipulated with the addition of the manip space as shown in Figure \ref{fig:step2}.
-
-<!-- step 2 disp code. -->
-```{r step2CODEEXAMPLE, eval=F, echo=T}
-f_mvar  <- 5  ## manip var number
-
-view_manip_space(basis = f_basis, 
-                 manip_var = f_mvar, 
-                 labels = colnames(f_data))
-```
-
-Now we are ready to perform a manual tour on the selected variable.
-Use \code{play\_manual\_tour()} to perform the algorithm as discussed above, in section \ref{sec:algorithm}.
-This is the animated equivalent of Figure \ref{fig:step3}.
-
-<!-- step 3 disp code. -->
-```{r step3CODEEXAMPLE, eval=F, echo=T}
+f_mvar  <- 5                                              ## set manip var
 angle_speed <- .26
-
-play_manual_tour(data = f_data,
-                 basis = f_basis, 
+play_manual_tour(data = f_data,                           ## Generates animation
+                 basis = f_basis,                         ## using plotly
                  manip_var = f_mvar, 
@@
 
+<!--After becoming familiar with this space, select a manip var, the variable to change the contributions of.
+Use \code{view\_manip\_space()} to view the new space with a dimension orthogonal to the projection plane where the manip var has a full contribution.
+This illustrates how the manip var is manipulated with the addition of the manip space as shown in Figure \ref{fig:step2}.
+
+Now we are ready to perform a manual tour on the selected variable.
+Use \code{play\_manual\_tour()} to perform the algorithm as discussed above, in section \ref{sec:algorithm}.
+This is the animated equivalent of Figure \ref{fig:step3}.
+-->
+
+The `play_manual_tour` function is a composite function for user convenience that does these operations:
+
+```{r eval=FALSE, echo=TRUE}
+manual_tour()
+array2df()
+render_plotly()
+```
+
+<!--
 This concludes the content of the algorithm section, however, lets cover animating paths generated in \pkg{tourr}.
@@
 )
+```
+-->
+
+### Extras
+
+```{r eval=FALSE}
+view_basis(basis = f_basis, 
+           data = f_data,
+           labels = colnames(f_data))
+```
+<!-- step 2 disp code. -->
+```{r step2CODEEXAMPLE, eval=F, echo=T}
+
+view_manip_space(basis = f_basis, 
+                 manip_var = f_mvar, 
+                 labels = colnames(f_data))
 ```

### 2019-05-31 paper/spinifex-paper.tex
done with usage section

@@
 
-\hypertarget{extras}{%
-\subsubsection{Extras}\label{extras}}
+\noindent This will generate an html animation using plotly. Switching
+the renderer to \texttt{gganimate} is an option, which will also create
+an html animation, or an animated gif. Each of these formats allows for
+the animation to be made available on a web site, or directly visible in
+an html formatted document.
+
+\hypertarget{making-illustrations}{%
+\subsubsection{Making illustrations}\label{making-illustrations}}
+
+The function \texttt{view\_basis} can be used to show a projection, just
+the basis, or with the data overlaid. For example, the plots in Figures
+\ref{fig:step0} and \ref{fig:step3} were made with code similar to this:
 
@@
 
+\noindent An illustration of the manip space (as shown in Figure
+\ref{fig:step2}) is made with the \texttt{view\_manip\_space} function,
+as follows:
+
 \begin{Schunk}
@@
 \end{Schunk}
-
-\hypertarget{rendering-and-sharing}{%
-\subsection{Rendering and sharing}\label{rendering-and-sharing}}
-
-The \pkg{tourr} package utilizes \pkg{base} graphics for the display of
-tours. \pkg{spinifex} allows tours to be rendered in \pkg{plotly} as an
-HTML5 object or \pkg{gganimate} as GIF or MP4 files. Sharing of
-animations is not trivial, especially static formats like print and PDF.
-Even with dynamic display capturing the correct resolution and aspect
-ratio can be challenging, while many formats quickly bloat file sizes
-limiting sharing options. Keep in mind hosting options and exporting
-functions offered in \pkg{plotly}, \pkg{gganimate} and \pkg{tourr}.
-
-\hypertarget{storage}{%
-\subsection{Storage}\label{storage}}
-
-Storing each data point for every frame of the animation is redundant.
-Just as operations are performed on the bases, so too should tour paths
-be stored as bases and a single instance of the data. Consider a radial
-manual tour, we can store the salient features in 3 bases, where
-\(\phi\) is at its starting, minimum, and maximum values. The frames in
-between can be interpolated by supplying angular speed. With the use of
-the \code{tourr::save\_history()} function, the target bases can be
-saved. From there geodesic interpolation can be used to populate the
-intermittent frames. This type of interpolation should not be used on
-manual tours, which have already been initialized into a 3D manip space
-where direct linear interpolation is appropriate.
 

### 2019-05-31 paper/spinifex_paper.Rmd
done with usage section

@@
 
+\noindent This will generate an html animation using plotly.
+Switching the renderer to `gganimate` is an option, which will also create an html animation, or an animated gif.
+Each of these formats allows for the animation to be made available on a web site, or directly visible in an html formatted document.
+
 <!--
@@
 
-### Extras
+### Making illustrations
+
+The function `view_basis` can be used to show a projection, just the basis, or with the data overlaid.
+For example, the plots in Figures \ref{fig:step0} and \ref{fig:step3} were made with code similar to this:
 
@@
 ```
+
+\noindent An illustration of the manip space (as shown in Figure \ref{fig:step2}) is made with the `view_manip_space` function, as follows:
+
 <!-- step 2 disp code. -->
@@
 
+<!--
 ## Rendering and sharing
@@
 This type of interpolation should not be used on manual tours, which have already been initialized into a 3D manip space where direct linear interpolation is appropriate.
+-->
 

### 2019-05-31 paper/spinifex-paper.tex
done with application

@@
 for a single variable to be controlled. The variable is rotated fully
-into the projection with a coefficient of 1 or -1, or completely out of
-the projection with a coefficient of 0. The resulting sequence of
-projections can be displayed using animation, with functions from either
-the \pkg{plotly} and \pkg{gganimate} packages. By varying the
-coefficient of a single variable, it is possible to explore the
+into the projection, or completely out of the projection. The resulting
+sequence of projections can be displayed using animation, with functions
+from either the \pkg{plotly} and \pkg{gganimate} packages. By varying
+the coefficient of a single variable, it is possible to explore the
 sensitivity of structure in the projection to that variable. It is
 particularly useful when used with a projection pursuit guided tour to
-simplify and understand the solution. The use of the manual tour is
-illustrated using a problem from particle physics.
+simplify and understand the solution. The use of the manual tour for
+studying the sensitivity of structure, in a projection, to specific
+variable contributions, is illustrated using particle physics data.
 }
@@
 glimpse of the high-dimensional space. An alternative approach is to use
-a tour of dynamic linear projections, to look at many different
-low-dimensional projections. Tours can be considered to extend the
-dimensionality of visualization, which is important as data and models
-exist in more than 3D. The package \CRANpkg{tourr} (Wickham et al.
+a tour (Asimov \protect\hyperlink{ref-asimov_grand_1985}{1985}) of
+dynamic linear projections, to look at many different low-dimensional
+projections. Tours can be considered to extend the dimensionality of
+visualization, which is important as data and models exist in more than
+3D. The package \CRANpkg{tourr} (Wickham et al.
 \protect\hyperlink{ref-wickham_tourr_2011}{2011}) provides a platform
@@
 
-In a recent paper, Wang et al.
-(\protect\hyperlink{ref-wang_mapping_2018}{2018}), the authors aggregate
-and visualize the sensitivity of hadronic experiments to nucleon
-structure. The authors introduce a new tool, PDFSense, to aid in the
-visualization of parton distribution functions (PDF). The
-parameter-space of these experiments lies in 56 dimensions,
-\(\delta \in \mathbb{R}^{56}\), and are visualized as 3D subspaces of
-the 10 first principal components in linear (PCA) and non-linear (t-SNE)
-embeddings.
+Wang et al. (\protect\hyperlink{ref-wang_mapping_2018}{2018}) introduce
+a new tool, PDFSense, to visualize the sensitivity of hadronic
+experiments to nucleon structure. The parameter-space of these
+experiments lies in 56 dimensions, \(\delta \in \mathbb{R}^{56}\), and
+are visualized as 3D subspaces of the 10 first principal components in
+linear (PCA) and non-linear (t-SNE) embeddings.
 
-Using the same data, another study, Cook, Laa, and Valencia
-(\protect\hyperlink{ref-cook_dynamical_2018}{2018}), applied grand tours
-(Asimov \protect\hyperlink{ref-asimov_grand_1985}{1985}) to the same
-subspaces. Grand tours are dynamic linear projections of high
-dimensional spaces where basis sets are selected at random and animated
-with geodesic interpolation of the intermediate frames. Because of the
-change in the basis, or orientation to the subspace, tours are able to
-better resolve the distribution shape of clusters, intra-cluster detail,
-and better outlier detection than the use of PDFSense \& TFEP
-(TensorFlow embedded projections) or traditional static embeddings.
-Before applying manual tours the structure of the data is discussed.
+Cook, Laa, and Valencia
+(\protect\hyperlink{ref-cook_dynamical_2018}{2018}) illustrates how to
+learn more about the structures using a grand tour. Tours are able to
+better resolve the shape of clusters, intra-cluster detail, and better
+outlier detection than PDFSense \& TFEP (TensorFlow embedded
+projections) or traditional static embeddings. This example builds from
+here, illustrating how the manual tour can be used to examine the
+sensitivity of structure in a projection to different parameters. The
+specific 2D projections passed to the manual tour were provided from
+Cook, Laa, and Valencia
+(\protect\hyperlink{ref-cook_dynamical_2018}{2018}).
 
@@
 \protect\hyperlink{ref-cook_dynamical_2018}{2018}). The data within this
-4D embedding is subset down to ATLAS7old and ATLAS7new to focus in on
-two groups with a reasonable number of observations that occupy
-different parts of the subspace. Radial manual tours controlling
-contributions from PC4 and PC3 are shown in Figure
-\ref{fig:JetClusterGood} and Figure \ref{fig:JetClusterBad}
-respectively. These variables are selected to contrast the information
-conveyed by different manip variables. Links to dynamic HTML5 animations
-controlling each of the four variables are also provided.
+4D embedding is further subsetted, to ATLAS7old and ATLAS7new, in order
+to focus on two groups, with a reasonable number of observations, that
+occupy different parts of the subspace. Radial manual tours controlling
+contributions from PC4 and PC3 are shown in Figures
+\ref{fig:JetClusterGood} and \ref{fig:JetClusterBad}, respectively. The
+difference in shape can be interpreted as the experiments probing
+different phase-spaces. Back-transforming the principal components to
+the original variables can be done for more detailed interpretation.
 
-When manipulating PC4, there is a clear difference in the parameter
-space spanned by the experiment types ATLAS7new and ATLAS7old.
-Specifically, the variation of ATLAS7new (green) becomes more singular.
-The experiments are probing different parameter space and PC4 is
-important to demonstrate this. Yet, when PC3 is manipulated there is no
-clear indication that the different experiments probe different
-parameter space. Performing a radial manual tour on PC4 is more
-insightful than for PC3. Radial manual tours manipulating each of the
-principal components in the jet cluster can be viewed by following the
-links:
+When PC4 is removed from the projection (Figure
+\ref{fig:JetClusterGood}), the difference between the two groups is
+removed, indicating that it is important for distinguishing experiments.
+However, removing PC3 from the projection (Figure
+\ref{fig:JetClusterBad}), has no effect on the structure, indicating it
+is not important for distinguishing experiments. Animations for the
+remaining PCs can be viewed at the following links:
 \href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc1/}{PC1},
@@
 \href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc4/}{PC4}.
+It can be seen that only PC4 is important for viewing the difference in
+these two experiments.
 
@@
 
-\caption[A radial manual tour of PC4 within the jet cluster]{A radial manual tour of PC4 within the jet cluster.
-Colored by experiment type: ATLAS7new in green and ATLAS7old in orange.
-When PC4 fully/negligibly contributes to the projection ATLAS7new (green) spans the same space as the orange points.
-During the intermediate frames, the ATLAS7new is compressed in the direction radial to PC4.
-The differece in distribution shape demonstrates the experiments probe different phase-space, which has a linear mapping back to the original variables for interpretation and further exploration.
-An HTML5 version can be viewed at https://nspyrison.netlify.com/thesis/jetcluster\_manualtour\_pc4/.}\label{fig:JetClusterGood}
+\caption[Snapshots of a radial manual tour of PC4 focusing on the jet cluster, with color indicating experiment type]{Snapshots of a radial manual tour of PC4 focusing on the jet cluster, with color indicating experiment type: ATLAS7new (green) and ATLAS7old (orange).
+When PC4 is removed from the projection (frame 10) there is little difference between the groups, suggesting that PC4 is important for distinguishing the experiments.
+The animation can be viewed at https://nspyrison.netlify.com/thesis/jetcluster\_manualtour\_pc4/.}\label{fig:JetClusterGood}
 \end{figure}
@@
 
-\caption[A radial manual tour of PC4 within the jet cluster]{A radial manual tour of PC4 within the jet cluster.
-Colored by experiment type: ATLAS7new in green and ATLAS7old in orange.
-Data from ATLAS7new (green) spans mostly the same space as ALTLAS7old (orange) with no evident difference in cluster structure across varying contributions of PC3.
-An HTML5 version can be viewed at https://nspyrison.netlify.com/thesis/jetcluster\_manualtour\_pc3/.}\label{fig:JetClusterBad}
+\caption[Snapshots of a radial manual tour of PC3 focusing on the jet cluster, with color indicating experiment type]{Snapshots of a radial manual tour of PC3 focusing on the jet cluster, with color indicating experiment type: ATLAS7new (green) and ATLAS7old (orange).
+When the contribution of PC3 is changed there is little change to the structure of the two groups, suggesting that PC3 is not important for distinguishing the experiments.
+The animation can be viewed at https://nspyrison.netlify.com/thesis/jetcluster\_manualtour\_pc3/.}\label{fig:JetClusterBad}
 \end{figure}
@@
 
-A different space is used to explore the DIS cluster; specifically the
-first six principal components, which explains 48\% of the variation
-contained within the aggregated data (Cook, Laa, and Valencia
-\protect\hyperlink{ref-cook_dynamical_2018}{2018}). Radial manual tours
-are performed on PC6 and PC2 in Figure \ref{fig:DISclusterGood} and
-Figure \ref{fig:DISclusterBad} respectively.
+Following Cook, Laa, and Valencia
+(\protect\hyperlink{ref-cook_dynamical_2018}{2018}), to explore the DIS
+cluster, PCA is recomputed and the first six principal components,
+explaining 48\% of the variation is used. The contributions of PC6 and
+PC2 are explored in Figures \ref{fig:DISclusterGood} and
+\ref{fig:DISclusterBad}, respectively. Three experiments are examined:
+DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange).
 
-The selection of the manip variable is important, as the manip spaces
-convey substantially different information. The manual tour of PC6
-offers information about the dimensionality, shape, and orientations of
-the different experiment classes. PC6 is particularly important to
-describe the variations of DIS HERA1+2 and charm SIDIS observations,
-Whereas manipulating the contributions of PC2 only shows a subset of the
-dimensionality and shape information. Manipulating the contributions of
-PC6 turned out to be much more insightful than PC2. This result might
-seem counter-intuitive at first as PC2 should explain much more of the
-variation in the data. However, features and structures in the data
-regularly reside in smaller dimensionality which can be overlooked when
-optimizing on full sample statistics. DIS cluster manual tours
-manipulating each of the principal components can be viewed from the
-links:
+Both PC2 and PC6 contribute to the projection similarly. When PC6 is
+rotated into the projection, variation in the DIS HERA1+2 is greatly
+reduced. When PC2 is removed from the projection, dimuon SIDIS becomes
+more clearly distinct. Even though both variables contribute similarly
+to the original projection, their contributions have quite different
+effects on the structure of each cluster, and the distinction between
+clusters. Animations of all of the principal components can be viewed
+from the links:
 \href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc1/}{PC1},
@@
 
-\caption[A radial manual tour manipulating the contribution of PC6 within the DIS cluster]{A radial manual tour manipulating the contribution of PC6 within the DIS cluster.
-Points are colored by experiment type: DIS HERA1+2 in green, dimuon SIDIS in purple, and charm SIDIS in orange.
-The cluster DIS HERA1+2 (green) is distributed in a cross-shaped plane, charm SIDIS (orange) occupies the center space of this cross.
-As the contribution of PC6 becomes whole the distributions of DIS HERA1+2 (green) and charm SIDIS (orange) become singular but offset by a small angle.
-Less evident is the linear dimuon SIDIS (purple) observations approaching the line of view for intermediate values of PC6.
-An HTML5 version can be viewed at https://nspyrison.netlify.com/thesis/discluster\_manualtour\_pc6/.}\label{fig:DISclusterGood}
+\caption[Snapshots of a radial manual tour exploring the contribution of PC6 for the DIS cluster, with color indicating experiment type]{Snapshots of a radial manual tour exploring the contribution of PC6 for the DIS cluster, with color indicating experiment type: DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange).
+DIS HERA1+2 is distributed in a cross-shaped plane, charm SIDIS occupies the center of this cross, and dimuon SIDIS is a linear cluster crossing DIS HERA1+2.
+As PC6's contribution is increased, DIS HERA1+2 becomes almost singular in one direction (frame 5), indicating that this experiment has very little variability in the direction of PC6.
+The animation can be viewed at https://nspyrison.netlify.com/thesis/discluster\_manualtour\_pc6/.}\label{fig:DISclusterGood}
 \end{figure}
@@
 
-\caption[A radial manual tour manipulating the contribution of PC2 within the DIS cluster]{A radial manual tour manipulating the contribution of PC2 within the DIS cluster.
-Points are colored by experiment type: DIS HERA1+2 in green, dimuon SIDIS in purple, and charm SIDIS in orange.
-The plane of cross distributed DIS HERA data (green) and a nearly orthogonal jet of dimuon SIDIS (purple) is present.
-This jet does extend more in the plane of view when the contribution of PC2 is full, giving insight to its orientation.
-However, less information about the shape of DIS HERA (green) and charm SIDIS (orange) is available compared to PC6 as the manip var.
-An HTML5 version can be viewed at https://nspyrison.netlify.com/thesis/discluster\_manualtour\_pc2/.}\label{fig:DISclusterBad}
+\caption[Snapshots of a radial manual tour exploring the contribution of PC2 for the DIS cluster, with color indicating experiment type]{Snapshots of a radial manual tour exploring the contribution of PC2 for the DIS cluster, with color indicating experiment type: DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange).
+As PC2's contribution is decreased, dimuon SIDIS becomes more distinguishable from the other two clusters (frames 10-14), indicating that in its absence PC2 is important.
+The animation can be viewed at https://nspyrison.netlify.com/thesis/discluster\_manualtour\_pc2/.}\label{fig:DISclusterBad}
 \end{figure}

### 2019-05-31 paper/spinifex_paper.Rmd
done with application

@@
 This paper describes a new package, \pkg{spinifex}, to provide a manual tour, that allows the coefficient for a single variable to be controlled.
-The variable is rotated fully into the projection with a coefficient of 1 or -1, or completely out of the projection with a coefficient of 0.
+The variable is rotated fully into the projection, or completely out of the projection.
 The resulting sequence of projections can be displayed using animation, with functions from either the \pkg{plotly} and \pkg{gganimate} packages.
@@
 It is particularly useful when used with a projection pursuit guided tour to simplify and understand the solution.
-The use of the manual tour is illustrated using a problem from particle physics.
+The use of the manual tour for studying the sensitivity of structure, in a projection, to specific variable contributions, is illustrated using particle physics data.
 
@@
 These are useful for finding relationships between multiple variables, but they are limited because they show only a glimpse of the high-dimensional space.
-An alternative approach is to use a tour of dynamic linear projections, to look at many different low-dimensional projections.
+An alternative approach is to use a tour [@asimov_grand_1985] of dynamic linear projections, to look at many different low-dimensional projections.
 Tours can be considered to extend the dimensionality of visualization, which is important as data and models exist in more than 3D.
@@
 <!-- Introduction of data and original paper -->
-In a recent paper, @wang_mapping_2018, the authors aggregate and visualize the sensitivity of hadronic experiments to nucleon structure.
-The authors introduce a new tool, PDFSense, to aid in the visualization of parton distribution functions (PDF).
+@wang_mapping_2018 introduce a new tool, PDFSense, to visualize the sensitivity of hadronic experiments to nucleon structure.
 The parameter-space of these experiments lies in 56 dimensions, $\delta \in \mathbb{R}^{56}$, and are visualized as 3D subspaces of the 10 first principal components in linear (PCA) and non-linear (t-SNE) embeddings.
@@
 <!-- grand tours on the same data -->
-Using the same data, another study, @cook_dynamical_2018, applied grand tours [@asimov_grand_1985] to the same subspaces.
-Grand tours are dynamic linear projections of high dimensional spaces where basis sets are selected at random and animated with geodesic interpolation of the intermediate frames.
-Because of the change in the basis, or orientation to the subspace, tours are able to better resolve the distribution shape of clusters, intra-cluster detail, and better outlier detection than the use of PDFSense & TFEP (TensorFlow embedded projections) or traditional static embeddings.
-Before applying manual tours the structure of the data is discussed.
+@cook_dynamical_2018 illustrates how to learn more about the structures using a grand tour.
+Tours are able to better resolve the shape of clusters, intra-cluster detail, and better outlier detection than PDFSense & TFEP (TensorFlow embedded projections) or traditional static embeddings.
+This example builds from here, illustrating how the manual tour can be used to examine the sensitivity of structure in a projection to different parameters.
+The specific 2D projections passed to the manual tour were provided from @cook_dynamical_2018.
 
@@
 The jet cluster resides in a smaller dimensionality than the full set of experiments with four principal components explaining 95% of the variation in the cluster [@cook_dynamical_2018].
-The data within this 4D embedding is subset down to ATLAS7old and ATLAS7new to focus in on two groups with a reasonable number of observations that occupy different parts of the subspace.
-Radial manual tours controlling contributions from PC4 and PC3 are shown in Figure \ref{fig:JetClusterGood} and Figure \ref{fig:JetClusterBad} respectively.
-These variables are selected to contrast the information conveyed by different manip variables.
-Links to dynamic HTML5 animations controlling each of the four variables are also provided.
+The data within this 4D embedding is further subsetted, to ATLAS7old and ATLAS7new, in order to focus on two groups, with a reasonable number of observations, that occupy different parts of the subspace.
+Radial manual tours controlling contributions from PC4 and PC3 are shown in Figures \ref{fig:JetClusterGood} and  \ref{fig:JetClusterBad}, respectively.
+The difference in shape can be interpreted as the experiments probing different phase-spaces.
+Back-transforming the principal components to the original variables can be done for more detailed interpretation.
+<!--These variables are selected to contrast the information conveyed by different manip variables. Links to dynamic HTML5 animations controlling each of the four variables are also provided.-->
 
 <!-- discussion of findings and which is more insightful -->
-When manipulating PC4, there is a clear difference in the parameter space spanned by the experiment types ATLAS7new and ATLAS7old.
+When PC4 is removed from the projection (Figure \ref{fig:JetClusterGood}), the difference between the two groups is removed, indicating that it is important for distinguishing experiments.
+However, removing PC3 from the projection (Figure \ref{fig:JetClusterBad}), has no effect on the structure, indicating it is not important for distinguishing experiments.
+<!--, there is a clear difference in the parameter space spanned by the experiment types ATLAS7new and ATLAS7old.
 Specifically, the variation of ATLAS7new (green) becomes more singular.
@@
 Yet, when PC3 is manipulated there is no clear indication that the different experiments probe different parameter space.
-Performing a radial manual tour on PC4 is more insightful than for PC3.
-Radial manual tours manipulating each of the principal components in the jet cluster can be viewed by following the links: [PC1](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc1/), [PC2](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc2/), [PC3](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/), and [PC4](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc4/).
+Performing a radial manual tour on PC4 is more insightful than for PC3.-->
+Animations for the remaining PCs can be viewed at the following links: [PC1](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc1/), [PC2](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc2/), [PC3](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/), and [PC4](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc4/).
+It can be seen that only PC4 is important for viewing the difference in these two experiments.
 
-```{r JetClusterGood, echo=F, warning=F, fig.height=7.0, fig.width=5.83, out.height='7in', out.width='5.83in',                                     fig.cap="A radial manual tour of PC4 within the jet cluster.
-Colored by experiment type: ATLAS7new in green and ATLAS7old in orange.
-When PC4 fully/negligibly contributes to the projection ATLAS7new (green) spans the same space as the orange points.
-During the intermediate frames, the ATLAS7new is compressed in the direction radial to PC4.
-The differece in distribution shape demonstrates the experiments probe different phase-space, which has a linear mapping back to the original variables for interpretation and further exploration.
-An HTML5 version can be viewed at https://nspyrison.netlify.com/thesis/jetcluster\\_manualtour\\_pc4/."}
+```{r JetClusterGood, echo=F, warning=F, fig.height=7.0, fig.width=5.83, out.height='7in', out.width='5.83in',                                     fig.cap="Snapshots of a radial manual tour of PC4 focusing on the jet cluster, with color indicating experiment type: ATLAS7new (green) and ATLAS7old (orange).
+When PC4 is removed from the projection (frame 10) there is little difference between the groups, suggesting that PC4 is important for distinguishing the experiments.
+The animation can be viewed at https://nspyrison.netlify.com/thesis/jetcluster\\_manualtour\\_pc4/."}
 # Jet cluster, Fig 7 of cook_dynamical_2018, subset to ATLAS7old and ATLAS7new
@@
 
-```{r JetClusterBad, echo=F, warning=F, fig.height=7.0, fig.width=5.83, out.height='7in', out.width='5.83in',                                       fig.cap = "A radial manual tour of PC4 within the jet cluster.
-Colored by experiment type: ATLAS7new in green and ATLAS7old in orange.
-Data from ATLAS7new (green) spans mostly the same space as ALTLAS7old (orange) with no evident difference in cluster structure across varying contributions of PC3.
-An HTML5 version can be viewed at https://nspyrison.netlify.com/thesis/jetcluster\\_manualtour\\_pc3/."}
+```{r JetClusterBad, echo=F, warning=F, fig.height=7.0, fig.width=5.83, out.height='7in', out.width='5.83in',                                       fig.cap = "Snapshots of a radial manual tour of PC3 focusing on the jet cluster, with color indicating experiment type: ATLAS7new (green) and ATLAS7old (orange).
+When the contribution of PC3 is changed there is little change to the structure of the two groups, suggesting that PC3 is not important for distinguishing the experiments.
+The animation can be viewed at https://nspyrison.netlify.com/thesis/jetcluster\\_manualtour\\_pc3/."}
 # Jet cluster, Fig 7 of cook_dynalical_2018, subset to ATLAS7old and ATLAS7new
@@
 <!-- introduce DIS cluster -->
-A different space is used to explore the DIS cluster; specifically the first six principal components, which explains 48% of the variation contained within the aggregated data [@cook_dynamical_2018].
-Radial manual tours are performed on PC6 and PC2 in Figure \ref{fig:DISclusterGood} and Figure \ref{fig:DISclusterBad} respectively.
+Following @cook_dynamical_2018, to explore the DIS cluster, PCA is recomputed and the first six principal components, explaining 48% of the variation is used.
+The contributions of PC6 and PC2 are explored in Figures \ref{fig:DISclusterGood} and  \ref{fig:DISclusterBad}, respectively.
+Three experiments are examined: DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange).
 
 <!-- comparison of the DIS cluster -->
-The selection of the manip variable is important, as the manip spaces convey substantially different information.
-The manual tour of PC6 offers information about the dimensionality, shape, and orientations of the different experiment classes.
-PC6 is particularly important to describe the variations of DIS HERA1+2 and charm SIDIS observations,  Whereas manipulating the contributions of PC2 only shows a subset of the dimensionality and shape information.
-Manipulating the contributions of PC6 turned out to be much more insightful than PC2.
-This result might seem counter-intuitive at first as PC2 should explain much more of the variation in the data.
-However, features and structures in the data regularly reside in smaller dimensionality which can be overlooked when optimizing on full sample statistics.
-DIS cluster manual tours manipulating each of the principal components can be viewed from the links: [PC1](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc1/), [PC2](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc2/), [PC3](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc3/), [PC4](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc4/), [PC5](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc5/), and [PC6](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/).
+Both PC2 and PC6 contribute to the projection similarly.
+When PC6 is rotated into the projection, variation in the DIS HERA1+2 is greatly reduced.
+When PC2 is removed from the projection, dimuon SIDIS becomes more clearly distinct.
+Even though both variables contribute similarly to the original projection, their contributions have quite different effects on the structure of each cluster, and the distinction between clusters.
+Animations of all of the principal components can be viewed from the links: [PC1](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc1/), [PC2](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc2/), [PC3](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc3/), [PC4](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc4/), [PC5](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc5/), and [PC6](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/).
 
-```{r DISclusterGood, echo=F, warning=F, fig.height=7.0, fig.width=5.83, out.height='7in', out.width='5.83in',                                            fig.cap = "A radial manual tour manipulating the contribution of PC6 within the DIS cluster.
-Points are colored by experiment type: DIS HERA1+2 in green, dimuon SIDIS in purple, and charm SIDIS in orange.
-The cluster DIS HERA1+2 (green) is distributed in a cross-shaped plane, charm SIDIS (orange) occupies the center space of this cross.
-As the contribution of PC6 becomes whole the distributions of DIS HERA1+2 (green) and charm SIDIS (orange) become singular but offset by a small angle.
-Less evident is the linear dimuon SIDIS (purple) observations approaching the line of view for intermediate values of PC6.
-An HTML5 version can be viewed at https://nspyrison.netlify.com/thesis/discluster\\_manualtour\\_pc6/."}
+```{r DISclusterGood, echo=F, warning=F, fig.height=7.0, fig.width=5.83, out.height='7in', out.width='5.83in',                                            fig.cap = "Snapshots of a radial manual tour exploring the contribution of PC6 for the DIS cluster, with color indicating experiment type: DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange).
+DIS HERA1+2 is distributed in a cross-shaped plane, charm SIDIS occupies the center of this cross, and dimuon SIDIS is a linear cluster crossing DIS HERA1+2.
+As PC6's contribution is increased, DIS HERA1+2 becomes almost singular in one direction (frame 5), indicating that this experiment has very little variability in the direction of PC6.
+The animation can be viewed at https://nspyrison.netlify.com/thesis/discluster\\_manualtour\\_pc6/."}
 
@@
 
-```{r DISclusterBad, echo=F, warning=F, fig.height=7.0, fig.width=5.83, out.height='7in', out.width='5.83in',                                            fig.cap = "A radial manual tour manipulating the contribution of PC2 within the DIS cluster.
-Points are colored by experiment type: DIS HERA1+2 in green, dimuon SIDIS in purple, and charm SIDIS in orange.
-The plane of cross distributed DIS HERA data (green) and a nearly orthogonal jet of dimuon SIDIS (purple) is present.
-This jet does extend more in the plane of view when the contribution of PC2 is full, giving insight to its orientation.
-However, less information about the shape of DIS HERA (green) and charm SIDIS (orange) is available compared to PC6 as the manip var.
-An HTML5 version can be viewed at https://nspyrison.netlify.com/thesis/discluster\\_manualtour\\_pc2/."}
+```{r DISclusterBad, echo=F, warning=F, fig.height=7.0, fig.width=5.83, out.height='7in', out.width='5.83in',                                            fig.cap = "Snapshots of a radial manual tour exploring the contribution of PC2 for the DIS cluster, with color indicating experiment type: DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange).
+As PC2's contribution is decreased, dimuon SIDIS becomes more distinguishable from the other two clusters (frames 10-14), indicating that in its absence PC2 is important.
+The animation can be viewed at https://nspyrison.netlify.com/thesis/discluster\\_manualtour\\_pc2/."}
 ##DIS cluster, fig 8 of cook_dynamical_2018

### 2019-05-31 paper/spinifex-paper.tex
done with all

@@
 
-Tours, which are a dynamic linear projection of numeric multivariate
-data, play an important role in data visualization; they extend the
-dimensionality of visuals while data- and parameter-spaces become ever
-larger. This research has modified the algorithm producing manual tours
-which and has made this functionality available in package
-\pkg{spinifex}. The package adds to \pkg{tourr}, extending the graphics
-offerings that can be used to display tours.
+Dynamic linear projections of numeric multivariate data, tours, play an
+important role in data visualization; they extend the dimensionality of
+visuals to peek into high-dimensional data and parameter spaces. This
+research has taken the manual tour algorithm, specifically the radial
+rotation, used in GGobi (Swayne et al.
+\protect\hyperlink{ref-swayne_ggobi:_2003}{2003}) to interactively
+rotate a variable into or out of a 2D projection, and modified it to
+create an animation that performs the same task. It is most useful for
+examining the importance of variables, and how the structure in the
+projection is sensitive or not to specific variables. This functionality
+available in package \CRANpkg{spinifex}. The work complements the
+methods available in the \CRANpkg{tourr} package.
 
-Radial manual tours were applied to a dataset across different
-experiments of hadronic collisions. The importance of selecting the
-correct manip var, as demonstrated by comparing tours of varying amounts
-of structural information. Manual tours, by controlling the contribution
-of the manip var on the projection, enable analysts to explore the
-sensitivity that variable has on the structure of the projection. This
-information can be used by domain experts to identify which variables
-are probing which parameter spaces and how sensitive structural features
-are to different variables. This insight can be indispensable for
-variable inclusion/exclusion, and in higher-level decisions, such as
-meta-analysis, to suggest directions of future research.
+This work was motivated by problems in physics, and thus the usage was
+illustrated on data comparing experiments of hadronic collisions, to
+explore the sensitivity of cluster structure to different principal
+components. These tools can be applied quite broadly, to many
+multivariate data analysis problems.
 
-Future research on the algorithm would include extending it for use in
-3D projections. The addition of another dimension theoretically allows
-for improved perception. In some frameworks, such as the game engine
-\textbf{unity}, this would allow for the exploration in immersive
-virtual reality or mixed reality, which may further allow for a better
-perception of structure and aid in higher-dimensional function
-visualization. Functions with many parameters suffer from the same
-dimensionality problem as data while their possible values lie on a
-surface of values rather than discrete points. Occulation, or the closer
-surface blocking further surfaces, will likely be an issue that may be
-alleviated by the use of wire mesh, changing opacity, or looking at
-sections of the projections known as prosections (Furnas and Buja
-\protect\hyperlink{ref-furnas_prosection_1994}{1994}).
+The manual tour is constrained in the sense that the effect of one
+variable is dependent on the contributions of other variables in the
+manip space. However, this can be useful to simplify a projection, to
+remove variables without affecting the visible structure. Defining a
+manual rotation in high dimensions is possible using Givens rotations
+and Householder reflections as outlined in Buja et al.
+(\protect\hyperlink{ref-buja_computational_2005}{2005}). This would
+provide more flexible manual rotation, but more difficulty for a user,
+because they have the choice (too much choice) of which directions to
+move.
 
-The \pkg{tourr} package provides many other geometric displays with the
-\code{tourr::display\_*()} family. These geometric options could be
-integrated into the \pkg{ggplot2} framework for display on \pkg{plotly}
-and \pkg{gganimate}. Additionally, the \CRANpkg{animation} package Xie
-et al. (\protect\hyperlink{ref-xie_animation:_2018}{2018}) could be
-implemented for another graphics framework. However, \pkg{animation}
-builds from \pkg{base} graphics while \pkg{spinifex} utilizes
-\pkg{ggplot2} graphics, a significant paradigm shift, which may have a
-low return on investment.
+Another future research topic could be to extend the algorithm for use
+on 3D projections. With the current popularity and availability of 3D
+virtual displays, this may benefit the detection and understanding of
+higher dimensional structure, or enable the examination of functions.
 
-The Givens rotations and Householder reflections as outlined in Buja et
-al. (\protect\hyperlink{ref-buja_computational_2005}{2005}) could also
-be added. Currently, Gram-Schmidt is the only form of frame
-interpolation used. In a Givens rotation, the \(x\) and \(y\) components
-(for example \(\theta~= 0,~pi/2\)) of the in-plane rotation are
-calculated separately and would be applied sequentially to produce the
-radial rotation. Householder reflections define reflection axes to
-project points on to the axes to generate rotations.
-
-Having script-only interaction with tours causes a significant barrier
-to entry. To a lesser extent, \pkg{plotly} offers some static
-interactions with the contained object, such as tooltips, brushing, and
-linking without communicating back to the R console. The development of
-a dynamic graphical user interface, perhaps with the use of a
-\CRANpkg{shiny} (Chang et al.
-\protect\hyperlink{ref-chang_shiny:_2018}{2018}) application, would
-mitigate the barrier to entry, allow for more rapid analysis, and offer
-an approachable demo tool. The user could easily switch between
-variables to control, adjust the interpolation step angle, or flag/save
-specific frame basis sets.
+Having a graphical user interface would be useful for making it easier
+and more accessible to a general audience. This is possible to implement
+using \CRANpkg{shiny} (Chang et al.
+\protect\hyperlink{ref-chang_shiny:_2018}{2018}). The primary purposes
+of the interface would be to allow the user to interactively change the
+manip variable easily, and the interpolation step for more or less
+detailed views.
 
@@
 
-\leavevmode\hypertarget{ref-furnas_prosection_1994}{}%
-Furnas, George W., and Andreas Buja. 1994. ``Prosection Views:
-Dimensional Inference Through Sections and Projections.'' \emph{Journal
-of Computational and Graphical Statistics} 3 (4): 323--53.
-\url{https://doi.org/10.2307/1390897}.
-
 \leavevmode\hypertarget{ref-kirkpatrick_optimization_1983}{}%
@@
 
+\leavevmode\hypertarget{ref-swayne_ggobi:_2003}{}%
+Swayne, Deborah F, Duncan Temple Lang, Andreas Buja, and Dianne Cook.
+2003. ``GGobi: Evolving from XGobi into an Extensible Framework for
+Interactive Data Visualization.'' \emph{Computational Statistics \& Data
+Analysis}, Data Visualization, 43 (4): 423--44.
+\url{https://doi.org/10.1016/S0167-9473(02)00286-4}.
+
 \leavevmode\hypertarget{ref-wang_mapping_2018}{}%
@@
 
-\leavevmode\hypertarget{ref-xie_animation:_2018}{}%
-Xie, Yihui, Christian Mueller, Lijia Yu, and Weicheng Zhu. 2018.
-\emph{Animation: A Gallery of Animations in Statistics and Utilities to
-Create Animations}. \url{https://yihui.name/animation}.
-
 \bibliography{spinifex\_paper.bib}

### 2019-05-31 paper/spinifex_paper.Rmd
done with all

@@
 <!-- summary of spinifex  -->
-Tours, which are a dynamic linear projection of numeric multivariate data, play an important role in data visualization; they extend the dimensionality of visuals while data- and parameter-spaces become ever larger.
-This research has modified the algorithm producing manual tours which and has made this functionality available in package \pkg{spinifex}.
-The package adds to \pkg{tourr}, extending the graphics offerings that can be used to display tours.
+Dynamic linear projections of numeric multivariate data, tours, play an important role in data visualization; they extend the dimensionality of visuals to peek into high-dimensional data and parameter spaces.
+This research has taken the manual tour algorithm, specifically the radial rotation, used in GGobi [@swayne_ggobi:_2003] to interactively rotate a variable into or out of a 2D projection, and modified it to create an animation that performs the same task.
+It is most useful for examining the importance of variables, and how the structure in the projection is sensitive or not to specific variables.
+This functionality available in package \CRANpkg{spinifex}.
+The work complements the methods available in the \CRANpkg{tourr} package.
 
 <!-- summary of application  -->
-Radial manual tours were applied to a dataset across different experiments of hadronic collisions.
-The importance of selecting the correct manip var, as demonstrated by comparing tours of varying amounts of structural information.
-Manual tours, by controlling the contribution of the manip var on the projection, enable analysts to explore the sensitivity that variable has on the structure of the projection.
-This information can be used by domain experts to identify which variables are probing which parameter spaces and how sensitive structural features are to different variables.
-This insight can be indispensable for variable inclusion/exclusion, and in higher-level decisions, such as meta-analysis, to suggest directions of future research.
+This work was motivated by problems in physics, and thus the usage was illustrated on data comparing experiments of hadronic collisions, to explore the sensitivity of cluster structure to different principal components.
+These tools can be applied quite broadly, to many multivariate data analysis problems.
+<!--The importance of selecting the correct manip var, as demonstrated by comparing tours of varying amounts of structural information.
+Manual tours, by controlling the contribution of the manip var on the projection, enable analysts to explore the sensitivity that variable has on the structure of the projection.This information can be used by domain experts to identify which variables are probing which parameter spaces and how sensitive structural features are to different variables.
+This insight can be indispensable for variable inclusion/exclusion, and in higher-level decisions, such as meta-analysis, to suggest directions of future research.-->
+
+<!-- Constraints -->
+The manual tour is constrained in the sense that the effect of one variable is dependent on the contributions of other variables in the manip space.
+However, this can be useful to simplify a projection, to remove variables without affecting the visible structure.
+Defining a manual rotation in high dimensions is possible using Givens rotations and Householder reflections as outlined in @buja_computational_2005.
+This would provide more flexible manual rotation, but more difficulty for a user, because they have the choice (too much choice) of which directions to move.
+<!--Currently, Gram-Schmidt is the only form of frame interpolation used.
+In a Givens rotation, the $x$ and $y$ components (for example $\theta~= 0,~pi/2$) of the in-plane rotation are calculated separately and would be applied sequentially to produce the radial rotation.
+Householder reflections define reflection axes to project points on to the axes to generate rotations.-->
 
 <!-- 3D and function vis -->
-Future research on the algorithm would include extending it for use in 3D projections.
-The addition of another dimension theoretically allows for improved perception.
+Another future research topic could be to extend the algorithm for use on 3D projections.
+With the current popularity and availability of 3D virtual displays, this may benefit the detection and understanding of higher dimensional structure, or enable the examination of functions.
+
+<!--The addition of another dimension theoretically allows for improved perception.
 In some frameworks, such as the game engine **unity**, this would allow for the exploration in immersive virtual reality or mixed reality, which may further allow for a better perception of structure and aid in higher-dimensional function visualization.
 Functions with many parameters suffer from the same dimensionality problem as data while their possible values lie on a surface of values rather than discrete points.
-Occulation, or the closer surface blocking further surfaces, will likely be an issue that may be alleviated by the use of wire mesh, changing opacity, or looking at sections of the projections known as prosections [@furnas_prosection_1994].
+Occulation, or the closer surface blocking further surfaces, will likely be an issue that may be alleviated by the use of wire mesh, changing opacity, or looking at sections of the projections known as prosections [@furnas_prosection_1994].-->
 
-<!-- geometric and graphics display -->
+<!-- geometric and graphics display 
 The \pkg{tourr} package provides many other geometric displays with the \code{tourr::display\_*()} family.
@@
 Additionally, the \CRANpkg{animation} package @xie_animation:_2018 could be implemented for another graphics framework.
-However, \pkg{animation} builds from \pkg{base} graphics while \pkg{spinifex} utilizes \pkg{ggplot2} graphics, a significant paradigm shift, which may have  a low return on investment.
+However, \pkg{animation} builds from \pkg{base} graphics while \pkg{spinifex} utilizes \pkg{ggplot2} graphics, a significant paradigm shift, which may have  a low return on investment.-->
 
-<!-- interpolation -->
+<!-- interpolation 
 The Givens rotations and Householder reflections as outlined in @buja_computational_2005 could also be added.
@@
 In a Givens rotation, the $x$ and $y$ components (for example $\theta~= 0,~pi/2$) of the in-plane rotation are calculated separately and would be applied sequentially to produce the radial rotation.
-Householder reflections define reflection axes to project points on to the axes to generate rotations.
+Householder reflections define reflection axes to project points on to the axes to generate rotations.-->
 
 <!-- dynamic interaction/gui -->
-Having script-only interaction with tours causes a significant barrier to entry.
+Having a graphical user interface would be useful for making it easier and more accessible to a general audience.
+This is possible to implement using \CRANpkg{shiny} [@chang_shiny:_2018].
+The primary purposes of the interface would be to allow the user to interactively change the manip variable easily, and the interpolation step for more or less detailed views.
+
+<!--Having script-only interaction with tours causes a significant barrier to entry.
 To a lesser extent, \pkg{plotly} offers some static interactions with the contained object, such as tooltips, brushing, and linking without communicating back to the R console.
 The development of a dynamic graphical user interface, perhaps with the use of a \CRANpkg{shiny} [@chang_shiny:_2018] application, would mitigate the barrier to entry, allow for more rapid analysis, and offer an approachable demo tool.
-The user could easily switch between variables to control, adjust the interpolation step angle, or flag/save specific frame basis sets.
+The user could easily switch between variables to control, adjust the interpolation step angle, or flag/save specific frame basis sets.-->
 

### 2019-06-04 paper/spinifex-paper.tex
initial edits

@@
 
-The algorithm operates entirely on projection bases, and incorporates
-the data only when making the projected data plots.
+The algorithm operates entirely on projection bases and incorporates the
+data only when making the projected data plots.
 
@@
 \ref{fig:step0} shows the initial projection of the data. The left panel
-displays the projection basis (\(\textbf{B}\)), and can be used as a
+displays the projection basis (\(\textbf{B}\)) and can be used as a
 visual guide of the magnitude and direction that each variable
@@
 \(\textbf{Y}_{[n,~2]} ~=~ \textbf{X}_{[n,~p]} \textbf{B}_{[p,~2]}\). The
-colour and shape of points is mapped to the flea species. This plot is
+color and shape of points are mapped to the flea species. This plot is
 made using the \code{view\_basis()} function in \CRANpkg{spinifex},
@@
 these axes, the rotation matrix, \(\textbf{R}\) can be written as
-follows, using Rodrigues' rotation formula (originally pubished in
+follows, using Rodrigues' rotation formula (originally published in
 Rodrigues (\protect\hyperlink{ref-rodrigues_lois_1840}{1840})):
@@
 \item
-  Step towards \(0\), where the manip var is a completely in the
+  Step towards \(0\), where the manip var is completely in the
   projection.
@@
 In each of the steps 3-5, a small step may be added to ensure that the
-end points of \(\phi\) (0, \(\pi/2\)) are reached.
+endpoints of \(\phi\) (0, \(\pi/2\)) are reached.
 
@@
 steps. The projection axes are displayed on the top half, which
-correspond to the projected data in the bottom half. When aede2 is
+corresponds to the projected data in the bottom half. When aede2 is
 removed from the projection, the purple cluster overlaps with the green
@@
 the renderer to \texttt{gganimate} is an option, which will also create
-an html animation, or an animated gif. Each of these formats allows for
+an html animation or an animated gif. Each of these formats allows for
 the animation to be made available on a web site, or directly visible in
@@
 (\protect\hyperlink{ref-buja_computational_2005}{2005}). This would
-provide more flexible manual rotation, but more difficulty for a user,
+provide more flexible manual rotation, but more difficult for a user
 because they have the choice (too much choice) of which directions to
@@
 virtual displays, this may benefit the detection and understanding of
-higher dimensional structure, or enable the examination of functions.
+the higher dimensional structure, or enable the examination of
+functions.
 
@@
 }
-\href{mailto:Nicholas.Spyrison@monash.edu}{\nolinkurl{Nicholas.Spyrison@monash.edu}}
+\href{mailto:nicholas.spyrison@monash.edu}{\nolinkurl{nicholas.spyrison@monash.edu}}
 

### 2019-06-04 paper/spinifex_paper.Rmd
initial edits

@@
     - Faculty of Information Technology
-    email:  Nicholas.Spyrison@monash.edu
+    email:  nicholas.spyrison@monash.edu
   - name: Dianne Cook
@@
 - $\textbf{RM} = \textbf{M} * \textbf{R}$, the result of the manip space after the application of the rotation matrix.
-- $\textbf{P} = \textbf{X} * \textbf{RM}$, the data projection, the data as projected through the rotated manip space.
+- $\textbf{Y} = \textbf{X} * \textbf{RM}$, the data as projected through the rotated manip space.
 
-XXX Clean up the notation, not all of this is needed or useful. The shadow of $e$ in the projection may need some notation. The projected data should be $Y$ not $P$. 
+XXX Clean up the notation, not all of this is needed or useful. The shadow of $e$ in the projection may need some notation. The projected data should be $Y$, not $P$. 
 -->
@@
 <!-- operate on bases -->
-The algorithm operates entirely on projection bases, and incorporates the data only when making the projected data plots.
+The algorithm operates entirely on projection bases and incorporates the data only when making the projected data plots.
 
@@
 Figure \ref{fig:step0} shows the initial projection of the data.
-The left panel displays the projection basis ($\textbf{B}$), and can be used as a visual guide of the magnitude and direction that each variable contributes to the projection.
+The left panel displays the projection basis ($\textbf{B}$) and can be used as a visual guide of the magnitude and direction that each variable contributes to the projection.
 The right panel shows the projected data, $\textbf{Y}_{[n,~2]} ~=~ \textbf{X}_{[n,~p]} \textbf{B}_{[p,~2]}$.
-The colour and shape of points is mapped to the flea species.
+The color and shape of points are mapped to the flea species.
 This plot is made using the  \code{view\_basis()} function in \CRANpkg{spinifex}, which generates a \CRANpkg{ggplot2} [@wickham_ggplot2:_2016] object.
@@
 Out-of-projection rotation, $\phi$, is the rotation around an axis on the $XY$ plane, $\textbf{U}$, orthogonal to $\textbf{e}$.
-Given these axes, the rotation matrix, $\textbf{R}$ can be written as follows, using Rodrigues' rotation formula (originally pubished in @rodrigues_lois_1840):
+Given these axes, the rotation matrix, $\textbf{R}$ can be written as follows, using Rodrigues' rotation formula (originally published in @rodrigues_lois_1840):
 
@@
 To use these for sensitivity analysis, the radial rotation is built into an animation where the manip var is rotated fully into the projection, completely out, and then back to the initial value.
-This involves allowing $\phi$, to vary between 0 and $\pi/2$, call the steps $\phi_i$.
+This involves allowing $\phi$ to vary between 0 and $\pi/2$, call the steps $\phi_i$.
 
@@
 (Note: Using angle increment, rather than a number of steps, to control the movement, is consistent with the tour algorithm as implemented in the \pkg{tourr}).
-3. Step towards $0$, where the manip var is a completely in the projection.
+3. Step towards $0$, where the manip var is completely in the projection.
 4. Step towards $\pi/2$, where the manip variable has no contribution.
@@
 
-In each of the steps 3-5, a small step may be added to ensure that the end points of $\phi$ (0, $\pi/2$) are reached.
+In each of the steps 3-5, a small step may be added to ensure that the endpoints of $\phi$ (0, $\pi/2$) are reached.
 
@@
 Figure \ref{fig:step4} illustrates a manual tour sequence having 15 steps.
-The projection axes are displayed on the top half, which correspond to the projected data in the bottom half.
+The projection axes are displayed on the top half, which corresponds to the projected data in the bottom half.
 When aede2 is removed from the projection, the purple cluster overlaps with the green cluster.
@@
 \noindent This will generate an html animation using plotly.
-Switching the renderer to `gganimate` is an option, which will also create an html animation, or an animated gif.
+Switching the renderer to `gganimate` is an option, which will also create an html animation or an animated gif.
 Each of these formats allows for the animation to be made available on a web site, or directly visible in an html formatted document.
@@
 <!--The importance of selecting the correct manip var, as demonstrated by comparing tours of varying amounts of structural information.
-Manual tours, by controlling the contribution of the manip var on the projection, enable analysts to explore the sensitivity that variable has on the structure of the projection.This information can be used by domain experts to identify which variables are probing which parameter spaces and how sensitive structural features are to different variables.
+Manual tours, by controlling the contribution of the manip var on the projection, enable analysts to explore the sensitivity that variable has on the structure of the projection.
+This information can be used by domain experts to identify which variables are probing which parameter spaces and how sensitive structural features are to different variables.
 This insight can be indispensable for variable inclusion/exclusion, and in higher-level decisions, such as meta-analysis, to suggest directions of future research.-->
@@
 Defining a manual rotation in high dimensions is possible using Givens rotations and Householder reflections as outlined in @buja_computational_2005.
-This would provide more flexible manual rotation, but more difficulty for a user, because they have the choice (too much choice) of which directions to move.
+This would provide more flexible manual rotation, but more difficult for a user because they have the choice (too much choice) of which directions to move.
 <!--Currently, Gram-Schmidt is the only form of frame interpolation used.
@@
 Another future research topic could be to extend the algorithm for use on 3D projections.
-With the current popularity and availability of 3D virtual displays, this may benefit the detection and understanding of higher dimensional structure, or enable the examination of functions.
+With the current popularity and availability of 3D virtual displays, this may benefit the detection and understanding of the higher dimensional structure, or enable the examination of functions.
 

### 2019-06-05 paper/spinifex-paper.tex
formating: \pkg, \code, reference update and house keeping

@@
   3D rotations within the manip space, \(\textbf{M}\).
+\item
+  \(\textbf{Y}\), the resulting projection of the data through the manip
+  space, \(\textbf{M}\), and rotation matrix, \(\textbf{R}\). 
 \end{itemize}
@@
 color and shape of points are mapped to the flea species. This plot is
-made using the \code{view\_basis()} function in \CRANpkg{spinifex},
-which generates a \CRANpkg{ggplot2} (Wickham
+made using the \code{view\_basis()} function in \pkg{spinifex}, which
+generates a \CRANpkg{ggplot2} (Wickham
 \protect\hyperlink{ref-wickham_ggplot2:_2016}{2016}) object.
@@
   e_k &\leftarrow 1 \\ 
-  \textbf{e}^*   &\leftarrow \textbf{e} - \langle \textbf{e}, \textbf{B}_1 \rangle \textbf{B}_1 - \langle \textbf{e}, \textbf{B}_2 \rangle \textbf{B}_2 \\ 
+  \textbf{e}^*_{[p,~1]} &= \textbf{e} - \langle \textbf{e}, \textbf{B}_1 \rangle \textbf{B}_1 - \langle \textbf{e}, \textbf{B}_2 \rangle \textbf{B}_2 \\ 
   \textbf{M}_{[p,~3]} &= (\textbf{B}_1,\textbf{B}_2,\textbf{e}^*)
@@
     \begin{bmatrix}
-      1&0&0 \\ 
-      0&1&0 \\ 
-      0&0&1 \\
+      1 & 0 & 0 \\ 
+      0 & 1 & 0 \\ 
+      0 & 0 & 1 \\
     \end{bmatrix} +
@@
 into the projection, completely out, and then back to the initial value.
-This involves allowing \(\phi\), to vary between 0 and \(\pi/2\), call
-the steps \(\phi_i\).
+This involves allowing \(\phi\) to vary between \(0\) and \(\pi/2\),
+call the steps \(\phi_i\).
 
@@
   \(\phi_1 = \cos^{-1}{\sqrt{B_{k1}^2+B_{k2}^2}}\),
-  \(\theta=\tan^{-1}\frac{B_{k2}}{B_{k1}}\). \(\phi_1\) is the angle
-  between \(\textbf{e}\) and its shadow in \(\textbf{B}\). Check that
-  \(\phi_1\) is between 0 and \(\pi/2\), and if not, transform it into
-  this range, by subtracting or adding \(\pi\). (XXX Nick needs to check
-  this math, and the ways to check and fix if angle produced by the
-  calculation given fall in the right range.)
+  \(\theta = \tan^{-1}\frac{B_{k2}}{B_{k1}}\). \(\phi_1\) is the angle
+  between \(\textbf{e}\) and its shadow on \(\textbf{B}\). 
 \item
@@
 In each of the steps 3-5, a small step may be added to ensure that the
-endpoints of \(\phi\) (0, \(\pi/2\)) are reached.
+endpoints of \(\phi\) (\(0\), \(\pi/2\)) are reached.
 
 \hypertarget{sec:display}{%
-\subsubsection{Step 4) Projecting the data}\label{sec:display}}
+\subsubsection{Step 5) Projecting the data}\label{sec:display}}
 
@@
 \begin{align*}
-  \textbf{Y}_{[n,~3]} &= \textbf{X}_{[n,~p]} \textbf{M}_{[p,~3]} \textbf{R}^{(i)}_{[3,3]}
+  \textbf{Y}^{(i)}_{[n,~3]} &= \textbf{X}_{[n,~p]} \textbf{M}_{[p,~3]} \textbf{R}^{(i)}_{[3,3]}
 \end{align*}
@@
 
-The \CRANpkg{spinifex} is available from CRAN, and can be installed by:
+The \pkg{spinifex} is available from CRAN, and can be installed by:
 
@@
 
-Using the \texttt{flea} data from the \CRANpkg{tourr} package, we will
+Using the \code{flea} data from the \pkg{tourr} package, we will
 illustrate generating a manual tour to explore the sensitivity of the
-cluster structure is to the variable \texttt{aede2}.
+cluster structure is to the variable aede2.
 
@@
 
-The \texttt{play\_manual\_tour} function is a composite function for
+The \code{play\_manual\_tour()} function is a composite function for
 user convenience that does these operations:
@@
 \noindent This will generate an html animation using plotly. Switching
-the renderer to \texttt{gganimate} is an option, which will also create
-an html animation or an animated gif. Each of these formats allows for
-the animation to be made available on a web site, or directly visible in
-an html formatted document.
+the renderer to \code{gganimate} is an option, which will also create an
+html animation or an animated gif. Each of these formats allows for the
+animation to be made available on a web site, or directly visible in an
+html formatted document.
 
@@
 \ref{fig:JetClusterBad}), has no effect on the structure, indicating it
-is not important for distinguishing experiments. Animations for the
-remaining PCs can be viewed at the following links:
+is not important for distinguishing experiments.
+
+Animations for the remaining PCs can be viewed at the following links:
 \href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc1/}{PC1},
@@
 projection is sensitive or not to specific variables. This functionality
-available in package \CRANpkg{spinifex}. The work complements the
-methods available in the \CRANpkg{tourr} package.
+available in package \pkg{spinifex}. The work complements the methods
+available in the \pkg{tourr} package.
 
@@
 This article was created in R (R Core Team
-\protect\hyperlink{ref-r_core_team_r:_2018}{2018}), using
+\protect\hyperlink{ref-r_core_team_r:_2019}{2019}), using
 \CRANpkg{knitr} (Xie \protect\hyperlink{ref-stodden_knitr:_2014}{2014})
@@
 
-\leavevmode\hypertarget{ref-r_core_team_r:_2018}{}%
-R Core Team. 2018. \emph{R: A Language and Environment for Statistical
+\leavevmode\hypertarget{ref-r_core_team_r:_2019}{}%
+R Core Team. 2019. \emph{R: A Language and Environment for Statistical
 Computing}. Vienna, Austria: R Foundation for Statistical Computing.

### 2019-06-05 paper/spinifex_paper.Rmd
formating: \pkg, \code, reference update and house keeping

@@
 It takes the current contribution of the chosen variable, and using rotation brings this variable fully into the projection, completely removes it, and then returns to the original position.
-<!--This is an overview of the steps:
-
-1. Provided with a 2D projection, choose a variable to manipulate. This is called the "manip var".
-2. Create a 3D manip space, where the manip variable has the full contribution.
-3. Generate a rotation sequence which increases the norm of the coefficient to 1 and zeros it.
-
-More detail is provided in the following sections. Function names are provided, but code structure and usage is discussed in section \ref{sec:usage}.
--->
 
@@
 - $\textbf{R}$, the 3D rotation matrix, for performing unconstrained 3D rotations within the manip space, $\textbf{M}$.
+- $\textbf{Y}$, the resulting projection of the data through the manip space, $\textbf{M}$, and rotation matrix, $\textbf{R}$.
 <!--
-- $\textbf{RM} = \textbf{M} * \textbf{R}$, the result of the manip space after the application of the rotation matrix.
-- $\textbf{Y} = \textbf{X} * \textbf{RM}$, the data as projected through the rotated manip space.
-
-XXX Clean up the notation, not all of this is needed or useful. The shadow of $e$ in the projection may need some notation. The projected data should be $Y$, not $P$. 
+XXX Clean up the notation, not all of this is needed or useful. The shadow of $e$ in the projection may need some notation. 
 -->
@@
 The color and shape of points are mapped to the flea species.
-This plot is made using the  \code{view\_basis()} function in \CRANpkg{spinifex}, which generates a \CRANpkg{ggplot2} [@wickham_ggplot2:_2016] object.
+This plot is made using the \code{view\_basis()} function in \pkg{spinifex}, which generates a \CRANpkg{ggplot2} [@wickham_ggplot2:_2016] object.
 
@@
 Generating a sequence of values for the rotation angles produces a path for the rotation of the manip space.
-<!--This defines the (orthonormally-constrained) rotation on the basis coefficients.-->
 
@@
 
-<!-- rotate manip space -->
-<!--Defining the axis of rotation for $\phi$ and the application of Rodrigues' rotation formula results in the rotation matrix, the rotation group SO(3).
-Continue with the rotation of the manip space, post-multiply the space by the above rotation matrix producing **RM**, the rotated manip space.
-This is basis values, the linear combination of variables, for one frame of the animation, repeat this process over a sequence of values for $\phi$ to produce an animation.
--->
-
-<!-- **For each value of ** $\phi_i$ **:** -->
-
-<!-- \begin{align*} -->
-<!--   \textbf{RM}_{[p,~3,~i]} &= \textbf{M}_{[p,~3]} ~*~ \textbf{R}_{[3,~3,~i]}  -->
-<!-- \end{align*} -->
-
 ### Step 4) Creating an animation of the radial rotation
@@
 To use these for sensitivity analysis, the radial rotation is built into an animation where the manip var is rotated fully into the projection, completely out, and then back to the initial value.
-This involves allowing $\phi$ to vary between 0 and $\pi/2$, call the steps $\phi_i$.
-
-<!--Apply the transformation $\phi_i \leftarrow \phi_i-\phi_1$ such that $\phi$ is relative to the projection plane.
-When $\phi$ is 0, $|\textbf{B}_k| = 1$, and when $\phi$ is $\pi/2$, the manip var has no contribution to the projection.-->
+This involves allowing $\phi$ to vary between $0$ and $\pi/2$, call the steps $\phi_i$.
 
@@
 1.
-Set initial value of $\phi_1$ and $\theta$: $\phi_1 = \cos^{-1}{\sqrt{B_{k1}^2+B_{k2}^2}}$, $\theta=\tan^{-1}\frac{B_{k2}}{B_{k1}}$.
-$\phi_1$ is the angle between $\textbf{e}$ and its shadow in $\textbf{B}$.
-Check that  $\phi_1$ is between 0 and $\pi/2$, and if not, transform it into this range, by subtracting or adding $\pi$.
-(XXX Nick needs to check this math, and the ways to check and fix if angle produced by the calculation given fall in the right range.)
+Set initial value of $\phi_1$ and $\theta$: $\phi_1 = \cos^{-1}{\sqrt{B_{k1}^2+B_{k2}^2}}$, $\theta = \tan^{-1}\frac{B_{k2}}{B_{k1}}$.
+$\phi_1$ is the angle between $\textbf{e}$ and its shadow on $\textbf{B}$.
+<!-- phi_1 is actually 0, rebrand phi_1 as phi_offset, phi_alpha? -->
 2.
@@
 
-In each of the steps 3-5, a small step may be added to ensure that the endpoints of $\phi$ (0, $\pi/2$) are reached.
+In each of the steps 3-5, a small step may be added to ensure that the endpoints of $\phi$ ($0$, $\pi/2$) are reached.
 
-### Step 4) Projecting the data {#sec:display}
+### Step 5) Projecting the data {#sec:display}
 
@@
 Tours are typically viewed as an animation.
-The animation of this tour can be viewed online at https://nspyrison.netlify.com/thesis/flea_manualtour_mvar5/.
+The animation of this tour can be viewed online at \url{https://nspyrison.netlify.com/thesis/flea_manualtour_mvar5/}.
 The page may take a moment to load.
@@
 
-<!--
-### Checks that need to be made
-
-The initial value of $\phi$ is determined by the angle between $\textbf{e}$ and its shadow in $\textbf{B}$.
-
-relative to its initial value.
-The transformation $\phi_i$ - $\phi_1$ is useful to think about $\phi$ relative to the basis plane.
-Additionally, the value of $\phi$ may be out of phase by a factor of $\pi$.
-If the rotation does not behave as expected these are the first places to check.
--->
-
 # Package structure and functionality {#sec:usage}
@@
 
-The \CRANpkg{spinifex} is available from CRAN, and can be installed by: 
+The \pkg{spinifex} is available from CRAN, and can be installed by: 
 ```{r installCODEEXAMPLE, eval=F, echo=T}
@@
 
-<!--
-The primary functions that will aid users in performing their own manual tours are broken into the following groupings: tour preparation, tour animation, and utility.
-In the following section describes and demonstrates the usage of the above functions following the order in the algorithm.
-The utility functions are called by the tour animation functions, passed to the argument \code{render\_type}.
-Usage is demonstrated below and in the example documentation.-->
-
 ```{r functionsTable, echo=F}
@@
   c("", "", ""),
-  # c("utility", "col_of", "maps a variable to the point color"),
-  # c("utility", "pch_of", "maps a variable to the point character (shape)"),
   c("specialty", "view_basis", "displays the reference frame of a given basis"),
@@
 
-Using the `flea` data from the \CRANpkg{tourr} package, we will illustrate generating a manual tour to explore the sensitivity of the cluster structure is to the variable `aede2`.
-
-<!--
-We'll start by initializing values including a standardized data set (numeric columns only), a starting basis, a categorical variable for point aesthetics (optional), and a manip var.
-To get bearings on the projection, start by observing the reference axes of the basis with \code{view\_basis()} producing Figure \ref{fig:step0}.-->
+Using the \code{flea} data from the \pkg{tourr} package, we will illustrate generating a manual tour to explore the sensitivity of the cluster structure is to the variable aede2.
 
@@
 
-<!--After becoming familiar with this space, select a manip var, the variable to change the contributions of.
-Use \code{view\_manip\_space()} to view the new space with a dimension orthogonal to the projection plane where the manip var has a full contribution.
-This illustrates how the manip var is manipulated with the addition of the manip space as shown in Figure \ref{fig:step2}.
-
-Now we are ready to perform a manual tour on the selected variable.
-Use \code{play\_manual\_tour()} to perform the algorithm as discussed above, in section \ref{sec:algorithm}.
-This is the animated equivalent of Figure \ref{fig:step3}.
--->
-
-The `play_manual_tour` function is a composite function for user convenience that does these operations:
+The \code{play\_manual\_tour()} function is a composite function for user convenience that does these operations:
 
@@
 \noindent This will generate an html animation using plotly.
-Switching the renderer to `gganimate` is an option, which will also create an html animation or an animated gif.
+Switching the renderer to \code{gganimate} is an option, which will also create an html animation or an animated gif.
 Each of these formats allows for the animation to be made available on a web site, or directly visible in an html formatted document.
-
-<!--
-This concludes the content of the algorithm section, however, lets cover animating paths generated in \pkg{tourr}.
-Animate the previously generated guided tour path via \code{play\_tour\_path()}.
-Utility functions can also be passed as arguments into either of the tour animation function to change the resulting graphics object, set \code{render\_type = render\_gganimate} to view the animation as a GIF.
-
-```{r guidedTourCODEEXAMPLE, eval=F, echo=T}
-play_tour_path(tour_path = f_path,
-               data = f_data,
-               render_type = render_gganimate, 
-               angle = angle_speed
-)
-```
--->
 
@@
 Back-transforming the principal components to the original variables can be done for more detailed interpretation.
-<!--These variables are selected to contrast the information conveyed by different manip variables. Links to dynamic HTML5 animations controlling each of the four variables are also provided.-->
 
@@
 However, removing PC3 from the projection (Figure \ref{fig:JetClusterBad}), has no effect on the structure, indicating it is not important for distinguishing experiments.
-<!--, there is a clear difference in the parameter space spanned by the experiment types ATLAS7new and ATLAS7old.
-Specifically, the variation of ATLAS7new (green) becomes more singular.
-The experiments are probing different parameter space and PC4 is important to demonstrate this.
-Yet, when PC3 is manipulated there is no clear indication that the different experiments probe different parameter space.
-Performing a radial manual tour on PC4 is more insightful than for PC3.-->
+
 Animations for the remaining PCs can be viewed at the following links: [PC1](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc1/), [PC2](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc2/), [PC3](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/), and [PC4](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc4/).
@@
 It is most useful for examining the importance of variables, and how the structure in the projection is sensitive or not to specific variables.
-This functionality available in package \CRANpkg{spinifex}.
-The work complements the methods available in the \CRANpkg{tourr} package.
+This functionality available in package \pkg{spinifex}.
+The work complements the methods available in the \pkg{tourr} package.
 
@@
 These tools can be applied quite broadly, to many multivariate data analysis problems.
-<!--The importance of selecting the correct manip var, as demonstrated by comparing tours of varying amounts of structural information.
-Manual tours, by controlling the contribution of the manip var on the projection, enable analysts to explore the sensitivity that variable has on the structure of the projection.
-This information can be used by domain experts to identify which variables are probing which parameter spaces and how sensitive structural features are to different variables.
-This insight can be indispensable for variable inclusion/exclusion, and in higher-level decisions, such as meta-analysis, to suggest directions of future research.-->
 
@@
 To a lesser extent, \pkg{plotly} offers some static interactions with the contained object, such as tooltips, brushing, and linking without communicating back to the R console.
-The development of a dynamic graphical user interface, perhaps with the use of a \CRANpkg{shiny} [@chang_shiny:_2018] application, would mitigate the barrier to entry, allow for more rapid analysis, and offer an approachable demo tool.
+The development of a dynamic graphical user interface, perhaps with the use of a \pkg{shiny} [@chang_shiny:_2018] application, would mitigate the barrier to entry, allow for more rapid analysis, and offer an approachable demo tool.
 The user could easily switch between variables to control, adjust the interpolation step angle, or flag/save specific frame basis sets.-->
@@
 
-This article was created in R [@r_core_team_r:_2018], using \CRANpkg{knitr} [@stodden_knitr:_2014] and \CRANpkg{rmarkdown} [@xie_r_2018], with code generating the examples inline.
+This article was created in R [@r_core_team_r:_2019], using \CRANpkg{knitr} [@stodden_knitr:_2014] and \CRANpkg{rmarkdown} [@xie_r_2018], with code generating the examples inline.
 The source files for this article be found at [github.com/nspyrison/spinifex_paper/](https://github.com/nspyrison/spinifex_paper/).

### 2019-06-05 paper/spinifex_paper.Rmd
install dev version; angle step size now fixed with small correction rather than rounding. angle args adjusted.

@@
 abstract: >
-  Dynamic low-dimensional linear projections of multivariate data collectively known as **tours** provide an important tool for exploring multivariate data and models.
+  Dynamic low-dimensional linear projections of multivariate data collectively known as *tours* provide an important tool for exploring multivariate data and models.
 The R package \pkg{tourr} provides functions for several types of tours: grand, guided, little, local and frozen.
@@
   if (axes != "off") {
-    gg <- gg + ggplot2::geom_path(
+    gg <- 
+      gg + ggplot2::geom_path(
         data = circ, mapping = ggplot2::aes(x = x, y = y), 
-      color = "grey80", size = 0.3, inherit.aes = F) + 
+        color = "grey80", size = 0.3, inherit.aes = F
+      ) + 
       ggplot2::geom_segment(
-        data = basis, mapping = ggplot2::aes(x = V1, y = V2, xend = zero, yend = zero), color = col) + 
+        data = basis, color = col, 
+        mapping = ggplot2::aes(x = V1, y = V2, xend = zero, yend = zero)
+      ) + 
       ggplot2::geom_text(
         data = basis, mapping = ggplot2::aes(x = V1, y = V2, label = labels), 
-        size = 4, color = col, hjust = 1.1, vjust=.4)
+        size = 4, color = col, hjust = 1.1, vjust=.4
+      )
   }
@@
             as.matrix(basis)) - 0.5)
-        gg <- gg + ggplot2::geom_point(data = proj, mapping = ggplot2::aes(x = V1, 
-            y = V2))
+        gg <- gg + ggplot2::geom_point(data = proj, 
+                                       mapping = ggplot2::aes(x = V1, y = V2))
     }
@@
 ## Steps
+
 ### Step 0) Set up
@@
 "}
-f_angle <- .26
+f_angle <- .29
 f_mtour <- manual_tour(f_bas, manip_var = f_mvar, angle = f_angle)
@@
 p1 <- view_basis2(f_mtour[,,1],  colnames(f_dat), manip_var = f_mvar) + ggtitle("1 phi=0.78")
-p2 <- view_basis2(f_mtour[,,6],  colnames(f_dat), manip_var = f_mvar) + ggtitle("2  phi=0")
+p2 <- view_basis2(f_mtour[,,5],  colnames(f_dat), manip_var = f_mvar) + ggtitle("2  phi=0")
 p3 <- view_basis2(f_mtour[,,12], colnames(f_dat), manip_var = f_mvar) + ggtitle("3 phi=1.57")
@@
 
-<!--### Sequence of $\phi_i$-->
-
+<!-- Sequence of $\phi_i$ -->
 1.
 Set initial value of $\phi_1$ and $\theta$: $\phi_1 = \cos^{-1}{\sqrt{B_{k1}^2+B_{k2}^2}}$, $\theta = \tan^{-1}\frac{B_{k2}}{B_{k1}}$.
-$\phi_1$ is the angle between $\textbf{e}$ and its shadow on $\textbf{B}$.
-<!-- phi_1 is actually 0, rebrand phi_1 as phi_offset, phi_alpha? -->
+$\phi_1$ is the angle between $\textbf{e}$ and its shadow in $\textbf{B}$.
 2.
@@
 
-<!-- plot 2 var in seq for animation -->
+<!-- plot XY in seq for animation -->
 \noindent where $\textbf{R}^{(i)}_{[3,3]}$ is the incremental rotation matrix, using $\phi_i$.
@@
 ## Functions
-<!-- in terms of work flow. -->
 
+<!-- in terms of work flow -->
 Table \ref{tab:functionsTable} lists the primary functions and their purpose.
@@
 jet_mvar  <- 4
-jet_ang   <- .26
+jet_ang   <- .315
 jet_mtour <- manual_tour(basis = jet_bas, manip_var = jet_mvar, angle = jet_ang)
@@
 DIS_mvar  <- 6
-DIS_ang   <- .27
+DIS_ang   <- .32
 DIS_mtour <- manual_tour(basis = DIS_bas, manip_var = DIS_mvar, angle = DIS_ang)

### 2019-06-05 paper/spinifex-paper.tex
install dev version; angle step size now fixed with small correction rather than rounding. angle args adjusted.
hard copy edits.

@@
 Dynamic low-dimensional linear projections of multivariate data
-collectively known as \textbf{tours} provide an important tool for
+collectively known as \emph{tours} provide an important tool for
 exploring multivariate data and models. The R package \pkg{tourr}
@@
   \(\theta = \tan^{-1}\frac{B_{k2}}{B_{k1}}\). \(\phi_1\) is the angle
-  between \(\textbf{e}\) and its shadow on \(\textbf{B}\). 
+  between \(\textbf{e}\) and its shadow in \(\textbf{B}\).
 \item

### 2019-06-05 paper/spinifex_paper.Rmd
hard copy edits.

@@
 - $\textbf{M}$ is a $p \times 3$ matrix, defining the 3D subspace where data rotation occurs, and is called the manip(ulation) space. 
-- $\theta$, the angle of in-projection-plane rotation, for example, on the reference axes; $c_\theta, s_\theta$ are the cosine and sine.
-- $\phi$, the angle of out-of-projection-plane rotation, into the manip space; $c_\phi, s_\phi$ are the cosine and sine. The initial value for animation purposes is $\phi_1$.
+- $\theta$, the angle of in-projection rotation, for example, on the reference axes; $c_\theta, s_\theta$ are the cosine and sine.
+- $\phi$, the angle of out-of-projection rotation, into the manip space; $c_\phi, s_\phi$ are the cosine and sine. The initial value for animation purposes is $\phi_1$.
 - $\textbf{U}$, the axis of rotation for out-of-projection rotation.
@@
 For a radial tour, fix $\theta$, the angle describing rotation within the projection plane, and compute a sequence for $\phi$, defining movement out of the plane.
-This will change the initial value, $\phi_1$, the angle between $\textbf{e}$ and its shadow in $\textbf{B}$, to a maximum of $0$ (manip var fully in projection), then to a minimum of $\pi/2$ (manip var out of projection), before returning to $\phi_1$.
+This will change $\phi$ from the initial value, $\phi_1$, the angle between $\textbf{e}$ and its shadow in $\textbf{B}$, to a maximum of $0$ (manip var fully in projection), then to a minimum of $\pi/2$ (manip var out of projection), before returning to $\phi_1$.
 
@@
 
-<!--The addition of another dimension theoretically allows for improved perception.
-In some frameworks, such as the game engine **unity**, this would allow for the exploration in immersive virtual reality or mixed reality, which may further allow for a better perception of structure and aid in higher-dimensional function visualization.
-Functions with many parameters suffer from the same dimensionality problem as data while their possible values lie on a surface of values rather than discrete points.
-Occulation, or the closer surface blocking further surfaces, will likely be an issue that may be alleviated by the use of wire mesh, changing opacity, or looking at sections of the projections known as prosections [@furnas_prosection_1994].-->
-
 <!-- geometric and graphics display 
@@
 
-<!-- interpolation 
-The Givens rotations and Householder reflections as outlined in @buja_computational_2005 could also be added.
-Currently, Gram-Schmidt is the only form of frame interpolation used.
-In a Givens rotation, the $x$ and $y$ components (for example $\theta~= 0,~pi/2$) of the in-plane rotation are calculated separately and would be applied sequentially to produce the radial rotation.
-Householder reflections define reflection axes to project points on to the axes to generate rotations.-->
-
 <!-- dynamic interaction/gui -->
@@
 The primary purposes of the interface would be to allow the user to interactively change the manip variable easily, and the interpolation step for more or less detailed views.
-
-<!--Having script-only interaction with tours causes a significant barrier to entry.
-To a lesser extent, \pkg{plotly} offers some static interactions with the contained object, such as tooltips, brushing, and linking without communicating back to the R console.
-The development of a dynamic graphical user interface, perhaps with the use of a \pkg{shiny} [@chang_shiny:_2018] application, would mitigate the barrier to entry, allow for more rapid analysis, and offer an approachable demo tool.
-The user could easily switch between variables to control, adjust the interpolation step angle, or flag/save specific frame basis sets.-->
 

### 2019-06-05 paper/spinifex-paper.tex
rest of hard copy edits (+knit)

@@
 \item
-  \(\theta\), the angle of in-projection-plane rotation, for example, on
-  the reference axes; \(c_\theta, s_\theta\) are the cosine and sine.
+  \(\theta\), the angle of in-projection rotation, for example, on the
+  reference axes; \(c_\theta, s_\theta\) are the cosine and sine.
 \item
-  \(\phi\), the angle of out-of-projection-plane rotation, into the
-  manip space; \(c_\phi, s_\phi\) are the cosine and sine. The initial
-  value for animation purposes is \(\phi_1\).
+  \(\phi\), the angle of out-of-projection rotation, into the manip
+  space; \(c_\phi, s_\phi\) are the cosine and sine. The initial value
+  for animation purposes is \(\phi_1\).
 \item
@@
 the projection plane, and compute a sequence for \(\phi\), defining
-movement out of the plane. This will change the initial value,
-\(\phi_1\), the angle between \(\textbf{e}\) and its shadow in
+movement out of the plane. This will change \(\phi\) from the initial
+value, \(\phi_1\), the angle between \(\textbf{e}\) and its shadow in
 \(\textbf{B}\), to a maximum of \(0\) (manip var fully in projection),
@@
 cluster, PCA is recomputed and the first six principal components,
-explaining 48\% of the variation is used. The contributions of PC6 and
-PC2 are explored in Figures \ref{fig:DISclusterGood} and
+explaining 48\% of the full sample variation, is used. The contributions
+of PC6 and PC2 are explored in Figures \ref{fig:DISclusterGood} and
 \ref{fig:DISclusterBad}, respectively. Three experiments are examined:

### 2019-06-05 paper/spinifex_paper.Rmd
rest of hard copy edits (+knit)

@@
 <!--
-XXX Clean up the notation, not all of this is needed or useful. The shadow of $e$ in the projection may need some notation. 
+XXX Clean up the notation, not all of this is needed or useful. 
+- The shadow of $e$ in the projection may need some notation. 
+- R is really an unconstrained rotation? (constrained rotation mentioned twice, what is the differenece?
 -->
@@
 <!-- introduce DIS cluster -->
-Following @cook_dynamical_2018, to explore the DIS cluster, PCA is recomputed and the first six principal components, explaining 48% of the variation is used.
+Following @cook_dynamical_2018, to explore the DIS cluster, PCA is recomputed and the first six principal components, explaining 48% of the full sample variation, is used.
 The contributions of PC6 and PC2 are explored in Figures \ref{fig:DISclusterGood} and  \ref{fig:DISclusterBad}, respectively.

### 2019-06-05 paper/spinifex-paper.tex
grammarly and knit

@@
   \(\textbf{M}\) is a \(p \times 3\) matrix, defining the 3D subspace
-  where data rotation occurs, and is called the manip(ulation) space.
+  where data rotation occurs and is called the manip(ulation) space.
 \item
@@
 
-Wang et al. (\protect\hyperlink{ref-wang_mapping_2018}{2018}) introduce
+Wang et al. (\protect\hyperlink{ref-wang_mapping_2018}{2018}) introduces
 a new tool, PDFSense, to visualize the sensitivity of hadronic

### 2019-06-05 paper/spinifex_paper.Rmd
grammarly and knit

@@
 - $\textbf{e}$, a 1D basis vector of length $p$, with 1 in the $k$-th position and 0 elsewhere.
-- $\textbf{M}$ is a $p \times 3$ matrix, defining the 3D subspace where data rotation occurs, and is called the manip(ulation) space. 
+- $\textbf{M}$ is a $p \times 3$ matrix, defining the 3D subspace where data rotation occurs and is called the manip(ulation) space. 
 - $\theta$, the angle of in-projection rotation, for example, on the reference axes; $c_\theta, s_\theta$ are the cosine and sine.
@@
 - The shadow of $e$ in the projection may need some notation. 
-- R is really an unconstrained rotation? (constrained rotation mentioned twice, what is the differenece?
+- R is really an unconstrained rotation? (constrained rotation mentioned twice, what is the difference?
 -->
@@
 
-<!-- in terms of work flow -->
+<!-- in terms of workflow -->
 Table \ref{tab:functionsTable} lists the primary functions and their purpose.
@@
 <!-- Introduction of data and original paper -->
-@wang_mapping_2018 introduce a new tool, PDFSense, to visualize the sensitivity of hadronic experiments to nucleon structure.
+@wang_mapping_2018 introduces a new tool, PDFSense, to visualize the sensitivity of hadronic experiments to nucleon structure.
 The parameter-space of these experiments lies in 56 dimensions, $\delta \in \mathbb{R}^{56}$, and are visualized as 3D subspaces of the 10 first principal components in linear (PCA) and non-linear (t-SNE) embeddings.

### 2019-06-07 paper/spinifex_paper.Rmd
final edits, figures without boarders

@@
 
-  # BASIS
+  # BASIS AXES
   gg1 <-
@@
                             label = paste0("frame: ",slide))) +
+    # OPTIONS
     theme_void() +
     scale_color_brewer(palette = "Dark2") + 
-    # coord_fixed() +
-    theme(legend.position="none",
-          panel.border = element_rect(colour = "black", fill = NA))
+    theme(legend.position="none")
+          #,panel.border = element_rect(colour = "black", fill = NA))
 
@@
 - $\textbf{M}$ is a $p \times 3$ matrix, defining the 3D subspace where data rotation occurs and is called the manip(ulation) space.
+- $\textbf{R}$, the 3D rotation matrix, for performing unconstrained 3D rotations within the manip space, $\textbf{M}$.
 - $\theta$, the angle of in-projection rotation, for example, on the reference axes; $c_\theta, s_\theta$ are the cosine and sine.
 - $\phi$, the angle of out-of-projection rotation, into the manip space; $c_\phi, s_\phi$ are the cosine and sine. The initial value for animation purposes is $\phi_1$.
-- $\textbf{U}$, the axis of rotation for out-of-projection rotation.
-- $\textbf{R}$, the 3D rotation matrix, for performing unconstrained 3D rotations within the manip space, $\textbf{M}$.
+- $\textbf{U}$, the axis of rotation for out-of-projection rotation orthogonal to $\textbf{e}$.
 - $\textbf{Y}$, the resulting projection of the data through the manip space, $\textbf{M}$, and rotation matrix, $\textbf{R}$.

### 2019-06-10 zRotationMatrix.r
coverletter + cleanup

@@ -1,158 +1,0 @@
-## scratchpad for changign rotation matrix
-# https://en.wikipedia.org/wiki/3D_rotation_group
-# https://en.wikipedia.org/wiki/Rotation_matrix#Basic_rotations
-
-library(spinifex)
-
-set.seed(20190425)
-f_dat  <- tourr::rescale(flea[,1:6])
-f_cat  <- factor(flea$species)
-f_bas  <- basis_random(6,2)
-f_mvar <- 4
-
-view_manip_space(basis = f_bas,manip_var = f_mvar)
-#play_manual_tour(data = f_dat, basis = f_bas, manip_var = f_mvar,
-#                 col = col_of(flea$species), axes = "bottomleft")
-
-load_funcs <- T
-if (load_funcs) # overwrites the spinifex functions: rotate_manip_space and manual_tour
-{
-  #rotate_manip_space
-  rotate_manip_space <- function(manip_space,
-                                    theta=NULL,
-                                    phi=NULL,
-                                    omega=NULL # if null, use Cook's 2 angle rotation matrix.
-  )
-  {
-    # if(is.null(theta)) theta = 1
-    # if(is.null(phi))   phi   = 2
-    s_theta <- sin(theta)
-    c_theta <- cos(theta)
-    s_phi <- sin(phi)
-    c_phi <- cos(phi)
-    if(!is.null(omega))
-      {s_omega <- sin(omega)
-      c_omega <- cos(omega)}
-
-    if(is.null(omega)){ # if omega not passed, use Cook defined (2 angle) rotation.
-      R <- matrix(c(c_theta^2 * c_phi + s_theta^2,
-                    -c_theta * s_theta * (1 - c_phi),
-                    -c_theta * s_phi,  # 3
-                    -c_theta * s_theta * (1 - c_phi),
-                    s_theta^2 * c_phi + c_theta^2,
-                    -s_theta * s_phi, # 6
-                    c_theta * s_phi,
-                    s_theta * s_phi, c_phi), # 9
-                  nrow = 3, ncol = 3,
-                  byrow = TRUE)
-    } else { #if omega is passed, use 3 euler angle rotation
-      r_theta <- matrix(c(1,
-                      0,
-                      0,  # 3
-                      0,
-                      c_theta,
-                      -s_theta, # 6
-                      0,
[…]

### 2019-06-28 paper/spinifex_paper.Rmd
Hard coded holes basis. touched up n_slides!=15 warning. RJ form needs work.

@@
 knitr::opts_chunk$set(
-  fig.align = "center", echo = TRUE, collapse = TRUE,
-  message = FALSE, warning = FALSE, error = FALSE,
-  cache = F, cache.lazy = FALSE
+  fig.align = "center", 
+  echo = TRUE, 
+  collapse = TRUE,
+  message = FALSE, 
+  warning = FALSE, 
+  error = FALSE,
+  cache = FALSE, 
+  cache.lazy = FALSE
 )
@@
   if (n_slides != 15) 
-    stop(paste0("stop: n_slides !=15 !!!!! n_slides = ", n_slides))
+    stop(paste0("stop: n_slides !=15 !!! Check the angle step size. n_slides = ", n_slides))
 
@@
 f_cat  <- factor(flea$species)
-f_path <- save_history(f_dat, guided_tour(holes()))
-f_bas  <- matrix(f_path[,, max(dim(f_path)[3])], ncol=2)
+## hard code a basis. tourr doesn't fix results with set.seed().
+f_path <-  c(.693, -.022, .082, -.119, .706, .023, 
+             -.070, .438, .405, .515, .103, .604)
+f_bas <- matrix(f_path, ncol=2)
+# f_path <- save_history(f_dat, guided_tour(holes()))
+# f_bas  <- matrix(f_path[,, max(dim(f_path)[3])], ncol=2)
 f_mvar <- 5
@@
 if(dim(f_mtour)[3] != 15) message(
-  paste0("!!!!! step3 n_slides != 15 !!!!! Please knit again. n_slides = ", dim(f_mtour)[3]))
+  paste0("!!! n_slides != 15 !!! Check angle step size. n_slides = ", dim(f_mtour)[3], ". angle = ", f_angle))
 p1 <- view_basis2(f_mtour[,,1],  colnames(f_dat), manip_var = f_mvar) + ggtitle("1 phi=0.78")

### 2019-06-28 paper/spyrison-cook.Rmd
/paper/ clean up for RJ resubmission.

--- 2019-06-28 paper/spinifex_paper.Rmd
@@
 toc: false
-bibliography: spinifex_paper.bib
+bibliography: spyrison-cook.bib
 output: rticles::rjournal_article
@@
 \bibliography{spinifex_paper}
-<!-- note that Di/Ursula have different setup and options and use static images.  -->
+<!-- tourrd3 example has different setup and options and uses static images.  -->
 ```{r setup_paper, include=FALSE}
@@
   error = FALSE,
-  cache = FALSE, 
+  cache = TRUE, # needed for log and figure out?
   cache.lazy = FALSE

### 2019-06-28 paper/spyrison-cook.tex
/paper/ clean up for RJ resubmission.

--- 2019-06-05 paper/spinifex-paper.tex
@@
 \item
+  \(\textbf{R}\), the 3D rotation matrix, for performing unconstrained
+  3D rotations within the manip space, \(\textbf{M}\).
+\item
   \(\theta\), the angle of in-projection rotation, for example, on the
@@
 \item
-  \(\textbf{U}\), the axis of rotation for out-of-projection rotation.
-\item
-  \(\textbf{R}\), the 3D rotation matrix, for performing unconstrained
-  3D rotations within the manip space, \(\textbf{M}\).
+  \(\textbf{U}\), the axis of rotation for out-of-projection rotation
+  orthogonal to \(\textbf{e}\).
 \item
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{spinifex_paper_files/figure-latex/step0-1}
+{\centering \includegraphics[width=0.7\linewidth]{spyrison-cook_files/figure-latex/step0-1}
 
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{spinifex_paper_files/figure-latex/step2-1}
+{\centering \includegraphics[width=0.7\linewidth]{spyrison-cook_files/figure-latex/step2-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{spinifex_paper_files/figure-latex/step3-1}
+{\centering \includegraphics[width=1\linewidth]{spyrison-cook_files/figure-latex/step3-1}
 
@@
 
-{\centering \includegraphics[width=5.83in,height=7in]{spinifex_paper_files/figure-latex/step4-1}
+{\centering \includegraphics[width=5.83in,height=7in]{spyrison-cook_files/figure-latex/step4-1}
 
@@
 
-{\centering \includegraphics[width=5.83in,height=7in]{spinifex_paper_files/figure-latex/JetClusterGood-1}
+{\centering \includegraphics[width=5.83in,height=7in]{spyrison-cook_files/figure-latex/JetClusterGood-1}
 
@@
 
-{\centering \includegraphics[width=5.83in,height=7in]{spinifex_paper_files/figure-latex/JetClusterBad-1}
+{\centering \includegraphics[width=5.83in,height=7in]{spyrison-cook_files/figure-latex/JetClusterBad-1}
 
@@
 
-{\centering \includegraphics[width=5.83in,height=7in]{spinifex_paper_files/figure-latex/DISclusterGood-1}
+{\centering \includegraphics[width=5.83in,height=7in]{spyrison-cook_files/figure-latex/DISclusterGood-1}
 
@@
 
-{\centering \includegraphics[width=5.83in,height=7in]{spinifex_paper_files/figure-latex/DISclusterBad-1}
+{\centering \includegraphics[width=5.83in,height=7in]{spyrison-cook_files/figure-latex/DISclusterBad-1}
 
@@
 
-\bibliography{spinifex\_paper.bib}
+\bibliography{spyrison-cook.bib}
 

### 2019-06-28 _RJformat_troubleshooting.r

@@ -1,0 +1,33 @@
+
+tools::texi2pdf(file="paper/RJwrapper.tex",clean = T)
+## Error in texi2dvi(file = file, pdf = TRUE, clean = clean, quiet = quiet,  :
+## unable to run 'pdflatex' on 'paper/spinifex-paper.tex'
+
+Sys.which("pdflatex") # tinytex...pdflatex.exe
+Sys.getenv("PATH") #. large list
+
+### I think this path is not the same use case.
+library(msa)
+#msa::msaPrettyPrint("paper/spinifex-paper.tex", output="tex", showConsensus = "none", askForOverwrite=FALSE, verbose=FALSE,
+#               file = "tex_file")
+?msa::msaPrettyPrint
+
+## read sequences
+filepath <- system.file("examples", "exampleAA.fasta", package="msa")
+mySeqs <- readAAStringSet(filepath)
+
+## call unified interface msa() for default method (ClustalW) and
+## default parameters
+myAlignment <- msa(mySeqs)
+
+## show resulting LaTeX code with default settings
+msaPrettyPrint(myAlignment, output="asis", askForOverwrite=FALSE)
+
+## create PDF file according to some custom settings
+tmpFile <- tempfile(pattern="msa", tmpdir=".", fileext=".pdf")
+tmpFile
+msaPrettyPrint(myAlignment, file=tmpFile, output="pdf",
+               showNames="left", showNumbering="none", showLogo="top",
+               showConsensus="bottom", logoColors="rasmol",
+               verbose=FALSE, askForOverwrite=FALSE)
+

### 2019-11-21 paper/spyrison-cook.Rmd
typos, edits, local function recency.

@@
 Each of these can be viewed dynamically, or saved into a data object for animation.
-This paper describes a new package, \pkg{spinifex}, to provide a manual tour, that allows the coefficient for a single variable to be controlled.
+This paper describes a new package, \pkg{spinifex}, which provides a manual tour of multivariate data where the projection coefficient of a single variable is controlled.
 The variable is rotated fully into the projection, or completely out of the projection.
-The resulting sequence of projections can be displayed using animation, with functions from either the \pkg{plotly} and \pkg{gganimate} packages.
+The resulting sequence of projections can be displayed as an animation, with functions from either the \pkg{plotly} or \pkg{gganimate} packages.
 By varying the coefficient of a single variable, it is possible to explore the sensitivity of structure in the projection to that variable.
-It is particularly useful when used with a projection pursuit guided tour to simplify and understand the solution.
-The use of the manual tour for studying the sensitivity of structure, in a projection, to specific variable contributions, is illustrated using particle physics data.
+This is particularly useful when used with a projection pursuit guided tour to simplify and understand the solution.
+The use of the manual tour is applied particle physics data to illustrate the sensitivity of structure in a projection to specific variable contributions.
 
@@
 # going to facets loses control of output size, simpler to just live with this.
-array2static <- function(.m_tour, .data, .m_var, .cat, .mag =2.2)
-{
+array2static <- function(.m_tour, .data, .m_var, .cat, .mag = 2.2){
+  ##' for @EXAMPLE 
+  if (F){ #Example run:
+    set.seed(20190425); f_dat <- tourr::rescale(flea[,1:6]); f_cat <- factor(flea$species);
+    f_path <- c(.693, -.022, .082, -.119, .706, .023, -.070, .438, .405, .515, .103, .604); 
+    f_bas <- matrix(f_path, ncol=2); f_mvar <- 5; f_angle <- .29;
+    f_mtour <- manual_tour(f_bas, manip_var = f_mvar, angle = f_angle);
+    .m_tour <- f_mtour; .data <- f_dat; .m_var <- f_mvar; .cat <- f_cat; .mag = 2.2;
+  }
+
   slides       <- array2df(array = .m_tour, data = .data)
@@
   basis_grid <- merge(x = basis_slides, y = grid_t, by = "slide", all = TRUE)
+  ## grid_t$x becomes x.y . ie. colname.tbl
   # CROSS JOIN
@@
     # AXES LINE SEGMETNS
-    geom_segment(aes(x = V1 + x, y = V2 + y, xend = x, yend = y),
+    geom_segment(aes(x = x.x + x.y, y = y.x + y.y, xend = x.y, yend = y.y),
                  color = col_v, size = siz_v) +
     # AXES TEXT LABELS
-    geom_text(aes(x = V1 + x, y = V2 + y, label = lab_abbr),
+    geom_text(aes(x = x.x + x.y, y = y.x + y.y, label = lab),
               color = col_v, vjust = "outward", hjust = "outward") +
     # AXES FRAME NUM
-    geom_text(aes(x = x - .7, y = y + 1.1, 
+    geom_text(aes(x = x.y - .7, y = y.y + 1.1, 
                   label = paste0("frame: ",slide)), color = "grey50") +
@@
     geom_point(data = data_grid, size = .7,
-               mapping = aes(x = V1 + x, y = V2 + y, color = cat),
+               mapping = aes(x = x.x + x.y, y = y.x + y.y, color = cat),
                shape = as.integer(cat) + 15) +
@@
     geom_text(data = data_grid, color = "grey50",
-              mapping = aes(x = x - .7, y = y + 1.1, 
+              mapping = aes(x = x.y - .7, y = y.y + 1.1, 
                             label = paste0("frame: ",slide))) +
@@
 Exploring multivariate spaces is a challenging task, increasingly so as dimensionality increases.
-Traditionally, static low-dimensional projections are used to display multivariate data in two dimensions, such as principal component analysis, linear discriminant spaces or projection pursuit.
+Traditionally, static low-dimensional projections are used to display multivariate data in two dimensions including principal component analysis, linear discriminant spaces or projection pursuit.
 These are useful for finding relationships between multiple variables, but they are limited because they show only a glimpse of the high-dimensional space.
-An alternative approach is to use a tour [@asimov_grand_1985] of dynamic linear projections, to look at many different low-dimensional projections.
+An alternative approach is to use a tour [@asimov_grand_1985] of dynamic linear projections to look at many different low-dimensional projections.
 Tours can be considered to extend the dimensionality of visualization, which is important as data and models exist in more than 3D.
 The package \CRANpkg{tourr} [@wickham_tourr_2011] provides a platform for generating tours.
-It has many types of tours available, and many types of displays possible.
+It can produce a variety of tours, each paried with a varitey of possible displays.
 A user can make a grand, guided, little, local or frozen tour, and display the resulting projected data as a scatterplot, density plot, histogram, or even as Chernoff faces if the projection dimension is more than 3.
@@
 The direct optimization of a function allows guided tours to rapidly identify interesting projection features given the relatively large parameter-space.
-After a projection of interest is identified an analyst can then use the "finer brush" of the manual tour, by controlling the contributions of individual variables to explore the sensitivity they have on the structure visible in the projection.
+After a projection of interest is identified, an analyst can then use the "finer brush" of the manual tour to controll the contributions of individual variables to explore the sensitivity they have on the structure visible in the projection.
 
@@
 - $\textbf{R}$, the 3D rotation matrix, for performing unconstrained 3D rotations within the manip space, $\textbf{M}$.
-- $\theta$, the angle of in-projection rotation, for example, on the reference axes; $c_\theta, s_\theta$ are the cosine and sine.
-- $\phi$, the angle of out-of-projection rotation, into the manip space; $c_\phi, s_\phi$ are the cosine and sine. The initial value for animation purposes is $\phi_1$.
+- $\theta$, the angle of in-projection rotation, for example, on the reference axes; $c_\theta, s_\theta$ are its cosine and sine.
+- $\phi$, the angle of out-of-projection rotation, into the manip space; $c_\phi, s_\phi$ are itscosine and sine. The initial value for animation purposes is $\phi_1$.
 - $\textbf{U}$, the axis of rotation for out-of-projection rotation orthogonal to $\textbf{e}$.
@@
 <!-- operate on bases -->
-The algorithm operates entirely on projection bases and incorporates the data only when making the projected data plots.
+The algorithm operates entirely on projection bases and incorporates the data only when making the projected data plots, in light of efficiency.
 
@@
 The flea data (@lubischew_use_1962), available in the \pkg{tourr} package, is used to illustrate the algorithm.
-The data contains 74 observations on 6 variables, which are physical measurements made on beetles.
+The data contains 74 observations on 6 variables, which are physical measurements made on flea beetles.
 Each observation belongs to one of three species.
@@
 A suggested way to start is to identify an interesting projection using a projection pursuit guided tour.
-<!--In a guided tour, the projection sequence is selected by optimizing an index function via hill-climbing on the projection space.--> Here the holes index is used to find a 2D projection of the flea data, which shows three separated groups.
+Here the holes index is used to find a 2D projection of the flea data, which shows three separated species groups.
 Figure \ref{fig:step0} shows the initial projection of the data.
@@
 # Adjust centering to make axes and data side by side
-view_basis(f_bas, labels = colnames(f_dat)) +
+view_basis(f_bas, lab = colnames(f_dat)) +
   geom_point(data = f_proj,
@@
 <!-- select a manip var-->
-In Figure \ref{fig:step0} the contribution of the variables tars1 and aede2 mostly contrast the contribution of the other four variables.
+In figure \ref{fig:step0} the contribution of the variables tars1 and aede2 mostly contrast the contribution of the other four variables.
 These two variables combined contribute in the direction of the projection where the purple cluster is separated from the other two clusters.
 The variable aede2 is selected as the manip var, the variable to be controlled in the tour.
-The question that will be explored, is, how important this variable is to the separation of the clusters.
+The question that will be explored is: how important is this variable to the separation of the clusters in this projection?
 
@@
 <!-- Zero Vect, manip sp -->
-Initialize the coordinate basis vector as a zero vector $\textbf{e}$ of length $p$, and set the $k$-th element to 1.
+Initialize the coordinate basis vector as a zero vector, $\textbf{e}$, of length $p$, and set the $k$-th element to 1.
 In the example data, aede2 is the fifth variable in the data, so $k=5$, set $e_5=1$.
@@
 The manip space provides a 3D projection from $p$-dimensional space, where the coefficient of the manip var can range completely between [0, 1].
-3D rotation can be used to rotate the manip variable completely into or completely out of a 2D projection.
+This 3D space serves as the medium that to rotate the projection basis relative to the selected manipulation variable.
 Figure \ref{fig:step2} illustrates this 3D manip space with the manip var highlighted.
@@
 
-```{r step2, echo=F, out.width='70%',                                           fig.cap = "Illustration of a 3D manip space, constructed to change the contribution of the variable aede2 in the example data.
-The red circle indicates a unit sphere.
-The 2D projection is represented by the blue circle, with the projection coefficients represented by grey lines and text.
-The manip var axis, in red, has length 1 touching the sphere, extends the projection to a third dimension.
-The shadow of this axis (blue) is its contribution in the original 2D projection.
-Illustrated with the \\code{view\\_manip\\_space()} function."}
+```{r step2, echo=F, out.width='70%',                                           fig.cap = "Illustration of a 3D manip space, this space is rotated effectively changing the contribution of the manip variable, aede2 in the example data.
+The blue circle and variable map lies on the projection plane.
+The red circle, orthogonal to the projection plane, illustrates the manipulation space and how the manip var can be controlled and how this effects the variable contribution back onto the projection plane.
+The other variables are omited from the manipulation dimension for simplicity.
+Picturing the other variables in that dimension reveals the intuition that rotating one variable performs a constrained rotation on the others.
+This is illustrated with the \\code{view\\_manip\\_space()} function."}
 view_manip_space2(basis = f_bas, manip_var = f_mvar, labels = colnames(f_dat)) +
@@
 <!-- illustration of axis manip -->
-The basis vector corresponding to the manip var (red line in Figure \ref{fig:step2}), can be operated like a lever.
+The basis vector corresponding to the manip var (red line in Figure \ref{fig:step2}), can be operated like a lever anchored to the origin.
 This is the process of the manual control, that rotates the manip variable into and out of the 2D projection (Figure \ref{fig:step3}).
@@
 Set initial value of $\phi_1$ and $\theta$: $\phi_1 = \cos^{-1}{\sqrt{B_{k1}^2+B_{k2}^2}}$, $\theta = \tan^{-1}\frac{B_{k2}}{B_{k1}}$.
-$\phi_1$ is the angle between $\textbf{e}$ and its shadow in $\textbf{B}$.
+Where $\phi_1$ is the angle between $\textbf{e}$ and its shadow in $\textbf{B}$.
 2.
@@
 (Note: Using angle increment, rather than a number of steps, to control the movement, is consistent with the tour algorithm as implemented in the \pkg{tourr}).
-3. Step towards $0$, where the manip var is completely in the projection.
-4. Step towards $\pi/2$, where the manip variable has no contribution.
+3. Step towards $0$, where the manip var is completely in the projection plane.
+4. Step towards $\pi/2$, where the manip variable has no contribution to the projection.
 5. Step back to $\phi_1$.
 
-In each of the steps 3-5, a small step may be added to ensure that the endpoints of $\phi$ ($0$, $\pi/2$) are reached.
+In each of the steps 3-5, a small step may be added to ensure that the endpoints of $\phi$ ($0$, $\pi/2$, $\phi_1$) are reached.
 
@@
 Show the projected data for each frame in sequence to form an animation.
-<!--The remaining manip space variable can be linked to point aesthetic (such as size or color) to produce depth cues used in conjunction with the $XY$ scatterplot.-->
 
@@
 When aede2 is removed from the projection, the purple cluster overlaps with the green cluster.
-This suggests that aede2 is important for distinguishing this species.
+This suggests that aede2 is important for distinguishing between these species.
 
@@
 The axis for aede2 increases in contribution to the projection, from its initial value to 1, decreasing to 0 and then returning to the initial value.
-This affects the separatiion between the purple and green clusters.
+This effects the separation between the purple and green clusters.
 This shows that aede2 is important for distinguishing the purple species, because the separation disappears when aede2 is not contributing to the projection.
@@
   c("", "", ""),
-  c("render", "render_", "constructs the ggplot object to feed to animation"),
-  c("render", "render_plotly", "converts the ggplot object to a plotly animation"),
-  c("render", "render_gganimate", "converts the ggplot object to a gganimate  animation"),
-  c("render", "array2df", "turns a tour path array into long form, for plotting"),
+  c("render", "array2df", "turn the tour path array into long form, for plotting"),
+  c("render", "render_", "render long form as a ggplot2 objection for animation"),
+  c("render", "render_plotly", "render the animation as a plotly object (default)"),
+  c("render", "render_gganimate", "render the animation as a gganimate object"),
   c("", "", ""),
-  c("animation", "play_tour_path", "animates given tour path"),
-  c("animation", "play_manual_tour", "animates the manual tour algorithm"),
+  c("animation", "play_tour_path", "composite function animating the specified tour path"),
+  c("animation", "play_manual_tour", "composite function animating the specified manual tour"),
   c("", "", ""),
-  c("specialty", "view_basis", "displays the reference frame of a given basis"),
+  c("specialty", "view_basis", "display the reference frame of a given basis"),
   c("specialty", "view_manip_space", "illustrative display of any manip space")
@@
 library(spinifex)
-f_data  <- tourr::rescale(flea[, 1:6])                    ## standardize data
-f_path  <- save_history(f_data, guided_tour(holes()))     ## guided tour using holes index
+f_data  <- tourr::rescale(flea[, 1:6])                    ## standardized flea data
+f_path  <- save_history(f_data, guided_tour(holes()))     ## guided tour path, holes index
 f_basis <- matrix(f_path[,, max(dim(f_path)[3])], ncol=2) ## optimal projection
 f_cat   <- factor(flea$species)                           ## categorical var
-f_mvar  <- 5                                              ## set manip var
-angle_speed <- .26
-play_manual_tour(data = f_data,                           ## Generates animation
-                 basis = f_basis,                         ## using plotly
+f_mvar  <- 5                                              ## set manip var number
+angle_speed <- .26                                        ## anglular dist between slides
+play_manual_tour(data = f_data,                           ## generates animation
+                 basis = f_basis,                         ##   plotly obj (default)
                  manip_var = f_mvar, 
@@
 
-The \code{play\_manual\_tour()} function is a composite function for user convenience that does these operations:
+\noindent The \code{play\_manual\_tour()} function is a composite function for user convenience that does these operations:
 
@@
 
-The function `view_basis` can be used to show a projection, just the basis, or with the data overlaid.
+The function `view_basis` can be used to show a projection of the basis, or with the data overlaid.
 For example, the plots in Figures \ref{fig:step0} and \ref{fig:step3} were made with code similar to this:
@@
 ```{r step2CODEEXAMPLE, eval=F, echo=T}
-
 view_manip_space(basis = f_basis, 
@@
 ```
-
-<!--
-## Rendering and sharing
-
-The \pkg{tourr} package utilizes \pkg{base} graphics for the display of tours.
-\pkg{spinifex} allows tours to be rendered in \pkg{plotly} as an HTML5 object or \pkg{gganimate} as GIF or MP4 files.
-Sharing of animations is not trivial, especially static formats like print and PDF.
-Even with dynamic display capturing the correct resolution and aspect ratio can be challenging, while many formats quickly bloat file sizes limiting sharing options.
-Keep in mind hosting options and exporting functions offered in \pkg{plotly}, \pkg{gganimate} and \pkg{tourr}.
-
-## Storage
-
-Storing each data point for every frame of the animation is redundant.
-Just as operations are performed on the bases, so too should tour paths be stored as bases and a single instance of the data.
-Consider a radial manual tour, we can store the salient features in 3 bases, where $\phi$ is at its starting, minimum, and maximum values.
-The frames in between can be interpolated by supplying angular speed.
-With the use of the \code{tourr::save\_history()} function, the target bases can be saved.
-From there geodesic interpolation can be used to populate the intermittent frames.
-This type of interpolation should not be used on manual tours, which have already been initialized into a 3D manip space where direct linear interpolation is appropriate.
--->
 

### 2019-11-21 paper/spyrison-cook.tex
typos, edits, local function recency.

@@
 a data object for animation. This paper describes a new package,
-\pkg{spinifex}, to provide a manual tour, that allows the coefficient
-for a single variable to be controlled. The variable is rotated fully
-into the projection, or completely out of the projection. The resulting
-sequence of projections can be displayed using animation, with functions
-from either the \pkg{plotly} and \pkg{gganimate} packages. By varying
-the coefficient of a single variable, it is possible to explore the
-sensitivity of structure in the projection to that variable. It is
-particularly useful when used with a projection pursuit guided tour to
-simplify and understand the solution. The use of the manual tour for
-studying the sensitivity of structure, in a projection, to specific
-variable contributions, is illustrated using particle physics data.
+\pkg{spinifex}, which provides a manual tour of multivariate data where
+the projection coefficient of a single variable is controlled. The
+variable is rotated fully into the projection, or completely out of the
+projection. The resulting sequence of projections can be displayed as an
+animation, with functions from either the \pkg{plotly} or
+\pkg{gganimate} packages. By varying the coefficient of a single
+variable, it is possible to explore the sensitivity of structure in the
+projection to that variable. This is particularly useful when used with
+a projection pursuit guided tour to simplify and understand the
+solution. The use of the manual tour is applied particle physics data to
+illustrate the sensitivity of structure in a projection to specific
+variable contributions.
 }
@@
 dimensionality increases. Traditionally, static low-dimensional
-projections are used to display multivariate data in two dimensions,
-such as principal component analysis, linear discriminant spaces or
+projections are used to display multivariate data in two dimensions
+including principal component analysis, linear discriminant spaces or
 projection pursuit. These are useful for finding relationships between
@@
 a tour (Asimov \protect\hyperlink{ref-asimov_grand_1985}{1985}) of
-dynamic linear projections, to look at many different low-dimensional
+dynamic linear projections to look at many different low-dimensional
 projections. Tours can be considered to extend the dimensionality of
@@
 \protect\hyperlink{ref-wickham_tourr_2011}{2011}) provides a platform
-for generating tours. It has many types of tours available, and many
-types of displays possible. A user can make a grand, guided, little,
-local or frozen tour, and display the resulting projected data as a
-scatterplot, density plot, histogram, or even as Chernoff faces if the
-projection dimension is more than 3.
+for generating tours. It can produce a variety of tours, each paried
+with a varitey of possible displays. A user can make a grand, guided,
+little, local or frozen tour, and display the resulting projected data
+as a scatterplot, density plot, histogram, or even as Chernoff faces if
+the projection dimension is more than 3.
 
@@
 interesting projection features given the relatively large
-parameter-space. After a projection of interest is identified an analyst
-can then use the ``finer brush'' of the manual tour, by controlling the
-contributions of individual variables to explore the sensitivity they
-have on the structure visible in the projection.
+parameter-space. After a projection of interest is identified, an
+analyst can then use the ``finer brush'' of the manual tour to controll
+the contributions of individual variables to explore the sensitivity
+they have on the structure visible in the projection.
 
@@
   \(\theta\), the angle of in-projection rotation, for example, on the
-  reference axes; \(c_\theta, s_\theta\) are the cosine and sine.
+  reference axes; \(c_\theta, s_\theta\) are its cosine and sine.
 \item
   \(\phi\), the angle of out-of-projection rotation, into the manip
-  space; \(c_\phi, s_\phi\) are the cosine and sine. The initial value
+  space; \(c_\phi, s_\phi\) are itscosine and sine. The initial value
   for animation purposes is \(\phi_1\).
@@
 The algorithm operates entirely on projection bases and incorporates the
-data only when making the projected data plots.
+data only when making the projected data plots, in light of efficiency.
 
@@
 contains 74 observations on 6 variables, which are physical measurements
-made on beetles. Each observation belongs to one of three species.
+made on flea beetles. Each observation belongs to one of three species.
 
@@
 pursuit guided tour. Here the holes index is used to find a 2D
-projection of the flea data, which shows three separated groups. Figure
-\ref{fig:step0} shows the initial projection of the data. The left panel
-displays the projection basis (\(\textbf{B}\)) and can be used as a
-visual guide of the magnitude and direction that each variable
+projection of the flea data, which shows three separated species groups.
+Figure \ref{fig:step0} shows the initial projection of the data. The
+left panel displays the projection basis (\(\textbf{B}\)) and can be
+used as a visual guide of the magnitude and direction that each variable
 contributes to the projection. The right panel shows the projected data,
@@
 
-In Figure \ref{fig:step0} the contribution of the variables tars1 and
+In figure \ref{fig:step0} the contribution of the variables tars1 and
 aede2 mostly contrast the contribution of the other four variables.
@@
 clusters. The variable aede2 is selected as the manip var, the variable
-to be controlled in the tour. The question that will be explored, is,
-how important this variable is to the separation of the clusters.
+to be controlled in the tour. The question that will be explored is: how
+important is this variable to the separation of the clusters in this
+projection?
 
@@
 
-Initialize the coordinate basis vector as a zero vector \(\textbf{e}\)
+Initialize the coordinate basis vector as a zero vector, \(\textbf{e}\),
 of length \(p\), and set the \(k\)-th element to 1. In the example data,
@@
 where the coefficient of the manip var can range completely between
-{[}0, 1{]}. 3D rotation can be used to rotate the manip variable
-completely into or completely out of a 2D projection. Figure
+{[}0, 1{]}. This 3D space serves as the medium that to rotate the
+projection basis relative to the selected manipulation variable. Figure
 \ref{fig:step2} illustrates this 3D manip space with the manip var
@@
 
-\caption[Illustration of a 3D manip space, constructed to change the contribution of the variable aede2 in the example data]{Illustration of a 3D manip space, constructed to change the contribution of the variable aede2 in the example data.
-The red circle indicates a unit sphere.
-The 2D projection is represented by the blue circle, with the projection coefficients represented by grey lines and text.
-The manip var axis, in red, has length 1 touching the sphere, extends the projection to a third dimension.
-The shadow of this axis (blue) is its contribution in the original 2D projection.
-Illustrated with the \code{view\_manip\_space()} function.}\label{fig:step2}
+\caption[Illustration of a 3D manip space, this space is rotated effectively changing the contribution of the manip variable, aede2 in the example data]{Illustration of a 3D manip space, this space is rotated effectively changing the contribution of the manip variable, aede2 in the example data.
+The blue circle and variable map lies on the projection plane.
+The red circle, orthogonal to the projection plane, illustrates the manipulation space and how the manip var can be controlled and how this effects the variable contribution back onto the projection plane.
+The other variables are omited from the manipulation dimension for simplicity.
+Picturing the other variables in that dimension reveals the intuition that rotating one variable performs a constrained rotation on the others.
+This is illustrated with the \code{view\_manip\_space()} function.}\label{fig:step2}
 \end{figure}
@@
 The basis vector corresponding to the manip var (red line in Figure
-\ref{fig:step2}), can be operated like a lever. This is the process of
-the manual control, that rotates the manip variable into and out of the
-2D projection (Figure \ref{fig:step3}). As the variable contribution is
-controlled, the manip space rotates, and the projection onto the
-horizontal projection plane correspondingly changes. This is a manual
-tour. Generating a sequence of values for the rotation angles produces a
-path for the rotation of the manip space.
+\ref{fig:step2}), can be operated like a lever anchored to the origin.
+This is the process of the manual control, that rotates the manip
+variable into and out of the 2D projection (Figure \ref{fig:step3}). As
+the variable contribution is controlled, the manip space rotates, and
+the projection onto the horizontal projection plane correspondingly
+changes. This is a manual tour. Generating a sequence of values for the
+rotation angles produces a path for the rotation of the manip space.
 
@@
   \(\phi_1 = \cos^{-1}{\sqrt{B_{k1}^2+B_{k2}^2}}\),
-  \(\theta = \tan^{-1}\frac{B_{k2}}{B_{k1}}\). \(\phi_1\) is the angle
-  between \(\textbf{e}\) and its shadow in \(\textbf{B}\).
+  \(\theta = \tan^{-1}\frac{B_{k2}}{B_{k1}}\). Where \(\phi_1\) is the
+  angle between \(\textbf{e}\) and its shadow in \(\textbf{B}\).
 \item
@@
   Step towards \(0\), where the manip var is completely in the
-  projection.
+  projection plane.
 \item
-  Step towards \(\pi/2\), where the manip variable has no contribution.
+  Step towards \(\pi/2\), where the manip variable has no contribution
+  to the projection.
 \item
@@
 In each of the steps 3-5, a small step may be added to ensure that the
-endpoints of \(\phi\) (\(0\), \(\pi/2\)) are reached.
+endpoints of \(\phi\) (\(0\), \(\pi/2\), \(\phi_1\)) are reached.
 
@@
 removed from the projection, the purple cluster overlaps with the green
-cluster. This suggests that aede2 is important for distinguishing this
-species.
+cluster. This suggests that aede2 is important for distinguishing
+between these species.
 
@@
 \begin{Schunk}
-\begin{table}[t]
+\begin{table}
 
@@
 \centering
-\begin{tabular}{lll}
+\begin{tabular}[t]{lll}
 \toprule
@@
  &  & \\
-render & render\_ & constructs the ggplot object to feed to animation\\
-render & render\_plotly & converts the ggplot object to a plotly animation\\
-render & render\_gganimate & converts the ggplot object to a gganimate  animation\\
-render & array2df & turns a tour path array into long form, for plotting\\
+render & array2df & turn the tour path array into long form, for plotting\\
+render & render\_ & render long form as a ggplot2 objection for animation\\
+render & render\_plotly & render the animation as a plotly object (default)\\
+render & render\_gganimate & render the animation as a gganimate object\\
  &  & \\
-animation & play\_tour\_path & animates given tour path\\
-animation & play\_manual\_tour & animates the manual tour algorithm\\
+animation & play\_tour\_path & composite function animating the specified tour path\\
+animation & play\_manual\_tour & composite function animating the specified manual tour\\
  &  & \\
-specialty & view\_basis & displays the reference frame of a given basis\\
+specialty & view\_basis & display the reference frame of a given basis\\
 specialty & view\_manip\_space & illustrative display of any manip space\\
@@
 
-The \code{play\_manual\_tour()} function is a composite function for
-user convenience that does these operations:
+\noindent The \code{play\_manual\_tour()} function is a composite
+function for user convenience that does these operations:
 
@@
 
-The function \texttt{view\_basis} can be used to show a projection, just
+The function \texttt{view\_basis} can be used to show a projection of
 the basis, or with the data overlaid. For example, the plots in Figures
@@
 \begin{Sinput}
-
 view_manip_space(basis = f_basis, 

### 2019-11-21 paper/spyrison-cook.Rmd
Grammerly edits and reviewer response. TODO: application section edits

@@
 The package \CRANpkg{tourr} [@wickham_tourr_2011] provides a platform for generating tours.
-It can produce a variety of tours, each paried with a varitey of possible displays.
+It can produce a variety of tours, each paired with a variety of possible displays.
 A user can make a grand, guided, little, local or frozen tour, and display the resulting projected data as a scatterplot, density plot, histogram, or even as Chernoff faces if the projection dimension is more than 3.
@@
 One way to identify "interesting" features is with the use of a guided tour [@cook_grand_1995].
-Guided tours select a very specific path, which approaches a projection that optimizes an objective-function.
-The optimization is conducted in a manner similar to simulated annealing [@kirkpatrick_optimization_1983].
+Guided tours select a very specific path, which approaches a projection that optimizes an objective function.
+The optimization used to guide the tour is simulated annealing [@kirkpatrick_optimization_1983].
 The direct optimization of a function allows guided tours to rapidly identify interesting projection features given the relatively large parameter-space.
-After a projection of interest is identified, an analyst can then use the "finer brush" of the manual tour to controll the contributions of individual variables to explore the sensitivity they have on the structure visible in the projection.
+After a projection of interest is identified, an analyst can then use the "finer brush" of the manual tour to control the contributions of individual variables to explore the sensitivity they have on the structure visible in the projection.
 
@@
 - $\theta$, the angle of in-projection rotation, for example, on the reference axes; $c_\theta, s_\theta$ are its cosine and sine.
-- $\phi$, the angle of out-of-projection rotation, into the manip space; $c_\phi, s_\phi$ are itscosine and sine. The initial value for animation purposes is $\phi_1$.
+- $\phi$, the angle of out-of-projection rotation, into the manip space; $c_\phi, s_\phi$ are its cosine and sine. The initial value for animation purposes is $\phi_1$.
 - $\textbf{U}$, the axis of rotation for out-of-projection rotation orthogonal to $\textbf{e}$.
 - $\textbf{Y}$, the resulting projection of the data through the manip space, $\textbf{M}$, and rotation matrix, $\textbf{R}$.
-<!--
-XXX Clean up the notation, not all of this is needed or useful. 
-- The shadow of $e$ in the projection may need some notation. 
-- R is really an unconstrained rotation? (constrained rotation mentioned twice, what is the difference?
--->
 
@@
 The manip space provides a 3D projection from $p$-dimensional space, where the coefficient of the manip var can range completely between [0, 1].
-This 3D space serves as the medium that to rotate the projection basis relative to the selected manipulation variable.
+This 3D space serves as the medium to rotate the projection basis relative to the selected manipulation variable.
 Figure \ref{fig:step2} illustrates this 3D manip space with the manip var highlighted.
@@
 The blue circle and variable map lies on the projection plane.
-The red circle, orthogonal to the projection plane, illustrates the manipulation space and how the manip var can be controlled and how this effects the variable contribution back onto the projection plane.
-The other variables are omited from the manipulation dimension for simplicity.
+The red circle, orthogonal to the projection plane, illustrates the manipulation space and how the manip var can be controlled and how this affects the variable contribution back onto the projection plane.
+The other variables are omitted from the manipulation dimension for simplicity.
 Picturing the other variables in that dimension reveals the intuition that rotating one variable performs a constrained rotation on the others.
@@
 @cook_dynamical_2018 illustrates how to learn more about the structures using a grand tour.
-Tours are able to better resolve the shape of clusters, intra-cluster detail, and better outlier detection than PDFSense & TFEP (TensorFlow embedded projections) or traditional static embeddings.
+Tours can better resolve the shape of clusters, intra-cluster detail, and better outlier detection than PDFSense & TFEP (TensorFlow embedded projections) or traditional static embeddings.
 This example builds from here, illustrating how the manual tour can be used to examine the sensitivity of structure in a projection to different parameters.
@@
 The difference in shape can be interpreted as the experiments probing different phase-spaces.
-Back-transforming the principal components to the original variables can be done for more detailed interpretation.
+Back-transforming the principal components to the original variables can be done for a more detailed interpretation.
 
@@
 When PC4 is removed from the projection (Figure \ref{fig:JetClusterGood}), the difference between the two groups is removed, indicating that it is important for distinguishing experiments.
-However, removing PC3 from the projection (Figure \ref{fig:JetClusterBad}), has no effect on the structure, indicating it is not important for distinguishing experiments.
+However, removing PC3 from the projection (Figure \ref{fig:JetClusterBad}), does not affect the structure, indicating it is not important for distinguishing experiments.
 
@@
 <!-- introduce DIS cluster -->
-Following @cook_dynamical_2018, to explore the DIS cluster, PCA is recomputed and the first six principal components, explaining 48% of the full sample variation, is used.
+Following @cook_dynamical_2018, to explore the DIS cluster, PCA is recomputed and the first six principal components, explaining 48% of the full sample variation, are used.
 The contributions of PC6 and PC2 are explored in Figures \ref{fig:DISclusterGood} and  \ref{fig:DISclusterBad}, respectively.

### 2019-11-21 paper/spyrison-cook.tex
Grammerly edits and reviewer response. TODO: application section edits

@@
 
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
+
 \bibliography{spinifex_paper}
@@
 \protect\hyperlink{ref-wickham_tourr_2011}{2011}) provides a platform
-for generating tours. It can produce a variety of tours, each paried
-with a varitey of possible displays. A user can make a grand, guided,
+for generating tours. It can produce a variety of tours, each paired
+with a variety of possible displays. A user can make a grand, guided,
 little, local or frozen tour, and display the resulting projected data
@@
 tours select a very specific path, which approaches a projection that
-optimizes an objective-function. The optimization is conducted in a
-manner similar to simulated annealing (Kirkpatrick, Gelatt, and Vecchi
+optimizes an objective function. The optimization used to guide the tour
+is simulated annealing (Kirkpatrick, Gelatt, and Vecchi
 \protect\hyperlink{ref-kirkpatrick_optimization_1983}{1983}). The direct
@@
 parameter-space. After a projection of interest is identified, an
-analyst can then use the ``finer brush'' of the manual tour to controll
+analyst can then use the ``finer brush'' of the manual tour to control
 the contributions of individual variables to explore the sensitivity
@@
   \(\phi\), the angle of out-of-projection rotation, into the manip
-  space; \(c_\phi, s_\phi\) are itscosine and sine. The initial value
+  space; \(c_\phi, s_\phi\) are its cosine and sine. The initial value
   for animation purposes is \(\phi_1\).
@@
 where the coefficient of the manip var can range completely between
-{[}0, 1{]}. This 3D space serves as the medium that to rotate the
-projection basis relative to the selected manipulation variable. Figure
+{[}0, 1{]}. This 3D space serves as the medium to rotate the projection
+basis relative to the selected manipulation variable. Figure
 \ref{fig:step2} illustrates this 3D manip space with the manip var
@@
 (\protect\hyperlink{ref-cook_dynamical_2018}{2018}) illustrates how to
-learn more about the structures using a grand tour. Tours are able to
-better resolve the shape of clusters, intra-cluster detail, and better
-outlier detection than PDFSense \& TFEP (TensorFlow embedded
-projections) or traditional static embeddings. This example builds from
-here, illustrating how the manual tour can be used to examine the
-sensitivity of structure in a projection to different parameters. The
-specific 2D projections passed to the manual tour were provided from
-Cook, Laa, and Valencia
-(\protect\hyperlink{ref-cook_dynamical_2018}{2018}).
+learn more about the structures using a grand tour. Tours can better
+resolve the shape of clusters, intra-cluster detail, and better outlier
+detection than PDFSense \& TFEP (TensorFlow embedded projections) or
+traditional static embeddings. This example builds from here,
+illustrating how the manual tour can be used to examine the sensitivity
+of structure in a projection to different parameters. The specific 2D
+projections passed to the manual tour were provided from Cook, Laa, and
+Valencia (\protect\hyperlink{ref-cook_dynamical_2018}{2018}).
 
@@
 different phase-spaces. Back-transforming the principal components to
-the original variables can be done for more detailed interpretation.
+the original variables can be done for a more detailed interpretation.
 
@@
 However, removing PC3 from the projection (Figure
-\ref{fig:JetClusterBad}), has no effect on the structure, indicating it
+\ref{fig:JetClusterBad}), does not affect the structure, indicating it
 is not important for distinguishing experiments.
@@
 cluster, PCA is recomputed and the first six principal components,
-explaining 48\% of the full sample variation, is used. The contributions
-of PC6 and PC2 are explored in Figures \ref{fig:DISclusterGood} and
-\ref{fig:DISclusterBad}, respectively. Three experiments are examined:
-DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange).
+explaining 48\% of the full sample variation, are used. The
+contributions of PC6 and PC2 are explored in Figures
+\ref{fig:DISclusterGood} and \ref{fig:DISclusterBad}, respectively.
+Three experiments are examined: DIS HERA1+2 (green), dimuon SIDIS
+(purple), and charm SIDIS (orange).
 

### 2019-11-22 paper/spyrison-cook.Rmd
CHECK POINT before returning accepted paper with edits.

@@
 This example builds from here, illustrating how the manual tour can be used to examine the sensitivity of structure in a projection to different parameters.
-The specific 2D projections passed to the manual tour were provided from @cook_dynamical_2018.
+The specific 2D projections passed to the manual tour were provided in their work.
 
@@
 The data has a hierarchical structure with top-level clusters; DIS, VBP, and jet.
-Each cluster is a particular class of experiments, each with many experimental datasets which, in turn, have many observations.
+Each cluster is a particular class of experiments, each with many experimental datasets which,  each have many observations of their own.
 In consideration of data density, we conduct manual tours on subsets of the DIS and jet clusters.
@@
 The jet cluster resides in a smaller dimensionality than the full set of experiments with four principal components explaining 95% of the variation in the cluster [@cook_dynamical_2018].
-The data within this 4D embedding is further subsetted, to ATLAS7old and ATLAS7new, in order to focus on two groups, with a reasonable number of observations, that occupy different parts of the subspace.
+The data within this 4D embedding is further subsetted, to ATLAS7old and ATLAS7new, to focus on two groups that occupy different parts of the subspace.
 Radial manual tours controlling contributions from PC4 and PC3 are shown in Figures \ref{fig:JetClusterGood} and  \ref{fig:JetClusterBad}, respectively.
@@
 <!-- discussion of findings and which is more insightful -->
-When PC4 is removed from the projection (Figure \ref{fig:JetClusterGood}), the difference between the two groups is removed, indicating that it is important for distinguishing experiments.
-However, removing PC3 from the projection (Figure \ref{fig:JetClusterBad}), does not affect the structure, indicating it is not important for distinguishing experiments.
-
+When PC4 is removed from the projection (Figure \ref{fig:JetClusterGood}) the difference between the two groups is removed, indicating that it is important for distinguishing experiments.
+However, removing PC3 from the projection (Figure \ref{fig:JetClusterBad}) does not affect the structure, indicating it is not important for distinguishing experiments.
 Animations for the remaining PCs can be viewed at the following links: [PC1](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc1/), [PC2](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc2/), [PC3](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/), and [PC4](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc4/).
@@
 
-```{r JetClusterGood, echo=F, warning=F, fig.height=7.0, fig.width=5.83, out.height='7in', out.width='5.83in',                                     fig.cap="Snapshots of a radial manual tour of PC4 focusing on the jet cluster, with color indicating experiment type: ATLAS7new (green) and ATLAS7old (orange).
+```{r JetClusterGood, echo=F, warning=F, fig.height=7.0, fig.width=5.83, out.height='7in', out.width='5.83in',                                     fig.cap="Snapshots of a radial manual tour of PC4 within the jet cluster, with color indicating experiment type: ATLAS7new (green) and ATLAS7old (orange).
 When PC4 is removed from the projection (frame 10) there is little difference between the groups, suggesting that PC4 is important for distinguishing the experiments.
@@
 
-```{r JetClusterBad, echo=F, warning=F, fig.height=7.0, fig.width=5.83, out.height='7in', out.width='5.83in',                                       fig.cap = "Snapshots of a radial manual tour of PC3 focusing on the jet cluster, with color indicating experiment type: ATLAS7new (green) and ATLAS7old (orange).
-When the contribution of PC3 is changed there is little change to the structure of the two groups, suggesting that PC3 is not important for distinguishing the experiments.
+```{r JetClusterBad, echo=F, warning=F, fig.height=7.0, fig.width=5.83, out.height='7in', out.width='5.83in',                                       fig.cap = "Snapshots of a radial manual tour of PC3 within the jet cluster, with color indicating experiment type: ATLAS7new (green) and ATLAS7old (orange).
+When the contribution from PC3 is changed there is little change to the structure of the two groups, suggesting that PC3 is not important for distinguishing the experiments.
 The animation can be viewed at https://nspyrison.netlify.com/thesis/jetcluster\\_manualtour\\_pc3/."}
@@
 
-```{r DISclusterGood, echo=F, warning=F, fig.height=7.0, fig.width=5.83, out.height='7in', out.width='5.83in',                                            fig.cap = "Snapshots of a radial manual tour exploring the contribution of PC6 for the DIS cluster, with color indicating experiment type: DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange).
+```{r DISclusterGood, echo=F, warning=F, fig.height=7.0, fig.width=5.83, out.height='7in', out.width='5.83in',                                            fig.cap = "Snapshots of a radial manual tour exploring the sensitivity PC6 has on the structure of the DIS cluster, with color indicating experiment type: DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange).
 DIS HERA1+2 is distributed in a cross-shaped plane, charm SIDIS occupies the center of this cross, and dimuon SIDIS is a linear cluster crossing DIS HERA1+2.
-As PC6's contribution is increased, DIS HERA1+2 becomes almost singular in one direction (frame 5), indicating that this experiment has very little variability in the direction of PC6.
+As the contribution of PC6 is increased, DIS HERA1+2 becomes almost singular in one direction (frame 5), indicating that this experiment has very little variability in the direction of PC6.
 The animation can be viewed at https://nspyrison.netlify.com/thesis/discluster\\_manualtour\\_pc6/."}
@@
 
-```{r DISclusterBad, echo=F, warning=F, fig.height=7.0, fig.width=5.83, out.height='7in', out.width='5.83in',                                            fig.cap = "Snapshots of a radial manual tour exploring the contribution of PC2 for the DIS cluster, with color indicating experiment type: DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange).
-As PC2's contribution is decreased, dimuon SIDIS becomes more distinguishable from the other two clusters (frames 10-14), indicating that in its absence PC2 is important.
+```{r DISclusterBad, echo=F, warning=F, fig.height=7.0, fig.width=5.83, out.height='7in', out.width='5.83in',                                            fig.cap = "Snapshots of a radial manual tour exploring the sensitivity PC2 to the structure of the DIS cluster, with color indicating experiment type: DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange).
+As contribution from PC2 is decreased, dimuon SIDIS becomes more distinguishable from the other two clusters (frames 10-14), indicating that in its absence PC2 is important.
 The animation can be viewed at https://nspyrison.netlify.com/thesis/discluster\\_manualtour\\_pc2/."}
@@
 This work was motivated by problems in physics, and thus the usage was illustrated on data comparing experiments of hadronic collisions, to explore the sensitivity of cluster structure to different principal components.
-These tools can be applied quite broadly, to many multivariate data analysis problems.
+These tools can be applied quite broadly to many multivariate data analysis problems.
 
@@
 The manual tour is constrained in the sense that the effect of one variable is dependent on the contributions of other variables in the manip space.
-However, this can be useful to simplify a projection, to remove variables without affecting the visible structure.
+However, this can be useful to simplify a projection by removing variables without affecting the visible structure.
 Defining a manual rotation in high dimensions is possible using Givens rotations and Householder reflections as outlined in @buja_computational_2005.
 This would provide more flexible manual rotation, but more difficult for a user because they have the choice (too much choice) of which directions to move.
-<!--Currently, Gram-Schmidt is the only form of frame interpolation used.
-In a Givens rotation, the $x$ and $y$ components (for example $\theta~= 0,~pi/2$) of the in-plane rotation are calculated separately and would be applied sequentially to produce the radial rotation.
-Householder reflections define reflection axes to project points on to the axes to generate rotations.-->
 
@@
 
-This article was created in R [@r_core_team_r:_2019], using \CRANpkg{knitr} [@stodden_knitr:_2014] and \CRANpkg{rmarkdown} [@xie_r_2018], with code generating the examples inline.
+<!-- removed [@r_core_team_r:_2019], as per journal reqs -->
+This article was created in R, using \CRANpkg{knitr} [@stodden_knitr:_2014] and \CRANpkg{rmarkdown} [@xie_r_2018], with code generating the examples inline.
 The source files for this article be found at [github.com/nspyrison/spinifex_paper/](https://github.com/nspyrison/spinifex_paper/).

### 2019-11-22 paper/spyrison-cook.tex
CHECK POINT before returning accepted paper with edits.

@@
 of structure in a projection to different parameters. The specific 2D
-projections passed to the manual tour were provided from Cook, Laa, and
-Valencia (\protect\hyperlink{ref-cook_dynamical_2018}{2018}).
+projections passed to the manual tour were provided in their work.
 
@@
 and jet. Each cluster is a particular class of experiments, each with
-many experimental datasets which, in turn, have many observations. In
-consideration of data density, we conduct manual tours on subsets of the
-DIS and jet clusters. This explores the sensitivity of the structure to
-each of the variables in turn and we present the subjectively best and
-worst variable to manipulate for identifying dimensionality of the
-clusters and describing the span of the clusters.
+many experimental datasets which, each have many observations of their
+own. In consideration of data density, we conduct manual tours on
+subsets of the DIS and jet clusters. This explores the sensitivity of
+the structure to each of the variables in turn and we present the
+subjectively best and worst variable to manipulate for identifying
+dimensionality of the clusters and describing the span of the clusters.
 
@@
 \protect\hyperlink{ref-cook_dynamical_2018}{2018}). The data within this
-4D embedding is further subsetted, to ATLAS7old and ATLAS7new, in order
-to focus on two groups, with a reasonable number of observations, that
-occupy different parts of the subspace. Radial manual tours controlling
-contributions from PC4 and PC3 are shown in Figures
+4D embedding is further subsetted, to ATLAS7old and ATLAS7new, to focus
+on two groups that occupy different parts of the subspace. Radial manual
+tours controlling contributions from PC4 and PC3 are shown in Figures
 \ref{fig:JetClusterGood} and \ref{fig:JetClusterBad}, respectively. The
@@
 When PC4 is removed from the projection (Figure
-\ref{fig:JetClusterGood}), the difference between the two groups is
+\ref{fig:JetClusterGood}) the difference between the two groups is
 removed, indicating that it is important for distinguishing experiments.
 However, removing PC3 from the projection (Figure
-\ref{fig:JetClusterBad}), does not affect the structure, indicating it
-is not important for distinguishing experiments.
-
-Animations for the remaining PCs can be viewed at the following links:
+\ref{fig:JetClusterBad}) does not affect the structure, indicating it is
+not important for distinguishing experiments. Animations for the
+remaining PCs can be viewed at the following links:
 \href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc1/}{PC1},
@@
 
-\caption[Snapshots of a radial manual tour of PC4 focusing on the jet cluster, with color indicating experiment type]{Snapshots of a radial manual tour of PC4 focusing on the jet cluster, with color indicating experiment type: ATLAS7new (green) and ATLAS7old (orange).
+\caption[Snapshots of a radial manual tour of PC4 within the jet cluster, with color indicating experiment type]{Snapshots of a radial manual tour of PC4 within the jet cluster, with color indicating experiment type: ATLAS7new (green) and ATLAS7old (orange).
 When PC4 is removed from the projection (frame 10) there is little difference between the groups, suggesting that PC4 is important for distinguishing the experiments.
@@
 
-\caption[Snapshots of a radial manual tour of PC3 focusing on the jet cluster, with color indicating experiment type]{Snapshots of a radial manual tour of PC3 focusing on the jet cluster, with color indicating experiment type: ATLAS7new (green) and ATLAS7old (orange).
-When the contribution of PC3 is changed there is little change to the structure of the two groups, suggesting that PC3 is not important for distinguishing the experiments.
+\caption[Snapshots of a radial manual tour of PC3 within the jet cluster, with color indicating experiment type]{Snapshots of a radial manual tour of PC3 within the jet cluster, with color indicating experiment type: ATLAS7new (green) and ATLAS7old (orange).
+When the contribution from PC3 is changed there is little change to the structure of the two groups, suggesting that PC3 is not important for distinguishing the experiments.
 The animation can be viewed at https://nspyrison.netlify.com/thesis/jetcluster\_manualtour\_pc3/.}\label{fig:JetClusterBad}
@@
 
-\caption[Snapshots of a radial manual tour exploring the contribution of PC6 for the DIS cluster, with color indicating experiment type]{Snapshots of a radial manual tour exploring the contribution of PC6 for the DIS cluster, with color indicating experiment type: DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange).
+\caption[Snapshots of a radial manual tour exploring the sensitivity PC6 has on the structure of the DIS cluster, with color indicating experiment type]{Snapshots of a radial manual tour exploring the sensitivity PC6 has on the structure of the DIS cluster, with color indicating experiment type: DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange).
 DIS HERA1+2 is distributed in a cross-shaped plane, charm SIDIS occupies the center of this cross, and dimuon SIDIS is a linear cluster crossing DIS HERA1+2.
-As PC6's contribution is increased, DIS HERA1+2 becomes almost singular in one direction (frame 5), indicating that this experiment has very little variability in the direction of PC6.
+As the contribution of PC6 is increased, DIS HERA1+2 becomes almost singular in one direction (frame 5), indicating that this experiment has very little variability in the direction of PC6.
 The animation can be viewed at https://nspyrison.netlify.com/thesis/discluster\_manualtour\_pc6/.}\label{fig:DISclusterGood}
@@
 
-\caption[Snapshots of a radial manual tour exploring the contribution of PC2 for the DIS cluster, with color indicating experiment type]{Snapshots of a radial manual tour exploring the contribution of PC2 for the DIS cluster, with color indicating experiment type: DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange).
-As PC2's contribution is decreased, dimuon SIDIS becomes more distinguishable from the other two clusters (frames 10-14), indicating that in its absence PC2 is important.
+\caption[Snapshots of a radial manual tour exploring the sensitivity PC2 to the structure of the DIS cluster, with color indicating experiment type]{Snapshots of a radial manual tour exploring the sensitivity PC2 to the structure of the DIS cluster, with color indicating experiment type: DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange).
+As contribution from PC2 is decreased, dimuon SIDIS becomes more distinguishable from the other two clusters (frames 10-14), indicating that in its absence PC2 is important.
 The animation can be viewed at https://nspyrison.netlify.com/thesis/discluster\_manualtour\_pc2/.}\label{fig:DISclusterBad}
@@
 explore the sensitivity of cluster structure to different principal
-components. These tools can be applied quite broadly, to many
+components. These tools can be applied quite broadly to many
 multivariate data analysis problems.
@@
 variable is dependent on the contributions of other variables in the
-manip space. However, this can be useful to simplify a projection, to
-remove variables without affecting the visible structure. Defining a
+manip space. However, this can be useful to simplify a projection by
+removing variables without affecting the visible structure. Defining a
 manual rotation in high dimensions is possible using Givens rotations
@@
 
-This article was created in R (R Core Team
-\protect\hyperlink{ref-r_core_team_r:_2019}{2019}), using
-\CRANpkg{knitr} (Xie \protect\hyperlink{ref-stodden_knitr:_2014}{2014})
-and \CRANpkg{rmarkdown} (Xie, Allaire, and Grolemund
+This article was created in R, using \CRANpkg{knitr} (Xie
+\protect\hyperlink{ref-stodden_knitr:_2014}{2014}) and
+\CRANpkg{rmarkdown} (Xie, Allaire, and Grolemund
 \protect\hyperlink{ref-xie_r_2018}{2018}), with code generating the
@@
 \url{http://github.com/thomasp85/gganimate}.
-
-\leavevmode\hypertarget{ref-r_core_team_r:_2019}{}%
-R Core Team. 2019. \emph{R: A Language and Environment for Statistical
-Computing}. Vienna, Austria: R Foundation for Statistical Computing.
-\url{https://www.R-project.org/}.
 

### 2020-08-24 paper/response_Rjournal_reviewer1.txt
post proofing changes

> Overall recommendation: accept, giving the authors the opportunity for final wordsmithing for clarity.

The paper has been has been revised including: simplifying sentance structure, minor changes for clarity, typos corrected. Some details comment in line below.

> I like the organization of the paper.  It seems like about the right amount of detail in terms of the background, discussion of the algorithm (with nice supporting figures), the basic package description, and examples.  The examples, in particular, are important and difficult for a paper of this type; I appreciate the care that went into the presentation.  The animations on the web were also appreciated.

Thank you, glad to hear the work is appreciated.

> I like the Github repository, and glanced through the package source.  It does conform to recommended package design/standards, including roxygen, etc…

Thanks for the brief review.

> I think the paper is generally well-written.  I found myself occasionally distracted by sentence structure. Some sentences are complex with qualifiers and additional phrases added with lots of commas (,).  As an example from the abstract, “This paper describes a new package, spinifex, to provide a manual tour, that allows the coefficient for a single variable to be controlled.”

> I can’t really argue it’s wrong, but it caused me to stop reading and think. I’d argue such a distraction doesn’t help.  Something like, “This paper describes a new package, spinifex, which provides manual tours of multivariate data by allowing the coefficient for a single variable to be controlled.” I can’t argue it’s a big deal and won’t argue that my proposal works or is that much better.  But as a reader, I’m interested in the subject matter and find it frustrating if I have to spend effort parsing a sentence.  Basically, any sentence with more than 3-4 commas could be studied.

> Here’s another: it really feels awkward to me.  I’d be inclined to use a comma after “identified” and do something else or eliminate the comma after “tour”.      “After a projection of interest is identified an analyst can then use the “finer brush” of the manual tour, by controlling the contributions of individual variables to explore the sensitivity they have on the structure visible in the projection. “

Ageed, these instances have been fixed and the whole paper has been reviewed with an eye out for sentance structure and over-use of commas.

> These are just examples.  Not a big deal, and frankly this is FAR better-written and presented than most papers I review: thank you!  Any further tightening before publication would help, but the subject and scholarship seems appropriate for publication in the R Journal.

Thank you for reviewing and offering the suggestions with examples. The feel as this round of edits did improve clarity, and simplfy sentances and clean up a few typos.

> Note a typo on the Github instructions: library(“spinifiex”) of course won’t work.

Corrected along with a few other typos.

### 2020-08-24 paper/spyrison-cook.Rmd
post proofing changes

@@
 The use of the manual tour is applied particle physics data to illustrate the sensitivity of structure in a projection to specific variable contributions.
-
 keywords: exploratory data analysis, projection pursuit, high dimensional data, data visualization, statistical graphics, data science
@@
 output: rticles::rjournal_article
-  #bookdown::pdf_book:
-   # base_format: rticles::rjournal_article #see sandbox for hard code solution.
+  #bookdown::pdf_book ## alternative
 ---
-\bibliography{spinifex_paper}
-<!-- tourrd3 example has different setup and options and uses static images.  -->
 ```{r setup_paper, include=FALSE}
@@
 )
-library(spinifex)
+if (F){ ## From reproducing, please use this dev ver:
+  remotes::install_github("nspyrison/spinifex", 
+                          ref = "da75976588dc008e21f01ce6a57d9778d0ed6933")
+}
+library(tourr)
+library(spinifex) ## Use dev version listed above!
 library(ggplot2)
@@
 ## FUNCTION FOR STATIC OUTPPUT, 
-# going to facets loses control of output size, simpler to just live with this.
-array2static <- function(.m_tour, .data, .m_var, .cat, .mag = 2.2){
-  ##' for @EXAMPLE 
-  if (F){ #Example run:
-    set.seed(20190425); f_dat <- tourr::rescale(flea[,1:6]); f_cat <- factor(flea$species);
-    f_path <- c(.693, -.022, .082, -.119, .706, .023, -.070, .438, .405, .515, .103, .604); 
-    f_bas <- matrix(f_path, ncol=2); f_mvar <- 5; f_angle <- .29;
-    f_mtour <- manual_tour(f_bas, manip_var = f_mvar, angle = f_angle);
-    .m_tour <- f_mtour; .data <- f_dat; .m_var <- f_mvar; .cat <- f_cat; .mag = 2.2;
-  }
+# going to facets loses control of multiple geoms and output size, simpler to just live with this.
+array2ggfacets <- function(tour_array, data, m_var, class, margin = 2.2){
+  n_frames <- dim(tour_array)[3]
+  if (n_frames != 15) 
+    stop(paste0("stop: n_frames != 15!!! Check the angle step size. n_frames = ", n_frames))
 
-  slides       <- array2df(array = .m_tour, data = .data)
-  basis_slides <- slides$basis_slides
-  data_slides  <- slides$data_slides
-  n_slides     <- max(basis_slides$slide)
-  if (n_slides != 15) 
-    stop(paste0("stop: n_slides !=15 !!! Check the angle step size. n_slides = ", n_slides))
+  ## Initialize
+  frames       <- array2df(array = tour_array, data = data)
+  basis_frames <- frames$basis_frames
+  data_frames  <- frames$data_frames
+  p            <- nrow(basis_frames) / n_frames
 
-  # Initialize
   ## manip var asethetics
-  p             <- nrow(basis_slides) / n_slides
   col_v        <- rep("grey80", p)
-  col_v[.m_var] <- "blue"
-  col_v         <- rep(col_v, n_slides)
+  col_v[m_var] <- "blue"
+  col_v        <- rep(col_v, n_frames)
   siz_v        <- rep(0.3, p)
-  siz_v[.m_var] <- 1
-  siz_v         <- rep(siz_v, n_slides)
-  cat           <- rep(as.factor(.cat), n_slides)
+  siz_v[m_var] <- 1
+  siz_v        <- rep(siz_v, n_frames)
+  cat          <- rep(as.factor(class), n_frames)
 
@@
   circ[nrow(circ)+1, ] <- NA
-  ## data asethetics
-  data_slides <- data.frame(data_slides, cat = rep(.cat, n_slides))
+  ## Data asethetics
+  data_frames <- data.frame(data_frames, class = rep(class, n_frames))
+  grid_b <- grid_t <- data.frame(
+    frame = 1:n_frames, x = margin * rep(1:5, 3), y = margin * rep(3:1, each = 5))
+  grid_t$y <- grid_t$y + max(grid_t$y)
+  ## OUTER JOIN
+  basis_grid <- merge(x = basis_frames, y = grid_t, by = "frame", all = TRUE)
+  ## CROSS JOIN
+  circ_grid  <- merge(x = circ, y = grid_t, by = NULL)
+  ## OUTER JOIN
+  data_grid  <- merge(x = data_frames, y = grid_b, by = "frame", all = TRUE)
 
-  grid_b <- grid_t <-
-    data.frame(slide = 1:n_slides, 
-               x = .mag*rep(1:5, 3), y = .mag*rep(3:1, each = 5))
-  grid_t$y <- grid_t$y + max(grid_t$y)
-  # OUTER JOIN
-  basis_grid <- merge(x = basis_slides, y = grid_t, by = "slide", all = TRUE)
-  ## grid_t$x becomes x.y . ie. colname.tbl
-  # CROSS JOIN
-  circ_grid  <- merge(x = circ, y = grid_t, by = NULL)
-  # OUTER JOIN
-  data_grid  <- merge(x = data_slides, y = grid_b, by = "slide", all = TRUE)
-
-  # BASIS AXES
-  gg1 <-
-    ggplot(data = basis_grid) +
-    # AXES LINE SEGMETNS
+  ##### RENDER
+  ## SETUP
+  gg <- ggplot(data = basis_grid) +
+    ## AXES LINE SEGMETNS
     geom_segment(aes(x = x.x + x.y, y = y.x + y.y, xend = x.y, yend = y.y),
                  color = col_v, size = siz_v) +
-    # AXES TEXT LABELS
+    ## AXES TEXT LABELS
     geom_text(aes(x = x.x + x.y, y = y.x + y.y, label = lab),
               color = col_v, vjust = "outward", hjust = "outward") +
-    # AXES FRAME NUM
+    ## AXES FRAME NUM
     geom_text(aes(x = x.y - .7, y = y.y + 1.1, 
-                  label = paste0("frame: ",slide)), color = "grey50") +
-    # AXES CIRCLE PATH
+                  label = paste0("frame: ",frame)), color = "grey50") +
+    ## AXES CIRCLE PATH
     suppressWarnings( # Suppress for "Removed 1 rows containing missing values."
-      geom_path(data = circ_grid, color = "grey80",
-                mapping = aes(x = x+c_x, y = y+c_y))
+      geom_path(data = circ_grid, color = "grey50",
+                mapping = aes(x = x + c_x, y = y + c_y))
     )
 
-  # PROJECTION
-  gg2 <- gg1 +
-    # PROJ DATA POINTS
+  ## PROJECTION
+  gg <- gg +
+    ## PROJ DATA POINTS
     geom_point(data = data_grid, size = .7,
-               mapping = aes(x = x.x + x.y, y = y.x + y.y, color = cat),
-               shape = as.integer(cat) + 15) +
-    # PROJ DATA FRAME NUM
+               mapping = aes(x = x.x + x.y, y = y.x + y.y, 
+                             color = class, shape = class)) +
+    ## FACET FRAME NUM
     geom_text(data = data_grid, color = "grey50",
               mapping = aes(x = x.y - .7, y = y.y + 1.1, 
-                            label = paste0("frame: ",slide))) +
-    # OPTIONS
+                            label = paste0("frame: ",frame))) +
     theme_void() +
     scale_color_brewer(palette = "Dark2") + 
-    theme(legend.position="none")
-          #,panel.border = element_rect(colour = "black", fill = NA))
+    theme(legend.position = "none")
 
-  gg2
-}
-
-view_basis2 <- function(basis, labels = paste0("V", 1:nrow(basis)),
-                        manip_var, data = NULL, axes = "center"){
-  basis <- as.data.frame(basis)
-  angle <- seq(0, 2 * pi, length = 360)
-  circ <- data.frame(x = cos(angle), y = sin(angle))
-  zero <- set_axes_position(0, axes)
-  basis <- set_axes_position(basis, axes)
-  circ <- set_axes_position(circ, axes)
-  col <- rep("black", nrow(basis))
-  col[manip_var] <- "blue"
-
-  gg <- ggplot2::ggplot() + ggplot2::scale_color_brewer(palette = "Dark2") + 
-    ggplot2::theme_void() + ggplot2::theme(legend.position = "none") + 
-    ggplot2::coord_fixed()
-  if (axes != "off") {
-    gg <- 
-      gg + ggplot2::geom_path(
-        data = circ, mapping = ggplot2::aes(x = x, y = y), 
-        color = "grey80", size = 0.3, inherit.aes = F
-      ) + 
-      ggplot2::geom_segment(
-        data = basis, color = col, 
-        mapping = ggplot2::aes(x = V1, y = V2, xend = zero, yend = zero)
-      ) + 
-      ggplot2::geom_text(
-        data = basis, mapping = ggplot2::aes(x = V1, y = V2, label = labels), 
-        size = 4, color = col, hjust = 1.1, vjust=.4
-      )
-  }
-  if (!is.null(data)) {
-        proj <- as.data.frame(tourr::rescale(as.matrix(data) %*% 
-            as.matrix(basis)) - 0.5)
-        gg <- gg + ggplot2::geom_point(data = proj, 
-                                       mapping = ggplot2::aes(x = V1, y = V2))
-    }
+  ## Return
   gg
@@
                                manip_col = "blue",
-                               tilt = pi *
-                                 5 / 12,
+                               tilt = pi * 5 / 12,
                                z_col = "red",
@@
 {
-    find_angle <-
-      function(a, b)
+  #### Finds the angle between two vectors
+  find_angle <- function(a, b)
     acos(sum(a * b) / (sqrt(sum(a * a)) * sqrt(sum(b * b))) )
-    make_curve <-
-      function(ang_st = 0,
+  #### Makes a df semi-circle, with 1 row per degree
+  make_curve <- function(ang_st = 0,
                          ang_stop = 2 * pi) {
-        # angle is atan(y/x)
-        angle <- seq(ang_st, ang_stop,
-                     length = round(360 / (2 * pi) * abs(ang_st - ang_stop)) )
+      degrees <- round(360 / (2 * pi) * abs(ang_st - ang_stop))
+      angle <- seq(ang_st, ang_stop, length = degrees)
       data.frame(x = cos(angle), y = sin(angle), z = sin(angle))
     }
-    xyz <- function(x) { colnames(x)<-c("x","y","z");x }
 
+  ## Initialize
   p <- nrow(basis)
+  m_sp   <- as.data.frame(create_manip_space(basis, manip_var))
+  colnames(m_sp) <- c("x", "y", "z")
+  m_sp_r <- m_sp %>% mutate(y = y * cos(tilt), z = z * sin(tilt))
+  mvar   <- m_sp[manip_var, ]
+  mvar_r <- m_sp_r[manip_var, ]
+
   col_v <- rep("grey80", p)
@@
   circ_r <- make_curve() %>%
-      mutate( y=y*cos(tilt), z=z*sin(tilt))
-    m_sp   <- xyz(as.data.frame(create_manip_space(basis, manip_var)))
-    m_sp_r <- m_sp %>% mutate( y=y*cos(tilt), z=z*sin(tilt))
-    mvar   <- m_sp[manip_var, ]
-    mvar_r <- m_sp_r[manip_var, ]
+    mutate(y = y * cos(tilt), z = z * sin(tilt))
 
@@
   phi_offset  <- find_angle(c(mvar_r$x, mvar_r$y), c(1,0))
-    phi_curve_r <- .4*make_curve(ang_st = phi_offset, ang_stop = phi_ang) %>%
+  phi_curve_r <- .4 * make_curve(ang_st = phi_offset, ang_stop = phi_ang) %>%
+    mutate(y = y * cos(tilt), z = z * sin(tilt))
+
+  thata_ang     <- find_angle(c(mvar$x, mvar$y),c(1, 0))
+  theta_curve_r <- .5 * make_curve(ang_st = 0, ang_stop = thata_ang) %>%
     mutate( y=y*cos(tilt), z=z*sin(tilt))
 
-    thata_ang     <- find_angle(c(mvar$x, mvar$y),c(1, 0))
-    theta_curve_r <- .5*make_curve(ang_st = 0, ang_stop = thata_ang) %>%
-      mutate( y=y*cos(tilt), z=z*sin(tilt))
-
+  ## Render
   gg <-
@@
     ggplot2::coord_fixed() +
+    ## Circle
     ggplot2::geom_path(data = circ_r,
                        mapping = ggplot2::aes(x = x, y = y),
-                           color = manip_col,
-                           size = 0.3,
-                           inherit.aes = F
-        ) +
-        #projection plane:
+                       color = manip_col, size = 0.3, inherit.aes = FALSE) +
+    ## Projection plane:
     ggplot2::geom_segment(data = m_sp_r,
@@
                           ),
-                              size = siz_v,
-                              colour = col_v
-        ) +
+                          size = siz_v, colour = col_v) +
     ggplot2::geom_text(data = m_sp_r,
                        mapping = ggplot2::aes(x = x, y = y, label = labels),
-                           size = 4,
-                           colour = col_v,
+                       size = 4, colour = col_v,
+                       vjust = "outward", hjust = "outward") +
+    ## Z direction
+    ggplot2::geom_path(data = circ_r,
+                       mapping = ggplot2::aes(x = x, y = z),
+                       color = z_col, size = 0.3, inherit.aes = FALSE) +
+    ggplot2::geom_segment(data = mvar_r,
+                          mapping = ggplot2::aes(x = x,y = z,
+                                                 xend = 0,yend = 0),
+                          size = 1, colour = z_col) +
+    ggplot2::geom_segment(data = mvar_r,
+                          mapping = ggplot2::aes(x = x, y = z,
+                                                 xend = x,yend = y),
+                          size = 0.3, colour = "grey80", linetype = 2) +
+    ggplot2::geom_text(data = mvar_r,
+                       mapping = ggplot2::aes(x = x, y = z, 
+                                              label = labels[manip_var]),
+                       size = 4, colour = z_col,
                        vjust = "outward",
@@
     ) +
-        #Z direction
-        ggplot2::geom_path(data = circ_r,
-                           mapping = ggplot2::aes(x = x, y = z),
-                           color = z_col,
-                           size = 0.3,
-                           inherit.aes = F
-        ) +
-        ggplot2::geom_segment(data = mvar_r,
-                              mapping = ggplot2::aes(x = x,y = z,
-                                                     xend = 0,yend = 0),
-                              size = 1,
-                              colour = z_col
-        ) +
-        ggplot2::geom_segment(data = mvar_r,
-                              mapping = ggplot2::aes(x = x, y = z,
-                                                     xend = x,yend = y),
-                              size = 0.3,
-                              colour = "grey80",
-                              linetype = 2
-        ) +
-        ggplot2::geom_text(data = mvar_r,
-                           mapping = ggplot2::aes(x = x, y = z, label = labels[manip_var]),
-                           size = 4,
-                           colour = z_col,
-                           vjust = "outward",
-                           hjust = "outward"
-        ) +
-        # label angles
+    ## Label angles
     geom_path(data = phi_curve_r,
               mapping = aes(x=x, y=z),
-                  color = z_col,
-                  size = 0.2
-        ) +
+              color = z_col, size = 0.2) +
     geom_text(data = 1.2 * phi_curve_r[ceiling(nrow(phi_curve_r)/2), ],
               mapping = ggplot2::aes(x=x, y=z, label = "phi"),
-                  color = z_col,
-                  parse = T,
-                  size = 4
-        ) +
+              color = z_col, size = 4, parse = TRUE) +
     geom_path(data = theta_curve_r,
@@
               color = manip_col,
-                  size = 0.2
-        ) +
+              size = 0.2) +
     geom_text(data = 1.2 * theta_curve_r[ceiling(nrow(theta_curve_r)/2), ],
               mapping = ggplot2::aes(x=x, y=y-.02, label = "theta"),
-                  color = manip_col,
-                  parse = T,
-                  size = 4
-        )
+              color = z_col, size = 4, parse = TRUE)
+
+  ## Return
   gg
@@
 Tours can be considered to extend the dimensionality of visualization, which is important as data and models exist in more than 3D.
-The package \CRANpkg{tourr} [@wickham_tourr_2011] provides a platform for generating tours.
+The package \CRANpkg{tourr} [@wickham_tourr:_2011] provides a platform for generating tours.
 It can produce a variety of tours, each paired with a variety of possible displays.
@@
 Section \ref{sec:algorithm} describes the algorithm used to perform a radial manual tour as implemented in the package \CRANpkg{spinifex}.
-Section \ref{sec:display} explains how to generate an animation of the manual tour using the animation frameworks offered by \CRANpkg{plotly} [@sievert_plotly_2018] and \CRANpkg{gganimate} [@pedersen_gganimate:_2019].
+Section \ref{sec:display} explains how to generate an animation of the manual tour using the animation frameworks offered by \CRANpkg{plotly} [@sievert_interactive_2020] and \CRANpkg{gganimate} [@pedersen_gganimate_2020].
 Package functionality and code usage following the order applied in the algorithm follows in section \ref{sec:usage}.
@@
 The color and shape of points are mapped to the flea species.
-This plot is made using the \code{view\_basis()} function in \pkg{spinifex}, which generates a \CRANpkg{ggplot2} [@wickham_ggplot2:_2016] object.
+This plot is made using the \code{view\_basis()} function in \pkg{spinifex}, which generates a \CRANpkg{ggplot2} [@wickham_ggplot2_2016] object.
 
@@
 ## hard code a basis. tourr doesn't fix results with set.seed().
-f_path <-  c(.693, -.022, .082, -.119, .706, .023, 
-             -.070, .438, .405, .515, .103, .604)
-f_bas <- matrix(f_path, ncol=2)
-# f_path <- save_history(f_dat, guided_tour(holes()))
-# f_bas  <- matrix(f_path[,, max(dim(f_path)[3])], ncol=2)
+f_bas <-  c(.693, -.022, .082, -.119, .706, .023, 
+             -.070, .438, .405, .515, .103, .604) %>% 
+  matrix( ncol=2) %>% 
+  tourr::orthonormalise()
 f_mvar <- 5
@@
 # Adjust centering to make axes and data side by side
-view_basis(f_bas, lab = colnames(f_dat)) +
-  geom_point(data = f_proj,
-             mapping = aes(x = X1 + .75, y = X2 - .5, color = f_cat),
-             pch = as.integer(f_cat) + 15) +
-  theme(panel.border = element_rect(colour = "black", fill = NA))
+oblique_frame(f_bas, data = f_dat, manip_var = f_mvar, lab = colnames(f_dat),
+              color = f_cat, shape = f_cat) + 
+  coord_fixed() + theme_void() + 
+  theme(legend.position = "none")
 ```
@@
 This is illustrated with the \\code{view\\_manip\\_space()} function."}
-view_manip_space2(basis = f_bas, manip_var = f_mvar, labels = colnames(f_dat)) +
-    theme(panel.border = element_rect(colour = "black", fill = NA))
+view_manip_space2(basis = f_bas, manip_var = f_mvar, lab = colnames(f_dat))
 ```
@@
 
-```{r step3, echo=F, warning=F, fig.height=2, fig.width=8, out.width='100%',  fig.cap = "Snapshots of a radial manual tour manipulating aede2: (1) original projection, (2) full contribution, (3) zero contribution, (4) back to original.
+```{r step3, echo=F, warning=F, out.width='100%',  fig.cap = "Snapshots of a radial manual tour manipulating aede2: (1) original projection, (2) full contribution, (3) zero contribution, (4) back to original.
 "}
@@
 if(dim(f_mtour)[3] != 15) message(
-  paste0("!!! n_slides != 15 !!! Check angle step size. n_slides = ", dim(f_mtour)[3], ". angle = ", f_angle))
-p1 <- view_basis2(f_mtour[,,1],  colnames(f_dat), manip_var = f_mvar) + ggtitle("1 phi=0.78")
-p2 <- view_basis2(f_mtour[,,5],  colnames(f_dat), manip_var = f_mvar) + ggtitle("2  phi=0")
-p3 <- view_basis2(f_mtour[,,12], colnames(f_dat), manip_var = f_mvar) + ggtitle("3 phi=1.57")
-p4 <- view_basis2(f_mtour[,,15], colnames(f_dat), manip_var = f_mvar) + ggtitle("4 phi=0.78")
-grid.arrange(p1, p2, p3, p4, ncol=4)
+  paste0("!!! n_frames != 15 !!! Check angle step size. n_frames = ", dim(f_mtour)[3], ". angle = ", f_angle))
+
+p1 <- oblique_frame(f_mtour[,, 1],  lab = colnames(f_dat), manip_var = f_mvar) +
+  ggtitle("(1) phi = 0.78") + coord_fixed() +
+  theme_void() + theme(legend.position = "none")
+p2 <- oblique_frame(f_mtour[,, 5],  lab = colnames(f_dat), manip_var = f_mvar) +
+  ggtitle("(2) phi = 0") + coord_fixed() +
+  theme_void() + theme(legend.position = "none")
+p3 <- oblique_frame(f_mtour[,, 12], lab = colnames(f_dat), manip_var = f_mvar) +
+  ggtitle("(3) phi = 1.57") + coord_fixed() +
+  theme_void() + theme(legend.position = "none")
+p4 <- oblique_frame(f_mtour[,, 15], lab = colnames(f_dat), manip_var = f_mvar) +
+  ggtitle("(4) phi = 0.78") + coord_fixed() +
+  theme_void() + theme(legend.position = "none")
+grid.arrange(p1, p2, p3, p4, ncol = 4)
 ```
@@
 
-```{r step4, echo=F, warning=F, fig.height=7.0, fig.width=5.83, out.height='7in', out.width='5.83in',                                             fig.cap = "Radial manual tour manipulating aede2 of standardized flea data.
+```{r step4, echo=F, warning=F, out.height='7in', out.width='5.83in',                                           fig.cap = "Radial manual tour manipulating aede2 of standardized flea data.
 The axis for aede2 increases in contribution to the projection, from its initial value to 1, decreasing to 0 and then returning to the initial value.
@@
 #play_manual_tour(f_dat, f_bas, f_mvar, col = f_cat, angle = f_angle)
-array2static(.m_tour = f_mtour, .data = f_dat, .m_var = f_mvar, .cat = f_cat)
+array2ggfacets(tour_array = f_mtour, data = f_dat, m_var = f_mvar, class = f_cat)
 ```
 
-# Package structure and functionality {#sec:usage}
+# Package structure and functionality
 
@@
 ```{r installCODEEXAMPLE, eval=F, echo=T}
+## Install from CRAN
 install.package("spinifex")
+## Load into session
 library("spinifex")
@@
 
-\noindent Also see the vignette for examples of usage:
+\noindent Also see the shiny app for understandign and the vignette for basic usage:
 
-```
+```{r installCODEEXAMPLE2, eval=F, echo=T}
+## Shiny app for visualizing basic application
+run_app("intro")
+## View the code vignette
 vignette("spinifex_vignette")
@@
 
-```
+```{r installCODEEXAMPLE3, eval=F, echo=T}
+## Optionally install latest developmention version from GitHub
 remotes::install_github("nspyrison/spinifex")
@@
   c("", "", ""),
-  c("specialty", "view_basis", "display the reference frame of a given basis"),
+  c("specialty", "oblique_basis", "table of the rotated basis and manip space"),
+  c("specialty", "oblique_frame", "display the reference axes of a given basis"),
   c("specialty", "view_manip_space", "illustrative display of any manip space")
@@
 
-## Usage
+## Usage {#sec:usage}
 
@@
 library(spinifex)
-f_data  <- tourr::rescale(flea[, 1:6])                    ## standardized flea data
-f_path  <- save_history(f_data, guided_tour(holes()))     ## guided tour path, holes index
-f_basis <- matrix(f_path[,, max(dim(f_path)[3])], ncol=2) ## optimal projection
-f_cat   <- factor(flea$species)                           ## categorical var
-f_mvar  <- 5                                              ## set manip var number
-angle_speed <- .26                                        ## anglular dist between slides
-play_manual_tour(data = f_data,                           ## generates animation
-                 basis = f_basis,                         ##   plotly obj (default)
+## Standardized flea data
+f_data <- tourr::rescale(flea[, 1:6])
+## Guided tour path, holes index
+f_path <- save_history(f_data, guided_tour(holes()))
+## Local extrema found
+f_basis <- matrix(f_path[,, max(dim(f_path)[3])], ncol=2)
+## Categorical class variable
+f_cat <- factor(flea$species)
+## Manip var, number of the variable to alter
+f_mvar <- 5
+## Anglular dist between frames (radians)
+step_size <- .26
+## Render and play animate, as plotly object by default
+play_manual_tour(data = f_data,
+                 basis = f_basis,
                  manip_var = f_mvar, 
-                 angle = angle_speed,
+                 angle = step_size,
                  col = f_cat,
@@
 
-\noindent The \code{play\_manual\_tour()} function is a composite function for user convenience that does these operations:
-
-```{r eval=FALSE, echo=TRUE}
-manual_tour()
-array2df()
-render_plotly()
-```
-
-\noindent This will generate an html animation using plotly.
-Switching the renderer to \code{gganimate} is an option, which will also create an html animation or an animated gif.
+\noindent The \code{play\_manual\_tour()} function is a composite function handling interaction between \code{manual\_tour()}, \code{array2df()}, and \code{render\_plotly()}.
+This will generate an html animation using plotly.
+Switching the renderer to \code{render\_gganimate()} alternatively creates an animated gif.
 Each of these formats allows for the animation to be made available on a web site, or directly visible in an html formatted document.
@@
 ```{r eval=FALSE}
-view_basis(basis = f_basis, 
+## View a basis and projected data 
+oblique_frame(basis = f_basis, 
               data = f_data,
-           labels = colnames(f_data))
+              color = f_cat,
+              shape = f_cat)
 ```
@@
 ```{r step2CODEEXAMPLE, eval=F, echo=T}
+## Displays the projection plane and manipulation space for the 
 view_manip_space(basis = f_basis, 
                  manip_var = f_mvar, 
-                 labels = colnames(f_data))
+                 lab = colnames(f_data))
 ```
@@
 
-```{r JetClusterGood, echo=F, warning=F, fig.height=7.0, fig.width=5.83, out.height='7in', out.width='5.83in',                                     fig.cap="Snapshots of a radial manual tour of PC4 within the jet cluster, with color indicating experiment type: ATLAS7new (green) and ATLAS7old (orange).
+```{r JetClusterGood, echo=F, warning=F, out.height='7in', out.width='5.83in',                                    fig.cap="Snapshots of a radial manual tour of PC4 within the jet cluster, with color indicating experiment type: ATLAS7new (green) and ATLAS7old (orange).
 When PC4 is removed from the projection (frame 10) there is little difference between the groups, suggesting that PC4 is important for distinguishing the experiments.
@@
 
-array2static(.m_tour = jet_mtour, .data = jet_dat,
-             .m_var = jet_mvar, .cat = jet_cat)
+array2ggfacets(tour_array = jet_mtour, data = jet_dat,
+             m_var = jet_mvar, class = jet_cat)
 # play_manual_tour(data = jet_dat, basis = jet_bas, manip_var = jet_mvar,
@@
 
-```{r JetClusterBad, echo=F, warning=F, fig.height=7.0, fig.width=5.83, out.height='7in', out.width='5.83in',                                       fig.cap = "Snapshots of a radial manual tour of PC3 within the jet cluster, with color indicating experiment type: ATLAS7new (green) and ATLAS7old (orange).
+```{r JetClusterBad, echo=F, warning=F, out.height='7in', out.width='5.83in',                                       fig.cap = "Snapshots of a radial manual tour of PC3 within the jet cluster, with color indicating experiment type: ATLAS7new (green) and ATLAS7old (orange).
 When the contribution from PC3 is changed there is little change to the structure of the two groups, suggesting that PC3 is not important for distinguishing the experiments.
@@
 
-array2static(.m_tour = jet_mtour, .data = jet_dat,
-             .m_var = jet_mvar, .cat = jet_cat)
+array2ggfacets(tour_array = jet_mtour, data = jet_dat,
+             m_var = jet_mvar, class = jet_cat)
 ```
@@
 
-```{r DISclusterGood, echo=F, warning=F, fig.height=7.0, fig.width=5.83, out.height='7in', out.width='5.83in',                                            fig.cap = "Snapshots of a radial manual tour exploring the sensitivity PC6 has on the structure of the DIS cluster, with color indicating experiment type: DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange).
+```{r DISclusterGood, echo=F, warning=F, out.height='7in', out.width='5.83in',                                          fig.cap = "Snapshots of a radial manual tour exploring the sensitivity PC6 has on the structure of the DIS cluster, with color indicating experiment type: DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange).
 DIS HERA1+2 is distributed in a cross-shaped plane, charm SIDIS occupies the center of this cross, and dimuon SIDIS is a linear cluster crossing DIS HERA1+2.
@@
 
-array2static(.m_tour = DIS_mtour, .data = DIS_dat,
-             .m_var = DIS_mvar, .cat = DIS_cat)
+array2ggfacets(tour_array = DIS_mtour, data = DIS_dat,
+             m_var = DIS_mvar, class = DIS_cat)
 # play_manual_tour(data = DIS_dat, basis = DIS_bas, manip_var = DIS_mvar,
@@
 
-```{r DISclusterBad, echo=F, warning=F, fig.height=7.0, fig.width=5.83, out.height='7in', out.width='5.83in',                                            fig.cap = "Snapshots of a radial manual tour exploring the sensitivity PC2 to the structure of the DIS cluster, with color indicating experiment type: DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange).
+```{r DISclusterBad, echo=F, warning=F, out.height='7in', out.width='5.83in',                                       fig.cap = "Snapshots of a radial manual tour exploring the sensitivity PC2 to the structure of the DIS cluster, with color indicating experiment type: DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange).
 As contribution from PC2 is decreased, dimuon SIDIS becomes more distinguishable from the other two clusters (frames 10-14), indicating that in its absence PC2 is important.
@@
 
-array2static(.m_tour = DIS_mtour, .data = DIS_dat,
-             .m_var = DIS_mvar, .cat = DIS_cat)
+array2ggfacets(tour_array = DIS_mtour, data = DIS_dat,
+             m_var = DIS_mvar, class = DIS_cat)
 ```
@@
 
-<!-- geometric and graphics display 
-The \pkg{tourr} package provides many other geometric displays with the \code{tourr::display\_*()} family.
-These geometric options could be integrated into the \pkg{ggplot2} framework for display on \pkg{plotly} and \pkg{gganimate}.
-Additionally, the \CRANpkg{animation} package @xie_animation:_2018 could be implemented for another graphics framework.
-However, \pkg{animation} builds from \pkg{base} graphics while \pkg{spinifex} utilizes \pkg{ggplot2} graphics, a significant paradigm shift, which may have  a low return on investment.-->
-
 <!-- dynamic interaction/gui -->
 Having a graphical user interface would be useful for making it easier and more accessible to a general audience.
-This is possible to implement using \CRANpkg{shiny} [@chang_shiny:_2018].
+This is possible to implement using \CRANpkg{shiny} [@chang_shiny_2020].
 The primary purposes of the interface would be to allow the user to interactively change the manip variable easily, and the interpolation step for more or less detailed views.
@@
 
-<!-- removed [@r_core_team_r:_2019], as per journal reqs -->
-This article was created in R, using \CRANpkg{knitr} [@stodden_knitr:_2014] and \CRANpkg{rmarkdown} [@xie_r_2018], with code generating the examples inline.
+This article was created in R, using \CRANpkg{knitr} [@xie_knitr_2020] and \CRANpkg{rmarkdown} [@allaire_rmarkdown_2020], with code generating the examples inline.
 The source files for this article be found at [github.com/nspyrison/spinifex_paper/](https://github.com/nspyrison/spinifex_paper/).

### 2020-08-24 paper/spyrison-cook.tex
post proofing changes

@@
 
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
-
-\bibliography{spinifex_paper}
-
 \hypertarget{introduction}{%
@@
 glimpse of the high-dimensional space. An alternative approach is to use
-a tour (Asimov \protect\hyperlink{ref-asimov_grand_1985}{1985}) of
-dynamic linear projections to look at many different low-dimensional
-projections. Tours can be considered to extend the dimensionality of
-visualization, which is important as data and models exist in more than
-3D. The package \CRANpkg{tourr} (Wickham et al.
-\protect\hyperlink{ref-wickham_tourr_2011}{2011}) provides a platform
-for generating tours. It can produce a variety of tours, each paired
-with a variety of possible displays. A user can make a grand, guided,
-little, local or frozen tour, and display the resulting projected data
-as a scatterplot, density plot, histogram, or even as Chernoff faces if
-the projection dimension is more than 3.
+a tour \citep{asimov_grand_1985} of dynamic linear projections to look
+at many different low-dimensional projections. Tours can be considered
+to extend the dimensionality of visualization, which is important as
+data and models exist in more than 3D. The package \CRANpkg{tourr}
+\citep{wickham_tourr:_2011} provides a platform for generating tours. It
+can produce a variety of tours, each paired with a variety of possible
+displays. A user can make a grand, guided, little, local or frozen tour,
+and display the resulting projected data as a scatterplot, density plot,
+histogram, or even as Chernoff faces if the projection dimension is more
+than 3.
 
 This work adds a manual tour to the collection. The manual tour was
-described in Cook and Buja
-(\protect\hyperlink{ref-cook_manual_1997}{1997}) and allows a user to
-control the projection coefficients of a selected variable in a 2D
-projection. The manipulation of these coefficients allows the analyst to
-explore their sensitivity to the structure within the projection. As
-manual tours operate on only one variable at a time, they are
-particularly useful once a feature of interest has been identified.
+described in \citet{cook_manual_1997} and allows a user to control the
+projection coefficients of a selected variable in a 2D projection. The
+manipulation of these coefficients allows the analyst to explore their
+sensitivity to the structure within the projection. As manual tours
+operate on only one variable at a time, they are particularly useful
+once a feature of interest has been identified.
 
 One way to identify ``interesting'' features is with the use of a guided
-tour (Cook et al. \protect\hyperlink{ref-cook_grand_1995}{1995}). Guided
-tours select a very specific path, which approaches a projection that
-optimizes an objective function. The optimization used to guide the tour
-is simulated annealing (Kirkpatrick, Gelatt, and Vecchi
-\protect\hyperlink{ref-kirkpatrick_optimization_1983}{1983}). The direct
-optimization of a function allows guided tours to rapidly identify
-interesting projection features given the relatively large
-parameter-space. After a projection of interest is identified, an
-analyst can then use the ``finer brush'' of the manual tour to control
-the contributions of individual variables to explore the sensitivity
-they have on the structure visible in the projection.
+tour \citep{cook_grand_1995}. Guided tours select a very specific path,
+which approaches a projection that optimizes an objective function. The
+optimization used to guide the tour is simulated annealing
+\citep{kirkpatrick_optimization_1983}. The direct optimization of a
+function allows guided tours to rapidly identify interesting projection
+features given the relatively large parameter-space. After a projection
+of interest is identified, an analyst can then use the ``finer brush''
+of the manual tour to control the contributions of individual variables
+to explore the sensitivity they have on the structure visible in the
+projection.
 
@@
 generate an animation of the manual tour using the animation frameworks
-offered by \CRANpkg{plotly} (Sievert
-\protect\hyperlink{ref-sievert_plotly_2018}{2018}) and
-\CRANpkg{gganimate} (Pedersen and Robinson
-\protect\hyperlink{ref-pedersen_gganimate:_2019}{2019}). Package
+offered by \CRANpkg{plotly} \citep{sievert_interactive_2020} and
+\CRANpkg{gganimate} \citep{pedersen_gganimate_2020}. Package
 functionality and code usage following the order applied in the
@@
 analysis applied to multivariate data collected on high energy physics
-experiments (Wang et al.
-\protect\hyperlink{ref-wang_mapping_2018}{2018}). Section
-\ref{sec:discussion} summarizes this paper and discusses potential
-future directions.
+experiments \citep{wang_mapping_2018}. Section \ref{sec:discussion}
+summarizes this paper and discusses potential future directions.
 
@@
 The algorithm to conduct a manual tour interactively, by recording
-mouse/cursor motion, is described in detail in Cook and Buja
-(\protect\hyperlink{ref-cook_manual_1997}{1997}). Movement can be in any
-direction and magnitude, but it can also be constrained in several ways:
+mouse/cursor motion, is described in detail in \citet{cook_manual_1997}.
+Movement can be in any direction and magnitude, but it can also be
+constrained in several ways:
 
@@
 
-The flea data (Lubischew
-(\protect\hyperlink{ref-lubischew_use_1962}{1962})), available in the
-\pkg{tourr} package, is used to illustrate the algorithm. The data
-contains 74 observations on 6 variables, which are physical measurements
-made on flea beetles. Each observation belongs to one of three species.
+The flea data (\citet{lubischew_use_1962}), available in the \pkg{tourr}
+package, is used to illustrate the algorithm. The data contains 74
+observations on 6 variables, which are physical measurements made on
+flea beetles. Each observation belongs to one of three species.
 
@@
 made using the \code{view\_basis()} function in \pkg{spinifex}, which
-generates a \CRANpkg{ggplot2} (Wickham
-\protect\hyperlink{ref-wickham_ggplot2:_2016}{2016}) object.
+generates a \CRANpkg{ggplot2} \citep{wickham_ggplot2_2016} object.
 
@@
 follows, using Rodrigues' rotation formula (originally published in
-Rodrigues (\protect\hyperlink{ref-rodrigues_lois_1840}{1840})):
+\citet{rodrigues_lois_1840}):
 
@@
 
-\hypertarget{sec:usage}{%
-\section{Package structure and functionality}\label{sec:usage}}
+\hypertarget{package-structure-and-functionality}{%
+\section{Package structure and
+functionality}\label{package-structure-and-functionality}}
 
@@
 \begin{Sinput}
+## Install from CRAN
 install.package("spinifex")
+## Load into session
 library("spinifex")
@@
 
-\noindent Also see the vignette for examples of usage:
+\noindent Also see the shiny app for understandign and the vignette for
+basic usage:
 
-\begin{verbatim}
+\begin{Schunk}
+\begin{Sinput}
+## Shiny app for visualizing basic application
+run_app("intro")
+## View the code vignette
 vignette("spinifex_vignette")
-\end{verbatim}
+\end{Sinput}
+\end{Schunk}
 
@@
 
-\begin{verbatim}
+\begin{Schunk}
+\begin{Sinput}
+## Optionally install latest developmention version from GitHub
 remotes::install_github("nspyrison/spinifex")
-\end{verbatim}
+\end{Sinput}
+\end{Schunk}
 
@@
  &  & \\
-specialty & view\_basis & display the reference frame of a given basis\\
+specialty & oblique\_basis & table of the rotated basis and manip space\\
+specialty & oblique\_frame & display the reference axes of a given basis\\
 specialty & view\_manip\_space & illustrative display of any manip space\\
@@
 
-\hypertarget{usage}{%
-\subsection{Usage}\label{usage}}
+\hypertarget{sec:usage}{%
+\subsection{Usage}\label{sec:usage}}
 
@@
 \noindent The \code{play\_manual\_tour()} function is a composite
-function for user convenience that does these operations:
-
-\begin{Schunk}
-\begin{Sinput}
-manual_tour()
-array2df()
-render_plotly()
-\end{Sinput}
-\end{Schunk}
-
-\noindent This will generate an html animation using plotly. Switching
-the renderer to \code{gganimate} is an option, which will also create an
-html animation or an animated gif. Each of these formats allows for the
-animation to be made available on a web site, or directly visible in an
-html formatted document.
+function handling interaction between \code{manual\_tour()},
+\code{array2df()}, and \code{render\_plotly()}. This will generate an
+html animation using plotly. Switching the renderer to
+\code{render\_gganimate()} alternatively creates an animated gif. Each
+of these formats allows for the animation to be made available on a web
+site, or directly visible in an html formatted document.
 
@@
 \begin{Sinput}
-view_basis(basis = f_basis, 
+## View a basis and projected data 
+oblique_frame(basis = f_basis, 
               data = f_data,
-           labels = colnames(f_data))
+              color = f_cat,
+              shape = f_cat)
 \end{Sinput}
@@
 \begin{Sinput}
+## Displays the projection plane and manipulation space for the 
 view_manip_space(basis = f_basis, 
                  manip_var = f_mvar, 
-                 labels = colnames(f_data))
+                 lab = colnames(f_data))
 \end{Sinput}
@@
 
-Wang et al. (\protect\hyperlink{ref-wang_mapping_2018}{2018}) introduces
-a new tool, PDFSense, to visualize the sensitivity of hadronic
-experiments to nucleon structure. The parameter-space of these
-experiments lies in 56 dimensions, \(\delta \in \mathbb{R}^{56}\), and
-are visualized as 3D subspaces of the 10 first principal components in
-linear (PCA) and non-linear (t-SNE) embeddings.
+\citet{wang_mapping_2018} introduces a new tool, PDFSense, to visualize
+the sensitivity of hadronic experiments to nucleon structure. The
+parameter-space of these experiments lies in 56 dimensions,
+\(\delta \in \mathbb{R}^{56}\), and are visualized as 3D subspaces of
+the 10 first principal components in linear (PCA) and non-linear (t-SNE)
+embeddings.
 
-Cook, Laa, and Valencia
-(\protect\hyperlink{ref-cook_dynamical_2018}{2018}) illustrates how to
-learn more about the structures using a grand tour. Tours can better
-resolve the shape of clusters, intra-cluster detail, and better outlier
-detection than PDFSense \& TFEP (TensorFlow embedded projections) or
-traditional static embeddings. This example builds from here,
-illustrating how the manual tour can be used to examine the sensitivity
-of structure in a projection to different parameters. The specific 2D
-projections passed to the manual tour were provided in their work.
+\citet{cook_dynamical_2018} illustrates how to learn more about the
+structures using a grand tour. Tours can better resolve the shape of
+clusters, intra-cluster detail, and better outlier detection than
+PDFSense \& TFEP (TensorFlow embedded projections) or traditional static
+embeddings. This example builds from here, illustrating how the manual
+tour can be used to examine the sensitivity of structure in a projection
+to different parameters. The specific 2D projections passed to the
+manual tour were provided in their work.
 
@@
 experiments with four principal components explaining 95\% of the
-variation in the cluster (Cook, Laa, and Valencia
-\protect\hyperlink{ref-cook_dynamical_2018}{2018}). The data within this
-4D embedding is further subsetted, to ATLAS7old and ATLAS7new, to focus
-on two groups that occupy different parts of the subspace. Radial manual
-tours controlling contributions from PC4 and PC3 are shown in Figures
-\ref{fig:JetClusterGood} and \ref{fig:JetClusterBad}, respectively. The
-difference in shape can be interpreted as the experiments probing
-different phase-spaces. Back-transforming the principal components to
-the original variables can be done for a more detailed interpretation.
+variation in the cluster \citep{cook_dynamical_2018}. The data within
+this 4D embedding is further subsetted, to ATLAS7old and ATLAS7new, to
+focus on two groups that occupy different parts of the subspace. Radial
+manual tours controlling contributions from PC4 and PC3 are shown in
+Figures \ref{fig:JetClusterGood} and \ref{fig:JetClusterBad},
+respectively. The difference in shape can be interpreted as the
+experiments probing different phase-spaces. Back-transforming the
+principal components to the original variables can be done for a more
+detailed interpretation.
 
@@
 
-Following Cook, Laa, and Valencia
-(\protect\hyperlink{ref-cook_dynamical_2018}{2018}), to explore the DIS
-cluster, PCA is recomputed and the first six principal components,
-explaining 48\% of the full sample variation, are used. The
-contributions of PC6 and PC2 are explored in Figures
-\ref{fig:DISclusterGood} and \ref{fig:DISclusterBad}, respectively.
-Three experiments are examined: DIS HERA1+2 (green), dimuon SIDIS
-(purple), and charm SIDIS (orange).
+Following \citet{cook_dynamical_2018}, to explore the DIS cluster, PCA
+is recomputed and the first six principal components, explaining 48\% of
+the full sample variation, are used. The contributions of PC6 and PC2
+are explored in Figures \ref{fig:DISclusterGood} and
+\ref{fig:DISclusterBad}, respectively. Three experiments are examined:
+DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange).
 
@@
 research has taken the manual tour algorithm, specifically the radial
-rotation, used in GGobi (Swayne et al.
-\protect\hyperlink{ref-swayne_ggobi:_2003}{2003}) to interactively
+rotation, used in GGobi \citep{swayne_ggobi:_2003} to interactively
 rotate a variable into or out of a 2D projection, and modified it to
@@
 manual rotation in high dimensions is possible using Givens rotations
-and Householder reflections as outlined in Buja et al.
-(\protect\hyperlink{ref-buja_computational_2005}{2005}). This would
-provide more flexible manual rotation, but more difficult for a user
-because they have the choice (too much choice) of which directions to
-move.
+and Householder reflections as outlined in
+\citet{buja_computational_2005}. This would provide more flexible manual
+rotation, but more difficult for a user because they have the choice
+(too much choice) of which directions to move.
 
@@
 and more accessible to a general audience. This is possible to implement
-using \CRANpkg{shiny} (Chang et al.
-\protect\hyperlink{ref-chang_shiny:_2018}{2018}). The primary purposes
-of the interface would be to allow the user to interactively change the
+using \CRANpkg{shiny} \citep{chang_shiny_2020}. The primary purposes of
+the interface would be to allow the user to interactively change the
 manip variable easily, and the interpolation step for more or less
@@
 
-This article was created in R, using \CRANpkg{knitr} (Xie
-\protect\hyperlink{ref-stodden_knitr:_2014}{2014}) and
-\CRANpkg{rmarkdown} (Xie, Allaire, and Grolemund
-\protect\hyperlink{ref-xie_r_2018}{2018}), with code generating the
-examples inline. The source files for this article be found at
+This article was created in R, using \CRANpkg{knitr}
+\citep{xie_knitr_2020} and \CRANpkg{rmarkdown}
+\citep{allaire_rmarkdown_2020}, with code generating the examples
+inline. The source files for this article be found at
 \href{https://github.com/nspyrison/spinifex_paper/}{github.com/nspyrison/spinifex\_paper/}.
@@
 \href{https://github.com/nspyrison/spinifex/}{github.com/nspyrison/spinifex/}.
-
-\hypertarget{bibliography}{%
-\section*{Bibliography}\label{bibliography}}
-\addcontentsline{toc}{section}{Bibliography}
-
-\hypertarget{refs}{}
-\leavevmode\hypertarget{ref-asimov_grand_1985}{}%
-Asimov, Daniel. 1985. ``The Grand Tour: A Tool for Viewing
-Multidimensional Data.'' \emph{SIAM Journal on Scientific and
-Statistical Computing} 6 (1): 128--43.
-\url{https://doi.org/https://doi.org/10.1137/0906011}.
-
-\leavevmode\hypertarget{ref-buja_computational_2005}{}%
-Buja, Andreas, Dianne Cook, Daniel Asimov, and Catherine Hurley. 2005.
-``Computational Methods for High-Dimensional Rotations in Data
-Visualization.'' In \emph{Handbook of Statistics}, 24:391--413.
-Elsevier. \url{https://doi.org/10.1016/S0169-7161(04)24014-7}.
-
-\leavevmode\hypertarget{ref-chang_shiny:_2018}{}%
-Chang, Winston, Joe Cheng, J. J. Allaire, Yihui Xie, and Jonathan
-McPherson. 2018. \emph{Shiny: Web Application Framework for R}.
-\url{https://CRAN.R-project.org/package=shiny}.
-
-\leavevmode\hypertarget{ref-cook_manual_1997}{}%
-Cook, Dianne, and Andreas Buja. 1997. ``Manual Controls for
-High-Dimensional Data Projections.'' \emph{Journal of Computational and
-Graphical Statistics} 6 (4): 464--80.
-\url{https://doi.org/10.2307/1390747}.
-
-\leavevmode\hypertarget{ref-cook_grand_1995}{}%
-Cook, Dianne, Andreas Buja, Javier Cabrera, and Catherine Hurley. 1995.
-``Grand Tour and Projection Pursuit.'' \emph{Journal of Computational
-and Graphical Statistics} 4 (3): 155.
-\url{https://doi.org/10.2307/1390844}.
-
-\leavevmode\hypertarget{ref-cook_dynamical_2018}{}%
-Cook, Dianne, Ursula Laa, and German Valencia. 2018. ``Dynamical
-Projections for the Visualization of PDFSense Data.'' \emph{Eur. Phys.
-J. C} 78 (9): 742. \url{https://doi.org/10.1140/epjc/s10052-018-6205-2}.
-
-\leavevmode\hypertarget{ref-kirkpatrick_optimization_1983}{}%
-Kirkpatrick, Scott, C. Daniel Gelatt, and Mario P. Vecchi. 1983.
-``Optimization by Simulated Annealing.'' \emph{Science} 220 (4598):
-671--80. \url{https://doi.org/10.1126/science.220.4598.671}.
-
-\leavevmode\hypertarget{ref-lubischew_use_1962}{}%
-Lubischew, Alexander A. 1962. ``On the Use of Discriminant Functions in
-Taxonomy.'' \emph{Biometrics}, 455--77.
-\url{https://doi.org/10.2307/2527894}.
-
-\leavevmode\hypertarget{ref-pedersen_gganimate:_2019}{}%
-Pedersen, Thomas Lin, and David Robinson. 2019. \emph{Gganimate: A
-Grammar of Animated Graphics}.
-\url{http://github.com/thomasp85/gganimate}.
-
-\leavevmode\hypertarget{ref-rodrigues_lois_1840}{}%
-Rodrigues, Olinde. 1840. \emph{Des Lois Géométriques Qui Régissent Les
-Déplacements d'un Système Solide Dans L'espace: Et de La Variation Des
-Cordonnées Provenant de Ces Déplacements Considérés Indépendamment Des
-Causes Qui Peuvent Les Produire}.
-
-\leavevmode\hypertarget{ref-sievert_plotly_2018}{}%
-Sievert, Carson. 2018. \emph{Plotly for R}.
-\url{https://plotly-book.cpsievert.me}.
-
-\leavevmode\hypertarget{ref-swayne_ggobi:_2003}{}%
-Swayne, Deborah F, Duncan Temple Lang, Andreas Buja, and Dianne Cook.
-2003. ``GGobi: Evolving from XGobi into an Extensible Framework for
-Interactive Data Visualization.'' \emph{Computational Statistics \& Data
-Analysis}, Data Visualization, 43 (4): 423--44.
-\url{https://doi.org/10.1016/S0167-9473(02)00286-4}.
-
-\leavevmode\hypertarget{ref-wang_mapping_2018}{}%
-Wang, Bo-Ting, T. J. Hobbs, Sean Doyle, Jun Gao, Tie-Jiun Hou, Pavel M.
-Nadolsky, and Fredrick I. Olness. 2018. ``Mapping the Sensitivity of
-Hadronic Experiments to Nucleon Structure.'' \emph{Physical Review D} 98
-(9): 094030. \url{https://doi.org/10.1103/PhysRevD.98.094030}.
-
-\leavevmode\hypertarget{ref-wickham_ggplot2:_2016}{}%
-Wickham, Hadley. 2016. \emph{Ggplot2: Elegant Graphics for Data
-Analysis}. Springer-Verlag New York. \url{http://ggplot2.org}.
-
-\leavevmode\hypertarget{ref-wickham_tourr_2011}{}%
-Wickham, Hadley, Dianne Cook, Heike Hofmann, and Andreas Buja. 2011.
-``\textbf{Tourr} : An \emph{R} Package for Exploring Multivariate Data
-with Projections.'' \emph{Journal of Statistical Software} 40 (2).
-\url{https://doi.org/10.18637/jss.v040.i02}.
-
-\leavevmode\hypertarget{ref-stodden_knitr:_2014}{}%
-Xie, Yihui. 2014. ``Knitr: A Comprehensive Tool for Reproducible
-Research in R.'' In \emph{Implementing Reproducible Computational
-Research}, edited by Victoria Stodden, Friedrich Leisch, and Roger D.
-Peng. Chapman; Hall/CRC.
-\url{http://www.crcpress.com/product/isbn/9781466561595}.
-
-\leavevmode\hypertarget{ref-xie_r_2018}{}%
-Xie, Yihui, J. J. Allaire, and Garrett Grolemund. 2018. \emph{R
-Markdown: The Definitive Guide}. Boca Raton, Florida: Chapman; Hall/CRC.
-\url{https://bookdown.org/yihui/rmarkdown}.
 

### 2020-08-24 paper/spyrison-cook.Rmd
caching issue, view_basis to oblique_frame

@@
   error      = FALSE,
-  cache      = TRUE, # needed for log and figure out?
+  cache      = FALSE, # needed for log and figure out?
   cache.lazy = FALSE
@@
                                z_col = "red",
-                               labels = paste0("V", 1:nrow(basis)))
+                               lab = paste0("V", 1:nrow(basis)))
 {
@@
     ggplot2::geom_text(data = m_sp_r,
-                       mapping = ggplot2::aes(x = x, y = y, label = labels),
+                       mapping = ggplot2::aes(x = x, y = y, label = lab),
                        size = 4, colour = col_v,
@@
                        mapping = ggplot2::aes(x = x, y = z, 
-                                              label = labels[manip_var]),
+                                              label = lab[manip_var]),
                        size = 4, colour = z_col,
@@
     geom_path(data = phi_curve_r,
-              mapping = aes(x=x, y=z),
+              mapping = aes(x = x, y = z),
               color = z_col, size = 0.2) +
-    geom_text(data = 1.2 * phi_curve_r[ceiling(nrow(phi_curve_r)/2), ],
+    geom_text(data = 1.2 * phi_curve_r[ceiling(nrow(phi_curve_r) / 2), ],
               mapping = ggplot2::aes(x=x, y=z, label = "phi"),
@@
     geom_text(data = 1.2 * theta_curve_r[ceiling(nrow(theta_curve_r)/2), ],
-              mapping = ggplot2::aes(x=x, y=y-.02, label = "theta"),
-              color = z_col, size = 4, parse = TRUE)
+              mapping = ggplot2::aes(x = x, y = y - .02, label = "theta"),
+              color = manip_col, size = 4, parse = TRUE)
 
@@
 
-The function `view_basis` can be used to show a projection of the basis, or with the data overlaid.
+The function `oblique_frame` can be used to show a projection of the basis, or with the data overlaid.
 For example, the plots in Figures \ref{fig:step0} and \ref{fig:step3} were made with code similar to this:

### 2020-08-25 paper/spyrison-cook.Rmd
Code touch up, gifs saved to /output/.

@@
   error      = FALSE,
-  cache      = FALSE, # needed for log and figure out?
+  cache      = FALSE,
   cache.lazy = FALSE
@@
   remotes::install_github("nspyrison/spinifex", 
-                          ref = "da75976588dc008e21f01ce6a57d9778d0ed6933")
+                          ref = "dc8244c7e3f470fe3eee86f5c577c0dae21136c3")
 }
@@
 - $k$, is the index of the variable to manipulate, called the "manip var".
-<!--- $|\textbf{B}_k|$, the norm of the manip var, $k$, the length of this basis vector between 0 and 1 inclusive.
--->
+
 - $\textbf{e}$, a 1D basis vector of length $p$, with 1 in the $k$-th position and 0 elsewhere.
@@
 
-```{r step0, echo=F, results='hide', out.width='70%',                        fig.cap = "Initial 2D projection: representation of the basis (left) and resulting data projection (right) of standardized flea data.
+```{r step0, echo=F, results='hide', out.width='70%',                                 fig.cap="Initial 2D projection: representation of the basis (left) and resulting data projection (right) of standardized flea data.
 The color and shape of data points are mapped to beetle species.
@@
 f_dat <- tourr::rescale(flea[,1:6])
-f_cat  <- factor(flea$species)
+f_clas <- factor(flea$species)
 ## hard code a basis. tourr doesn't fix results with set.seed().
@@
 oblique_frame(f_bas, data = f_dat, manip_var = f_mvar, lab = colnames(f_dat),
-              color = f_cat, shape = f_cat) + 
+              color = f_clas, shape = f_clas) + 
   coord_fixed() + theme_void() + 
@@
 
-```{r step2, echo=F, out.width='70%',                                                fig.cap = "Illustration of a 3D manip space, this space is rotated effectively changing the contribution of the manip variable, aede2 in the example data.
+```{r step2, echo=F, out.width='70%',                                                fig.cap="Illustration of a 3D manip space, this space is rotated effectively changing the contribution of the manip variable, aede2 in the example data.
 The blue circle and variable map lies on the projection plane.
@@
 
-```{r step3, echo=F, warning=F, out.width='100%',  fig.cap = "Snapshots of a radial manual tour manipulating aede2: (1) original projection, (2) full contribution, (3) zero contribution, (4) back to original.
+```{r step3, echo=F, warning=F, out.width='100%',                                     fig.cap="Snapshots of a radial manual tour manipulating aede2: (1) original projection, (2) full contribution, (3) zero contribution, (4) back to original.
 "}
@@
 
-```{r step4, echo=F, warning=F, out.height='7in', out.width='5.83in',                                           fig.cap = "Radial manual tour manipulating aede2 of standardized flea data.
+```{r step4, echo=F, warning=F, out.height='7in', out.width='5.83in',                 fig.cap="Radial manual tour manipulating aede2 of standardized flea data.
 The axis for aede2 increases in contribution to the projection, from its initial value to 1, decreasing to 0 and then returning to the initial value.
@@
 An animation can be viewed at https://nspyrison.netlify.com/thesis/flea\\_manualtour\\_mvar5/."}
-#play_manual_tour(f_dat, f_bas, f_mvar, col = f_cat, angle = f_angle)
-array2ggfacets(tour_array = f_mtour, data = f_dat, m_var = f_mvar, class = f_cat)
+## Figure
+array2ggfacets(tour_array = f_mtour, data = f_dat, m_var = f_mvar, class = f_clas)
+
+## Saving gif for the supplementary material:
+if (FALSE){
+  play_manual_tour(basis = f_bas, data = f_dat, manip_var = f_mvar,
+                   color = f_clas, shape = f_clas,
+                   ggtheme = list(theme_void(), 
+                                  ggtitle("flea data, radial tour of aede2")),
+                   fps = 8, start_pause = .5, end_pause = 1,
+                   render_type = render_gganimate,
+                   gif_filename = "flea_manualtour_mvar5.gif", 
+                   gif_path = "./output")
+}
 ```
@@
 ## Categorical class variable
-f_cat <- factor(flea$species)
+f_clas <- factor(flea$species)
 ## Manip var, number of the variable to alter
@@
                  angle = step_size,
-                 col = f_cat,
-                 pch = f_cat)
+                 col = f_clas,
+                 pch = f_clas)
 ```
@@
               data = f_data,
-              color = f_cat,
-              shape = f_cat)
+              color = f_clas,
+              shape = f_clas)
 ```
@@
 jet_dat   <- tourr::rescale(JetCluster_sub[, 1:4])
-jet_cat   <- factor(JetCluster_sub$exp)
+jet_clas   <- factor(JetCluster_sub$exp)
 jet_mvar  <- 4
@@
 jet_mtour <- manual_tour(basis = jet_bas, manip_var = jet_mvar, angle = jet_ang)
-#dim(jet_mtour)
 
+## Figure
 array2ggfacets(tour_array = jet_mtour, data = jet_dat,
-             m_var = jet_mvar, class = jet_cat)
-# play_manual_tour(data = jet_dat, basis = jet_bas, manip_var = jet_mvar,
-#                  angle = jet_ang, axes = "bottomleft", col = jet_cat)
+               m_var = jet_mvar, class = jet_clas)
+
+## Saving gif for the supplementary material:
+if (FALSE){
+  play_manual_tour(basis = jet_bas, data = jet_dat, manip_var = jet_mvar,
+                   color = jet_clas, shape = jet_clas,
+                   ggtheme = list(theme_void(), 
+                                  ggtitle("jet cluster, radial tour of PC4")),
+                   fps = 8, start_pause = .5, end_pause = 1,
+                   render_type = render_gganimate,
+                   gif_filename = "jetcluster_manualtour_pc4.gif", 
+                   gif_path = "./output")
+}
 
@@
 
-```{r JetClusterBad, echo=F, warning=F, out.height='7in', out.width='5.83in',                                       fig.cap = "Snapshots of a radial manual tour of PC3 within the jet cluster, with color indicating experiment type: ATLAS7new (green) and ATLAS7old (orange).
+```{r JetClusterBad, echo=F, warning=F, out.height='7in', out.width='5.83in',         fig.cap="Snapshots of a radial manual tour of PC3 within the jet cluster, with color indicating experiment type: ATLAS7new (green) and ATLAS7old (orange).
 When the contribution from PC3 is changed there is little change to the structure of the two groups, suggesting that PC3 is not important for distinguishing the experiments.
@@
 
+## Figure
 array2ggfacets(tour_array = jet_mtour, data = jet_dat,
-             m_var = jet_mvar, class = jet_cat)
+               m_var = jet_mvar, class = jet_clas)
+
+## Saving gif for the supplementary material:
+if (FALSE){
+  play_manual_tour(basis = jet_bas, data = jet_dat, manip_var = jet_mvar,
+                   color = jet_clas, shape = jet_clas,
+                   ggtheme = list(theme_void(), 
+                                 ggtitle("jet cluster, radial tour of PC3")),
+                   fps = 8, start_pause = .5, end_pause = 1,
+                   render_type = render_gganimate,
+                   gif_filename = "jetcluster_manualtour_pc3.gif", 
+                   gif_path = "./output")
+}
 ```
@@
 
-```{r DISclusterGood, echo=F, warning=F, out.height='7in', out.width='5.83in',                                          fig.cap = "Snapshots of a radial manual tour exploring the sensitivity PC6 has on the structure of the DIS cluster, with color indicating experiment type: DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange).
+```{r DISclusterGood, echo=F, warning=F, out.height='7in', out.width='5.83in',        fig.cap="Snapshots of a radial manual tour exploring the sensitivity PC6 has on the structure of the DIS cluster, with color indicating experiment type: DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange).
 DIS HERA1+2 is distributed in a cross-shaped plane, charm SIDIS occupies the center of this cross, and dimuon SIDIS is a linear cluster crossing DIS HERA1+2.
@@
 DIS_dat   <- tourr::rescale(grDIScenter[, 1:6])
-DIS_cat   <- factor(grDIScenter$disID)
+DIS_clas   <- factor(grDIScenter$disID)
 DIS_mvar  <- 6
@@
 
+## Figure
 array2ggfacets(tour_array = DIS_mtour, data = DIS_dat,
-             m_var = DIS_mvar, class = DIS_cat)
-# play_manual_tour(data = DIS_dat, basis = DIS_bas, manip_var = DIS_mvar,
-#                  angle = DIS_ang, axes="bottomleft", col = DIS_cat)
+               m_var = DIS_mvar, class = DIS_clas)
+
+## Saving gif for the supplementary material
+if (FALSE){
+  play_manual_tour(basis = DIS_bas, data = DIS_dat, manip_var = DIS_mvar,
+                   color = DIS_clas, shape = DIS_clas,
+                   ggtheme = list(theme_void(), 
+                                  ggtitle("DIS cluster, radial tour of PC6")),
+                   fps = 8, start_pause = .5, end_pause = 1,
+                   render_type = render_gganimate,
+                   gif_filename = "discluster_manualtour_pc6.gif", 
+                   gif_path = "./output")
+}
 
@@
 
-```{r DISclusterBad, echo=F, warning=F, out.height='7in', out.width='5.83in',                                       fig.cap = "Snapshots of a radial manual tour exploring the sensitivity PC2 to the structure of the DIS cluster, with color indicating experiment type: DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange).
+```{r DISclusterBad, echo=F, warning=F, out.height='7in', out.width='5.83in',         fig.cap="Snapshots of a radial manual tour exploring the sensitivity PC2 to the structure of the DIS cluster, with color indicating experiment type: DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange).
 As contribution from PC2 is decreased, dimuon SIDIS becomes more distinguishable from the other two clusters (frames 10-14), indicating that in its absence PC2 is important.
@@
 
+## Figure
 array2ggfacets(tour_array = DIS_mtour, data = DIS_dat,
-             m_var = DIS_mvar, class = DIS_cat)
+               m_var = DIS_mvar, class = DIS_clas)
+
+## Saving gif for the supplementary material
+if (FALSE){
+  play_manual_tour(basis = DIS_bas, data = DIS_dat, manip_var = DIS_mvar,
+                   color = DIS_clas, shape = DIS_clas,
+                   ggtheme = list(theme_void(), 
+                                  ggtitle("jet cluster, radial tour of PC2")),
+                   fps = 8, start_pause = .5, end_pause = 1,
+                   render_type = render_gganimate,
+                   gif_filename = "discluster_manualtour_pc2.gif", 
+                   gif_path = "./output")
+}
 ```

### 2020-08-25 paper/spyrison-cook.tex
caching issue, view_basis to oblique_frame
Code touch up, gifs saved to /output/.

@@
 \begin{itemize}
-\tightlist
 \item
@@
                  angle = step_size,
-                 col = f_cat,
-                 pch = f_cat)
+                 col = f_clas,
+                 pch = f_clas)
 \end{Sinput}
@@
 
-The function \texttt{view\_basis} can be used to show a projection of
+The function \texttt{oblique\_frame} can be used to show a projection of
 the basis, or with the data overlaid. For example, the plots in Figures
@@
               data = f_data,
-              color = f_cat,
-              shape = f_cat)
+              color = f_clas,
+              shape = f_clas)
 \end{Sinput}

### 2020-08-26 paper/gifs/discluster_radialtour_pc2.rmd
gifs to seprate htmls

---
title: "Supplemental material -- figure 8 animation: DIS cluster, radial tour on PC2"
output: html_document
---

```{r, echo=FALSE, fig.cap="Animation of figure 8. Snapshots of a radial manual tour exploring the sensitivity PC2 to the structure of the DIS cluster, with color indicating experiment type: DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange). As contribution from PC2 is decreased, dimuon SIDIS becomes more distinguishable from the other two clusters (frames 10-14), indicating that in its absence PC2 is important."}
knitr::include_graphics("../gifs/discluster_radialtour_pc2.gif")
```

### 2020-08-26 paper/gifs/discluster_radialtour_pc6.rmd
gifs to seprate htmls

---
title: "Supplemental material -- figure 7 animation: DIS cluster, radial tour on PC6"
output: html_document
---

```{r, echo=FALSE, fig.cap="Animation of figure 7. Snapshots of a radial manual tour exploring the sensitivity PC6 has on the structure of the DIS cluster, with color indicating experiment type: DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange). DIS HERA1+2 is distributed in a cross-shaped plane, charm SIDIS occupies the center of this cross, and dimuon SIDIS is a linear cluster crossing DIS HERA1+2. As the contribution of PC6 is increased, DIS HERA1+2 becomes almost singular in one direction (frame 5), indicating that this experiment has very little variability in the direction of PC6."}
knitr::include_graphics("../gifs/discluster_radialtour_pc6.gif")
```

### 2020-08-26 paper/gifs/flea_radialtour_mvar5.rmd
gifs to seprate htmls

---
title: "Supplemental material -- figure 4 animation: flea data, radial tour on aede2"
output: html_document
---

```{r, echo=FALSE, fig.cap="Animation of figure 4. Radial manual tour manipulating aede2 of standardized flea data. The axis for aede2 increases in contribution to the projection, from its initial value to 1, decreasing to 0 and then returning to the initial value. This effects the separation between the purple and green clusters. This shows that aede2 is important for distinguishing the purple species, because the separation disappears when aede2 is not contributing to the projection."}
knitr::include_graphics("../gifs/flea_radialtour_mvar5.gif")
```

### 2020-08-26 paper/gifs/jetcluster_radialtour_pc3.rmd
gifs to seprate htmls

---
title: "Supplemental material -- figure 6 animation: jet cluster, radial tour on PC3"
output: html_document
---

```{r, echo=FALSE, fig.cap="Animation of figure 6. Snapshots of a radial manual tour of PC3 within the jet cluster, with color indicating experiment type: ATLAS7new (green) and ATLAS7old (orange).  When the contribution from PC3 is changed there is little change to the structure of the two groups, suggesting that PC3 is not important for distinguishing the experiments."}
knitr::include_graphics("../gifs/jetcluster_radialtour_pc3.gif")
```

### 2020-08-26 paper/gifs/jetcluster_radialtour_pc4.rmd
gifs to seprate htmls

---
title: "Supplemental material -- figure 5 animation: jet cluster, radial tour on PC4"
output: html_document
---

```{r, echo=FALSE, fig.cap="Animation of figure 5. Snapshots of a radial manual tour of PC4 within the jet cluster, with color indicating experiment type: ATLAS7new (green) and ATLAS7old (orange). When PC4 is removed from the projection (frame 10) there is little difference between the groups, suggesting that PC4 is important for distinguishing the experiments."}
knitr::include_graphics("../gifs/jetcluster_radialtour_pc4.gif")
```

### 2020-08-26 paper/spyrison-cook.Rmd
gifs to seprate htmls

@@
   remotes::install_github("nspyrison/spinifex", 
-                          ref = "dc8244c7e3f470fe3eee86f5c577c0dae21136c3")
+                          ref = "176a5e9184e794109579e741cddc82931ceb0dfb")
 }
@@
 # Adjust centering to make axes and data side by side
-oblique_frame(f_bas, data = f_dat, manip_var = f_mvar, lab = colnames(f_dat),
+view_frame(f_bas, data = f_dat, manip_var = f_mvar, lab = colnames(f_dat),
               color = f_clas, shape = f_clas) + 
@@
 
-p1 <- oblique_frame(f_mtour[,, 1],  lab = colnames(f_dat), manip_var = f_mvar) +
+p1 <- view_frame(f_mtour[,, 1],  lab = colnames(f_dat), manip_var = f_mvar) +
   ggtitle("(1) phi = 0.78") + coord_fixed() +
   theme_void() + theme(legend.position = "none")
-p2 <- oblique_frame(f_mtour[,, 5],  lab = colnames(f_dat), manip_var = f_mvar) +
+p2 <- view_frame(f_mtour[,, 5],  lab = colnames(f_dat), manip_var = f_mvar) +
   ggtitle("(2) phi = 0") + coord_fixed() +
   theme_void() + theme(legend.position = "none")
-p3 <- oblique_frame(f_mtour[,, 12], lab = colnames(f_dat), manip_var = f_mvar) +
+p3 <- view_frame(f_mtour[,, 12], lab = colnames(f_dat), manip_var = f_mvar) +
   ggtitle("(3) phi = 1.57") + coord_fixed() +
   theme_void() + theme(legend.position = "none")
-p4 <- oblique_frame(f_mtour[,, 15], lab = colnames(f_dat), manip_var = f_mvar) +
+p4 <- view_frame(f_mtour[,, 15], lab = colnames(f_dat), manip_var = f_mvar) +
   ggtitle("(4) phi = 0.78") + coord_fixed() +
@@
                    color = f_clas, shape = f_clas,
-                   ggtheme = list(theme_void(), 
-                                  ggtitle("flea data, radial tour of aede2")),
+                   ggtheme = list(theme_void(), theme(legend.position = "none")),
                    fps = 8, start_pause = .5, end_pause = 1,
                    render_type = render_gganimate,
-                   gif_filename = "flea_manualtour_mvar5.gif", 
-                   gif_path = "./output")
+                   gif_filename = "flea_radialtour_mvar5.gif", 
+                   gif_path = "./gifs")
 }
@@
   c("", "", ""),
-  c("specialty", "oblique_basis", "table of the rotated basis and manip space"),
-  c("specialty", "oblique_frame", "display the reference axes of a given basis"),
+  c("specialty", "print_manip_space", "table of the rotated basis and manip space"),
+  c("specialty", "view_frame", "display the reference axes of a given basis"),
   c("specialty", "view_manip_space", "illustrative display of any manip space")
@@
 
-The function `oblique_frame` can be used to show a projection of the basis, or with the data overlaid.
+The function `view_frame` can be used to show a projection of the basis, or with the data overlaid.
 For example, the plots in Figures \ref{fig:step0} and \ref{fig:step3} were made with code similar to this:
@@
 ## View a basis and projected data 
-oblique_frame(basis = f_basis, 
+view_frame(basis = f_basis, 
               data = f_data,
@@
                    color = jet_clas, shape = jet_clas,
-                   ggtheme = list(theme_void(), 
-                                  ggtitle("jet cluster, radial tour of PC4")),
+                   ggtheme = list(theme_void(), theme(legend.position = "none")),
                    fps = 8, start_pause = .5, end_pause = 1,
                    render_type = render_gganimate,
-                   gif_filename = "jetcluster_manualtour_pc4.gif", 
-                   gif_path = "./output")
+                   gif_filename = "jetcluster_radialtour_pc4.gif", 
+                   gif_path = "./gifs")
 }
@@
                    color = jet_clas, shape = jet_clas,
-                   ggtheme = list(theme_void(), 
-                                 ggtitle("jet cluster, radial tour of PC3")),
+                   ggtheme = list(theme_void(), theme(legend.position = "none")),
                    fps = 8, start_pause = .5, end_pause = 1,
                    render_type = render_gganimate,
-                   gif_filename = "jetcluster_manualtour_pc3.gif", 
-                   gif_path = "./output")
+                   gif_filename = "jetcluster_radialtour_pc3.gif", 
+                   gif_path = "./gifs")
 }
@@
                    color = DIS_clas, shape = DIS_clas,
-                   ggtheme = list(theme_void(), 
-                                  ggtitle("DIS cluster, radial tour of PC6")),
+                   ggtheme = list(theme_void(), theme(legend.position = "none")),
                    fps = 8, start_pause = .5, end_pause = 1,
                    render_type = render_gganimate,
-                   gif_filename = "discluster_manualtour_pc6.gif", 
-                   gif_path = "./output")
+                   gif_filename = "discluster_radialtour_pc6.gif", 
+                   gif_path = "./gifs")
 }
@@
                    color = DIS_clas, shape = DIS_clas,
-                   ggtheme = list(theme_void(), 
-                                  ggtitle("jet cluster, radial tour of PC2")),
+                   ggtheme = list(theme_void(), theme(legend.position = "none")),
                    fps = 8, start_pause = .5, end_pause = 1,
                    render_type = render_gganimate,
-                   gif_filename = "discluster_manualtour_pc2.gif", 
-                   gif_path = "./output")
+                   gif_filename = "discluster_radialtour_pc2.gif", 
+                   gif_path = "./gifs")
 }

### 2020-08-26 paper/spyrison-cook.tex
gifs to seprate htmls

@@
  &  & \\
-specialty & oblique\_basis & table of the rotated basis and manip space\\
-specialty & oblique\_frame & display the reference axes of a given basis\\
+specialty & print\_manip\_space & table of the rotated basis and manip space\\
+specialty & view\_frame & display the reference axes of a given basis\\
 specialty & view\_manip\_space & illustrative display of any manip space\\
@@
 
-The function \texttt{oblique\_frame} can be used to show a projection of
+The function \texttt{view\_frame} can be used to show a projection of
 the basis, or with the data overlaid. For example, the plots in Figures

### 2020-08-26 spinifex_RJpostProofing_20200824/spyrison-cook.Rmd
gifs and rmds for all pcs, not just figures. and their zip

--- 2020-08-26 paper/spyrison-cook.Rmd
@@
 
-## Saving gif for the supplementary material:
+## Saving gif of radial tour for each PC; for the supplementary material
 if (FALSE){
-  play_manual_tour(basis = jet_bas, data = jet_dat, manip_var = jet_mvar,
+  for(i in 1:ncol(jet_dat)){
+    fn <- paste0("jetcluster_radialtour_pc", i, ".gif")
+    play_manual_tour(basis = jet_bas, data = jet_dat, manip_var = i,
                      color = jet_clas, shape = jet_clas,
@@
                      render_type = render_gganimate,
-                   gif_filename = "jetcluster_radialtour_pc4.gif", 
+                     gif_filename = fn, 
                      gif_path = "./gifs")
+  }
 }
@@
                m_var = jet_mvar, class = jet_clas)
-
-## Saving gif for the supplementary material:
-if (FALSE){
-  play_manual_tour(basis = jet_bas, data = jet_dat, manip_var = jet_mvar,
-                   color = jet_clas, shape = jet_clas,
-                   ggtheme = list(theme_void(), theme(legend.position = "none")),
-                   fps = 8, start_pause = .5, end_pause = 1,
-                   render_type = render_gganimate,
-                   gif_filename = "jetcluster_radialtour_pc3.gif", 
-                   gif_path = "./gifs")
-}
 ```
@@
 
-## Saving gif for the supplementary material
+## Saving gif of radial tour for each PC; for the supplementary material
 if (FALSE){
-  play_manual_tour(basis = DIS_bas, data = DIS_dat, manip_var = DIS_mvar,
+  for(i in 1:ncol(DIS_dat)){
+    fn <- paste0("discluster_radialtour_pc", i, ".gif")
+    play_manual_tour(basis = DIS_bas, data = DIS_dat, manip_var = i,
                      color = DIS_clas, shape = DIS_clas,
@@
                      render_type = render_gganimate,
-                   gif_filename = "discluster_radialtour_pc6.gif", 
+                     gif_filename = fn, 
                      gif_path = "./gifs")
+  }
 }

### 2020-09-12 paper/spyrison-cook.Rmd
changes for CRAN ver 0.2.5

@@
 ```{r setup_paper, include=FALSE}
+if (F){ ## From reproducing, please use this dev ver:
+  install_version("ggplot2", version = "0.2.5", repos = "http://cran.us.r-project.org")
+}
+
 knitr::opts_chunk$set(
@@
 )
-if (F){ ## From reproducing, please use this dev ver:
-  remotes::install_github("nspyrison/spinifex", 
-                          ref = "176a5e9184e794109579e741cddc82931ceb0dfb")
-}
+
 library(tourr)
@@
 library(dplyr)
+
+my_theme <- list(scale_color_brewer(palette = "Dark2"),
+                 theme_void(), 
+                 theme(legend.position ="none"))
 ```
@@
   frames       <- array2df(array = tour_array, data = data)
-  basis_frames <- frames$basis_frames
-  data_frames  <- frames$data_frames
+  basis_frames <- frames$basis_slides
+  data_frames  <- frames$data_slides
   p            <- nrow(basis_frames) / n_frames
@@
   data_frames <- data.frame(data_frames, class = rep(class, n_frames))
+  colnames(data_frames)  <- c("x", "y", "frame", "class")
+  colnames(basis_frames) <- c("x", "y", "frame", "lab")
   grid_b <- grid_t <- data.frame(
@@
     geom_text(aes(x = x.y - .7, y = y.y + 1.1, 
-                  label = paste0("frame: ",frame)), color = "grey50") +
+                  label = paste0("frame: ", frame)), color = "grey50") +
     ## AXES CIRCLE PATH
@@
                             label = paste0("frame: ",frame))) +
-    theme_void() +
-    scale_color_brewer(palette = "Dark2") + 
-    theme(legend.position = "none")
+    my_theme
 
@@
   gg <-
-    ggplot2::ggplot() + ggplot2::scale_color_brewer(palette = "Dark2") +
-    ggplot2::theme_void() + ggplot2::theme(legend.position = "none") +
-    ggplot2::coord_fixed() +
+    ggplot2::ggplot() + my_theme +
     ## Circle
@@
 # Adjust centering to make axes and data side by side
-view_frame(f_bas, data = f_dat, manip_var = f_mvar, lab = colnames(f_dat),
+oblique_frame(f_bas, data = f_dat, manip_var = f_mvar, lab = colnames(f_dat),
               color = f_clas, shape = f_clas) + 
-  coord_fixed() + theme_void() + 
-  theme(legend.position = "none")
+  my_theme
 ```
@@
   paste0("!!! n_frames != 15 !!! Check angle step size. n_frames = ", dim(f_mtour)[3], ". angle = ", f_angle))
+ggproto
+z <- data.frame(0,0,0,0,0,0)
 
-p1 <- view_frame(f_mtour[,, 1],  lab = colnames(f_dat), manip_var = f_mvar) +
-  ggtitle("(1) phi = 0.78") + coord_fixed() +
-  theme_void() + theme(legend.position = "none")
-p2 <- view_frame(f_mtour[,, 5],  lab = colnames(f_dat), manip_var = f_mvar) +
-  ggtitle("(2) phi = 0") + coord_fixed() +
-  theme_void() + theme(legend.position = "none")
-p3 <- view_frame(f_mtour[,, 12], lab = colnames(f_dat), manip_var = f_mvar) +
-  ggtitle("(3) phi = 1.57") + coord_fixed() +
-  theme_void() + theme(legend.position = "none")
-p4 <- view_frame(f_mtour[,, 15], lab = colnames(f_dat), manip_var = f_mvar) +
-  ggtitle("(4) phi = 0.78") + coord_fixed() +
-  theme_void() + theme(legend.position = "none")
+p1 <- oblique_frame(f_mtour[,, 1], data = z, lab = colnames(f_dat), manip_var = f_mvar) +
+  ggtitle("(1) phi = 0.78") + my_theme
+p2 <- oblique_frame(f_mtour[,, 5], data = z, lab = colnames(f_dat), manip_var = f_mvar) +
+  ggtitle("(2) phi = 0") + my_theme
+p3 <- oblique_frame(f_mtour[,, 12], data = z, lab = colnames(f_dat), manip_var = f_mvar) +
+  ggtitle("(3) phi = 1.57") + my_theme
+p4 <- oblique_frame(f_mtour[,, 15], data = z, lab = colnames(f_dat), manip_var = f_mvar) +
+  ggtitle("(4) phi = 0.78") + my_theme
 grid.arrange(p1, p2, p3, p4, ncol = 4)
@@
   c("specialty", "print_manip_space", "table of the rotated basis and manip space"),
-  c("specialty", "view_frame", "display the reference axes of a given basis"),
+  c("specialty", "oblique_frame", "display the reference axes of a given basis"),
   c("specialty", "view_manip_space", "illustrative display of any manip space")
@@
 
-The function `view_frame` can be used to show a projection of the basis, or with the data overlaid.
+The function `oblique_frame` can be used to show a projection of the basis, or with the data overlaid.
 For example, the plots in Figures \ref{fig:step0} and \ref{fig:step3} were made with code similar to this:
@@
 ## View a basis and projected data 
-view_frame(basis = f_basis, 
+oblique_frame(basis = f_basis, 
               data = f_data,

### 2020-09-12 spinifex_RJpostProofing_20200912/spyrison-cook.tex
and associated zips

--- 2020-08-25 paper/spyrison-cook.tex
@@
 \begin{Schunk}
+\begin{Soutput}
+#> function (`_class` = NULL, `_inherit` = NULL, ...) 
+#> {
+#>     e <- new.env(parent = emptyenv())
+#>     members <- list(...)
+#>     if (length(members) != sum(nzchar(names(members)))) {
+#>         abort("All members of a ggproto object must be named.")
+#>     }
+#>     if (length(members) > 0) {
+#>         list2env(members, envir = e)
+#>     }
+#>     `_inherit` <- substitute(`_inherit`)
+#>     env <- parent.frame()
+#>     find_super <- function() {
+#>         eval(`_inherit`, env, NULL)
+#>     }
+#>     super <- find_super()
+#>     if (!is.null(super)) {
+#>         if (!is.ggproto(super)) {
+#>             abort("`_inherit` must be a ggproto object.")
+#>         }
+#>         e$super <- find_super
+#>         class(e) <- c(`_class`, class(super))
+#>     }
+#>     else {
+#>         class(e) <- c(`_class`, "ggproto", "gg")
+#>     }
+#>     e
+#> }
+#> <bytecode: 0x000000001d298b08>
+#> <environment: namespace:ggplot2>
+\end{Soutput}
 \begin{figure}
@@
  &  & \\
-specialty & oblique\_basis & table of the rotated basis and manip space\\
+specialty & print\_manip\_space & table of the rotated basis and manip space\\
 specialty & oblique\_frame & display the reference axes of a given basis\\

### 2020-09-12 paper/spyrison-cook.Rmd
final fixes for the published version

@@
   cache      = FALSE,
-  cache.lazy = FALSE
+  cache.lazy = FALSE,
+  fig.retina = 3
 )
@@
 
-```{r step2, echo=F, out.width='70%',                                                fig.cap="Illustration of a 3D manip space, this space is rotated effectively changing the contribution of the manip variable, aede2 in the example data.
+```{r step2, echo=F, out.width='50%', fig.width=3, fig.height=3,  fig.cap="Illustration of a 3D manip space, this space is rotated effectively changing the contribution of the manip variable, aede2 in the example data.
 The blue circle and variable map lies on the projection plane.
@@
 
-```{r step3, echo=F, warning=F, out.width='100%',                                     fig.cap="Snapshots of a radial manual tour manipulating aede2: (1) original projection, (2) full contribution, (3) zero contribution, (4) back to original.
+```{r step3, echo=F, warning=F, out.width='100%', fig.width=8, fig.height=2, fig.cap="Snapshots of a radial manual tour manipulating aede2: (1) original projection, (2) full contribution, (3) zero contribution, (4) back to original.
 "}
@@
 Tours are typically viewed as an animation.
-The animation of this tour can be viewed online at \url{https://nspyrison.netlify.com/thesis/flea_manualtour_mvar5/}.
+The animation of this tour can be viewed online at \url{https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/flea_radialtour_mvar5.gif}.
 The page may take a moment to load.
@@
 This effects the separation between the purple and green clusters.
-This shows that aede2 is important for distinguishing the purple species, because the separation disappears when aede2 is not contributing to the projection.
-An animation can be viewed at https://nspyrison.netlify.com/thesis/flea\\_manualtour\\_mvar5/."}
+This shows that aede2 is important for distinguishing the purple species, because the separation disappears when aede2 is not contributing to the projection."}
 ## Figure
@@
 However, removing PC3 from the projection (Figure \ref{fig:JetClusterBad}) does not affect the structure, indicating it is not important for distinguishing experiments.
-Animations for the remaining PCs can be viewed at the following links: [PC1](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc1/), [PC2](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc2/), [PC3](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/), and [PC4](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc4/).
+Animations for the remaining PCs can be viewed at the following links: [PC1](https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/jetcluster_manualtour_pc1.gif), [PC2](https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/jetcluster_manualtour_pc2.gif), [PC3](https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/jetcluster_manualtour_pc3.gif), and [PC4](https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/jetcluster_manualtour_pc4.gif).
 It can be seen that only PC4 is important for viewing the difference in these two experiments.
@@
 ```{r JetClusterGood, echo=F, warning=F, out.height='7in', out.width='5.83in', fig.cap="Snapshots of a radial manual tour of PC4 within the jet cluster, with color indicating experiment type: ATLAS7new (green) and ATLAS7old (orange).
-When PC4 is removed from the projection (frame 10) there is little difference between the groups, suggesting that PC4 is important for distinguishing the experiments.
-The animation can be viewed at https://nspyrison.netlify.com/thesis/jetcluster\\_manualtour\\_pc4/."}
+When PC4 is removed from the projection (frame 10) there is little difference between the groups, suggesting that PC4 is important for distinguishing the experiments."}
 # Jet cluster, Fig 7 of cook_dynamical_2018, subset to ATLAS7old and ATLAS7new
@@
 ```{r JetClusterBad, echo=F, warning=F, out.height='7in', out.width='5.83in', fig.cap="Snapshots of a radial manual tour of PC3 within the jet cluster, with color indicating experiment type: ATLAS7new (green) and ATLAS7old (orange).
-When the contribution from PC3 is changed there is little change to the structure of the two groups, suggesting that PC3 is not important for distinguishing the experiments.
-The animation can be viewed at https://nspyrison.netlify.com/thesis/jetcluster\\_manualtour\\_pc3/."}
+When the contribution from PC3 is changed there is little change to the structure of the two groups, suggesting that PC3 is not important for distinguishing the experiments."}
 # Jet cluster, Fig 7 of cook_dynalical_2018, subset to ATLAS7old and ATLAS7new
@@
 Even though both variables contribute similarly to the original projection, their contributions have quite different effects on the structure of each cluster, and the distinction between clusters.
-Animations of all of the principal components can be viewed from the links: [PC1](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc1/), [PC2](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc2/), [PC3](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc3/), [PC4](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc4/), [PC5](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc5/), and [PC6](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/).
+Animations of all of the principal components can be viewed from the links: [PC1](https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/discluster_manualtour_pc1.gif), [PC2](https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/discluster_manualtour_pc2.gif), [PC3](https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/discluster_manualtour_pc3.gif), [PC4](https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/discluster_manualtour_pc4.gif), [PC5](https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/discluster_manualtour_pc5.gif), and [PC6](https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/discluster_manualtour_pc6.gif).
 
@@
 DIS HERA1+2 is distributed in a cross-shaped plane, charm SIDIS occupies the center of this cross, and dimuon SIDIS is a linear cluster crossing DIS HERA1+2.
-As the contribution of PC6 is increased, DIS HERA1+2 becomes almost singular in one direction (frame 5), indicating that this experiment has very little variability in the direction of PC6.
-The animation can be viewed at https://nspyrison.netlify.com/thesis/discluster\\_manualtour\\_pc6/."}
+As the contribution of PC6 is increased, DIS HERA1+2 becomes almost singular in one direction (frame 5), indicating that this experiment has very little variability in the direction of PC6."}
 
@@
 ```{r DISclusterBad, echo=F, warning=F, out.height='7in', out.width='5.83in', fig.cap="Snapshots of a radial manual tour exploring the sensitivity PC2 to the structure of the DIS cluster, with color indicating experiment type: DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange).
-As contribution from PC2 is decreased, dimuon SIDIS becomes more distinguishable from the other two clusters (frames 10-14), indicating that in its absence PC2 is important.
-The animation can be viewed at https://nspyrison.netlify.com/thesis/discluster\\_manualtour\\_pc2/."}
+As contribution from PC2 is decreased, dimuon SIDIS becomes more distinguishable from the other two clusters (frames 10-14), indicating that in its absence PC2 is important."}
 ##DIS cluster, fig 8 of cook_dynamical_2018
@@
 The source files for this article be found at [github.com/nspyrison/spinifex_paper/](https://github.com/nspyrison/spinifex_paper/).
+The animated gifs can also be viewed at this site, and also in the supplementary material for this paper.
 The source code for the \pkg{spinifex} package can be found at [github.com/nspyrison/spinifex/](https://github.com/nspyrison/spinifex/).

### 2020-09-12 paper/spyrison-cook.tex
final fixes for the published version

--- 2020-08-25 paper/spyrison-cook.tex
@@
 Projections of Multivariate Data}
-\author{by Nicholas Spyrison, Dianne Cook}
+\author{by Nicholas Spyrison and Dianne Cook}
 
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{spyrison-cook_files/figure-latex/step2-1}
+{\centering \includegraphics[width=0.5\linewidth]{spyrison-cook_files/figure-latex/step2-1}
 
@@
 \begin{Schunk}
+\begin{Soutput}
+#> function (`_class` = NULL, `_inherit` = NULL, ...) 
+#> {
+#>     e <- new.env(parent = emptyenv())
+#>     members <- list(...)
+#>     if (length(members) != sum(nzchar(names(members)))) {
+#>         abort("All members of a ggproto object must be named.")
+#>     }
+#>     if (length(members) > 0) {
+#>         list2env(members, envir = e)
+#>     }
+#>     `_inherit` <- substitute(`_inherit`)
+#>     env <- parent.frame()
+#>     find_super <- function() {
+#>         eval(`_inherit`, env, NULL)
+#>     }
+#>     super <- find_super()
+#>     if (!is.null(super)) {
+#>         if (!is.ggproto(super)) {
+#>             abort("`_inherit` must be a ggproto object.")
+#>         }
+#>         e$super <- find_super
+#>         class(e) <- c(`_class`, class(super))
+#>     }
+#>     else {
+#>         class(e) <- c(`_class`, "ggproto", "gg")
+#>     }
+#>     e
+#> }
+#> <bytecode: 0x7fcdaa8d99c8>
+#> <environment: namespace:ggplot2>
+\end{Soutput}
 \begin{figure}
@@
 can be viewed online at
-\url{https://nspyrison.netlify.com/thesis/flea_manualtour_mvar5/}. The
-page may take a moment to load. Animations can be produced using the
+\url{https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/flea_radialtour_mvar5.gif}.
+The page may take a moment to load. Animations can be produced using the
 function \code{play\_manual\_tour()}.
@@
 This effects the separation between the purple and green clusters.
-This shows that aede2 is important for distinguishing the purple species, because the separation disappears when aede2 is not contributing to the projection.
-An animation can be viewed at https://nspyrison.netlify.com/thesis/flea\_manualtour\_mvar5/.}\label{fig:step4}
+This shows that aede2 is important for distinguishing the purple species, because the separation disappears when aede2 is not contributing to the projection.}\label{fig:step4}
 \end{figure}
@@
 remaining PCs can be viewed at the following links:
-\href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc1/}{PC1},
-\href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc2/}{PC2},
-\href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/}{PC3},
+\href{https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/jetcluster_manualtour_pc1.gif}{PC1},
+\href{https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/jetcluster_manualtour_pc2.gif}{PC2},
+\href{https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/jetcluster_manualtour_pc3.gif}{PC3},
 and
-\href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc4/}{PC4}.
+\href{https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/jetcluster_manualtour_pc4.gif}{PC4}.
 It can be seen that only PC4 is important for viewing the difference in
@@
 \caption[Snapshots of a radial manual tour of PC4 within the jet cluster, with color indicating experiment type]{Snapshots of a radial manual tour of PC4 within the jet cluster, with color indicating experiment type: ATLAS7new (green) and ATLAS7old (orange).
-When PC4 is removed from the projection (frame 10) there is little difference between the groups, suggesting that PC4 is important for distinguishing the experiments.
-The animation can be viewed at https://nspyrison.netlify.com/thesis/jetcluster\_manualtour\_pc4/.}\label{fig:JetClusterGood}
+When PC4 is removed from the projection (frame 10) there is little difference between the groups, suggesting that PC4 is important for distinguishing the experiments.}\label{fig:JetClusterGood}
 \end{figure}
@@
 \caption[Snapshots of a radial manual tour of PC3 within the jet cluster, with color indicating experiment type]{Snapshots of a radial manual tour of PC3 within the jet cluster, with color indicating experiment type: ATLAS7new (green) and ATLAS7old (orange).
-When the contribution from PC3 is changed there is little change to the structure of the two groups, suggesting that PC3 is not important for distinguishing the experiments.
-The animation can be viewed at https://nspyrison.netlify.com/thesis/jetcluster\_manualtour\_pc3/.}\label{fig:JetClusterBad}
+When the contribution from PC3 is changed there is little change to the structure of the two groups, suggesting that PC3 is not important for distinguishing the experiments.}\label{fig:JetClusterBad}
 \end{figure}
@@
 from the links:
-\href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc1/}{PC1},
-\href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc2/}{PC2},
-\href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc3/}{PC3},
-\href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc4/}{PC4},
-\href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc5/}{PC5},
+\href{https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/discluster_manualtour_pc1.gif}{PC1},
+\href{https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/discluster_manualtour_pc2.gif}{PC2},
+\href{https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/discluster_manualtour_pc3.gif}{PC3},
+\href{https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/discluster_manualtour_pc4.gif}{PC4},
+\href{https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/discluster_manualtour_pc5.gif}{PC5},
 and
-\href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/}{PC6}.
+\href{https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/discluster_manualtour_pc6.gif}{PC6}.
 
@@
 DIS HERA1+2 is distributed in a cross-shaped plane, charm SIDIS occupies the center of this cross, and dimuon SIDIS is a linear cluster crossing DIS HERA1+2.
-As the contribution of PC6 is increased, DIS HERA1+2 becomes almost singular in one direction (frame 5), indicating that this experiment has very little variability in the direction of PC6.
-The animation can be viewed at https://nspyrison.netlify.com/thesis/discluster\_manualtour\_pc6/.}\label{fig:DISclusterGood}
+As the contribution of PC6 is increased, DIS HERA1+2 becomes almost singular in one direction (frame 5), indicating that this experiment has very little variability in the direction of PC6.}\label{fig:DISclusterGood}
 \end{figure}
@@
 \caption[Snapshots of a radial manual tour exploring the sensitivity PC2 to the structure of the DIS cluster, with color indicating experiment type]{Snapshots of a radial manual tour exploring the sensitivity PC2 to the structure of the DIS cluster, with color indicating experiment type: DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange).
-As contribution from PC2 is decreased, dimuon SIDIS becomes more distinguishable from the other two clusters (frames 10-14), indicating that in its absence PC2 is important.
-The animation can be viewed at https://nspyrison.netlify.com/thesis/discluster\_manualtour\_pc2/.}\label{fig:DISclusterBad}
+As contribution from PC2 is decreased, dimuon SIDIS becomes more distinguishable from the other two clusters (frames 10-14), indicating that in its absence PC2 is important.}\label{fig:DISclusterBad}
 \end{figure}
@@
 \href{https://github.com/nspyrison/spinifex_paper/}{github.com/nspyrison/spinifex\_paper/}.
-The source code for the \pkg{spinifex} package can be found at
+The animated gifs can also be viewed at this site, and also in the
+supplementary material for this paper. The source code for the
+\pkg{spinifex} package can be found at
 \href{https://github.com/nspyrison/spinifex/}{github.com/nspyrison/spinifex/}.
@@
 Nicholas Spyrison\\
-Monash University\\
+Monash University\\%
 Faculty of Information Technology\\
+%
+%
+%
+\\\href{mailto:nicholas.spyrison@monash.edu}{\nolinkurl{nicholas.spyrison@monash.edu}}
 }
-\href{mailto:nicholas.spyrison@monash.edu}{\nolinkurl{nicholas.spyrison@monash.edu}}
 
@@
 Dianne Cook\\
-Monash University\\
+Monash University\\%
 Department of Econometrics and Business Statistics\\
+%
+%
+%
+\\\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}
 }
-\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}

### 2020-09-12 paper/spyrison-cook.Rmd
bibliography needed fixing

@@
 
-```{r step4, echo=F, warning=F, out.height='7in', out.width='5.83in', fig.cap="Radial manual tour manipulating aede2 of standardized flea data.
+```{r step4, echo=F, warning=F, out.height='7in', out.width='5.83in', out.width="100%", fig.cap="Radial manual tour manipulating aede2 of standardized flea data.
 The axis for aede2 increases in contribution to the projection, from its initial value to 1, decreasing to 0 and then returning to the initial value.

### 2020-09-12 paper/spyrison-cook.tex
bibliography needed fixing

@@
 #> }
-#> <bytecode: 0x7fcdaa8d99c8>
+#> <bytecode: 0x7fde89fe7e28>
 #> <environment: namespace:ggplot2>
@@
 
-{\centering \includegraphics[width=5.83in,height=7in]{spyrison-cook_files/figure-latex/step4-1}
+{\centering \includegraphics[width=1\linewidth,height=7in]{spyrison-cook_files/figure-latex/step4-1}
 
